### 2016-05-08 paper/paper.tex
Converts to LATeX

\documentclass[10pt,twocolumn]{article}

\usepackage{times}
\usepackage{fullpage}

\begin{document}

\title{Malacology: A Programmable Storage System Built on Ceph}
\author{Paper \# 123}
\date{}
\maketitle
\thispagestyle{empty}

\begin{abstract}
Storage systems are caught between rapidly changing data processing systems and
the increasing speed of storage devices. This puts tremendous pressure on
storage systems to adapt both in terms of their interfaces and their
performance. But adapting storage systems can be difficult because unprincipled
changes might jeopardize years of code-hardening and performance optimization
efforts that were necessary for users to entrust their data to the storage
system.  We introduce  Malacology, a prototype programmable storage system to
explore how existing abstractions of common services found in storage systems
can be leveraged to address new data processing systems and the increasing
speed of storage devices. This approach allows unprecedented flexibility for
storage systems to evolve without sacrificing the robustness of its
code-hardened subsystems.  We illustrate the advantages and challenges of
programmability by constructing two services out of existing abstractions: a
file system metadata load balancer and a high-performance distributed shared-log
that leverages flash devices.  RESULTS
\end{abstract}

\section{Introduction}\label{introduction}

A storage system implements abstractions designed to persistently store
data and must exhibit a high level of correctness to prevent data loss.
Storage systems have evolved around storage devices that often were
orders of magnitude slower than CPU and memory and therefore can
dominate overall performance if not used carefully. Over the last few
decades members of the storage systems community have developed
ingenious strategies to meet correctness requirements while somewhat
hiding the latency of traditional storage media. To avoid lock-in by a
particular vendor, users of storage systems have preferred systems with
highly standardized APIs and lowest common denominator abstract data
types such as blocks of bytes and byte stream files.

A number of recent developments are disrupting traditional storage
systems: (1) the falling prices of flash storage and the availability of
new types of non-volatile memory that are orders of magnitude faster
than traditional spinning media are moving overall performance
bottlenecks away from storage devices to CPUs and networking, and
pressure storage systems to shorten their code paths and incorporate new
optimizations; (2) demand for managing structured data and flexible
consistency semantics at scale pressure big data processing systems to
use storage abstractions that can meet these demands; and (3)
production-quality scalable storage systems available as open source
software have established and are continuing to establish new, \emph{de
facto} API standards at a faster pace than traditional standards bodies.

These three trends put evolutionary pressure on storage systems and
raise the question whether there are principles that storage systems
designers can follow to evolve storage systems efficiently and without
jeopardizing years of code-hardening and performance optimization
efforts that are important for users to continue to entrust their data
to the storage system.

In this paper we investigate an approach that focusses on generalizing
existing storage system resources, services, and abstractions that in
generalized form can be used to \emph{program} new services. By doing so
one can reuse subsystems and their optimizations and leverage their
established correctness, robustness, and efficiency. We will refer to
this programmability as \emph{programmable storage}, which differs from
\emph{active storage} (the injection and execution of arbitrary code in
a storage system) and \emph{software-defined storage} (the control of
thin-provisioning of storage).

To illustrate the benefits and challenges of this approach we examine
the programmability of Ceph \autocite{weil_ceph_2006}, the increasingly
popular, production-level open-source distributed storage system.
Something of a storage swiss army knife, Ceph supports file, block, and
object interfaces simultaneously in a single cluster. By introducing
programmability concepts into Ceph, we can build new services by
carefully exposing internal storage services to applications.

\begin{figure}[htbp]
\centering
\includegraphics{figures/overview.png}
\caption{A Ceph installation consists of a cluster running mostly
storage daemons (OSDs) to ensure data durability, a few monitor daemons
(MONs) to reach consensus of the cluster state and to ensure its
consistent versioning, and a few metadata service daemons (MDSs) for
scalable metadata service. Ceph's APIs include object, block, file
abstractions. Malacology enables the programmability of these
abstractions and services. To illustrate Malacology we implement two
services on Ceph, Zlog and Mantle, by programming Ceph's durability,
consistent versioning, consensus, and metadata subsystems.
\label{fig:overview}}
\end{figure}

For example, Ceph \autocite{weil_ceph_2006} contains a cluster of
monitoring nodes (MONs) that use PAXOS to maintain a consistent view of
system-wide metadata such as data distribution parameters and cluster
membership. The RADOS object storage system is a cluster of storage
devices (OSDs) that provide Ceph with data durability and integrity
using replication, erasure-coding, and scrubbing
\autocite{weil_rados_2007}. The data distrubiton metadata maintained by
the cluster monitors is used by RADOS servers and clients to provide
consistent reads and writes, and both RADOS and the monitoring services
are used by a cluster of file system metadata servers that provide POSIX
semantics using sophisticated indexing and distributed locking services.
We contend that re-using and re-purposing these code-hardened subsystems
is paramount to successfully adapting storage systems to new APIs and
new storage devices without losing the benefits from years of
code-hardening work.

To illustrate the benefits and challenges of this approach we have
designed and evaluated Malacology, a programmable storage system for
constructing new functionality by re-purposing existing subsystem
abstractions contained in Ceph. We build the framework on Ceph by
leveraging a broad spectrum of existing services, including distributed
locking and caching services provides by MDSs, durability and logical
storage devices provides by RADOS, and propogation of consistent cluster
state provided by the monitoring service. As shown in Figure
\ref{fig:overview}, this framework is expressive enough to provide the
functionality necessary for implementing new services. Our contributions
are:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  a programmable storage system implementation that re-uses and extends
  existing abstractions. It includes:

  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    inode abstractions for sequencer services
  \item
    consensus abstractions for policy versioning
  \item
    storage objects for persisting policies
  \item
    interface classes for new logical storage/metadata devices
  \end{enumerate}
\item
  example systems that use this framework

  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    shared log service based on CORFU \autocite{balakrishnan_corfu_2012}
  \item
    metadata load balancer based on Mantle
    \autocite{sevilla:sc15-mantle}
  \end{enumerate}
\end{itemize}

First, we describe existing techniques for bridging the gap between the
performance of storage systems and the requirements of applications
(\S\ref{problem}). Next, we discuss the ``distributed systems'' inspired
abstractions employed by Ceph (\S\ref{background}) and allude to their
re-usability as components in other services. The next section
enumerates the abstractions that Malacology exposes in the
implementation section (\S\ref{implementation}). We then show how we use
Malacology to synthesize entirely new storage services on an existing
system through configuration and small changes (\S\ref{services}). We
conclude by evaluating real-world use cases (\S\ref{evaluation}).

\section{Highly Tailored and Application-Specifc Storage
Systems}\label{highly-tailored-and-application-specifc-storage-systems}

\label{problem}

A consequence of complicated data management frameworks and faster
devices is that the storage system cannot meet the needs of the
general-purpose storage system. Workarounds for helping the application
meet its performance goals roughly fall into one of three categories:
``bolt-on'' services, application changes, and storage changes.

\subsection{\texorpdfstring{``Bolt-on''
services}{Bolt-on services}}\label{bolt-on-services}

So called ``bolt-on'' services are 3rd party systems that are integrated
into the system to provide functionality and improve performance for
common use-cases (e.g.~a metadata service). In addition to sacrificing
simplicity, ``bolt-on'' services often duplicate functionality and
execute redundant code, unnecessarily increasing the likelihood of bugs
or worse, introducing unpredictable performance problems. For example,
we are developing a distributed shared-commit log on Ceph called ZLog.
ZLog uses a ``sequencer'' to distribute tokens to clients that want to
append to the end of the log. When designing the sequencer, we
considered bolting on Zookeeper \cite{} as a service. Zookeeper would
satisfy our requirements by proiding a network and transaction stack but
it would perform poorly because Zookeeper persists its data and the
sequencer can be in-memory and volatile. A Zookeeper sequencer would
adhere to the pre-conceived notions of storage, namely that data in
storage systems should be safe from data loss, but would introduce
unecessary scalability limits \cite{Neha HDFS namenode example}.

\subsection{Application Changes}\label{application-changes}

The second approach to adapting to a storage system deficiency is to
extend the responsibility of the application. This means changing the
application itself by adding more data management intelligence or
domain-specific middleware (e.g., a new data layout). For instance, an
application may change itself to exploit data locality or I/O
parallelism in a distributed storage system.

For example, SciHadoop \autocites[
]{buck:hpc2011-scihadoop}{buck:sc2013-scidr} changes both the Hadoop
application and the Hadoop framework itself to leverage the structure of
scientific (3D array-based data, more specifically) to increase
performance, locality, and the number of early results. This is not a
bad proposition, but creates a coupling that is highly tied to the
underlying physical properties of the system, making it difficult to
adapt to future changes at the storage system level.

\subsection{Storage Changes}\label{storage-changes}

When these two approaches fail to meet the needs of the application,
developers turn their attention to the storage system itself.
Traditionally, storage system behaviour can be altered using two
techniques: tuning the system or introducing changes to the system. The
difficulty of both of these approaches has given rise to a third
technique called active storage.

\subsubsection{Tuning}\label{tuning}

Tuning large and complicated systems is difficult because of the number
of parameters (e.g., Hadoop 2.7.1 has 210 tunables) and because the
tunables are difficult to understand or quantify (e.g., Ceph tunables
\autocite{sevilla:sc15-mantle}). To succesfully tune the system, the
developer must have domain and system specific knowledge. Without this
intimate knowledge, the tuning turns into somewhat of a black art, where
the only way to figure out the best settings is trial and error.

Auto-tuning techniques attempt to find a good solution among a huge
space of available system configurations. However, in practice
auto-tuning is limited to only the configuration tunables that the
storage system exposes (e.g.~block size) and can be overwhelmed with too
many parameters. Starfish \autocite{herodotou:cidr2011-starfish} made an
attempt of this for Hadoop but this technique would need to severely
limit the space of parameters in order to feasible, similar to the
approach used in \autocite{behzad:sc2013-autotuning}. For instance,
auto-tuning may be capable of identifying instances in which new data
layouts would benefit a workload, but unless the system can provide such
a transformation, the option is left off the table.

\subsubsection{Software Changes}\label{software-changes}

As a last resort, the application developer can introduce changes to the
storage system itself. This endeavour is especially difficult because
the developer must first familiarate themselves with the storage system,
and even if they manage to make the necessary changes, they must get
their changes upstream for that specific project or become mantainer for
their in-house version. Being a maintainer means they re-base their
versions against new versions of the storage system and address any bugs
for their branched code. A successful example of this is the work that
focused on HDFS scalability for metadata-intensive workloads
\autocite{shvachko_hdfs_2010}. This has lead to modifications to its
architecture or API \autocite{balmin:sigmod2012-clydesdale} to improve
performance.

\begin{table}[ht]
\caption{
    A variety of RADOS object storage classes exist to expose interfaces
    to applications. \# is the number of methods that use these categories.
    \label{table:objclasses}}
  \begin{center}
  \begin{tabular}{l|l|l}
    Category & Specialization& \# \\ \hline
    \multirow{2}{*}{Locking} & Shared & \multirow{2}{*}{6} \\
                             & Exclusive & \\ \hdashline
    \multirow{3}{*}{Logging} & Replica & 3 \\
                             & State & 4 \\
                             & Timestamped & 4 \\ \hdashline
    \multirow{4}{*}{Metadata Managment} 
                             & RADOS Block Device  & 37 \\
                             & RADOS Gatway & 27 \\
                             & User & 5 \\
                             & Version & 5 \\ \hdashline
    Garbage Collection       & Reference Counting & 4 \\
\end{tabular}
\end{center}
\end{table}

\subsubsection{Hybrid Approaches}\label{hybrid-approaches}

As an alternative, some systems allow developers to extend the
interfaces of the storage devices to perform arbitrary operations on
data stored on disk. In Ceph, object interfaces are C/C++ plugins that
add new functionality to the OSDs. As an example use-case, ZLog needs to
store metadata information with each object so for each operation, it
checks the epoch value and write capability (2 metadata reads) and
writes the index (metadata write). Figure 1 shows the throughput
(\emph{y} axis) over time (\emph{x} axis) of two OSD implementations for
storing metadata: in the header of the byte stream (data) vs.~in the
object extended attributes (XATTR). The speed for appending data without
any metadata operations (data raw) is also shown as a baseline for
comparison. For append-heavy workloads storing metadata as a header in
the data performs about 1.5x better than storing metadata as an extended
attribute.

\begin{figure}[htbp]
\centering
\includegraphics{figures/cls_xattr-vs-data.png}
\caption{{[}\href{https://github.com/michaelsevilla/malacology-popper/blob/master/experiments/figure1/visualize.ipynb}{source}{]}
When appending data to objects, the object class that stores metadata in
a header in the data byte stream (data) performs 1.5x better than the
object class that stores metdata in the extended attributes of the
object (XATTR); it is almost as fast as appending data without updating
metadata (data raw).}
\end{figure}

Object interface classese are used in production Ceph environments;
Figure \ref{fig:obj-int-dev-growth} shows a dramatic growth in the use
of co-designed interfaces in the Ceph community since 2010. Figure
\ref{fig:obj-int-dev-churn} examines this growth in interfaces further
by showing the lines of code that are changed (\emph{y} axis) over time
(\emph{x} axis). The prevalence of blue dots indicates that users
frequently update their interfaces close to release cycles and that
massive changes come in bunches. This interface churn reflects both the
popularity of object interfaces and the complexity of the processing
being done by the OSDs. What is most remarkable about Figure
\ref{fig:obj-int-dev-churn} is that this trend contradicts the notion
that API changes are a burden for users. The types of object interfaces,
which are enumerated in Table \ref{table:objclasses}, show that this
active storage development is present throughough the Ceph software
stack.

\begin{figure}[htbp]
\centering
\includegraphics{figures/obj-int-dev-growth.png}
\caption{Since 2010, the growth in the number of co-designed object
storage interfaces in Ceph has been accelerating. This plot is the
number of object classes (a group of interfaces), and the total number
of methods (the actual API end-points).\label{fig:obj-int-dev-growth}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/obj-int-dev-churn.png}
\caption{Source code changes over time indicate the dynamic interface
development will need to support a high frequency of interface churn.
Each dot represents a Ceph commit with a corresponding number of lines
of code changed. \label{fig:obj-int-dev-churn}}
\end{figure}

The popularity of the active storage component of Ceph hint at three
trends in the Ceph community: (1) increasingly, the default
algorithms/tunables of the storage system are insufficient for the
application's performance goals, (2) programmers are becoming more aware
of their application's behavior, and (3) programmers know how to manage
resources to improve performance. Programmers gravitate towards object
interfaces because it gives them ability to tell the storage system
about their application: if it is CPU or IO bound, if it has locality,
if its size has the potential to overload a single proxy node, etc. The
programmers know what the problem is and how to solve it, but until
object interfaces, had no way to tell the storage system how to handle
their data.

These implementations use the active storage interfaces in Ceph. We
refer to this as \emph{storage programmability} which is a method by
which an application communicates its requirements to the storage system
in a way that allows the application to realize a new behavior without
sacrificing the correctness of the underlying system. This evaluation
and resulting implementation would have been a burden without object
classes.

\section{Re-Usable Components in
Ceph}\label{re-usable-components-in-ceph}

\label{background}

Ceph is a production-quality distributed system and in this section we
touch on some of the subsystems it uses to provide a general-purpose
storage system. In the next section, we describe how we leverage these
subsystems to build new services.

\subsection{Durability}\label{durability}

Ceph provides storage by striping and replicating data across RADOS
\autocite{weil_rados_2007}, the reliable distributed object store. RADOS
uses many techniques to ensure that data is not corrupted or lost, such
as erasure coding, replication, and data scrubbing. Furthermore, many of
these techniques try to be autonomous so that work is distributed across
the cluster. For example, when placement groups change the OSDs
rebalance and re-shard data in the background in a process called
placement group splitting.

\subsection{Consistency/Versioning of Cluster
State}\label{consistencyversioning-of-cluster-state}

Ceph needs to keep track of cluster state and it does this with separate
``services'': client authentication, logging, MDS maps, MON maps, OSD
maps and placement group (PG) maps. These services are all managed in
the MONs and they all talk to a PAXOS instance; this PAXOS instance goes
and talks to other PAXOS instances on other MONs so that they can all
agree on the correct version of the service; essentially they reach an
agreement about what the state of the cluster is.

\subsection{Active Storage}\label{active-storage}

Active storage attempts to push computation closer to the data by
exposing interfaces to execute code remotely. For example, the storage
server could process data before sending data over the network or
sorting it. The active storage component in Ceph is called ``object
interface classes'' (or \texttt{cls} for short). The Ceph object
interface classes can be loaded at runtime and customized with
user-defined functionality. The basic idea is shown in
Figure\textasciitilde{}\ref{fig:object-interfaces}, where the
\texttt{libcls\_md5.so} shared library performs the MD5 hash on an
object at the oSD instead of transferring data over the network. This
ability to carry out arbitrary operations on objects stored on OSDs
helps applications improve performance by optimizing things like network
round trips, data movement, and remote resources which may be idle.

Object storage interfaces are compiled into shared libraries and loaded
into a running OSD daemon using \texttt{dlopen()}. Because the shared
library has to link against symbols found in the running executable, the
interfaces are stored on the local file systems of each OSD so they can
be versioned and distributed alongside the same Ceph binaries. This
policy is for safety, since the OSDs could be on different distros that
provide different versioning capabilities. Also, this approach allows
bug fixes to co-evolve with the rest of the Ceph installation.

Although customizable OSD interfaces are powerful, the current
implementation has drawbacks. OSD interfaces are written in C/C++ and
compiled into a shared library, so the developer must account for
different target architectures. Second, C/C++ provides more
functionality than is necessary since the snippets of interface code
mainly set policies and perform simple operations. Finally, developing
these interfaces in C/C++ has high overhead. The developer must learn
how the OSD dynamically loads the shared libraries (\emph{e.g.}, OSDs
rely on a strict naming convention to find shared libraries), how to get
their interfaces compiled using the Ceph \texttt{make} system, and how
to debug issues that are not related to their specific interface
(\emph{e.g.}, the OSD cannot find the shared library) - this learning
curve is unacceptable for non-Ceph developers, especially since most
interfaces are one-off solutions specific to their applications.

Ceph has a whole infrastructure for getting dynamic code into the OSD.
The \texttt{ClassHandler} class safely opens object interfaces, executes
commands defined by the shared library, and fails gracefully with
helpful errors if anything goes wrong. Instead of injecting strings
directly into Ceph (like we did with the original Mantle
implementation), the \texttt{ClassHandler} class safely opens shared
code, even code with dependencies that are not in Ceph itself (e.g.,
Lua). With the \texttt{ClassHandler}, we get the safety and robustness
of loading dynamic code, the ease of transferring state between object
interfaces and the oSD internals, integration with testing and
correctness suites, and \texttt{struct}s for interface data and
handlers.

While we consider active storage to be an excellent example of
programmability, what separates our proposal from previous work is the
observation that so much \emph{more} of the storage system can be reused
to construct advanced, domain-specific interfaces.

\section{Malacology Implementation}\label{malacology-implementation}

\label{implementation}

To enable the programmability of Ceph subsystems, Malacology relies
heavily on interface classes for introducing new functionality into the
MDSs and OSDs. Malacology introduces 3 components to manage
object/metadata interfaces:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  additional interface hooks and classes for OSDs and MDSs using the
  existing Ceph \texttt{cls} infrastructure
\item
  durable interface classes using the Ceph object store
\item
  versioned and consistent interfaces using Ceph's consensus service
\end{enumerate}

In the following section, we will discuss the existing Ceph
infrastructure and the changes necessary for Malacology. Our focus will
be to demonstrate the viability and usefulness of the Malacology
appraoch with example implementations that due to limited time for the
submission of this paper are not necessarily fully optimized.

\subsection{Additional Interface Hooks and
Classes}\label{additional-interface-hooks-and-classes}

Malacology adds new hooks to Ceph and a new interface class for both the
MDS and OSD that uses Lua
\autocite{ierusalimschy:lua15}\autocite{ierusalimschy:spe96}. Our
framework has added a mechanism for defining and running object and
metadata balancer classes using Lua. Our Lua bindings expose functions
and symbols both ways; the host program can call functions defined in
Lua and the Lua scripts can call functions defined in native C++. These
bindings are merged upstream. We choose Lua for 4 reasons: performance,
portabability, size, and and security.

Lua is a fast scripting language. It was designed to be an embedded
language and the LuaJIT virtual machines boasts near-native performance
\autocite{pall:luajit15}\autocite{grawinkel_pdsw12}. Lua is frequently
used in game engines to set policies but we use it here because most of
the user-defined object classes in Ceph are policies as well (see
\S\ref{table:objclasses}). Separating policy from mechanism is a driving
factor in using Lua.

\begin{figure}[htbp]
\centering
\includegraphics{figures/programmability-framework.png}
\caption{Clients add functionality by injecting Lua scripts with a new
monitor (M) command. After the ``proposal'' (using PAXOS terminology) is
accepted by the cluster of monitors, the script is redundantly and
consistently distributed to OSDs. \label{fig:programmability-framework}}
\end{figure}

For object interfaces, clients send Lua classes to the OSD and then they
can invoke operations in that class remotely on the OSD. In reality,
this feature is a statically loaded object class written in C/C++ that
runs dynamically defined object interfaces, where clients access the
classes using an \texttt{exec()} wrapper. Although sending the class
with each client request is stateless (which has its own set of nice
features), it is costly for network bandwidth and difficult to organize
at the application level. Also, while the execute wrapper simplifies the
implementation, it burdens the applications, since they need to be
recompiled to use the \texttt{exec()} function.

Our framework has a mechanism for defining and running object and
metadata balancer classes with scripting. Scripting is useful in large
systems for 3 reasons: - accelerates deployment: tweaking policies, like
changing a threshold or metric calculation, does not require a full
recompile or system start-up. - transparency: if the bindings carefully
expose the correct metrics and functions, the user does not need to
learn the internal functions, variables, or types of the system. -
portability: scripts can be exchanged across platforms and system
versions without compatibility issues.

This effectively decouples policy from mechanism, which helps future
designers explore trade-offs and isolates policy development from
code-hardened systems. Decoupling policy mechanism is not a new
technique but designing the storage system with programmability as a
first-class citizen is credits policy/mechanism separation as the
biggest reason for the success of the block allocation mechanisms in the
Fast File System\textasciitilde{}\cite{mckusick:fast2015-FFS} yet there
are no mechanisms for transparently exposing and modifying the
underlying logic in modern file systems.

Malacology has Lua bindings that expose functions and symbols both ways;
the host program can call functions defined in Lua and the Lua scripts
can call functions defined in native C++. These bindings are merged
upstream. We choose Lua for 4 reasons: performance, portabability, size,
and security.

\subsubsection{Using Lua}\label{using-lua}

Lua is a performs well as an embeddable scripting language. Although it
is an interpreted language, the bindings are designed to make the
exchange of variables and functions amongst languages straightforward.
The LuaJIT virtual machine boasts near-native performance, making it a
logical choice for scriptability frameworks in systems research
\autocite{neto:dls14-luaos}. In storage systems, it has been effectively
used both on \autocites[
]{grawinkel:pdsw2012-lua}{watkins2013:bdmc13-in-vivo} and off
\autocite{sevilla:sc15-mantle} the critical path, where performance is
important. Lua is frequently used in game engines to set policies but we
use it here because most of the user-defined classes in Ceph are
policies as well! We do not want to provide specific implementations,
like pulling data from objects or transferring them over the network,
but instead strive to say \emph{what to do} with the data once we have
it. Separating policy from mechanism is a driving factor in using Lua.

Because Lua is interpreted, it is also portable. The portability allows
functions to be shipped around the cluster without having to compile or
reason about compatibility issues. Traditionally, intefaces are written
in C/C++, so they must be compiled into shared libraries before being
injected into the daemon at runtime. Shared libraries need to be
compiled on the host they will be run on to accomodate different
architectures and dependendencies. Embedding the Lua interpretator, on
the other hand, gives us the flexbility to store the actual Lua code in
other places, like RADOS, so we can enjoy all the other properties
provided by that particular back-end.

The other two advantages, that Lua leaves a small memory footprint and
that is is secure, are discussed at length in \autocites[
]{ierusalimschy_programming_2006}{neto:dls14-luaos}. In short, Lua
provides relatively robust sandbox capabilities for protecting the
system against some malicious cases, but does not protect the system
from poor policies. Since Malacology does not use these features (yet),
we will avoid discussion here.

\subsubsection{Generalizing the Lua VM}\label{generalizing-the-lua-vm}

We generalize the Lua VM since it will be used in both the OSD and MDS.
We put the core sandbox wrapper for the Lua object interface in a common
directory in Ceph and link against it. This core Lua wrapper contains
just the dependencies, symbols, and functions, need to run the Lua VM.
Some of these functions include \texttt{clslua\_log()} for transferring
logs to the daemon logs and \texttt{clslua\_pcall()} for calling Lua
functions.

Dependencies, symbols, and functions specific to the object or balancer
interface are put in the \texttt{cls} or \texttt{bal} directories,
respectively. For example, the Lua object interface uses functions that
have placement group filters, cryptography functions, and object
metadata (\emph{it i.e.} \texttt{cxx\_omap\_getvals()}, object data,
object extended attributes, and object versions -- all these are
dependencies, symbols, and functions are part of the OSD process but not
the MDS process. As a result, we put interface code that uses these in
\texttt{cls} directory. With this scheme, The OSD will \texttt{dlopen()}
the shared library created by Lua object interface shared library while
the MDS will \texttt{dlopen()} the Lua balancer interface shared
library; both shared libraries will the Lua core.

Both the object and balancer interfaces define functions and attach them
to the core Lua sandbox. For example, the Lua object storage interface
class attaches data, extended attribute, object map, and object version
functions, to the the Lua core functions; the exact functions are shown
in Listing\textasciitilde{}\ref{lst:attach_functions}. The advantages of
this is that we avoid duplicating code, we provide a framework for
putting Lua code in other parts of the system, and we remove components
and APIs that are too integrated with the OSD.

\subsubsection{Generalizing the Class
Handler}\label{generalizing-the-class-handler}

Our framework can also load Lua interfaces from the local file system --
the same technique as the C/C++ object interfaces. First, the OSD looks
for C/C++ shared library. If the OSD cannot find the file, it looks for
a Lua script of the same name. The script is read into a string of the
C++ object class; this string is later forwarded to the native Lua
class. When a requests comes in, the method name is passed with the
\texttt{exec()} function and the corresponding function in the Lua class
is called. Now clients need to only send their Lua object class once and
the OSDs will store them locally. Also, clients can execute any Lua
handler they want.

\subsection{Storing Interfaces in
RADOS}\label{storing-interfaces-in-rados}

For balancer interfaces, clients put balancers into RADOS and the CephFS
metadata balancer invokes the operations in the user-defined class
remotely on the MDS.

{[}7:31{]} Nothing would stop C++ from being stashed in objects and
recompiled on whatever platform they are being loaded on dynamically
(like a DB compiles each query). There just hasn't been a need for such
a feature. The next blog post, which I didn't at all allude to in the
one you are reading, shows how the Lua scripts can be put into the OSD
map and then monitors manage and version the script, distributing them
to each OSD. Stored in objects vs stored in the OSD map is a debate, I
guess. I am not sure which makes more sense.

\subsection{Versioning Interfaces with
MONs}\label{versioning-interfaces-with-mons}

We added a command,
\texttt{ceph\ osd\ pool\ set-class\ \textless{}pool\textgreater{}\ \textless{}class\textgreater{}\ \textless{}script\textgreater{}},
that the user uses to inject the Lua object interface into the cluster.
By leveraging the monitor daemons, we get the consistency from the
monitors' version management, the distribution from the monitors' data
structures (which are already distributed), and the durability from the
robustness of the monitors and persistence with Paxos. The
implementation uses the placement group pool data structure which
maintains the policies (\emph{e.g.}, erasure coding) and organization
details (\emph{e.g.}, snapshots) for each pool. While stuffing the
information into the monitor map, the structure that describes the the
monitor topology, was an option, the placement group pool allows finer
grainer control over different typoes of data. In regards to the
consistency, each time the user injects a new Lua object class it enters
the monitor quorum as a proposal; updates to the placement group pool
need to wait until accepted, at which point the state will be propogated
and versioned. We hooked up the monitor command to the \texttt{cls\_lua}
class by having the OSD pull the script from the placement group instead
of trying to dynamically load the shared library with \texttt{dlopen()}
or the Lua script.

Advantages:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  lets us store/manage interfaces
\item
  does most of the work (versioning, consistency, durabiltiy) for us
\item
  gives a new abstraction for dealing/mutating interfaces
\end{itemize}

The cls-lua branch lets users write object classes in Lua, the lua-rados
client library lets users write applications that talk to RADOS with
Lua, and the cls-client uses the lua-rados library to send Lua object
classes to the OSDs equipped with the LuaJIT VM.

\subsection{Specifying the Lua Class}\label{specifying-the-lua-class}

Maintain versions and consistency

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  cls-lua branch: write object classes in Lua
\item
  lua-rados: talk to RADOS with Lua
\item
  cls-client: use lua-rados + cls-lua branch to send Lua object classes
  to OSDs equipped with LuaJIT VM
\end{itemize}

\section{Services Built on
Malacology}\label{services-built-on-malacology}

\label{services}

\begin{table}
\begin{tabular}{ | l | l | l | l | l |}
\hline
Cap Owner & Failure & ZLog Recovery \\ \hline
mds & mds & mds \\ \hline
client & mds & xfer client cap \\ \hline
mds & client & n/a \\ \hline
client & client & mds \\
\hline
\end{tabular}
\caption{Failure and recovery scenarios.}
\label{t:mds-zlog-fail}
\end{table}

\subsection{Mantle: A Programmable Metadata Load
Balancer}\label{mantle-a-programmable-metadata-load-balancer}

Many distributed file systems decouple metadata and data I/O so that
these services can scale independently \autocites[
]{alam:pdsw2011-metadata-scaling}[ ]{ghemawat:sosp2003-gfs}[
]{hildebrand:msst2005-pnfs}[ ]{weil_ceph_2006}[
]{welch:fast2008-panasas}{shvachko:login2012-hdfs-scalability}. Despite
this optimization, scaling the metadata services is still difficult
because metadata accesses impose small and frequent requests on the
underlying storage syste \autocite{roselli:atec2000-FS-workloads}. Many
techniques for designing the metadata services have been proposed to
accommodate this workload: Lustre
\autocite{konstantinos:pdsw2014-lustre-metadata}, GFS
\autocite{ghemawat:sosp2003-gfs}, and HDFS
\autocite{shvachko:login2012-hdfs-scalability} keep all metadata on one
server; GIGA+ \autocite{patil:fast2011-giga}, IndexFS
\autocite{ren:sc2014-indexfs}, Lazy Hybrid
\autocite{brandt:msst2003-lh}, GPFS \autocite{schmuck:fast2002-gpfs},
and pNFS \autocite{hildebrand:supercomputing2006-pNFS} hash the file
system namespace across a dedicated cluster and cache metadata; Panasas
\autocite{welch:fast2008-panasas} and CephFS
\autocite{weil:sc2004-dyn-metadata} partition the file system namespace
into \# d assign them to servers. These systems have novel mechanisms
for Mantle \autocite{sevilla:sc15-mantle} is a programmable metadata
balancer that separates the metadata balancing policies from their
mechanisms. Admistrators inject code to change how the metadata cluster
distributes metadata. In the paper, we showed how to implement a single
node metadata service, a distributed metadata services with hashing, and
a distributed metadata service with dynamic subtree partitioning.

The Ceph team wants to merge Mantle because the scriptability is useful
for debugging, controlling the metadata balancer, and examining
trade-offs for different balancers. Unfortunately, this research quality
system is not as robust as Ceph and the Ceph team wants more safety,
durability, and consistency for the new functionality. For example, the
original Mantle API is fragile: the API is not enforced by the runtime,
the system allows injectable strings through the admin daemon,
constructing the Lua script in C++ is clunky, and the administrator can
inject really bad policies (e.g., while 1) that brings the whole system
down.

\begin{figure}[htbp]
\centering
\includegraphics{figures/overview-mantle.png}
\caption{This is Mantle.}
\end{figure}

\subsection{ZLog: A Distributed Shared Commit
Log}\label{zlog-a-distributed-shared-commit-log}

\begin{figure}[htbp]
\centering
\includegraphics{figures/overview-zlog.png}
\caption{This is ZLog.}
\end{figure}

\subsubsection{Consistency}\label{consistency}

\subsubsection{Striping Strategies}\label{striping-strategies}

\subsubsection{Tiering}\label{tiering}

Malacology uses the same Active and Typed Storage module presented in
DataMods \autocite{watkins_datamods_2012}; Asynchronous Service and File
Manifolds can be implemented with small changes to the Malacology
framework, namely asynchronous object calls and Lua stubs in the inode,
respectively.

\subsection{MDS as a sequencer}\label{mds-as-a-sequencer}

\subsubsection{Sequencer recovery with
MDS}\label{sequencer-recovery-with-mds}

\subsubsection{MDS Policy for Revoking
Capability}\label{mds-policy-for-revoking-capability}

The MDS has a queue of locks that it uses to revoke capabilities from
the clients. If we put a Mantle-style hook here, we can control when
those revoke messages go out.

\subsubsection{Mantle Balancing Policies to Migrate
Sequencers}\label{mantle-balancing-policies-to-migrate-sequencers}

\section{Evaluation}\label{evaluation}

1 Experiments: 2 3 1. mdtest separate directories 4 - just based on
default traffic 5 - kill one of the MDSs 6 7 2. filebench profiles 8 -
adaptable vs.~greedy vs.~fill\&spill 9 10 3. sequencer balancer 11 -
just based on default traffic 12 13 3. OpenSSL compile
\textasciitilde{}\\\label{evaluation}

\subsection{Mantle}\label{mantle}

\subsubsection{Experiment: FileBench}\label{experiment-filebench}

\subsubsection{Experiment: OpenSSL
Compile}\label{experiment-openssl-compile}

\subsection{MDS as ZLog Sequencer}\label{mds-as-zlog-sequencer}

\subsubsection{Experiment: Object Class
Programmability}\label{experiment-object-class-programmability}

Add quote from CORFU paper that talks about what they had to hack

\subsubsection{Experiment: Multi-Client
Burstiness}\label{experiment-multi-client-burstiness}

\begin{figure}[htbp]
\centering
\includegraphics{figures/caps-delay-thruput.png}
\caption{Forcing the client to drop their capabilities later (delay)
improves throughput}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/caps-delay-latency.png}
\caption{But if the clients hold their capabilities longer, it hurts
latency for everyone.}
\end{figure}

\subsubsection{CUT THIS Experiment: Client v. MDS caps +
Failures}\label{cut-this-experiment-client-v.-mds-caps-failures}

\subsubsection{Experiment: Mantle Sequencer
Balancer}\label{experiment-mantle-sequencer-balancer}

\section{Conclusion and Future Work}\label{conclusion-and-future-work}

Programmable storage is a viable method for eliminating duplication of
complex error prone software that are used as workarounds for storage
system deficiencies. However, this duplication has real-world problems
related to reliability. We propose that system expose their services in
a safe way allowing application developers to customize system behavior
to meet their needs while not sacrificing correctness.

We are intend to pursue this work towards the goal of constructing a set
of customization points that allow a wide variety of storage system
services to be configured on-the-fly in existing systems. This work is
one point along that path in which we have looked an a target
special-purpose storage system. Ultimately we want to utilize
declarative methods for expressing new services.

In conclusion, this paper should be accepted.

\section{Bibliography}\label{bibliography}

\end{document}

### 2016-05-08 paper/paper.tex
Adds preamble to latex file

@@
 \usepackage{times}
-\usepackage{fullpage}
-
-\begin{document}
+\usepackage{multirow}
+\usepackage{arydshln}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+    \usepackage{xltxtra,xunicode}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage{biblatex}
+\bibliography{refs}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={Malacology: A Programmable Storage System Built on Ceph},
+            colorlinks=true,
+            citecolor=blue,
+            urlcolor=blue,
+            linkcolor=black,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\setcounter{secnumdepth}{5}
 
@@
 \date{}
+\begin{document}
+
 \maketitle
-\thispagestyle{empty}
 
@@
 append to the end of the log. When designing the sequencer, we
-considered bolting on Zookeeper \cite{} as a service. Zookeeper would
+considered bolting on Zookeeper \autocite{} as a service. Zookeeper would
 satisfy our requirements by proiding a network and transaction stack but
@@
 storage systems should be safe from data loss, but would introduce
-unecessary scalability limits \cite{Neha HDFS namenode example}.
+unecessary scalability limits \autocite{Neha HDFS namenode example}.
 
@@
 biggest reason for the success of the block allocation mechanisms in the
-Fast File System\textasciitilde{}\cite{mckusick:fast2015-FFS} yet there
+Fast File System\textasciitilde{}\autocite{mckusick:fast2015-FFS} yet there
 are no mechanisms for transparently exposing and modifying the
@@
 
-\section{Bibliography}\label{bibliography}
+\printbibliography
 

### 2016-05-08 paper/paper.tex
paper: section 2
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-A consequence of complicated data management frameworks and faster
-devices is that the storage system cannot meet the needs of the
-general-purpose storage system. Workarounds for helping the application
-meet its performance goals roughly fall into one of three categories:
-``bolt-on'' services, application changes, and storage changes.
+A consequence of complicated data management requirements and faster storage
+devices is that storage systems may fail to meet the needs of applications.
+Application workarounds for meeting performance goals roughly fall into one of
+three categories: ``bolt-on'' services, application changes, and changes to
+the storage system itself.
 
@@
 
-So called ``bolt-on'' services are 3rd party systems that are integrated
-into the system to provide functionality and improve performance for
-common use-cases (e.g.~a metadata service). In addition to sacrificing
-simplicity, ``bolt-on'' services often duplicate functionality and
-execute redundant code, unnecessarily increasing the likelihood of bugs
-or worse, introducing unpredictable performance problems. For example,
-we are developing a distributed shared-commit log on Ceph called ZLog.
-ZLog uses a ``sequencer'' to distribute tokens to clients that want to
-append to the end of the log. When designing the sequencer, we
-considered bolting on Zookeeper \autocite{} as a service. Zookeeper would
-satisfy our requirements by proiding a network and transaction stack but
-it would perform poorly because Zookeeper persists its data and the
-sequencer can be in-memory and volatile. A Zookeeper sequencer would
-adhere to the pre-conceived notions of storage, namely that data in
-storage systems should be safe from data loss, but would introduce
-unecessary scalability limits \autocite{Neha HDFS namenode example}.
+So called ``bolt-on'' services are third-party solutions integrated into a
+system to provide functionality and improve performance for common use-cases
+(e.g. a metadata service). The introduction of additional services may provide
+simplicity and performance enhancements, but ``bolt-on`` services often
+duplicate functionality, unnecessarily increasing the likelihood of bugs or
+introducing unpredictable performance problems. For example, we are developing
+a high-performance distributed shared-log on Ceph called ZLog, based on the
+CORFU protocol~\cite{corfu}. Serialization at high-performance is achieved using a
+``sequencer'' service that maintains a volatile, in-memory counter used to assign a log
+position to each client that appends to the log. One option for building
+a distributed log is to use PAXOS, or an out-of-the-box solution such as
+Zookeeper. However, these systems adhere to the pre-coneived notions of
+storage, namely that data is safe from data loss, which is exactly the feature
+that causes traditional approaches to run at significantly reduced speed.
 
@@
 
-The second approach to adapting to a storage system deficiency is to
+The second approach that applications may use to adapt to a storage system
+limitation is to
 extend the responsibility of the application. This means changing the
 application itself by adding more data management intelligence or
-domain-specific middleware (e.g., a new data layout). For instance, an
-application may change itself to exploit data locality or I/O
+using domain-specific middleware (e.g., a new data layout). For instance, an
+application may be changed to exploit data locality or I/O
 parallelism in a distributed storage system.
@@
 application and the Hadoop framework itself to leverage the structure of
-scientific (3D array-based data, more specifically) to increase
+scientific data (3D array-based data, more specifically) to increase
 performance, locality, and the number of early results. This is not a
@@
 
-\subsection{Storage Changes}\label{storage-changes}
+\subsection{Storage System Changes}\label{storage-changes}
 
@@
 tunables are difficult to understand or quantify (e.g., Ceph tunables
-\autocite{sevilla:sc15-mantle}). To succesfully tune the system, the
+\autocite{sevilla:sc15-mantle}). To succesfully tune the system, a
 developer must have domain and system specific knowledge. Without this
@@
 
-Auto-tuning techniques attempt to find a good solution among a huge
-space of available system configurations. However, in practice
-auto-tuning is limited to only the configuration tunables that the
-storage system exposes (e.g.~block size) and can be overwhelmed with too
-many parameters. Starfish \autocite{herodotou:cidr2011-starfish} made an
-attempt of this for Hadoop but this technique would need to severely
-limit the space of parameters in order to feasible, similar to the
-approach used in \autocite{behzad:sc2013-autotuning}. For instance,
-auto-tuning may be capable of identifying instances in which new data
-layouts would benefit a workload, but unless the system can provide such
-a transformation, the option is left off the table.
+Auto-tuning techniques attempt to find a good solution among a huge space of
+available system configurations. However, in practice auto-tuning is limited
+to only the configuration tunables that the storage system exposes and can be
+overwhelmed with too many parameters.  For instance, auto-tuning may be
+capable of identifying instances in which new data layouts would benefit a
+workload, but unless the system can provide such a transformation, the option
+is left off the table.  Starfish \autocite{herodotou:cidr2011-starfish} made
+an attempt of this for Hadoop but this technique would need to severely limit
+the space of parameters in order to feasible, similar to the approach used in
+\autocite{behzad:sc2013-autotuning}.
 
@@
 
-As a last resort, the application developer can introduce changes to the
-storage system itself. This endeavour is especially difficult because
-the developer must first familiarate themselves with the storage system,
-and even if they manage to make the necessary changes, they must get
-their changes upstream for that specific project or become mantainer for
-their in-house version. Being a maintainer means they re-base their
-versions against new versions of the storage system and address any bugs
-for their branched code. A successful example of this is the work that
-focused on HDFS scalability for metadata-intensive workloads
+As a last resort, the application developer can alter the storage system
+itself. In the best case a developer is using an open-source system in which
+case one must become familiar with often enormous code-bases before being able
+to make changes, but then face the task of approaching a community to
+integrate changes, or be forced to maintain an in-house version and risk
+maintaining features that become incompatible.  A successful example of this
+is the work that focused on HDFS scalability for metadata-intensive workloads
 \autocite{shvachko_hdfs_2010}. This has lead to modifications to its
@@
 
-\begin{table}[ht]
-\caption{
-    A variety of RADOS object storage classes exist to expose interfaces
-    to applications. \# is the number of methods that use these categories.
-    \label{table:objclasses}}
-  \begin{center}
-  \begin{tabular}{l|l|l}
-    Category & Specialization& \# \\ \hline
-    \multirow{2}{*}{Locking} & Shared & \multirow{2}{*}{6} \\
-                             & Exclusive & \\ \hdashline
-    \multirow{3}{*}{Logging} & Replica & 3 \\
-                             & State & 4 \\
-                             & Timestamped & 4 \\ \hdashline
-    \multirow{4}{*}{Metadata Managment} 
-                             & RADOS Block Device  & 37 \\
-                             & RADOS Gatway & 27 \\
-                             & User & 5 \\
-                             & Version & 5 \\ \hdashline
-    Garbage Collection       & Reference Counting & 4 \\
-\end{tabular}
-\end{center}
-\end{table}
+Alternatively, some systems have formal support for introducing additional
+functionality, often in the form of a plugin infrastructure. Ceph provides the
+ability to create domain-specific logical object interfaces by composing
+existing interfaces. These interfaces are written in C++ and are statically
+loaded into the system. An example usage may be to atomically combine a
+``write'' with an index update to create a content-addressable object
+interface.
 
-\subsubsection{Hybrid Approaches}\label{hybrid-approaches}
-
-As an alternative, some systems allow developers to extend the
-interfaces of the storage devices to perform arbitrary operations on
-data stored on disk. In Ceph, object interfaces are C/C++ plugins that
-add new functionality to the OSDs. As an example use-case, ZLog needs to
+\being{comment}
+As an example use-case, ZLog needs to
 store metadata information with each object so for each operation, it
 checks the epoch value and write capability (2 metadata reads) and
-writes the index (metadata write). Figure 1 shows the throughput
+writes the index (metadata write).
+\end{comment}
+
+Figure 1 shows the throughput
 (\emph{y} axis) over time (\emph{x} axis) of two OSD implementations for
@@
 
-Object interface classese are used in production Ceph environments;
-Figure \ref{fig:obj-int-dev-growth} shows a dramatic growth in the use
-of co-designed interfaces in the Ceph community since 2010. Figure
-\ref{fig:obj-int-dev-churn} examines this growth in interfaces further
-by showing the lines of code that are changed (\emph{y} axis) over time
-(\emph{x} axis). The prevalence of blue dots indicates that users
-frequently update their interfaces close to release cycles and that
-massive changes come in bunches. This interface churn reflects both the
-popularity of object interfaces and the complexity of the processing
-being done by the OSDs. What is most remarkable about Figure
-\ref{fig:obj-int-dev-churn} is that this trend contradicts the notion
-that API changes are a burden for users. The types of object interfaces,
-which are enumerated in Table \ref{table:objclasses}, show that this
-active storage development is present throughough the Ceph software
-stack.
+The object interface facility provided by Ceph is more than an
+proof-of-concept active storage implementation.  Figure
+\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
+domain-specific interfaces in the Ceph community since 2010.  What is most
+remarkable is that this trend contradicts the notion that API changes are a
+burden for users. Rather it appears that a gap in existing interfaces are
+being addressed through programmability. In fact, Table~\ref{table:objclasses}
+categorizes existing interfaces and we clearly see a trend towards reusable
+services.
 
-\begin{figure}[htbp]
+\begin{figure}[ht]
 \centering
@@
 
+\begin{table}[ht]
+\centering
+  \begin{tabular}{l|l|l}
+    Category & Specialization& \# \\ \hline
+    \multirow{2}{*}{Locking} & Shared & \multirow{2}{*}{6} \\
+                             & Exclusive & \\ \hdashline
+    \multirow{3}{*}{Logging} & Replica & 3 \\
+                             & State & 4 \\
+                             & Timestamped & 4 \\ \hdashline
+    \multirow{4}{*}{Metadata Managment} 
+                             & RADOS Block Device  & 37 \\
+                             & RADOS Gatway & 27 \\
+                             & User & 5 \\
+                             & Version & 5 \\ \hdashline
+    Garbage Collection       & Reference Counting & 4 \\
+\end{tabular}
+\caption{A variety of RADOS object storage classes exist to expose interfaces
+    to applications. \# is the number of methods that use these categories.
+}
+\label{table:objclasses}
+\end{table}
+
+\begin{comment}
 \begin{figure}[htbp]
@@
 \end{figure}
+\end{comment}
 
@@
 Ceph provides storage by striping and replicating data across RADOS
-\autocite{weil_rados_2007}, the reliable distributed object store. RADOS
+~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
 uses many techniques to ensure that data is not corrupted or lost, such

### 2016-05-08 paper/paper.tex
deleted:    out/main.pdf 	modified:   paper.tex 	More changes in the intro

--- 2016-05-08 paper/paper.tex
@@
 Something of a storage swiss army knife, Ceph supports file, block, and
-object interfaces simultaneously in a single cluster. By introducing
-programmability concepts into Ceph, we can build new services by
-carefully exposing internal storage services to applications.
+object interfaces simultaneously in a single cluster. Ceph's Reliable Autonomous Distributed Object Storage (RADOS) system is a cluster of storage
+devices (OSDs) that provide Ceph with data durability and integrity
+using replication, erasure-coding, and scrubbing
+\autocite{weil_rados_2007}. 
+% By introducing
+% programmability concepts into Ceph, we can build new services by
+% carefully exposing internal storage services to applications.
 
@@
 
-For example, Ceph \autocite{weil_ceph_2006} contains a cluster of
+For example, a Ceph cluster contains a small number of
 monitoring nodes (MONs) that use PAXOS to maintain a consistent view of
 system-wide metadata such as data distribution parameters and cluster
-membership. The RADOS object storage system is a cluster of storage
-devices (OSDs) that provide Ceph with data durability and integrity
-using replication, erasure-coding, and scrubbing
-\autocite{weil_rados_2007}. The data distrubiton metadata maintained by
+membership. The data distribution metadata maintained by
 the cluster monitors is used by RADOS servers and clients to provide
@@
 locking and caching services provides by MDSs, durability and logical
-storage devices provides by RADOS, and propogation of consistent cluster
-state provided by the monitoring service. As shown in Figure
-\ref{fig:overview}, this framework is expressive enough to provide the
+storage devices provided by RADOS, and propogation of consistent cluster
+state provided by the monitoring service (see \ref{fig:overview}). As we will show in this paper, this framework is expressive enough to provide the
 functionality necessary for implementing new services. Our contributions
@@
 
+\paragraph{}
+
 \begin{itemize}
 \itemsep1pt\parskip0pt\parsep0pt
-\item
-  a programmable storage system implementation that re-uses and extends
-  existing abstractions. It includes:
+\item A programmable storage system implementation that re-uses and extends existing abstractions. It includes:
 
@@
   \item
-    inode abstractions for sequencer services
+    Extending file system metadata by a new type of inode that together with Ceph's capability subsystem implements high-performance, non-durable access serialization
   \item
-    consensus abstractions for policy versioning
+    Re-using Ceph's PAXOS-based consensus infrastructure for versioning of metadata service load-balancing policies
   \item
-    storage objects for persisting policies
+    Re-using storage objects for persisting policies
   \item
-    interface classes for new logical storage/metadata devices
+    Extending object interface classes for new logical storage/metadata devices
   \end{enumerate}
 \item
-  example systems that use this framework
+  Two example implementations and evaluations for feasibility of services that use this framework
 
@@
   \item
-    shared log service based on CORFU \autocite{balakrishnan_corfu_2012}
+    A high-performance distributed shared log service called Zlog which is based on CORFU~\autocite{balakrishnan_corfu_2012}
   \item
-    metadata load balancer based on Mantle
-    \autocite{sevilla:sc15-mantle}
+    A re-implementation of the programmable Mantle metadata load balancing service using Malacology~\autocite{sevilla:sc15-mantle}
+  \item 
+    An application of the Mantle metadata load balancing service to manage 
   \end{enumerate}

### 2016-05-08 paper/paper.tex
modified:   paper.tex 	Fixed cite and bibliography commands

@@
 }{}
-\usepackage{biblatex}
-\bibliography{refs}
+%\usepackage{biblatex}
+%\bibliography{paper}
 \usepackage{graphicx}
@@
 To illustrate the benefits and challenges of this approach we examine
-the programmability of Ceph \autocite{weil_ceph_2006}, the increasingly
+the programmability of Ceph~\cite{weil_ceph_2006}, the increasingly
 popular, production-level open-source distributed storage system.
@@
 devices (OSDs) that provide Ceph with data durability and integrity
-using replication, erasure-coding, and scrubbing
-\autocite{weil_rados_2007}. 
+using replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. 
 % By introducing
@@
   \item
-    A high-performance distributed shared log service called Zlog which is based on CORFU~\autocite{balakrishnan_corfu_2012}
+    A high-performance distributed shared log service called Zlog which is based on CORFU~\cite{balakrishnan_corfu_2012}
   \item
-    A re-implementation of the programmable Mantle metadata load balancing service using Malacology~\autocite{sevilla:sc15-mantle}
+    A re-implementation of the programmable Mantle metadata load balancing service using Malacology~\cite{sevilla:sc15-mantle}
   \item 
@@
 (e.g. a metadata service). The introduction of additional services may provide
-simplicity and performance enhancements, but ``bolt-on`` services often
+simplicity and performance enhancements, but ``bolt-on'' services often
 duplicate functionality, unnecessarily increasing the likelihood of bugs or
@@
 
-For example, SciHadoop \autocites[
-]{buck:hpc2011-scihadoop}{buck:sc2013-scidr} changes both the Hadoop
+For example, SciHadoop~\cite{buck:hpc2011-scihadoop,buck:sc2013-scidr} changes both the Hadoop
 application and the Hadoop framework itself to leverage the structure of
@@
 of parameters (e.g., Hadoop 2.7.1 has 210 tunables) and because the
-tunables are difficult to understand or quantify (e.g., Ceph tunables
-\autocite{sevilla:sc15-mantle}). To succesfully tune the system, a
+tunables are difficult to understand or quantify (e.g., Ceph tunables~\cite{sevilla:sc15-mantle}). To succesfully tune the system, a
 developer must have domain and system specific knowledge. Without this
@@
 workload, but unless the system can provide such a transformation, the option
-is left off the table.  Starfish \autocite{herodotou:cidr2011-starfish} made
+is left off the table.  Starfish~\cite{herodotou:cidr2011-starfish} made
 an attempt of this for Hadoop but this technique would need to severely limit
-the space of parameters in order to feasible, similar to the approach used in
-\autocite{behzad:sc2013-autotuning}.
+the space of parameters in order to feasible, similar to the approach used in~\cite{behzad:sc2013-autotuning}.
 
@@
 maintaining features that become incompatible.  A successful example of this
-is the work that focused on HDFS scalability for metadata-intensive workloads
-\autocite{shvachko_hdfs_2010}. This has lead to modifications to its
-architecture or API \autocite{balmin:sigmod2012-clydesdale} to improve
+is the work that focused on HDFS scalability for metadata-intensive workloads~\cite{shvachko_hdfs_2010}. This has lead to modifications to its
+architecture or API~\cite{balmin:sigmod2012-clydesdale} to improve
 performance.
@@
 
-\being{comment}
+\if{false}
 As an example use-case, ZLog needs to
@@
 writes the index (metadata write).
-\end{comment}
+\fi
 
@@
 
-\begin{comment}
+\if{false}
 \begin{figure}[htbp]
@@
 \end{figure}
-\end{comment}
+\fi
 
@@
 
-Ceph provides storage by striping and replicating data across RADOS
-~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
+Ceph provides storage by striping and replicating data across RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
 uses many techniques to ensure that data is not corrupted or lost, such
@@
 Malacology adds new hooks to Ceph and a new interface class for both the
-MDS and OSD that uses Lua
-\autocite{ierusalimschy:lua15}\autocite{ierusalimschy:spe96}. Our
+MDS and OSD that uses Lua~\cite{ierusalimschy:lua15,ierusalimschy:spe96}. Our
 framework has added a mechanism for defining and running object and
@@
 Lua is a fast scripting language. It was designed to be an embedded
-language and the LuaJIT virtual machines boasts near-native performance
-\autocite{pall:luajit15}\autocite{grawinkel_pdsw12}. Lua is frequently
+language and the LuaJIT virtual machines boasts near-native performance~\cite{pall:luajit15,grawinkel_pdsw12}. Lua is frequently
 used in game engines to set policies but we use it here because most of
@@
 biggest reason for the success of the block allocation mechanisms in the
-Fast File System\textasciitilde{}\autocite{mckusick:fast2015-FFS} yet there
+Fast File System~\cite{mckusick:fast2015-FFS} yet there
 are no mechanisms for transparently exposing and modifying the
@@
 The LuaJIT virtual machine boasts near-native performance, making it a
-logical choice for scriptability frameworks in systems research
-\autocite{neto:dls14-luaos}. In storage systems, it has been effectively
-used both on \autocites[
-]{grawinkel:pdsw2012-lua}{watkins2013:bdmc13-in-vivo} and off
-\autocite{sevilla:sc15-mantle} the critical path, where performance is
+logical choice for scriptability frameworks in systems research~\cite{neto:dls14-luaos}. In storage systems, it has been effectively
+used both on \cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo} and off
+\cite{sevilla:sc15-mantle} the critical path, where performance is
 important. Lua is frequently used in game engines to set policies but we
@@
 The other two advantages, that Lua leaves a small memory footprint and
-that is is secure, are discussed at length in \autocites[
-]{ierusalimschy_programming_2006}{neto:dls14-luaos}. In short, Lua
+that is is secure, are discussed at length in~\cite{ierusalimschy_programming_2006,neto:dls14-luaos}. In short, Lua
 provides relatively robust sandbox capabilities for protecting the
@@
 Many distributed file systems decouple metadata and data I/O so that
-these services can scale independently \autocites[
-]{alam:pdsw2011-metadata-scaling}[ ]{ghemawat:sosp2003-gfs}[
-]{hildebrand:msst2005-pnfs}[ ]{weil_ceph_2006}[
-]{welch:fast2008-panasas}{shvachko:login2012-hdfs-scalability}. Despite
+these services can scale independently~\cite{alam:pdsw2011-metadata-scaling,ghemawat:sosp2003-gfs,hildebrand:msst2005-pnfs,weil_ceph_2006,welch:fast2008-panasas,shvachko:login2012-hdfs-scalability}.
+Despite
 this optimization, scaling the metadata services is still difficult
 because metadata accesses impose small and frequent requests on the
-underlying storage syste \autocite{roselli:atec2000-FS-workloads}. Many
+underlying storage syste~\cite{roselli:atec2000-FS-workloads}. Many
 techniques for designing the metadata services have been proposed to
-accommodate this workload: Lustre
-\autocite{konstantinos:pdsw2014-lustre-metadata}, GFS
-\autocite{ghemawat:sosp2003-gfs}, and HDFS
-\autocite{shvachko:login2012-hdfs-scalability} keep all metadata on one
-server; GIGA+ \autocite{patil:fast2011-giga}, IndexFS
-\autocite{ren:sc2014-indexfs}, Lazy Hybrid
-\autocite{brandt:msst2003-lh}, GPFS \autocite{schmuck:fast2002-gpfs},
-and pNFS \autocite{hildebrand:supercomputing2006-pNFS} hash the file
-system namespace across a dedicated cluster and cache metadata; Panasas
-\autocite{welch:fast2008-panasas} and CephFS
-\autocite{weil:sc2004-dyn-metadata} partition the file system namespace
+accommodate this workload: Lustre~\cite{konstantinos:pdsw2014-lustre-metadata}, GFS~\cite{ghemawat:sosp2003-gfs}, and HDFS~\cite{shvachko:login2012-hdfs-scalability} keep all metadata on one
+server; GIGA+~\cite{patil:fast2011-giga}, IndexFS~\cite{ren:sc2014-indexfs}, Lazy Hybrid~\cite{brandt:msst2003-lh}, GPFS~\cite{schmuck:fast2002-gpfs},
+and pNFS~\cite{hildebrand:supercomputing2006-pNFS} hash the file
+system namespace across a dedicated cluster and cache metadata; Panasas~\cite{welch:fast2008-panasas} and CephFS~\cite{weil:sc2004-dyn-metadata} partition the file system namespace
 into \# d assign them to servers. These systems have novel mechanisms
-for Mantle \autocite{sevilla:sc15-mantle} is a programmable metadata
+for Mantle~\cite{sevilla:sc15-mantle} is a programmable metadata
 balancer that separates the metadata balancing policies from their
@@
 Malacology uses the same Active and Typed Storage module presented in
-DataMods \autocite{watkins_datamods_2012}; Asynchronous Service and File
+DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File
 Manifolds can be implemented with small changes to the Malacology
@@
 
-\printbibliography
+\bibliography{paper}
+\bibliographystyle{plain}
+%\printbibliography
 

### 2016-05-08 paper/paper.tex
full page package
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>
modified:   paper.tex 	More changes to the intro

@@
   \item
-    Extending file system metadata by a new type of inode that together with Ceph's capability subsystem implements high-performance, non-durable access serialization
+    Extending file system metadata by a new type of ``sequencer'' inode that together with Ceph's capability subsystem implements high-performance, non-durable access serialization
   \item
@@
   \item 
-    An application of the Mantle metadata load balancing service to manage 
+    An application of the Mantle metadata load balancing service to manage the placement of new sequencer inodes.
   \end{enumerate}

### 2016-05-09 paper/paper.tex
sec 2 updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>
Adds Section 4

@@
 
+\usepackage{fullpage}
 \usepackage{times}
@@
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
 }{}
-%\usepackage{biblatex}
-%\bibliography{paper}
+
 \usepackage{graphicx}
+\usepackage[small]{caption}
+\usepackage{pstricks}
 \makeatletter
@@
 The object interface facility provided by Ceph is more than an
-proof-of-concept active storage implementation.  Figure
+a proof-of-concept active storage implementation.  Figure
 \ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
@@
 remarkable is that this trend contradicts the notion that API changes are a
-burden for users. Rather it appears that a gap in existing interfaces are
+burden for users.
+Rather it appears that a gap in existing interfaces are
 being addressed through programmability. In fact, Table~\ref{table:objclasses}
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/implementation.png}
+\caption{Malacology re-uses Ceph subsystems to manage interface classes.
+command. After the proposal is accepted by the cluster of monitors, the
+script is redundantly and consistently distributed to OSDs.
+\label{fig:implementation}}
+\end{figure}
+
 \section{Malacology Implementation}\label{malacology-implementation}
@@
 
-To enable the programmability of Ceph subsystems, Malacology relies
-heavily on interface classes for introducing new functionality into the
-MDSs and OSDs. Malacology introduces 3 components to manage
-object/metadata interfaces:
+To enable the re-use of Ceph subsystems, Malacology relies heavily on
+interface classes for introducing new functionality into the MDSs and
+OSDs. Malacology introduces 3 components to manage object/metadata
+interfaces:
 
@@
 \item
-  versioned and consistent interfaces using Ceph's consensus service
+  versioned and consistent interfaces using the Ceph MONs
 \end{enumerate}
 
-In the following section, we will discuss the existing Ceph
-infrastructure and the changes necessary for Malacology. Our focus will
-be to demonstrate the viability and usefulness of the Malacology
-appraoch with example implementations that due to limited time for the
-submission of this paper are not necessarily fully optimized.
+Figure \ref{implementation} shows how interfaces can be loaded into
+either the MDS or OSD (1), how interfaces are persisted to object
+storage (2), and how they are versioned by the MONs (3). In the
+following section, we will discuss the existing Ceph infrastructure and
+the changes necessary for Malacology. We emphasize here that these
+implementations are not necessarily optimized but they help demonstrate
+the power of programmability.
 
-\subsection{Additional Interface Hooks and
-Classes}\label{additional-interface-hooks-and-classes}
-
-Malacology adds new hooks to Ceph and a new interface class for both the
-MDS and OSD that uses Lua~\cite{ierusalimschy:lua15,ierusalimschy:spe96}. Our
-framework has added a mechanism for defining and running object and
-metadata balancer classes using Lua. Our Lua bindings expose functions
-and symbols both ways; the host program can call functions defined in
-Lua and the Lua scripts can call functions defined in native C++. These
-bindings are merged upstream. We choose Lua for 4 reasons: performance,
-portabability, size, and and security.
-
-Lua is a fast scripting language. It was designed to be an embedded
-language and the LuaJIT virtual machines boasts near-native performance~\cite{pall:luajit15,grawinkel_pdsw12}. Lua is frequently
-used in game engines to set policies but we use it here because most of
-the user-defined object classes in Ceph are policies as well (see
-\S\ref{table:objclasses}). Separating policy from mechanism is a driving
-factor in using Lua.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/programmability-framework.png}
-\caption{Clients add functionality by injecting Lua scripts with a new
-monitor (M) command. After the ``proposal'' (using PAXOS terminology) is
-accepted by the cluster of monitors, the script is redundantly and
-consistently distributed to OSDs. \label{fig:programmability-framework}}
-\end{figure}
+\subsection{Additional Interface
+Frameworks}\label{additional-interface-frameworks}
 
@@
 
-Our framework has a mechanism for defining and running object and
-metadata balancer classes with scripting. Scripting is useful in large
-systems for 3 reasons: - accelerates deployment: tweaking policies, like
-changing a threshold or metric calculation, does not require a full
-recompile or system start-up. - transparency: if the bindings carefully
-expose the correct metrics and functions, the user does not need to
-learn the internal functions, variables, or types of the system. -
+Although these OSD interfaces are powerful, the current implementation
+has drawbacks. OSD interfaces are written in C/C++ and compiled into a
+shared library, so the developer must account for different target
+architectures. Second, C/C++ provides more functionality than is
+necessary since the snippets of interface code mainly set policies and
+perform simple operations. Finally, developing these interfaces in C/C++
+has high overhead. The developer must learn how the OSD dynamically
+loads the shared libraries (\emph{e.g.}, OSDs rely on a strict naming
+convention to find shared libraries), how to get their interfaces
+compiled using the Ceph \texttt{make} system, and how to debug issues
+that are not related to their specific interface (\emph{e.g.}, the OSD
+cannot find the shared library) - this learning curve is unacceptable
+for non-Ceph developers, especially since most interfaces are one-off
+solutions specific to their applications.
+
+\subsubsection{New Interface Classes}\label{new-interface-classes}
+
+Our framework has added a mechanism for defining and running object and
+metadata balancer classes using the Lua scripting language. Scripting is
+useful in large systems for 3 reasons:
+
+\begin{itemize}
+\item
+  accelerates deployment: tweaking policies, like changing a threshold
+  or metric calculation, does not require a full recompile or system
+  start-up.
+\item
+  transparency: if the bindings carefully expose the correct metrics and
+  functions, the user does not need to learn the internal functions,
+  variables, or types of the system.
+\item
   portability: scripts can be exchanged across platforms and system
   versions without compatibility issues.
+\end{itemize}
 
-This effectively decouples policy from mechanism, which helps future
-designers explore trade-offs and isolates policy development from
-code-hardened systems. Decoupling policy mechanism is not a new
-technique but designing the storage system with programmability as a
-first-class citizen is credits policy/mechanism separation as the
-biggest reason for the success of the block allocation mechanisms in the
-Fast File System~\cite{mckusick:fast2015-FFS} yet there
-are no mechanisms for transparently exposing and modifying the
-underlying logic in modern file systems.
+Our Lua bindings expose functions and symbols both ways; the host
+program can call functions defined in Lua and the Lua scripts can call
+functions defined in native C++. These bindings are merged upstream. Lua
+is frequently used in game engines to set policies but we use it here
+because most of the user-defined classes in Ceph are policies as well!
+We do not want to provide specific implementations, like pulling data
+from objects or transferring them over the network, but instead strive
+to say \emph{what to do} with the data once we have it. Separating
+policy from mechanism is a driving factor in using Lua. We choose Lua
+for 4 reasons: performance, portabability, size, and security.
 
-Malacology has Lua bindings that expose functions and symbols both ways;
-the host program can call functions defined in Lua and the Lua scripts
-can call functions defined in native C++. These bindings are merged
-upstream. We choose Lua for 4 reasons: performance, portabability, size,
-and security.
-
-\subsubsection{Using Lua}\label{using-lua}
-
-Lua is a performs well as an embeddable scripting language. Although it
-is an interpreted language, the bindings are designed to make the
-exchange of variables and functions amongst languages straightforward.
-The LuaJIT virtual machine boasts near-native performance, making it a
-logical choice for scriptability frameworks in systems research~\cite{neto:dls14-luaos}. In storage systems, it has been effectively
-used both on \cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo} and off
-\cite{sevilla:sc15-mantle} the critical path, where performance is
-important. Lua is frequently used in game engines to set policies but we
-use it here because most of the user-defined classes in Ceph are
-policies as well! We do not want to provide specific implementations,
-like pulling data from objects or transferring them over the network,
-but instead strive to say \emph{what to do} with the data once we have
-it. Separating policy from mechanism is a driving factor in using Lua.
+Lua performs well as an embeddable scripting language. Although it is an
+interpreted language, the bindings are designed to make the exchange of
+variables and functions amongst languages straightforward. The LuaJIT
+virtual machine boasts near-native performance, making it a logical
+choice for scriptability frameworks in systems research
+\autocite{neto:dls14-luaos}. In storage systems, it has been effectively
+used both on \autocites[
+]{grawinkel:pdsw2012-lua}{watkins2013:bdmc13-in-vivo} and off
+\autocite{sevilla:sc15-mantle} the critical path, where performance is
+important.
 
@@
 The other two advantages, that Lua leaves a small memory footprint and
-that is is secure, are discussed at length in~\cite{ierusalimschy_programming_2006,neto:dls14-luaos}. In short, Lua
+that it is secure, are discussed at length in \autocites[
+]{ierusalimschy_programming_2006}{neto:dls14-luaos}. In short, Lua
 provides relatively robust sandbox capabilities for protecting the
@@
 
-\subsubsection{Generalizing the Lua VM}\label{generalizing-the-lua-vm}
+\subsection{New Interface Hooks}\label{new-interface-hooks}
+
+Malacogy adds interface hooks to the MDS in the load balancing and
+capability subsystems. In contrast to the Mantle implementation, which
+had a large function hard-coded into the balancer subsystem, Malacology
+embeds code in the MDS using OSD subsystems. Ceph has a whole
+infrastructure for getting dynamic code into the OSD. The
+\texttt{ClassHandler} class safely opens object interfaces, executes
+commands defined by the shared library, and fails gracefully with
+helpful errors if anything goes wrong. Instead of injecting strings
+directly into Ceph (like we did with the original Mantle
+implementation), the \texttt{ClassHandler} class safely opens shared
+code, even code with dependencies that are not in Ceph itself (e.g.,
+Lua).
+
+With the \texttt{ClassHandler}, the MDS gets the safety and robustness
+of loading dynamic code, the ease of transferring state between object
+interfaces and the OSD internals, integration with testing and
+correctness suites, and \texttt{struct}s for interface data and
+handlers. Using the OSD class handler hooks also gives the MDS the
+ability to load interfaces from the local file system.
 
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/cls-osd-mds.png}
+\caption{To generalize the LuaVM for use in both the OSD and MDS,
+Malacology detaches the core functionality into a Lua core module. The
+VM is dynamically loaded first (1) and then interfaces can execute (2).
+\label{fig:cls-osd-mds}}
+\end{figure}
+
 Both the object and balancer interfaces define functions and attach them
@@
 class attaches data, extended attribute, object map, and object version
-functions, to the the Lua core functions; the exact functions are shown
-in Listing\textasciitilde{}\ref{lst:attach_functions}. The advantages of
-this is that we avoid duplicating code, we provide a framework for
-putting Lua code in other parts of the system, and we remove components
-and APIs that are too integrated with the OSD.
-
-\subsubsection{Generalizing the Class
-Handler}\label{generalizing-the-class-handler}
-
-Our framework can also load Lua interfaces from the local file system --
-the same technique as the C/C++ object interfaces. First, the OSD looks
-for C/C++ shared library. If the OSD cannot find the file, it looks for
-a Lua script of the same name. The script is read into a string of the
-C++ object class; this string is later forwarded to the native Lua
-class. When a requests comes in, the method name is passed with the
-\texttt{exec()} function and the corresponding function in the Lua class
-is called. Now clients need to only send their Lua object class once and
-the OSDs will store them locally. Also, clients can execute any Lua
-handler they want.
+functions, to the Lua core functions, as shown in Figure
+\ref{fig:cls-osd-mds}. The advantages of this is that we avoid
+duplicating code, we provide a framework for putting Lua code in other
+parts of the system, and we remove components and APIs that are too
+integrated with the OSD.
 
@@
 
-For balancer interfaces, clients put balancers into RADOS and the CephFS
-metadata balancer invokes the operations in the user-defined class
-remotely on the MDS.
+Traditionally, object storage interfaces are compiled into shared
+libraries and loaded into a running OSD daemon using \texttt{dlopen()}.
+Because the shared library has to link against symbols found in the
+running executable, the interfaces are stored on the local file systems
+of each OSD so they can be versioned and distributed alongside the same
+Ceph binaries. This policy is for safety, since the OSDs could be on
+different distros that provide different versioning capabilities. Also,
+this approach allows bug fixes to co-evolve with the rest of the Ceph
+installation.
 
-{[}7:31{]} Nothing would stop C++ from being stashed in objects and
-recompiled on whatever platform they are being loaded on dynamically
-(like a DB compiles each query). There just hasn't been a need for such
-a feature. The next blog post, which I didn't at all allude to in the
-one you are reading, shows how the Lua scripts can be put into the OSD
-map and then monitors manage and version the script, distributing them
-to each OSD. Stored in objects vs stored in the OSD map is a debate, I
-guess. I am not sure which makes more sense.
+Because of the portability of Lua and because the Lua virtual machine is
+compiled and loaded as a shared library, the MDS/OSD can store the
+balancers in a more redundant location. For our implementation, we store
+balancers in RADOS and the daemons first search RADOS, then the local
+file system for their interface classes.
 
@@
 
-We added a command,
-\texttt{ceph\ osd\ pool\ set-class\ \textless{}pool\textgreater{}\ \textless{}class\textgreater{}\ \textless{}script\textgreater{}},
-that the user uses to inject the Lua object interface into the cluster.
-By leveraging the monitor daemons, we get the consistency from the
-monitors' version management, the distribution from the monitors' data
+We added two new commands to Ceph command line client so the user can
+inject interface versions into the cluster::
+
+\noindent \texttt{ceph\ osd\ pool\ set-class\ \textless{}pool\textgreater{}\ \textless{}class\textgreater{}\ \textless{}script\textgreater{}}
+
+\noindent \texttt{ceph\ mds\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}
+
+By leveraging the monitor daemons, our services get the consistency from
+the monitors version management, the distribution from the monitors data
 structures (which are already distributed), and the durability from the
@@
 monitor topology, was an option, the placement group pool allows finer
-grainer control over different typoes of data. In regards to the
-consistency, each time the user injects a new Lua object class it enters
-the monitor quorum as a proposal; updates to the placement group pool
-need to wait until accepted, at which point the state will be propogated
-and versioned. We hooked up the monitor command to the \texttt{cls\_lua}
-class by having the OSD pull the script from the placement group instead
-of trying to dynamically load the shared library with \texttt{dlopen()}
-or the Lua script.
+grain control over different types of data.
 
-Advantages:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  lets us store/manage interfaces
-\item
-  does most of the work (versioning, consistency, durabiltiy) for us
-\item
-  gives a new abstraction for dealing/mutating interfaces
-\end{itemize}
+In regards to the consistency, each time the user injects a new Lua
+object class it enters the monitor quorum as a proposal; updates to the
+placement group pool need to wait until accepted, at which point the
+state will be propogated and versioned. We hooked up the monitor command
+to the \texttt{cls\_lua} class by having the OSD pull the script from
+the placement group instead of trying to dynamically load the shared
+library with \texttt{dlopen()} or the Lua script.
 
@@
 classes to the OSDs equipped with the LuaJIT VM.
-
-\subsection{Specifying the Lua Class}\label{specifying-the-lua-class}
-
-Maintain versions and consistency
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  cls-lua branch: write object classes in Lua
-\item
-  lua-rados: talk to RADOS with Lua
-\item
-  cls-client: use lua-rados + cls-lua branch to send Lua object classes
-  to OSDs equipped with LuaJIT VM
-\end{itemize}
 

### 2016-05-09 paper/paper.tex
Additional packages to help reduce white space
Removes autocites

@@
 choice for scriptability frameworks in systems research
-\autocite{neto:dls14-luaos}. In storage systems, it has been effectively
-used both on \autocites[
-]{grawinkel:pdsw2012-lua}{watkins2013:bdmc13-in-vivo} and off
-\autocite{sevilla:sc15-mantle} the critical path, where performance is
+\cite{neto:dls14-luaos}. In storage systems, it has been effectively
+used both on \cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo} and off
+\cite{sevilla:sc15-mantle} the critical path, where performance is
 important.
@@
 The other two advantages, that Lua leaves a small memory footprint and
-that it is secure, are discussed at length in \autocites[
-]{ierusalimschy_programming_2006}{neto:dls14-luaos}. In short, Lua
+that it is secure, are discussed at length in 
+\cite{ierusalimschy_programming_2006,neto:dls14-luaos}. In short, Lua
 provides relatively robust sandbox capabilities for protecting the

### 2016-05-09 paper/paper.tex
logical storage devices
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-\subsection{Durability}\label{durability}
+\subsection{Durability}
+\label{sec:durability}
 
@@
 
-\subsection{Active Storage}\label{active-storage}
+\subsection{Logical Storage Devices}
+\label{active-storage}
 
-Active storage attempts to push computation closer to the data by
-exposing interfaces to execute code remotely. For example, the storage
-server could process data before sending data over the network or
-sorting it. The active storage component in Ceph is called ``object
-interface classes'' (or \texttt{cls} for short). The Ceph object
-interface classes can be loaded at runtime and customized with
-user-defined functionality. The basic idea is shown in
-Figure\textasciitilde{}\ref{fig:object-interfaces}, where the
-\texttt{libcls\_md5.so} shared library performs the MD5 hash on an
-object at the oSD instead of transferring data over the network. This
-ability to carry out arbitrary operations on objects stored on OSDs
-helps applications improve performance by optimizing things like network
-round trips, data movement, and remote resources which may be idle.
+Ceph provides a rich plugin framework for defining new interfaces to objects
+stored within RADOS, effectively allowing developers to use co-design
+techniques to define logical storage devices tailored to their application
+domain.
 
-Object storage interfaces are compiled into shared libraries and loaded
-into a running OSD daemon using \texttt{dlopen()}. Because the shared
-library has to link against symbols found in the running executable, the
-interfaces are stored on the local file systems of each OSD so they can
-be versioned and distributed alongside the same Ceph binaries. This
-policy is for safety, since the OSDs could be on different distros that
-provide different versioning capabilities. Also, this approach allows
-bug fixes to co-evolve with the rest of the Ceph installation.
+An object interface plugin is structured in a similar way to that of an RPC in
+which a developer creates a named function within the cluster that clients may
+invoke upon an object. In Ceph these functions are defined by a composition of
+existing native interfaces that include reading and writing bytestream data
+and controlling snapshots and cloning, all of which may be transactionally
+composed with full access to a sorted key-value service allowing applications
+to fine tune storage interfaces, such as using a read-modify-write guard to
+protect shared data.
 
-Although customizable OSD interfaces are powerful, the current
-implementation has drawbacks. OSD interfaces are written in C/C++ and
-compiled into a shared library, so the developer must account for
-different target architectures. Second, C/C++ provides more
-functionality than is necessary since the snippets of interface code
-mainly set policies and perform simple operations. Finally, developing
-these interfaces in C/C++ has high overhead. The developer must learn
-how the OSD dynamically loads the shared libraries (\emph{e.g.}, OSDs
-rely on a strict naming convention to find shared libraries), how to get
-their interfaces compiled using the Ceph \texttt{make} system, and how
-to debug issues that are not related to their specific interface
-(\emph{e.g.}, the OSD cannot find the shared library) - this learning
-curve is unacceptable for non-Ceph developers, especially since most
-interfaces are one-off solutions specific to their applications.
-
-Ceph has a whole infrastructure for getting dynamic code into the OSD.
-The \texttt{ClassHandler} class safely opens object interfaces, executes
-commands defined by the shared library, and fails gracefully with
-helpful errors if anything goes wrong. Instead of injecting strings
-directly into Ceph (like we did with the original Mantle
-implementation), the \texttt{ClassHandler} class safely opens shared
-code, even code with dependencies that are not in Ceph itself (e.g.,
-Lua). With the \texttt{ClassHandler}, we get the safety and robustness
-of loading dynamic code, the ease of transferring state between object
-interfaces and the oSD internals, integration with testing and
-correctness suites, and \texttt{struct}s for interface data and
-handlers.
-
-While we consider active storage to be an excellent example of
-programmability, what separates our proposal from previous work is the
-observation that so much \emph{more} of the storage system can be reused
-to construct advanced, domain-specific interfaces.
+Although this feature is powerful, the current implementation has drawbacks.
+Supporting only C/C++ for object interface developers, Ceph requires
+distribution of compiled binaries for the correct architecture, adding a
+large barrier of entry for developers and system administrators. Second,
+having no way to dynamically unload modules, any changes require a full
+restart of an OSD which may have serious performance impacts due to a loss
+of cached data. And finally, the security limitations of the framework
+limit the use of object classes to all but those with administrative level
+access and technique expertise.
 
@@
 
-Malacology uses the same Active and Typed Storage module presented in
-DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File
-Manifolds can be implemented with small changes to the Malacology
-framework, namely asynchronous object calls and Lua stubs in the inode,
-respectively.
-
 \subsection{MDS as a sequencer}\label{mds-as-a-sequencer}
@@
 
+\section{Related Work}
+
+Malacology uses the same Active and Typed Storage module presented in
+DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File
+Manifolds can be implemented with small changes to the Malacology
+framework, namely asynchronous object calls and Lua stubs in the inode,
+respectively.
+
 \section{Conclusion and Future Work}\label{conclusion-and-future-work}

### 2016-05-09 paper/paper.tex
sec 3
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-\section{Re-Usable Components in
-Ceph}\label{re-usable-components-in-ceph}
+\section{Re-Usable Components in Ceph}\label{re-usable-components-in-ceph}
 
@@
 storage system. In the next section, we describe how we leverage these
-subsystems to build new services.
+subsystems to build new services. This is the an exhuastive list of services.
+
+\subsection{Cluster State Management}
+\label{consistencyversioning-of-cluster-state}
+
+A consistent view of cluster state among Ceph daemons and clients is critical
+to provide strong consistency guarnatees to clients.  Ceph maintains cluster
+state information in per-subsystem ``maps'' (e.g. OSD map, MDS map) that
+record membership and status information. A PAXOS-based monitoring service is
+responsible integrating state changes into cluster maps, and responding to
+requests from out-of-date clients.  As a fundamental building block of many
+system designs, consensus abstractions such as PAXOS are a common technique
+for maintaining consistent data versions, and are a useful system to expose.
+While Ceph does not expose this service directly, it does allow limited use by
+clients through a key-value service intended for low-volume, small size
+configuration information.
 
@@
 
-Ceph provides storage by striping and replicating data across RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-uses many techniques to ensure that data is not corrupted or lost, such
-as erasure coding, replication, and data scrubbing. Furthermore, many of
-these techniques try to be autonomous so that work is distributed across
-the cluster. For example, when placement groups change the OSDs
-rebalance and re-shard data in the background in a process called
-placement group splitting.
+Ceph provides storage by striping and replicating data across
+RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
+uses many techniques to ensure that data is not corrupted or lost, such as
+erasure coding, replication, and data scrubbing. Furthermore, many of these
+techniques try to be autonomous so that work is distributed across the
+cluster. For example, when placement groups change the OSDs rebalance and
+re-shard data in the background in a process called placement group splitting.
 
-\subsection{Consistency/Versioning of Cluster
-State}\label{consistencyversioning-of-cluster-state}
-
-Ceph needs to keep track of cluster state and it does this with separate
-``services'': client authentication, logging, MDS maps, MON maps, OSD
-maps and placement group (PG) maps. These services are all managed in
-the MONs and they all talk to a PAXOS instance; this PAXOS instance goes
-and talks to other PAXOS instances on other MONs so that they can all
-agree on the correct version of the service; essentially they reach an
-agreement about what the state of the cluster is.
+In order to reduce load on the monitoring service, Ceph OSDs use a gossip
+protocol to efficiently propogate changes to cluster maps throughout the
+system, and automonmously initiate recovery mechansims when failures are
+discovered.
 
@@
 
-\section{Malacology Implementation}\label{malacology-implementation}
+\subsection{File System Indexing and Locking}
 
+The distributed metadata service in Ceph provides clients with a POSIX file
+system abstraction. Distributed file systems are complex, but fundamentally
+provide durable hierarhical indexing and distributed locking service.  The
+locking service implements a capabilities that express what data and metadata
+clients are allowed to access as well as what they may cache locally.  While
+designed for a fixed file abstraction, indexing, locking, and caching are all
+common services that are useful to large sets of applications.
+
+\section{Malacology Implementation}
 \label{implementation}

### 2016-05-09 paper/paper.tex
fix whitespace
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>
add fig
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>
describe prop delay graph
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 \usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-    \usepackage{xltxtra,xunicode}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
 % use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
 
 \usepackage{graphicx}
-\usepackage[small]{caption}
+\usepackage[font=footnotesize,labelfont=bf]{caption}
 \usepackage{pstricks}
@@
 \setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
+%\setlength{\parskip}{6pt plus 2pt minus 1pt}
 \setlength{\emergencystretch}{3em}  % prevent overfull lines
@@
 are:
-
-\paragraph{}
 
@@
 
+In Section~\ref{} we described an interface management service that is
+responsible for the durable and consistent cluster-wide distribution of
+interface versions. We evaluate the costs of such a service by embedding
+interface definitions in a Ceph cluster membership data structure that is
+managed by a Paxos instance, and distributed using a scalable, peer-to-peer
+gossip protocol. We assume an architecture that consists of a light-weight
+interface that enforces the semantics of interface management resulting in one
+or more interface updates that must be consistent across the cluster. The
+following experiments focus on the cost of achieving distribution of consistent
+a interface version.
+
+\begin{figure}[h]
+\centering
+\includegraphics{figures/iface-update-delay.pdf}
+\caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
+interface commit. Each of the 120 OSDs used a RAM-based object store, while the
+3 OSDs used an HDD-backed object store.}
+\label{fig:propdelay}
+\end{figure}
+
+Figure~\ref{fig:propdelay} shows the CDF of the latency of distributing and
+installing an interface update on all OSDs in the cluster. The latency is
+defined as the elapsed time following the Paxos proposal for an interface
+update until each OSDs makes the update live (he cost of the Paxos proposal is
+configurable and is discussed below). The latency measurements were taken on
+the nodes running Ceph server daemons (the Paxos servers and OSDs nodes), and
+thus excludes the client round-trip cost. In each of the experiments 1000
+interface updates were observed.
+
+The first experiment shown in Figure~\ref{fig:propdelay} illustrates a lower
+bound cost for updates in a large cluster by avoiding disk I/O for storing
+interface updates locally within each OSD.  In the experiment labeled "120 OSD
+(RAM)" a cluster of 120 OSDs (10 OSDs x 12 servers) using an in-memory data
+store were deployed, showing a latency of less than 54 ms with a probability of
+90\% and a worst case latency of 194 ms. The curve labeled "3 OSD (HDD)" is a
+smaller cluster where each of the three OSDs uses a spinning disk as the object
+store that saves a copy of the cluster map ({\bf is this I/O before its made
+live?}), and shows a latency of 37 ms with a 90\% probability and a worst case
+latency of 255 ms. Since the cost of a disk I/O is larger than the same I/O
+executed against the in-memory device, the predominant cost in a lightly loaded
+cluster appears to be the number of OSDs.
+
+The real cost of interface management, in addition to cluster-wide propagation
+of interface updates, includes the network round-trip to the interface
+management service, the Paxos commit protocol, and other factors such as system
+load. By default Paxos proposals occur periodically with a 1 second interval in
+order to accumulate updates. In a minimum realistic quorum of 3 monitors using
+HDD-based storage, we were able to decrease this interval to an average of 222
+ms.
+
 \begin{figure}[htbp]

### 2016-05-09 paper/paper.tex
zlog notes
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-\subsection{ZLog: A Distributed Shared Commit
-Log}\label{zlog-a-distributed-shared-commit-log}
+\subsection{ZLog: A Distributed Shared-Log}
+
+The CORFU protocol is enforced using logical storage devices.
+
+The lua implementation of the interfaces are encoded into cluster maps
+for automatic verisoning and dissemtation across the cluster.
 
@@
 
-\subsubsection{Consistency}\label{consistency}
+log management is supported by mds using a special directory layout
+and file types.
 
-\subsubsection{Striping Strategies}\label{striping-strategies}
+The zlog sequencer is volatile, shared resource that requires serialized
+access. in corfu the sequencer is a fast, centeralized network counter
+resource.
 
-\subsubsection{Tiering}\label{tiering}
+Initially we saught to directly model the sequencer in ceph as a non-exclusive,
+non-cacheable resource, forcing clients to perform round-trip access the
+resource at the authoratiave MDS for the inode representing the sequencer.
+However, we found that the capapbility system in Ceph is used to batch access
+to cacheable resources, providing clients best-effort access to the capability
+and allowing them to burst. While unexpected, this alternative mode provides
+a different approach to sequencer implementation.
 
-\subsection{MDS as a sequencer}\label{mds-as-a-sequencer}
+In Section we quantify the trade-offs of throughput and latency for the
+centeralized sequencer and the batching mechanism based on capabilities.
 

### 2016-05-09 paper/paper.tex
paper: add implementation table
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-05-09 paper/paper.tex
@@
 
+\begin{figure}[tb]
+\centering
+\includegraphics{figures/overview.png}
+\caption{A Ceph installation consists of a cluster running mostly
+storage daemons (OSDs) to ensure data durability, a few monitor daemons
+(MONs) to reach consensus of the cluster state and to ensure its
+consistent versioning, and a few metadata service daemons (MDSs) for
+scalable metadata service. Ceph's APIs include object, block, file
+abstractions. Malacology enables the programmability of these
+abstractions and services. To illustrate Malacology we implement two
+services on Ceph, Zlog and Mantle, by programming Ceph's durability,
+consistent versioning, consensus, and metadata subsystems.
+\label{fig:overview}}
+\end{figure}
+
 These three trends put evolutionary pressure on storage systems and
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/overview.png}
-\caption{A Ceph installation consists of a cluster running mostly
-storage daemons (OSDs) to ensure data durability, a few monitor daemons
-(MONs) to reach consensus of the cluster state and to ensure its
-consistent versioning, and a few metadata service daemons (MDSs) for
-scalable metadata service. Ceph's APIs include object, block, file
-abstractions. Malacology enables the programmability of these
-abstractions and services. To illustrate Malacology we implement two
-services on Ceph, Zlog and Mantle, by programming Ceph's durability,
-consistent versioning, consensus, and metadata subsystems.
-\label{fig:overview}}
-\end{figure}
-
 For example, a Ceph cluster contains a small number of
@@
 
-\section{Services Built on
-Malacology}\label{services-built-on-malacology}
-
+\section{Services Built on Malacology}\label{services-built-on-malacology}
 \label{services}
@@
 \begin{table}
-\begin{tabular}{ | l | l | l | l | l |}
-\hline
-Cap Owner & Failure & ZLog Recovery \\ \hline
-mds & mds & mds \\ \hline
-client & mds & xfer client cap \\ \hline
-mds & client & n/a \\ \hline
-client & client & mds \\
-\hline
+\centering
+\begin{tabular}{  l | l | l    }
+\textbf{Service} & \textbf{Ceph Component} & \textbf{Used for...}  \\ \hline
+Mantle  & RADOS    & policy durability \\
+        & MON maps & policy versioning \\ \hline
+ZLog    & Logical Storage & CORFU protocol  \\ 
+        & MON maps & interface install \\ \hline
+Seq.    & MDS file types & log management \\ 
+        & MDS fault toler. & seq. fault toler. \\
 \end{tabular}
-\caption{Failure and recovery scenarios.}
-\label{t:mds-zlog-fail}
+\caption{To implement the Mantle and ZLog services, we leverage Ceph components. Malacology uses these subsystems because they provide powerful abstractions.}
+\label{table:implementation}
 \end{table}

### 2016-05-09 paper/paper.tex

--- 2016-05-09 paper/paper.tex
@@
 % carefully exposing internal storage services to applications.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/overview.png}
-\caption{A Ceph installation consists of a cluster running mostly
-storage daemons (OSDs) to ensure data durability, a few monitor daemons
-(MONs) to reach consensus of the cluster state and to ensure its
-consistent versioning, and a few metadata service daemons (MDSs) for
-scalable metadata service. Ceph's APIs include object, block, file
-abstractions. Malacology enables the programmability of these
-abstractions and services. To illustrate Malacology we implement two
-services on Ceph, Zlog and Mantle, by programming Ceph's durability,
-consistent versioning, consensus, and metadata subsystems.
-\label{fig:overview}}
-\end{figure}
 

### 2016-05-09 paper/paper.tex
zlog notes
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-05-09 paper/paper.tex
@@
 
-\subsection{ZLog: A Distributed Shared-Log}
+\subsection{ZLog: A Fast Distributed Shared-Log}
 
@@
 and file types.
+
+\subsubsection{Sequencer}
 
@@
 
-\subsubsection{Sequencer recovery with
-MDS}\label{sequencer-recovery-with-mds}
-
-\subsubsection{MDS Policy for Revoking
-Capability}\label{mds-policy-for-revoking-capability}
-
 The MDS has a queue of locks that it uses to revoke capabilities from
@@
 
-\subsubsection{Mantle Balancing Policies to Migrate
-Sequencers}\label{mantle-balancing-policies-to-migrate-sequencers}
+Mantle is used to implement a zlog file type specific policy for maximizing
+throughput to the sequencer by migration the resource between metadata servers.
+
+Sequencer recovery is handled by tapping into existing recovery mechanisms
+that recover inode state. file type specific recovery augments the existing
+recovery procedure.
 

### 2016-05-09 paper/paper.tex
add zlog seq eval
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-\subsection{MDS as ZLog Sequencer}\label{mds-as-zlog-sequencer}
+\subsection{ZLog Sequencer}
 
-\subsubsection{Experiment: Object Class
-Programmability}\label{experiment-object-class-programmability}
+We evaluated the performance of the ZLog Sequencer implemented in several
+different configurations by showing a trade-off between aggregate throughput
+achieved (shown in Figure~\ref{fig:captp}), and latency distribution (shown in
+Figure~\ref{fig:capcdf}).
 
-Add quote from CORFU paper that talks about what they had to hack
+In Figure~\ref{fig:captp} the highest throughput labeled $1C$ shows that a
+single client caching the sequencer resource may obtain new sequences at a rate
+of nearly 6 million positions per second. This is not surprising given that all
+state is cached locally in the client. When this resource is shared between two
+clients, the MDS uses a best-effort approach to granting the sequencer resource
+to each client (value $2C,BE$). This sharing has a large impact on performance,
+as shown by the nearly three orders of magnitude drop off in throughput.
+
+While a best-effort approach to sharing capabilities makes sense for a file
+system client, throughput drops significantly with two clients as the majority
+of the time neither client has the capapbitility as it ping-pongs between
+clients. This observation reveals a tuning parameter: how long a client
+maintains its capapbility lease. In throughputs labeled $2C,<time>$ we show
+aggregate throughput as we decrease the amount of time that a client holds its
+lease before responding to a request for the capapbility.  At lease times as
+low as 10ms throughputs achieve a significant proportion of the locally cached
+throughput.
+
+While the batching mechanism can achieve high-throughput, it is not appropriate
+for applications that require predictable latency. The throughput in
+Figure~\ref{fig:captp} labeled $2C,ZLog$ corresponds to a centeralized
+sequencer implementation. While throughput is lower for two clients (because of
+the synchronous network round-trip required for each request),
+Figure~\ref{fig:capcdf} shows that this approach achieves more predicable
+latency compared to all other approaches except for the single client locally
+cached mode.
 
@@
 \includegraphics{figures/caps-delay-thruput.png}
-\caption{Forcing the client to drop their capabilities later (delay)
-improves throughput}
+\caption{Sequencer throughput by reusing various services.
+The highest performance is achieved using a single client with
+exclusive, cacheable priviledge. Round-robin sharing of the sequencer
+resource is affected by the amount of time the resource is held,
+with best-effort performing the worst.}
+\label{fig:captp}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/caps-delay-latency.png}
+\caption{But if the clients hold their capabilities longer, it hurts
+latency for everyone.}
+\label{fig:capcdf}
 \end{figure}
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/caps-delay-latency.png}
-\caption{But if the clients hold their capabilities longer, it hurts
-latency for everyone.}
-\end{figure}
+\subsubsection{Experiment: Object Class
+Programmability}\label{experiment-object-class-programmability}
+
+Add quote from CORFU paper that talks about what they had to hack
 

### 2016-05-09 paper/paper.tex
updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-\section{Services Built on
-Malacology}\label{services-built-on-malacology}
-
+\section{Services Built on Malacology}
 \label{services}
 
-\begin{table}
-\begin{tabular}{ | l | l | l | l | l |}
-\hline
-Cap Owner & Failure & ZLog Recovery \\ \hline
-mds & mds & mds \\ \hline
-client & mds & xfer client cap \\ \hline
-mds & client & n/a \\ \hline
-client & client & mds \\
-\hline
-\end{tabular}
-\caption{Failure and recovery scenarios.}
-\label{t:mds-zlog-fail}
-\end{table}
+In this section we describe two services built on top of Malacology. The first
+is Mantle, a framework for dynamically specifying metadata load balancing
+policies. The second system, ZLog, is a high-performance distributed shared-log.
+In addition to these services, we'll demonstrate how we combine ZLog and Mantle
+to implement service-aware metadata load balancing policies.
 
-\subsection{Mantle: A Programmable Metadata Load
-Balancer}\label{mantle-a-programmable-metadata-load-balancer}
+\subsection{Mantle: A Programmable Metadata Load Balancer}
 
@@
 
-\subsection{ZLog Sequencer}
+\subsection{ZLog}
+
+\subsubsection{Sequencer}
 
@@
 as shown by the nearly three orders of magnitude drop off in throughput.
+
+\begin{figure}[h]
+\centering
+\includegraphics{figures/caps-delay-thruput.png}
+\caption{Sequencer throughput by reusing various services.
+The highest performance is achieved using a single client with
+exclusive, cacheable priviledge. Round-robin sharing of the sequencer
+resource is affected by the amount of time the resource is held,
+with best-effort performing the worst.}
+\label{fig:captp}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics{figures/caps-delay-latency.png}
+\caption{But if the clients hold their capabilities longer, it hurts
+latency for everyone.}
+\label{fig:capcdf}
+\end{figure}
 
@@
 
-\subsubsection{Experiment: Multi-Client
-Burstiness}\label{experiment-multi-client-burstiness}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/caps-delay-thruput.png}
-\caption{Sequencer throughput by reusing various services.
-The highest performance is achieved using a single client with
-exclusive, cacheable priviledge. Round-robin sharing of the sequencer
-resource is affected by the amount of time the resource is held,
-with best-effort performing the worst.}
-\label{fig:captp}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/caps-delay-latency.png}
-\caption{But if the clients hold their capabilities longer, it hurts
-latency for everyone.}
-\label{fig:capcdf}
-\end{figure}
+\subsubsection{Interface Propogation}
 
@@
 
-\subsubsection{CUT THIS Experiment: Client v. MDS caps +
-Failures}\label{cut-this-experiment-client-v.-mds-caps-failures}
-
-\subsubsection{Experiment: Mantle Sequencer
-Balancer}\label{experiment-mantle-sequencer-balancer}
-
 \section{Related Work}

### 2016-05-09 paper/paper.tex
paper: fix fig 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-05-09 paper/paper.tex
@@
 
+\begin{figure}[tb]
+\centering
+\includegraphics{figures/overview.png}
+\caption{A Ceph installation consists of a cluster running mostly
+storage daemons (OSDs) to ensure data durability, a few monitor daemons
+(MONs) to reach consensus of the cluster state and to ensure its
+consistent versioning, and a few metadata service daemons (MDSs) for
+scalable metadata service. Ceph's APIs include object, block, file
+abstractions. Malacology enables the programmability of these
+abstractions and services (indicated by bold arrows). To illustrate Malacology we implement two
+services on Ceph, Zlog and Mantle, by programming Ceph's durability,
+consistent versioning, consensus, and metadata subsystems.
+\label{fig:overview}}
+\end{figure}
+
 These three trends put evolutionary pressure on storage systems and
@@
 
-\subsection{ZLog: A Fast Distributed Shared-Log}
+\subsection{ZLog: A Distributed Shared-Log}
 
@@
 and file types.
-
-\subsubsection{Sequencer}
 
@@
 
+\subsubsection{Sequencer recovery with
+MDS}\label{sequencer-recovery-with-mds}
+
+\subsubsection{MDS Policy for Revoking
+Capability}\label{mds-policy-for-revoking-capability}
+
 The MDS has a queue of locks that it uses to revoke capabilities from
@@
 
-Mantle is used to implement a zlog file type specific policy for maximizing
-throughput to the sequencer by migration the resource between metadata servers.
-
-Sequencer recovery is handled by tapping into existing recovery mechanisms
-that recover inode state. file type specific recovery augments the existing
-recovery procedure.
+\subsubsection{Mantle Balancing Policies to Migrate
+Sequencers}\label{mantle-balancing-policies-to-migrate-sequencers}
 

### 2016-05-09 paper/paper.tex

--- 2016-05-09 paper/paper.tex
@@
 
+\begin{table}
+\centering
+\begin{tabular}{  l | l | l    }
+\textbf{Service} & \textbf{Ceph Component} & \textbf{Used for...}  \\ \hline
+Mantle  & RADOS    & policy durability \\
+        & MON maps & policy versioning \\ \hline
+ZLog    & Logical Storage & CORFU protocol  \\ 
+        & MON maps & interface install \\ \hline
+Seq.    & MDS file types & log management \\ 
+        & MDS fault toler. & seq. fault toler. \\
+\end{tabular}
+\caption{To implement the Mantle and ZLog services, we leverage Ceph components. Malacology uses these subsystems because they provide powerful abstractions.}
+\label{table:implementation}
+\end{table}
+
 In this section we describe two services built on top of Malacology. The first

### 2016-05-09 paper/paper.tex
modified:   paper.tex 	Polished section 2

--- 2016-05-09 paper/paper.tex
@@
 
-\begin{figure}[tb]
-\centering
-\includegraphics{figures/overview.png}
-\caption{A Ceph installation consists of a cluster running mostly
-storage daemons (OSDs) to ensure data durability, a few monitor daemons
-(MONs) to reach consensus of the cluster state and to ensure its
-consistent versioning, and a few metadata service daemons (MDSs) for
-scalable metadata service. Ceph's APIs include object, block, file
-abstractions. Malacology enables the programmability of these
-abstractions and services (indicated by bold arrows). To illustrate Malacology we implement two
-services on Ceph, Zlog and Mantle, by programming Ceph's durability,
-consistent versioning, consensus, and metadata subsystems.
-\label{fig:overview}}
-\end{figure}
-
 These three trends put evolutionary pressure on storage systems and
@@
 % carefully exposing internal storage services to applications.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/overview.png}
+\caption{A Ceph installation consists of a cluster running mostly
+storage daemons (OSDs) to ensure data durability, a few monitor daemons
+(MONs) to reach consensus of the cluster state and to ensure its
+consistent versioning, and a few metadata service daemons (MDSs) for
+scalable metadata service. Ceph's APIs include object, block, file
+abstractions. Malacology enables the programmability of these
+abstractions and services. To illustrate Malacology we implement two
+services on Ceph, Zlog and Mantle, by programming Ceph's durability,
+consistent versioning, consensus, and metadata subsystems.
+\label{fig:overview}}
+\end{figure}
 
@@
 a high-performance distributed shared-log on Ceph called ZLog, based on the
-CORFU protocol~\cite{corfu}. Serialization at high-performance is achieved using a
-``sequencer'' service that maintains a volatile, in-memory counter used to assign a log
-position to each client that appends to the log. One option for building
+CORFU protocol~\cite{corfu}.
+Without going into greater detail, the CORFU protocol by separating the concern of correctness from the concern of performance enables serialization using a \emph{volatile}
+``sequencer'' service. This high-performance in-memory counter is used to assign a log
+position to each client that appends to the log.
+The sequencer is purely a performance optimization and does not have a role in correctness (if the sequencer fails, the tail of the log can always be retrieved by other means).
+Thus, it is the volatiity of the sequence that enables high performance.
+A ``bolt-on'' option for building
 a distributed log is to use PAXOS, or an out-of-the-box solution such as
@@
 scientific data (3D array-based data, more specifically) to increase
-performance, locality, and the number of early results. This is not a
-bad proposition, but creates a coupling that is highly tied to the
+performance, locality, and the number of early results. %This is not a
+%bad proposition, but 
+Changing applications creates a coupling that is highly tied to the
 underlying physical properties of the system, making it difficult to
@@
 
-When these two approaches fail to meet the needs of the application,
+When both the bolt-on approach and changing applications fail to meet the needs of the application,
 developers turn their attention to the storage system itself.
@@
 tunables are difficult to understand or quantify (e.g., Ceph tunables~\cite{sevilla:sc15-mantle}). To succesfully tune the system, a
-developer must have domain and system specific knowledge. Without this
+developer must have domain and system-specific knowledge. Without this
 intimate knowledge, the tuning turns into somewhat of a black art, where
@@
 existing interfaces. These interfaces are written in C++ and are statically
-loaded into the system. An example usage may be to atomically combine a
-``write'' with an index update to create a content-addressable object
-interface.
+loaded into the system. %An example usage may be to atomically combine a
+%``write'' with an index update to create a content-addressable object
+%interface.
 
@@
 
-Figure 1 shows the throughput
-(\emph{y} axis) over time (\emph{x} axis) of two OSD implementations for
-storing metadata: in the header of the byte stream (data) vs.~in the
+For example, Figure~\ref{fig:datavsxattr} compares the throughput
+(\emph{y} axis) over time (\emph{x} axis) of two OSD implementations of the same storage object interface for
+storing metadata: either in the header of the byte stream (data) or in the
 object extended attributes (XATTR). The speed for appending data without
@@
 metadata (data raw).}
+\label{fig:datavsxattr}
 \end{figure}
@@
 The object interface facility provided by Ceph is more than an
-a proof-of-concept active storage implementation.  Figure
-\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
+a proof-of-concept active storage implementation.  Figure~\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
 domain-specific interfaces in the Ceph community since 2010.  What is most
@@
 number of object classes (a group of interfaces), and the total number
-of methods (the actual API end-points).\label{fig:obj-int-dev-growth}}
+of methods (the actual API end-points).}
+\label{fig:obj-int-dev-growth}
 \end{figure}
@@
 resources to improve performance. Programmers gravitate towards object
-interfaces because it gives them ability to tell the storage system
+interfaces because it gives them the ability to tell the storage system
 about their application: if it is CPU or IO bound, if it has locality,
@@
 programmers know what the problem is and how to solve it, but until
-object interfaces, had no way to tell the storage system how to handle
+the ability to modify object interfaces, they had no way to express to the storage system how to handle
 their data.
 
+\if{false}
 These implementations use the active storage interfaces in Ceph. We
@@
 classes.
+\fi
 
@@
 
-\section{Services Built on Malacology}\label{services-built-on-malacology}
+\section{Services Built on
+Malacology}\label{services-built-on-malacology}
+
 \label{services}
@@
 \begin{table}
-\centering
-\begin{tabular}{  l | l | l    }
-\textbf{Service} & \textbf{Ceph Component} & \textbf{Used for...}  \\ \hline
-Mantle  & RADOS    & policy durability \\
-        & MON maps & policy versioning \\ \hline
-ZLog    & Logical Storage & CORFU protocol  \\ 
-        & MON maps & interface install \\ \hline
-Seq.    & MDS file types & log management \\ 
-        & MDS fault toler. & seq. fault toler. \\
+\begin{tabular}{ | l | l | l | l | l |}
+\hline
+Cap Owner & Failure & ZLog Recovery \\ \hline
+mds & mds & mds \\ \hline
+client & mds & xfer client cap \\ \hline
+mds & client & n/a \\ \hline
+client & client & mds \\
+\hline
 \end{tabular}
-\caption{To implement the Mantle and ZLog services, we leverage Ceph components. Malacology uses these subsystems because they provide powerful abstractions.}
-\label{table:implementation}
+\caption{Failure and recovery scenarios.}
+\label{t:mds-zlog-fail}
 \end{table}

### 2016-05-09 paper/paper.tex

@@
 \begin{table}
-\begin{tabular}{ | l | l | l | l | l |}
-\hline
-Cap Owner & Failure & ZLog Recovery \\ \hline
-mds & mds & mds \\ \hline
-client & mds & xfer client cap \\ \hline
-mds & client & n/a \\ \hline
-client & client & mds \\
-\hline
+\centering
+\begin{tabular}{  l | l | l    }
+\textbf{Service} & \textbf{Ceph Component} & \textbf{Used for...}  \\ \hline
+Mantle  & RADOS    & policy durability \\
+        & MON maps & policy versioning \\ \hline
+ZLog    & Logical Storage & CORFU protocol  \\ 
+        & MON maps & interface install \\ \hline
+Seq.    & MDS file types & log management \\ 
+        & MDS fault toler. & seq. fault toler. \\
 \end{tabular}
-\caption{Failure and recovery scenarios.}
-\label{t:mds-zlog-fail}
+\caption{To implement the Mantle and ZLog services, we leverage Ceph components. Malacology uses these subsystems because they provide powerful abstractions.}
+\label{table:implementation}
 \end{table}
+
+In this section we describe two services built on top of Malacology. The first
+is Mantle, a framework for dynamically specifying metadata load balancing
+policies. The second system, ZLog, is a high-performance distributed shared-log.
+In addition to these services, we'll demonstrate how we combine ZLog and Mantle
+to implement service-aware metadata load balancing policies.
 

### 2016-05-09 paper/paper.tex
modified:   paper.tex 	Better formatting of contributions in intro

@@
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item A programmable storage system implementation that re-uses and extends existing abstractions. It includes:
+\paragraph*{A programmable storage system implementation} that re-uses and extends existing abstractions. It includes:
 
   \begin{enumerate}
-  \def\labelenumi{\arabic{enumi}.}
-  \itemsep1pt\parskip0pt\parsep0pt
+%  \def\labelenumi{\arabic{enumi}.}
+%  \itemsep1pt\parskip0pt\parsep0pt
   \item
-    Extending file system metadata by a new type of ``sequencer'' inode that together with Ceph's capability subsystem implements high-performance, non-durable access serialization
+    An extension of file system metadata by a new type of ``sequencer'' inode that together with Ceph's capability subsystem implements high-performance, non-durable access serialization
   \item
-    Re-using Ceph's PAXOS-based consensus infrastructure for versioning of metadata service load-balancing policies
+    The use of Ceph's PAXOS-based consensus infrastructure for versioning of metadata service load-balancing policies
   \item
-    Re-using storage objects for persisting policies
+    The use of storage objects for persisting policies
   \item
-    Extending object interface classes for new logical storage/metadata devices
+    An extension of the object interface classes for new logical storage/metadata devices
   \end{enumerate}
-\item
-  Two example implementations and evaluations for feasibility of services that use this framework
+
+\paragraph*{Two example implementations and evaluations} for feasibility of services that use this framework
 
   \begin{enumerate}
-  \def\labelenumi{\arabic{enumi}.}
-  \itemsep1pt\parskip0pt\parsep0pt
+%  \def\labelenumi{\arabic{enumi}.}
+%  \itemsep1pt\parskip0pt\parsep0pt
   \item
@@
   \item
-    A re-implementation of the programmable Mantle metadata load balancing service using Malacology~\cite{sevilla:sc15-mantle}
+    A re-implementation of the programmable Mantle metadata load balancing service~\cite{sevilla:sc15-mantle} using Malacology
   \item 
@@
   \end{enumerate}
-\end{itemize}
 

### 2016-05-09 paper/paper.tex
Update paper.tex

@@
 The sequencer is purely a performance optimization and does not have a role in correctness (if the sequencer fails, the tail of the log can always be retrieved by other means).
-Thus, it is the volatiity of the sequence that enables high performance.
+Thus, it is the volatility of the sequencer that enables high performance.
 A ``bolt-on'' option for building
 a distributed log is to use PAXOS, or an out-of-the-box solution such as
-Zookeeper. However, these systems adhere to the pre-coneived notions of
+Zookeeper. However, these systems adhere to the preconceived notions of
 storage, namely that data is safe from data loss, which is exactly the feature

### 2016-05-09 paper/paper.tex
Update paper.tex

@@
 is left off the table.  Starfish~\cite{herodotou:cidr2011-starfish} made
-an attempt of this for Hadoop but this technique would need to severely limit
-the space of parameters in order to feasible, similar to the approach used in~\cite{behzad:sc2013-autotuning}.
+an attempt at this for Hadoop but this technique would need to severely limit
+the space of parameters in order to be feasible, similar to the approach used in~\cite{behzad:sc2013-autotuning}.
 

### 2016-05-09 paper/paper.tex
Update paper.tex

--- 2016-05-09 paper/paper.tex
@@
 As an example use-case, ZLog needs to
-store metadata information with each object so for each operation, it
+store metadata information with each object, so for each operation, it
 checks the epoch value and write capability (2 metadata reads) and
@@
 any metadata operations (data raw) is also shown as a baseline for
-comparison. For append-heavy workloads storing metadata as a header in
+comparison. For append-heavy workloads, storing metadata as a header in
 the data performs about 1.5x better than storing metadata as an extended
@@
 
-The object interface facility provided by Ceph is more than an
+The object interface facility provided by Ceph is more than
 a proof-of-concept active storage implementation.  Figure~\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
@@
 A consistent view of cluster state among Ceph daemons and clients is critical
-to provide strong consistency guarnatees to clients.  Ceph maintains cluster
+to provide strong consistency guarantees to clients.  Ceph maintains cluster
 state information in per-subsystem ``maps'' (e.g. OSD map, MDS map) that
 record membership and status information. A PAXOS-based monitoring service is
-responsible integrating state changes into cluster maps, and responding to
+responsible for integrating state changes into cluster maps, and responding to
 requests from out-of-date clients.  As a fundamental building block of many
@@
 techniques try to be autonomous so that work is distributed across the
-cluster. For example, when placement groups change the OSDs rebalance and
+cluster. For example, when placement groups change, the OSDs rebalance and
 re-shard data in the background in a process called placement group splitting.

### 2016-05-09 paper/paper.tex
sec 4 updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-05-09 paper/paper.tex
@@
 
-To enable the re-use of Ceph subsystems, Malacology relies heavily on
-interface classes for introducing new functionality into the MDSs and
-OSDs. Malacology introduces 3 components to manage object/metadata
-interfaces:
-
+In this section we describe the extensions to existing Ceph services that
+allow for reusability.
+
+{Lua Integration.} Because of limited space we describe our extensions
+generally, but we have explicitly added the ability to use Lua to specify
+the customizations. Lua is great because..., and let's us save and version
+interfaces.
+
+Our framework has added a mechanism for defining and running object and
+metadata balancer classes using the Lua scripting language. Scripting is
+useful in large systems for 3 reasons:
+
+\begin{itemize}
+\item
+  accelerates deployment: tweaking policies, like changing a threshold
+  or metric calculation, does not require a full recompile or system
+  start-up.
+\item
+  transparency: if the bindings carefully expose the correct metrics and
+  functions, the user does not need to learn the internal functions,
+  variables, or types of the system.
+\item
+  portability: scripts can be exchanged across platforms and system
+  versions without compatibility issues.
+\end{itemize}
+
+Lua performs well as an embeddable scripting language. Although it is an
+interpreted language, the bindings are designed to make the exchange of
+variables and functions amongst languages straightforward. The LuaJIT
+virtual machine boasts near-native performance, making it a logical
+choice for scriptability frameworks in systems research
+\cite{neto:dls14-luaos}. In storage systems, it has been effectively
+used both on \cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo} and off
+\cite{sevilla:sc15-mantle} the critical path, where performance is
+important.
+
+The other two advantages, that Lua leaves a small memory footprint and
+that it is secure, are discussed at length in 
+\cite{ierusalimschy_programming_2006,neto:dls14-luaos}. In short, Lua
+provides relatively robust sandbox capabilities for protecting the
+
+\subsection{Metadata Services}
+
+{\bf Balancing Policy.}
+Avoiding hard-coded balancing policies are important. First, it is difficult,
+nay impossible, to select a one-size-fits-all policy. System administrators
+may want to partition load based on SLAs, etc... Applications are often in
+a better position to know what to do.
+
+Mantle is a framework that allows balancing policies to be written in Lua,
+and supports such policies be specified at run-time.
+
+{\bf Capabilities.}
+The metadata service tracks the state of clients, including the data they
+may have cached and any locks that they hold. The metadata service is
+responsible responding to client requests for capabilities by first
+revoking caps from other clients. Currently clients voluntarily
+release their capabilities, and the MDS maintains a queue of requests.
+Both clients and MDSs use a best-effort approach to handling capabilities.
+
+In our current we retain the best-effort, volutary release of caps within
+clients, but generalize the queueing policies within the MDS to allow for
+centeralized policies such as fairness or priority.
+
+{\bf File Types.}
+We can reference PDSW12 DataMods, and talk about any new stuff here that we
+want to.
+
+\subsection{Object Storage}
+
+Lua integration. Overhead?
+
+Although these OSD interfaces are powerful, the current implementation
+has drawbacks. OSD interfaces are written in C/C++ and compiled into a
+shared library, so the developer must account for different target
+architectures. Second, C/C++ provides more functionality than is
+necessary since the snippets of interface code mainly set policies and
+perform simple operations. Finally, developing these interfaces in C/C++
+has high overhead. The developer must learn how the OSD dynamically
+
+\subsection{Cluster State}
+
+Map modification for propogation of state. We allow applications to augment
+these data structures with application-specific metdata. In order to keep the
+size of these data structures small application state may be stored elsewhere
+and the references stored within the data structure.
+
+\subsection{Programmability and Durability}
+
+{\bf Recovery.}
+Application specific recovery is critical.
+
+{\bf Durability.}
+
+\noindent \texttt{ceph\ mds\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}
+\noindent \texttt{ceph\ osd\ pool\ set-class\ \textless{}pool\textgreater{}\ \textless{}class\textgreater{}\ \textless{}script\textgreater{}}
+
+Metadata service policies and object storage interfaces are stored durability
+within RADOS and managed by storing references with the object maps. Since
+the cluster already propogates a consistent view of these data structures,
+we use this service to automatically install interfaces in OSDs, and install
+policies within the MDS such that clients and daemons are synchronzied on
+correct implementations without restarting.
+
+\iffalse
 \begin{enumerate}
@@
 had a large function hard-coded into the balancer subsystem, Malacology
-embeds code in the MDS using OSD subsystems. Ceph has a whole
+embeds code in the MDS using OSD subsystems.
+
+Ceph has a whole
 infrastructure for getting dynamic code into the OSD. The
@@
 library; both shared libraries will the Lua core.
+\fi
 
@@
 
+\iffalse
 Both the object and balancer interfaces define functions and attach them
@@
 
-\subsection{Storing Interfaces in
-RADOS}\label{storing-interfaces-in-rados}
+\subsection{Programmability and Durability}
 
@@
 classes to the OSDs equipped with the LuaJIT VM.
-
-\section{Services Built on
-Malacology}\label{services-built-on-malacology}
+\fi
+
+\section{Services Built on Malacology}\label{services-built-on-malacology}
 

### 2016-05-09 paper/paper.tex

--- 2016-05-09 paper/paper.tex
@@
 is left off the table.  Starfish~\cite{herodotou:cidr2011-starfish} made
-an attempt at this for Hadoop but this technique would need to severely limit
-the space of parameters in order to be feasible, similar to the approach used in~\cite{behzad:sc2013-autotuning}.
+an attempt of this for Hadoop but this technique would need to severely limit
+the space of parameters in order to feasible, similar to the approach used in~\cite{behzad:sc2013-autotuning}.
 
@@
 provide durable hierarhical indexing and distributed locking service.  The
-locking service implements a capabilities that express what data and metadata
+locking service implements capabilities that express what data and metadata
 clients are allowed to access as well as what they may cache locally.  While
@@
 functions to be shipped around the cluster without having to compile or
-reason about compatibility issues. Traditionally, intefaces are written
+reason about compatibility issues. Traditionally, interfaces are written
 in C/C++, so they must be compiled into shared libraries before being
@@
 the MDS process. As a result, we put interface code that uses these in
-\texttt{cls} directory. With this scheme, The OSD will \texttt{dlopen()}
+\texttt{cls} directory. With this scheme, the OSD will \texttt{dlopen()}
 the shared library created by Lua object interface shared library while
@@
 details (\emph{e.g.}, snapshots) for each pool. While stuffing the
-information into the monitor map, the structure that describes the the
+information into the monitor map, the structure that describes the
 monitor topology, was an option, the placement group pool allows finer
@@
 The zlog sequencer is volatile, shared resource that requires serialized
-access. in corfu the sequencer is a fast, centeralized network counter
+access. In Corfu the sequencer is a fast, centeralized network counter
 resource.
 
-Initially we saught to directly model the sequencer in ceph as a non-exclusive,
+Initially we saught to directly model the sequencer in Ceph as a non-exclusive,
 non-cacheable resource, forcing clients to perform round-trip access the
 resource at the authoratiave MDS for the inode representing the sequencer.
-However, we found that the capapbility system in Ceph is used to batch access
+However, we found that the capability system in Ceph is used to batch access
 to cacheable resources, providing clients best-effort access to the capability

### 2016-05-09 paper/paper.tex
sec4
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-05-09 paper/paper.tex
@@
 
-In this section we describe the extensions to existing Ceph services that
-allow for reusability.
-
-{Lua Integration.} Because of limited space we describe our extensions
-generally, but we have explicitly added the ability to use Lua to specify
-the customizations. Lua is great because..., and let's us save and version
-interfaces.
-
-Our framework has added a mechanism for defining and running object and
-metadata balancer classes using the Lua scripting language. Scripting is
-useful in large systems for 3 reasons:
-
-\begin{itemize}
-\item
-  accelerates deployment: tweaking policies, like changing a threshold
-  or metric calculation, does not require a full recompile or system
-  start-up.
-\item
-  transparency: if the bindings carefully expose the correct metrics and
-  functions, the user does not need to learn the internal functions,
-  variables, or types of the system.
-\item
-  portability: scripts can be exchanged across platforms and system
-  versions without compatibility issues.
-\end{itemize}
-
-Lua performs well as an embeddable scripting language. Although it is an
-interpreted language, the bindings are designed to make the exchange of
-variables and functions amongst languages straightforward. The LuaJIT
-virtual machine boasts near-native performance, making it a logical
-choice for scriptability frameworks in systems research
-\cite{neto:dls14-luaos}. In storage systems, it has been effectively
-used both on \cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo} and off
-\cite{sevilla:sc15-mantle} the critical path, where performance is
-important.
-
-The other two advantages, that Lua leaves a small memory footprint and
-that it is secure, are discussed at length in 
-\cite{ierusalimschy_programming_2006,neto:dls14-luaos}. In short, Lua
-provides relatively robust sandbox capabilities for protecting the
+In this section we describe how Malacology exposes and builds upon existing
+services within Ceph. While we strive for implementation independent
+approaches within Malacology, our prototype implementation makes heavy use of
+the embedded language Lua to enable dynamic features without relying on
+complex and heavy-weight code generation techniques.
+
+{\bf Lua integration.}
+Supporting dynamic reconfiguration and programmabiilty is critical in real
+systems that must adapt to a changing environment where potentially
+conflicting policies arise. We've adopted Lua as a scripting language to
+implement dynamic components in Malacology.  Lua is a portable, embedded
+scripting language that offers superious performance and productivity
+trade-offs, including a JIT-based implementation that is well known for near
+native performance. Lua has been used extensively in game engines, and systems
+research \cite{neto:dls14-luaos}, including storage systems where it has been
+effectively used both on
+\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo} and off
+\cite{sevilla:sc15-mantle} the performance critical path.
 

### 2016-05-09 paper/paper.tex
indent?
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 \urlstyle{same}  % don't use monospace font for urls
-\setlength{\parindent}{0pt}
+%\setlength{\parindent}{0pt}
 %\setlength{\parskip}{6pt plus 2pt minus 1pt}
@@
   \item 
-    An application of the Mantle metadata load balancing service to manage the placement of new sequencer inodes.
+    An application of the Mantle metadata load balancing service to manage the placement of new sequencer inodes
   \end{enumerate}

### 2016-05-09 paper/paper.tex

@@
 \urlstyle{same}  % don't use monospace font for urls
-%\setlength{\parindent}{0pt}
+\setlength{\parindent}{0pt}
 %\setlength{\parskip}{6pt plus 2pt minus 1pt}
@@
   \item 
-    An application of the Mantle metadata load balancing service to manage the placement of new sequencer inodes
+    An application of the Mantle metadata load balancing service to manage the placement of new sequencer inodes.
   \end{enumerate}

### 2016-05-09 paper/paper.tex
stuff
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-05-09 paper/paper.tex
@@
 % using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\setkeys{Gin}{width=\maxwidth,keepaspectratio}
 \ifxetex
@@
 %\setlength{\parskip}{6pt plus 2pt minus 1pt}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
+%\setlength{\emergencystretch}{3em}  % prevent overfull lines
 \setcounter{secnumdepth}{5}
@@
 
+% todo:
+%  - ellude to the need for policies to be avialable quickly to support
+%    our experiment about propogtation delay
+%
+
 {\bf Balancing Policy.}
@@
 \subsection{Object Storage}
+
+% todo:
+%  - ellude to the need for interfaces to be avialable quickly to support
+%    our experiment about propogtation delay
+%
 
@@
 
-\subsection{Mantle: A Programmable Metadata Load
-Balancer}\label{mantle-a-programmable-metadata-load-balancer}
+\subsection{Mantle: A Programmable Metadata Load Balancer}
 
@@
 
+The second service implemented on Malacology is ZLog, and a high-performance
+distributed shared-log that is based on the CORFU protocol~\cite{corfu}. A
+shared-log is a general, yet powerful abstraction used to build many
+distributed systems. However, existing implementations that rely on consensus
+algorithms such as PAXOS funnel I/O through a single point introducing a
+bottleneck that restricts throughput. In contrast, the CORFU protocol is able
+to achieve high throughput using a \emph{volatile} in-memory network counter
+service (called a \emph{sequencer}) that distributes log positions to clients,
+and may run at very high rates. Failures in CORFU are handled by a recovery
+algorithm that recomputes the sequencer state using custom storage interfaces
+to discover the tail of the log.
+
+A full dexription ...
+
+that computes the
+tail of the log by contacting each storage device
+
+Two i
+
+The CORFU protocol handles failures using a combination of
+
+Failures in CORFU are handled using a recovery protocol that res
+
 The CORFU protocol is enforced using logical storage devices.

### 2016-05-09 paper/paper.tex
Adds references to Intro
Fixes #47

@@
 Storage systems have evolved around storage devices that often were
-orders of magnitude slower than CPU and memory and therefore can
+orders of magnitude slower than CPU and memory, and therefore could
 dominate overall performance if not used carefully. Over the last few
@@
 ingenious strategies to meet correctness requirements while somewhat
-hiding the latency of traditional storage media. To avoid lock-in by a
+hiding the latency of traditional storage media~\cite{brewer_disks_2016}. To avoid lock-in by a
 particular vendor, users of storage systems have preferred systems with
 highly standardized APIs and lowest common denominator abstract data
-types such as blocks of bytes and byte stream files.
+types such as blocks of bytes and byte stream files~\cite{armbrust_view_2010}.
 
@@
 pressure storage systems to shorten their code paths and incorporate new
-optimizations; (2) demand for managing structured data and flexible
+optimizations~\cite{gray_tape_2007,gray_flash_2008}; (2) demand for managing structured data and flexible
 consistency semantics at scale pressure big data processing systems to
-use storage abstractions that can meet these demands; and (3)
+use storage abstractions that can meet these demands~\cite{apache_contributors_parquet_2014}; and (3)
 production-quality scalable storage systems available as open source
 software have established and are continuing to establish new, \emph{de
-facto} API standards at a faster pace than traditional standards bodies.
+facto} API standards at a faster pace than traditional standards bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
 
@@
 Something of a storage swiss army knife, Ceph supports file, block, and
-object interfaces simultaneously in a single cluster. Ceph's Reliable Autonomous Distributed Object Storage (RADOS) system is a cluster of storage
+object interfaces simultaneously in a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous 
+Distributed Object Storage (RADOS) system is a cluster of storage
 devices (OSDs) that provide Ceph with data durability and integrity

### 2016-05-09 paper/paper.tex
sec5
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-05-09 paper/paper.tex
@@
 
-The second service implemented on Malacology is ZLog, and a high-performance
+The second service implemented on Malacology is ZLog, a high-performance
 distributed shared-log that is based on the CORFU protocol~\cite{corfu}. A
@@
 bottleneck that restricts throughput. In contrast, the CORFU protocol is able
-to achieve high throughput using a \emph{volatile} in-memory network counter
-service (called a \emph{sequencer}) that distributes log positions to clients,
-and may run at very high rates. Failures in CORFU are handled by a recovery
-algorithm that recomputes the sequencer state using custom storage interfaces
-to discover the tail of the log.
-
-A full dexription ...
-
-that computes the
-tail of the log by contacting each storage device
-
-Two i
-
-The CORFU protocol handles failures using a combination of
-
-Failures in CORFU are handled using a recovery protocol that res
-
-The CORFU protocol is enforced using logical storage devices.
-
-The lua implementation of the interfaces are encoded into cluster maps
-for automatic verisoning and dissemtation across the cluster.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/overview-zlog.png}
-\caption{This is ZLog.}
-\end{figure}
+to achieve high throughput network counter service (called a \emph{sequencer})
+that avoids all persistent I/O in the common case.
+
+While a full description of the CORFU system is beyond the scope of this
+paper, we'll briefly describe the storage interface, sequencer, and recovery
+protocol that form the core components of CORFU that we implement within
+Malacology:
+
+{\bf Sequencer.}
+High-performance in CORFU is achived using a sequencer service that assigns
+client log positions by reading from a volatile, in-memory counter which can
+run at a very high throughput and at low-latency. Since the sequencer is
+centralized, ensuring seiralizability in the common case is trivial.  The
+primary challenge in CORFU is handling the failure of the sequencer in a way
+that guarantees correctness.  Failures of the sequencer service in CORFU are
+handled by a recovery algorithm that recomputes the new sequencer state using
+an application-specific custom storage interface to discover the tail of the
+log and invalidate stale client requests using an epoch-based protocol.
+
+{\bf Storage Interface.} The storage interface is a critical component in the
+CORFU design. Storage devices must provide a \emph{write}-once, random
+\emph{read} interface for reading and writing log entries. The key to
+correctness in CORFU lies with the enforcement of up-to-date epoch tags on
+client requests; requests tagged with out-of-date epoch values are rejected,
+and clients are expected to request a new tail from the sequencer after
+refreshing state from an auxillary service. This mechanism forms the basis for
+sequencer recovery.
+
+{\bf Sequencer Recovery.}
+In order to repopulate the cached tail value during recovery of a sequencer,
+the maximum position in the log must be obtained. To do this, the storage
+interface exposes an additional \emph{seal} method that atomically installs a
+new epoch value and returns the maximum log position that has been written.
+Since the recovery protocol does not inform clients of a new epoch value until
+the end of the process, the validity of the tail can be guaranteed despite the
+iterative process of contacting storage devices.  Recovery of the sequencer
+process itself may be handled in many ways, such as leader election using an
+auxillary service such as PAXOS.
+
+\subsubsection{Log Management}
 
@@
 recovery procedure.
+
+%\begin{figure}[htbp]
+%\centering
+%\includegraphics{figures/overview-zlog.png}
+%\caption{This is ZLog.}
+%\end{figure}
+
+\subsection{Interface Installation}
+
+The lua implementation of the interfaces are encoded into cluster maps
+for automatic verisoning and dissemtation across the cluster.
 

### 2016-05-09 paper/paper.tex
paper: add load balancing problem to section 3
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-05-09 paper/paper.tex
@@
 Storage systems have evolved around storage devices that often were
-orders of magnitude slower than CPU and memory, and therefore could
+orders of magnitude slower than CPU and memory and therefore can
 dominate overall performance if not used carefully. Over the last few
@@
 ingenious strategies to meet correctness requirements while somewhat
-hiding the latency of traditional storage media~\cite{brewer_disks_2016}. To avoid lock-in by a
+hiding the latency of traditional storage media. To avoid lock-in by a
 particular vendor, users of storage systems have preferred systems with
 highly standardized APIs and lowest common denominator abstract data
-types such as blocks of bytes and byte stream files~\cite{armbrust_view_2010}.
+types such as blocks of bytes and byte stream files.
 
@@
 pressure storage systems to shorten their code paths and incorporate new
-optimizations~\cite{gray_tape_2007,gray_flash_2008}; (2) demand for managing structured data and flexible
+optimizations; (2) demand for managing structured data and flexible
 consistency semantics at scale pressure big data processing systems to
-use storage abstractions that can meet these demands~\cite{apache_contributors_parquet_2014}; and (3)
+use storage abstractions that can meet these demands; and (3)
 production-quality scalable storage systems available as open source
 software have established and are continuing to establish new, \emph{de
-facto} API standards at a faster pace than traditional standards bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
+facto} API standards at a faster pace than traditional standards bodies.
 
@@
 Something of a storage swiss army knife, Ceph supports file, block, and
-object interfaces simultaneously in a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous 
-Distributed Object Storage (RADOS) system is a cluster of storage
+object interfaces simultaneously in a single cluster. Ceph's Reliable Autonomous Distributed Object Storage (RADOS) system is a cluster of storage
 devices (OSDs) that provide Ceph with data durability and integrity
@@
 
+\subsection{Load Balancing}
+
+% Straw man example
+Migrating resources is an important part of load balancing but understanding the effects of different configurations and parameters is difficult.
+Systems already virtualize and migrate memory; soon systems will be able to migrate other resources, like CPU, disk, and memory.
+Figuring out how, when, and where to move these resources depends on complex metrics like utilization, configuration, and workload.
+
+These problems are general, distributed systems load balancing problems and share many of the same properties as the POSIX metadata wall~\cite{alam:pdsw2011-metadata-scaling,ghemawat:sosp2003-gfs,hildebrand:msst2005-pnfs,weil_ceph_2006,welch:fast2008-panasas,shvachko:login2012-hdfs-scalability}.
+Many distributed file systems decouple metadata and data I/O so that
+these services can scale independently. Despite
+this optimization, scaling the metadata services is still difficult
+because metadata accesses impose small and frequent requests on the
+underlying storage system~\cite{roselli:atec2000-FS-workloads}. Many
+techniques for designing the metadata services have been proposed to
+accommodate this workload: Lustre~\cite{konstantinos:pdsw2014-lustre-metadata}, GFS~\cite{ghemawat:sosp2003-gfs}, and HDFS~\cite{shvachko:login2012-hdfs-scalability} keep all metadata on one
+server; GIGA+~\cite{patil:fast2011-giga}, IndexFS~\cite{ren:sc2014-indexfs}, Lazy Hybrid~\cite{brandt:msst2003-lh}, GPFS~\cite{schmuck:fast2002-gpfs},
+and pNFS~\cite{hildebrand:supercomputing2006-pNFS} hash the file
+system namespace across a dedicated cluster and cache metadata; Panasas~\cite{welch:fast2008-panasas} and CephFS~\cite{weil:sc2004-dyn-metadata} use subtree partitioning to carve up the namespace and distribute it across a cluster.
+
+CephFS moves units of the namespace called directory fragments and assigns them to servers using  its own metadata load balancer.
+It then uses a hard-coded policy to balance these fragments using a secularized load metric based on CPU, workload, and file system operation metrics~\cite{sevilla:sc15-mantle}.
+
 \section{Malacology Implementation}
@@
 
-Many distributed file systems decouple metadata and data I/O so that
-these services can scale independently~\cite{alam:pdsw2011-metadata-scaling,ghemawat:sosp2003-gfs,hildebrand:msst2005-pnfs,weil_ceph_2006,welch:fast2008-panasas,shvachko:login2012-hdfs-scalability}.
-Despite
-this optimization, scaling the metadata services is still difficult
-because metadata accesses impose small and frequent requests on the
-underlying storage syste~\cite{roselli:atec2000-FS-workloads}. Many
-techniques for designing the metadata services have been proposed to
-accommodate this workload: Lustre~\cite{konstantinos:pdsw2014-lustre-metadata}, GFS~\cite{ghemawat:sosp2003-gfs}, and HDFS~\cite{shvachko:login2012-hdfs-scalability} keep all metadata on one
-server; GIGA+~\cite{patil:fast2011-giga}, IndexFS~\cite{ren:sc2014-indexfs}, Lazy Hybrid~\cite{brandt:msst2003-lh}, GPFS~\cite{schmuck:fast2002-gpfs},
-and pNFS~\cite{hildebrand:supercomputing2006-pNFS} hash the file
-system namespace across a dedicated cluster and cache metadata; Panasas~\cite{welch:fast2008-panasas} and CephFS~\cite{weil:sc2004-dyn-metadata} partition the file system namespace
-into \# d assign them to servers. These systems have novel mechanisms
-for Mantle~\cite{sevilla:sc15-mantle} is a programmable metadata
-balancer that separates the metadata balancing policies from their
-mechanisms. Admistrators inject code to change how the metadata cluster
+Mantle~\cite{sevilla:sc15-mantle} is a programmable metadata
+balancer for Ceph that separates the metadata balancing policies from their
+mechanisms. Administrators inject code to change how the metadata cluster
 distributes metadata. In the paper, we showed how to implement a single
 node metadata service, a distributed metadata services with hashing, and
-a distributed metadata service with dynamic subtree partitioning.
+a distributed metadata service with dynamic subtree ipartitioning.
 
@@
 down.
+
+We implemented it using the components from section 3. I will tell you about them here.
 

### 2016-05-09 paper/paper.tex
sec5
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-05-09 paper/paper.tex
@@
 The second service implemented on Malacology is ZLog, a high-performance
-distributed shared-log that is based on the CORFU protocol~\cite{corfu}. A
+distributed shared-log that is based on the CORFU
+protocol~\cite{balakrishnan_corfu_2012}. A
 shared-log is a general, yet powerful abstraction used to build many
@@
 bottleneck that restricts throughput. In contrast, the CORFU protocol is able
-to achieve high throughput network counter service (called a \emph{sequencer})
-that avoids all persistent I/O in the common case.
+to achieve high throughput using a network counter service, called a 
+\emph{sequencer}, that avoids all persistent I/O in the common case.
 
 While a full description of the CORFU system is beyond the scope of this
-paper, we'll briefly describe the storage interface, sequencer, and recovery
-protocol that form the core components of CORFU that we implement within
-Malacology:
+paper, we'll briefly describe the custom storage device interface, sequencer
+service, and recovery protocol that form the core components of CORFU that we
+implement within Malacology:
 
@@
 High-performance in CORFU is achived using a sequencer service that assigns
-client log positions by reading from a volatile, in-memory counter which can
-run at a very high throughput and at low-latency. Since the sequencer is
-centralized, ensuring seiralizability in the common case is trivial.  The
+log positions to clients by reading from a volatile, in-memory counter which
+can run at a very high throughput and at low-latency. And since the sequencer
+is centralized, ensuring seiralizability in the common case is trivial.  The
 primary challenge in CORFU is handling the failure of the sequencer in a way
-that guarantees correctness.  Failures of the sequencer service in CORFU are
+that preserves correctness. Failure of the sequencer service in CORFU is
 handled by a recovery algorithm that recomputes the new sequencer state using
 an application-specific custom storage interface to discover the tail of the
-log and invalidate stale client requests using an epoch-based protocol.
+log, while simulatenously invalidating stale client requests using an
+epoch-based protocol.
 
 {\bf Storage Interface.} The storage interface is a critical component in the
-CORFU design. Storage devices must provide a \emph{write}-once, random
-\emph{read} interface for reading and writing log entries. The key to
-correctness in CORFU lies with the enforcement of up-to-date epoch tags on
-client requests; requests tagged with out-of-date epoch values are rejected,
-and clients are expected to request a new tail from the sequencer after
-refreshing state from an auxillary service. This mechanism forms the basis for
-sequencer recovery.
+CORFU design. Clients independently map log positions onto storage devices,
+while storage devices provide an intelligent \emph{write}-once, random
+\emph{read} interface for accessing log entries. The key to correctness in
+CORFU lies with the enforcement of up-to-date epoch tags on client requests;
+requests tagged with out-of-date epoch values are rejected, and clients are
+expected to request a new tail from the sequencer after refreshing state from
+an auxillary service. This mechanism forms the basis for sequencer recovery.
 
@@
 
-\subsubsection{Log Management}
-
-log management is supported by mds using a special directory layout
-and file types.
-
 \subsubsection{Sequencer}
+
+We implement the sequencer service in Malacology as an application-specific
+file type that attaches a 64-bit value to the inode state within the Ceph
+metadata service.
 
@@
 
-%\begin{figure}[htbp]
-%\centering
-%\includegraphics{figures/overview-zlog.png}
-%\caption{This is ZLog.}
-%\end{figure}
-
-\subsection{Interface Installation}
+\subsubsection{Log Management}
+
+log management is supported by mds using a special directory layout
+and file types.
+
+\subsubsection{Storage Interface Management}
+
+lua snippet
 

### 2016-05-09 experiments/ceph-deploy/run.sh
experiment: add results for sequencer migrate
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,12 +1,0 @@
-#!/bin/bash
-
-set -ex
-
-for i in 0 1 2 3 4 5; do
-  for w in test; do
-    for s in ceph-ext4; do
-      ANSIBLE_LOG_PATH="logs/$w-$s-$i.log" \
-      ansible-playbook --skip-tags "with_pkg" $s.yml workloads/$w.yml
-    done
-  done
-done

### 2016-05-09 experiments/mds-zlog-seq-migrate/run.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+
+set -ex
+
+ANSIBLE_LOG_PATH="logs/ceph-zeqr.log" \
+ansible-playbook --forks 50 --skip-tags "with_pkg" \
+ceph.yml workloads/zeqr.yml workloads/zeqr-mantle.yml collect.yml

### 2016-05-09 paper/paper.tex
Adds some straw man examples to section 4

@@
 CORFU protocol~\cite{corfu}.
-Without going into greater detail, the CORFU protocol by separating the concern of correctness from the concern of performance enables serialization using a \emph{volatile}
+Without going into greater detail, the CORFU protocol, by separating the concern of correctness from the concern of performance, enables serialization using a \emph{volatile}
 ``sequencer'' service. This high-performance in-memory counter is used to assign a log
@@
 
-Ceph is a production-quality distributed system and in this section we
+Ceph is a production-quality distributed storage system and in this section we
 touch on some of the subsystems it uses to provide a general-purpose
-storage system. In the next section, we describe how we leverage these
-subsystems to build new services. This is the an exhuastive list of services.
+storage system. In the next section, we describe how Malacology exposes these
+subsystems in order to enable the creation of new services.
 
@@
 
-A consistent view of cluster state among Ceph daemons and clients is critical
+% straw man example
+Keeping track of the state of a distributed system is an essential part of any 
+successful service and a necessary component in order to diagnose and detect 
+failures, when they occur. In the case of Ceph, a consistent view of cluster 
+state among server daemons and clients is critical
 to provide strong consistency guarantees to clients. Ceph maintains cluster
@@
 
-Ceph provides a rich plugin framework for defining new interfaces to objects
+Allowing users to define custom code that executes in the storage nodes is a 
+useful technique for offloading computation closer to where the data is stored 
+in order to significantly reduce the amount of data moved between storage 
+servers and clients. Ceph provides a rich plugin framework for defining new 
+interfaces to objects
 stored within RADOS, effectively allowing developers to use co-design
@@
 a better position to know what to do.
-
-Mantle is a framework that allows balancing policies to be written in Lua,
-and supports such policies be specified at run-time.
 
@@
 mechanisms. Administrators inject code to change how the metadata cluster
-distributes metadata. In the paper, we showed how to implement a single
-node metadata service, a distributed metadata services with hashing, and
-a distributed metadata service with dynamic subtree ipartitioning.
+distributes metadata. In ~\cite{sevilla:sc15-mantle} the authors showed how 
+using Mantle, one can implement a single node metadata service, a distributed 
+metadata services with hashing, and a distributed metadata service with dynamic 
+subtree partitioning.
 
@@
 
+Mantle is a framework that allows balancing policies to be written in Lua,
+and supports such policies be specified at run-time.
 We implemented it using the components from section 3. I will tell you about them here.

### 2016-05-10 paper/paper.tex
food and stuff
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-05-09 paper/paper.tex
@@
 We implement the sequencer service in Malacology as an application-specific
-file type that attaches a 64-bit value to the inode state within the Ceph
-metadata service.
-
-The zlog sequencer is volatile, shared resource that requires serialized
-access. in corfu the sequencer is a fast, centeralized network counter
-resource.
-
-Initially we saught to directly model the sequencer in ceph as a non-exclusive,
-non-cacheable resource, forcing clients to perform round-trip access the
-resource at the authoratiave MDS for the inode representing the sequencer.
-However, we found that the capapbility system in Ceph is used to batch access
-to cacheable resources, providing clients best-effort access to the capability
-and allowing them to burst. While unexpected, this alternative mode provides
-a different approach to sequencer implementation.
-
-In Section we quantify the trade-offs of throughput and latency for the
-centeralized sequencer and the batching mechanism based on capabilities.
-
-The MDS has a queue of locks that it uses to revoke capabilities from
-the clients. If we put a Mantle-style hook here, we can control when
-those revoke messages go out.
-
-Mantle is used to implement a zlog file type specific policy for maximizing
-throughput to the sequencer by migration the resource between metadata servers.
-
-Sequencer recovery is handled by tapping into existing recovery mechanisms
-that recover inode state. file type specific recovery augments the existing
-recovery procedure.
-
-\subsubsection{Log Management}
-
-log management is supported by mds using a special directory layout
-and file types.
-
-\subsubsection{Storage Interface Management}
-
-lua snippet
-
-The lua implementation of the interfaces are encoded into cluster maps
-for automatic verisoning and dissemtation across the cluster.
+file type that associates a small amount of metadata, including the 64-bit log
+tail position, to the inode state managed by the Ceph metadata service.  Since
+a sequencer is associated with a unique log, this approach has the added
+benefit of allowing the metadata service to also handle naming, by
+representing each log service instance within the standard POSIX hierarchical
+namespace.
+
+{\bf Physical design.}
+The next component of our design is the low-level configuration of the
+sequencer inode type that achieves the correct semantics.  Initially we sought
+to directly model the sequencer service in Ceph as a non-exclusive,
+non-cacheable resource, forcing clients to perform a round-trip access to the
+resource at the authoratiave metadata server for the inode representing the
+sequencer.  Interestingly, we found that the capapbility system in Ceph uses a
+strategy to reduce metadata service load by allowing multiple clients that
+open a shared file to temporarily cache the inode resource, resulting in a
+round-robin, best-effort batching behavior. When a single client is accessing
+the sequencer resource it is able to cache state locally without interruption.
+
+While unexpected, this discovery allowed us to explore an implementation
+strategy that we had not previously considered. In particular, for bursty
+clients and clients that can tolerate high latency, this mode of operation may
+allow a system to achieve much higher throughput than a system with a
+centralized sequencer service.
+
+In Section we quantify the trade-offs of throughput and latency for an
+approach based on a round-robin batching mode, and compare this mode to one in
+which the metadata server mediates access to the sequencer state when in it is
+being shared among multiple clients.
+
+{\bf Capability policies.}
+We use capabilitity policies in the metadata service to tune the amount of
+time that clients may cache the sequencer resource. This allows us to control
+the trade-off between latency and throughput beyond the standard best-effort
+policy.
+
+{\bf Balancing policies.}
+While the metadata service is a shared resource, authority over metadata
+resources can be physically moved between servers using a dynamic, ZLog-aware
+policy expressed in Mantle. For instance, one or more metadata servers may be
+dedicated to supporting ZLog sequencer load. When the system load is low all
+resources are spread evenly, but non-ZLog resources may be migrated to
+best-effort servers when sequencer performance suffers.
+
+{\bf Recovery.}
+What?
 

### 2016-05-10 paper/paper.tex
SAUSAGE
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-\subsection{MDS as ZLog Sequencer}\label{mds-as-zlog-sequencer}
-
-\subsubsection{Experiment: Object Class
-Programmability}\label{experiment-object-class-programmability}
-
-Add quote from CORFU paper that talks about what they had to hack
-
-\subsubsection{Experiment: Multi-Client
-Burstiness}\label{experiment-multi-client-burstiness}
-
-\begin{figure}[htbp]
+\subsection{ZLog}
+
+\subsubsection{Sequencer}
+
+We evaluated the performance of the ZLog Sequencer implemented in several
+different configurations by showing a trade-off between aggregate throughput
+achieved (shown in Figure~\ref{fig:captp}), and latency distribution (shown in
+Figure~\ref{fig:capcdf}).
+
+In Figure~\ref{fig:captp} the highest throughput labeled $1C$ shows that a
+single client caching the sequencer resource may obtain new sequences at a rate
+of nearly 6 million positions per second. This is not surprising given that all
+state is cached locally in the client. When this resource is shared between two
+clients, the MDS uses a best-effort approach to granting the sequencer resource
+to each client (value $2C,BE$). This sharing has a large impact on performance,
+as shown by the nearly three orders of magnitude drop off in throughput.
+
+\begin{figure}[h]
 \centering
 \includegraphics{figures/caps-delay-thruput.png}
-\caption{Forcing the client to drop their capabilities later (delay)
-improves throughput}
+\caption{Sequencer throughput by reusing various services.
+The highest performance is achieved using a single client with
+exclusive, cacheable priviledge. Round-robin sharing of the sequencer
+resource is affected by the amount of time the resource is held,
+with best-effort performing the worst.}
+\label{fig:captp}
 \end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics{figures/caps-delay-latency.png}
+\caption{But if the clients hold their capabilities longer, it hurts
+latency for everyone.}
+\label{fig:capcdf}
+\end{figure}
+
+While a best-effort approach to sharing capabilities makes sense for a file
+system client, throughput drops significantly with two clients as the majority
+of the time neither client has the capapbitility as it ping-pongs between
+clients. This observation reveals a tuning parameter: how long a client
+maintains its capapbility lease. In throughputs labeled $2C,<time>$ we show
+aggregate throughput as we decrease the amount of time that a client holds its
+lease before responding to a request for the capapbility.  At lease times as
+low as 10ms throughputs achieve a significant proportion of the locally cached
+throughput.
+
+While the batching mechanism can achieve high-throughput, it is not appropriate
+for applications that require predictable latency. The throughput in
+Figure~\ref{fig:captp} labeled $2C,ZLog$ corresponds to a centeralized
+sequencer implementation. While throughput is lower for two clients (because of
+the synchronous network round-trip required for each request),
+Figure~\ref{fig:capcdf} shows that this approach achieves more predicable
+latency compared to all other approaches except for the single client locally
+cached mode.
+
+\subsubsection{Interface Propogation}
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/caps-delay-latency.png}
-\caption{But if the clients hold their capabilities longer, it hurts
-latency for everyone.}
-\end{figure}
-
-\subsubsection{CUT THIS Experiment: Client v. MDS caps +
-Failures}\label{cut-this-experiment-client-v.-mds-caps-failures}
-
-\subsubsection{Experiment: Mantle Sequencer
-Balancer}\label{experiment-mantle-sequencer-balancer}
+%\subsubsection{Experiment: Object Class
+%Programmability}\label{experiment-object-class-programmability}
+%
+%Add quote from CORFU paper that talks about what they had to hack
+%
+%\subsubsection{Experiment: Multi-Client
+%Burstiness}\label{experiment-multi-client-burstiness}
+%
+%\begin{figure}[htbp]
+%\centering
+%\includegraphics{figures/caps-delay-thruput.png}
+%\caption{Forcing the client to drop their capabilities later (delay)
+%improves throughput}
+%\end{figure}
 

### 2016-05-10 paper/paper.tex
commit comment
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 generalized form can be used to \emph{program} new services. By doing so
-one can reuse subsystems and their optimizations and leverage their
+one can reuse subsystems and leverage their optimizations,
 established correctness, robustness, and efficiency. We will refer to
@@
 designed and evaluated Malacology, a programmable storage system for
-constructing new functionality by re-purposing existing subsystem
-abstractions contained in Ceph. We build the framework on Ceph by
+constructing new services by re-purposing existing subsystem
+abstractions contained in Ceph. We build the framework in Ceph by
 leveraging a broad spectrum of existing services, including distributed
-locking and caching services provides by MDSs, durability and logical
+locking and caching services provided by MDSs, durability and logical
 storage devices provided by RADOS, and propogation of consistent cluster
@@
 The second approach that applications may use to adapt to a storage system
-limitation is to
-extend the responsibility of the application. This means changing the
-application itself by adding more data management intelligence or
+limitation is to extend the responsibility of the application. This means
+changing the application itself by adding more data management intelligence or
 using domain-specific middleware (e.g., a new data layout). For instance, an
-application may be changed to exploit data locality or I/O
-parallelism in a distributed storage system.
-
-For example, SciHadoop~\cite{buck:hpc2011-scihadoop,buck:sc2013-scidr} changes both the Hadoop
-application and the Hadoop framework itself to leverage the structure of
-scientific data (3D array-based data, more specifically) to increase
-performance, locality, and the number of early results. %This is not a
+application may be changed to exploit data locality or I/O parallelism in a
+distributed storage system.
+In SciHadoop~\cite{buck:hpc2011-scihadoop,buck:sc2013-scidr} both the Hadoop
+application and the Hadoop framework itself are changed to leverage the structure of
+scientific data (3D array-based data, more specifically) to
+improve locality, performance, and increase the number of early results. %This is not a
 %bad proposition, but 
-Changing applications creates a coupling that is highly tied to the
+However, changing applications creates a coupling that is highly tied to the
 underlying physical properties of the system, making it difficult to
@@
 Traditionally, storage system behaviour can be altered using two
-techniques: tuning the system or introducing changes to the system. The
-difficulty of both of these approaches has given rise to a third
-technique called active storage.
+techniques: tuning the system or introducing changes to the system.
+%The
+%difficulty of both of these approaches has given rise to a third
+%technique called active storage.
 
@@
 The object interface facility provided by Ceph is more than
-a proof-of-concept active storage implementation.  Figure~\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
+a proof-of-concept implementation resembling active storage.  Figure~\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
 domain-specific interfaces in the Ceph community since 2010.  What is most
@@
 
-The popularity of the active storage component of Ceph hint at three
+The popularity of the custom object interface facility of Ceph hint at three
 trends in the Ceph community: (1) increasingly, the default
@@
 system is not as robust as Ceph and the Ceph team wants more safety,
-durability, and consistency for the new functionality. For example, the
+durability, and consistency for the new services. For example, the
 original Mantle API is fragile: the API is not enforced by the runtime,

### 2016-05-10 paper/paper.tex
Edits to section 4 intro and 4.1

@@
 To illustrate the benefits and challenges of this approach we examine
-the programmability of Ceph~\cite{weil_ceph_2006}, the increasingly
+the programmability of Ceph~\cite{weil:osdi2006-ceph}, the increasingly
 popular, production-level open-source distributed storage system.
@@
 a high-performance distributed shared-log on Ceph called ZLog, based on the
-CORFU protocol~\cite{corfu}.
+CORFU protocol~\cite{balakrishnan_corfu_2012}.
 Without going into greater detail, the CORFU protocol, by separating the concern of correctness from the concern of performance, enables serialization using a \emph{volatile}
@@
 
-\section{Re-Usable Components in Ceph}\label{re-usable-components-in-ceph}
+\section{Components of Ceph Suitable for Reusability}\label{re-usable-components-in-ceph}
 
@@
 storage system. In the next section, we describe how Malacology exposes these
-subsystems in order to enable the creation of new services.
+subsystems in order to make them reusable and enable the creation of new services on top of them.
 
@@
 for maintaining consistent data versions, and are a useful system to expose.
+
 While Ceph does not expose this service directly, it does allow limited use by
 clients through a key-value service intended for low-volume, small size
-configuration information.
+configuration information. Unfortunately, while a key-value service is
+generally useful, it doesn't provide many of the useful services hidden within
+the monitoring framework that would be required for applications with more
+demanding requirements.
 
@@
 
-Although this feature is powerful, the current implementation has drawbacks.
-Supporting only C/C++ for object interface developers, Ceph requires
-distribution of compiled binaries for the correct architecture, adding a
-large barrier of entry for developers and system administrators. Second,
-having no way to dynamically unload modules, any changes require a full
-restart of an OSD which may have serious performance impacts due to a loss
-of cached data. And finally, the security limitations of the framework
-limit the use of object classes to all but those with administrative level
-access and technique expertise.
-
 \begin{figure}[htbp]
@@
 
-\subsection{Load Balancing}
+\subsection{Metadata Load Balancing}
 
@@
 
-In this section we describe how Malacology exposes and builds upon existing
-services within Ceph. While we strive for implementation independent
-approaches within Malacology, our prototype implementation makes heavy use of
-the embedded language Lua to enable dynamic features without relying on
-complex and heavy-weight code generation techniques.
-
-{\bf Lua integration.}
-Supporting dynamic reconfiguration and programmabiilty is critical in real
-systems that must adapt to a changing environment where potentially
-conflicting policies arise. We've adopted Lua as a scripting language to
-implement dynamic components in Malacology.  Lua is a portable, embedded
+In this section we describe how Malacology builds upon existing
+services within Ceph. The main idea behind Malacology is to take the subsystems 
+described in the previous section and expose APIs on top of them that allow 
+users to dynamically inject code, allowing them to customize the behavior of 
+these components. Supporting dynamic reconfiguration and programmability 
+efficiently is critical in real systems that must adapt to a changing 
+environment where potentially
+conflicting policies arise. We've adopted Lua as the dynamic runtime  to
+implement dynamic components in Malacology.
+
+There are multiple reasons that make Lua an attractive runtime for implementing 
+Malacology. Lua is a portable, embedded
 scripting language that offers superious performance and productivity
 trade-offs, including a JIT-based implementation that is well known for near
-native performance. Lua has been used extensively in game engines, and systems
-research \cite{neto:dls14-luaos}, including storage systems where it has been
-effectively used both on
-\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo} and off
-\cite{sevilla:sc15-mantle} the performance critical path.
+native performance. Additionally, Lua has been used extensively in game engines, 
+and systems research \cite{neto:dls14-luaos}, including storage systems where it 
+has been effectively used both on
+\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010} and 
+off
+\cite{sevilla:sc15-mantle} the performance critical path. Finally, the 
+flexibility of the runtime allows to sandbox applications in order to address 
+security and performance concerns. We now describe how Malacology exposes each 
+subcomponent in Ceph.
+
+\subsection{Object Storage}
+
+% todo:
+%  - ellude to the need for interfaces to be avialable quickly to support
+%    our experiment about propogtation delay
+%
+
+The implementation of Ceph's object abstraction, although powerful, has some 
+drawbacks.
+Supporting only C/C++ for object interface developers, Ceph requires
+distribution of compiled binaries for the correct architecture, adding a
+large barrier of entry for developers and system administrators. Second,
+having no way to dynamically unload modules, any changes require a full
+restart of an OSD which may have serious performance impacts due to a loss
+of cached data. And finally, the security limitations of the framework
+limit the use of object classes to all but those with administrative level
+access and technique expertise.
+
+By embedding the Lua runtime, Malacology makes the base RADOS API available in a 
+dynamic environment, resulting in a object storage API with economy of 
+expression, which at the same time provides the full set of features of the base 
+object class. New object interfaces that are expressed in thousands of lines of 
+code can be implemented in approximately an order of magnitude less 
+code~\cite{geambasu_comet_2010}.
 
@@
 {\bf Balancing Policy.}
-Avoiding hard-coded balancing policies are important. First, it is difficult,
+
+Avoiding hard-coded balancing policies is important. First, it is difficult,
 nay impossible, to select a one-size-fits-all policy. System administrators
@@
 want to.
-
-\subsection{Object Storage}
-
-% todo:
-%  - ellude to the need for interfaces to be avialable quickly to support
-%    our experiment about propogtation delay
-%
-
-Lua integration. Overhead?
-
-Although these OSD interfaces are powerful, the current implementation
-has drawbacks. OSD interfaces are written in C/C++ and compiled into a
-shared library, so the developer must account for different target
-architectures. Second, C/C++ provides more functionality than is
-necessary since the snippets of interface code mainly set policies and
-perform simple operations. Finally, developing these interfaces in C/C++
-has high overhead. The developer must learn how the OSD dynamically
 

### 2016-05-10 paper/paper.tex
rel work
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>
Changes to section 1 and 2

--- 2016-05-10 paper/paper.tex
@@
 
-\section{Introduction}\label{introduction}
+\section{Introduction}
+\label{introduction}
 
@@
 To illustrate the benefits and challenges of this approach we examine
-the programmability of Ceph~\cite{weil:osdi2006-ceph}, the increasingly
+the programmability of Ceph~\cite{weil:osdi2006-ceph,ceph}, the increasingly
 popular, production-level open-source distributed storage system.
@@
 storage devices provided by RADOS, and propogation of consistent cluster
-state provided by the monitoring service (see \ref{fig:overview}). As we will show in this paper, this framework is expressive enough to provide the
+state provided by the monitoring service (see Figure~\ref{fig:overview}). As we will show in this paper, this framework is expressive enough to provide the
 functionality necessary for implementing new services. Our contributions
@@
 
-\paragraph*{A programmable storage system implementation} that re-uses and extends existing abstractions. It includes:
+\paragraph*{A programmable storage system} that re-uses and extends existing abstractions and services of Ceph, a common production-level storage system. It includes:
 
@@
   \item
-    An extension of file system metadata by a new type of ``sequencer'' inode that together with Ceph's capability subsystem implements high-performance, non-durable access serialization
+    An extension of file system metadata by a new type of ``sequencer'' inode that together with Ceph's capability subsystem implements high-performance, non-durable access serialization:
   \item
@@
 
-\paragraph*{Two example implementations and evaluations} for feasibility of services that use this framework
+\paragraph*{A distributed shared log service} to demonstrate the feasibility of services using the programmable storage approach:
 
@@
   \item 
-    An application of the Mantle metadata load balancing service to manage the placement of new sequencer inodes
+    An application of Mantle to manage load-balancing of sequencers.
   \end{enumerate}
 
-First, we describe existing techniques for bridging the gap between the
-performance of storage systems and the requirements of applications
+First, we describe existing techniques for adapting storage systems
 (\S\ref{problem}). Next, we discuss the ``distributed systems'' inspired
@@
 
-\section{Highly Tailored and Application-Specifc Storage
-Systems}\label{highly-tailored-and-application-specifc-storage-systems}
+\section{Adapting Storage Systems}
 
@@
 devices is that storage systems may fail to meet the needs of applications.
-Application workarounds for meeting performance goals roughly fall into one of
-three categories: ``bolt-on'' services, application changes, and changes to
-the storage system itself.
-
-\subsection{\texorpdfstring{``Bolt-on''
-services}{Bolt-on services}}\label{bolt-on-services}
-
-So called ``bolt-on'' services are third-party solutions integrated into a
+Short of changing the appalication itself, approaches for meeting performance goals roughly fall into ``bolt-on'' services and changes to the storage system itself.
+
+\subsection{Bolt-on services}
+\label{bolt-on-services}
+
+Bolt-on services are third-party solutions integrated into a
 system to provide functionality and improve performance for common use-cases
@@
 a high-performance distributed shared-log on Ceph called ZLog, based on the
-CORFU protocol~\cite{balakrishnan_corfu_2012}.
-Without going into greater detail, the CORFU protocol, by separating the concern of correctness from the concern of performance, enables serialization using a \emph{volatile}
+CORFU protocol~\cite{balakrishnan_corfu_2012}. The CORFU protocol separates the concern of correctness from the concern of performance, enables serialization using a \emph{volatile}
 ``sequencer'' service. This high-performance in-memory counter is used to assign a log
@@
 The sequencer is purely a performance optimization and does not have a role in correctness (if the sequencer fails, the tail of the log can always be retrieved by other means).
-Thus, it is the volatility of the sequencer that enables high performance.
+Thus, it is the volatility of the sequencer that enables high performance (see Section~\ref{sec:zlog} for more detail).
 A ``bolt-on'' option for building
@@
 
+\if{false}
 \subsection{Application Changes}\label{application-changes}
@@
 adapt to future changes at the storage system level.
+\fi
 
@@
 
-When both the bolt-on approach and changing applications fail to meet the needs of the application,
-developers turn their attention to the storage system itself.
-Traditionally, storage system behaviour can be altered using two
+Storage system behaviour can be altered using two
 techniques: tuning the system or introducing changes to the system.
@@
 
-As a last resort, the application developer can alter the storage system
+To go beyond tuning one has to alter the storage system
 itself. In the best case a developer is using an open-source system in which
@@
 \subsection{Mantle: A Programmable Metadata Load Balancer}
+\label{sec:mantle}
 
@@
 \subsection{ZLog: A Fast Distributed Shared-Log}
+\label{sec:zlog}
 

### 2016-05-10 paper/paper.tex
paper: add final graphs for Mantle
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-05-10 paper/paper.tex
@@
 
-\subsection{Mantle}\label{mantle}
-
-\subsubsection{Experiment: FileBench}\label{experiment-filebench}
-
-\subsubsection{Experiment: OpenSSL
-Compile}\label{experiment-openssl-compile}
-
 \subsection{ZLog}
@@
 ms.
+
+\subsection{Mantle}
+
+\subsection{MDS as a Sequencer}
+\begin{figure}[h]
+\centering
+\includegraphics{figures/mantle-seq-thruput.png}
+\caption{A ZLog sequencer that uses the CephFS metadata cluster enjoys all the benefits of the MDS recovery ("Sequencer Recovery") and load balancing protocols (``Mantle Load Balancing'').
+The ``2 Seq'' curves are the throughputs of 1 sequencer when an additional sequencer is running in the background generating load.
+The ``Sequencer Baseline'' has no background sequencers.}
+\label{fig:mantle-seq-thruput.png}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics{figures/mantle-cpu-utilization-annotate.png}
+\caption{The sequencer using Mantle in Figure~\ref{fig:mantle-seq-thruput.png} migrates to a new MDS immediately because its uses a policy designed for bursty workloads.
+This is a trace of the same workload running twice: once with a single MDS and once with Mantle.}
+\label{fig:mantle-cpu-utilization}
+\end{figure}
+
+\subsection{Load Balancing}
+
+\subsection{Recovery}
 

### 2016-05-10 paper/paper.tex
Removed section 2 and moved empirical evidence into intro

--- 2016-05-10 paper/paper.tex
@@
 
+\if{false}
 For example, a Ceph cluster contains a small number of
@@
 code-hardening work.
-
-To illustrate the benefits and challenges of this approach we have
-designed and evaluated Malacology, a programmable storage system for
-constructing new services by re-purposing existing subsystem
-abstractions contained in Ceph. We build the framework in Ceph by
-leveraging a broad spectrum of existing services, including distributed
-locking and caching services provided by MDSs, durability and logical
-storage devices provided by RADOS, and propogation of consistent cluster
-state provided by the monitoring service (see Figure~\ref{fig:overview}). As we will show in this paper, this framework is expressive enough to provide the
-functionality necessary for implementing new services. Our contributions
-are:
-
-\paragraph*{A programmable storage system} that re-uses and extends existing abstractions and services of Ceph, a common production-level storage system. It includes:
-
-  \begin{enumerate}
-%  \def\labelenumi{\arabic{enumi}.}
-%  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    An extension of file system metadata by a new type of ``sequencer'' inode that together with Ceph's capability subsystem implements high-performance, non-durable access serialization:
-  \item
-    The use of Ceph's PAXOS-based consensus infrastructure for versioning of metadata service load-balancing policies
-  \item
-    The use of storage objects for persisting policies
-  \item
-    An extension of the object interface classes for new logical storage/metadata devices
-  \end{enumerate}
-
-\paragraph*{A distributed shared log service} to demonstrate the feasibility of services using the programmable storage approach:
-
-  \begin{enumerate}
-%  \def\labelenumi{\arabic{enumi}.}
-%  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    A high-performance distributed shared log service called Zlog which is based on CORFU~\cite{balakrishnan_corfu_2012}
-  \item
-    A re-implementation of the programmable Mantle metadata load balancing service~\cite{sevilla:sc15-mantle} using Malacology
-  \item 
-    An application of Mantle to manage load-balancing of sequencers.
-  \end{enumerate}
-
-First, we describe existing techniques for adapting storage systems
-(\S\ref{problem}). Next, we discuss the ``distributed systems'' inspired
-abstractions employed by Ceph (\S\ref{background}) and allude to their
-re-usability as components in other services. The next section
-enumerates the abstractions that Malacology exposes in the
-implementation section (\S\ref{implementation}). We then show how we use
-Malacology to synthesize entirely new storage services on an existing
-system through configuration and small changes (\S\ref{services}). We
-conclude by evaluating real-world use cases (\S\ref{evaluation}).
-
-\section{Adapting Storage Systems}
-
-\label{problem}
-
-A consequence of complicated data management requirements and faster storage
-devices is that storage systems may fail to meet the needs of applications.
-Short of changing the appalication itself, approaches for meeting performance goals roughly fall into ``bolt-on'' services and changes to the storage system itself.
-
-\subsection{Bolt-on services}
-\label{bolt-on-services}
-
-Bolt-on services are third-party solutions integrated into a
-system to provide functionality and improve performance for common use-cases
-(e.g. a metadata service). The introduction of additional services may provide
-simplicity and performance enhancements, but ``bolt-on'' services often
-duplicate functionality, unnecessarily increasing the likelihood of bugs or
-introducing unpredictable performance problems. For example, we are developing
-a high-performance distributed shared-log on Ceph called ZLog, based on the
-CORFU protocol~\cite{balakrishnan_corfu_2012}. The CORFU protocol separates the concern of correctness from the concern of performance, enables serialization using a \emph{volatile}
-``sequencer'' service. This high-performance in-memory counter is used to assign a log
-position to each client that appends to the log.
-The sequencer is purely a performance optimization and does not have a role in correctness (if the sequencer fails, the tail of the log can always be retrieved by other means).
-Thus, it is the volatility of the sequencer that enables high performance (see Section~\ref{sec:zlog} for more detail).
-A ``bolt-on'' option for building
-a distributed log is to use PAXOS, or an out-of-the-box solution such as
-Zookeeper. However, these systems adhere to the preconceived notions of
-storage, namely that data is safe from data loss, which is exactly the feature
-that causes traditional approaches to run at significantly reduced speed.
-
-\if{false}
-\subsection{Application Changes}\label{application-changes}
-
-The second approach that applications may use to adapt to a storage system
-limitation is to extend the responsibility of the application. This means
-changing the application itself by adding more data management intelligence or
-using domain-specific middleware (e.g., a new data layout). For instance, an
-application may be changed to exploit data locality or I/O parallelism in a
-distributed storage system.
-In SciHadoop~\cite{buck:hpc2011-scihadoop,buck:sc2013-scidr} both the Hadoop
-application and the Hadoop framework itself are changed to leverage the structure of
-scientific data (3D array-based data, more specifically) to
-improve locality, performance, and increase the number of early results. %This is not a
-%bad proposition, but 
-However, changing applications creates a coupling that is highly tied to the
-underlying physical properties of the system, making it difficult to
-adapt to future changes at the storage system level.
 \fi
 
-\subsection{Storage System Changes}\label{storage-changes}
-
-Storage system behaviour can be altered using two
-techniques: tuning the system or introducing changes to the system.
-%The
-%difficulty of both of these approaches has given rise to a third
-%technique called active storage.
-
-\subsubsection{Tuning}\label{tuning}
-
-Tuning large and complicated systems is difficult because of the number
-of parameters (e.g., Hadoop 2.7.1 has 210 tunables) and because the
-tunables are difficult to understand or quantify (e.g., Ceph tunables~\cite{sevilla:sc15-mantle}). To succesfully tune the system, a
-developer must have domain and system-specific knowledge. Without this
-intimate knowledge, the tuning turns into somewhat of a black art, where
-the only way to figure out the best settings is trial and error.
-
-Auto-tuning techniques attempt to find a good solution among a huge space of
-available system configurations. However, in practice auto-tuning is limited
-to only the configuration tunables that the storage system exposes and can be
-overwhelmed with too many parameters.  For instance, auto-tuning may be
-capable of identifying instances in which new data layouts would benefit a
-workload, but unless the system can provide such a transformation, the option
-is left off the table.  Starfish~\cite{herodotou:cidr2011-starfish} made
-an attempt at this for Hadoop but this technique would need to severely limit
-the space of parameters in order to be feasible, similar to the approach used in~\cite{behzad:sc2013-autotuning}.
-
-\subsubsection{Software Changes}\label{software-changes}
-
-To go beyond tuning one has to alter the storage system
-itself. In the best case a developer is using an open-source system in which
-case one must become familiar with often enormous code-bases before being able
-to make changes, but then face the task of approaching a community to
-integrate changes, or be forced to maintain an in-house version and risk
-maintaining features that become incompatible.  A successful example of this
-is the work that focused on HDFS scalability for metadata-intensive workloads~\cite{shvachko_hdfs_2010}. This has lead to modifications to its
-architecture or API~\cite{balmin:sigmod2012-clydesdale} to improve
-performance.
-
-Alternatively, some systems have formal support for introducing additional
-functionality, often in the form of a plugin infrastructure. Ceph provides the
-ability to create domain-specific logical object interfaces by composing
+Ceph also provides empirical evidence that developers embrace programmable storage: Ceph already supports the
+implementation of domain-specific logical object interfaces by composing
 existing interfaces. These interfaces are written in C++ and are statically
-loaded into the system. %An example usage may be to atomically combine a
-%``write'' with an index update to create a content-addressable object
-%interface.
-
-\if{false}
-As an example use-case, ZLog needs to
-store metadata information with each object, so for each operation, it
-checks the epoch value and write capability (2 metadata reads) and
-writes the index (metadata write).
-\fi
-
-For example, Figure~\ref{fig:datavsxattr} compares the throughput
-(\emph{y} axis) over time (\emph{x} axis) of two OSD implementations of the same storage object interface for
-storing metadata: either in the header of the byte stream (data) or in the
-object extended attributes (XATTR). The speed for appending data without
-any metadata operations (data raw) is also shown as a baseline for
-comparison. For append-heavy workloads, storing metadata as a header in
-the data performs about 1.5x better than storing metadata as an extended
-attribute.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/cls_xattr-vs-data.png}
-\caption{{[}\href{https://github.com/michaelsevilla/malacology-popper/blob/master/experiments/figure1/visualize.ipynb}{source}{]}
-When appending data to objects, the object class that stores metadata in
-a header in the data byte stream (data) performs 1.5x better than the
-object class that stores metdata in the extended attributes of the
-object (XATTR); it is almost as fast as appending data without updating
-metadata (data raw).}
-\label{fig:datavsxattr}
-\end{figure}
-
-The object interface facility provided by Ceph is more than
-a proof-of-concept implementation resembling active storage.  Figure~\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
+loaded into the system. Figure~\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
 domain-specific interfaces in the Ceph community since 2010.  What is most
@@
 
-\if{false}
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/obj-int-dev-churn.png}
-\caption{Source code changes over time indicate the dynamic interface
-development will need to support a high frequency of interface churn.
-Each dot represents a Ceph commit with a corresponding number of lines
-of code changed. \label{fig:obj-int-dev-churn}}
-\end{figure}
-\fi
-
 The popularity of the custom object interface facility of Ceph hint at three
@@
 
-\if{false}
-These implementations use the active storage interfaces in Ceph. We
-refer to this as \emph{storage programmability} which is a method by
-which an application communicates its requirements to the storage system
-in a way that allows the application to realize a new behavior without
-sacrificing the correctness of the underlying system. This evaluation
-and resulting implementation would have been a burden without object
-classes.
-\fi
+To illustrate the benefits and challenges of this approach we have
+designed and evaluated Malacology, a programmable storage system for
+constructing new services by re-purposing existing subsystem
+abstractions contained in Ceph. We build the framework in Ceph by
+leveraging a broad spectrum of existing services, including distributed
+locking and caching services provided by MDSs, durability and logical
+storage devices provided by RADOS, and propogation of consistent cluster
+state provided by the monitoring service (see Figure~\ref{fig:overview}). As we will show in this paper, this framework is expressive enough to provide the
+functionality necessary for implementing new services. Our contributions
+are:
+
+\paragraph*{A programmable storage system} that re-uses and extends existing abstractions and services of Ceph, a common production-level storage system. It includes:
+
+  \begin{enumerate}
+%  \def\labelenumi{\arabic{enumi}.}
+%  \itemsep1pt\parskip0pt\parsep0pt
+  \item
+    An extension of file system metadata by a new type of ``sequencer'' inode that together with Ceph's capability subsystem implements high-performance, non-durable access serialization:
+  \item
+    The use of Ceph's PAXOS-based consensus infrastructure for versioning of metadata service load-balancing policies
+  \item
+    The use of storage objects for persisting policies
+  \item
+    An extension of the object interface classes for new logical storage/metadata devices
+  \end{enumerate}
+
+\paragraph*{A distributed shared log service} to demonstrate the feasibility of services using the programmable storage approach:
+
+  \begin{enumerate}
+%  \def\labelenumi{\arabic{enumi}.}
+%  \itemsep1pt\parskip0pt\parsep0pt
+  \item
+    A high-performance distributed shared log service called Zlog which is based on CORFU~\cite{balakrishnan_corfu_2012}
+  \item
+    A re-implementation of the programmable Mantle metadata load balancing service~\cite{sevilla:sc15-mantle} using Malacology
+  \item 
+    An application of Mantle to manage load-balancing of sequencers.
+  \end{enumerate}
+
+First, we describe existing techniques for adapting storage systems
+(\S\ref{sec:evidence}). Next, we discuss the ``distributed systems'' inspired
+abstractions employed by Ceph (\S\ref{background}) and allude to their
+re-usability as components in other services. The next section
+enumerates the abstractions that Malacology exposes in the
+implementation section (\S\ref{implementation}). We then show how we use
+Malacology to synthesize entirely new storage services on an existing
+system through configuration and small changes (\S\ref{services}). We
+conclude by evaluating real-world use cases (\S\ref{evaluation}).
 
@@
 
+% IOStack: Software-Defined Object Storage
+% sRoute: Treating the Storage Stack Like a Network
+   % - lots of good refs to check out
+% Information and control in gray-box systems
+
 Malacology uses the same Active and Typed Storage module presented in

### 2016-05-10 paper/paper.tex
Beginnings of joining "Components of Ceph..." with "Malacology Implementation"

@@
 
-\section{Components of Ceph Suitable for Reusability}\label{re-usable-components-in-ceph}
-
-\label{background}
-
-Ceph is a production-quality distributed storage system and in this section we
-touch on some of the subsystems it uses to provide a general-purpose
-storage system. In the next section, we describe how Malacology exposes these
-subsystems in order to make them reusable and enable the creation of new services on top of them.
+\section{Malacology}
+\label{sec:malacology}
+
+Malacology is both a prototype for a programmable storage system and a design approach to evolve storage systems efficiently and without
+jeopardizing years of code-hardening and performance optimization
+efforts.
+The guiding principle is to re-use existing services and extend them so that these services can be \emph{programmed}.
+We accomplish programmability of a service by exporting bindings (or ``hooks'') for an interpreted programming language so that programming can occur without having to restart the storage system (see also below,~\S\ref{sec:durability}).
+
+There are multiple reasons that make Lua an attractive runtime for implementing 
+Malacology. Lua is a portable, embedded
+scripting language that offers superious performance and productivity
+trade-offs, including a JIT-based implementation that is well known for near
+native performance. Additionally, Lua has been used extensively in game engines, 
+and systems research \cite{neto:dls14-luaos}, including storage systems where it 
+has been effectively used both on
+\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010} and 
+off
+\cite{sevilla:sc15-mantle} the performance critical path. Finally, the 
+flexibility of the runtime allows to sandbox applications in order to address 
+security and performance concerns. We now describe how Malacology exposes each 
+subcomponent in Ceph.
+
+We will now discuss the subsystems Ceph uses to provide a general-purpose
+storage system and how Malacology makes them programmable.
+
+\subsection{Durability}
+\label{sec:durability}
+
+Ceph provides storage by striping and replicating data across
+RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
+uses many techniques to ensure that data is not corrupted or lost, such as
+erasure coding, replication, and data scrubbing. Furthermore, many of these
+techniques try to be autonomous so that work is distributed across the
+cluster. For example, when placement groups change, the OSDs rebalance and
+re-shard data in the background in a process called placement group splitting.
+
+In order to reduce load on the monitoring service, Ceph OSDs use a gossip
+protocol to efficiently propagate changes to cluster maps throughout the
+system, and automonmously initiate recovery mechansims when failures are
+discovered.
+
+\paragraph*{Malacology:} The implementation of Ceph's object abstraction, although powerful, has some 
+drawbacks.
+Supporting only C/C++ for object interface developers, Ceph requires
+distribution of compiled binaries for the correct architecture, adding a
+large barrier of entry for developers and system administrators. Second,
+having no way to dynamically unload modules, any changes require a full
+restart of an OSD which may have serious performance impacts due to a loss
+of cached data. And finally, the security limitations of the framework
+limit the use of object classes to all but those with administrative level
+access and technique expertise.
+
+By embedding the Lua runtime, Malacology makes the base RADOS API available in a 
+dynamic environment, resulting in a object storage API with economy of 
+expression, which at the same time provides the full set of features of the base 
+object class. New object interfaces that are expressed in thousands of lines of 
+code can be implemented in approximately an order of magnitude less 
+code~\cite{geambasu_comet_2010}.
 
@@
 
-\subsection{Durability}
-\label{sec:durability}
-
-Ceph provides storage by striping and replicating data across
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-uses many techniques to ensure that data is not corrupted or lost, such as
-erasure coding, replication, and data scrubbing. Furthermore, many of these
-techniques try to be autonomous so that work is distributed across the
-cluster. For example, when placement groups change, the OSDs rebalance and
-re-shard data in the background in a process called placement group splitting.
-
-In order to reduce load on the monitoring service, Ceph OSDs use a gossip
-protocol to efficiently propogate changes to cluster maps throughout the
-system, and automonmously initiate recovery mechansims when failures are
-discovered.
+\paragraph*{Malacology:} Map modification for propogation of state. We allow applications to augment
+these data structures with application-specific metdata. In order to keep the
+size of these data structures small application state may be stored elsewhere
+and the references stored within the data structure.
 
@@
 
+\paragraph*{Malacology:}
+
 \subsection{File System Indexing and Locking}
@@
 system abstraction. Distributed file systems are complex, but fundamentally
-provide durable hierarhical indexing and distributed locking service.  The
+provide durable hierarchical indexing and distributed locking service.  The
 locking service implements a capabilities that express what data and metadata

### 2016-05-10 paper/paper.tex
Added two more "\paragraph*{Malacology:}"
paper: added new Mantle service details... so tired...
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-05-10 paper/paper.tex
@@
 %    our experiment about propogtation delay
-%
-
-{\bf Balancing Policy.}
-
-Avoiding hard-coded balancing policies is important. First, it is difficult,
-nay impossible, to select a one-size-fits-all policy. System administrators
-may want to partition load based on SLAs, etc... Applications are often in
-a better position to know what to do.
-
-{\bf Capabilities.}
-The metadata service tracks the state of clients, including the data they
-may have cached and any locks that they hold. The metadata service is
-responsible responding to client requests for capabilities by first
+% Original: inject strings
+% New: use C++ bindings
+
+Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.
+
+\subsubsection{Balancing Policy}
+
+For CephFS, designing a general POSIX metadata load balancer is difficult becuase the workloads are diverse and quickly changing.
+With CephFS, it is especially complicated because the system has variable sized migration units (i.e.
+file system namespace subtrees), which can be shipped anywhere (i.e., to any MDS of any capacity) at any time for any reason.
+Furthermore, the original balancer was designed with hard-coded policies and tunables.
+Both angles were difficult to understand.
+It is understandable why it was implemented this way -- it is difficult to pick a one-size-fits-all policy, system administrators may want to partition load based on SLAs, or applications are better equipped to make these decisions, etc.
+-- but we choose instead to expose a balancing API at strategic points in the balancing logic using Lua hooks.
+
+These hooks embed Lua code in the MDS using the
+OSD subsystems.
+By re-using the OSD interface classes, the MDS gets the safety and robustness of loading dynamic code (including graceful failures and helpful errors), the ease of transferring state between object
+interfaces and the OSD internals, integration with testing and
+correctness suites, and \texttt{struct}s for interface data and
+handlers.
+
+To facilitate the use of Lua in both the MDS and OSD, we modularize the LuaVM in a core sandbox wrapper and link against it, as shown in Figure~\ref{fig:cls-osd-mds}. This core Lua wrapper contains
+just the dependencies, symbols, and functions, needed to run the Lua VM and strips away all the functionality inherent to object classes (like placement group filters, cryptography functions, object metadata,  object data, object extended attributes, and object versions).
+Some functions of Ceph's more helpful subsystems, like the logging facilities for the daemons, were re-used in the MDS.
+With this scheme, The OSD will dynamically open the shared library created by Lua object interface shared library while the MDS will dynamically open the Lua balancer interface shared
+library; both shared libraries will the Lua core.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/cls-osd-mds.png}
+\caption{To generalize the LuaVM for use in both the OSD and MDS,
+Malacology detaches the core functionality into a Lua core module. The
+VM is dynamically loaded first (1) and then interfaces can execute (2).
+\label{fig:cls-osd-mds}}
+\end{figure}
+
+Both the object and balancer interfaces define functions and attach them
+to the core Lua sandbox. For example, the Lua object storage interface
+class attaches data, extended attribute, object map, and object version
+functions, to the Lua core functions, as shown in Figure
+\ref{fig:cls-osd-mds}. The advantages of this is that we avoid
+duplicating code, we provide a framework for putting Lua code in other
+parts of the system, and we remove components and APIs that are too
+integrated with the OSD.
+
+\subsubsection{Capabilities}
+
+The metadata service also manages sessions with its clients. This gives both subsystems (MDSs and clients) the ability cache data and grab locks. The metadata service is
+responsible for responding to client requests for capabilities by first
 revoking caps from other clients. Currently clients voluntarily
 release their capabilities, and the MDS maintains a queue of requests.
-Both clients and MDSs use a best-effort approach to handling capabilities.
-
-In our current we retain the best-effort, volutary release of caps within
+Both clients and MDSs use a best-effort approach to revoking or releasing capabilities.
+
+In our current we retain the best-effort, voluntary release of caps within
 clients, but generalize the queueing policies within the MDS to allow for
@@
 
-\noindent \texttt{ceph\ mds\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}
-\noindent \texttt{ceph\ osd\ pool\ set-class\ \textless{}pool\textgreater{}\ \textless{}class\textgreater{}\ \textless{}script\textgreater{}}
-
 Metadata service policies and object storage interfaces are stored durability
 within RADOS and managed by storing references with the object maps. Since
-the cluster already propogates a consistent view of these data structures,
+the cluster already propagates a consistent view of these data structures,
 we use this service to automatically install interfaces in OSDs, and install
@@
 
-Malacogy adds interface hooks to the MDS in the load balancing and
-capability subsystems. In contrast to the Mantle implementation, which
-had a large function hard-coded into the balancer subsystem, Malacology
-embeds code in the MDS using OSD subsystems.
-
-Ceph has a whole
-infrastructure for getting dynamic code into the OSD. The
-\texttt{ClassHandler} class safely opens object interfaces, executes
-commands defined by the shared library, and fails gracefully with
-helpful errors if anything goes wrong. Instead of injecting strings
-directly into Ceph (like we did with the original Mantle
-implementation), the \texttt{ClassHandler} class safely opens shared
-code, even code with dependencies that are not in Ceph itself (e.g.,
-Lua).
-
-With the \texttt{ClassHandler}, the MDS gets the safety and robustness
-of loading dynamic code, the ease of transferring state between object
-interfaces and the OSD internals, integration with testing and
-correctness suites, and \texttt{struct}s for interface data and
-handlers. Using the OSD class handler hooks also gives the MDS the
-ability to load interfaces from the local file system.
-
-We generalize the Lua VM since it will be used in both the OSD and MDS.
-We put the core sandbox wrapper for the Lua object interface in a common
-directory in Ceph and link against it. This core Lua wrapper contains
-just the dependencies, symbols, and functions, need to run the Lua VM.
-Some of these functions include \texttt{clslua\_log()} for transferring
-logs to the daemon logs and \texttt{clslua\_pcall()} for calling Lua
-functions.
-
-Dependencies, symbols, and functions specific to the object or balancer
-interface are put in the \texttt{cls} or \texttt{bal} directories,
-respectively. For example, the Lua object interface uses functions that
-have placement group filters, cryptography functions, and object
-metadata (\emph{it i.e.} \texttt{cxx\_omap\_getvals()}, object data,
-object extended attributes, and object versions -- all these are
-dependencies, symbols, and functions are part of the OSD process but not
-the MDS process. As a result, we put interface code that uses these in
-\texttt{cls} directory. With this scheme, The OSD will \texttt{dlopen()}
-the shared library created by Lua object interface shared library while
-the MDS will \texttt{dlopen()} the Lua balancer interface shared
-library; both shared libraries will the Lua core.
 \fi
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/cls-osd-mds.png}
-\caption{To generalize the LuaVM for use in both the OSD and MDS,
-Malacology detaches the core functionality into a Lua core module. The
-VM is dynamically loaded first (1) and then interfaces can execute (2).
-\label{fig:cls-osd-mds}}
-\end{figure}
-
 \iffalse
-Both the object and balancer interfaces define functions and attach them
-to the core Lua sandbox. For example, the Lua object storage interface
-class attaches data, extended attribute, object map, and object version
-functions, to the Lua core functions, as shown in Figure
-\ref{fig:cls-osd-mds}. The advantages of this is that we avoid
-duplicating code, we provide a framework for putting Lua code in other
-parts of the system, and we remove components and APIs that are too
-integrated with the OSD.
 
@@
 
+\subsection{Re-Implementation}
+
 The Ceph team wants to merge Mantle because the scriptability is useful
@@
 
-Mantle is a framework that allows balancing policies to be written in Lua,
-and supports such policies be specified at run-time.
-We implemented it using the components from section 3. I will tell you about them here.
+The API for Malacology is the same as the one exposed in~\cite{sevilla:sc15-mantle} but the hooks are put in different places using the interface hooks described in Section \S\ref{}.
+
+\subsection{Policy Versioning}
+
+For Mantle, we added new monitor command to change the metadata load balancers:
+
+\noindent \texttt{ mds\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}
+%\noindent \texttt{ceph\ osd\ pool\ set-class\ \textless{}pool\textgreater{}\ \textless{}class\textgreater{}\ \textless{}script\textgreater{}}
+
+This command assigns the current version of the metadata load balancer.
+In our case, it is simply the name of the balancer.
+The MONs take the version and store it in the MDS map.
+Similar to the PG map or MON maps, the MDS map has all the information about the topology and state of the MDS cluster.
+Leveraging the MONs helps ease the burden on the MDS to propagate the policy and ensure that everyone agrees.
+
+There are Ceph components that are obstacles to this approach, namely the size of the  MDS map, API changes, and security.
+The Ceph engineers want to avoid using the maps to exchange information; the maps should be lightweight and fast because the MONs need to be able to version MDS cluster state quickly.
+API changes can also be a burden, especially for the MDS map, since all clients need to have the same version of the CephFS library in order to properly interact with CephFS.
+Finally, Ceph tries to avoid injecting strings into the cluster because they are insecure and buggy.
+
+\subsection{Policy Durability}
+
+-
+
+-
+
+-
 

### 2016-05-10 paper/paper.tex

@@
 common services that are useful to large sets of applications.
+
+\paragraph*{Malacology:}
 
@@
 It then uses a hard-coded policy to balance these fragments using a secularized load metric based on CPU, workload, and file system operation metrics~\cite{sevilla:sc15-mantle}.
+
+\paragraph*{Malacology:}
 

### 2016-05-10 paper/paper.tex
Changes order of object storage discussion

@@
 flexibility of the runtime allows to sandbox applications in order to address 
-security and performance concerns. We now describe how Malacology exposes each 
-subcomponent in Ceph.
+security and performance concerns.
 
@@
 
-\subsection{Durability}
-\label{sec:durability}
-
-Ceph provides storage by striping and replicating data across
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-uses many techniques to ensure that data is not corrupted or lost, such as
-erasure coding, replication, and data scrubbing. Furthermore, many of these
-techniques try to be autonomous so that work is distributed across the
-cluster. For example, when placement groups change, the OSDs rebalance and
-re-shard data in the background in a process called placement group splitting.
-
-In order to reduce load on the monitoring service, Ceph OSDs use a gossip
-protocol to efficiently propagate changes to cluster maps throughout the
-system, and automonmously initiate recovery mechansims when failures are
-discovered.
+\subsection{Logical Storage Devices}
+\label{active-storage}
+
+Allowing users to define custom code that executes in the storage nodes is a 
+useful technique for offloading computation closer to where the data is stored 
+in order to significantly reduce the amount of data moved between storage 
+servers and clients.
+Ceph provides object storage via 
+RADOS~\cite{weil_rados_2007}, the reliable distributed object store. 
+Ceph provides a rich plugin framework for defining new interfaces to objects
+stored within RADOS, effectively allowing developers to use co-design
+techniques to define logical storage devices tailored to their application
+domain.
+
+An object interface plugin is structured in a similar way to that of an RPC in
+which a developer creates a named function within the cluster that clients may
+invoke upon an object. In Ceph these functions are defined by a composition of
+existing native interfaces that include reading and writing bytestream data
+and controlling snapshots and cloning, all of which may be transactionally
+composed with full access to a sorted key-value service allowing applications
+to fine tune storage interfaces, such as using a read-modify-write guard to
+protect shared data.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/implementation.png}
+\caption{Malacology re-uses Ceph subsystems to manage interface classes.
+command. After the proposal is accepted by the cluster of monitors, the
+script is redundantly and consistently distributed to OSDs.
+\label{fig:implementation}}
+\end{figure}
 
@@
 code~\cite{geambasu_comet_2010}.
+
+\subsection{Durability}
+\label{sec:durability}
+
+Ceph provides storage by striping and replicating data across
+RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
+uses many techniques to ensure that data is not corrupted or lost, such as
+erasure coding, replication, and data scrubbing. Furthermore, many of these
+techniques try to be autonomous so that work is distributed across the
+cluster. For example, when placement groups change, the OSDs rebalance and
+re-shard data in the background in a process called placement group splitting.
+
+In order to reduce load on the monitoring service, Ceph OSDs use a gossip
+protocol to efficiently propagate changes to cluster maps throughout the
+system, and autonomously initiate recovery mechanisms when failures are
+discovered.
+
+\paragraph*{Malacology:}
 
@@
 \paragraph*{Malacology:} Map modification for propogation of state. We allow applications to augment
-these data structures with application-specific metdata. In order to keep the
+these data structures with application-specific metadata. In order to keep the
 size of these data structures small application state may be stored elsewhere
 and the references stored within the data structure.
-
-\subsection{Logical Storage Devices}
-\label{active-storage}
-
-Allowing users to define custom code that executes in the storage nodes is a 
-useful technique for offloading computation closer to where the data is stored 
-in order to significantly reduce the amount of data moved between storage 
-servers and clients. Ceph provides a rich plugin framework for defining new 
-interfaces to objects
-stored within RADOS, effectively allowing developers to use co-design
-techniques to define logical storage devices tailored to their application
-domain.
-
-An object interface plugin is structured in a similar way to that of an RPC in
-which a developer creates a named function within the cluster that clients may
-invoke upon an object. In Ceph these functions are defined by a composition of
-existing native interfaces that include reading and writing bytestream data
-and controlling snapshots and cloning, all of which may be transactionally
-composed with full access to a sorted key-value service allowing applications
-to fine tune storage interfaces, such as using a read-modify-write guard to
-protect shared data.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/implementation.png}
-\caption{Malacology re-uses Ceph subsystems to manage interface classes.
-command. After the proposal is accepted by the cluster of monitors, the
-script is redundantly and consistently distributed to OSDs.
-\label{fig:implementation}}
-\end{figure}
-
-\paragraph*{Malacology:}
 

### 2016-05-10 paper/paper.tex
Merge section 3 into 2
Also removes commented out text to avoid confusion

@@
 % carefully exposing internal storage services to applications.
-
-\if{false}
-For example, a Ceph cluster contains a small number of
-monitoring nodes (MONs) that use PAXOS to maintain a consistent view of
-system-wide metadata such as data distribution parameters and cluster
-membership. The data distribution metadata maintained by
-the cluster monitors is used by RADOS servers and clients to provide
-consistent reads and writes, and both RADOS and the monitoring services
-are used by a cluster of file system metadata servers that provide POSIX
-semantics using sophisticated indexing and distributed locking services.
-We contend that re-using and re-purposing these code-hardened subsystems
-is paramount to successfully adapting storage systems to new APIs and
-new storage devices without losing the benefits from years of
-code-hardening work.
-\fi
 
@@
 
+{\bf Recovery.}
+Application specific recovery is critical.
+
 \subsection{Durability}
@@
 
-\paragraph*{Malacology:}
+\paragraph*{Malacology:} Metadata service policies and object storage interfaces are stored durability
+within RADOS and managed by storing references with the object maps. Since
+the cluster already propagates a consistent view of these data structures,
+we use this service to automatically install interfaces in OSDs, and install
+policies within the MDS such that clients and daemons are synchronzied on
+correct implementations without restarting.
 
@@
 
-\paragraph*{Malacology:} Map modification for propogation of state. We allow applications to augment
+\paragraph*{Malacology:} Malacology exposes the monitoring service in two ways. 
+First, Malacology APIs for object storage (\S\ref{active-storage}) and the MDS 
+(\S\ref{malacology:mds}) have functions available to them that allow them to 
+store application-specific data in the highly-available, consistent store 
+associated to the Ceph monitors. As mentioned
+Map modification for propogation of state. We allow applications to augment
 these data structures with application-specific metadata. In order to keep the
@@
 
-\paragraph*{Malacology:}
+\paragraph*{Malacology:} Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.
+Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.
+
+These hooks embed Lua code in the MDS using the
+OSD subsystems.
+By re-using the OSD interface classes, the MDS gets the safety and robustness of loading dynamic code (including graceful failures and helpful errors), the ease of transferring state between object
+interfaces and the OSD internals, integration with testing and
+correctness suites, and \texttt{struct}s for interface data and
+handlers.
+
+The metadata service also manages sessions with its clients. This gives both subsystems (MDSs and clients) the ability cache data and grab locks. The metadata service is
+responsible for responding to client requests for capabilities by first
+revoking caps from other clients. Currently clients voluntarily
+release their capabilities, and the MDS maintains a queue of requests.
+Both clients and MDSs use a best-effort approach to revoking or releasing capabilities.
+
+In our current we retain the best-effort, voluntary release of caps within
+clients, but generalize the queueing policies within the MDS to allow for
+centeralized policies such as fairness or priority.
+
+{\bf File Types.}
+We can reference PDSW12 DataMods, and talk about any new stuff here that we
+want to.
 
 \subsection{Metadata Load Balancing}
+\label{malacology:mds}
 
@@
 
-\section{Malacology Implementation}
-\label{implementation}
-
-In this section we describe how Malacology builds upon existing
-services within Ceph. The main idea behind Malacology is to take the subsystems 
-described in the previous section and expose APIs on top of them that allow 
-users to dynamically inject code, allowing them to customize the behavior of 
-these components. Supporting dynamic reconfiguration and programmability 
-efficiently is critical in real systems that must adapt to a changing 
-environment where potentially
-conflicting policies arise. We've adopted Lua as the dynamic runtime  to
-implement dynamic components in Malacology.
-
-There are multiple reasons that make Lua an attractive runtime for implementing 
-Malacology. Lua is a portable, embedded
-scripting language that offers superious performance and productivity
-trade-offs, including a JIT-based implementation that is well known for near
-native performance. Additionally, Lua has been used extensively in game engines, 
-and systems research \cite{neto:dls14-luaos}, including storage systems where it 
-has been effectively used both on
-\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010} and 
-off
-\cite{sevilla:sc15-mantle} the performance critical path. Finally, the 
-flexibility of the runtime allows to sandbox applications in order to address 
-security and performance concerns. We now describe how Malacology exposes each 
-subcomponent in Ceph.
-
-\subsection{Object Storage}
-
-% todo:
-%  - ellude to the need for interfaces to be avialable quickly to support
-%    our experiment about propogtation delay
-%
-
-The implementation of Ceph's object abstraction, although powerful, has some 
-drawbacks.
-Supporting only C/C++ for object interface developers, Ceph requires
-distribution of compiled binaries for the correct architecture, adding a
-large barrier of entry for developers and system administrators. Second,
-having no way to dynamically unload modules, any changes require a full
-restart of an OSD which may have serious performance impacts due to a loss
-of cached data. And finally, the security limitations of the framework
-limit the use of object classes to all but those with administrative level
-access and technique expertise.
-
-By embedding the Lua runtime, Malacology makes the base RADOS API available in a 
-dynamic environment, resulting in a object storage API with economy of 
-expression, which at the same time provides the full set of features of the base 
-object class. New object interfaces that are expressed in thousands of lines of 
-code can be implemented in approximately an order of magnitude less 
-code~\cite{geambasu_comet_2010}.
-
-\subsection{Metadata Services}
-
-% todo:
-%  - ellude to the need for policies to be avialable quickly to support
-%    our experiment about propogtation delay
-% Original: inject strings
-% New: use C++ bindings
-
-Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.
-
-\subsubsection{Balancing Policy}
-
 For CephFS, designing a general POSIX metadata load balancer is difficult becuase the workloads are diverse and quickly changing.
@@
 -- but we choose instead to expose a balancing API at strategic points in the balancing logic using Lua hooks.
-
-These hooks embed Lua code in the MDS using the
-OSD subsystems.
-By re-using the OSD interface classes, the MDS gets the safety and robustness of loading dynamic code (including graceful failures and helpful errors), the ease of transferring state between object
-interfaces and the OSD internals, integration with testing and
-correctness suites, and \texttt{struct}s for interface data and
-handlers.
-
-To facilitate the use of Lua in both the MDS and OSD, we modularize the LuaVM in a core sandbox wrapper and link against it, as shown in Figure~\ref{fig:cls-osd-mds}. This core Lua wrapper contains
-just the dependencies, symbols, and functions, needed to run the Lua VM and strips away all the functionality inherent to object classes (like placement group filters, cryptography functions, object metadata,  object data, object extended attributes, and object versions).
-Some functions of Ceph's more helpful subsystems, like the logging facilities for the daemons, were re-used in the MDS.
-With this scheme, The OSD will dynamically open the shared library created by Lua object interface shared library while the MDS will dynamically open the Lua balancer interface shared
-library; both shared libraries will the Lua core.
 
@@
 
-\subsubsection{Capabilities}
-
-The metadata service also manages sessions with its clients. This gives both subsystems (MDSs and clients) the ability cache data and grab locks. The metadata service is
-responsible for responding to client requests for capabilities by first
-revoking caps from other clients. Currently clients voluntarily
-release their capabilities, and the MDS maintains a queue of requests.
-Both clients and MDSs use a best-effort approach to revoking or releasing capabilities.
-
-In our current we retain the best-effort, voluntary release of caps within
-clients, but generalize the queueing policies within the MDS to allow for
-centeralized policies such as fairness or priority.
-
-{\bf File Types.}
-We can reference PDSW12 DataMods, and talk about any new stuff here that we
-want to.
-
-\subsection{Cluster State}
-
-Map modification for propogation of state. We allow applications to augment
-these data structures with application-specific metdata. In order to keep the
-size of these data structures small application state may be stored elsewhere
-and the references stored within the data structure.
-
-\subsection{Programmability and Durability}
-
-{\bf Recovery.}
-Application specific recovery is critical.
-
-{\bf Durability.}
-
-Metadata service policies and object storage interfaces are stored durability
-within RADOS and managed by storing references with the object maps. Since
-the cluster already propagates a consistent view of these data structures,
-we use this service to automatically install interfaces in OSDs, and install
-policies within the MDS such that clients and daemons are synchronzied on
-correct implementations without restarting.
-
-\iffalse
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  additional interface hooks and classes for OSDs and MDSs using the
-  existing Ceph \texttt{cls} infrastructure
-\item
-  durable interface classes using the Ceph object store
-\item
-  versioned and consistent interfaces using the Ceph MONs
-\end{enumerate}
-
-Figure \ref{implementation} shows how interfaces can be loaded into
-either the MDS or OSD (1), how interfaces are persisted to object
-storage (2), and how they are versioned by the MONs (3). In the
-following section, we will discuss the existing Ceph infrastructure and
-the changes necessary for Malacology. We emphasize here that these
-implementations are not necessarily optimized but they help demonstrate
-the power of programmability.
-
-\subsection{Additional Interface
-Frameworks}\label{additional-interface-frameworks}
-
-For object interfaces, clients send Lua classes to the OSD and then they
-can invoke operations in that class remotely on the OSD. In reality,
-this feature is a statically loaded object class written in C/C++ that
-runs dynamically defined object interfaces, where clients access the
-classes using an \texttt{exec()} wrapper. Although sending the class
-with each client request is stateless (which has its own set of nice
-features), it is costly for network bandwidth and difficult to organize
-at the application level. Also, while the execute wrapper simplifies the
-implementation, it burdens the applications, since they need to be
-recompiled to use the \texttt{exec()} function.
-
-Although these OSD interfaces are powerful, the current implementation
-has drawbacks. OSD interfaces are written in C/C++ and compiled into a
-shared library, so the developer must account for different target
-architectures. Second, C/C++ provides more functionality than is
-necessary since the snippets of interface code mainly set policies and
-perform simple operations. Finally, developing these interfaces in C/C++
-has high overhead. The developer must learn how the OSD dynamically
-loads the shared libraries (\emph{e.g.}, OSDs rely on a strict naming
-convention to find shared libraries), how to get their interfaces
-compiled using the Ceph \texttt{make} system, and how to debug issues
-that are not related to their specific interface (\emph{e.g.}, the OSD
-cannot find the shared library) - this learning curve is unacceptable
-for non-Ceph developers, especially since most interfaces are one-off
-solutions specific to their applications.
-
-\subsubsection{New Interface Classes}\label{new-interface-classes}
-
-Our framework has added a mechanism for defining and running object and
-metadata balancer classes using the Lua scripting language. Scripting is
-useful in large systems for 3 reasons:
-
-\begin{itemize}
-\item
-  accelerates deployment: tweaking policies, like changing a threshold
-  or metric calculation, does not require a full recompile or system
-  start-up.
-\item
-  transparency: if the bindings carefully expose the correct metrics and
-  functions, the user does not need to learn the internal functions,
-  variables, or types of the system.
-\item
-  portability: scripts can be exchanged across platforms and system
-  versions without compatibility issues.
-\end{itemize}
-
-Our Lua bindings expose functions and symbols both ways; the host
-program can call functions defined in Lua and the Lua scripts can call
-functions defined in native C++. These bindings are merged upstream. Lua
-is frequently used in game engines to set policies but we use it here
-because most of the user-defined classes in Ceph are policies as well!
-We do not want to provide specific implementations, like pulling data
-from objects or transferring them over the network, but instead strive
-to say \emph{what to do} with the data once we have it. Separating
-policy from mechanism is a driving factor in using Lua. We choose Lua
-for 4 reasons: performance, portabability, size, and security.
-
-Lua performs well as an embeddable scripting language. Although it is an
-interpreted language, the bindings are designed to make the exchange of
-variables and functions amongst languages straightforward. The LuaJIT
-virtual machine boasts near-native performance, making it a logical
-choice for scriptability frameworks in systems research
-\cite{neto:dls14-luaos}. In storage systems, it has been effectively
-used both on \cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo} and off
-\cite{sevilla:sc15-mantle} the critical path, where performance is
-important.
-
-Because Lua is interpreted, it is also portable. The portability allows
-functions to be shipped around the cluster without having to compile or
-reason about compatibility issues. Traditionally, intefaces are written
-in C/C++, so they must be compiled into shared libraries before being
-injected into the daemon at runtime. Shared libraries need to be
-compiled on the host they will be run on to accomodate different
-architectures and dependendencies. Embedding the Lua interpretator, on
-the other hand, gives us the flexbility to store the actual Lua code in
-other places, like RADOS, so we can enjoy all the other properties
-provided by that particular back-end.
-
-The other two advantages, that Lua leaves a small memory footprint and
-that it is secure, are discussed at length in 
-\cite{ierusalimschy_programming_2006,neto:dls14-luaos}. In short, Lua
-provides relatively robust sandbox capabilities for protecting the
-system against some malicious cases, but does not protect the system
-from poor policies. Since Malacology does not use these features (yet),
-we will avoid discussion here.
-
-\subsection{New Interface Hooks}\label{new-interface-hooks}
-
-\fi
-
-\iffalse
-
-\subsection{Programmability and Durability}
-
-Traditionally, object storage interfaces are compiled into shared
-libraries and loaded into a running OSD daemon using \texttt{dlopen()}.
-Because the shared library has to link against symbols found in the
-running executable, the interfaces are stored on the local file systems
-of each OSD so they can be versioned and distributed alongside the same
-Ceph binaries. This policy is for safety, since the OSDs could be on
-different distros that provide different versioning capabilities. Also,
-this approach allows bug fixes to co-evolve with the rest of the Ceph
-installation.
-
-Because of the portability of Lua and because the Lua virtual machine is
-compiled and loaded as a shared library, the MDS/OSD can store the
-balancers in a more redundant location. For our implementation, we store
-balancers in RADOS and the daemons first search RADOS, then the local
-file system for their interface classes.
-
-\subsection{Versioning Interfaces with
-MONs}\label{versioning-interfaces-with-mons}
-
-We added two new commands to Ceph command line client so the user can
-inject interface versions into the cluster::
-
-\noindent \texttt{ceph\ osd\ pool\ set-class\ \textless{}pool\textgreater{}\ \textless{}class\textgreater{}\ \textless{}script\textgreater{}}
-
-\noindent \texttt{ceph\ mds\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}
-
-By leveraging the monitor daemons, our services get the consistency from
-the monitors version management, the distribution from the monitors data
-structures (which are already distributed), and the durability from the
-robustness of the monitors and persistence with Paxos. The
-implementation uses the placement group pool data structure which
-maintains the policies (\emph{e.g.}, erasure coding) and organization
-details (\emph{e.g.}, snapshots) for each pool. While stuffing the
-information into the monitor map, the structure that describes the the
-monitor topology, was an option, the placement group pool allows finer
-grain control over different types of data.
-
-In regards to the consistency, each time the user injects a new Lua
-object class it enters the monitor quorum as a proposal; updates to the
-placement group pool need to wait until accepted, at which point the
-state will be propogated and versioned. We hooked up the monitor command
-to the \texttt{cls\_lua} class by having the OSD pull the script from
-the placement group instead of trying to dynamically load the shared
-library with \texttt{dlopen()} or the Lua script.
-
-The cls-lua branch lets users write object classes in Lua, the lua-rados
-client library lets users write applications that talk to RADOS with
-Lua, and the cls-client uses the lua-rados library to send Lua object
-classes to the OSDs equipped with the LuaJIT VM.
-\fi
+To facilitate the use of Lua in both the MDS and OSD, we modularize the LuaVM in a core sandbox wrapper and link against it, as shown in Figure~\ref{fig:cls-osd-mds}. This core Lua wrapper contains
+just the dependencies, symbols, and functions, needed to run the Lua VM and strips away all the functionality inherent to object classes (like placement group filters, cryptography functions, object metadata,  object data, object extended attributes, and object versions).
+Some functions of Ceph's more helpful subsystems, like the logging facilities for the daemons, were re-used in the MDS.
+With this scheme, The OSD will dynamically open the shared library created by Lua object interface shared library while the MDS will dynamically open the Lua balancer interface shared
+library; both shared libraries will the Lua core.
 

### 2016-05-10 paper/paper.tex
new outline for sec 2
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>
Update paper.tex
line 406: "In our current we"
line 724: update until each OSD makes the update live (the cost of the Paxos proposal is

@@
 By embedding the Lua runtime, Malacology makes the base RADOS API available in a 
-dynamic environment, resulting in a object storage API with economy of 
+dynamic environment, resulting in an object storage API with economy of 
 expression, which at the same time provides the full set of features of the base 
@@
 
-\paragraph*{Malacology:} Metadata service policies and object storage interfaces are stored durability
+\paragraph*{Malacology:} Metadata service policies and object storage interfaces are stored durably
 within RADOS and managed by storing references with the object maps. Since
@@
 {\bf Capability policies.}
-We use capabilitity policies in the metadata service to tune the amount of
+We use capability policies in the metadata service to tune the amount of
 time that clients may cache the sequencer resource. This allows us to control
@@
 system client, throughput drops significantly with two clients as the majority
-of the time neither client has the capapbitility as it ping-pongs between
+of the time neither client has the capability as it ping-pongs between
 clients. This observation reveals a tuning parameter: how long a client
-maintains its capapbility lease. In throughputs labeled $2C,<time>$ we show
+maintains its capability lease. In throughputs labeled $2C,<time>$ we show
 aggregate throughput as we decrease the amount of time that a client holds its
-lease before responding to a request for the capapbility.  At lease times as
+lease before responding to a request for the capability.  At lease times as
 low as 10ms throughputs achieve a significant proportion of the locally cached
@@
 for applications that require predictable latency. The throughput in
-Figure~\ref{fig:captp} labeled $2C,ZLog$ corresponds to a centeralized
+Figure~\ref{fig:captp} labeled $2C,ZLog$ corresponds to a centralized
 sequencer implementation. While throughput is lower for two clients (because of
@@
 system deficiencies. However, this duplication has real-world problems
-related to reliability. We propose that system expose their services in
-a safe way allowing application developers to customize system behavior
+related to reliability. We propose that systems expose their services in
+a safe way, allowing application developers to customize system behavior
 to meet their needs while not sacrificing correctness.
@@
 services to be configured on-the-fly in existing systems. This work is
-one point along that path in which we have looked an a target
+one point along that path in which we have looked at a target
 special-purpose storage system. Ultimately we want to utilize

### 2016-05-10 paper/paper.tex
sec2.2
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-05-10 paper/paper.tex
@@
 \label{introduction}
+\label{sec:intro}
 
@@
 
-\subsection{Logical Storage Devices}
-\label{active-storage}
-
-Allowing users to define custom code that executes in the storage nodes is a 
-useful technique for offloading computation closer to where the data is stored 
-in order to significantly reduce the amount of data moved between storage 
-servers and clients.
-Ceph provides object storage via 
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. 
-Ceph provides a rich plugin framework for defining new interfaces to objects
-stored within RADOS, effectively allowing developers to use co-design
-techniques to define logical storage devices tailored to their application
-domain.
-
-An object interface plugin is structured in a similar way to that of an RPC in
-which a developer creates a named function within the cluster that clients may
-invoke upon an object. In Ceph these functions are defined by a composition of
-existing native interfaces that include reading and writing bytestream data
-and controlling snapshots and cloning, all of which may be transactionally
-composed with full access to a sorted key-value service allowing applications
-to fine tune storage interfaces, such as using a read-modify-write guard to
-protect shared data.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/implementation.png}
-\caption{Malacology re-uses Ceph subsystems to manage interface classes.
-command. After the proposal is accepted by the cluster of monitors, the
-script is redundantly and consistently distributed to OSDs.
-\label{fig:implementation}}
-\end{figure}
-
-\paragraph*{Malacology:}  The implementation of Ceph's object abstraction, although powerful, has some 
-drawbacks.
-Supporting only C/C++ for object interface developers, Ceph requires
-distribution of compiled binaries for the correct architecture, adding a
-large barrier of entry for developers and system administrators. Second,
-having no way to dynamically unload modules, any changes require a full
-restart of an OSD which may have serious performance impacts due to a loss
-of cached data. And finally, the security limitations of the framework
-limit the use of object classes to all but those with administrative level
-access and technique expertise.
-
-By embedding the Lua runtime, Malacology makes the base RADOS API available in a 
-dynamic environment, resulting in a object storage API with economy of 
-expression, which at the same time provides the full set of features of the base 
-object class. New object interfaces that are expressed in thousands of lines of 
-code can be implemented in approximately an order of magnitude less 
-code~\cite{geambasu_comet_2010}.
-
-{\bf Recovery.}
-Application specific recovery is critical.
-
-\subsection{Durability}
-\label{sec:durability}
-
-Ceph provides storage by striping and replicating data across
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-uses many techniques to ensure that data is not corrupted or lost, such as
-erasure coding, replication, and data scrubbing. Furthermore, many of these
-techniques try to be autonomous so that work is distributed across the
-cluster. For example, when placement groups change, the OSDs rebalance and
-re-shard data in the background in a process called placement group splitting.
-
-In order to reduce load on the monitoring service, Ceph OSDs use a gossip
-protocol to efficiently propagate changes to cluster maps throughout the
-system, and autonomously initiate recovery mechanisms when failures are
-discovered.
-
-\paragraph*{Malacology:} Metadata service policies and object storage interfaces are stored durability
-within RADOS and managed by storing references with the object maps. Since
-the cluster already propagates a consistent view of these data structures,
-we use this service to automatically install interfaces in OSDs, and install
-policies within the MDS such that clients and daemons are synchronzied on
-correct implementations without restarting.
-
 \subsection{Cluster State Management}
@@
 
-\subsection{File System Indexing and Locking}
+\subsection{Logical Storage Devices}
+\label{active-storage}
 
-The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction. Distributed file systems are complex, but fundamentally
-provide durable hierarchical indexing and distributed locking service.  The
-locking service implements a capabilities that express what data and metadata
-clients are allowed to access as well as what they may cache locally.  While
-designed for a fixed file abstraction, indexing, locking, and caching are all
-common services that are useful to large sets of applications.
+Briefly described in Section~\section{sec:intro}, Ceph supports the creation
+of application-specific object interfaces~\cite{weil_rados_2007}. The ability
+to offload computation can reduce data movement, and transactional interfaces can
+significantly simplify construction of complex storage interfaces that require
+uncoordinated parallel access.
 
-\paragraph*{Malacology:} Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.
-Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.
+An object interface is a plugin structured in a similar way to that of an RPC
+in which a developer creates a named function within the cluster that clients
+may invoke. In the case of Ceph each function is implicitly run within the
+context of an object specified when the function is invoked. Developers of
+object interfaces express behavior by creating a composition of native
+interfaces or other custom object interfaces, and handle serialization of
+function input and output. A wide range of native interfaces are available to
+developers, such as reading and writing to a bytestream, controlling snapshots
+and clones, all of which may be transactionally composed with access to a
+sorted key-value service allowing applications to create rich interfaces, such
+as using a read-modify-write guard to protect a shared index defined over
+content store in a byte stream.
 
-These hooks embed Lua code in the MDS using the
-OSD subsystems.
-By re-using the OSD interface classes, the MDS gets the safety and robustness of loading dynamic code (including graceful failures and helpful errors), the ease of transferring state between object
-interfaces and the OSD internals, integration with testing and
-correctness suites, and \texttt{struct}s for interface data and
-handlers.
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/implementation.png}
+\caption{Malacology re-uses Ceph subsystems to manage interface classes.
+command. After the proposal is accepted by the cluster of monitors, the
+script is redundantly and consistently distributed to OSDs.
+\label{fig:implementation}}
+\end{figure}
 
-The metadata service also manages sessions with its clients. This gives both subsystems (MDSs and clients) the ability cache data and grab locks. The metadata service is
-responsible for responding to client requests for capabilities by first
-revoking caps from other clients. Currently clients voluntarily
-release their capabilities, and the MDS maintains a queue of requests.
-Both clients and MDSs use a best-effort approach to revoking or releasing capabilities.
+\paragraph*{Malacology:} The implementation of Ceph's object abstraction,
+although powerful, has some drawbacks. Supporting only C/C++ for object
+interface developers, Ceph requires distribution of compiled binaries for the
+correct architecture, adding a large barrier of entry for developers and
+system administrators. Second, having no way to dynamically unload modules,
+any changes require a full restart of an OSD which may have serious
+performance impacts due to a loss of cached data~\cite{facebook-mysql}. And
+finally, the security limitations of the framework limit the use of object
+classes to all but those with administrative level access and deep technical
+expertise.
 
-In our current we retain the best-effort, voluntary release of caps within
-clients, but generalize the queueing policies within the MDS to allow for
-centeralized policies such as fairness or priority.
+To address these concerns, Malacology takes advantage of community contributed
+Lua extensions to Ceph allowing new object interfaces to be dynamically loaded
+into the system and modified at runtime, resulting in a object storage API
+with economy of expression, which at the same time provides the full set of
+features of the base object class~\cite{bdmc}. New object interfaces that are
+expressed in thousands of lines of code can be implemented in approximately an
+order of magnitude less code~\cite{geambasu_comet_2010}.
 
-{\bf File Types.}
-We can reference PDSW12 DataMods, and talk about any new stuff here that we
-want to.
+All of these modifications work in tandem to implement the desired behavior,
+and are typically co-designed together such that each depends on the other to
+behave as expected. In highly dynamic environments such as a distributed
+storage system, changes to the system propagate at variable rates, requiring
+all components of the system to be able to adapt. Thus, system services
+realized through extensibility should be versioned together as a group,
+exposing the version to allow introspection and domain-specific treatment of
+version mismatch scenarios.
 
-\subsection{Metadata Load Balancing}
+And finally, it is important to not underestimate the importance of the
+preservation of the modifications that enable a new system service. In many
+(if not all) cases, interfaces defining access to data are just as important
+as the data itself by virtue of inherently providing structural context. Thus,
+all components of a system service must be kept to the same standards of
+protection as data in the system itself.
+
+To this end we have extended the monitoring service using the framework
+described in Section~\ref{sec:mon} to provide an interface for registration
+and verification of dynamically defined object interfaces. Using this service
+guarantees that interface definitions are not only made durable, but are
+transparently and consistently propogated throughput the cluster and that
+clients are properly synchronousized with the latest interfaces.
+
+{\bf Recovery.} Application specific recovery is critical.
+
+\subsection{Distributed Metadata Services}
 \label{malacology:mds}
@@
 library; both shared libraries will the Lua core.
+
+\subsection{File System Indexing and Locking}
+
+The distributed metadata service in Ceph provides clients with a POSIX file
+system abstraction. Distributed file systems are complex, but fundamentally
+provide durable hierarchical indexing and distributed locking service.  The
+locking service implements a capabilities that express what data and metadata
+clients are allowed to access as well as what they may cache locally.  While
+designed for a fixed file abstraction, indexing, locking, and caching are all
+common services that are useful to large sets of applications.
+
+\paragraph*{Malacology:} Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.
+Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.
+
+These hooks embed Lua code in the MDS using the
+OSD subsystems.
+By re-using the OSD interface classes, the MDS gets the safety and robustness of loading dynamic code (including graceful failures and helpful errors), the ease of transferring state between object
+interfaces and the OSD internals, integration with testing and
+correctness suites, and \texttt{struct}s for interface data and
+handlers.
+
+The metadata service also manages sessions with its clients. This gives both subsystems (MDSs and clients) the ability cache data and grab locks. The metadata service is
+responsible for responding to client requests for capabilities by first
+revoking caps from other clients. Currently clients voluntarily
+release their capabilities, and the MDS maintains a queue of requests.
+Both clients and MDSs use a best-effort approach to revoking or releasing capabilities.
+
+In our current we retain the best-effort, voluntary release of caps within
+clients, but generalize the queueing policies within the MDS to allow for
+centeralized policies such as fairness or priority.
+
+{\bf File Types.}
+We can reference PDSW12 DataMods, and talk about any new stuff here that we
+want to.
+
+\subsection{Durability}
+\label{sec:durability}
+
+Ceph provides storage by striping and replicating data across
+RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
+uses many techniques to ensure that data is not corrupted or lost, such as
+erasure coding, replication, and data scrubbing. Furthermore, many of these
+techniques try to be autonomous so that work is distributed across the
+cluster. For example, when placement groups change, the OSDs rebalance and
+re-shard data in the background in a process called placement group splitting.
+
+In order to reduce load on the monitoring service, Ceph OSDs use a gossip
+protocol to efficiently propagate changes to cluster maps throughout the
+system, and autonomously initiate recovery mechanisms when failures are
+discovered.
+
+\paragraph*{Malacology:} Metadata service policies and object storage interfaces are stored durability
+within RADOS and managed by storing references with the object maps. Since
+the cluster already propagates a consistent view of these data structures,
+we use this service to automatically install interfaces in OSDs, and install
+policies within the MDS such that clients and daemons are synchronzied on
+correct implementations without restarting.
 

### 2016-05-10 paper/paper.tex
typo
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>
clarify
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-Briefly described in Section~\section{sec:intro}, Ceph supports the creation
+Briefly described in Section~\ref{sec:intro}, Ceph supports the creation
 of application-specific object interfaces~\cite{weil_rados_2007}. The ability
@@
 
-All of these modifications work in tandem to implement the desired behavior,
-and are typically co-designed together such that each depends on the other to
-behave as expected. In highly dynamic environments such as a distributed
-storage system, changes to the system propagate at variable rates, requiring
-all components of the system to be able to adapt. Thus, system services
-realized through extensibility should be versioned together as a group,
-exposing the version to allow introspection and domain-specific treatment of
-version mismatch scenarios.
+{\bf Interface management.} All of these modifications work in tandem to
+implement the desired behavior, and are typically co-designed together such
+that each depends on the other to behave as expected. In highly dynamic
+environments such as a distributed storage system, changes to the system
+propagate at variable rates, requiring all components of the system to be able
+to adapt. Thus, system services realized through extensibility should be
+versioned together as a group, exposing the version to allow introspection and
+domain-specific treatment of version mismatch scenarios.
 

### 2016-05-10 paper/paper.tex
paper: evaluation mantle part 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-05-10 paper/paper.tex
@@
 \usepackage{ifxetex,ifluatex}
+\usepackage{amsmath}
+\usepackage{mathptmx}
+\usepackage{graphicx}
+\usepackage{listings}
+
 % use upquote if available, for straight quotes in verbatim environments
@@
 
+\if{false}
+For example, a Ceph cluster contains a small number of
+monitoring nodes (MONs) that use PAXOS to maintain a consistent view of
+system-wide metadata such as data distribution parameters and cluster
+membership. The data distribution metadata maintained by
+the cluster monitors is used by RADOS servers and clients to provide
+consistent reads and writes, and both RADOS and the monitoring services
+are used by a cluster of file system metadata servers that provide POSIX
+semantics using sophisticated indexing and distributed locking services.
+We contend that re-using and re-purposing these code-hardened subsystems
+is paramount to successfully adapting storage systems to new APIs and
+new storage devices without losing the benefits from years of
+code-hardening work.
+\fi
+
 Ceph also provides empirical evidence that developers embrace programmable storage: Ceph already supports the
@@
 By embedding the Lua runtime, Malacology makes the base RADOS API available in a 
-dynamic environment, resulting in an object storage API with economy of 
+dynamic environment, resulting in a object storage API with economy of 
 expression, which at the same time provides the full set of features of the base 
@@
 code~\cite{geambasu_comet_2010}.
-
-{\bf Recovery.}
-Application specific recovery is critical.
 
@@
 
-\paragraph*{Malacology:} Metadata service policies and object storage interfaces are stored durably
-within RADOS and managed by storing references with the object maps. Since
-the cluster already propagates a consistent view of these data structures,
-we use this service to automatically install interfaces in OSDs, and install
-policies within the MDS such that clients and daemons are synchronzied on
-correct implementations without restarting.
+\paragraph*{Malacology:}
 
@@
 
-\paragraph*{Malacology:} Malacology exposes the monitoring service in two ways. 
-First, Malacology APIs for object storage (\S\ref{active-storage}) and the MDS 
-(\S\ref{malacology:mds}) have functions available to them that allow them to 
-store application-specific data in the highly-available, consistent store 
-associated to the Ceph monitors. As mentioned
-Map modification for propogation of state. We allow applications to augment
+\paragraph*{Malacology:} Map modification for propogation of state. We allow applications to augment
 these data structures with application-specific metadata. In order to keep the
@@
 
-\paragraph*{Malacology:} Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.
-Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.
-
-These hooks embed Lua code in the MDS using the
-OSD subsystems.
-By re-using the OSD interface classes, the MDS gets the safety and robustness of loading dynamic code (including graceful failures and helpful errors), the ease of transferring state between object
-interfaces and the OSD internals, integration with testing and
-correctness suites, and \texttt{struct}s for interface data and
-handlers.
-
-The metadata service also manages sessions with its clients. This gives both subsystems (MDSs and clients) the ability cache data and grab locks. The metadata service is
-responsible for responding to client requests for capabilities by first
-revoking caps from other clients. Currently clients voluntarily
-release their capabilities, and the MDS maintains a queue of requests.
-Both clients and MDSs use a best-effort approach to revoking or releasing capabilities.
-
-In our current we retain the best-effort, voluntary release of caps within
-clients, but generalize the queueing policies within the MDS to allow for
-centeralized policies such as fairness or priority.
-
-{\bf File Types.}
-We can reference PDSW12 DataMods, and talk about any new stuff here that we
-want to.
+\paragraph*{Malacology:}
 
 \subsection{Metadata Load Balancing}
-\label{malacology:mds}
 
@@
 
+\section{Malacology Implementation}
+\label{implementation}
+
+In this section we describe how Malacology builds upon existing
+services within Ceph. The main idea behind Malacology is to take the subsystems 
+described in the previous section and expose APIs on top of them that allow 
+users to dynamically inject code, allowing them to customize the behavior of 
+these components. Supporting dynamic reconfiguration and programmability 
+efficiently is critical in real systems that must adapt to a changing 
+environment where potentially
+conflicting policies arise. We've adopted Lua as the dynamic runtime  to
+implement dynamic components in Malacology.
+
+There are multiple reasons that make Lua an attractive runtime for implementing 
+Malacology. Lua is a portable, embedded
+scripting language that offers superious performance and productivity
+trade-offs, including a JIT-based implementation that is well known for near
+native performance. Additionally, Lua has been used extensively in game engines, 
+and systems research \cite{neto:dls14-luaos}, including storage systems where it 
+has been effectively used both on
+\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010} and 
+off
+\cite{sevilla:sc15-mantle} the performance critical path. Finally, the 
+flexibility of the runtime allows to sandbox applications in order to address 
+security and performance concerns. We now describe how Malacology exposes each 
+subcomponent in Ceph.
+
+\subsection{Object Storage}
+
+% todo:
+%  - ellude to the need for interfaces to be avialable quickly to support
+%    our experiment about propogtation delay
+%
+
+The implementation of Ceph's object abstraction, although powerful, has some 
+drawbacks.
+Supporting only C/C++ for object interface developers, Ceph requires
+distribution of compiled binaries for the correct architecture, adding a
+large barrier of entry for developers and system administrators. Second,
+having no way to dynamically unload modules, any changes require a full
+restart of an OSD which may have serious performance impacts due to a loss
+of cached data. And finally, the security limitations of the framework
+limit the use of object classes to all but those with administrative level
+access and technique expertise.
+
+By embedding the Lua runtime, Malacology makes the base RADOS API available in a 
+dynamic environment, resulting in a object storage API with economy of 
+expression, which at the same time provides the full set of features of the base 
+object class. New object interfaces that are expressed in thousands of lines of 
+code can be implemented in approximately an order of magnitude less 
+code~\cite{geambasu_comet_2010}.
+
+\subsection{Metadata Services}
+
+% todo:
+%  - ellude to the need for policies to be avialable quickly to support
+%    our experiment about propogtation delay
+% Original: inject strings
+% New: use C++ bindings
+
+Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.
+
+\subsubsection{Balancing Policy}
+
 For CephFS, designing a general POSIX metadata load balancer is difficult becuase the workloads are diverse and quickly changing.
@@
 -- but we choose instead to expose a balancing API at strategic points in the balancing logic using Lua hooks.
+
+These hooks embed Lua code in the MDS using the
+OSD subsystems.
+By re-using the OSD interface classes, the MDS gets the safety and robustness of loading dynamic code (including graceful failures and helpful errors), the ease of transferring state between object
+interfaces and the OSD internals, integration with testing and
+correctness suites, and \texttt{struct}s for interface data and
+handlers.
+
+To facilitate the use of Lua in both the MDS and OSD, we modularize the LuaVM in a core sandbox wrapper and link against it, as shown in Figure~\ref{fig:cls-osd-mds}. This core Lua wrapper contains
+just the dependencies, symbols, and functions, needed to run the Lua VM and strips away all the functionality inherent to object classes (like placement group filters, cryptography functions, object metadata,  object data, object extended attributes, and object versions).
+Some functions of Ceph's more helpful subsystems, like the logging facilities for the daemons, were re-used in the MDS.
+With this scheme, The OSD will dynamically open the shared library created by Lua object interface shared library while the MDS will dynamically open the Lua balancer interface shared
+library; both shared libraries will the Lua core.
 
@@
 
-To facilitate the use of Lua in both the MDS and OSD, we modularize the LuaVM in a core sandbox wrapper and link against it, as shown in Figure~\ref{fig:cls-osd-mds}. This core Lua wrapper contains
-just the dependencies, symbols, and functions, needed to run the Lua VM and strips away all the functionality inherent to object classes (like placement group filters, cryptography functions, object metadata,  object data, object extended attributes, and object versions).
-Some functions of Ceph's more helpful subsystems, like the logging facilities for the daemons, were re-used in the MDS.
-With this scheme, The OSD will dynamically open the shared library created by Lua object interface shared library while the MDS will dynamically open the Lua balancer interface shared
-library; both shared libraries will the Lua core.
+\subsubsection{Capabilities}
+
+The metadata service also manages sessions with its clients. This gives both subsystems (MDSs and clients) the ability cache data and grab locks. The metadata service is
+responsible for responding to client requests for capabilities by first
+revoking caps from other clients. Currently clients voluntarily
+release their capabilities, and the MDS maintains a queue of requests.
+Both clients and MDSs use a best-effort approach to revoking or releasing capabilities.
+
+In our current we retain the best-effort, voluntary release of caps within
+clients, but generalize the queueing policies within the MDS to allow for
+centeralized policies such as fairness or priority.
+
+{\bf File Types.}
+We can reference PDSW12 DataMods, and talk about any new stuff here that we
+want to.
+
+\subsection{Cluster State}
+
+Map modification for propogation of state. We allow applications to augment
+these data structures with application-specific metdata. In order to keep the
+size of these data structures small application state may be stored elsewhere
+and the references stored within the data structure.
+
+\subsection{Programmability and Durability}
+
+{\bf Recovery.}
+Application specific recovery is critical.
+
+{\bf Durability.}
+
+Metadata service policies and object storage interfaces are stored durability
+within RADOS and managed by storing references with the object maps. Since
+the cluster already propagates a consistent view of these data structures,
+we use this service to automatically install interfaces in OSDs, and install
+policies within the MDS such that clients and daemons are synchronzied on
+correct implementations without restarting.
+
+\iffalse
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  additional interface hooks and classes for OSDs and MDSs using the
+  existing Ceph \texttt{cls} infrastructure
+\item
+  durable interface classes using the Ceph object store
+\item
+  versioned and consistent interfaces using the Ceph MONs
+\end{enumerate}
+
+Figure \ref{implementation} shows how interfaces can be loaded into
+either the MDS or OSD (1), how interfaces are persisted to object
+storage (2), and how they are versioned by the MONs (3). In the
+following section, we will discuss the existing Ceph infrastructure and
+the changes necessary for Malacology. We emphasize here that these
+implementations are not necessarily optimized but they help demonstrate
+the power of programmability.
+
+\subsection{Additional Interface
+Frameworks}\label{additional-interface-frameworks}
+
+For object interfaces, clients send Lua classes to the OSD and then they
+can invoke operations in that class remotely on the OSD. In reality,
+this feature is a statically loaded object class written in C/C++ that
+runs dynamically defined object interfaces, where clients access the
+classes using an \texttt{exec()} wrapper. Although sending the class
+with each client request is stateless (which has its own set of nice
+features), it is costly for network bandwidth and difficult to organize
+at the application level. Also, while the execute wrapper simplifies the
+implementation, it burdens the applications, since they need to be
+recompiled to use the \texttt{exec()} function.
+
+Although these OSD interfaces are powerful, the current implementation
+has drawbacks. OSD interfaces are written in C/C++ and compiled into a
+shared library, so the developer must account for different target
+architectures. Second, C/C++ provides more functionality than is
+necessary since the snippets of interface code mainly set policies and
+perform simple operations. Finally, developing these interfaces in C/C++
+has high overhead. The developer must learn how the OSD dynamically
+loads the shared libraries (\emph{e.g.}, OSDs rely on a strict naming
+convention to find shared libraries), how to get their interfaces
+compiled using the Ceph \texttt{make} system, and how to debug issues
+that are not related to their specific interface (\emph{e.g.}, the OSD
+cannot find the shared library) - this learning curve is unacceptable
+for non-Ceph developers, especially since most interfaces are one-off
+solutions specific to their applications.
+
+\subsubsection{New Interface Classes}\label{new-interface-classes}
+
+Our framework has added a mechanism for defining and running object and
+metadata balancer classes using the Lua scripting language. Scripting is
+useful in large systems for 3 reasons:
+
+\begin{itemize}
+\item
+  accelerates deployment: tweaking policies, like changing a threshold
+  or metric calculation, does not require a full recompile or system
+  start-up.
+\item
+  transparency: if the bindings carefully expose the correct metrics and
+  functions, the user does not need to learn the internal functions,
+  variables, or types of the system.
+\item
+  portability: scripts can be exchanged across platforms and system
+  versions without compatibility issues.
+\end{itemize}
+
+Our Lua bindings expose functions and symbols both ways; the host
+program can call functions defined in Lua and the Lua scripts can call
+functions defined in native C++. These bindings are merged upstream. Lua
+is frequently used in game engines to set policies but we use it here
+because most of the user-defined classes in Ceph are policies as well!
+We do not want to provide specific implementations, like pulling data
+from objects or transferring them over the network, but instead strive
+to say \emph{what to do} with the data once we have it. Separating
+policy from mechanism is a driving factor in using Lua. We choose Lua
+for 4 reasons: performance, portabability, size, and security.
+
+Lua performs well as an embeddable scripting language. Although it is an
+interpreted language, the bindings are designed to make the exchange of
+variables and functions amongst languages straightforward. The LuaJIT
+virtual machine boasts near-native performance, making it a logical
+choice for scriptability frameworks in systems research
+\cite{neto:dls14-luaos}. In storage systems, it has been effectively
+used both on \cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo} and off
+\cite{sevilla:sc15-mantle} the critical path, where performance is
+important.
+
+Because Lua is interpreted, it is also portable. The portability allows
+functions to be shipped around the cluster without having to compile or
+reason about compatibility issues. Traditionally, intefaces are written
+in C/C++, so they must be compiled into shared libraries before being
+injected into the daemon at runtime. Shared libraries need to be
+compiled on the host they will be run on to accomodate different
+architectures and dependendencies. Embedding the Lua interpretator, on
+the other hand, gives us the flexbility to store the actual Lua code in
+other places, like RADOS, so we can enjoy all the other properties
+provided by that particular back-end.
+
+The other two advantages, that Lua leaves a small memory footprint and
+that it is secure, are discussed at length in 
+\cite{ierusalimschy_programming_2006,neto:dls14-luaos}. In short, Lua
+provides relatively robust sandbox capabilities for protecting the
+system against some malicious cases, but does not protect the system
+from poor policies. Since Malacology does not use these features (yet),
+we will avoid discussion here.
+
+\subsection{New Interface Hooks}\label{new-interface-hooks}
+
+\fi
+
+\iffalse
+
+\subsection{Programmability and Durability}
+
+Traditionally, object storage interfaces are compiled into shared
+libraries and loaded into a running OSD daemon using \texttt{dlopen()}.
+Because the shared library has to link against symbols found in the
+running executable, the interfaces are stored on the local file systems
+of each OSD so they can be versioned and distributed alongside the same
+Ceph binaries. This policy is for safety, since the OSDs could be on
+different distros that provide different versioning capabilities. Also,
+this approach allows bug fixes to co-evolve with the rest of the Ceph
+installation.
+
+Because of the portability of Lua and because the Lua virtual machine is
+compiled and loaded as a shared library, the MDS/OSD can store the
+balancers in a more redundant location. For our implementation, we store
+balancers in RADOS and the daemons first search RADOS, then the local
+file system for their interface classes.
+
+\subsection{Versioning Interfaces with
+MONs}\label{versioning-interfaces-with-mons}
+
+We added two new commands to Ceph command line client so the user can
+inject interface versions into the cluster::
+
+\noindent \texttt{ceph\ osd\ pool\ set-class\ \textless{}pool\textgreater{}\ \textless{}class\textgreater{}\ \textless{}script\textgreater{}}
+
+\noindent \texttt{ceph\ mds\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}
+
+By leveraging the monitor daemons, our services get the consistency from
+the monitors version management, the distribution from the monitors data
+structures (which are already distributed), and the durability from the
+robustness of the monitors and persistence with Paxos. The
+implementation uses the placement group pool data structure which
+maintains the policies (\emph{e.g.}, erasure coding) and organization
+details (\emph{e.g.}, snapshots) for each pool. While stuffing the
+information into the monitor map, the structure that describes the the
+monitor topology, was an option, the placement group pool allows finer
+grain control over different types of data.
+
+In regards to the consistency, each time the user injects a new Lua
+object class it enters the monitor quorum as a proposal; updates to the
+placement group pool need to wait until accepted, at which point the
+state will be propogated and versioned. We hooked up the monitor command
+to the \texttt{cls\_lua} class by having the OSD pull the script from
+the placement group instead of trying to dynamically load the shared
+library with \texttt{dlopen()} or the Lua script.
+
+The cls-lua branch lets users write object classes in Lua, the lua-rados
+client library lets users write applications that talk to RADOS with
+Lua, and the cls-client uses the lua-rados library to send Lua object
+classes to the OSDs equipped with the LuaJIT VM.
+\fi
 
@@
 
-Mantle~\cite{sevilla:sc15-mantle} is a programmable metadata
+Mantle is a programmable metadata
 balancer for Ceph that separates the metadata balancing policies from their
@@
 {\bf Capability policies.}
-We use capability policies in the metadata service to tune the amount of
+We use capabilitity policies in the metadata service to tune the amount of
 time that clients may cache the sequencer resource. This allows us to control
@@
 \section{Evaluation}\label{evaluation}
-
-1 Experiments: 2 3 1. mdtest separate directories 4 - just based on
-default traffic 5 - kill one of the MDSs 6 7 2. filebench profiles 8 -
-adaptable vs.~greedy vs.~fill\&spill 9 10 3. sequencer balancer 11 -
-just based on default traffic 12 13 3. OpenSSL compile
-\textasciitilde{}\\\label{evaluation}
+We have conducted our evaluation on CloudLab as well as our own 20 node cluster.
+We follow the ``Popper Convention"~\cite{} so as to try and make our experiments as reproducible as possible.
+We have a GitHub repository with our source code,  experimental environment, configuration parameters, and workloads in a GitHub repository.
+Due to the double-blind policy, we have temporarily disabled the links to our graphs and experimental details.
+
+Much of this work focuses on the programmability of Ceph and the services that
+can be layered on top of Malacology. As such, we focus on the performance of the 
+components and subsystems exposed by the framework. Detailed system-wide performance
+measurements and analysis is forthcoming for both Mantle and ZLog~\cite{}, as we try
+to find the best metadata balancing policies and physical design implementation trade-offs
+respectively.
 
@@
 system client, throughput drops significantly with two clients as the majority
-of the time neither client has the capability as it ping-pongs between
+of the time neither client has the capapbitility as it ping-pongs between
 clients. This observation reveals a tuning parameter: how long a client
-maintains its capability lease. In throughputs labeled $2C,<time>$ we show
+maintains its capapbility lease. In throughputs labeled $2C,<time>$ we show
 aggregate throughput as we decrease the amount of time that a client holds its
-lease before responding to a request for the capability.  At lease times as
+lease before responding to a request for the capapbility.  At lease times as
 low as 10ms throughputs achieve a significant proportion of the locally cached
@@
 for applications that require predictable latency. The throughput in
-Figure~\ref{fig:captp} labeled $2C,ZLog$ corresponds to a centralized
+Figure~\ref{fig:captp} labeled $2C,ZLog$ corresponds to a centeralized
 sequencer implementation. While throughput is lower for two clients (because of
@@
 
-\subsection{Mantle}
-
 \subsection{MDS as a Sequencer}
+
+The following experiments demonstrate the feasibility of using the Ceph MDS
+cluster as a sequencer. The goal of the workloads is to saturate the MDS
+cluster, so we design sequencers uses a more heavy-weight file system
+operation. The experiments use our internal cluster and there are 15 clients
+and between 1 and 3 MDS nodes.
+
+WE NEED SOMETHING HERE DESCRIBING WHY AN INODE IS A GOOD ABSTRACTION
+
 \begin{figure}[h]
@@
 \includegraphics{figures/mantle-seq-thruput.png}
-\caption{A ZLog sequencer that uses the CephFS metadata cluster enjoys all the benefits of the MDS recovery ("Sequencer Recovery") and load balancing protocols (``Mantle Load Balancing'').
-The ``2 Seq'' curves are the throughputs of 1 sequencer when an additional sequencer is running in the background generating load.
-The ``Sequencer Baseline'' has no background sequencers.}
+\caption{A ZLog sequencer that uses the CephFS MDS enjoys the load balancing
+capabilities of the metadata cluster. The ``Sequencer Baseline'' has no
+background sequencers, while the the ``2 Seq'' curves are the throughputs
+of 1 sequencer when an additional sequencer is running in the background
+generating load.} 
 \label{fig:mantle-seq-thruput.png}
 \end{figure}
+
+\subsubsection{Load Balancing}
+
+We use a balancer policy specifically designed for the sequencer. The balancer aggressively sheds half its load to its neighbor as soon as it can.  It is based
+off the ``Greedy Spill Even" approach in the Mantle~\cite{sevilla:sc15-mantle} paper
+but is reproduced here for the reader's convenience.
+
+\begin{figure}[h]
+\begin{lstlisting}
+  -- Metadata server load
+  mdsload = MDSs[i]["all"]
+
+  -- When policy
+  if MDSs[whoami]["load"]>.01 and 
+    MDSs[whoami+1]["load"]<.01 then
+
+  -- Where policy
+  targets[whoami+1]=allmetaload/2
+\end{lstlisting}
+    \caption{The sequencer balancing policy that works best for bursty workloads. \label{listing:greedy-spill}}
+\end{figure}
+
+Figure~\ref{fig:mantle-seq-thruput.png} gives a macro-level view of the
+performance of the Ceph components exposed by Malacology. Each curve shows the
+throughput from the perspective of one sequencer ({\it y} axis) over time ({\it
+x} axis). The ``Sequencer Baseline'' curve is one sequencer serving 7 clients.
+The sequencer operates on an MDS cluster where balancing is turned off and
+there are no other workloads running. ``2 Seq; No Load Balancing" adds another
+sequencer of equal load running in the background. The performance of the
+background sequencer is omitted. ``2 Seq; Mantle Load Balancing" uses a 3 MDS
+node cluster to provide the sequencer services, again with a background
+sequencer service. Finally, the ``Sequencer Recovery" curve shows the
+performance of the sequencer in a multi-MDS cluster with 2 standby MDS nodes.
+At 60 seconds, MDS 0 is killed.
+
+Overall, the clients achieve lower performance than the capability experiments
+in Figure~\ref{fig:captp} because the sequencer uses a slower metadata
+operation to manage the capability.  ``Sequencer Baseline" is the best
+performance for this particular workload because there is no contention.
+Performance is stable and the MDS is underloaded because we do not see the
+spike at the end like the other curves. It appears the bottleneck is somewhere
+with the clients -- perhaps it is our 1GBit/s network link. ``2 Seq; No Load
+Balancing" shows an overloaded sequencer because the performance of our one
+sequencer is low and unstable (high variation in performance). We also deduce
+that it is overloaded by comparing to the ``2 Seq; Mantle Load Balancing"
+curve, which shows improved performance when load is migrated. At 10 seconds, 
+the sequencer balancer immediately starts balancing and decides that two
+sequencer streams should be split onto two MDSs. It migrates to its neighbor as
+per the policy.
+
+\textbf{The Malacology Principle:} Re-uses the Ceph metadata migration, load
+balancing, cache coherence, and capabilities subsystems of the MDS. Applying
+Mantle's load balancing policy shows 2\(\times\) performance improvent and x\%
+of the baseline best-case performance.
+
+\subsubsection{Recovery}
+
+- Describe the graph
+
+- We get recovery and load balancing with minimal implementation
+
+\subsection{Load Balancing}
 
@@
 
-\subsection{Load Balancing}
+- Sequencer specific policy
+
+- Describe the graph
+
+- We get tunable policies
 
@@
 system deficiencies. However, this duplication has real-world problems
-related to reliability. We propose that systems expose their services in
-a safe way, allowing application developers to customize system behavior
+related to reliability. We propose that system expose their services in
+a safe way allowing application developers to customize system behavior
 to meet their needs while not sacrificing correctness.
@@
 services to be configured on-the-fly in existing systems. This work is
-one point along that path in which we have looked at a target
+one point along that path in which we have looked an a target
 special-purpose storage system. Ultimately we want to utilize

### 2016-05-10 paper/paper.tex

--- 2016-05-10 paper/paper.tex
@@
 \section{Evaluation}\label{evaluation}
+We have conducted our evaluation on CloudLab as well as our own 20 node cluster.
+We follow the ``Popper Convention"~\cite{} so as to try and make our experiments as reproducible as possible.
+We have a GitHub repository with our source code,  experimental environment, configuration parameters, and workloads in a GitHub repository.
+Due to the double-blind policy, we have temporarily disabled the links to our graphs and experimental details.
 
-1 Experiments: 2 3 1. mdtest separate directories 4 - just based on
-default traffic 5 - kill one of the MDSs 6 7 2. filebench profiles 8 -
-adaptable vs.~greedy vs.~fill\&spill 9 10 3. sequencer balancer 11 -
-just based on default traffic 12 13 3. OpenSSL compile
-\textasciitilde{}\\\label{evaluation}
+Much of this work focuses on the programmability of Ceph and the services that
+can be layered on top of Malacology. As such, we focus on the performance of the 
+components and subsystems exposed by the framework. Detailed system-wide performance
+measurements and analysis is forthcoming for both Mantle and ZLog~\cite{}, as we try
+to find the best metadata balancing policies and physical design implementation trade-offs
+respectively.
 
@@
 
-\subsection{Mantle}
+\subsection{MDS as a Sequencer}
 
-\subsection{MDS as a Sequencer}
+The following experiments demonstrate the feasibility of using the Ceph MDS
+cluster as a sequencer. The goal of the workloads is to saturate the MDS
+cluster, so we design sequencers uses a more heavy-weight file system
+operation. The experiments use our internal cluster and there are 15 clients
+and between 1 and 3 MDS nodes.
+
+WE NEED SOMETHING HERE DESCRIBING WHY AN INODE IS A GOOD ABSTRACTION
+
 \begin{figure}[h]

### 2016-05-10 paper/paper.tex
sec2.3
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-05-10 paper/paper.tex
@@
 
-% Straw man example
-Migrating resources is an important part of load balancing but understanding the effects of different configurations and parameters is difficult.
-Systems already virtualize and migrate memory; soon systems will be able to migrate other resources, like CPU, disk, and memory.
-Figuring out how, when, and where to move these resources depends on complex metrics like utilization, configuration, and workload.
+The distributed metadata service in Ceph provides clients with a POSIX file
+system abstraction~\cite{cephfs}. Distributed file systems are complex, but
+fundamentally provide hierarchical indexing and distributed locking services.
+In Ceph, the locking service implements a capability-based system that express
+what data and metadata clients are allowed to access as well as what state
+they may cache and modify locally. While designed for a fixed file
+abstraction, indexing, locking, and caching are all common services that are
+useful to a broad spectrum of applications.
 
-These problems are general, distributed systems load balancing problems and share many of the same properties as the POSIX metadata wall~\cite{alam:pdsw2011-metadata-scaling,ghemawat:sosp2003-gfs,hildebrand:msst2005-pnfs,weil_ceph_2006,welch:fast2008-panasas,shvachko:login2012-hdfs-scalability}.
-Many distributed file systems decouple metadata and data I/O so that
-these services can scale independently. Despite
-this optimization, scaling the metadata services is still difficult
-because metadata accesses impose small and frequent requests on the
-underlying storage system~\cite{roselli:atec2000-FS-workloads}. Many
-techniques for designing the metadata services have been proposed to
-accommodate this workload: Lustre~\cite{konstantinos:pdsw2014-lustre-metadata}, GFS~\cite{ghemawat:sosp2003-gfs}, and HDFS~\cite{shvachko:login2012-hdfs-scalability} keep all metadata on one
-server; GIGA+~\cite{patil:fast2011-giga}, IndexFS~\cite{ren:sc2014-indexfs}, Lazy Hybrid~\cite{brandt:msst2003-lh}, GPFS~\cite{schmuck:fast2002-gpfs},
-and pNFS~\cite{hildebrand:supercomputing2006-pNFS} hash the file
-system namespace across a dedicated cluster and cache metadata; Panasas~\cite{welch:fast2008-panasas} and CephFS~\cite{weil:sc2004-dyn-metadata} use subtree partitioning to carve up the namespace and distribute it across a cluster.
-
-CephFS moves units of the namespace called directory fragments and assigns them to servers using  its own metadata load balancer.
-It then uses a hard-coded policy to balance these fragments using a secularized load metric based on CPU, workload, and file system operation metrics~\cite{sevilla:sc15-mantle}.
+Ceph addresses the challenge of balancing load across the metadata cluster
+using load balancing policies that compute balancing metrics based on system
+state (e.g. CPU and memory utilization) and statistics collected by the
+cluster (e.g. the popularity of an inode). Distributed applications that share
+centralized resources (e.g. a database or directory) face similar challenges
+which are often solved using application-specific sharding.
 
 \paragraph*{Malacology:}
+Malacogy adds interface hooks in strategic spots in the metadata service to
+control load balancing, capability subsystems, and for defining new file
+types allowing the indexing service to be repurposed for other types of
+named resources.
 
-For CephFS, designing a general POSIX metadata load balancer is difficult becuase the workloads are diverse and quickly changing.
-With CephFS, it is especially complicated because the system has variable sized migration units (i.e.
-file system namespace subtrees), which can be shipped anywhere (i.e., to any MDS of any capacity) at any time for any reason.
-Furthermore, the original balancer was designed with hard-coded policies and tunables.
-Both angles were difficult to understand.
-It is understandable why it was implemented this way -- it is difficult to pick a one-size-fits-all policy, system administrators may want to partition load based on SLAs, or applications are better equipped to make these decisions, etc.
--- but we choose instead to expose a balancing API at strategic points in the balancing logic using Lua hooks.
+{\bf Load balancing.}
+For CephFS, designing a general POSIX metadata load balancer is difficult
+becuase the workloads are diverse and quickly changing. With CephFS, it is
+especially complicated because the system has variable sized migration units
+(i.e. file system namespace subtrees), which can be shipped anywhere (i.e., to
+any MDS of any capacity) at any time for any reason. Furthermore, the original
+balancer was designed with hard-coded policies and tunables. Both angles were
+difficult to understand. It is understandable why it was implemented this way
+-- it is difficult to pick a one-size-fits-all policy, system administrators
+may want to partition load based on SLAs, or applications are better equipped
+to make these decisions, etc. -- but we choose instead to expose a balancing
+API at strategic points in the balancing logic using Lua hooks.
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/cls-osd-mds.png}
-\caption{To generalize the LuaVM for use in both the OSD and MDS,
-Malacology detaches the core functionality into a Lua core module. The
-VM is dynamically loaded first (1) and then interfaces can execute (2).
-\label{fig:cls-osd-mds}}
-\end{figure}
+These problems are general, distributed systems load balancing problems and
+share many of the same properties as the POSIX metadata
+wall~\cite{alam:pdsw2011-metadata-scaling,ghemawat:sosp2003-gfs,hildebrand:msst2005-pnfs,weil_ceph_2006,welch:fast2008-panasas,shvachko:login2012-hdfs-scalability}.
+Many distributed file systems decouple metadata and data I/O so that these
+services can scale independently. Despite this optimization, scaling the
+metadata services is still difficult because metadata accesses impose small
+and frequent requests on the underlying storage
+system~\cite{roselli:atec2000-FS-workloads}. Many techniques for designing the
+metadata services have been proposed to accommodate this workload:
+Lustre~\cite{konstantinos:pdsw2014-lustre-metadata},
+GFS~\cite{ghemawat:sosp2003-gfs}, and
+HDFS~\cite{shvachko:login2012-hdfs-scalability} keep all metadata on one
+server; GIGA+~\cite{patil:fast2011-giga}, IndexFS~\cite{ren:sc2014-indexfs},
+Lazy Hybrid~\cite{brandt:msst2003-lh}, GPFS~\cite{schmuck:fast2002-gpfs}, and
+pNFS~\cite{hildebrand:supercomputing2006-pNFS} hash the file system namespace
+across a dedicated cluster and cache metadata;
+Panasas~\cite{welch:fast2008-panasas} and
+CephFS~\cite{weil:sc2004-dyn-metadata} use subtree partitioning to carve up
+the namespace and distribute it across a cluster.
+
+CephFS moves units of the namespace called directory fragments and assigns
+them to servers using  its own metadata load balancer. It then uses a
+hard-coded policy to balance these fragments using a secularized load metric
+based on CPU, workload, and file system operation
+metrics~\cite{sevilla:sc15-mantle}.
 
@@
 
-To facilitate the use of Lua in both the MDS and OSD, we modularize the LuaVM in a core sandbox wrapper and link against it, as shown in Figure~\ref{fig:cls-osd-mds}. This core Lua wrapper contains
-just the dependencies, symbols, and functions, needed to run the Lua VM and strips away all the functionality inherent to object classes (like placement group filters, cryptography functions, object metadata,  object data, object extended attributes, and object versions).
-Some functions of Ceph's more helpful subsystems, like the logging facilities for the daemons, were re-used in the MDS.
-With this scheme, The OSD will dynamically open the shared library created by Lua object interface shared library while the MDS will dynamically open the Lua balancer interface shared
-library; both shared libraries will the Lua core.
+To facilitate the use of Lua in both the MDS and OSD, we modularize the LuaVM
+in a core sandbox wrapper and link against it, as shown in
+Figure~\ref{fig:cls-osd-mds}. This core Lua wrapper contains just the
+dependencies, symbols, and functions, needed to run the Lua VM and strips away
+all the functionality inherent to object classes (like placement group
+filters, cryptography functions, object metadata,  object data, object
+extended attributes, and object versions). Some functions of Ceph's more
+helpful subsystems, like the logging facilities for the daemons, were re-used
+in the MDS.  With this scheme, The OSD will dynamically open the shared
+library created by Lua object interface shared library while the MDS will
+dynamically open the Lua balancer interface shared library; both shared
+libraries will the Lua core.
 
-\subsection{File System Indexing and Locking}
-
-The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction. Distributed file systems are complex, but fundamentally
-provide durable hierarchical indexing and distributed locking service.  The
-locking service implements a capabilities that express what data and metadata
-clients are allowed to access as well as what they may cache locally.  While
-designed for a fixed file abstraction, indexing, locking, and caching are all
-common services that are useful to large sets of applications.
-
-\paragraph*{Malacology:} Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.
-Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.
-
-These hooks embed Lua code in the MDS using the
-OSD subsystems.
-By re-using the OSD interface classes, the MDS gets the safety and robustness of loading dynamic code (including graceful failures and helpful errors), the ease of transferring state between object
-interfaces and the OSD internals, integration with testing and
-correctness suites, and \texttt{struct}s for interface data and
-handlers.
-
-The metadata service also manages sessions with its clients. This gives both subsystems (MDSs and clients) the ability cache data and grab locks. The metadata service is
-responsible for responding to client requests for capabilities by first
-revoking caps from other clients. Currently clients voluntarily
-release their capabilities, and the MDS maintains a queue of requests.
-Both clients and MDSs use a best-effort approach to revoking or releasing capabilities.
+{\bf Capabilities.}
+The metadata service also manages sessions with its clients. This gives both
+subsystems (MDSs and clients) the ability cache data and grab locks. The
+metadata service is responsible for responding to client requests for
+capabilities by first revoking caps from other clients. Currently clients
+voluntarily release their capabilities, and the MDS maintains a queue of
+requests.  Both clients and MDSs use a best-effort approach to revoking or
+releasing capabilities.
 
@@
 want to.
+
+{\bf Policy management.} A new interface is added to the metadata service
+subsystem using the monitoring framework to support management of interfaces,
+which ensures durability and distribution of policies in a consistent way.
+
+\begin{figure}[h]
+\centering
+\includegraphics{figures/cls-osd-mds.png}
+\caption{To generalize the LuaVM for use in both the OSD and MDS,
+Malacology detaches the core functionality into a Lua core module. The
+VM is dynamically loaded first (1) and then interfaces can execute (2).
+\label{fig:cls-osd-mds}}
+\end{figure}
+
+% Straw man example
+%Migrating resources is an important part of load balancing but understanding
+%the effects of different configurations and parameters is difficult. Systems
+%already virtualize and migrate memory; soon systems will be able to migrate
+%other resources, like CPU, disk, and memory. Figuring out how, when, and where
+%to move these resources depends on complex metrics like utilization,
+%configuration, and workload.
+
+%These hooks embed Lua code in the MDS using the OSD subsystems. By re-using
+%the OSD interface classes, the MDS gets the safety and robustness of loading
+%dynamic code (including graceful failures and helpful errors), the ease of
+%transferring state between object interfaces and the OSD internals,
+%integration with testing and correctness suites, and \texttt{struct}s for
+%interface data and handlers.
 

### 2016-05-10 paper/paper.tex
Added beginnings of a Challenges section.

@@
 conclude by evaluating real-world use cases (\S\ref{evaluation}).
+
+\section{Challenges}
+\label{sec:challenges}
+
+The challenges of introducing programmability into existing storage systems: (1) Programmability while keeping storage system available, (2) Finding and factoring out hard-wired policies and optimizations, (3) Bugs that were hidden because mechanisms were only exercised according to hard-wired policies and not in their full generality they were thought of -- programmability into a storage system turned out to be a great debugging tool.
 

### 2016-05-10 paper/paper.tex
3.5
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-05-10 paper/paper.tex
@@
 paper, we'll briefly describe the custom storage device interface, sequencer
-service, and recovery protocol that form the core components of CORFU that we
-implement within Malacology:
+service, and recovery protocol, and how these services are instantiated in the
+Malacology framework.
 
-{\bf Sequencer.}
+\subsubsection{Sequencer}
+
 High-performance in CORFU is achived using a sequencer service that assigns
@@
 
-{\bf Storage Interface.} The storage interface is a critical component in the
-CORFU design. Clients independently map log positions onto storage devices,
-while storage devices provide an intelligent \emph{write}-once, random
-\emph{read} interface for accessing log entries. The key to correctness in
-CORFU lies with the enforcement of up-to-date epoch tags on client requests;
-requests tagged with out-of-date epoch values are rejected, and clients are
-expected to request a new tail from the sequencer after refreshing state from
-an auxillary service. This mechanism forms the basis for sequencer recovery.
+We implement the sequencer service in Malacology as an application-specific
+file type that associates a small amount of metadata, including the 64-bit log
+tail position, to the inode state managed by the Ceph metadata service. Since
+a sequencer is associated with a unique log, this approach has the added
+benefit of allowing the metadata service to also handle naming, by
+representing each log service instance within the standard POSIX hierarchical
+namespace.
 
-{\bf Sequencer Recovery.}
+{\bf Sequencer interface.} The sequencer resource supports the ability to
+\emph{read()} the current tail value as well get the \emph{next()} position in
+the log which also atomically increments the tail position. We implement this
+functionality as methods on ZLog inode resources. The primary challenge in
+mapping the sequencer resource to the metadata service is handling
+serialization correctly.
+
+Initially we sought to directly model the sequencer service in Ceph as a
+non-exclusive, non-cacheable resource, forcing clients to perform a round-trip
+access to the resource at the authoratiave metadata server for the sequencer
+inode.  Interestingly, we found that the capapbility system in Ceph uses a
+strategy to reduce metadata service load by allowing multiple clients that
+open a shared file to temporarily cache the inode resource, resulting in a
+round-robin, best-effort batching behavior. When a single client is accessing
+the sequencer resource it is able to cache state locally without interruption.
+
+While unexpected, this discovery allowed us to explore an implementation
+strategy that we had not previously considered. In particular, for bursty
+clients, and clients that can tolerate increased latency, this mode of operation may
+allow a system to achieve much higher throughput than a system with a
+centralized sequencer service.
+We utilize the programmability of the metadata service to define a new policy
+for handling capabilities that controls the amount of time that clients are
+able to cache the sequencer resource. This allows an administrator or
+application to control the trade-off between latency and throughput beyond the
+standard best-effort policy that is present in Ceph by default.
+
+In Section~\cite{evaluation} we quantify the trade-offs of throughput and
+latency for an approach based on a round-robin batching mode, and compare this
+mode to one in which the metadata server mediates access to the sequencer
+state when in it is being shared among multiple clients.
+
+{\bf Balancing policies.}
+As opposed to the batching mode for controlling access to the sequencer
+resource, more predictable latency can be achieved by treating the sequencer
+inode as a shared non-cacheable resource, forcing clients to make a round-trip
+to the metadata service. However, the shared nature of the metadata service
+may prevent the sequencer from achieving maximum throughput. To provide a
+solution to this issue we have taken advantage of the programmability of the
+metadata service load balancing to construct a service-specific load balancing
+policy. As opposed to a generic balancing policy that may strive for uniform
+load distribution, a ZLog-specific policy may utilize knowledge of inode types
+to migrate the sequencer service to provisioned hardware during periods of
+contention or high demand.
+
+\subsubsection{Storage Interface}
+
+The storage interface is a critical component in the CORFU design. Clients
+independently map log positions that they have obtained from a sequencer
+service (described in detail in the next seciton) onto storage devices, while
+storage devices provide an intelligent \emph{write}-once, random \emph{read}
+interface for accessing log entries. The key to correctness in CORFU lies with
+the enforcement of up-to-date epoch tags on client requests; requests tagged
+with out-of-date epoch values are rejected, and clients are expected to
+request a new tail from the sequencer after refreshing state from an auxillary
+service.  This mechanism forms the basis for sequencer recovery.
+
 In order to repopulate the cached tail value during recovery of a sequencer,
@@
 auxillary service such as PAXOS.
-
-\subsubsection{Sequencer}
-
-We implement the sequencer service in Malacology as an application-specific
-file type that associates a small amount of metadata, including the 64-bit log
-tail position, to the inode state managed by the Ceph metadata service.  Since
-a sequencer is associated with a unique log, this approach has the added
-benefit of allowing the metadata service to also handle naming, by
-representing each log service instance within the standard POSIX hierarchical
-namespace.
-
-{\bf Physical design.}
-The next component of our design is the low-level configuration of the
-sequencer inode type that achieves the correct semantics.  Initially we sought
-to directly model the sequencer service in Ceph as a non-exclusive,
-non-cacheable resource, forcing clients to perform a round-trip access to the
-resource at the authoratiave metadata server for the inode representing the
-sequencer.  Interestingly, we found that the capapbility system in Ceph uses a
-strategy to reduce metadata service load by allowing multiple clients that
-open a shared file to temporarily cache the inode resource, resulting in a
-round-robin, best-effort batching behavior. When a single client is accessing
-the sequencer resource it is able to cache state locally without interruption.
-
-While unexpected, this discovery allowed us to explore an implementation
-strategy that we had not previously considered. In particular, for bursty
-clients and clients that can tolerate high latency, this mode of operation may
-allow a system to achieve much higher throughput than a system with a
-centralized sequencer service.
-
-In Section we quantify the trade-offs of throughput and latency for an
-approach based on a round-robin batching mode, and compare this mode to one in
-which the metadata server mediates access to the sequencer state when in it is
-being shared among multiple clients.
-
-{\bf Capability policies.}
-We use capabilitity policies in the metadata service to tune the amount of
-time that clients may cache the sequencer resource. This allows us to control
-the trade-off between latency and throughput beyond the standard best-effort
-policy.
-
-{\bf Balancing policies.}
-While the metadata service is a shared resource, authority over metadata
-resources can be physically moved between servers using a dynamic, ZLog-aware
-policy expressed in Mantle. For instance, one or more metadata servers may be
-dedicated to supporting ZLog sequencer load. When the system load is low all
-resources are spread evenly, but non-ZLog resources may be migrated to
-best-effort servers when sequencer performance suffers.
-
-{\bf Recovery.}
-What?
 

### 2016-05-10 paper/paper.tex

@@
 
+\section{Challenges}
+\label{sec:challenges}
+
+The challenges of introducing programmability into existing storage systems: (1) Programmability while keeping storage system available, (2) Finding and factoring out hard-wired policies and optimizations, (3) Bugs that were hidden because mechanisms were only exercised according to hard-wired policies and not in their full generality they were thought of -- programmability into a storage system turned out to be a great debugging tool.
+
 \section{Malacology}

### 2016-05-10 paper/paper.tex
Adds Malacology subsection to Cluster State Mgmt

@@
 failures, when they occur. In the case of Ceph, a consistent view of cluster 
-state among server daemons and clients is critical
-to provide strong consistency guarantees to clients. Ceph maintains cluster
-state information in per-subsystem ``maps'' (e.g. OSD map, MDS map) that
-record membership and status information. A PAXOS-based monitoring service is
-responsible for integrating state changes into cluster maps, and responding to
-requests from out-of-date clients. As a fundamental building block of many
-system designs, consensus abstractions such as PAXOS are a common technique
-for maintaining consistent data versions, and are a useful system to expose.
-
-While Ceph does not expose this service directly, it does allow limited use by
-clients through a key-value service intended for low-volume, small size
-configuration information. Unfortunately, while a key-value service is
+state among server daemons and clients is critical to provide strong consistency 
+guarantees to clients. Ceph maintains cluster state information in per-subsystem 
+``maps'' (e.g. OSD map, MDS map) that record membership and status information. 
+A PAXOS-based monitoring service (termed, the Monitor) is responsible for 
+integrating state changes into cluster maps, responding to requests from 
+out-of-date clients and synchronizing members of the cluster whenever there is a 
+change in a map so that they all observe the same system state. As a fundamental 
+building block of many system designs, consensus abstractions such as PAXOS are 
+a common technique for maintaining consistent data versions, and are a useful 
+system to expose.
+
+The default behavior of the monitor can be seen as a PAXOS-based notification 
+system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing 
+clients to identify when new values (termed epochs in Ceph) are associated to 
+given maps. While Ceph does not expose this service directly, it does allow 
+limited use by clients through a key-value service intended for low-volume, 
+small size read-only configuration information. While a key-value service is 
 generally useful, it doesn't provide many of the useful services hidden within 
 the monitoring framework that would be required for applications with more 
-demanding requirements.
-
-\paragraph*{Malacology:} Malacology exposes the monitoring service in two ways. 
-First, Malacology APIs for object storage (\S\ref{active-storage}) and the MDS 
-(\S\ref{malacology:mds}) have functions available to them that allow them to 
-store application-specific data in the highly-available, consistent store 
-associated to the Ceph monitors. As mentioned
-Map modification for propogation of state. We allow applications to augment
-these data structures with application-specific metadata. In order to keep the
-size of these data structures small application state may be stored elsewhere
-and the references stored within the data structure.
+demanding requirements, such as creating arbitrary maps or associating 
+application-specific logic that can be executed for particular maps (or values 
+in a map).
+
+\paragraph*{Malacology:} Malacology exposes Ceph's PAXOS-based monitoring 
+service in two ways. First, a Lua VM is embedded in the monitors that implements 
+a generic API for creating maps and associating arbitrary data to them. This API 
+also allows clients to specify small snippets of Lua code that can be executed 
+on monitors. As a consequence of this, applications can define simple but useful 
+domain-specific logic to the PAXOS quorum, such as authorization control (just 
+specific clients can write new values) or to trigger actions based on specific 
+values (e.g. sanitize values). Since the monitor is intended to be out of 
+high-performance I/O paths, a general guideline is to make use of this 
+functionality infrequently and to assign small values to maps. Malacology itself 
+makes use of this functionality to version object interfaces defined in OSDs 
+(\S\ref{active-storage}) and load balancing policies in MDSs 
+(\S\ref{malacology:mds}).
 

### 2016-05-10 paper/paper.tex
Missed this small edit

@@
 \paragraph*{Malacology:} Malacology exposes Ceph's PAXOS-based monitoring 
-service in two ways. First, a Lua VM is embedded in the monitors that implements 
-a generic API for creating maps and associating arbitrary data to them. This API 
-also allows clients to specify small snippets of Lua code that can be executed 
-on monitors. As a consequence of this, applications can define simple but useful 
+service by embedding a Lua VM in the monitors that implements a generic API for 
+creating maps and associating arbitrary data to them. This API also allows 
+clients to specify small snippets of Lua code that can be executed on monitors. 
+As a consequence of this, applications can define simple but useful 
 domain-specific logic to the PAXOS quorum, such as authorization control (just 

### 2016-05-10 paper/paper.tex

@@
 
-WE NEED SOMETHING HERE DESCRIBING WHY AN INODE IS A GOOD ABSTRACTION
-
-\begin{figure}[h]
+\begin{figure}[t!]
 \centering
@@
 of 1 sequencer when an additional sequencer is running in the background
-generating load.} 
-\label{fig:mantle-seq-thruput.png}
+generating load.}\label{fig:mantle-seq-thruput}
 \end{figure}
@@
 
-\begin{figure}[h]
+\begin{figure}[]
 \begin{lstlisting}
@@
 
-Figure~\ref{fig:mantle-seq-thruput.png} gives a macro-level view of the
-performance of the Ceph components exposed by Malacology. Each curve shows the
-throughput from the perspective of one sequencer ({\it y} axis) over time ({\it
-x} axis). The ``Sequencer Baseline'' curve is one sequencer serving 7 clients.
-The sequencer operates on an MDS cluster where balancing is turned off and
-there are no other workloads running. ``2 Seq; No Load Balancing" adds another
-sequencer of equal load running in the background. The performance of the
-background sequencer is omitted. ``2 Seq; Mantle Load Balancing" uses a 3 MDS
-node cluster to provide the sequencer services, again with a background
-sequencer service. Finally, the ``Sequencer Recovery" curve shows the
-performance of the sequencer in a multi-MDS cluster with 2 standby MDS nodes.
-At 60 seconds, MDS 0 is killed.
+\begin{figure}[t!]
+\centering
+\includegraphics{figures/mantle-cpu-utilization-annotate.png}
+\caption{The sequencer using Mantle in Figure~\ref{fig:mantle-seq-thruput.png}
+migrates to a new MDS immediately because its uses a policy designed for bursty
+workloads. This is a trace of the same workload running twice: once with a
+single MDS and once with Mantle.}
+\label{fig:mantle-cpu-utilization}
+\end{figure}
+
+Figure~\ref{fig:mantle-seq-thruput} gives a macro-level view of the performance
+of the Ceph components exposed by Malacology. Each curve shows the throughput
+from the perspective of one sequencer (\(y\) axis) over time (\(x\) axis).  The
+``Sequencer Baseline'' curve is one sequencer serving 7 clients.  The sequencer
+operates on an MDS cluster where balancing is turned off and there are no other
+workloads running. ``No Load Balancing" adds another sequencer of equal load
+running in the background. The performance of the background sequencer is
+omitted. The last two policies, ``Sequencer Balancing Policy" and
+``Conservative Balancing Policy", show the performance of the sequencer (with a
+background sequencer running) when using Mantle with a 3 MDS node cluster.
+These policies differ by the threshold at which they start balancing:
+``Sequencer Balancing Policy" starts balancing immediately and ``Conservative
+Balancing Policy" waits for 120 seconds.
+
+Figure~\ref{fig:mantle-cpu-utilization} gives a time series graph to show the
+behavior of the same workload, 2 sequencers, with balancing turned on and off.
+The curves are the throughputs (\(y\) axis) over time (\(x\) axis) and the graph
+has been partitioned into the two cluster setups. The CPU utilization is
+confined to one core the MDS is not multithreaded, as it has been shown that
+metadata performance does not improve with multiple
+threads~\cite{konstantinos:pdsw2014-lustre-metadata}.
 
@@
 balancing, cache coherence, and capabilities subsystems of the MDS. Applying
-Mantle's load balancing policy shows 2\(\times\) performance improvent and x\%
-of the baseline best-case performance.
+Mantle's load balancing policy shows 2.02\(\times\) performance improvement and
+16\% degradation in performance over the baseline, even though the baseline has
+half as many clients.
 
@@
 
-- Describe the graph
-
-- We get recovery and load balancing with minimal implementation
-
-\subsection{Load Balancing}
-
-\begin{figure}[h]
-\centering
-\includegraphics{figures/mantle-cpu-utilization-annotate.png}
-\caption{The sequencer using Mantle in Figure~\ref{fig:mantle-seq-thruput.png} migrates to a new MDS immediately because its uses a policy designed for bursty workloads.
-This is a trace of the same workload running twice: once with a single MDS and once with Mantle.}
-\label{fig:mantle-cpu-utilization}
-\end{figure}
-
-- Sequencer specific policy
-
-- Describe the graph
-
-- We get tunable policies
-
-\subsection{Recovery}
+\begin{figure}[t]
+\centering
+\includegraphics{figures/recovery-seq-thruput.png}
+\caption{The sequencer using Mantle also inherits all the recovery protocols of
+the Ceph metadata cluster. The throughput of the sequencer drops when we
+terminal MDS 0. The performance quickly resumes and the sequencer is only down
+for 15 seconds.}
+\label{fig:recovery-seq-thruput}
+\end{figure}
+
+This experiment shows how a sequencer handing out tokens can use the Ceph MDS
+recovery protocols to recover. For this workload, two sequencers are running on
+a metadata cluster {\it without} multiple active MDS nodes. In this cluster,
+MDS 1 and MDS 2 are in ``standby" mode, meaning one of them will step in should
+the head MDS 0 fail. At the 60 second mark, we manually kill MDS 0.
+
+Figure~\ref{fig:recovery-seq-thruput} shows the sequencer throughput (\(y\)
+axis) over time (\(x\) axis). When MDS 0 dies, the MON cluster notices and
+quickly assigns a new MDS to handle the namespace. This is when MDS 1 takes
+over the sequencer duties and starts handing out tokens.
+
+\textbf{The Malacology Principle:} Re-uses the Ceph metadata protocols for
+failure detection and recovery. The sequencer recovers in 15 seconds on a new
+MDS and it maintains the same throughput that it had before the failure.
 

### 2016-05-10 paper/paper.tex
One more challenge in "Challenges"

--- 2016-05-10 paper/paper.tex
@@
 
-The challenges of introducing programmability into existing storage systems: (1) Programmability while keeping storage system available, (2) Finding and factoring out hard-wired policies and optimizations, (3) Bugs that were hidden because mechanisms were only exercised according to hard-wired policies and not in their full generality they were thought of -- programmability into a storage system turned out to be a great debugging tool.
+The challenges of introducing programmability into existing storage systems: (1) Programmability while keeping storage system available, (2) Finding and factoring out hard-wired policies and optimizations, (3) Bugs that were hidden because mechanisms were only exercised according to hard-wired policies and not in their full generality they were thought of -- programmability into a storage system turned out to be a great debugging tool, (4) version control and propagation of updates throughout the system without interrupting production workloads, (5) ...
 
@@
 failures, when they occur. In the case of Ceph, a consistent view of cluster 
-state among server daemons and clients is critical to provide strong consistency 
-guarantees to clients. Ceph maintains cluster state information in per-subsystem 
-``maps'' (e.g. OSD map, MDS map) that record membership and status information. 
-A PAXOS-based monitoring service (termed, the Monitor) is responsible for 
-integrating state changes into cluster maps, responding to requests from 
-out-of-date clients and synchronizing members of the cluster whenever there is a 
-change in a map so that they all observe the same system state. As a fundamental 
-building block of many system designs, consensus abstractions such as PAXOS are 
-a common technique for maintaining consistent data versions, and are a useful 
-system to expose.
-
-The default behavior of the monitor can be seen as a PAXOS-based notification 
-system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing 
-clients to identify when new values (termed epochs in Ceph) are associated to 
-given maps. While Ceph does not expose this service directly, it does allow 
-limited use by clients through a key-value service intended for low-volume, 
-small size read-only configuration information. While a key-value service is 
+state among server daemons and clients is critical
+to provide strong consistency guarantees to clients. Ceph maintains cluster
+state information in per-subsystem ``maps'' (e.g. OSD map, MDS map) that
+record membership and status information. A PAXOS-based monitoring service is
+responsible for integrating state changes into cluster maps, and responding to
+requests from out-of-date clients. As a fundamental building block of many
+system designs, consensus abstractions such as PAXOS are a common technique
+for maintaining consistent data versions, and are a useful system to expose.
+
+While Ceph does not expose this service directly, it does allow limited use by
+clients through a key-value service intended for low-volume, small size
+configuration information. Unfortunately, while a key-value service is
 generally useful, it doesn't provide many of the useful services hidden within
 the monitoring framework that would be required for applications with more
-demanding requirements, such as creating arbitrary maps or associating 
-application-specific logic that can be executed for particular maps (or values 
-in a map).
-
-\paragraph*{Malacology:} Malacology exposes Ceph's PAXOS-based monitoring 
-service by embedding a Lua VM in the monitors that implements a generic API for 
-creating maps and associating arbitrary data to them. This API also allows 
-clients to specify small snippets of Lua code that can be executed on monitors. 
-As a consequence of this, applications can define simple but useful 
-domain-specific logic to the PAXOS quorum, such as authorization control (just 
-specific clients can write new values) or to trigger actions based on specific 
-values (e.g. sanitize values). Since the monitor is intended to be out of 
-high-performance I/O paths, a general guideline is to make use of this 
-functionality infrequently and to assign small values to maps. Malacology itself 
-makes use of this functionality to version object interfaces defined in OSDs 
-(\S\ref{active-storage}) and load balancing policies in MDSs 
-(\S\ref{malacology:mds}).
+demanding requirements.
+
+\paragraph*{Malacology:} Malacology exposes the monitoring service in two ways. 
+First, Malacology APIs for object storage (\S\ref{active-storage}) and the MDS 
+(\S\ref{malacology:mds}) have functions available to them that allow them to 
+store application-specific data in the highly-available, consistent store 
+associated to the Ceph monitors. As mentioned
+Map modification for propogation of state. We allow applications to augment
+these data structures with application-specific metadata. In order to keep the
+size of these data structures small application state may be stored elsewhere
+and the references stored within the data structure.
 
@@
 paper, we'll briefly describe the custom storage device interface, sequencer
-service, and recovery protocol, and how these services are instantiated in the
-Malacology framework.
-
-\subsubsection{Sequencer}
-
+service, and recovery protocol that form the core components of CORFU that we
+implement within Malacology:
+
+{\bf Sequencer.}
 High-performance in CORFU is achived using a sequencer service that assigns
@@
 
-We implement the sequencer service in Malacology as an application-specific
-file type that associates a small amount of metadata, including the 64-bit log
-tail position, to the inode state managed by the Ceph metadata service. Since
-a sequencer is associated with a unique log, this approach has the added
-benefit of allowing the metadata service to also handle naming, by
-representing each log service instance within the standard POSIX hierarchical
-namespace.
-
-{\bf Sequencer interface.} The sequencer resource supports the ability to
-\emph{read()} the current tail value as well get the \emph{next()} position in
-the log which also atomically increments the tail position. We implement this
-functionality as methods on ZLog inode resources. The primary challenge in
-mapping the sequencer resource to the metadata service is handling
-serialization correctly.
-
-Initially we sought to directly model the sequencer service in Ceph as a
-non-exclusive, non-cacheable resource, forcing clients to perform a round-trip
-access to the resource at the authoratiave metadata server for the sequencer
-inode.  Interestingly, we found that the capapbility system in Ceph uses a
-strategy to reduce metadata service load by allowing multiple clients that
-open a shared file to temporarily cache the inode resource, resulting in a
-round-robin, best-effort batching behavior. When a single client is accessing
-the sequencer resource it is able to cache state locally without interruption.
-
-While unexpected, this discovery allowed us to explore an implementation
-strategy that we had not previously considered. In particular, for bursty
-clients, and clients that can tolerate increased latency, this mode of operation may
-allow a system to achieve much higher throughput than a system with a
-centralized sequencer service.
-We utilize the programmability of the metadata service to define a new policy
-for handling capabilities that controls the amount of time that clients are
-able to cache the sequencer resource. This allows an administrator or
-application to control the trade-off between latency and throughput beyond the
-standard best-effort policy that is present in Ceph by default.
-
-In Section~\cite{evaluation} we quantify the trade-offs of throughput and
-latency for an approach based on a round-robin batching mode, and compare this
-mode to one in which the metadata server mediates access to the sequencer
-state when in it is being shared among multiple clients.
-
-{\bf Balancing policies.}
-As opposed to the batching mode for controlling access to the sequencer
-resource, more predictable latency can be achieved by treating the sequencer
-inode as a shared non-cacheable resource, forcing clients to make a round-trip
-to the metadata service. However, the shared nature of the metadata service
-may prevent the sequencer from achieving maximum throughput. To provide a
-solution to this issue we have taken advantage of the programmability of the
-metadata service load balancing to construct a service-specific load balancing
-policy. As opposed to a generic balancing policy that may strive for uniform
-load distribution, a ZLog-specific policy may utilize knowledge of inode types
-to migrate the sequencer service to provisioned hardware during periods of
-contention or high demand.
-
-\subsubsection{Storage Interface}
-
-The storage interface is a critical component in the CORFU design. Clients
-independently map log positions that they have obtained from a sequencer
-service (described in detail in the next seciton) onto storage devices, while
-storage devices provide an intelligent \emph{write}-once, random \emph{read}
-interface for accessing log entries. The key to correctness in CORFU lies with
-the enforcement of up-to-date epoch tags on client requests; requests tagged
-with out-of-date epoch values are rejected, and clients are expected to
-request a new tail from the sequencer after refreshing state from an auxillary
-service.  This mechanism forms the basis for sequencer recovery.
-
+{\bf Storage Interface.} The storage interface is a critical component in the
+CORFU design. Clients independently map log positions onto storage devices,
+while storage devices provide an intelligent \emph{write}-once, random
+\emph{read} interface for accessing log entries. The key to correctness in
+CORFU lies with the enforcement of up-to-date epoch tags on client requests;
+requests tagged with out-of-date epoch values are rejected, and clients are
+expected to request a new tail from the sequencer after refreshing state from
+an auxillary service. This mechanism forms the basis for sequencer recovery.
+
+{\bf Sequencer Recovery.}
 In order to repopulate the cached tail value during recovery of a sequencer,
@@
 auxillary service such as PAXOS.
+
+\subsubsection{Sequencer}
+
+We implement the sequencer service in Malacology as an application-specific
+file type that associates a small amount of metadata, including the 64-bit log
+tail position, to the inode state managed by the Ceph metadata service.  Since
+a sequencer is associated with a unique log, this approach has the added
+benefit of allowing the metadata service to also handle naming, by
+representing each log service instance within the standard POSIX hierarchical
+namespace.
+
+{\bf Physical design.}
+The next component of our design is the low-level configuration of the
+sequencer inode type that achieves the correct semantics.  Initially we sought
+to directly model the sequencer service in Ceph as a non-exclusive,
+non-cacheable resource, forcing clients to perform a round-trip access to the
+resource at the authoratiave metadata server for the inode representing the
+sequencer.  Interestingly, we found that the capapbility system in Ceph uses a
+strategy to reduce metadata service load by allowing multiple clients that
+open a shared file to temporarily cache the inode resource, resulting in a
+round-robin, best-effort batching behavior. When a single client is accessing
+the sequencer resource it is able to cache state locally without interruption.
+
+While unexpected, this discovery allowed us to explore an implementation
+strategy that we had not previously considered. In particular, for bursty
+clients and clients that can tolerate high latency, this mode of operation may
+allow a system to achieve much higher throughput than a system with a
+centralized sequencer service.
+
+In Section we quantify the trade-offs of throughput and latency for an
+approach based on a round-robin batching mode, and compare this mode to one in
+which the metadata server mediates access to the sequencer state when in it is
+being shared among multiple clients.
+
+{\bf Capability policies.}
+We use capabilitity policies in the metadata service to tune the amount of
+time that clients may cache the sequencer resource. This allows us to control
+the trade-off between latency and throughput beyond the standard best-effort
+policy.
+
+{\bf Balancing policies.}
+While the metadata service is a shared resource, authority over metadata
+resources can be physically moved between servers using a dynamic, ZLog-aware
+policy expressed in Mantle. For instance, one or more metadata servers may be
+dedicated to supporting ZLog sequencer load. When the system load is low all
+resources are spread evenly, but non-ZLog resources may be migrated to
+best-effort servers when sequencer performance suffers.
+
+{\bf Recovery.}
+What?
 

### 2016-05-10 paper/paper.tex
Adds 'Policy Durability' section
This is short

@@
 
--
-
--
-
--
+While storing the code snippets that define the policies in the MDS map is 
+possible, it is not recommended as it is an expensive operation (a PAXOS voting 
+round). Thus, Malacology only stores policies IDs which correspond to an object 
+in a system-maintained pool. When MDSs read a policy, they first check latest 
+version from the MDS map, and then dereference the pointer by reading the 
+corresponding object in RADOS.
 

### 2016-05-10 paper/paper.tex
iface propgation updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-05-10 paper/paper.tex
@@
 
-In Section~\ref{} we described an interface management service that is
-responsible for the durable and consistent cluster-wide distribution of
-interface versions. We evaluate the costs of such a service by embedding
-interface definitions in a Ceph cluster membership data structure that is
-managed by a Paxos instance, and distributed using a scalable, peer-to-peer
-gossip protocol. We assume an architecture that consists of a light-weight
-interface that enforces the semantics of interface management resulting in one
-or more interface updates that must be consistent across the cluster. The
-following experiments focus on the cost of achieving distribution of consistent
-a interface version.
+In Section~\ref{consistencyversioning-of-cluster-state} we described an
+interface management service that is responsible for the durable and
+consistent cluster-wide distribution of interface versions. We evaluate the
+costs of such a service by embedding interface definitions in a Ceph cluster
+membership data structure that is managed by a PAXOS instance, and distributed
+using a scalable, peer-to-peer gossip protocol that is built into Ceph. We
+assume an architecture that consists of a light-weight interface that enforces
+the semantics of interface management resulting in one or more interface
+updates that must be consistent across the cluster. The following experiments
+focus on the cost of achieving distribution of consistent a interface version.
 
@@
 \centering
-\includegraphics{figures/iface-update-delay.pdf}
+\includegraphics[trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
 \caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
@@
 
-Figure~\ref{fig:propdelay} shows the CDF of the latency of distributing and
-installing an interface update on all OSDs in the cluster. The latency is
-defined as the elapsed time following the Paxos proposal for an interface
-update until each OSDs makes the update live (he cost of the Paxos proposal is
+Figure~\ref{fig:propdelay} shows the CDF of the latency of installing and
+distributing an interface update on all OSDs in the cluster. The latency is
+defined as the elapsed time following the PAXOS proposal for an interface
+update until each OSD makes the update live (the cost of the PAXOS proposal is
 configurable and is discussed below). The latency measurements were taken on
-the nodes running Ceph server daemons (the Paxos servers and OSDs nodes), and
-thus excludes the client round-trip cost. In each of the experiments 1000
-interface updates were observed.
+the nodes running Ceph server daemons, and thus exclude the client round-trip
+cost. In each of the experiments 1000 interface updates were observed.
 
@@
 (RAM)" a cluster of 120 OSDs (10 OSDs x 12 servers) using an in-memory data
-store were deployed, showing a latency of less than 54 ms with a probability of
-90\% and a worst case latency of 194 ms. The curve labeled "3 OSD (HDD)" is a
-smaller cluster where each of the three OSDs uses a spinning disk as the object
-store that saves a copy of the cluster map ({\bf is this I/O before its made
-live?}), and shows a latency of 37 ms with a 90\% probability and a worst case
-latency of 255 ms. Since the cost of a disk I/O is larger than the same I/O
-executed against the in-memory device, the predominant cost in a lightly loaded
-cluster appears to be the number of OSDs.
+store were deployed, showing a latency of less than 54 ms with a probability
+of 90\% and a worst case latency of 194 ms. The curve labeled "3 OSD (HDD)" is
+a smaller cluster where each of the three OSDs uses a spinning disk as the
+object store that saves a copy of the cluster map, and shows a latency of 37
+ms with a 90\% probability and a worst case latency of 255 ms. Since the cost
+of a disk I/O is larger than the same I/O executed against the in-memory
+device, the predominant cost in a lightly loaded cluster appears to be the
+number of OSDs.
 
@@
 of interface updates, includes the network round-trip to the interface
-management service, the Paxos commit protocol, and other factors such as system
-load. By default Paxos proposals occur periodically with a 1 second interval in
-order to accumulate updates. In a minimum realistic quorum of 3 monitors using
+management service, the PAXOS commit protocol, and other factors such as system
+load. By default PAXOS proposals occur periodically with a 1 second interval in
+order to accumulate updates. In a minimum, realistic quorum of 3 monitors using
 HDD-based storage, we were able to decrease this interval to an average of 222

### 2016-05-10 paper/paper.tex
mds
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 \paragraph*{Malacology:} Malacology exposes Ceph's PAXOS-based monitoring
-service by embedding a Lua VM in the monitors that implements a generic API for 
-creating maps and associating arbitrary data to them. This API also allows 
-clients to specify small snippets of Lua code that can be executed on monitors. 
+service by embedding a Lua VM in the monitors that implements a generic API
+for creating dynamic interfaces around existing subsystems that may associated
+additional state with a subsystem map.
 As a consequence of this, applications can define simple but useful 
@@
 clients are properly synchronousized with the latest interfaces.
-
-{\bf Recovery.} Application specific recovery is critical.
 

### 2016-05-10 paper/paper.tex
Quick review to Eval section

@@
 
-\subsection{Policy Durability}
+\subsection{Making Policies Durable}
 
@@
 We have conducted our evaluation on CloudLab as well as our own 20 node cluster.
-We follow the ``Popper Convention"~\cite{} so as to try and make our experiments as reproducible as possible.
+We follow the ``Popper Convention"~\cite{anonymous} so as to try and make our experiments as reproducible as possible.
 We have a GitHub repository with our source code,  experimental environment, configuration parameters, and workloads in a GitHub repository.
@@
 
+\textbf{The Malacology Approach:} Re-use the capability management features of the MDS.
 While the batching mechanism can achieve high-throughput, it is not appropriate
@@
 
+\textbf{The Malacology Approach:} Re-use Ceph's Monitor functionality to 
+propagate application-specific state that needs to be highly-available and 
+strongly consistent.
 The real cost of interface management, in addition to cluster-wide propagation
@@
 
-\subsection{MDS as a Sequencer}
+\subsection{MDS as a ZLog Sequencer}
 
@@
 
-\textbf{The Malacology Principle:} Re-uses the Ceph metadata migration, load
+\textbf{The Malacology Approach:} Re-use the Ceph metadata migration, load
 balancing, cache coherence, and capabilities subsystems of the MDS. Applying
@@
 
-\textbf{The Malacology Principle:} Re-uses the Ceph metadata protocols for
+\textbf{The Malacology Approach:} Re-use the Ceph metadata protocols for
 failure detection and recovery. The sequencer recovers in 15 seconds on a new

### 2016-05-10 paper/paper.tex
zlog intro
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-05-10 paper/paper.tex
@@
 
+We evaluate the mapping of ZLog onto Malacology by examining the costs of
+utilizing the metadata service as a sequencer, and propogating interface
+changes throughout the cluster using the monitoring service. We have omitted
+evaluation of the overhead costs of adding object interfaces themselves due to
+lack of space, and the fact that they are heavily tested and used already in
+production with acceptable performance, introducing a small constant overhead
+per operation but otherwise does not affect scalability.
+
 \subsubsection{Sequencer}

### 2016-05-10 paper/paper.tex
paper: fix Metadata Services section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-05-10 paper/paper.tex
@@
 \paragraph*{Malacology:} Malacology exposes Ceph's PAXOS-based monitoring 
-service by embedding a Lua VM in the monitors that implements a generic API
-for creating dynamic interfaces around existing subsystems that may associated
-additional state with a subsystem map.
+service by embedding a Lua VM in the monitors that implements a generic API for 
+creating maps and associating arbitrary data to them. This API also allows 
+clients to specify small snippets of Lua code that can be executed on monitors. 
 As a consequence of this, applications can define simple but useful 
@@
 
+{\bf Recovery.} Application specific recovery is critical.
+
 \subsection{Distributed Metadata Services}
@@
 The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction~\cite{cephfs}. Distributed file systems are complex, but
-fundamentally provide hierarchical indexing and distributed locking services.
-In Ceph, the locking service implements a capability-based system that express
-what data and metadata clients are allowed to access as well as what state
-they may cache and modify locally. While designed for a fixed file
-abstraction, indexing, locking, and caching are all common services that are
-useful to a broad spectrum of applications.
-
-Ceph addresses the challenge of balancing load across the metadata cluster
-using load balancing policies that compute balancing metrics based on system
-state (e.g. CPU and memory utilization) and statistics collected by the
-cluster (e.g. the popularity of an inode). Distributed applications that share
-centralized resources (e.g. a database or directory) face similar challenges
-which are often solved using application-specific sharding.
-
-\paragraph*{Malacology:}
-Malacogy adds interface hooks in strategic spots in the metadata service to
-control load balancing, capability subsystems, and for defining new file
-types allowing the indexing service to be repurposed for other types of
-named resources.
-
-{\bf Load balancing.}
-For CephFS, designing a general POSIX metadata load balancer is difficult
-becuase the workloads are diverse and quickly changing. With CephFS, it is
-especially complicated because the system has variable sized migration units
-(i.e. file system namespace subtrees), which can be shipped anywhere (i.e., to
-any MDS of any capacity) at any time for any reason. Furthermore, the original
-balancer was designed with hard-coded policies and tunables. Both angles were
-difficult to understand. It is understandable why it was implemented this way
--- it is difficult to pick a one-size-fits-all policy, system administrators
-may want to partition load based on SLAs, or applications are better equipped
-to make these decisions, etc. -- but we choose instead to expose a balancing
-API at strategic points in the balancing logic using Lua hooks.
-
-These problems are general, distributed systems load balancing problems and
-share many of the same properties as the POSIX metadata
-wall~\cite{alam:pdsw2011-metadata-scaling,ghemawat:sosp2003-gfs,hildebrand:msst2005-pnfs,weil_ceph_2006,welch:fast2008-panasas,shvachko:login2012-hdfs-scalability}.
-Many distributed file systems decouple metadata and data I/O so that these
-services can scale independently. Despite this optimization, scaling the
-metadata services is still difficult because metadata accesses impose small
-and frequent requests on the underlying storage
-system~\cite{roselli:atec2000-FS-workloads}. Many techniques for designing the
-metadata services have been proposed to accommodate this workload:
-Lustre~\cite{konstantinos:pdsw2014-lustre-metadata},
-GFS~\cite{ghemawat:sosp2003-gfs}, and
-HDFS~\cite{shvachko:login2012-hdfs-scalability} keep all metadata on one
-server; GIGA+~\cite{patil:fast2011-giga}, IndexFS~\cite{ren:sc2014-indexfs},
-Lazy Hybrid~\cite{brandt:msst2003-lh}, GPFS~\cite{schmuck:fast2002-gpfs}, and
-pNFS~\cite{hildebrand:supercomputing2006-pNFS} hash the file system namespace
-across a dedicated cluster and cache metadata;
-Panasas~\cite{welch:fast2008-panasas} and
-CephFS~\cite{weil:sc2004-dyn-metadata} use subtree partitioning to carve up
-the namespace and distribute it across a cluster.
-
-CephFS moves units of the namespace called directory fragments and assigns
-them to servers using  its own metadata load balancer. It then uses a
-hard-coded policy to balance these fragments using a secularized load metric
-based on CPU, workload, and file system operation
-metrics~\cite{sevilla:sc15-mantle}.
-
-Both the object and balancer interfaces define functions and attach them
-to the core Lua sandbox. For example, the Lua object storage interface
-class attaches data, extended attribute, object map, and object version
-functions, to the Lua core functions, as shown in Figure
-\ref{fig:cls-osd-mds}. The advantages of this is that we avoid
-duplicating code, we provide a framework for putting Lua code in other
-parts of the system, and we remove components and APIs that are too
-integrated with the OSD.
-
-To facilitate the use of Lua in both the MDS and OSD, we modularize the LuaVM
-in a core sandbox wrapper and link against it, as shown in
-Figure~\ref{fig:cls-osd-mds}. This core Lua wrapper contains just the
-dependencies, symbols, and functions, needed to run the Lua VM and strips away
-all the functionality inherent to object classes (like placement group
-filters, cryptography functions, object metadata,  object data, object
-extended attributes, and object versions). Some functions of Ceph's more
-helpful subsystems, like the logging facilities for the daemons, were re-used
-in the MDS.  With this scheme, The OSD will dynamically open the shared
-library created by Lua object interface shared library while the MDS will
-dynamically open the Lua balancer interface shared library; both shared
-libraries will the Lua core.
-
-{\bf Capabilities.}
-The metadata service also manages sessions with its clients. This gives both
-subsystems (MDSs and clients) the ability cache data and grab locks. The
-metadata service is responsible for responding to client requests for
-capabilities by first revoking caps from other clients. Currently clients
-voluntarily release their capabilities, and the MDS maintains a queue of
-requests.  Both clients and MDSs use a best-effort approach to revoking or
-releasing capabilities.
-
-In our current we retain the best-effort, voluntary release of caps within
-clients, but generalize the queueing policies within the MDS to allow for
+system abstraction~\cite{weil:sc2004-dyn-metadata}. In general, distributed
+file systems protect resources by providing hierarchical indexing and
+distributed locking services.  In Ceph, the locking service implements a
+capability-based system that express what data and metadata clients are allowed
+to access as well as what state they may cache and modify locally. While
+designed for a fixed file abstraction, indexing, locking, and caching are all
+common services that are useful to a broad spectrum of applications.
+
+Ceph also addresses the challenge of balancing metadata load using load
+balancing policies that compute metrics based on system state (e.g.  CPU and
+memory utilization) and statistics collected by the cluster (e.g. the
+popularity of an inode). Ceph uses dynamic subtree partitioning to move
+variable sized namespace subtrees. These units can be shipped anywhere (i.e.,
+to any MDS of any capacity) at any time for any reason. The original balancer
+was designed with hard-coded policies and tunables.  Distributed applications
+that share centralized resources (e.g. a database or directory) face similar
+challenges which are often solved using application-specific sharding.
+
+%The metadata cluster moves units of the namespace called directory fragments
+%and assigns them to servers using its own metadata load balancer. It then uses
+%a hard-coded policy to balance these fragments using a scalarized load metric
+%based on CPU, workload, and file system operation metrics. Although the CephFS
+%balancer has been shown to be complicated, its complexity is justifiable.
+%Having the flexibility to choose what to move, where to move it, and how much
+%to move is very powerful. The problem is that it is difficult to pick a
+%one-size-fits-all policy, system administrators may want to partition load
+%based on SLAs, or applications are better equipped to make these decisions,
+%etc. -- but we choose instead to expose a balancing API at strategic points in
+%the balancing logic using Lua hooks. ~\cite{sevilla:sc15-mantle}.
+
+\paragraph*{Malacology:} New interface hooks are added in strategic spots in
+the metadata service to control load balancing and capabilities, and for
+defining new file types allowing the indexing service to be repurposed for
+other types of named resources.
+
+{\bf Load balancing.} Mantle~\cite{sevilla:sc15-mantle}, a programmable
+metadata load balancer, has been re-implemented on Malacology so that its
+infrastructure can enjoy all the properties of other Ceph components. The load
+balancing logic has stayed largely the same but the infrastructure has been
+improved to safely load, manage, and persist its load balancing policies.
+
+{\bf Safely Loading Classes.} The original Mantle API is fragile: the API is
+not enforced by the runtime, the system allows injectable strings through the
+admin daemon, constructing the Lua script in C++ is clunky, and the
+administrator can inject really bad policies (e.g., while 1) that brings the
+whole system down. Malacology exposes the important subsystems of Ceph's
+logical storage device framework so that the MDS can safely and dynamically
+load policies specified as Lua classes.
+
+{\bf Policy management.} A new interface is added to the metadata service
+subsystem using the monitoring framework to support management of interfaces. This 
+ensures that the correct balancer version is consistently distributed and verified on
+all MDS nodes. Policies are also durably saved with Ceph's replication and redundancy.
+
+%These problems are general, distributed systems load balancing problems and
+%share many of the same properties as the POSIX metadata
+%wall~\cite{alam:pdsw2011-metadata-scaling,ghemawat:sosp2003-gfs,hildebrand:msst2005-pnfs,weil_ceph_2006,welch:fast2008-panasas,shvachko:login2012-hdfs-scalability}.
+%Many distributed file systems decouple metadata and data I/O so that these
+%services can scale independently. Despite this optimization, scaling the
+%metadata services is still difficult because metadata accesses impose small
+%and frequent requests on the underlying storage
+%system~\cite{roselli:atec2000-FS-workloads}. Many techniques for designing the
+%metadata services have been proposed to accommodate this workload:
+%Lustre~\cite{konstantinos:pdsw2014-lustre-metadata},
+%GFS~\cite{ghemawat:sosp2003-gfs}, and
+%HDFS~\cite{shvachko:login2012-hdfs-scalability} keep all metadata on one
+%server; GIGA+~\cite{patil:fast2011-giga}, IndexFS~\cite{ren:sc2014-indexfs},
+%Lazy Hybrid~\cite{brandt:msst2003-lh}, GPFS~\cite{schmuck:fast2002-gpfs}, and
+%pNFS~\cite{hildebrand:supercomputing2006-pNFS} hash the file system namespace
+%across a dedicated cluster and cache metadata;
+%Panasas~\cite{welch:fast2008-panasas} and
+%CephFS~\cite{weil:sc2004-dyn-metadata} use subtree partitioning to carve up
+%the namespace and distribute it across a cluster.
+%
+%Both the object and balancer interfaces define functions and attach them
+%to the core Lua sandbox. For example, the Lua object storage interface
+%class attaches data, extended attribute, object map, and object version
+%functions, to the Lua core functions, as shown in Figure
+%\ref{fig:cls-osd-mds}. The advantages of this is that we avoid
+%duplicating code, we provide a framework for putting Lua code in other
+%parts of the system, and we remove components and APIs that are too
+%integrated with the OSD.
+
+%To facilitate the use of Lua in both the MDS and OSD, we modularize the LuaVM
+%in a core sandbox wrapper and link against it, as shown in
+%Figure~\ref{fig:cls-osd-mds}. This core Lua wrapper contains just the
+%dependencies, symbols, and functions, needed to run the Lua VM and strips away
+%all the functionality inherent to object classes (like placement group
+%filters, cryptography functions, object metadata,  object data, object
+%extended attributes, and object versions). Some functions of Ceph's more
+%helpful subsystems, like the logging facilities for the daemons, were re-used
+%in the MDS.  With this scheme, The OSD will dynamically open the shared
+%library created by Lua object interface shared library while the MDS will
+%dynamically open the Lua balancer interface shared library; both shared
+%libraries will the Lua core.
+
+{\bf Capabilities.} The metadata service also manages sessions with its
+clients. This gives both subsystems (MDSs and clients) the ability to cache
+data and grab locks. The metadata service is responsible for responding to
+client requests for capabilities by first revoking caps from other clients.
+Currently, clients voluntarily release their capabilities, and the MDS maintains
+a queue of requests.  Both clients and MDSs use a best-effort approach to
+revoking or releasing capabilities.
+
+The clients retain the best-effort, voluntary release of capabilities, but in
+Malacology, the queuing policies are generalized within the MDS to allow for
 centeralized policies such as fairness or priority.
 
-{\bf File Types.}
-We can reference PDSW12 DataMods, and talk about any new stuff here that we
-want to.
-
-{\bf Policy management.} A new interface is added to the metadata service
-subsystem using the monitoring framework to support management of interfaces,
-which ensures durability and distribution of policies in a consistent way.
-
-\begin{figure}[h]
-\centering
-\includegraphics{figures/cls-osd-mds.png}
-\caption{To generalize the LuaVM for use in both the OSD and MDS,
-Malacology detaches the core functionality into a Lua core module. The
-VM is dynamically loaded first (1) and then interfaces can execute (2).
-\label{fig:cls-osd-mds}}
-\end{figure}
+{\bf File Types.} We leverage the same active and typed storage module
+presented in DataMods~\cite{watkins_datamods_2012} to assign state and a set of
+functions to a file (stored in the inode). This is valuable for
+metadata load balancing so that certain types of files can have a higher
+priority when making balancing decisions.
+
+%\begin{figure}[h]
+%\centering
+%\includegraphics{figures/cls-osd-mds.png}
+%\caption{To generalize the LuaVM for use in both the OSD and MDS,
+%Malacology detaches the core functionality into a Lua core module. The
+%VM is dynamically loaded first (1) and then interfaces can execute (2).
+%\label{fig:cls-osd-mds}}
+%\end{figure}
 
@@
 
-The Ceph team wants to merge Mantle because the scriptability is useful
-for debugging, controlling the metadata balancer, and examining
-trade-offs for different balancers. Unfortunately, this research quality
-system is not as robust as Ceph and the Ceph team wants more safety,
-durability, and consistency for the new services. For example, the
-original Mantle API is fragile: the API is not enforced by the runtime,
-the system allows injectable strings through the admin daemon,
-constructing the Lua script in C++ is clunky, and the administrator can
-inject really bad policies (e.g., while 1) that brings the whole system
-down.
-
-The API for Malacology is the same as the one exposed in~\cite{sevilla:sc15-mantle} but the hooks are put in different places using the interface hooks described in Section \S\ref{}.
+The Ceph team wants to merge Mantle because the scriptability is useful for
+debugging, controlling the metadata balancer, and examining trade-offs for
+different balancers. Unfortunately, this research quality system is not as
+robust as Ceph and the Ceph team wants more safety, durability, and consistency
+for the new services.  The API for Malacology is the same as the one exposed
+in~\cite{sevilla:sc15-mantle} but the hooks are put in different places.
 
@@
 
-While storing the code snippets that define the policies in the MDS map is 
-possible, it is not recommended as it is an expensive operation (a PAXOS voting 
-round). Thus, Malacology only stores policies IDs which correspond to an object 
-in a system-maintained pool. When MDSs read a policy, they first check latest 
-version from the MDS map, and then dereference the pointer by reading the 
-corresponding object in RADOS.
+-
+
+-
+
+-
 
@@
 
-We evaluate the mapping of ZLog onto Malacology by examining the costs of
-utilizing the metadata service as a sequencer, and propogating interface
-changes throughout the cluster using the monitoring service. We have omitted
-evaluation of the overhead costs of adding object interfaces themselves due to
-lack of space, and the fact that they are heavily tested and used already in
-production with acceptable performance, introducing a small constant overhead
-per operation but otherwise does not affect scalability.
-
 \subsubsection{Sequencer}
@@
 
-In Section~\ref{consistencyversioning-of-cluster-state} we described an
-interface management service that is responsible for the durable and
-consistent cluster-wide distribution of interface versions. We evaluate the
-costs of such a service by embedding interface definitions in a Ceph cluster
-membership data structure that is managed by a PAXOS instance, and distributed
-using a scalable, peer-to-peer gossip protocol that is built into Ceph. We
-assume an architecture that consists of a light-weight interface that enforces
-the semantics of interface management resulting in one or more interface
-updates that must be consistent across the cluster. The following experiments
-focus on the cost of achieving distribution of consistent a interface version.
+In Section~\ref{} we described an interface management service that is
+responsible for the durable and consistent cluster-wide distribution of
+interface versions. We evaluate the costs of such a service by embedding
+interface definitions in a Ceph cluster membership data structure that is
+managed by a Paxos instance, and distributed using a scalable, peer-to-peer
+gossip protocol. We assume an architecture that consists of a light-weight
+interface that enforces the semantics of interface management resulting in one
+or more interface updates that must be consistent across the cluster. The
+following experiments focus on the cost of achieving distribution of consistent
+a interface version.
 
@@
 \centering
-\includegraphics[trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
+\includegraphics{figures/iface-update-delay.pdf}
 \caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
@@
 
-Figure~\ref{fig:propdelay} shows the CDF of the latency of installing and
-distributing an interface update on all OSDs in the cluster. The latency is
-defined as the elapsed time following the PAXOS proposal for an interface
-update until each OSD makes the update live (the cost of the PAXOS proposal is
+Figure~\ref{fig:propdelay} shows the CDF of the latency of distributing and
+installing an interface update on all OSDs in the cluster. The latency is
+defined as the elapsed time following the Paxos proposal for an interface
+update until each OSDs makes the update live (he cost of the Paxos proposal is
 configurable and is discussed below). The latency measurements were taken on
-the nodes running Ceph server daemons, and thus exclude the client round-trip
-cost. In each of the experiments 1000 interface updates were observed.
+the nodes running Ceph server daemons (the Paxos servers and OSDs nodes), and
+thus excludes the client round-trip cost. In each of the experiments 1000
+interface updates were observed.
 
@@
 (RAM)" a cluster of 120 OSDs (10 OSDs x 12 servers) using an in-memory data
-store were deployed, showing a latency of less than 54 ms with a probability
-of 90\% and a worst case latency of 194 ms. The curve labeled "3 OSD (HDD)" is
-a smaller cluster where each of the three OSDs uses a spinning disk as the
-object store that saves a copy of the cluster map, and shows a latency of 37
-ms with a 90\% probability and a worst case latency of 255 ms. Since the cost
-of a disk I/O is larger than the same I/O executed against the in-memory
-device, the predominant cost in a lightly loaded cluster appears to be the
-number of OSDs.
+store were deployed, showing a latency of less than 54 ms with a probability of
+90\% and a worst case latency of 194 ms. The curve labeled "3 OSD (HDD)" is a
+smaller cluster where each of the three OSDs uses a spinning disk as the object
+store that saves a copy of the cluster map ({\bf is this I/O before its made
+live?}), and shows a latency of 37 ms with a 90\% probability and a worst case
+latency of 255 ms. Since the cost of a disk I/O is larger than the same I/O
+executed against the in-memory device, the predominant cost in a lightly loaded
+cluster appears to be the number of OSDs.
 
@@
 of interface updates, includes the network round-trip to the interface
-management service, the PAXOS commit protocol, and other factors such as system
-load. By default PAXOS proposals occur periodically with a 1 second interval in
-order to accumulate updates. In a minimum, realistic quorum of 3 monitors using
+management service, the Paxos commit protocol, and other factors such as system
+load. By default Paxos proposals occur periodically with a 1 second interval in
+order to accumulate updates. In a minimum realistic quorum of 3 monitors using
 HDD-based storage, we were able to decrease this interval to an average of 222
@@
 
-Malacology uses the same Active and Typed Storage module presented in
-DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File
-Manifolds can be implemented with small changes to the Malacology
-framework, namely asynchronous object calls and Lua stubs in the inode,
-respectively.
-
 \section{Conclusion and Future Work}\label{conclusion-and-future-work}

### 2016-05-10 paper/paper.tex

@@
 
-\subsection{Policy Durability}
-
--
-
--
-
--
+\subsection{Making Policies Durable}
+
+While storing the code snippets that define the policies in the MDS map is 
+possible, it is not recommended as it is an expensive operation (a PAXOS voting 
+round). Thus, Malacology only stores policies IDs which correspond to an object 
+in a system-maintained pool. When MDSs read a policy, they first check latest 
+version from the MDS map, and then dereference the pointer by reading the 
+corresponding object in RADOS.
 
@@
 \subsection{ZLog}
+
+We evaluate the mapping of ZLog onto Malacology by examining the costs of
+utilizing the metadata service as a sequencer, and propogating interface
+changes throughout the cluster using the monitoring service. We have omitted
+evaluation of the overhead costs of adding object interfaces themselves due to
+lack of space, and the fact that they are heavily tested and used already in
+production with acceptable performance, introducing a small constant overhead
+per operation but otherwise does not affect scalability.
 
@@
 (RAM)" a cluster of 120 OSDs (10 OSDs x 12 servers) using an in-memory data
-store were deployed, showing a latency of less than 54 ms with a probability of
-90\% and a worst case latency of 194 ms. The curve labeled "3 OSD (HDD)" is a
-smaller cluster where each of the three OSDs uses a spinning disk as the object
-store that saves a copy of the cluster map ({\bf is this I/O before its made
-live?}), and shows a latency of 37 ms with a 90\% probability and a worst case
-latency of 255 ms. Since the cost of a disk I/O is larger than the same I/O
-executed against the in-memory device, the predominant cost in a lightly loaded
-cluster appears to be the number of OSDs.
-
+store were deployed, showing a latency of less than 54 ms with a probability
+of 90\% and a worst case latency of 194 ms. The curve labeled "3 OSD (HDD)" is
+a smaller cluster where each of the three OSDs uses a spinning disk as the
+object store that saves a copy of the cluster map, and shows a latency of 37
+ms with a 90\% probability and a worst case latency of 255 ms. Since the cost
+of a disk I/O is larger than the same I/O executed against the in-memory
+device, the predominant cost in a lightly loaded cluster appears to be the
+number of OSDs.
+
+\textbf{The Malacology Approach:} Re-use Ceph's Monitor functionality to 
+propagate application-specific state that needs to be highly-available and 
+strongly consistent.
 The real cost of interface management, in addition to cluster-wide propagation
 of interface updates, includes the network round-trip to the interface
-management service, the Paxos commit protocol, and other factors such as system
-load. By default Paxos proposals occur periodically with a 1 second interval in
-order to accumulate updates. In a minimum realistic quorum of 3 monitors using
+management service, the PAXOS commit protocol, and other factors such as system
+load. By default PAXOS proposals occur periodically with a 1 second interval in
+order to accumulate updates. In a minimum, realistic quorum of 3 monitors using
 HDD-based storage, we were able to decrease this interval to an average of 222

### 2016-05-10 paper/paper.tex
spalling mystaces
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 file system metadata load balancer and a high-performance distributed shared-log
-that leverages flash devices.  RESULTS
+that leverages flash devices.
 \end{abstract}
@@
 production-quality scalable storage systems available as open source
-software have established and are continuing to establish new, \emph{de
-facto} API standards at a faster pace than traditional standards bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
+software have established and are continuing to establish new, \emph{de-facto} API standards at a faster pace than traditional standards bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
 
@@
 abstractions and services (indicated by bold arrows). To illustrate Malacology we implement two
-services on Ceph, Zlog and Mantle, by programming Ceph's durability,
+services on Ceph, ZLog and Mantle, by programming Ceph's durability,
 consistent versioning, consensus, and metadata subsystems.
@@
 
-In this paper we investigate an approach that focusses on generalizing
+In this paper we investigate an approach that focuses on generalizing
 existing storage system resources, services, and abstractions that in
@@
 locking and caching services provided by MDSs, durability and logical
-storage devices provided by RADOS, and propogation of consistent cluster
+storage devices provided by RADOS, and propagation of consistent cluster
 state provided by the monitoring service (see Figure~\ref{fig:overview}). As we will show in this paper, this framework is expressive enough to provide the
@@
   \item
-    A high-performance distributed shared log service called Zlog which is based on CORFU~\cite{balakrishnan_corfu_2012}
+    A high-performance distributed shared log service called ZLog which is based on CORFU~\cite{balakrishnan_corfu_2012}
   \item
@@
 Malacology. Lua is a portable, embedded
-scripting language that offers superious performance and productivity
+scripting language that offers superior performance and productivity
 trade-offs, including a JIT-based implementation that is well known for near
@@
 of application-specific object interfaces~\cite{weil_rados_2007}. The ability
-to offload computation can reduce data movement, and transactional interfaces can
+to offload computation can reduce data movement, and transnational interfaces can
 significantly simplify construction of complex storage interfaces that require
@@
 function input and output. A wide range of native interfaces are available to
-developers, such as reading and writing to a bytestream, controlling snapshots
+developers, such as reading and writing to a byte stream, controlling snapshots
 and clones, all of which may be transactionally composed with access to a
@@
 \caption{Malacology re-uses Ceph subsystems to manage interface classes.
-command. After the proposal is accepted by the cluster of monitors, the
+After the proposal is accepted by the cluster of monitors, the
 script is redundantly and consistently distributed to OSDs.
@@
 guarantees that interface definitions are not only made durable, but are
-transparently and consistently propogated throughput the cluster and that
-clients are properly synchronousized with the latest interfaces.
+transparently and consistently propagated throughput the cluster and that
+clients are properly synchronized with the latest interfaces.
 
@@
 the metadata service to control load balancing and capabilities, and for
-defining new file types allowing the indexing service to be repurposed for
+defining new file types allowing the indexing service to be re-purposed for
 other types of named resources.
@@
 {\bf Safely Loading Classes.} The original Mantle API is fragile: the API is
-not enforced by the runtime, the system allows injectable strings through the
+not enforced by the runtime, the system allows strings to be injected through the
 admin daemon, constructing the Lua script in C++ is clunky, and the
@@
 Malacology, the queuing policies are generalized within the MDS to allow for
-centeralized policies such as fairness or priority.
+centralized policies such as fairness or priority.
 
@@
 techniques try to be autonomous so that work is distributed across the
-cluster. For example, when placement groups change, the OSDs rebalance and
+cluster. For example, when placement groups change, the OSDs re-balance and
 re-shard data in the background in a process called placement group splitting.
@@
 we use this service to automatically install interfaces in OSDs, and install
-policies within the MDS such that clients and daemons are synchronzied on
+policies within the MDS such that clients and daemons are synchronized on
 correct implementations without restarting.
@@
 
-High-performance in CORFU is achived using a sequencer service that assigns
+High-performance in CORFU is achieved using a sequencer service that assigns
 log positions to clients by reading from a volatile, in-memory counter which
 can run at a very high throughput and at low-latency. And since the sequencer
-is centralized, ensuring seiralizability in the common case is trivial.  The
+is centralized, ensuring serializability in the common case is trivial.  The
 primary challenge in CORFU is handling the failure of the sequencer in a way
@@
 an application-specific custom storage interface to discover the tail of the
-log, while simulatenously invalidating stale client requests using an
+log, while simultaneously invalidating stale client requests using an
 epoch-based protocol.
@@
 non-exclusive, non-cacheable resource, forcing clients to perform a round-trip
-access to the resource at the authoratiave metadata server for the sequencer
-inode.  Interestingly, we found that the capapbility system in Ceph uses a
+access to the resource at the authoritative metadata server for the sequencer
+inode.  Interestingly, we found that the capability system in Ceph uses a
 strategy to reduce metadata service load by allowing multiple clients that
@@
 independently map log positions that they have obtained from a sequencer
-service (described in detail in the next seciton) onto storage devices, while
+service (described in detail in the next section) onto storage devices, while
 storage devices provide an intelligent \emph{write}-once, random \emph{read}
@@
 with out-of-date epoch values are rejected, and clients are expected to
-request a new tail from the sequencer after refreshing state from an auxillary
+request a new tail from the sequencer after refreshing state from an auxiliary
 service.  This mechanism forms the basis for sequencer recovery.
@@
 process itself may be handled in many ways, such as leader election using an
-auxillary service such as PAXOS.
+auxiliary service such as PAXOS.
 
@@
 We evaluate the mapping of ZLog onto Malacology by examining the costs of
-utilizing the metadata service as a sequencer, and propogating interface
+utilizing the metadata service as a sequencer, and propagating interface
 changes throughout the cluster using the monitoring service. We have omitted
@@
 The highest performance is achieved using a single client with
-exclusive, cacheable priviledge. Round-robin sharing of the sequencer
+exclusive, cacheable privilege. Round-robin sharing of the sequencer
 resource is affected by the amount of time the resource is held,
@@
 system client, throughput drops significantly with two clients as the majority
-of the time neither client has the capapbitility as it ping-pongs between
+of the time neither client has the capability as it ping-pongs between
 clients. This observation reveals a tuning parameter: how long a client
-maintains its capapbility lease. In throughputs labeled $2C,<time>$ we show
+maintains its capability lease. In throughput labeled $2C,<time>$ we show
 aggregate throughput as we decrease the amount of time that a client holds its
-lease before responding to a request for the capapbility.  At lease times as
-low as 10ms throughputs achieve a significant proportion of the locally cached
+lease before responding to a request for the capability.  At lease times as
+low as 10ms throughput achieve a significant proportion of the locally cached
 throughput.
@@
 
-\subsubsection{Interface Propogation}
+\subsubsection{Interface Propagation}
 
@@
 capabilities of the metadata cluster. The ``Sequencer Baseline'' has no
-background sequencers, while the the ``2 Seq'' curves are the throughputs
+background sequencers, while the ``2 Seq'' curves are the throughput
 of 1 sequencer when an additional sequencer is running in the background
@@
 behavior of the same workload, 2 sequencers, with balancing turned on and off.
-The curves are the throughputs (\(y\) axis) over time (\(x\) axis) and the graph
+The curves are the throughput (\(y\) axis) over time (\(x\) axis) and the graph
 has been partitioned into the two cluster setups. The CPU utilization is
-confined to one core the MDS is not multithreaded, as it has been shown that
+confined to one core the MDS is not multi-threaded, as it has been shown that
 metadata performance does not improve with multiple
@@
 performance for this particular workload because there is no contention.
-Performance is stable and the MDS is underloaded because we do not see the
+Performance is stable and the MDS is under-loaded because we do not see the
 spike at the end like the other curves. It appears the bottleneck is somewhere
@@
 
-We are intend to pursue this work towards the goal of constructing a set
+We are intending to pursue this work towards the goal of constructing a set
 of customization points that allow a wide variety of storage system
 services to be configured on-the-fly in existing systems. This work is
-one point along that path in which we have looked an a target
-special-purpose storage system. Ultimately we want to utilize
+one point along that path in which we have looked at targeting
+special-purpose storage systems. Ultimately we want to utilize
 declarative methods for expressing new services.
-
-In conclusion, this paper should be accepted.
 

### 2016-05-10 paper/paper.tex
paper: add result to abstract
Related work section

--- 2016-05-10 paper/paper.tex
@@
 The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction~\cite{weil:sc2004-dyn-metadata}. In general, distributed
-file systems protect resources by providing hierarchical indexing and
-distributed locking services.  In Ceph, the locking service implements a
-capability-based system that express what data and metadata clients are allowed
-to access as well as what state they may cache and modify locally. While
-designed for a fixed file abstraction, indexing, locking, and caching are all
-common services that are useful to a broad spectrum of applications.
-
-Ceph also addresses the challenge of balancing metadata load using load
-balancing policies that compute metrics based on system state (e.g.  CPU and
-memory utilization) and statistics collected by the cluster (e.g. the
-popularity of an inode). Ceph uses dynamic subtree partitioning to move
-variable sized namespace subtrees. These units can be shipped anywhere (i.e.,
-to any MDS of any capacity) at any time for any reason. The original balancer
-was designed with hard-coded policies and tunables.  Distributed applications
-that share centralized resources (e.g. a database or directory) face similar
-challenges which are often solved using application-specific sharding.
-
-%The metadata cluster moves units of the namespace called directory fragments
-%and assigns them to servers using its own metadata load balancer. It then uses
-%a hard-coded policy to balance these fragments using a scalarized load metric
-%based on CPU, workload, and file system operation metrics. Although the CephFS
-%balancer has been shown to be complicated, its complexity is justifiable.
-%Having the flexibility to choose what to move, where to move it, and how much
-%to move is very powerful. The problem is that it is difficult to pick a
-%one-size-fits-all policy, system administrators may want to partition load
-%based on SLAs, or applications are better equipped to make these decisions,
-%etc. -- but we choose instead to expose a balancing API at strategic points in
-%the balancing logic using Lua hooks. ~\cite{sevilla:sc15-mantle}.
-
-\paragraph*{Malacology:} New interface hooks are added in strategic spots in
-the metadata service to control load balancing and capabilities, and for
-defining new file types allowing the indexing service to be repurposed for
-other types of named resources.
-
-{\bf Load balancing.} Mantle~\cite{sevilla:sc15-mantle}, a programmable
-metadata load balancer, has been re-implemented on Malacology so that its
-infrastructure can enjoy all the properties of other Ceph components. The load
-balancing logic has stayed largely the same but the infrastructure has been
-improved to safely load, manage, and persist its load balancing policies.
-
-{\bf Safely Loading Classes.} The original Mantle API is fragile: the API is
-not enforced by the runtime, the system allows injectable strings through the
-admin daemon, constructing the Lua script in C++ is clunky, and the
-administrator can inject really bad policies (e.g., while 1) that brings the
-whole system down. Malacology exposes the important subsystems of Ceph's
-logical storage device framework so that the MDS can safely and dynamically
-load policies specified as Lua classes.
+system abstraction~\cite{cephfs}. Distributed file systems are complex, but
+fundamentally provide hierarchical indexing and distributed locking services.
+In Ceph, the locking service implements a capability-based system that express
+what data and metadata clients are allowed to access as well as what state
+they may cache and modify locally. While designed for a fixed file
+abstraction, indexing, locking, and caching are all common services that are
+useful to a broad spectrum of applications.
+
+Ceph addresses the challenge of balancing load across the metadata cluster
+using load balancing policies that compute balancing metrics based on system
+state (e.g. CPU and memory utilization) and statistics collected by the
+cluster (e.g. the popularity of an inode). Distributed applications that share
+centralized resources (e.g. a database or directory) face similar challenges
+which are often solved using application-specific sharding.
+
+\paragraph*{Malacology:}
+Malacogy adds interface hooks in strategic spots in the metadata service to
+control load balancing, capability subsystems, and for defining new file
+types allowing the indexing service to be repurposed for other types of
+named resources.
+
+{\bf Load balancing.}
+For CephFS, designing a general POSIX metadata load balancer is difficult
+becuase the workloads are diverse and quickly changing. With CephFS, it is
+especially complicated because the system has variable sized migration units
+(i.e. file system namespace subtrees), which can be shipped anywhere (i.e., to
+any MDS of any capacity) at any time for any reason. Furthermore, the original
+balancer was designed with hard-coded policies and tunables. Both angles were
+difficult to understand. It is understandable why it was implemented this way
+-- it is difficult to pick a one-size-fits-all policy, system administrators
+may want to partition load based on SLAs, or applications are better equipped
+to make these decisions, etc. -- but we choose instead to expose a balancing
+API at strategic points in the balancing logic using Lua hooks.
+
+These problems are general, distributed systems load balancing problems and
+share many of the same properties as the POSIX metadata
+wall~\cite{alam:pdsw2011-metadata-scaling,ghemawat:sosp2003-gfs,hildebrand:msst2005-pnfs,weil_ceph_2006,welch:fast2008-panasas,shvachko:login2012-hdfs-scalability}.
+Many distributed file systems decouple metadata and data I/O so that these
+services can scale independently. Despite this optimization, scaling the
+metadata services is still difficult because metadata accesses impose small
+and frequent requests on the underlying storage
+system~\cite{roselli:atec2000-FS-workloads}. Many techniques for designing the
+metadata services have been proposed to accommodate this workload:
+Lustre~\cite{konstantinos:pdsw2014-lustre-metadata},
+GFS~\cite{ghemawat:sosp2003-gfs}, and
+HDFS~\cite{shvachko:login2012-hdfs-scalability} keep all metadata on one
+server; GIGA+~\cite{patil:fast2011-giga}, IndexFS~\cite{ren:sc2014-indexfs},
+Lazy Hybrid~\cite{brandt:msst2003-lh}, GPFS~\cite{schmuck:fast2002-gpfs}, and
+pNFS~\cite{hildebrand:supercomputing2006-pNFS} hash the file system namespace
+across a dedicated cluster and cache metadata;
+Panasas~\cite{welch:fast2008-panasas} and
+CephFS~\cite{weil:sc2004-dyn-metadata} use subtree partitioning to carve up
+the namespace and distribute it across a cluster.
+
+CephFS moves units of the namespace called directory fragments and assigns
+them to servers using  its own metadata load balancer. It then uses a
+hard-coded policy to balance these fragments using a secularized load metric
+based on CPU, workload, and file system operation
+metrics~\cite{sevilla:sc15-mantle}.
+
+Both the object and balancer interfaces define functions and attach them
+to the core Lua sandbox. For example, the Lua object storage interface
+class attaches data, extended attribute, object map, and object version
+functions, to the Lua core functions, as shown in Figure
+\ref{fig:cls-osd-mds}. The advantages of this is that we avoid
+duplicating code, we provide a framework for putting Lua code in other
+parts of the system, and we remove components and APIs that are too
+integrated with the OSD.
+
+To facilitate the use of Lua in both the MDS and OSD, we modularize the LuaVM
+in a core sandbox wrapper and link against it, as shown in
+Figure~\ref{fig:cls-osd-mds}. This core Lua wrapper contains just the
+dependencies, symbols, and functions, needed to run the Lua VM and strips away
+all the functionality inherent to object classes (like placement group
+filters, cryptography functions, object metadata,  object data, object
+extended attributes, and object versions). Some functions of Ceph's more
+helpful subsystems, like the logging facilities for the daemons, were re-used
+in the MDS.  With this scheme, The OSD will dynamically open the shared
+library created by Lua object interface shared library while the MDS will
+dynamically open the Lua balancer interface shared library; both shared
+libraries will the Lua core.
+
+{\bf Capabilities.}
+The metadata service also manages sessions with its clients. This gives both
+subsystems (MDSs and clients) the ability cache data and grab locks. The
+metadata service is responsible for responding to client requests for
+capabilities by first revoking caps from other clients. Currently clients
+voluntarily release their capabilities, and the MDS maintains a queue of
+requests.  Both clients and MDSs use a best-effort approach to revoking or
+releasing capabilities.
+
+In our current we retain the best-effort, voluntary release of caps within
+clients, but generalize the queueing policies within the MDS to allow for
+centeralized policies such as fairness or priority.
+
+{\bf File Types.}
+We can reference PDSW12 DataMods, and talk about any new stuff here that we
+want to.
 
 {\bf Policy management.} A new interface is added to the metadata service
-subsystem using the monitoring framework to support management of interfaces. This 
-ensures that the correct balancer version is consistently distributed and verified on
-all MDS nodes. Policies are also durably saved with Ceph's replication and redundancy.
-
-%These problems are general, distributed systems load balancing problems and
-%share many of the same properties as the POSIX metadata
-%wall~\cite{alam:pdsw2011-metadata-scaling,ghemawat:sosp2003-gfs,hildebrand:msst2005-pnfs,weil_ceph_2006,welch:fast2008-panasas,shvachko:login2012-hdfs-scalability}.
-%Many distributed file systems decouple metadata and data I/O so that these
-%services can scale independently. Despite this optimization, scaling the
-%metadata services is still difficult because metadata accesses impose small
-%and frequent requests on the underlying storage
-%system~\cite{roselli:atec2000-FS-workloads}. Many techniques for designing the
-%metadata services have been proposed to accommodate this workload:
-%Lustre~\cite{konstantinos:pdsw2014-lustre-metadata},
-%GFS~\cite{ghemawat:sosp2003-gfs}, and
-%HDFS~\cite{shvachko:login2012-hdfs-scalability} keep all metadata on one
-%server; GIGA+~\cite{patil:fast2011-giga}, IndexFS~\cite{ren:sc2014-indexfs},
-%Lazy Hybrid~\cite{brandt:msst2003-lh}, GPFS~\cite{schmuck:fast2002-gpfs}, and
-%pNFS~\cite{hildebrand:supercomputing2006-pNFS} hash the file system namespace
-%across a dedicated cluster and cache metadata;
-%Panasas~\cite{welch:fast2008-panasas} and
-%CephFS~\cite{weil:sc2004-dyn-metadata} use subtree partitioning to carve up
-%the namespace and distribute it across a cluster.
-%
-%Both the object and balancer interfaces define functions and attach them
-%to the core Lua sandbox. For example, the Lua object storage interface
-%class attaches data, extended attribute, object map, and object version
-%functions, to the Lua core functions, as shown in Figure
-%\ref{fig:cls-osd-mds}. The advantages of this is that we avoid
-%duplicating code, we provide a framework for putting Lua code in other
-%parts of the system, and we remove components and APIs that are too
-%integrated with the OSD.
-
-%To facilitate the use of Lua in both the MDS and OSD, we modularize the LuaVM
-%in a core sandbox wrapper and link against it, as shown in
-%Figure~\ref{fig:cls-osd-mds}. This core Lua wrapper contains just the
-%dependencies, symbols, and functions, needed to run the Lua VM and strips away
-%all the functionality inherent to object classes (like placement group
-%filters, cryptography functions, object metadata,  object data, object
-%extended attributes, and object versions). Some functions of Ceph's more
-%helpful subsystems, like the logging facilities for the daemons, were re-used
-%in the MDS.  With this scheme, The OSD will dynamically open the shared
-%library created by Lua object interface shared library while the MDS will
-%dynamically open the Lua balancer interface shared library; both shared
-%libraries will the Lua core.
-
-{\bf Capabilities.} The metadata service also manages sessions with its
-clients. This gives both subsystems (MDSs and clients) the ability to cache
-data and grab locks. The metadata service is responsible for responding to
-client requests for capabilities by first revoking caps from other clients.
-Currently, clients voluntarily release their capabilities, and the MDS maintains
-a queue of requests.  Both clients and MDSs use a best-effort approach to
-revoking or releasing capabilities.
-
-The clients retain the best-effort, voluntary release of capabilities, but in
-Malacology, the queuing policies are generalized within the MDS to allow for
-centeralized policies such as fairness or priority.
-
-{\bf File Types.} We leverage the same active and typed storage module
-presented in DataMods~\cite{watkins_datamods_2012} to assign state and a set of
-functions to a file (stored in the inode). This is valuable for
-metadata load balancing so that certain types of files can have a higher
-priority when making balancing decisions.
-
-%\begin{figure}[h]
-%\centering
-%\includegraphics{figures/cls-osd-mds.png}
-%\caption{To generalize the LuaVM for use in both the OSD and MDS,
-%Malacology detaches the core functionality into a Lua core module. The
-%VM is dynamically loaded first (1) and then interfaces can execute (2).
-%\label{fig:cls-osd-mds}}
-%\end{figure}
+subsystem using the monitoring framework to support management of interfaces,
+which ensures durability and distribution of policies in a consistent way.
+
+\begin{figure}[h]
+\centering
+\includegraphics{figures/cls-osd-mds.png}
+\caption{To generalize the LuaVM for use in both the OSD and MDS,
+Malacology detaches the core functionality into a Lua core module. The
+VM is dynamically loaded first (1) and then interfaces can execute (2).
+\label{fig:cls-osd-mds}}
+\end{figure}
 
@@
 
-The Ceph team wants to merge Mantle because the scriptability is useful for
-debugging, controlling the metadata balancer, and examining trade-offs for
-different balancers. Unfortunately, this research quality system is not as
-robust as Ceph and the Ceph team wants more safety, durability, and consistency
-for the new services.  The API for Malacology is the same as the one exposed
-in~\cite{sevilla:sc15-mantle} but the hooks are put in different places.
+The Ceph team wants to merge Mantle because the scriptability is useful
+for debugging, controlling the metadata balancer, and examining
+trade-offs for different balancers. Unfortunately, this research quality
+system is not as robust as Ceph and the Ceph team wants more safety,
+durability, and consistency for the new services. For example, the
+original Mantle API is fragile: the API is not enforced by the runtime,
+the system allows injectable strings through the admin daemon,
+constructing the Lua script in C++ is clunky, and the administrator can
+inject really bad policies (e.g., while 1) that brings the whole system
+down.
+
+The API for Malacology is the same as the one exposed in~\cite{sevilla:sc15-mantle} but the hooks are put in different places using the interface hooks described in Section \S\ref{}.
 
@@
 
+Programmability of operating systems and networking resources, including distributed storage systems is not new, but we are not aware of work that makes generalization of existing services into programmable resourses a key principle in storage systems design.
+
+Programmable storage systems can be viewed as an infrastructure for creating abstractions to better separate policies from mechanisms.
+This idea is not new.
+Software-defined networks (SDNs) create such an abstraction by separating the control plane from the data plane (see for example~\cite{jain:sigcomm13}).
+This notion of control/data separation was also applied in software-defined storage (SDS)~\cite{arpaci:sosp01,thereska:sosp13,stefanovici:fast16}.
+Similarly, IOStack~\cite{gracia:internet16} is providing policy-based provisioning and filtering in OpenStack Swift.
+
+Another view of programmable storage systems is one of tailoring systems resources to applications~\cite{arpaci:sosp01}.
+Related work includes work around the Exokernel~\cite{engler:sosp95} and SPIN~\cite{bershad:sosp95} and Vino~\cite{seltzer:osdi96}, the latter two addressed the ability of injecting code into the kernel to specialize resource management.
+Another approach is to pass hints between the different layers of the IO stack to bridge the semantic gap between applications and storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}).
+
+Malacology uses the same Active and Typed Storage module presented in
+DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File
+Manifolds can be implemented with small changes to the Malacology
+framework, namely asynchronous object calls and Lua stubs in the inode,
+respectively.
+
 \section{Conclusion and Future Work}\label{conclusion-and-future-work}

### 2016-05-10 paper/paper.tex
Related work prose and references

--- 2016-05-10 paper/paper.tex
@@
 file system metadata load balancer and a high-performance distributed shared-log
-that leverages flash devices.
+that leverages flash devices. The evaluation demonstrates that our services 
+inherit desirable qualities of Ceph, including the ability to make trade-offs 
+for lease capabilities, balance load (2\(\times\) improvement), propagate interfaces 
+(average of 222ms) and recover (15 second down time).  
 \end{abstract}

### 2016-05-10 paper/paper.tex
New image for malacology implementation
Minor rewording to 3.1

@@
 \paragraph*{Malacology:} Malacology exposes Ceph's PAXOS-based monitoring
-service by embedding a Lua VM in the monitors that implements a generic API
-for creating dynamic interfaces around existing subsystems that may associated
-additional state with a subsystem map.
-As a consequence of this, applications can define simple but useful 
-domain-specific logic to the PAXOS quorum, such as authorization control (just 
-specific clients can write new values) or to trigger actions based on specific 
-values (e.g. sanitize values). Since the monitor is intended to be out of 
-high-performance I/O paths, a general guideline is to make use of this 
-functionality infrequently and to assign small values to maps. Malacology itself 
-makes use of this functionality to version object interfaces defined in OSDs 
+service implementing a generic API
+for adding arbitrary values to existing subsystem cluster maps, namely to the 
+OSD map and the MDS map. As a consequence of this, applications can define 
+simple but useful domain-specific logic to the PAXOS state machine, such as 
+authorization control (just specific clients can write new values) or to trigger 
+actions based on specific values (e.g. sanitize values). Since the monitor is 
+intended to be out of high-performance I/O paths, a general guideline is to make 
+use of this functionality infrequently and to assign small values to maps. 
+Malacology itself makes use of this functionality to register, version and 
+propagate dynamic code (Lua scripts) for new object interfaces defined in OSDs 
 (\S\ref{active-storage}) and load balancing policies in MDSs 
@@
 \includegraphics{figures/implementation.png}
-\caption{Malacology re-uses Ceph subsystems to manage interface classes.
-After the proposal is accepted by the cluster of monitors, the
-script is redundantly and consistently distributed to OSDs.
+\caption{Malacology re-uses Ceph subsystems to allow users to dynamically define 
+new object interfaces and load balancing policies. It achieves this by extending 
+the OSD and MDS subsystems with an embedded Lua VM. To manage newly dynamic 
+interface classes, it exposes the core functionality of the Monitor to allow 
+arbitrary data to be associated to ODS and MDS cluster maps. After new 
+interfaces are submitted to the monitor and the associated Paxos proposal is 
+accepted by the monitor quorum, the script is redundantly and consistently 
+distributed to OSDs so this snippets persist.
 \label{fig:implementation}}
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/overview-mantle.png}
-\caption{This is Mantle.}
-\end{figure}
-
 \subsection{ZLog: A Fast Distributed Shared-Log}

### 2016-05-10 paper/paper.tex
fix refs
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-First, we describe existing techniques for adapting storage systems
-(\S\ref{sec:evidence}). Next, we discuss the ``distributed systems'' inspired
-abstractions employed by Ceph (\S\ref{background}) and allude to their
-re-usability as components in other services. The next section
-enumerates the abstractions that Malacology exposes in the
-implementation section (\S\ref{implementation}). We then show how we use
-Malacology to synthesize entirely new storage services on an existing
-system through configuration and small changes (\S\ref{services}). We
-conclude by evaluating real-world use cases (\S\ref{evaluation}).
+The remainder of this paper is structured as follows. First we describe
+Malacology by presenting the subsystems within Ceph that we re-purpose, and
+briefly describe how those system are used within Malacology
+(\S\ref{sec:malacology}). Next we describe the services that we have
+constructed within the Malacology framework (\S\ref{sec:services}),
+and evaluate our ideas within our prototype implementation
+(\S\ref{evaluation}). Finally we discuss related work and conclude.
 
@@
 \subsection{Cluster State Management}
+\label{sec:mon}
 \label{consistencyversioning-of-cluster-state}
@@
 any changes require a full restart of an OSD which may have serious
-performance impacts due to a loss of cached data~\cite{facebook-mysql}. And
+performance impacts due to a loss of cached data. And
 finally, the security limitations of the framework limit the use of object
@@
 with economy of expression, which at the same time provides the full set of
-features of the base object class~\cite{bdmc}. New object interfaces that are
+features of the base object class. New object interfaces that are
 expressed in thousands of lines of code can be implemented in approximately an
@@
 
-\section{Services Built on Malacology}\label{services-built-on-malacology}
+\section{Services Built on Malacology}
+\label{sec:services}
+\label{services-built-on-malacology}
 
@@
 
-In Section~\cite{evaluation} we quantify the trade-offs of throughput and
+In Section~\ref{evaluation} we quantify the trade-offs of throughput and
 latency for an approach based on a round-robin batching mode, and compare this
@@
 
-\section{Evaluation}\label{evaluation}
+\section{Evaluation}
+\label{evaluation}
 We have conducted our evaluation on CloudLab as well as our own 20 node cluster.
@@
 \includegraphics{figures/mantle-cpu-utilization-annotate.png}
-\caption{The sequencer using Mantle in Figure~\ref{fig:mantle-seq-thruput.png}
+\caption{The sequencer using Mantle in Figure~\ref{fig:mantle-seq-thruput}
 migrates to a new MDS immediately because its uses a policy designed for bursty

### 2016-05-10 paper/paper.tex
Challenges section

--- 2016-05-10 paper/paper.tex
@@
 
-The challenges of introducing programmability into existing storage systems: (1) Programmability while keeping storage system available, (2) Finding and factoring out hard-wired policies and optimizations, (3) Bugs that were hidden because mechanisms were only exercised according to hard-wired policies and not in their full generality they were thought of -- programmability into a storage system turned out to be a great debugging tool, (4) version control and propagation of updates throughout the system without interrupting production workloads, (5) ...
+Establishing the infrastructure for programmability into existing services and abstractions of distributed storage systems is challenging, even if one assumes that the source code of the storage system and the necessary expertise for understanding it is available:
+
+\begin{itemized}
+
+\item Storage systems are generally required to be highly available so that any complete restarts of the storage system to reprogram them is usually unacceptable.
+
+\item Policies and optimizations are usually hard-wired into the services and one has to be careful when factoring them out not to introduce additional bugs.
+
+\item Mechanisms that are only exercised according to hard-wired policies and not in their full generality have hidden bugs that are revealed as soon as those mechanisms are governed by different policies.
+In our experience introducing programmability into a storage system proved to be a great debugging tool.
+
+\item Programmability, especially in live systems, implies changes that need to be carefully managed by the system itself, including versioning and propagation of those changes without affecting correctness.
+
+\end{itemize}
 

### 2016-05-10 paper/paper.tex
List
popper: do it
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-05-10 paper/paper.tex
@@
 \includegraphics{figures/obj-int-dev-growth.png}
-\caption{Since 2010, the growth in the number of co-designed object
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] Since 2010, the growth in the number of co-designed object
 storage interfaces in Ceph has been accelerating. This plot is the
@@
 
-\section{Challenges}
-\label{sec:challenges}
-
-The challenges of introducing programmability into existing storage systems: (1) Programmability while keeping storage system available, (2) Finding and factoring out hard-wired policies and optimizations, (3) Bugs that were hidden because mechanisms were only exercised according to hard-wired policies and not in their full generality they were thought of -- programmability into a storage system turned out to be a great debugging tool, (4) version control and propagation of updates throughout the system without interrupting production workloads, (5) ...
+%\section{Challenges}
+%\label{sec:challenges}
+%
+%The challenges of introducing programmability into existing storage systems: (1) Programmability while keeping storage system available, (2) Finding and factoring out hard-wired policies and optimizations, (3) Bugs that were hidden because mechanisms were only exercised according to hard-wired policies and not in their full generality they were thought of -- programmability into a storage system turned out to be a great debugging tool, (4) version control and propagation of updates throughout the system without interrupting production workloads, (5) ...
 
@@
 We follow the ``Popper Convention"~\cite{anonymous} so as to try and make our experiments as reproducible as possible.
-We have a GitHub repository with our source code,  experimental environment, configuration parameters, and workloads in a GitHub repository.
-Due to the double-blind policy, we have temporarily disabled the links to our graphs and experimental details.
+We have a GitHub repository with our source code, experimental environment, configuration parameters, and workloads but due to the double-blind policy, we have temporarily disabled the links to our graphs and experimental details.
 
@@
 components and subsystems exposed by the framework. Detailed system-wide performance
-measurements and analysis is forthcoming for both Mantle and ZLog~\cite{}, as we try
+measurements and analysis is forthcoming for both Mantle and ZLog, as we try
 to find the best metadata balancing policies and physical design implementation trade-offs
@@
 \includegraphics{figures/caps-delay-thruput.png}
-\caption{Sequencer throughput by reusing various services.
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] Sequencer throughput by re-using various services.
 The highest performance is achieved using a single client with
@@
 \includegraphics{figures/caps-delay-latency.png}
-\caption{But if the clients hold their capabilities longer, it hurts
-latency for everyone.}
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The latency distribution shows that if the clients hold their capabilities longer, performance decreases.
+Changing the delay via policy can help the sequencer and clients strike a balance between throughput and latency.}
 \label{fig:capcdf}
@@
 \includegraphics[trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
-\caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] Cluster-wide interface update latency, excluding the Paxos proposal cost for
 interface commit. Each of the 120 OSDs used a RAM-based object store, while the
@@
 \includegraphics{figures/mantle-seq-thruput.png}
-\caption{A ZLog sequencer that uses the CephFS MDS enjoys the load balancing
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] A ZLog sequencer that uses the CephFS MDS enjoys the load balancing
 capabilities of the metadata cluster. The ``Sequencer Baseline'' has no
@@
 \includegraphics{figures/mantle-cpu-utilization-annotate.png}
-\caption{The sequencer using Mantle in Figure~\ref{fig:mantle-seq-thruput}
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The sequencer using Mantle in Figure~\ref{fig:mantle-seq-thruput}
 migrates to a new MDS immediately because its uses a policy designed for bursty

### 2016-05-10 paper/paper.tex

@@
 %The challenges of introducing programmability into existing storage systems: (1) Programmability while keeping storage system available, (2) Finding and factoring out hard-wired policies and optimizations, (3) Bugs that were hidden because mechanisms were only exercised according to hard-wired policies and not in their full generality they were thought of -- programmability into a storage system turned out to be a great debugging tool, (4) version control and propagation of updates throughout the system without interrupting production workloads, (5) ...
+\section{Challenges}
+\label{sec:challenges}
+
+Establishing the infrastructure for programmability into existing services and abstractions of distributed storage systems is challenging, even if one assumes that the source code of the storage system and the necessary expertise for understanding it is available:
+
+\begin{itemize}
+
+\item Storage systems are generally required to be highly available so that any complete restarts of the storage system to reprogram them is usually unacceptable.
+
+\item Policies and optimizations are usually hard-wired into the services and one has to be careful when factoring them out not to introduce additional bugs.
+
+\item Mechanisms that are only exercised according to hard-wired policies and not in their full generality have hidden bugs that are revealed as soon as those mechanisms are governed by different policies.
+In our experience introducing programmability into a storage system proved to be a great debugging tool.
+
+\item Programmability, especially in live systems, implies changes that need to be carefully managed by the system itself, including versioning and propagation of those changes without affecting correctness.
+
+\end{itemize}
 

### 2016-10-03 paper/paper.tex
paper: jeff's section 4 comments; start decoupling Ceph
Signed-off-by: Michael Sevilla <michael.sevilla@hpe.com>

@@
             pdfauthor={},
-            pdftitle={Malacology: A Programmable Storage System Built on Ceph},
+            pdftitle={Malacology: A Programmable Storage System},
             colorlinks=true,
@@
 
-\title{Malacology: A Programmable Storage System Built on Ceph}
+\title{Malacology: A Programmable Storage System}
 \author{Paper \# 123}
@@
 that leverages flash devices. The evaluation demonstrates that our services 
-inherit desirable qualities of Ceph, including the ability to make trade-offs 
+inherit desirable qualities of the back-end storage system, including the ability to make trade-offs 
 for lease capabilities, balance load (2\(\times\) improvement), propagate interfaces 
@@
 \includegraphics{figures/overview.png}
-\caption{A Ceph installation consists of a cluster running mostly
-storage daemons (OSDs) to ensure data durability, a few monitor daemons
-(MONs) to reach consensus of the cluster state and to ensure its
-consistent versioning, and a few metadata service daemons (MDSs) for
-scalable metadata service. Ceph's APIs include object, block, file
-abstractions. Malacology enables the programmability of these
-abstractions and services (indicated by bold arrows). To illustrate Malacology we implement two
-services on Ceph, ZLog and Mantle, by programming Ceph's durability,
-consistent versioning, consensus, and metadata subsystems.
-\label{fig:overview}}
+
+\caption{Scalable storage systems consist of object storage daemons (OSD) which
+store data fragments, monitor daemons (MON) that maintain cluster state, and
+service-specific daemons such as file system metadata servers (MDS). Malacology
+enables the programmability of internal abstractions and services (indicated by
+bold arrows) in order to leverage the back-end system's durability, consistent
+versioning, consensus, and metadata subsystems.  With Malacology, we add two
+new services, ZLog and Mantle, that sit alongside the traditional user-facing
+APIs (file, block, and object) in the storage stack.  \label{fig:overview}}
 \end{figure}
@@
 
-We will now discuss the subsystems Ceph uses to provide a general-purpose
+We will now discuss the common subsystems used to manage
 storage system and how Malacology makes them programmable.
@@
 
-\subsection{Mantle: A Programmable Metadata Load Balancer}
+\subsection{Mantle: Programmable Load Balancer}
 \label{sec:mantle}
 
-Mantle is a programmable metadata
-balancer for Ceph that separates the metadata balancing policies from their
-mechanisms. Administrators inject code to change how the metadata cluster
-distributes metadata. In ~\cite{sevilla:sc15-mantle} the authors showed how 
-using Mantle, one can implement a single node metadata service, a distributed 
-metadata services with hashing, and a distributed metadata service with dynamic 
-subtree partitioning.
-
-\subsection{Re-Implementation}
-
-The Ceph team wants to merge Mantle because the scriptability is useful for
-debugging, controlling the metadata balancer, and examining trade-offs for
-different balancers. Unfortunately, this research quality system is not as
-robust as Ceph and the Ceph team wants more safety, durability, and consistency
-for the new services.  The API for Malacology is the same as the one exposed
-in~\cite{sevilla:sc15-mantle} but the hooks are put in different places.
-
-\subsection{Policy Versioning}
-
-For Mantle, we added new monitor command to change the metadata load balancers:
-
-\noindent \texttt{ mds\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}
-%\noindent \texttt{ceph\ osd\ pool\ set-class\ \textless{}pool\textgreater{}\ \textless{}class\textgreater{}\ \textless{}script\textgreater{}}
-
-This command assigns the current version of the metadata load balancer.
-In our case, it is simply the name of the balancer.
-The MONs take the version and store it in the MDS map.
-Similar to the PG map or MON maps, the MDS map has all the information about the topology and state of the MDS cluster.
-Leveraging the MONs helps ease the burden on the MDS to propagate the policy and ensure that everyone agrees.
-
-There are Ceph components that are obstacles to this approach, namely the size of the  MDS map, API changes, and security.
-The Ceph engineers want to avoid using the maps to exchange information; the maps should be lightweight and fast because the MONs need to be able to version MDS cluster state quickly.
-API changes can also be a burden, especially for the MDS map, since all clients need to have the same version of the CephFS library in order to properly interact with CephFS.
-Finally, Ceph tries to avoid injecting strings into the cluster because they are insecure and buggy.
-
-\subsection{Making Policies Durable}
-
-While storing the code snippets that define the policies in the MDS map is 
-possible, it is not recommended as it is an expensive operation (a PAXOS voting 
-round). Thus, Malacology only stores policies IDs which correspond to an object 
-in a system-maintained pool. When MDSs read a policy, they first check latest 
-version from the MDS map, and then dereference the pointer by reading the 
-corresponding object in RADOS.
+Mantle is a programmable load balancer that separates the metadata balancing
+policies from their mechanisms. Administrators inject code to change how the
+metadata cluster distributes metadata. In ~\cite{sevilla:sc15-mantle} the
+authors showed how using Mantle, one can implement a single node metadata
+service, a distributed metadata services with hashing, and a distributed
+metadata service with dynamic subtree partitioning.
+
+The original implementation was ``hard-coded" into Ceph and lacked robustness
+(lacking things such as versioning, durability, or reliable policy
+distribution).  Re-implemented using Malacology, Mantle now enjoys (1) the
+versioning of MONs and (2) the durability and distribution provided by RADOS.
+The version of the load balancer, which is stored in the MDS map, corresponds
+to an object name in RADOS that houses the load balancer code.
+
+\subsubsection{Versioning Balancer Policies}
+
+Ensuring that the version of the current load balancer is consistent across the
+MDS cluster was ignored in the original implementation. The user had to set the
+version on each individual MDS and it was trivial to make the version
+inconsistent across MDS nodes.
+
+With Malacology, Mantle stores the version of the current load balancer in the
+MDS map.  Similar to the PG map or MON maps, the MDS map has all the
+information about the topology and state of the MDS cluster.  Recall that the
+MONs propogate these maps across the cluster and ensure that the map versions
+are consistent with PAXOS.  While storing the code snippets that define the
+policies in the MDS map is possible, it is not recommended as it is an
+expensive operation (a PAXOS voting round).
+
+The user changes the version of the load balancer using a new CLI command:
+
+\noindent \texttt{ fs\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}
+
+%There are Ceph components that are obstacles to this approach, namely the size
+%of the  MDS map, API changes, and security. The Ceph engineers want to avoid
+%using the maps to exchange information; the maps should be lightweight and fast
+%because the MONs need to be able to version MDS cluster state quickly. API
+%changes can also be a burden, especially for the MDS map, since all clients
+%need to have the same version of the CephFS library in order to properly
+%interact with CephFS. Finally, Ceph tries to avoid injecting strings into the
+%cluster because they are insecure and buggy.
+
+\subsubsection{Making Balancer Policies Durable}
+
+The load balancer version described above corresponds to the name of an object
+in RADOS that holds the actual Lua balancing code. The objects containing load
+balancer code are stored in a system-maintained group of objects (i.e. a pool).
+When MDS nodes start balancing load, they first check the latest version from
+the MDS map and compare it to the balancer they have loaded. If the version has
+changed, they dereference the pointer to the balancer version by reading the
+corresponding object in RADOS. This is in contrast to the original balancer
+which stored load balancer code on the local file system -- a technique which
+is unreliable and results in inconsistencies if any MDS accidently modifies the
+code.
+
+The balancer pulls the Lua code from RADOS synchronously; asynchronous reads
+are not possible because of the architecture of the MDS. The synchronous
+behavior is not the default behavior for RADOS operations, so we achieve this
+with a timeout: if the asynchronous read does not come back within half the
+balancing tick interval the operation is cancelled and a Connection Timeout
+error is returned. By default, the balancing tick interval is 10 seconds, so
+Mantle will use a 5 second second timeout.
+
+This design allows Mantle to immediately return an error if anything
+RADOS-related goes wrong.  We use this implementation because we do not want to
+do a blocking OSD read from inside the global MDS lock. Doing so would bring
+down the MDS cluster if any of the OSDs are not responsive.
+
+Storing the balancers in RADOS is only possible because we use Lua as the
+language for writing balancer code. If we used a language that needs to be
+compiled, like the C++ object classes in the OSD, we would need to ensure
+binary compatibility, which is complicated by different operating systems,
+distributions, and compilers.
+
+\subsubsection{Logging, Debugging, and Warnings}
+
+In the original implementation, Mantle would log all errors, warnings, and
+debug messages to the local log on each MDS. To get the simplest status
+messages or to debug problems, the user would have to log into each MDS
+individually, look at the logs, and reason about their causality.
+
+With Malacology, Mantle re-usues the centralized logging features of the MON
+subsystem. Important errors, warnings, and info messages are sent with the MON
+beacon subsystem and appear in the monitor cluster log so instead of users
+going to each node, they can watch messages appear at the monitor. Messages are
+used discretly, so as not to spam the monitor with frivolous debugging but the
+large events, like balancer version changes or failed subsystems show up in the
+centralized log.
 

### 2016-10-04 paper/paper.tex
paper: fix line wrap
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 programmability by constructing two services out of existing abstractions: a
-file system metadata load balancer and a high-performance distributed shared-log
-that leverages flash devices. The evaluation demonstrates that our services 
-inherit desirable qualities of the back-end storage system, including the ability to make trade-offs 
-for lease capabilities, balance load (2\(\times\) improvement), propagate interfaces 
-(average of 222ms) and recover (15 second down time).  
+file system metadata load balancer and a high-performance distributed
+shared-log that leverages flash devices. The evaluation demonstrates that our
+services inherit desirable qualities of the back-end storage system, including
+the ability to make trade-offs for lease capabilities, balance load
+(2\(\times\) improvement), propagate interfaces (average of 222ms) and recover
+(15 second down time).  
 \end{abstract}
@@
 
-A storage system implements abstractions designed to persistently store
-data and must exhibit a high level of correctness to prevent data loss.
-Storage systems have evolved around storage devices that often were
-orders of magnitude slower than CPU and memory, and therefore could
-dominate overall performance if not used carefully. Over the last few
-decades members of the storage systems community have developed
-ingenious strategies to meet correctness requirements while somewhat
-hiding the latency of traditional storage media~\cite{brewer_disks_2016}. To avoid lock-in by a
-particular vendor, users of storage systems have preferred systems with
-highly standardized APIs and lowest common denominator abstract data
-types such as blocks of bytes and byte stream files~\cite{armbrust_view_2010}.
-
-A number of recent developments are disrupting traditional storage
-systems: (1) the falling prices of flash storage and the availability of
-new types of non-volatile memory that are orders of magnitude faster
-than traditional spinning media are moving overall performance
-bottlenecks away from storage devices to CPUs and networking, and
-pressure storage systems to shorten their code paths and incorporate new
-optimizations~\cite{gray_tape_2007,gray_flash_2008}; (2) demand for managing structured data and flexible
-consistency semantics at scale pressure big data processing systems to
-use storage abstractions that can meet these demands~\cite{apache_contributors_parquet_2014}; and (3)
-production-quality scalable storage systems available as open source
-software have established and are continuing to establish new, \emph{de-facto} API standards at a faster pace than traditional standards bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
+A storage system implements abstractions designed to persistently store data
+and must exhibit a high level of correctness to prevent data loss.  Storage
+systems have evolved around storage devices that often were orders of magnitude
+slower than CPU and memory, and therefore could dominate overall performance if
+not used carefully. Over the last few decades members of the storage systems
+community have developed ingenious strategies to meet correctness requirements
+while somewhat hiding the latency of traditional storage
+media~\cite{brewer_disks_2016}. To avoid lock-in by a particular vendor, users
+of storage systems have preferred systems with highly standardized APIs and
+lowest common denominator abstract data types such as blocks of bytes and byte
+stream files~\cite{armbrust_view_2010}.
+
+A number of recent developments are disrupting traditional storage systems: (1)
+the falling prices of flash storage and the availability of new types of
+non-volatile memory that are orders of magnitude faster than traditional
+spinning media are moving overall performance bottlenecks away from storage
+devices to CPUs and networking, and pressure storage systems to shorten their
+code paths and incorporate new
+optimizations~\cite{gray_tape_2007,gray_flash_2008}; (2) demand for managing
+structured data and flexible consistency semantics at scale pressure big data
+processing systems to use storage abstractions that can meet these
+demands~\cite{apache_contributors_parquet_2014}; and (3) production-quality
+scalable storage systems available as open source software have established and
+are continuing to establish new, \emph{de-facto} API standards at a faster pace
+than traditional standards
+bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
 
@@
 
-To illustrate the benefits and challenges of this approach we examine
-the programmability of Ceph~\cite{weil:osdi2006-ceph,ceph}, the increasingly
-popular, production-level open-source distributed storage system.
-Something of a storage swiss army knife, Ceph supports file, block, and
-object interfaces simultaneously in a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous 
-Distributed Object Storage (RADOS) system is a cluster of storage
-devices (OSDs) that provide Ceph with data durability and integrity
+To illustrate the benefits and challenges of this approach we examine the
+programmability of Ceph~\cite{weil:osdi2006-ceph,ceph}, the increasingly
+popular, production-level open-source distributed storage system.  Something of
+a storage swiss army knife, Ceph supports file, block, and object interfaces
+simultaneously in a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's
+Reliable Autonomous Distributed Object Storage (RADOS) system is a cluster of
+storage devices (OSDs) that provide Ceph with data durability and integrity
 using replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. 
@@
 
-Ceph also provides empirical evidence that developers embrace programmable storage: Ceph already supports the
-implementation of domain-specific logical object interfaces by composing
-existing interfaces. These interfaces are written in C++ and are statically
-loaded into the system. Figure~\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
-domain-specific interfaces in the Ceph community since 2010.  What is most
-remarkable is that this trend contradicts the notion that API changes are a
-burden for users.
-Rather it appears that a gap in existing interfaces are
+Ceph also provides empirical evidence that developers embrace programmable
+storage: Ceph already supports the implementation of domain-specific logical
+object interfaces by composing existing interfaces. These interfaces are
+written in C++ and are statically loaded into the system.
+Figure~\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production
+use of domain-specific interfaces in the Ceph community since 2010.  What is
+most remarkable is that this trend contradicts the notion that API changes are
+a burden for users.  Rather it appears that a gap in existing interfaces are
 being addressed through programmability. In fact, Table~\ref{table:objclasses}
@@
 \includegraphics{figures/obj-int-dev-growth.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] Since 2010, the growth in the number of co-designed object
-storage interfaces in Ceph has been accelerating. This plot is the
-number of object classes (a group of interfaces), and the total number
-of methods (the actual API end-points).}
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
+Since 2010, the growth in the number of co-designed object storage interfaces
+in Ceph has been accelerating. This plot is the number of object classes (a
+group of interfaces), and the total number of methods (the actual API
+end-points).}
 \label{fig:obj-int-dev-growth}
@@
 The popularity of the custom object interface facility of Ceph hint at three
-trends in the Ceph community: (1) increasingly, the default
-algorithms/tunables of the storage system are insufficient for the
-application's performance goals, (2) programmers are becoming more aware
-of their application's behavior, and (3) programmers know how to manage
-resources to improve performance. Programmers gravitate towards object
-interfaces because it gives them the ability to tell the storage system
-about their application: if it is CPU or IO bound, if it has locality,
-if its size has the potential to overload a single proxy node, etc. The
-programmers know what the problem is and how to solve it, but until
-the ability to modify object interfaces, they had no way to express to the storage system how to handle
-their data.
-
-To illustrate the benefits and challenges of this approach we have
-designed and evaluated Malacology, a programmable storage system for
-constructing new services by re-purposing existing subsystem
-abstractions contained in Ceph. We build the framework in Ceph by
-leveraging a broad spectrum of existing services, including distributed
-locking and caching services provided by MDSs, durability and logical
-storage devices provided by RADOS, and propagation of consistent cluster
-state provided by the monitoring service (see Figure~\ref{fig:overview}). As we will show in this paper, this framework is expressive enough to provide the
-functionality necessary for implementing new services. Our contributions
-are:
-
-\paragraph*{A programmable storage system} that re-uses and extends existing abstractions and services of Ceph, a common production-level storage system. It includes:
+trends in the Ceph community: (1) increasingly, the default algorithms/tunables
+of the storage system are insufficient for the application's performance goals,
+(2) programmers are becoming more aware of their application's behavior, and
+(3) programmers know how to manage resources to improve performance.
+Programmers gravitate towards object interfaces because it gives them the
+ability to tell the storage system about their application: if it is CPU or IO
+bound, if it has locality, if its size has the potential to overload a single
+proxy node, etc. The programmers know what the problem is and how to solve it,
+but until the ability to modify object interfaces, they had no way to express
+to the storage system how to handle their data.
+
+To illustrate the benefits and challenges of this approach we have designed and
+evaluated Malacology, a programmable storage system for constructing new
+services by re-purposing existing subsystem abstractions contained in Ceph. We
+build the framework in Ceph by leveraging a broad spectrum of existing
+services, including distributed locking and caching services provided by MDSs,
+durability and logical storage devices provided by RADOS, and propagation of
+consistent cluster state provided by the monitoring service (see
+Figure~\ref{fig:overview}). As we will show in this paper, this framework is
+expressive enough to provide the functionality necessary for implementing new
+services. Our contributions are:
+
+\paragraph*{A programmable storage system} that re-uses and extends existing
+abstractions and services of Ceph, a common production-level storage system. It
+includes:
 
@@
 %  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    An extension of file system metadata by a new type of ``sequencer'' inode that together with Ceph's capability subsystem implements high-performance, non-durable access serialization:
-  \item
-    The use of Ceph's PAXOS-based consensus infrastructure for versioning of metadata service load-balancing policies
-  \item
-    The use of storage objects for persisting policies
-  \item
-    An extension of the object interface classes for new logical storage/metadata devices
+
+\item An extension of file system metadata by a new type of ``sequencer''
+inode that together with Ceph's capability subsystem implements
+high-performance, non-durable access serialization:
+
+\item The use of Ceph's PAXOS-based consensus infrastructure for versioning
+of metadata service load-balancing policies
+
+\item The use of storage objects for persisting policies
+
+\item An extension of the object interface classes for new logical
+storage/metadata devices
+
 \end{enumerate}
 
-\paragraph*{A distributed shared log service} to demonstrate the feasibility of services using the programmable storage approach:
+\paragraph*{A distributed shared log service} to demonstrate the feasibility of
+services using the programmable storage approach:
 
 \begin{enumerate}
-%  \def\labelenumi{\arabic{enumi}.}
-%  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    A high-performance distributed shared log service called ZLog which is based on CORFU~\cite{balakrishnan_corfu_2012}
-  \item
-    A re-implementation of the programmable Mantle metadata load balancing service~\cite{sevilla:sc15-mantle} using Malacology
-  \item 
-    An application of Mantle to manage load-balancing of sequencers.
+%  \def\labelenumi{\arabic{enumi}.} \itemsep1pt\parskip0pt\parsep0pt
+\item A high-performance distributed shared log service called ZLog which is
+based on CORFU~\cite{balakrishnan_corfu_2012}
+
+\item A re-implementation of the programmable Mantle metadata load balancing
+service~\cite{sevilla:sc15-mantle} using Malacology
+
+\item An application of Mantle to manage load-balancing of sequencers.
+
 \end{enumerate}
@@
 (\S\ref{sec:malacology}). Next we describe the services that we have
-constructed within the Malacology framework (\S\ref{sec:services}),
-and evaluate our ideas within our prototype implementation
-(\S\ref{evaluation}). Finally we discuss related work and conclude.
+constructed within the Malacology framework (\S\ref{sec:services}), and
+evaluate our ideas within our prototype implementation (\S\ref{evaluation}).
+Finally we discuss related work and conclude.
 
@@
 %
-%The challenges of introducing programmability into existing storage systems: (1) Programmability while keeping storage system available, (2) Finding and factoring out hard-wired policies and optimizations, (3) Bugs that were hidden because mechanisms were only exercised according to hard-wired policies and not in their full generality they were thought of -- programmability into a storage system turned out to be a great debugging tool, (4) version control and propagation of updates throughout the system without interrupting production workloads, (5) ...
+%The challenges of introducing programmability into existing storage systems:
+%(1) Programmability while keeping storage system available, (2) Finding and
+%factoring out hard-wired policies and optimizations, (3) Bugs that were hidden
+%because mechanisms were only exercised according to hard-wired policies and
+%not in their full generality they were thought of -- programmability into a
+%storage system turned out to be a great debugging tool, (4) version control
+%and propagation of updates throughout the system without interrupting
+%production workloads, (5) ...
+
 \section{Challenges}
@@
 
-Establishing the infrastructure for programmability into existing services and abstractions of distributed storage systems is challenging, even if one assumes that the source code of the storage system and the necessary expertise for understanding it is available:
+Establishing the infrastructure for programmability into existing services and
+abstractions of distributed storage systems is challenging, even if one assumes
+that the source code of the storage system and the necessary expertise for
+understanding it is available:
 
@@
 
-\item Storage systems are generally required to be highly available so that any complete restarts of the storage system to reprogram them is usually unacceptable.
-
-\item Policies and optimizations are usually hard-wired into the services and one has to be careful when factoring them out not to introduce additional bugs.
-
-\item Mechanisms that are only exercised according to hard-wired policies and not in their full generality have hidden bugs that are revealed as soon as those mechanisms are governed by different policies.
-In our experience introducing programmability into a storage system proved to be a great debugging tool.
-
-\item Programmability, especially in live systems, implies changes that need to be carefully managed by the system itself, including versioning and propagation of those changes without affecting correctness.
+\item Storage systems are generally required to be highly available so that any
+complete restarts of the storage system to reprogram them is usually
+unacceptable.
+
+\item Policies and optimizations are usually hard-wired into the services and
+one has to be careful when factoring them out not to introduce additional bugs.
+
+\item Mechanisms that are only exercised according to hard-wired policies and
+not in their full generality have hidden bugs that are revealed as soon as
+those mechanisms are governed by different policies. In our experience
+introducing programmability into a storage system proved to be a great
+debugging tool.
+
+\item Programmability, especially in live systems, implies changes that need to
+be carefully managed by the system itself, including versioning and propagation
+of those changes without affecting correctness.
 
@@
 
-Malacology is both a prototype for a programmable storage system and a design approach to evolve storage systems efficiently and without
-jeopardizing years of code-hardening and performance optimization
-efforts.
-The guiding principle is to re-use existing services and extend them so that these services can be \emph{programmed}.
-We accomplish programmability of a service by exporting bindings (or ``hooks'') for an interpreted programming language so that programming can occur without having to restart the storage system (see also below,~\S\ref{sec:durability}).
+Malacology is both a prototype for a programmable storage system and a design
+approach to evolve storage systems efficiently and without jeopardizing years
+of code-hardening and performance optimization efforts. The guiding principle
+is to re-use existing services and extend them so that these services can be
+\emph{programmed}. We accomplish programmability of a service by exporting
+bindings (or ``hooks'') for an interpreted programming language so that
+programming can occur without having to restart the storage system (see also
+below,~\S\ref{sec:durability}).
 
 There are multiple reasons that make Lua an attractive runtime for implementing
-Malacology. Lua is a portable, embedded
-scripting language that offers superior performance and productivity
-trade-offs, including a JIT-based implementation that is well known for near
-native performance. Additionally, Lua has been used extensively in game engines, 
-and systems research \cite{neto:dls14-luaos}, including storage systems where it 
-has been effectively used both on
-\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010} and 
-off
-\cite{sevilla:sc15-mantle} the performance critical path. Finally, the 
+Malacology. Lua is a portable, embedded scripting language that offers superior
+performance and productivity trade-offs, including a JIT-based implementation
+that is well known for near native performance. Additionally, Lua has been used
+extensively in game engines, and systems research \cite{neto:dls14-luaos},
+including storage systems where it has been effectively used both on
+\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010}
+and off \cite{sevilla:sc15-mantle} the performance critical path. Finally, the
 flexibility of the runtime allows to sandbox applications in order to address
@@
 
-We will now discuss the common subsystems used to manage
-storage system and how Malacology makes them programmable.
+We will now discuss the common subsystems used to manage storage system and how
+Malacology makes them programmable.
 
@@
 {\bf Safely Loading Classes.} The original Mantle API is fragile: the API is
-not enforced by the runtime, the system allows strings to be injected through the
-admin daemon, constructing the Lua script in C++ is clunky, and the
+not enforced by the runtime, the system allows strings to be injected through
+the admin daemon, constructing the Lua script in C++ is clunky, and the
 administrator can inject really bad policies (e.g., while 1) that brings the
@@
 {\bf Policy management.} A new interface is added to the metadata service
-subsystem using the monitoring framework to support management of interfaces. This 
-ensures that the correct balancer version is consistently distributed and verified on
-all MDS nodes. Policies are also durably saved with Ceph's replication and redundancy.
+subsystem using the monitoring framework to support management of interfaces.
+This ensures that the correct balancer version is consistently distributed and
+verified on all MDS nodes. Policies are also durably saved with Ceph's
+replication and redundancy.
 
@@
 
-\paragraph*{Malacology:} Metadata service policies and object storage interfaces are stored durability
-within RADOS and managed by storing references with the object maps. Since
-the cluster already propagates a consistent view of these data structures,
-we use this service to automatically install interfaces in OSDs, and install
-policies within the MDS such that clients and daemons are synchronized on
-correct implementations without restarting.
+\paragraph*{Malacology:} Metadata service policies and object storage
+interfaces are stored durability within RADOS and managed by storing references
+with the object maps. Since the cluster already propagates a consistent view of
+these data structures, we use this service to automatically install interfaces
+in OSDs, and install policies within the MDS such that clients and daemons are
+synchronized on correct implementations without restarting.
 
@@
 \end{tabular}
-\caption{To implement the Mantle and ZLog services, we leverage Ceph components. Malacology uses these subsystems because they provide powerful abstractions.}
+\caption{To implement the Mantle and ZLog services, we leverage Ceph
+components. Malacology uses these subsystems because they provide powerful
+abstractions.}
 \label{table:implementation}
@@
 
-\section{Evaluation}
-\label{evaluation}
-We have conducted our evaluation on CloudLab as well as our own 20 node cluster.
-We follow the ``Popper Convention"~\cite{anonymous} so as to try and make our experiments as reproducible as possible.
-We have a GitHub repository with our source code, experimental environment, configuration parameters, and workloads but due to the double-blind policy, we have temporarily disabled the links to our graphs and experimental details.
+\section{Evaluation} \label{evaluation} We have conducted our evaluation on
+CloudLab as well as our own 20 node cluster. We follow the ``Popper
+Convention"~\cite{anonymous} so as to try and make our experiments as
+reproducible as possible. We have a GitHub repository with our source code,
+experimental environment, configuration parameters, and workloads but due to
+the double-blind policy, we have temporarily disabled the links to our graphs
+and experimental details.
 
 Much of this work focuses on the programmability of Ceph and the services that
-can be layered on top of Malacology. As such, we focus on the performance of the 
-components and subsystems exposed by the framework. Detailed system-wide performance
-measurements and analysis is forthcoming for both Mantle and ZLog, as we try
-to find the best metadata balancing policies and physical design implementation trade-offs
-respectively.
+can be layered on top of Malacology. As such, we focus on the performance of
+the components and subsystems exposed by the framework. Detailed system-wide
+performance measurements and analysis is forthcoming for both Mantle and ZLog,
+as we try to find the best metadata balancing policies and physical design
+implementation trade-offs respectively.
 
@@
 \includegraphics{figures/caps-delay-thruput.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] Sequencer throughput by re-using various services.
-The highest performance is achieved using a single client with
-exclusive, cacheable privilege. Round-robin sharing of the sequencer
-resource is affected by the amount of time the resource is held,
-with best-effort performing the worst.}
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
+Sequencer throughput by re-using various services.  The highest performance is
+achieved using a single client with exclusive, cacheable privilege. Round-robin
+sharing of the sequencer resource is affected by the amount of time the
+resource is held, with best-effort performing the worst.}
 \label{fig:captp}
@@
 \includegraphics{figures/caps-delay-latency.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The latency distribution shows that if the clients hold their capabilities longer, performance decreases.
-Changing the delay via policy can help the sequencer and clients strike a balance between throughput and latency.}
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
+latency distribution shows that if the clients hold their capabilities longer,
+performance decreases. Changing the delay via policy can help the sequencer and
+clients strike a balance between throughput and latency.}
 \label{fig:capcdf}
@@
 \includegraphics[trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] Cluster-wide interface update latency, excluding the Paxos proposal cost for
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
+Cluster-wide interface update latency, excluding the Paxos proposal cost for
 interface commit. Each of the 120 OSDs used a RAM-based object store, while the
@@
 propagate application-specific state that needs to be highly-available and
-strongly consistent.
-The real cost of interface management, in addition to cluster-wide propagation
-of interface updates, includes the network round-trip to the interface
-management service, the PAXOS commit protocol, and other factors such as system
-load. By default PAXOS proposals occur periodically with a 1 second interval in
-order to accumulate updates. In a minimum, realistic quorum of 3 monitors using
-HDD-based storage, we were able to decrease this interval to an average of 222
-ms.
+strongly consistent.  The real cost of interface management, in addition to
+cluster-wide propagation of interface updates, includes the network round-trip
+to the interface management service, the PAXOS commit protocol, and other
+factors such as system load. By default PAXOS proposals occur periodically with
+a 1 second interval in order to accumulate updates. In a minimum, realistic
+quorum of 3 monitors using HDD-based storage, we were able to decrease this
+interval to an average of 222 ms.
 
@@
 \includegraphics{figures/mantle-seq-thruput.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] A ZLog sequencer that uses the CephFS MDS enjoys the load balancing
-capabilities of the metadata cluster. The ``Sequencer Baseline'' has no
-background sequencers, while the ``2 Seq'' curves are the throughput
-of 1 sequencer when an additional sequencer is running in the background
-generating load.}\label{fig:mantle-seq-thruput}
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] A
+ZLog sequencer that uses the CephFS MDS enjoys the load balancing capabilities
+of the metadata cluster. The ``Sequencer Baseline'' has no background
+sequencers, while the ``2 Seq'' curves are the throughput of 1 sequencer when
+an additional sequencer is running in the background generating
+load.}\label{fig:mantle-seq-thruput}
 \end{figure}
@@
 
-We use a balancer policy specifically designed for the sequencer. The balancer aggressively sheds half its load to its neighbor as soon as it can.  It is based
-off the ``Greedy Spill Even" approach in the Mantle~\cite{sevilla:sc15-mantle} paper
-but is reproduced here for the reader's convenience.
+We use a balancer policy specifically designed for the sequencer. The balancer
+aggressively sheds half its load to its neighbor as soon as it can.  It is
+based off the ``Greedy Spill Even" approach in the
+Mantle~\cite{sevilla:sc15-mantle} paper but is reproduced here for the reader's
+convenience.
 
@@
 \end{lstlisting}
-    \caption{The sequencer balancing policy that works best for bursty workloads. \label{listing:greedy-spill}}
+\caption{The sequencer balancing policy that works best for bursty workloads.
+\label{listing:greedy-spill}}
 \end{figure}
@@
 \includegraphics{figures/mantle-cpu-utilization-annotate.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The sequencer using Mantle in Figure~\ref{fig:mantle-seq-thruput}
-migrates to a new MDS immediately because its uses a policy designed for bursty
-workloads. This is a trace of the same workload running twice: once with a
-single MDS and once with Mantle.}
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
+sequencer using Mantle in Figure~\ref{fig:mantle-seq-thruput} migrates to a new
+MDS immediately because its uses a policy designed for bursty workloads. This
+is a trace of the same workload running twice: once with a single MDS and once
+with Mantle.}
 \label{fig:mantle-cpu-utilization}
@@
 
-Programmability of operating systems and networking resources, including distributed storage systems is not new, but we are not aware of work that makes generalization of existing services into programmable resourses a key principle in storage systems design.
-
-Programmable storage systems can be viewed as an infrastructure for creating abstractions to better separate policies from mechanisms.
-This idea is not new.
-Software-defined networks (SDNs) create such an abstraction by separating the control plane from the data plane (see for example~\cite{jain:sigcomm13}).
-This notion of control/data separation was also applied in software-defined storage (SDS)~\cite{arpaci:sosp01,thereska:sosp13,stefanovici:fast16}.
-Similarly, IOStack~\cite{gracia:internet16} is providing policy-based provisioning and filtering in OpenStack Swift.
-
-Another view of programmable storage systems is one of tailoring systems resources to applications~\cite{arpaci:sosp01}.
-Related work includes work around the Exokernel~\cite{engler:sosp95} and SPIN~\cite{bershad:sosp95} and Vino~\cite{seltzer:osdi96}, the latter two addressed the ability of injecting code into the kernel to specialize resource management.
-Another approach is to pass hints between the different layers of the IO stack to bridge the semantic gap between applications and storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}).
+Programmability of operating systems and networking resources, including
+distributed storage systems is not new, but we are not aware of work that makes
+generalization of existing services into programmable resourses a key principle
+in storage systems design.
+
+Programmable storage systems can be viewed as an infrastructure for creating
+abstractions to better separate policies from mechanisms. This idea is not new.
+Software-defined networks (SDNs) create such an abstraction by separating the
+control plane from the data plane (see for example~\cite{jain:sigcomm13}). This
+notion of control/data separation was also applied in software-defined storage
+(SDS)~\cite{arpaci:sosp01,thereska:sosp13,stefanovici:fast16}. Similarly,
+IOStack~\cite{gracia:internet16} is providing policy-based provisioning and
+filtering in OpenStack Swift.
+
+Another view of programmable storage systems is one of tailoring systems
+resources to applications~\cite{arpaci:sosp01}. Related work includes work
+around the Exokernel~\cite{engler:sosp95} and SPIN~\cite{bershad:sosp95} and
+Vino~\cite{seltzer:osdi96}, the latter two addressed the ability of injecting
+code into the kernel to specialize resource management. Another approach is to
+pass hints between the different layers of the IO stack to bridge the semantic
+gap between applications and
+storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}).
 
@@
 
-Programmable storage is a viable method for eliminating duplication of
-complex error prone software that are used as workarounds for storage
-system deficiencies. However, this duplication has real-world problems
-related to reliability. We propose that system expose their services in
-a safe way allowing application developers to customize system behavior
-to meet their needs while not sacrificing correctness.
-
-We are intending to pursue this work towards the goal of constructing a set
-of customization points that allow a wide variety of storage system
-services to be configured on-the-fly in existing systems. This work is
-one point along that path in which we have looked at targeting
-special-purpose storage systems. Ultimately we want to utilize
-declarative methods for expressing new services.
+Programmable storage is a viable method for eliminating duplication of complex
+error prone software that are used as workarounds for storage system
+deficiencies. However, this duplication has real-world problems related to
+reliability. We propose that system expose their services in a safe way
+allowing application developers to customize system behavior to meet their
+needs while not sacrificing correctness.
+
+We are intending to pursue this work towards the goal of constructing a set of
+customization points that allow a wide variety of storage system services to be
+configured on-the-fly in existing systems. This work is one point along that
+path in which we have looked at targeting special-purpose storage systems.
+Ultimately we want to utilize declarative methods for expressing new services.
 

### 2016-10-05 paper/paper.tex
paper: separate intro and problem
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
+To illustrate the benefits and challenges of this approach we have designed and
+evaluated Malacology, a programmable storage system for constructing new
+services by re-purposing existing subsystem abstractions of the storage stack.
+We build the framework in Ceph by leveraging a broad spectrum of existing
+services, including distributed locking and caching services provided by
+metadata servers, durability and logical storage devices provided by the
+backend object store, and propagation of consistent cluster state provided by
+the monitoring service (see Figure~\ref{fig:overview}). As we will show in this
+paper, this framework is expressive enough to provide the functionality
+necessary for implementing new services. Our contributions are:
+
+\paragraph*{A programmable storage system} that re-uses and extends existing
+abstractions and services of Ceph, a common production-level storage system. It
+includes:
+
+\begin{enumerate}
+
+\item An extension of file system metadata by a new type of ``sequencer''
+inode that together with Ceph's capability subsystem implements
+high-performance, non-durable access serialization:
+
+\item The use of Ceph's PAXOS-based consensus infrastructure for versioning
+of metadata service load-balancing policies
+
+\item The use of storage objects for persisting policies
+
+\item An extension of the object interface classes for new logical
+storage/metadata devices
+
+\end{enumerate}
+
+\paragraph*{A distributed shared log service} to demonstrate the feasibility of
+services using the programmable storage approach:
+
+\begin{enumerate}
+
+\item A high-performance distributed shared log service called ZLog which is
+based on CORFU~\cite{balakrishnan_corfu_2012}
+
+\item A re-implementation of the programmable Mantle metadata load balancing
+service~\cite{sevilla:sc15-mantle} using Malacology
+
+\item An application of Mantle to manage load-balancing of sequencers.
+
+\end{enumerate}
+
+The remainder of this paper is structured as follows. First we describe
+Malacology by presenting the subsystems within Ceph that we re-purpose, and
+briefly describe how those system are used within Malacology
+(\S\ref{sec:malacology}). Next we describe the services that we have
+constructed within the Malacology framework (\S\ref{sec:services}), and
+evaluate our ideas within our prototype implementation (\S\ref{evaluation}).
+Finally we discuss related work and conclude.
+
+\section{Highly Tailored, Application-Specific Storage Stacks
+}\label{highly-tailored-and-application-specifc-storage-systems}
+
 To illustrate the benefits and challenges of this approach we examine the
-programmability of Ceph~\cite{weil:osdi2006-ceph,ceph}, the increasingly
-popular, production-level open-source distributed storage system.  Something of
-a storage swiss army knife, Ceph supports file, block, and object interfaces
-simultaneously in a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's
-Reliable Autonomous Distributed Object Storage (RADOS) system is a cluster of
-storage devices (OSDs) that provide Ceph with data durability and integrity
-using replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. 
-% By introducing
-% programmability concepts into Ceph, we can build new services by
-% carefully exposing internal storage services to applications.
+
+Today's storage stacks are large, monolothic systems built to ensure
+scalability and persistence -- a storage system that loses user's data is
+useless. Building storage stacks from the ground up for a specific purpose
+results in the best performance. For example, HDFS was designed specifically to
+server Hadoop's jobs and relaxes POSIX constraints~\cite{CITEME}.
+Alternatively, general-purpose storage stacks are built with the flexiblity to
+serve many applications by embedding tunable parameters to control low level
+implementation details like block size. Unfortunately, users want more control
+from the general-purpose storage stacks without going as far as building their
+storage system from the ground up.
+
+% does this belong?
+To demonstarte the need for programmability in storage systems, we examine
+Ceph's programmability. Ceph~\cite{weil:osdi2006-ceph,ceph}, the increasingly
+popular, production-level open-source distributed storage system, was
+originally built to be an HPC file system. But demands for storage systems that
+could slide in behind cloud infrastructures, like OpenStack, shifted the focus
+to Ceph's internal object store as a back end for large cloud block devices,
+S3-like objects, and file system workloads.  Something of a storage swiss army
+knife, Ceph supports file, block, and object interfaces simultaneously in a
+single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
+Distributed Object Storage (RADOS) system is a cluster of storage devices
+(OSDs) that provide Ceph with data durability and integrity using replication,
+erasure-coding, and scrubbing~\cite{weil_rados_2007}. 
+% By introducing programmability concepts into Ceph, we can build new services
+% by carefully exposing internal storage services to applications.
 
@@
 to the storage system how to handle their data.
-
-To illustrate the benefits and challenges of this approach we have designed and
-evaluated Malacology, a programmable storage system for constructing new
-services by re-purposing existing subsystem abstractions contained in Ceph. We
-build the framework in Ceph by leveraging a broad spectrum of existing
-services, including distributed locking and caching services provided by MDSs,
-durability and logical storage devices provided by RADOS, and propagation of
-consistent cluster state provided by the monitoring service (see
-Figure~\ref{fig:overview}). As we will show in this paper, this framework is
-expressive enough to provide the functionality necessary for implementing new
-services. Our contributions are:
-
-\paragraph*{A programmable storage system} that re-uses and extends existing
-abstractions and services of Ceph, a common production-level storage system. It
-includes:
-
-\begin{enumerate}
-%  \def\labelenumi{\arabic{enumi}.}
-%  \itemsep1pt\parskip0pt\parsep0pt
-
-\item An extension of file system metadata by a new type of ``sequencer''
-inode that together with Ceph's capability subsystem implements
-high-performance, non-durable access serialization:
-
-\item The use of Ceph's PAXOS-based consensus infrastructure for versioning
-of metadata service load-balancing policies
-
-\item The use of storage objects for persisting policies
-
-\item An extension of the object interface classes for new logical
-storage/metadata devices
-
-\end{enumerate}
-
-\paragraph*{A distributed shared log service} to demonstrate the feasibility of
-services using the programmable storage approach:
-
-\begin{enumerate}
-%  \def\labelenumi{\arabic{enumi}.} \itemsep1pt\parskip0pt\parsep0pt
-\item A high-performance distributed shared log service called ZLog which is
-based on CORFU~\cite{balakrishnan_corfu_2012}
-
-\item A re-implementation of the programmable Mantle metadata load balancing
-service~\cite{sevilla:sc15-mantle} using Malacology
-
-\item An application of Mantle to manage load-balancing of sequencers.
-
-\end{enumerate}
-
-The remainder of this paper is structured as follows. First we describe
-Malacology by presenting the subsystems within Ceph that we re-purpose, and
-briefly describe how those system are used within Malacology
-(\S\ref{sec:malacology}). Next we describe the services that we have
-constructed within the Malacology framework (\S\ref{sec:services}), and
-evaluate our ideas within our prototype implementation (\S\ref{evaluation}).
-Finally we discuss related work and conclude.
 

### 2016-10-06 paper/paper.tex
paper: emphasize hooks and outline
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \usepackage{listings}
+\usepackage{amssymb}
+\usepackage{pifont}
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
 
@@
 
-\subsection{Cluster State Management}
+\subsection{Versioning: Cluster State Management}
 \label{sec:mon}
@@
 \label{services}
-
 \begin{table}
 \centering
-\begin{tabular}{  l | l | l    }
-\textbf{Service} & \textbf{Ceph Component} & \textbf{Used for...}  \\ \hline
-Mantle  & RADOS    & policy durability \\
-        & MON maps & policy versioning \\ \hline
-ZLog    & Logical Storage & CORFU protocol  \\ 
-        & MON maps & interface install \\ \hline
-Seq.    & MDS file types & log management \\ 
-        & MDS fault toler. & seq. fault toler. \\
+\begin{tabular}{  l | l | l | l }
+\textbf{Service} & \text{*} & \textbf{Malacology Hook} & \textbf{Used to...}  \\ \hline
+Mantle           &          & Logical Storage  & implement metadata balancers \\
+                 & \cmark   & Versioning       & agree on current balancer \\ 
+                 & \cmark   & Capabilities     & control client write lock \\
+                 &          & Durability       & persist/distribute policy logic \\\hline
+ZLog             & \cmark   & Logical Storage  & implement CORFU protocol  \\ 
+                 & \cmark   & Versioning       & agree on ZLog version \\
+                 & \cmark   & Capabilities     & file types for log management \\
+                 &          & Durability       & persist log entries \\ \hline
+Seqr.            &          & Logical Storage  & load balancing \\
+                 & \cmark   & Versioning       & fault tolerance \\
+                 & \cmark   & Capabilities     & control client bursts \\ 
 \end{tabular}
-\caption{To implement the Mantle and ZLog services, we leverage Ceph
-components. Malacology uses these subsystems because they provide powerful
-abstractions.}
+\caption{Because Mantle and ZLog are built on Malacology, they inherit the
+robustness of the internal subsystems of the underlying storage stack.\\ For
+the \text{*} column, the \cmark s are hooks contributed in this paper;
+durability is provided by librados and Lua Policy (MDS) is already
+published~\cite{sevilla:sc15-mantle}. }
 \label{table:implementation}

### 2016-10-06 paper/paper.tex
paper: first crack at decoupling Ceph from intro - did a hack job on the contributions (more general) - adopted the term underlying storage instead of the word Ceph - split the intro into 2 section to shore up motivation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-10-05 paper/paper.tex
@@
 \begin{abstract}
-Storage systems are caught between rapidly changing data processing systems and
-the increasing speed of storage devices. This puts tremendous pressure on
-storage systems to adapt both in terms of their interfaces and their
-performance. But adapting storage systems can be difficult because unprincipled
-changes might jeopardize years of code-hardening and performance optimization
-efforts that were necessary for users to entrust their data to the storage
-system.  We introduce  Malacology, a prototype programmable storage system to
-explore how existing abstractions of common services found in storage systems
-can be leveraged to address new data processing systems and the increasing
-speed of storage devices. This approach allows unprecedented flexibility for
-storage systems to evolve without sacrificing the robustness of its
-code-hardened subsystems.  We illustrate the advantages and challenges of
+Storage systems are caught between rapidly changing data
+processing systems and the increasing speed of storage devices. This puts
+tremendous pressure on storage systems to adapt both in terms of their
+interfaces and their performance. But adapting storage systems can be difficult
+because unprincipled changes might jeopardize years of code-hardening and
+performance optimization efforts that were necessary for users to entrust their
+data to the storage system.  We introduce  Malacology, a prototype programmable
+storage system to explore how existing abstractions of common services found in
+storage systems can be leveraged to address new data processing systems and the
+increasing speed of storage devices. This approach allows unprecedented
+flexibility for storage systems to evolve without sacrificing the robustness of
+its code-hardened subsystems.  We illustrate the advantages and challenges of
 programmability by constructing two services out of existing abstractions: a
@@
 services inherit desirable qualities of the back-end storage system, including
-the ability to make trade-offs for lease capabilities, balance load
-(2\(\times\) improvement), propagate interfaces (average of 222ms) and recover
-(15 second down time).  
+the ability to make trade-offs for lease capabilities, balance load, propogate
+interfaces, and recover from failure.
 \end{abstract}
@@
 
-A number of recent developments are disrupting traditional storage systems: (1)
-the falling prices of flash storage and the availability of new types of
+A number of recent developments are disrupting traditional storage systems.
+First, the falling prices of flash storage and the availability of new types of
 non-volatile memory that are orders of magnitude faster than traditional
@@
 code paths and incorporate new
-optimizations~\cite{gray_tape_2007,gray_flash_2008}; (2) demand for managing
-structured data and flexible consistency semantics at scale pressure big data
-processing systems to use storage abstractions that can meet these
-demands~\cite{apache_contributors_parquet_2014}; and (3) production-quality
+optimizations~\cite{gray_tape_2007,gray_flash_2008}. Second, the demand for
+managing structured data and flexible consistency semantics at scale pressure
+big data processing systems to use storage abstractions that can meet these
+demands~\cite{apache_contributors_parquet_2014}. Finally, production-quality
 scalable storage systems available as open source software have established and
@@
 
-These three trends put evolutionary pressure on storage systems and
-raise the question whether there are principles that storage systems
-designers can follow to evolve storage systems efficiently and without
-jeopardizing years of code-hardening and performance optimization
-efforts that are important for users to continue to entrust their data
-to the storage system.
-
-In this paper we investigate an approach that focuses on generalizing
-existing storage system resources, services, and abstractions that in
-generalized form can be used to \emph{program} new services. By doing so
-one can reuse subsystems and leverage their optimizations,
-established correctness, robustness, and efficiency. We will refer to
-this programmability as \emph{programmable storage}, which differs from
-\emph{active storage} (the injection and execution of arbitrary code in
-a storage system) and \emph{software-defined storage} (the control of
-thin-provisioning of storage).
+These three trends put evolutionary pressure on storage systems and raise the
+question whether there are principles that storage systems designers can follow
+to evolve storage systems efficiently and without jeopardizing years of
+code-hardening and performance optimization efforts that are important for
+users to continue to entrust their data to the storage system.
+
+In this paper we investigate an approach that focuses on generalizing existing
+storage system resources, services, and abstractions that in a generalized form
+can be used to \emph{program} new services. By doing so one can reuse
+subsystems and leverage their optimizations, established correctness,
+robustness, and efficiency. We will refer to this programmability as
+\emph{programmable storage}, which differs from \emph{active storage} (the
+injection and execution of arbitrary code in a storage system) and
+\emph{software-defined storage} (the control of thin-provisioning of storage).
 
@@
 services by re-purposing existing subsystem abstractions of the storage stack.
-We build the framework in Ceph by leveraging a broad spectrum of existing
-services, including distributed locking and caching services provided by
-metadata servers, durability and logical storage devices provided by the
-backend object store, and propagation of consistent cluster state provided by
-the monitoring service (see Figure~\ref{fig:overview}). As we will show in this
-paper, this framework is expressive enough to provide the functionality
-necessary for implementing new services. Our contributions are:
+We build the framework in Ceph, a popular, open source software storage stack,
+by leveraging a broad spectrum of existing services, including distributed
+locking and caching services provided by metadata servers, durability and
+logical storage devices provided by the backend object store, and propagation
+of consistent cluster state provided by the monitoring service (see
+Figure~\ref{fig:overview}). As we will show in this paper, this framework is
+expressive enough to provide the functionality necessary for implementing new
+services. Our contributions are:
 
 \paragraph*{A programmable storage system} that re-uses and extends existing
-abstractions and services of Ceph, a common production-level storage system. It
-includes:
+abstractions and services of the underlying storage stack. It
+includes the following building blocks:
 
@@
 
-\item An extension of file system metadata by a new type of ``sequencer''
-inode that together with Ceph's capability subsystem implements
-high-performance, non-durable access serialization:
-
-\item The use of Ceph's PAXOS-based consensus infrastructure for versioning
-of metadata service load-balancing policies
-
-\item The use of storage objects for persisting policies
-
-\item An extension of the object interface classes for new logical
-storage/metadata devices
+\item Versioning interface that uses the underyling storage stack's consensus
+infrastructure
+
+\item Logical Device interface for adding functionality to daemons that manage
+storage/metadata devices, allowing users to push computation to the data
+
+\item An interface for mediating access to a shared resources, which exposes
+serialization primatives using capabilities, caching, and batching
+
+\item Load Balancing interface for managing resources across the cluster in
+multi-tenant environments
+
+\item Durability interface for persisting policies using the underyling storage
+stack's object store
 
@@
 
-\paragraph*{A distributed shared log service} to demonstrate the feasibility of
-services using the programmable storage approach:
+\paragraph*{Two distributed services} that demonstrate the feasibility of using
+the programmable storage approach:
 
@@
 
-\item An application of Mantle to manage load-balancing of sequencers.
-
 \end{enumerate}
 
-The remainder of this paper is structured as follows. First we describe
-Malacology by presenting the subsystems within Ceph that we re-purpose, and
-briefly describe how those system are used within Malacology
-(\S\ref{sec:malacology}). Next we describe the services that we have
-constructed within the Malacology framework (\S\ref{sec:services}), and
-evaluate our ideas within our prototype implementation (\S\ref{evaluation}).
-Finally we discuss related work and conclude.
+The remainder of this paper is structured as follows. First, we describe and
+motivate the need for programmable storage by describing current practices in
+the open source community. Next we describe Malacology by presenting the
+subsystems within the underlying storage system that we re-purpose, and briefly
+describe how those system are used within Malacology (\S\ref{sec:malacology}).
+Then we describe the services that we have constructed within the Malacology
+framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
+implementation (\S\ref{evaluation}).  Finally we discuss related work and
+conclude.
 
@@
 
-To illustrate the benefits and challenges of this approach we examine the
-
-Today's storage stacks are large, monolothic systems built to ensure
-scalability and persistence -- a storage system that loses user's data is
-useless. Building storage stacks from the ground up for a specific purpose
-results in the best performance. For example, HDFS was designed specifically to
-server Hadoop's jobs and relaxes POSIX constraints~\cite{CITEME}.
-Alternatively, general-purpose storage stacks are built with the flexiblity to
-serve many applications by embedding tunable parameters to control low level
-implementation details like block size. Unfortunately, users want more control
-from the general-purpose storage stacks without going as far as building their
-storage system from the ground up.
+Building storage stacks from the ground-up for a specific purpose results in
+the best performance. For example, HDFS was designed specifically to serve
+Hadoop's jobs, and uses techniques like relaxing POSIX constraints to achieve
+the best performance~\cite{CITEME}. Alternatively, general-purpose storage
+stacks are built with the flexiblity to serve many applications by exposing
+tunable parameters to control low level implementation details like block size.
+Unfortunately, users want more control from the general-purpose storage stacks
+without going as far as building their storage system from the ground up.
 
 % does this belong?
-To demonstarte the need for programmability in storage systems, we examine
-Ceph's programmability. Ceph~\cite{weil:osdi2006-ceph,ceph}, the increasingly
-popular, production-level open-source distributed storage system, was
-originally built to be an HPC file system. But demands for storage systems that
-could slide in behind cloud infrastructures, like OpenStack, shifted the focus
-to Ceph's internal object store as a back end for large cloud block devices,
-S3-like objects, and file system workloads.  Something of a storage swiss army
-knife, Ceph supports file, block, and object interfaces simultaneously in a
-single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
+%Ceph~\cite{weil:osdi2006-ceph,ceph}, the increasingly
+%popular, production-level open-source distributed storage system, was
+%originally built to be an HPC file system. But demands for storage systems that
+%could slide in behind cloud infrastructures, like OpenStack, shifted the focus
+%to Ceph's internal object store as a backend for large cloud block devices,
+%S3-like objects, and file system workloads.
+
+To demonstrate this trend in storage systems, we examine Ceph's
+programmability. Something of a storage swiss army knife, Ceph supports file,
+block, and object interfaces simultaneously in a single
+cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
 Distributed Object Storage (RADOS) system is a cluster of storage devices
 (OSDs) that provide Ceph with data durability and integrity using replication,
-erasure-coding, and scrubbing~\cite{weil_rados_2007}. 
+erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already has some
+programmability; the OSDs support the implementation of domain-specific logical
+object interfaces by composing existing interfaces. These interfaces are
+written in C++ and are statically loaded into the system.
+
 % By introducing programmability concepts into Ceph, we can build new services
@@
 
-Ceph also provides empirical evidence that developers embrace programmable
-storage: Ceph already supports the implementation of domain-specific logical
-object interfaces by composing existing interfaces. These interfaces are
-written in C++ and are statically loaded into the system.
-Figure~\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production
-use of domain-specific interfaces in the Ceph community since 2010.  What is
-most remarkable is that this trend contradicts the notion that API changes are
-a burden for users.  Rather it appears that a gap in existing interfaces are
-being addressed through programmability. In fact, Table~\ref{table:objclasses}
-categorizes existing interfaces and we clearly see a trend towards reusable
-services.
+The Ceph community provides empirical evidence that developers embrace
+programmable storage. Figure~\ref{fig:obj-int-dev-growth} shows a dramatic
+growth in the production use of domain-specific interfaces in the Ceph
+community since 2010.  What is most remarkable is that this trend contradicts
+the notion that API changes are a burden for users.  Rather it appears that a
+gap in existing interfaces are being addressed through programmability. In
+fact, Table~\ref{table:objclasses} categorizes existing interfaces and we
+clearly see a trend towards reusable services.
 
@@
 
-The popularity of the custom object interface facility of Ceph hint at three
-trends in the Ceph community: (1) increasingly, the default algorithms/tunables
-of the storage system are insufficient for the application's performance goals,
-(2) programmers are becoming more aware of their application's behavior, and
-(3) programmers know how to manage resources to improve performance.
-Programmers gravitate towards object interfaces because it gives them the
-ability to tell the storage system about their application: if it is CPU or IO
-bound, if it has locality, if its size has the potential to overload a single
-proxy node, etc. The programmers know what the problem is and how to solve it,
-but until the ability to modify object interfaces, they had no way to express
-to the storage system how to handle their data.
+The takeway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
+already trying to achieve programmability because their needs, whether they be
+related to performance, availability, consistency, correctness etc., are not
+satisfied by the underlying storage system. The popularity of the custom object
+interface facility of Ceph could be due to a number of reasons, such as the
+default algorithms/tunables of the storage system are insufficient for the
+application's performance goals, programmers want to exploit
+application-specific semantics, and/or programmers know how to manage resources
+to improve performance. Their solution is a way to work around the
+traditionally rigid storage APIs because custome object interfaces give
+programmers the ability to tell the storage system about their application: if
+it is CPU or IO bound, if it has locality, if its size has the potential to
+overload a single proxy node, etc.  The programmers know what the problem is
+and how to solve it, but until the ability to modify object interfaces, they
+had no way to express to the storage system how to handle their data.
+
+Our approach is to expose more of the commonly used, code-hardened subsystems
+of the underlying storage system as interfaces. The intent is that these
+interfaces, which can be as simple as a redirection to the persistent data
+store or as complicated as a strongly consistent directly service, should be
+used and re-used in many contexts to implement a wide range of services. By
+making programmability a `feature', rather than a `hack' or `workaround', we
+help standardize a development process that is largely ad-hoc.
 
@@
 \label{services}
-
 \begin{table}
 \centering
-\begin{tabular}{  l | l | l    }
-\textbf{Service} & \textbf{Ceph Component} & \textbf{Used for...}  \\ \hline
-Mantle  & RADOS    & policy durability \\
-        & MON maps & policy versioning \\ \hline
-ZLog    & Logical Storage & CORFU protocol  \\ 
-        & MON maps & interface install \\ \hline
-Seq.    & MDS file types & log management \\ 
-        & MDS fault toler. & seq. fault toler. \\
+\begin{tabular}{  l | l | l | l }
+\textbf{Service} & \text{*} & \textbf{Malacology Collection of Interfaces/APIS}       & \textbf{Used to...}  \\ \hline
+Mantle           &          & Load Balancing      & load balance metadata \\
+                 & \cmark   & Versioning          & agree on current balancer \\ 
+                 & \cmark   & Batching            & control client write lock \\
+                 &          & Durability          & persist policy logic \\\hline
+ZLog             & \cmark   & Logical Device      & implement CORFU \\ 
+                 & \cmark   & Versioning          & install ZLog interfaces \\
+                 &          & Durability          & persist log entries \\ \hline
+ZLog             &          & Load balancing      & load balance sequencers \\
+Seqr.            & \cmark   & Batching            & guard shared resource \\ %\hline
+                 %\multicolumn{4}{c}{} \\ 
+                 %\multicolumn{4}{c}{\#: interface in metadata server,     \(\mp\): interface in object storage server} \\ 
 \end{tabular}
-\caption{To implement the Mantle and ZLog services, we leverage Ceph
-components. Malacology uses these subsystems because they provide powerful
-abstractions.}
+\caption{Mantle and ZLog are built using Malacology hooks so they inherit the
+robustness of the subsystems of the underlying stack. For the \text{*} column,
+the \cmark s are hooks contributed in this paper; durability is already
+provided with a client library and the MDS Logical Device is already
+published~\cite{sevilla:sc15-mantle}. }
 \label{table:implementation}

### 2016-10-09 paper/paper.tex
sty: setup eurosys latex style
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>
paper: started fixing table
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-\documentclass[10pt,twocolumn]{article}
-
-\usepackage{fullpage}
-\usepackage{times}
+%\documentclass[10pt,twocolumn]{article}
+\documentclass[preprint]{sigplanconf-eurosys}
+
+%\usepackage{fullpage}
+%\usepackage{times}
 \usepackage{multirow}
 \usepackage{arydshln}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
+%\usepackage{amssymb,amsmath}
+%\usepackage{ifxetex,ifluatex}
 \usepackage{amsmath}
-\usepackage{mathptmx}
+%\usepackage{mathptmx}
 \usepackage{graphicx}
 \usepackage{listings}
-\usepackage{amssymb}
+%\usepackage{amssymb}
 \usepackage{pifont}
 \newcommand{\cmark}{\ding{51}}%
-\newcommand{\xmark}{\ding{55}}%
-
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-
-\usepackage{graphicx}
-\usepackage[font=footnotesize,labelfont=bf]{caption}
+%\newcommand{\xmark}{\ding{55}}%
+%
+%
+%
+%% use upquote if available, for straight quotes in verbatim environments
+%\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+%
+%\usepackage[font=footnotesize,labelfont=bf]{caption}
 \usepackage{pstricks}
@@
 \makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
+%% Scale images if necessary, so that they will not overflow the page
+%% margins by default, and it is still possible to overwrite the defaults
+%% using explicit options in \includegraphics[width, height, ...]{}
 \setkeys{Gin}{width=\maxwidth,keepaspectratio}
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
+
 \usepackage[unicode=true]{hyperref}
-\fi
 \hypersetup{breaklinks=true,
@@
 \urlstyle{same}  % don't use monospace font for urls
-%\setlength{\parindent}{0pt}
-%\setlength{\parskip}{6pt plus 2pt minus 1pt}
-%\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\setcounter{secnumdepth}{5}
+%%\setlength{\parindent}{0pt}
+%%\setlength{\parskip}{6pt plus 2pt minus 1pt}
+%%\setlength{\emergencystretch}{3em}  % prevent overfull lines
+%\setcounter{secnumdepth}{5}
 
 \title{Malacology: A Programmable Storage System}
-\author{Paper \# 123}
+\authorinfo{Paper ID: 123}
+           {Total Page: 12}
 \date{}
@@
 \label{sec:malacology}
+
+\begin{table*}
+\centering
+\begin{tabular}{  l | l | l | l | l}
+\textbf{Interface}            &
+\textbf{\text{*}}             &
+\textbf{Provides...}          &
+\textbf{Example in Storage} & \\ \hline
+Load Balancing  & migration \& monitoring primitives   & \\
+Versioning      & consistency primitives               & \\ 
+Shared Resource & serialization \& batching primitives
+                & \\
+Durability      & persistence primitives
+                & \\
+Logical Device  & transaction \& atomicity primitives 
+                & Swift/ZeroVM in situ storage \& compute [1] \\ 
+\end{tabular}
+\caption{The Malacology interfaces are commonly used abstractions used
+internally in storage stacks.}
+\label{table:implementation}
+\end{table*}
+% http://www.zerovm.org/zerocloud.html
 
@@
 \centering
-\begin{tabular}{  l | l | l | l }
-\textbf{Service} & \text{*} & \textbf{Malacology Collection of Interfaces/APIS}       & \textbf{Used to...}  \\ \hline
-Mantle           &          & Load Balancing      & load balance metadata \\
-                 & \cmark   & Versioning          & agree on current balancer \\ 
-                 & \cmark   & Batching            & control client write lock \\
-                 &          & Durability          & persist policy logic \\\hline
-ZLog             & \cmark   & Logical Device      & implement CORFU \\ 
-                 & \cmark   & Versioning          & install ZLog interfaces \\
-                 &          & Durability          & persist log entries \\ \hline
-ZLog             &          & Load balancing      & load balance sequencers \\
-Seqr.            & \cmark   & Batching            & guard shared resource \\ %\hline
-                 %\multicolumn{4}{c}{} \\ 
-                 %\multicolumn{4}{c}{\#: interface in metadata server,     \(\mp\): interface in object storage server} \\ 
+\begin{tabular}{  l | l | l | l}
+\textbf{Service}              &
+\textbf{Used to...}           &
+\textbf{Malacology Interface} & \\ \hline
+Mantle & load balance metadata     & Load Balancing  \\
+       & agree on current balancer & Versioning      \\ 
+       & control client write lock & Shared Resource \\
+       & persist policy logic      & Durability      \\ \hline
+ZLog   & implement CORFU           & Logical Device  \\ 
+       & install ZLog interfaces   & Versioning      \\
+       & persist log entries       & Durability      \\ \hline
+ZLog   & load balance sequencers   & Load Balancing  \\
+Seqr.  & guard shared resource     & Shared Resource \\ 
 \end{tabular}
-\caption{Mantle and ZLog are built using Malacology hooks so they inherit the
-robustness of the subsystems of the underlying stack. For the \text{*} column,
-the \cmark s are hooks contributed in this paper; durability is already
-provided with a client library and the MDS Logical Device is already
+\caption{Mantle and ZLog are built using Malacology interfaces so they inherit the
+robustness of the subsystems of the underlying stack. The \cmark s are
+interfaces contributed in this paper; durability is already provided with a
+client library and the MDS Logical Device is already
 published~\cite{sevilla:sc15-mantle}. }

### 2016-10-09 paper/paper.tex
paper: add some chub
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \textbf{Example in Storage} & \\ \hline
-Load Balancing  & migration \& monitoring primitives   & \\
-Versioning      & consistency primitives               & \\ 
-Shared Resource & serialization \& batching primitives
-                & \\
-Durability      & persistence primitives
-                & \\
-Logical Device  & transaction \& atomicity primitives 
-                & Swift/ZeroVM in situ storage \& compute [1] \\ 
+Versioning      & \cmark & consensus \& consistency primitives  & PAXOS in Chubby [2] \\ 
+Shared Resource & \cmark & serialization \& batching primitives & \\
+Logical Device  & \cmark & transaction \& atomicity primitives  & Swift/ZeroVM in situ storage \& compute [1] \\ 
+Durability      &        & persistence primitives               & \\
+Load Balancing  &        & migration \& monitoring primitives   & IndexFS~\cite{ren:sc2014-indexfs} \& GIGA+~\cite{patil:fast2011-giga} \\
 \end{tabular}
@@
 % http://www.zerovm.org/zerocloud.html
+% https://www.usenix.org/legacy/events/nsdi11/tech/full_papers/Bolosky.pdf
+% https://www.cs.utexas.edu/users/lorenzo/corsi/cs380d/papers/paper2-1.pdf
 

### 2016-10-10 images/build.sh
benchmark: make images popper compliant
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,29 @@
+#!/bin/bash
+
+set -ex
+
+# build builder image
+. docker-cephdev/aliases.sh
+docker build -t cephbuilder/ceph:jewel docker-cephdev/builder-base
+docker build -t cephbuilder/zlog-mantle:jewel docker-zlogmantledev
+
+# pull base image from ceph (we will layer on top of this)
+docker pull ceph/daemon:tag-build-master-jewel-ubuntu-14.04
+docker tag ceph/daemon:tag-build-master-jewel-ubuntu-14.04 ceph/daemon:jewel
+
+# setup temporary directory
+SRC="/tmp/ceph-daemon"
+mkdir $SRC || true
+cd $SRC
+
+# layer custom binaries on top of the base image
+dmake \
+  -e SHA1_OR_REF="remotes/origin/zlog-mantle" \
+  -e GIT_URL="https://github.com/michaelsevilla/ceph.git" \
+  -e RECONFIGURE="true" \
+  -e BUILD_THREADS=`grep processor /proc/cpuinfo | wc -l` \
+  -e CONFIGURE_FLAGS="-DWITH_TESTS=OFF" \
+  cephbuilder/zlog-mantle:jewel
+cd -
+
+docker tag ceph-heads/remotes/origin/zlog-mantle-base ceph/zlog-mantle:jewel

### 2016-10-10 images/docker-zlogmantledev/entrypoint.sh

@@ -1,0 +1,37 @@
+#!/bin/bash
+set -e
+set -x
+
+. build-common
+
+mkdir -p build
+cd build
+
+if [ "$RECONFIGURE" = "true" ] || [ "$CLEAN" = "true" ] ; then
+  cmake -DCMAKE_INSTALL_PREFIX:PATH=/usr "$CONFIGURE_FLAGS" ..
+fi
+
+cd /ceph/build
+make -j$BUILD_THREADS
+make install DESTDIR=./install
+
+# do this so we can link client against libcephfs (hack)
+make install
+
+if [ -z "$INSTALL_HEADERS" ] ; then
+  INSTALL_HEADERS=true
+fi
+
+if [ "$INSTALL_HEADERS" = "true" ] ; then
+  mkdir --parents /ceph/build/install/usr/include/rados/
+  cp -L /ceph/src/include/rados/* /ceph/build/install/usr/include/rados/
+fi
+
+cd /ceph/src/zlog
+make
+cp seq-client /ceph/build/install/usr/bin
+
+export CEPH_INSTALL_PATH=/ceph/build/install
+. generate_daemon_image
+
+exit 0

### 2016-10-10 images/push.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+
+set -ex
+
+if [ -z "$REGISTRY_IP" ]; then
+  echo "oops, you did not specify a registry IP"
+  exit 0
+fi
+
+docker tag ceph/zlog-mantle:jewel $REGISTRY_IP:5000/ceph/zlog-mantle:jewel
+docker push $REGISTRY_IP:5000/ceph/zlog-mantle:jewel

### 2016-10-11 paper/paper.tex

--- 2016-10-09 paper/paper.tex
@@
 \centering
-\begin{tabular}{  l | l | l | l | l}
-\textbf{Interface}            &
-\textbf{\text{*}}             &
-\textbf{Provides...}          &
-\textbf{Example in Storage} & \\ \hline
-Load Balancing  & migration \& monitoring primitives   & \\
-Versioning      & consistency primitives               & \\ 
-Shared Resource & serialization \& batching primitives
-                & \\
-Durability      & persistence primitives
-                & \\
-Logical Device  & transaction \& atomicity primitives 
-                & Swift/ZeroVM in situ storage \& compute [1] \\ 
+\title{Hi}
+\begin{tabular}{  l | l | l | l }
+\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
+\multicolumn{4}{c}{} \\
+\textbf{Abstraction}                    &
+\textbf{Example in Production Systems}  &
+\textbf{Example in Ceph}                &
+\textbf{Provided Primitives}            \\ \hline
+{\it Versioning}
+  & Zookeeper coordination~\cite{hunt_zookeeper_2010}, Chubby~\cite{website:chubby}
+  & cluster state management~\cite{website:ceph-mon}
+  & consensus \& consistency
+  \\
+{\it Shared Resource}
+  & MPI collective IO, burst
+  & POSIX metadata protocols
+  & serialization \& batching
+  \\
+{\it Logical Device}
+  & Swift in situ storage/compute~\cite{website:zerocloud}
+  & object interface classes~\cite{website:cls-lua}
+  & transaction \& atomicity
+  \\
+Durability
+  & S3/Swift interfaces (RESTful API)
+  & object store library~\cite{weil_rados_2007}
+  & interface to persistence
+  \\
+Load Balancing
+  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
+  & migration in metadata cluster~\cite{weil:sc2004-dyn-metadata}
+  & migration \& monitoring
+  \\
 \end{tabular}
-\caption{The Malacology interfaces are commonly used abstractions used
-internally in storage stacks.}
-\label{table:implementation}
+\caption{The same ``internal abstractions" are common across large-scale
+systems because they provide primitives that solve general distributed systems
+problems.  Here we list examples of what these internal abstractions are used
+for in ``production systems" and in Ceph.  Malacology provides these internal
+abstractions as interfaces that higher level applications can use; {\it
+italicized abstractions} are interfaces contributed in this paper.}
+\label{table:examples}
 \end{table*}
-% http://www.zerovm.org/zerocloud.html
 
@@
 \centering
-\begin{tabular}{  l | l | l | l}
+\begin{tabular}{  l | l | l }
 \textbf{Service}              &
 \textbf{Used to...}           &
-\textbf{Malacology Interface} & \\ \hline
+\textbf{Malacology Interface} \\ \hline
 Mantle & load balance metadata     & Load Balancing  \\
@@
 \caption{Mantle and ZLog are built using Malacology interfaces so they inherit the
-robustness of the subsystems of the underlying stack. The \cmark s are
-interfaces contributed in this paper; durability is already provided with a
-client library and the MDS Logical Device is already
-published~\cite{sevilla:sc15-mantle}. }
+robustness of the subsystems of the underlying stack.}
 \label{table:implementation}

### 2016-10-11 paper/paper.tex
paper: more jeff edits for motivation/intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 storage systems can be leveraged to address new data processing systems and the
-increasing speed of storage devices. This approach allows unprecedented
+increasing speed of storage devices. This approach allows a large degree of
 flexibility for storage systems to evolve without sacrificing the robustness of
 its code-hardened subsystems.  We illustrate the advantages and challenges of
-programmability by constructing two services out of existing abstractions: a
-file system metadata load balancer and a high-performance distributed
+programmability by composing existing primitives into two new higher-level
+services: a file system metadata load balancer and a high-performance distributed
 shared-log that leverages flash devices. The evaluation demonstrates that our
 services inherit desirable qualities of the back-end storage system, including
-the ability to make trade-offs for lease capabilities, balance load, propogate
+the ability to make trade-offs for latency/throughput using lease capabilities, balance load, propogate
 interfaces, and recover from failure.
@@
 not used carefully. Over the last few decades members of the storage systems
-community have developed ingenious strategies to meet correctness requirements
+community have developed clever strategies to meet correctness requirements
 while somewhat hiding the latency of traditional storage
@@
 
-In this paper we investigate an approach that focuses on generalizing existing
+In this paper we investigate an approach that focuses on generalizing and
+composing existing
 storage system resources, services, and abstractions that in a generalized form
-can be used to \emph{program} new services. By doing so one can reuse
+can be used to \emph{program} new services. By composing higher-level services
+over existing abstractions, one can reuse
 subsystems and leverage their optimizations, established correctness,
@@
 services by re-purposing existing subsystem abstractions of the storage stack.
-We build the framework in Ceph, a popular, open source software storage stack,
-by leveraging a broad spectrum of existing services, including distributed
-locking and caching services provided by metadata servers, durability and
-logical storage devices provided by the backend object store, and propagation
-of consistent cluster state provided by the monitoring service (see
+We build the framework in Ceph, a popular, open source software storage stack.
+We choose Ceph to demonstrate the concept of programmable stoarge because it
+offers a broad spectrum of existing services, including distributed locking and
+caching services provided by metadata servers, durability and logical storage
+devices provided by the backend object store, and propagation of consistent
+cluster state provided by the monitoring service (see
 Figure~\ref{fig:overview}). As we will show in this paper, this framework is
@@
 
-\paragraph*{A programmable storage system} that re-uses and extends existing
-abstractions and services of the underlying storage stack. It
-includes the following building blocks:
+{\it \textbf{We define a programmable storage system}} to be a storage system that re-uses and extends existing
+abstractions and services of the underlying software stack. Our prototype
+includes (but is not limited to) the following building blocks:
 
@@
 
-\paragraph*{Two distributed services} that demonstrate the feasibility of using
-the programmable storage approach:
+{\it \textbf{We implement two distributed services}} using programmable storage
+to demonstrate its feasability:
 
@@
 
-\item A re-implementation of the programmable Mantle metadata load balancing
-service~\cite{sevilla:sc15-mantle} using Malacology
+\item An implementation of the programmable Mantle metadata load balancing
+service~\cite{sevilla:sc15-mantle}
 

### 2016-10-12 paper/paper.tex
paper: fix zlog sequencers with mantle evaluation - remove mds since its an implementation detail - explain the scenario - emphasize that we are testing the load balancing interface
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\item An implementation of the programmable Mantle metadata load balancing
+\item An implementation of Mantle, the programmable load balancing
 service~\cite{sevilla:sc15-mantle}
@@
   & MPI collective IO, burst
-  & POSIX metadata protocols
+  & POSIX metadata protocols~\cite{CITEME}
   & serialization \& batching
@@
   & migration in metadata cluster~\cite{weil:sc2004-dyn-metadata}
-  & migration \& monitoring
+  & migration \& sampling
+  \\
+Beacon
+  & Apache RabbitMQ~\cite{CITEME}
+  & inter-daemon messages, centralized logging
+  & debugging \& monitoring
   \\
@@
 {\bf Load balancing.} Mantle~\cite{sevilla:sc15-mantle}, a programmable
-metadata load balancer, has been re-implemented on Malacology so that its
-infrastructure can enjoy all the properties of other Ceph components. The load
-balancing logic has stayed largely the same but the infrastructure has been
+metadata load balancer, has been implemented on Malacology so that its
+infrastructure can enjoy all the properties of existing internal abstractions. The load
+balancing logic is intact but the infrastructure has been
 improved to safely load, manage, and persist its load balancing policies.
@@
 administrator can inject really bad policies (e.g., while 1) that brings the
-whole system down. Malacology exposes the important subsystems of Ceph's
-logical storage device framework so that the MDS can safely and dynamically
-load policies specified as Lua classes.
+whole system down. Using the Malacology durability and beacon interfaces,
+Mantle can safely load policies from the object store and centralize
+warnings/errors, respectively.
 
@@
 
-\section{Evaluation} \label{evaluation} We have conducted our evaluation on
-CloudLab as well as our own 20 node cluster. We follow the ``Popper
-Convention"~\cite{anonymous} so as to try and make our experiments as
-reproducible as possible. We have a GitHub repository with our source code,
-experimental environment, configuration parameters, and workloads but due to
-the double-blind policy, we have temporarily disabled the links to our graphs
-and experimental details.
-
-Much of this work focuses on the programmability of Ceph and the services that
-can be layered on top of Malacology. As such, we focus on the performance of
-the components and subsystems exposed by the framework. Detailed system-wide
-performance measurements and analysis is forthcoming for both Mantle and ZLog,
-as we try to find the best metadata balancing policies and physical design
-implementation trade-offs respectively.
-
-\subsection{ZLog}
-
+\section{Evaluation}
+\label{evaluation}
+
+We wanted to implement CORFU and we had two choices: build it from the ground
+up or do it on top of Malacology. To show the feasability of building on
+Malacology our evaluation focuses on the performance of the internal
+abstractions that support the Malacogy interfaces. First, we benchmark
+scenarios where contention is at the sequencer by examining the intefaces used
+to map ZLog onto Malacology; specifically, we describe the sequencer
+implementation and the propogation logical device interfaces. Next, we
+benchmark scenarios when there are multiple logs by using Mantle to (1)
+balance sequencers across a cluster and (2) recover from failure.
+
+Detailed system-wide performance measurements and analysis is forthcoming for
+both Mantle and ZLog, as we try to find the best metadata balancing policies
+and physical design implementation trade-offs respectively.  Since this work
+focuses on the programmabilty of Malacology, the goal of the benchmarks is to
+show that the components and subsystems that support the Malacology interfaces
+provide reasonable performance.
+
+We have conducted our evaluation on CloudLab as well as our own 20 node
+cluster. We follow the ``Popper Convention"~\cite{anonymous} so as to try and
+make our experiments as reproducible as possible. We have a GitHub repository
+with our source code, experimental environment, configuration parameters, and
+workloads but due to the double-blind policy, we have temporarily disabled the
+links to our graphs and experimental details.
+
+\subsection{Mapping ZLog onto Malacology}
+\label{mapping-zlog-onto-malacology}
 We evaluate the mapping of ZLog onto Malacology by examining the costs of
@@
 
-\subsubsection{Sequencer}
+\subsubsection{Sequencer Implementation}
 
@@
 
-\subsection{MDS as a ZLog Sequencer}
-
-The following experiments demonstrate the feasibility of using the Ceph MDS
-cluster as a sequencer. The goal of the workloads is to saturate the MDS
-cluster, so we design sequencers uses a more heavy-weight file system
-operation. The experiments use our internal cluster and there are 15 clients
-and between 1 and 3 MDS nodes.
+\subsection{ZLog and Mantle in a Multi-tenant Environment}
+
+For scenarios with multiple shared logs and many clients, co-locating
+sequencers on the same physical node is not ideal. Unfortunately, building a
+load balancer that can migrate the shared resource (in this case, the resource
+that guards the current token) would require building subsystems for migrating
+resources, monitoring the workloads, monitoring physical nodes, partitioning
+resources, and managing multiple sequencers.
+
+The following experiments demonstrate the feasibility of using the mechanisms
+of the Malacology load balancing interface (e.g., migrating, monitoring, and
+partitioning load) to (1) alleviate load when multiple sequencers overwhelm a
+physical servers and (2) recover from failure.
 
@@
 \caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] A
-ZLog sequencer that uses the CephFS MDS enjoys the load balancing capabilities
-of the metadata cluster. The ``Sequencer Baseline'' has no background
-sequencers, while the ``2 Seq'' curves are the throughput of 1 sequencer when
-an additional sequencer is running in the background generating
-load.}\label{fig:mantle-seq-thruput}
+ZLog sequencer that uses the Malacology load balancing interface enjoys the
+load balancing features of the of the metadata cluster. The ``Sequencer
+Baseline'' has no background sequencers, while the rest of the curves show the
+throughput of 1 sequencer when an additional sequencer is running in the
+background generating load.}\label{fig:mantle-seq-thruput}
 \end{figure}
 
-\subsubsection{Load Balancing}
-
-We use a balancer policy specifically designed for the sequencer. The balancer
-aggressively sheds half its load to its neighbor as soon as it can.  It is
-based off the ``Greedy Spill Even" approach in the
-Mantle~\cite{sevilla:sc15-mantle} paper but is reproduced here for the reader's
-convenience.
+\subsubsection{Load Balancing Sequencers}
+
+We evaluate the load balancing interface by showing the performance
+implications of shedding load. The experiments use our internal cluster which
+uses 10 nodes to store objects, one node to monitor the cluster, and 3 nodes
+that can accomodate sequencers. The workload is two logs, one with 7 clients
+and the other with 8 clients; we chose this workload because this load is the
+point at which a single physical server gets overloaded.
+
+We use a balancer policy specifically designed for multiple sequencers. The
+balancer aggressively sheds half its load to its neighbor as soon as it can.
+It is based off the ``Greedy Spill Even" approach in the
+Mantle~\cite{sevilla:sc15-mantle} paper but is reproduced in
+Figure~\ref{listing:greedy-spill} for the reader's convenience.
 
@@
 sequencer using Mantle in Figure~\ref{fig:mantle-seq-thruput} migrates to a new
-MDS immediately because its uses a policy designed for bursty workloads. This
-is a trace of the same workload running twice: once with a single MDS and once
+server immediately because it uses a policy designed for bursty workloads. This
+is a trace of the same workload running twice: once with a single host server and once
 with Mantle.}
@@
 
-Figure~\ref{fig:mantle-seq-thruput} gives a macro-level view of the performance
-of the Ceph components exposed by Malacology. Each curve shows the throughput
-from the perspective of one sequencer (\(y\) axis) over time (\(x\) axis).  The
-``Sequencer Baseline'' curve is one sequencer serving 7 clients.  The sequencer
-operates on an MDS cluster where balancing is turned off and there are no other
-workloads running. ``No Load Balancing" adds another sequencer of equal load
-running in the background. The performance of the background sequencer is
-omitted. The last two policies, ``Sequencer Balancing Policy" and
-``Conservative Balancing Policy", show the performance of the sequencer (with a
-background sequencer running) when using Mantle with a 3 MDS node cluster.
-These policies differ by the threshold at which they start balancing:
+Figure~\ref{fig:mantle-seq-thruput} shows a performance profile of one of the
+sequencers as the load balancing interface migrates sequencers around the
+cluster to alleviate load.  Each curve shows the throughput (\(y\) axis) over
+time (\(x\) axis).  ``Sequencer Baseline'' is the performance of one sequencer
+on one physical server serving 7 clients - there are no other sequencers or
+logs in the system and the curve represents the maximum throughput of a
+sequencer on an underutilized physical server.  ``No Load Balancing" adds a
+second sequencer of equal load to the same physical server with the intent of
+overloading the single physical server; recall that the throughput for this
+second sequencer is omitted. Compared to the baseline, the overloaded physical
+server with 2 sequencers has more than a 2\(\times\) throughput degradation.
+
+The last two curves of Figure~\ref{fig:mantle-seq-thruput} show the performance
+improvements when sequencers are load balanced across the cluster. Both curves
+are the throughput of one of the sequencers when a background sequencer is
+running. These policies differ by the threshold at which they start balancing:
 ``Sequencer Balancing Policy" starts balancing immediately and ``Conservative
-Balancing Policy" waits for 120 seconds.
+Balancing Policy" waits for 120 seconds before migrating sequencers.
+
+The variation of the sequencer throughputs helps us determine which sequencers
+are on overloaded physical servers.  ``Sequencer Baseline" is the best
+performance for this particular workload because there is no contention.
+Performance is stable and the MDS is under-loaded because we do not see the
+spike at the end like the other curves.  ``No Load Balancing" shows an
+overloaded sequencer because the performance of the monitored sequencer is low
+and unstable (high variation in performance). We also deduce that it is
+overloaded by comparing to the ``Sequencer Balancing Policy" and ``Conservative
+Balancing Policy" curves, which show improved performance when load is
+migrated.
+
+%Overall, the clients achieve lower performance than the capability experiments
+%#in Figure~\ref{fig:captp} because the sequencer uses a slower metadata
+%operation to manage the capability.  It appears the bottleneck is somewhere
+%with the clients -- perhaps it is our 1GBit/s network link.
 
@@
 The curves are the throughput (\(y\) axis) over time (\(x\) axis) and the graph
-has been partitioned into the two cluster setups. The CPU utilization is
-confined to one core the MDS is not multi-threaded, as it has been shown that
-metadata performance does not improve with multiple
-threads~\cite{konstantinos:pdsw2014-lustre-metadata}.
-
-Overall, the clients achieve lower performance than the capability experiments
-in Figure~\ref{fig:captp} because the sequencer uses a slower metadata
-operation to manage the capability.  ``Sequencer Baseline" is the best
-performance for this particular workload because there is no contention.
-Performance is stable and the MDS is under-loaded because we do not see the
-spike at the end like the other curves. It appears the bottleneck is somewhere
-with the clients -- perhaps it is our 1GBit/s network link. ``2 Seq; No Load
-Balancing" shows an overloaded sequencer because the performance of our one
-sequencer is low and unstable (high variation in performance). We also deduce
-that it is overloaded by comparing to the ``2 Seq; Mantle Load Balancing"
-curve, which shows improved performance when load is migrated. At 10 seconds,
-the sequencer balancer immediately starts balancing and decides that two
-sequencer streams should be split onto two MDSs. It migrates to its neighbor as
-per the policy.
+has been partitioned into the two cluster setup stages. The CPU utilization is
+confined to one core because load balancer interface is not multi-threaded.
+Recall the the load balancing interface re-uses the POSIX metadata cluster
+migration mechanisms and as it has been shown that metadata performance does
+not improve with multiple threads~\cite{konstantinos:pdsw2014-lustre-metadata}.
 
 \textbf{The Malacology Approach:} Re-use the Ceph metadata migration, load
-balancing, cache coherence, and capabilities subsystems of the MDS. Applying
-Mantle's load balancing policy shows 2.02\(\times\) performance improvement and
-16\% degradation in performance over the baseline, even though the baseline has
-half as many clients.
+balancing, cache coherence, and capabilities subsystems of the metadata
+cluster. Applying Mantle's load balancing policy shows 2.02\(\times\)
+performance improvement and only a 16\% degradation in performance over the
+baseline, even though the baseline has half as many clients.
 

### 2016-10-12 paper/paper.tex
paper: fix graphs in mantle evaluation - touch up throughput graph - decided to leave the CPU graph because it shows no multi-threaded - fixed axes on both graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics{figures/mantle-seq-thruput.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] A
-ZLog sequencer that uses the Malacology load balancing interface enjoys the
-load balancing features of the of the metadata cluster. The ``Sequencer
-Baseline'' has no background sequencers, while the rest of the curves show the
-throughput of 1 sequencer when an additional sequencer is running in the
-background generating load.}\label{fig:mantle-seq-thruput}
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
+Performance profiles of a single sequencer running in a resource constrained
+environment show the benefits of using the load balancing interface.  ``1
+Sequencer Baseline'' is the throughput of one sequencer on an underloaded
+physical server; the other curves show how load balancing affects the
+throughput of one sequencer when sequencers for other logs are co-located on
+the same physical server.}\label{fig:mantle-seq-thruput}
 \end{figure}
@@
 
-We evaluate the load balancing interface by showing the performance
-implications of shedding load. The experiments use our internal cluster which
-uses 10 nodes to store objects, one node to monitor the cluster, and 3 nodes
-that can accomodate sequencers. The workload is two logs, one with 7 clients
-and the other with 8 clients; we chose this workload because this load is the
-point at which a single physical server gets overloaded.
+We evaluate the load balancing interface by showing how the cluster of servers
+housing sequencers alleviate load.  The experiments use our internal cluster
+which uses 10 nodes to store objects, one node to monitor the cluster, and 3
+nodes that can accomodate sequencers. The workload is two logs, one with 7
+clients and the other with 8 clients; we chose this workload because this load
+is the point at which a single physical server gets overloaded.
 
 We use a balancer policy specifically designed for multiple sequencers. The
-balancer aggressively sheds half its load to its neighbor as soon as it can.
+balancer aggressively sheds half its load to the next server as soon as it can.
 It is based off the ``Greedy Spill Even" approach in the
 Mantle~\cite{sevilla:sc15-mantle} paper but is reproduced in
-Figure~\ref{listing:greedy-spill} for the reader's convenience.
+Figure~\ref{listing:greedy-spill} for the reader's convenience. An MDS is a
+physical server capable of housing a sequencer.
 
@@
 \centering
-\includegraphics{figures/mantle-cpu-utilization-annotate.png}
+\includegraphics{figures/mantle-cpu-utilization.png}
 \caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
-sequencer using Mantle in Figure~\ref{fig:mantle-seq-thruput} migrates to a new
-server immediately because it uses a policy designed for bursty workloads. This
-is a trace of the same workload running twice: once with a single host server and once
-with Mantle.}
+CPU utilization for a sequencer in an environment with the ``Conservative
+Policy" load balancer shows that the load balancing service is not
+multi-threaded. The load balancing service is designed this way because the
+performance for a guarded resource will not improve with more cores.}
 \label{fig:mantle-cpu-utilization}
@@
 cluster to alleviate load.  Each curve shows the throughput (\(y\) axis) over
-time (\(x\) axis).  ``Sequencer Baseline'' is the performance of one sequencer
+time (\(x\) axis).  ``1 Sequencer Baseline'' is the performance of one sequencer
 on one physical server serving 7 clients - there are no other sequencers or
 logs in the system and the curve represents the maximum throughput of a
-sequencer on an underutilized physical server.  ``No Load Balancing" adds a
+sequencer on an underutilized physical server.  ``No Balancing" adds a
 second sequencer of equal load to the same physical server with the intent of
@@
 server with 2 sequencers has more than a 2\(\times\) throughput degradation.
+The stability of throughput curves (i.e. the variation on the \(y\) axis) indicate which sequencers
+are on overloaded physical servers.  ``1 Sequencer Baseline" is the best
+performance for this particular workload because there is no contention.
+Performance is stable and the MDS is under-loaded because we do not see the
+spike at the end like the other curves.  ``No Balancing" shows an
+overloaded sequencer because the performance of the monitored sequencer is low
+and unstable (high variation in performance).
 
@@
 improvements when sequencers are load balanced across the cluster. Both curves
-are the throughput of one of the sequencers when a background sequencer is
-running. These policies differ by the threshold at which they start balancing:
-``Sequencer Balancing Policy" starts balancing immediately and ``Conservative
-Balancing Policy" waits for 120 seconds before migrating sequencers.
-
-The variation of the sequencer throughputs helps us determine which sequencers
-are on overloaded physical servers.  ``Sequencer Baseline" is the best
-performance for this particular workload because there is no contention.
-Performance is stable and the MDS is under-loaded because we do not see the
-spike at the end like the other curves.  ``No Load Balancing" shows an
-overloaded sequencer because the performance of the monitored sequencer is low
-and unstable (high variation in performance). We also deduce that it is
-overloaded by comparing to the ``Sequencer Balancing Policy" and ``Conservative
-Balancing Policy" curves, which show improved performance when load is
-migrated.
+show the throughput of one of the sequencers when a second sequencer is
+running on the same physical server. The load balancing policies differ by the threshold at which they start balancing:
+``Sequencer Policy" starts balancing immediately and ``Conservative
+Policy" waits for 120 seconds before migrating sequencers.
 
@@
 
-Figure~\ref{fig:mantle-cpu-utilization} gives a time series graph to show the
-behavior of the same workload, 2 sequencers, with balancing turned on and off.
-The curves are the throughput (\(y\) axis) over time (\(x\) axis) and the graph
-has been partitioned into the two cluster setup stages. The CPU utilization is
-confined to one core because load balancer interface is not multi-threaded.
-Recall the the load balancing interface re-uses the POSIX metadata cluster
-migration mechanisms and as it has been shown that metadata performance does
-not improve with multiple threads~\cite{konstantinos:pdsw2014-lustre-metadata}.
+Figure~\ref{fig:mantle-cpu-utilization} shows how load balancing effects the
+physical servers by plotting the CPU utilization average over 1 minute (\(y\)
+axis) over time (\(x\) axis).  It shows the ``Conservative Policy" load
+balancer from the run in Figure~\ref{fig:mantle-seq-thruput}, so recall that
+resources are constrained during the job by co-locating a second sequencer on
+the same physical server.  CPU utilization spikes when physical servers receive
+load (at 2 minutes physical server 1 receives a sequencer from physical server
+0) but the CPU utilization never eclipses 50\%. The utilization is low because
+the load balancer interface is not multi-threaded.  Recall the the load
+balancing interface re-uses the POSIX metadata cluster migration mechanisms;
+serving file system metadata is single-threaded because it has been shown that
+metadata performance does not improve with multiple
+threads~\cite{konstantinos:pdsw2014-lustre-metadata}.
 

### 2016-10-13 paper/paper.tex
paper: more mantle eval touchup
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics{figures/mantle-seq-thruput.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
-Performance profiles of a single sequencer running in a resource constrained
-environment show the benefits of using the load balancing interface.  ``1
-Sequencer Baseline'' is the throughput of one sequencer on an underloaded
-physical server; the other curves show how load balancing affects the
-throughput of one sequencer when sequencers for other logs are co-located on
-the same physical server.}\label{fig:mantle-seq-thruput}
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
+load balancing interface alleviates load for multiple sequencers in
+resource-constrained environments.  ``Underloaded (Baseline)'' is the
+throughput of one sequencer running alone on a cluster; the other curves show
+how load balancing affects the throughput of one sequencer when sequencers for
+other logs are initially co-located on the same physical server (i.e. an
+"Overloaded" environment).}\label{fig:mantle-seq-thruput}
 \end{figure}
@@
 Mantle~\cite{sevilla:sc15-mantle} paper but is reproduced in
-Figure~\ref{listing:greedy-spill} for the reader's convenience. An MDS is a
-physical server capable of housing a sequencer.
+Figure~\ref{listing:greedy-spill} for the reader's convenience. Each line of
+code uses variables exposed by the Mantle API and is executed as a callback by
+Mantle.  ``Server load" lets the administrator control what metric constitutes
+load on a server; ``i" is a variable used by Mantle to iterate over servers and
+``req\_rate" is a metric based on the exponenential decay of the request rate.
+``When policy" is used by Mantle to set a conditional for kicking off migration;
+in this case, the cluster migrates load if the current server, ``whoami", has
+load (\texttt{>0.1}) and the next server (physical servers have ``ranks") has a
+small amount load (\texttt{<0.1}). Finally, the ``where policy" is used by
+Mantle to determine how much load to shed. The policy fills in a ``targets"
+array to tell the migrator how much load to send to each server; in this case,
+we shed half of the current load to the next server. The unit of load is a
+single sequencer for a ZLog log.
 
@@
 \begin{lstlisting}
-  -- Metadata server load
-  mdsload = MDSs[i]["all"]
+  -- Server load
+  load = server[i]["req_rate"]
 
   -- When policy
-  if MDSs[whoami]["load"]>.01 and 
-    MDSs[whoami+1]["load"]<.01 then
+  if server[whoami]["load"]>.01 and 
+    server[whoami+1]["load"]<.01 then
 
   -- Where policy
-  targets[whoami+1]=allmetaload/2
+  targets[whoami+1]=load/2
 \end{lstlisting}
@@
 cluster to alleviate load.  Each curve shows the throughput (\(y\) axis) over
-time (\(x\) axis).  ``1 Sequencer Baseline'' is the performance of one sequencer
+time (\(x\) axis).  ``Underloaded (Baseline)'' is the performance of one sequencer
 on one physical server serving 7 clients - there are no other sequencers or
 logs in the system and the curve represents the maximum throughput of a
-sequencer on an underutilized physical server.  ``No Balancing" adds a
+sequencer on an underutilized physical server.  ``Overloaded: No Balancing" adds a
 second sequencer of equal load to the same physical server with the intent of
@@
 The stability of throughput curves (i.e. the variation on the \(y\) axis) indicate which sequencers
-are on overloaded physical servers.  ``1 Sequencer Baseline" is the best
+are on overloaded physical servers.  ``Underloaded (Baseline)" is the best
 performance for this particular workload because there is no contention.
 Performance is stable and the MDS is under-loaded because we do not see the
-spike at the end like the other curves.  ``No Balancing" shows an
+spike at the end like the other curves.  ``Overloaded: No Balancing" shows an
 overloaded sequencer because the performance of the monitored sequencer is low
@@
 The last two curves of Figure~\ref{fig:mantle-seq-thruput} show the performance
-improvements when sequencers are load balanced across the cluster. Both curves
-show the throughput of one of the sequencers when a second sequencer is
-running on the same physical server. The load balancing policies differ by the threshold at which they start balancing:
-``Sequencer Policy" starts balancing immediately and ``Conservative
+improvements when sequencers are load balanced across the cluster.
+The load balancing policies differ by the threshold at which they start balancing:
+``Overloaded: Sequencer Policy" starts balancing immediately and ``Overloaded: Conservative
 Policy" waits for 120 seconds before migrating sequencers.

### 2016-10-13 paper/paper.tex
paper: add Noah's intro comments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 flexibility for storage systems to evolve without sacrificing the robustness of
-its code-hardened subsystems.  We illustrate the advantages and challenges of
+their code-hardened subsystems.  We illustrate the advantages and challenges of
 programmability by composing existing primitives into two new higher-level
@@
 services inherit desirable qualities of the back-end storage system, including
-the ability to make trade-offs for latency/throughput using lease capabilities, balance load, propogate
-interfaces, and recover from failure.
+the ability to balance load, efficiently propogate cluster metadata, recover
+from failure, and to make trade-offs for latency/throughput using leases.
 \end{abstract}
@@
 
-A number of recent developments are disrupting traditional storage systems.
+A number of recent developments have disrupted traditional storage systems.
 First, the falling prices of flash storage and the availability of new types of
@@
 These three trends put evolutionary pressure on storage systems and raise the
-question whether there are principles that storage systems designers can follow
+question of whether there are principles that storage systems designers can follow
 to evolve storage systems efficiently and without jeopardizing years of
@@
 can be used to \emph{program} new services. By composing higher-level services
-over existing abstractions, one can reuse
+over existing abstraction one can reuse
 subsystems and leverage their optimizations, established correctness,
@@
 services by re-purposing existing subsystem abstractions of the storage stack.
-We build the framework in Ceph, a popular, open source software storage stack.
-We choose Ceph to demonstrate the concept of programmable stoarge because it
+We build the Malacology in Ceph, a popular open source software storage stack.
+We choose Ceph to demonstrate the concept of programmable storage because it
 offers a broad spectrum of existing services, including distributed locking and
-caching services provided by metadata servers, durability and logical storage
-devices provided by the backend object store, and propagation of consistent
+caching services provided by metadata servers, durability and object interfaces
+provided by the backend object store, and propagation of consistent
 cluster state provided by the monitoring service (see
-Figure~\ref{fig:overview}). As we will show in this paper, this framework is
+Figure~\ref{fig:overview}). As we will show in this paper, Malacology is
 expressive enough to provide the functionality necessary for implementing new
@@
 
-{\it \textbf{We define a programmable storage system}} to be a storage system that re-uses and extends existing
-abstractions and services of the underlying software stack. Our prototype
-includes (but is not limited to) the following building blocks:
+{\it \textbf{The definition a programmable storage system}} to be "a storage
+system that allows exisiting abstractions to be re-used and composed to form
+new data and metadata management services whose operation is orothogonal to
+system correctness".  Our prototype includes (but is not limited to) the
+following building blocks:
 
@@
 
-\item Versioning interface that uses the underyling storage stack's consensus
-infrastructure
-
-\item Logical Device interface for adding functionality to daemons that manage
-storage/metadata devices, allowing users to push computation to the data
-
-\item An interface for mediating access to a shared resources, which exposes
-serialization primatives using capabilities, caching, and batching
-
-\item Load Balancing interface for managing resources across the cluster in
-multi-tenant environments
-
-\item Durability interface for persisting policies using the underyling storage
+\item An interface for managing strongly-consistent time-varying service
+metadata.
+
+\item An interface for installing and evolving domain-specific, cluster-wide data
+interfaces.
+
+\item An interface for managing access to shared resources using a variety of
+optimization strategies.
+
+\item An interface for load balancing resources across the cluster.
+
+\item A durability interface for persisting policies using the underyling storage
 stack's object store
@@
 
-\item A high-performance distributed shared log service called ZLog which is
-based on CORFU~\cite{balakrishnan_corfu_2012}
+\item A high-performance distributed shared log service called ZLog, that is an
+implementation of CORFU~\cite{balakrishnan_corfu_2012}
 
@@
 
-\section{Highly Tailored, Application-Specific Storage Stacks
+\section{Application-Specific Storage Stacks
 }\label{highly-tailored-and-application-specifc-storage-systems}

### 2016-10-13 experiments/mds-zlog-seq-migrate-master/run.sh
mantle: add experiments re-based on master - shows better throughput (similar to BRADOS story) - fixes bugs with rados cli - shows confusing throughput results involving path traversal
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,15 @@
+#!/bin/bash
+
+set -ex
+
+cd site
+ANSIBLE_LOG_PATH="logs/baseliner.log" \
+ansible-playbook --forks 50 --skip-tags "with_pkg"
+  ceph.yml \
+  ../workloads/zeqr-isolated.yml \
+  ../workloads/zeqr-nobalancing.yml \
+  mantle.yml \
+  ../workloads/zeqr-mantle.yml
+cd -
+
+tar czf ../mds-zlog-seq-migrate-master.tar.gz ../mds-zlog-seq-migrate-master

### 2016-10-13 paper/paper.tex
paper: add Noah's problem comments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 the best performance. For example, HDFS was designed specifically to serve
-Hadoop's jobs, and uses techniques like relaxing POSIX constraints to achieve
+Hadoop's jobs, and uses techniques like exposing data locality and relaxing POSIX constraints to achieve
 the best performance~\cite{CITEME}. Alternatively, general-purpose storage
-stacks are built with the flexiblity to serve many applications by exposing
-tunable parameters to control low level implementation details like block size.
-Unfortunately, users want more control from the general-purpose storage stacks
+stacks are built with the flexiblity to serve many applications by providing a
+variety of interfaces and tunable parameters. Unfortunately, managing competing
+forces in these systems is difficult and users
+want more control from the general-purpose storage stacks
 without going as far as building their storage system from the ground up.
@@
 
-To demonstrate this trend in storage systems, we examine Ceph's
-programmability. Something of a storage swiss army knife, Ceph supports file,
-block, and object interfaces simultaneously in a single
+To demonstrate this trend in storage systems we examine the state of
+programmability in Ceph. Something of a storage swiss army knife, Ceph simultaneously supports file,
+block, and object interfaces on a single
 cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
@@
 erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already has some
-programmability; the OSDs support the implementation of domain-specific logical
-object interfaces by composing existing interfaces. These interfaces are
+programmability; the OSDs support domain-specific object interfaces that are
+implemented by composing existing low-level storage interfaces that execute
+atomically. These interfaces are
 written in C++ and are statically loaded into the system.
@@
 growth in the production use of domain-specific interfaces in the Ceph
-community since 2010.  What is most remarkable is that this trend contradicts
-the notion that API changes are a burden for users.  Rather it appears that a
-gap in existing interfaces are being addressed through programmability. In
-fact, Table~\ref{table:objclasses} categorizes existing interfaces and we
-clearly see a trend towards reusable services.
+community since 2010. In that figure, classes are functional groupings of methods on storage
+objects.  What is most remarkable is that this trend contradicts the notion
+that API changes are a burden for users.  Rather it appears that a gap in
+existing interfaces are being addressed through programmability. In fact,
+Table~\ref{table:objclasses} categorizes existing interfaces and we clearly see
+a trend towards reusable services.
 
@@
 The takeway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
-already trying to achieve programmability because their needs, whether they be
-related to performance, availability, consistency, correctness etc., are not
-satisfied by the underlying storage system. The popularity of the custom object
+already trying to use programmability because their needs, whether they be
+related to performance, availability, consistency, convenience, etc., are not
+satisfied by the existing default set of interfaces. The popularity of the custom object
 interface facility of Ceph could be due to a number of reasons, such as the
@@
 to improve performance. Their solution is a way to work around the
-traditionally rigid storage APIs because custome object interfaces give
+traditionally rigid storage APIs because custom object interfaces give
 programmers the ability to tell the storage system about their application: if
@@
 interfaces, which can be as simple as a redirection to the persistent data
-store or as complicated as a strongly consistent directly service, should be
+store or as complicated as a strongly consistent directory service, should be
 used and re-used in many contexts to implement a wide range of services. By

### 2016-10-14 paper/paper.tex
paper: tidy up malacology section, fix figure 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics{figures/overview.png}
-
-\caption{Scalable storage systems consist of object storage daemons (OSD) which
-store data fragments, monitor daemons (MON) that maintain cluster state, and
-service-specific daemons such as file system metadata servers (MDS). Malacology
-enables the programmability of internal abstractions and services (indicated by
-bold arrows) in order to leverage the back-end system's durability, consistent
-versioning, consensus, and metadata subsystems.  With Malacology, we add two
-new services, ZLog and Mantle, that sit alongside the traditional user-facing
-APIs (file, block, and object) in the storage stack.  \label{fig:overview}}
+\caption{Scalable storage systems consist of storage daemons which store data
+fragments, monitor daemons that maintain cluster state, and service-specific
+daemons such as file system metadata servers. Malacology enables the
+programmability of internal abstractions (indicated by bold arrows) in order to
+re-use and compose existing subsystems.  With Malacology, we add two new
+services, ZLog and Mantle, that sit alongside the traditional user-facing APIs
+(file, block, and object) in the storage stack.  \label{fig:overview}}
 \end{figure}
@@
 
-{\it \textbf{The definition a programmable storage system}} to be "a storage
-system that allows exisiting abstractions to be re-used and composed to form
+{\it \textbf{The definition a programmable storage system}} to be ``a storage
+system that allows existing abstractions to be re-used and composed to form
 new data and metadata management services whose operation is orothogonal to
@@
 
-\item An interface for managing strongly-consistent time-varying service
-metadata.
-
-\item An interface for installing and evolving domain-specific, cluster-wide data
-interfaces.
-
-\item An interface for managing access to shared resources using a variety of
+\item An interface for managing strongly-consistent time-varying **service
+metadata**.
+
+\item An interface for installing and evolving domain-specific, cluster-wide **data
+interfaces**.
+
+\item An interface for managing access to **shared resources** using a variety of
 optimization strategies.
 
-\item An interface for load balancing resources across the cluster.
-
-\item A durability interface for persisting policies using the underyling storage
-stack's object store
+\item An interface for **load balancing** resources across the cluster.
+
+\item An interface for **durability** that persists policies using the underyling storage
+stack's object store.
 
@@
 
-\section{Application-Specific Storage Stacks
-}\label{highly-tailored-and-application-specifc-storage-systems}
+\section{Application-Specific Storage Stacks}
+\label{sec:application-specifc-storage-stacks}
 
@@
 
-The takeway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
+The takeaway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
 already trying to use programmability because their needs, whether they be
@@
 \centering
-\title{Hi}
 \begin{tabular}{  l | l | l | l }
@@
 \textbf{Provided Primitives}            \\ \hline
-{\it Versioning}
+{\it Service Metadata}
   & Zookeeper coordination~\cite{hunt_zookeeper_2010}, Chubby~\cite{website:chubby}
@@
   \\
-{\it Logical Device}
+{\it Object \& Data}
   & Swift in situ storage/compute~\cite{website:zerocloud}
@@
 
-\subsection{Versioning: Cluster State Management}
+\subsection{Service Metadata Interface}
 \label{sec:mon}
-\label{consistencyversioning-of-cluster-state}
+\label{service-metadata}
 
@@
 successful service and a necessary component in order to diagnose and detect
-failures, when they occur. In the case of Ceph, a consistent view of cluster 
-state among server daemons and clients is critical to provide strong consistency 
-guarantees to clients. Ceph maintains cluster state information in per-subsystem 
-``maps'' (e.g. OSD map, MDS map) that record membership and status information. 
-A PAXOS-based monitoring service (termed, the Monitor) is responsible for 
-integrating state changes into cluster maps, responding to requests from 
-out-of-date clients and synchronizing members of the cluster whenever there is a 
-change in a map so that they all observe the same system state. As a fundamental 
-building block of many system designs, consensus abstractions such as PAXOS are 
-a common technique for maintaining consistent data versions, and are a useful 
-system to expose.
+failures, when they occur. This is further complicated by variable propagation
+delays and heterogenous hardware in highly dynamic environments.
+
+%Thus, system
+%services realized through extensibility should be versioned together as a
+%group, exposing the version to allow introspection and domain-specific
+%treatment of version mismatch scenarios.
+
+In the case of Ceph, a consistent view of cluster state among server daemons
+and clients is critical to provide strong consistency guarantees to clients.
+Ceph maintains cluster state information in per-subsystem ``maps'' that record
+membership and status information.  A PAXOS-based monitoring service is
+responsible for integrating state changes into cluster maps, responding to
+requests from out-of-date clients and synchronizing members of the cluster
+whenever there is a change in a map so that they all observe the same system
+state. As a fundamental building block of many system designs, consensus
+abstractions such as PAXOS are a common technique for maintaining consistent
+data versions, and are a useful system to expose.
 
@@
 application-specific logic that can be executed for particular maps (or values
-in a map).
-
-\paragraph*{Malacology:} Malacology exposes Ceph's PAXOS-based monitoring
-service implementing a generic API
-for adding arbitrary values to existing subsystem cluster maps, namely to the 
-OSD map and the MDS map. As a consequence of this, applications can define 
-simple but useful domain-specific logic to the PAXOS state machine, such as 
-authorization control (just specific clients can write new values) or to trigger 
-actions based on specific values (e.g. sanitize values). Since the monitor is 
-intended to be out of high-performance I/O paths, a general guideline is to make 
-use of this functionality infrequently and to assign small values to maps. 
-Malacology itself makes use of this functionality to register, version and 
-propagate dynamic code (Lua scripts) for new object interfaces defined in OSDs 
-(\S\ref{active-storage}) and load balancing policies in MDSs 
-(\S\ref{malacology:mds}).
-
-\subsection{Logical Storage Devices}
-\label{active-storage}
-
-Briefly described in Section~\ref{sec:intro}, Ceph supports the creation
-of application-specific object interfaces~\cite{weil_rados_2007}. The ability
-to offload computation can reduce data movement, and transnational interfaces can
-significantly simplify construction of complex storage interfaces that require
-uncoordinated parallel access.
+in a map).  Since the monitor is intended to be out of high-performance I/O
+paths, a general guideline is to make use of this functionality infrequently
+and to assign small values to maps.
+
+\paragraph*{Malacology:} Malacology exposes a strongly-consistent view of
+time-varying service metadata as a service rather than a hidden internal
+component. Malacology provides a generic API for adding arbitrary values to
+existing subsystem cluster maps. As a consequence of this, applications can
+define simple but useful service-specific logic to the strongly-consistent
+interface, such as authorization control (just specific clients can write new
+values) or to trigger actions based on specific values (e.g. sanitize values).
+The higher level services we implement in \S\ref{sec:services} make use of this
+functionality to register, version and propagate dynamic code (Lua scripts) for
+new object interfaces defined in storage daemons (\S\ref{active-storage}) and
+policies in the load balancer \S\ref{malacology:mds}).  Using this service
+guarantees that interface definitions are not only made durable, but are
+transparently and consistently propagated throughput the cluster and that
+clients are properly synchronized with the latest interfaces.
+
+%And finally, it is important to not underestimate the importance of the
+%preservation of the modifications that enable a new system service. In many
+%(if not all) cases, interfaces defining access to data are just as important
+%as the data itself by virtue of inherently providing structural context. Thus,
+%all components of a system service must be kept to the same standards of
+%protection as data in the system itself.
+
+\subsection{Object \& Data Interface}
+\label{object-data-interface}
+
+Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
+supports the creation of application-specific object
+interfaces~\cite{weil_rados_2007}. The ability to offload computation can
+reduce data movement, and transactional interfaces can significantly simplify
+construction of complex storage interfaces that require uncoordinated parallel
+access.
 
@@
 may invoke. In the case of Ceph each function is implicitly run within the
-context of an object specified when the function is invoked. Developers of
+context of an object specified when the function is called. Developers of
 object interfaces express behavior by creating a composition of native
@@
 function input and output. A wide range of native interfaces are available to
-developers, such as reading and writing to a byte stream, controlling snapshots
-and clones, all of which may be transactionally composed with access to a
-sorted key-value service allowing applications to create rich interfaces, such
-as using a read-modify-write guard to protect a shared index defined over
-content store in a byte stream.
+developers, such as reading and writing to a byte stream, and controlling
+snapshots and clones. These object interfaces may be transactionally composed
+with access to a sorted key-value service allowing applications to create rich
+semantics. An example is using a read-modify-write guard to protect a shared
+index defined over content stored in a byte stream.
 
@@
 \includegraphics{figures/implementation.png}
-\caption{Malacology re-uses Ceph subsystems to allow users to dynamically define 
-new object interfaces and load balancing policies. It achieves this by extending 
-the OSD and MDS subsystems with an embedded Lua VM. To manage newly dynamic 
-interface classes, it exposes the core functionality of the Monitor to allow 
-arbitrary data to be associated to ODS and MDS cluster maps. After new 
-interfaces are submitted to the monitor and the associated Paxos proposal is 
-accepted by the monitor quorum, the script is redundantly and consistently 
-distributed to OSDs so this snippets persist.
-\label{fig:implementation}}
-\end{figure}
+\caption{Malacology re-uses Ceph subsystems to allow users to dynamically
+define new object/data interfaces and load balancing policies. It achieves this
+by extending the object storage daemon (OSD) and metadata server daemon (MDS)
+subsystems with an embedded Lua VM.  It uses the of the strongly-consistent
+service metadata interface to propogate the object interfaces and versions
+across the cluster.
+\label{fig:implementation}} \end{figure}
 
 \paragraph*{Malacology:} The implementation of Ceph's object abstraction,
-although powerful, has some drawbacks. Supporting only C/C++ for object
-interface developers, Ceph requires distribution of compiled binaries for the
-correct architecture, adding a large barrier of entry for developers and
-system administrators. Second, having no way to dynamically unload modules,
-any changes require a full restart of an OSD which may have serious
-performance impacts due to a loss of cached data. And
-finally, the security limitations of the framework limit the use of object
-classes to all but those with administrative level access and deep technical
-expertise.
-
-To address these concerns, Malacology takes advantage of community contributed
-Lua extensions to Ceph allowing new object interfaces to be dynamically loaded
-into the system and modified at runtime, resulting in a object storage API
-with economy of expression, which at the same time provides the full set of
-features of the base object class. New object interfaces that are
-expressed in thousands of lines of code can be implemented in approximately an
-order of magnitude less code~\cite{geambasu_comet_2010}.
-
-{\bf Interface management.} All of these modifications work in tandem to
-implement the desired behavior, and are typically co-designed together such
-that each depends on the other to behave as expected. In highly dynamic
-environments such as a distributed storage system, changes to the system
-propagate at variable rates, requiring all components of the system to be able
-to adapt. Thus, system services realized through extensibility should be
-versioned together as a group, exposing the version to allow introspection and
-domain-specific treatment of version mismatch scenarios.
-
-And finally, it is important to not underestimate the importance of the
-preservation of the modifications that enable a new system service. In many
-(if not all) cases, interfaces defining access to data are just as important
-as the data itself by virtue of inherently providing structural context. Thus,
-all components of a system service must be kept to the same standards of
-protection as data in the system itself.
-
-To this end we have extended the monitoring service using the framework
-described in Section~\ref{sec:mon} to provide an interface for registration
-and verification of dynamically defined object interfaces. Using this service
-guarantees that interface definitions are not only made durable, but are
-transparently and consistently propagated throughput the cluster and that
-clients are properly synchronized with the latest interfaces.
-
-\subsection{Distributed Metadata Services}
+although powerful, does not readily support programmability. Supporting only
+C/C++ for object interface developers, Ceph requires distribution of compiled
+binaries for the correct architecture, adding a large barrier of entry for
+developers and system administrators. Second, having no way to dynamically
+unload modules, any changes require a full restart of a storage daemon which
+may have serious performance impacts due to loss of cached data. And finally,
+the security limitations of the framework limit the use of object interfaces to
+all but those with administrative level access and deep technical expertise.
+
+To address these concerns, Malacology takes advantage of Lua extensions
+contributed by the Ceph community. This allows new object interfaces to be
+dynamically loaded into the system and modified at runtime, resulting in a
+object storage API with economy of expression, which at the same time provides
+the full set of features of the base object class. New object interfaces that
+are expressed in thousands of lines of code can be implemented in approximately
+an order of magnitude less code~\cite{geambasu_comet_2010}.
+
+All of these modifications work in tandem to implement the desired behavior,
+and are typically co-designed together such that each depends on the other to
+behave as expected.
+
+\subsection{Distributed Metadata Interfaces}
 \label{malacology:mds}
@@
 distributed locking services.  In Ceph, the locking service implements a
-capability-based system that express what data and metadata clients are allowed
-to access as well as what state they may cache and modify locally. While
-designed for a fixed file abstraction, indexing, locking, and caching are all
-common services that are useful to a broad spectrum of applications.
-
-Ceph also addresses the challenge of balancing metadata load using load
-balancing policies that compute metrics based on system state (e.g.  CPU and
-memory utilization) and statistics collected by the cluster (e.g. the
-popularity of an inode). Ceph uses dynamic subtree partitioning to move
-variable sized namespace subtrees. These units can be shipped anywhere (i.e.,
-to any MDS of any capacity) at any time for any reason. The original balancer
-was designed with hard-coded policies and tunables.  Distributed applications
-that share centralized resources (e.g. a database or directory) face similar
-challenges which are often solved using application-specific sharding.
+capability-based system that expresses what data and metadata clients are
+allowed to access as well as what state they may cache and modify locally.
+While designed for a fixed file abstraction, indexing, locking, and caching are
+all common services that are useful to a broad spectrum of applications.
+Distributed applications that share centralized resources (e.g. a database or
+directory) face similar challenges which are often solved using
+application-specific sharding.
+
+Ceph also addresses the challenge of balancing metadata load with a separate
+metadata cluster. This cluster uses load balancing policies to migrate
+directory inodes around the cluster to alleviate load. The policies use metrics
+based on system state (e.g.  CPU and memory utilization) and statistics
+collected by the cluster (e.g. the popularity of an inode). Ceph uses dynamic
+subtree partitioning to move variable sized namespace subtrees. These units can
+be shipped anywhere (i.e., to any metadata server of any capacity) at any time
+for any reason. The original balancer was designed with hard-coded policies and
+tunables.
 
@@
 
-\paragraph*{Malacology:} New interface hooks are added in strategic spots in
-the metadata service to control load balancing and capabilities, and for
-defining new file types allowing the indexing service to be re-purposed for
-other types of named resources.
+\paragraph*{Malacology:} interfaces are added in strategic spots in the
+metadata service for guarding shared resources and load balancing.
+
+{\bf Shared Resource Interface.} The metadata service also manages sessions
+with its clients. This gives both subsystems (metadata servers and clients) the
+ability to cache data and obtain locks. The metadata service is responsible for
+responding to client requests for capabilities by first revoking caps from
+other clients.  Currently, clients voluntarily release their capabilities, and
+the metadata server maintains a queue of requests.  Both clients and metadata
+servers use a best-effort approach to revoking or releasing capabilities.
+
+The clients retain the best-effort, voluntary release of capabilities, but in
+Malacology, the queuing policies are generalized within the MDS to allow for
+centralized policies such as fairness or priority.
 
@@
 metadata load balancer, has been implemented on Malacology so that its
-infrastructure can enjoy all the properties of existing internal abstractions. The load
-balancing logic is intact but the infrastructure has been
-improved to safely load, manage, and persist its load balancing policies.
-
-{\bf Safely Loading Classes.} The original Mantle API is fragile: the API is
-not enforced by the runtime, the system allows strings to be injected through
-the admin daemon, constructing the Lua script in C++ is clunky, and the
-administrator can inject really bad policies (e.g., while 1) that brings the
-whole system down. Using the Malacology durability and beacon interfaces,
-Mantle can safely load policies from the object store and centralize
-warnings/errors, respectively.
-
-{\bf Policy management.} A new interface is added to the metadata service
-subsystem using the monitoring framework to support management of interfaces.
-This ensures that the correct balancer version is consistently distributed and
-verified on all MDS nodes. Policies are also durably saved with Ceph's
-replication and redundancy.
+infrastructure can enjoy all the properties of existing internal abstractions.
+The load balancing logic is intact but the infrastructure has been improved to
+safely load, manage, and persist its load balancing policies.  Composing Mantle
+using the service metadata, durability and beacon interfaces, Mantle can safely
+version balancer policies, save balancer policies in the back end object store
+and centralize warnings/errors.
+
+%{\bf Policy management.} A new interface is added to the metadata service
+%subsystem using the monitoring framework to support management of interfaces.
+%This ensures that the correct balancer version is consistently distributed and
+%verified on all MDS nodes. Policies are also durably saved with Ceph's
+%replication and redundancy.
 
@@
 
-{\bf Capabilities.} The metadata service also manages sessions with its
-clients. This gives both subsystems (MDSs and clients) the ability to cache
-data and grab locks. The metadata service is responsible for responding to
-client requests for capabilities by first revoking caps from other clients.
-Currently, clients voluntarily release their capabilities, and the MDS maintains
-a queue of requests.  Both clients and MDSs use a best-effort approach to
-revoking or releasing capabilities.
-
-The clients retain the best-effort, voluntary release of capabilities, but in
-Malacology, the queuing policies are generalized within the MDS to allow for
-centralized policies such as fairness or priority.
-
-{\bf File Types.} We leverage the same active and typed storage module
-presented in DataMods~\cite{watkins_datamods_2012} to assign state and a set of
-functions to a file (stored in the inode). This is valuable for
-metadata load balancing so that certain types of files can have a higher
-priority when making balancing decisions.
+%{\bf File Types.} We leverage the same active and typed storage module
+%presented in DataMods~\cite{watkins_datamods_2012} to assign state and a set of
+%functions to a file (stored in the inode). This is valuable for
+%metadata load balancing so that certain types of files can have a higher
+%priority when making balancing decisions.
 
@@
 
-\subsection{Durability}
+\subsection{Durability Interface}
 \label{sec:durability}

### 2016-10-14 paper/paper.tex
paper: focus the challenges section, add transition
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 services inherit desirable qualities of the back-end storage system, including
-the ability to balance load, efficiently propogate cluster metadata, recover
+the ability to balance load, efficiently propagate cluster metadata, recover
 from failure, and to make trade-offs for latency/throughput using leases.
@@
 
-\item An interface for managing strongly-consistent time-varying **service
-metadata**.
-
-\item An interface for installing and evolving domain-specific, cluster-wide **data
-interfaces**.
-
-\item An interface for managing access to **shared resources** using a variety of
-optimization strategies.
-
-\item An interface for **load balancing** resources across the cluster.
-
-\item An interface for **durability** that persists policies using the underyling storage
-stack's object store.
+\item An interface for managing strongly-consistent time-varying \textbf{service
+metadata}.
+
+\item An interface for installing and evolving domain-specific, cluster-wide
+\textbf{data interfaces}.
+
+\item An interface for managing access to \textbf{shared resources} using a
+variety of optimization strategies.
+
+\item An interface for \textbf{load balancing} resources across the cluster.
+
+\item An interface for \textbf{durability} that persists policies using the
+underyling storage stack's object store.
 
@@
 
-Establishing the infrastructure for programmability into existing services and
+Implementing the infrastructure for programmability into existing services and
 abstractions of distributed storage systems is challenging, even if one assumes
@@
 one has to be careful when factoring them out not to introduce additional bugs.
-
-\item Mechanisms that are only exercised according to hard-wired policies and
+These policies and optimiations are usually cross-cutting solutions to concerns
+or trade-offs that cannot be fully explored yet (as they relate to workload or
+hardware). Given these policies and optimizations, decomposition of otherwise
+orthogonal internal abstractions can be difficult or dangerous.
+
+\item Mechanisms that are often only exercised according to hard-wired policies and
 not in their full generality have hidden bugs that are revealed as soon as
@@
 \end{itemize}
+
+To address these challenges we present Malacology. Malacology is both a
+prototype for a programmable storage system and a design approach to evolve
+storage systems efficiently and without jeopardizing years of code-hardening
+and performance optimization efforts.  Although Malacology uses
+the internal abstractions of the underlying storage system, including its
+subsystems, components, and implementations, we emphasize that our system
+still addresses the general challenges outlined above.
+
+The main challenge of designing a programmable storage system is choosing the
+right internal abstractions and picking the correct layers for exposing them.
+In this paper, we do not present an exhaustive list of possible internal
+abstractions nor do we contend that the abstractions we choose provide the best
+trade-offs for all applications. For example, if consensus is correctly
+exposed, one could implement versioning, serialization, or capabilities on top;
+but perhaps versioning, serialization, and capabilities would be better
+interfaces for a given set of applications. These questions are not answered in
+this paper and instead we focus on showing the feasibility of building such a
+system, given advances in the quality and robustness of today's storage stacks.
+
+Malacology implements and demonstrates the ideas of programmable storage in
+Ceph. We choose Ceph because it is a production quality system and because it
+is open source. The large developer community ensures that code is relatively
+robust and the visibility of the code lets us expose any interface we want. In
+the next section we describe the Ceph components that we expose as Malacology
+interfaces.
 
@@
 
-Malacology is both a prototype for a programmable storage system and a design
-approach to evolve storage systems efficiently and without jeopardizing years
-of code-hardening and performance optimization efforts. The guiding principle
+The guiding principle
 is to re-use existing services and extend them so that these services can be

### 2016-10-15 paper/paper.tex
:intro tweaks

--- 2016-10-13 paper/paper.tex
@@
 flexibility for storage systems to evolve without sacrificing the robustness of
-their code-hardened subsystems.  We illustrate the advantages and challenges of
+its code-hardened subsystems.  We illustrate the advantages and challenges of
 programmability by composing existing primitives into two new higher-level
@@
 services inherit desirable qualities of the back-end storage system, including
-the ability to balance load, efficiently propogate cluster metadata, recover
-from failure, and to make trade-offs for latency/throughput using leases.
+the ability to make trade-offs for latency/throughput using lease capabilities, balance load, propogate
+interfaces, and recover from failure.
 \end{abstract}
@@
 
-A number of recent developments have disrupted traditional storage systems.
+A number of recent developments are disrupting traditional storage systems.
 First, the falling prices of flash storage and the availability of new types of
@@
 code paths and incorporate new
-optimizations~\cite{gray_tape_2007,gray_flash_2008}. Second, the demand for
+optimizations~\cite{gray_tape_2007,gray_flash_2008}. 
+Second, the demand for
 managing structured data and flexible consistency semantics at scale pressure
 big data processing systems to use storage abstractions that can meet these
-demands~\cite{apache_contributors_parquet_2014}. Finally, production-quality
+demands~\cite{apache_contributors_parquet_2014}. 
+Finally, production-quality
 scalable storage systems available as open source software have established and
@@
 question of whether there are principles that storage systems designers can follow
-to evolve storage systems efficiently and without jeopardizing years of
-code-hardening and performance optimization efforts that are important for
-users to continue to entrust their data to the storage system.
+to evolve storage systems efficiently, without jeopardizing years of
+code-hardening and performance optimization efforts.
+% that are important for
+%users to continue to entrust their data to the storage system.
 
@@
 can be used to \emph{program} new services. By composing higher-level services
-over existing abstraction one can reuse
+over existing abstractions, one can reuse
 subsystems and leverage their optimizations, established correctness,
-robustness, and efficiency. We will refer to this programmability as
-\emph{programmable storage}, which differs from \emph{active storage} (the
-injection and execution of arbitrary code in a storage system) and
-\emph{software-defined storage} (the control of thin-provisioning of storage).
+robustness, and efficiency. 
+%We will refer to this programmability as
+%\emph{programmable storage}, which differs from \emph{active storage} (the
+%injection and execution of arbitrary code in a storage system) 
+%and
+%\emph{software-defined storage} (the control of thin-provisioning of storage).
+
+{\it \textbf{We define a programmable storage system}} to be a storage system that facilitates the 
+re-use and extension of existing storage abstractions and services provided by the underlying software stack,
+to enable the creation of new storage abstractions via composition.
+A programmable storage system is realized by exposing existing functionality (such as 
+metadata services and synchronization capabilities) as interfaces that can be ``glued together''
+using a high-level language.  Programmable storage differs from \emph{active storage},
+or the injection and execution of arbitrary code in a storage system.
 
@@
 services by re-purposing existing subsystem abstractions of the storage stack.
-We build the Malacology in Ceph, a popular open source software storage stack.
-We choose Ceph to demonstrate the concept of programmable storage because it
+We build the framework in Ceph, a popular, open source software storage stack.
+We choose Ceph to demonstrate the concept of programmable stoarge because it
 offers a broad spectrum of existing services, including distributed locking and
-caching services provided by metadata servers, durability and object interfaces
-provided by the backend object store, and propagation of consistent
+caching services provided by metadata servers, durability and logical storage
+devices provided by the backend object store, and propagation of consistent
 cluster state provided by the monitoring service (see
-Figure~\ref{fig:overview}). As we will show in this paper, Malacology is
+Figure~\ref{fig:overview}). As we will show in this paper, this framework is
 expressive enough to provide the functionality necessary for implementing new
@@
 
-{\it \textbf{The definition a programmable storage system}} to be "a storage
-system that allows exisiting abstractions to be re-used and composed to form
-new data and metadata management services whose operation is orothogonal to
-system correctness".  Our prototype includes (but is not limited to) the
-following building blocks:
+Our prototype includes a non-exhaustive set of interfaces that can be used as building blocks
+for constructing novel storage abstractions, including:
 
@@
 
-\item An interface for managing strongly-consistent time-varying service
-metadata.
-
-\item An interface for installing and evolving domain-specific, cluster-wide data
-interfaces.
-
-\item An interface for managing access to shared resources using a variety of
-optimization strategies.
-
-\item An interface for load balancing resources across the cluster.
-
-\item A durability interface for persisting policies using the underyling storage
+\item Versioning interface that uses the underyling storage stack's consensus
+infrastructure
+
+\item Logical Device interface for adding functionality to daemons that manage
+storage/metadata devices, allowing users to push computation to the data
+
+\item An interface for mediating access to a shared resources, which exposes
+serialization primatives using capabilities, caching, and batching
+
+\item Load Balancing interface for managing resources across the cluster in
+multi-tenant environments
+
+\item Durability interface for persisting policies using the underyling storage
 stack's object store
@@
 
-\item A high-performance distributed shared log service called ZLog, that is an
-implementation of CORFU~\cite{balakrishnan_corfu_2012}
+\item A high-performance distributed shared log service called ZLog which is
+based on CORFU~\cite{balakrishnan_corfu_2012}
 
@@
 
-\section{Application-Specific Storage Stacks
+\section{Highly Tailored, Application-Specific Storage Stacks
 }\label{highly-tailored-and-application-specifc-storage-systems}
@@
 the best performance. For example, HDFS was designed specifically to serve
-Hadoop's jobs, and uses techniques like exposing data locality and relaxing POSIX constraints to achieve
+Hadoop's jobs, and uses techniques like relaxing POSIX constraints to achieve
 the best performance~\cite{CITEME}. Alternatively, general-purpose storage
-stacks are built with the flexiblity to serve many applications by providing a
-variety of interfaces and tunable parameters. Unfortunately, managing competing
-forces in these systems is difficult and users
-want more control from the general-purpose storage stacks
+stacks are built with the flexiblity to serve many applications by exposing
+tunable parameters to control low level implementation details like block size.
+Unfortunately, users want more control from the general-purpose storage stacks
 without going as far as building their storage system from the ground up.
@@
 
-To demonstrate this trend in storage systems we examine the state of
-programmability in Ceph. Something of a storage swiss army knife, Ceph simultaneously supports file,
-block, and object interfaces on a single
+To demonstrate this trend in storage systems, we examine Ceph's
+programmability. Something of a storage swiss army knife, Ceph supports file,
+block, and object interfaces simultaneously in a single
 cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
@@
 erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already has some
-programmability; the OSDs support domain-specific object interfaces that are
-implemented by composing existing low-level storage interfaces that execute
-atomically. These interfaces are
+programmability; the OSDs support the implementation of domain-specific logical
+object interfaces by composing existing interfaces. These interfaces are
 written in C++ and are statically loaded into the system.
@@
 growth in the production use of domain-specific interfaces in the Ceph
-community since 2010. In that figure, classes are functional groupings of methods on storage
-objects.  What is most remarkable is that this trend contradicts the notion
-that API changes are a burden for users.  Rather it appears that a gap in
-existing interfaces are being addressed through programmability. In fact,
-Table~\ref{table:objclasses} categorizes existing interfaces and we clearly see
-a trend towards reusable services.
+community since 2010.  What is most remarkable is that this trend contradicts
+the notion that API changes are a burden for users.  Rather it appears that a
+gap in existing interfaces are being addressed through programmability. In
+fact, Table~\ref{table:objclasses} categorizes existing interfaces and we
+clearly see a trend towards reusable services.
 
@@
 The takeway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
-already trying to use programmability because their needs, whether they be
-related to performance, availability, consistency, convenience, etc., are not
-satisfied by the existing default set of interfaces. The popularity of the custom object
+already trying to achieve programmability because their needs, whether they be
+related to performance, availability, consistency, correctness etc., are not
+satisfied by the underlying storage system. The popularity of the custom object
 interface facility of Ceph could be due to a number of reasons, such as the
@@
 to improve performance. Their solution is a way to work around the
-traditionally rigid storage APIs because custom object interfaces give
+traditionally rigid storage APIs because custome object interfaces give
 programmers the ability to tell the storage system about their application: if
@@
 interfaces, which can be as simple as a redirection to the persistent data
-store or as complicated as a strongly consistent directory service, should be
+store or as complicated as a strongly consistent directly service, should be
 used and re-used in many contexts to implement a wide range of services. By
@@
 \includegraphics{figures/mantle-seq-thruput.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
-load balancing interface alleviates load for multiple sequencers in
-resource-constrained environments.  ``Underloaded (Baseline)'' is the
-throughput of one sequencer running alone on a cluster; the other curves show
-how load balancing affects the throughput of one sequencer when sequencers for
-other logs are initially co-located on the same physical server (i.e. an
-"Overloaded" environment).}\label{fig:mantle-seq-thruput}
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
+Performance profiles of a single sequencer running in a resource constrained
+environment show the benefits of using the load balancing interface.  ``1
+Sequencer Baseline'' is the throughput of one sequencer on an underloaded
+physical server; the other curves show how load balancing affects the
+throughput of one sequencer when sequencers for other logs are co-located on
+the same physical server.}\label{fig:mantle-seq-thruput}
 \end{figure}
@@
 Mantle~\cite{sevilla:sc15-mantle} paper but is reproduced in
-Figure~\ref{listing:greedy-spill} for the reader's convenience. Each line of
-code uses variables exposed by the Mantle API and is executed as a callback by
-Mantle.  ``Server load" lets the administrator control what metric constitutes
-load on a server; ``i" is a variable used by Mantle to iterate over servers and
-``req\_rate" is a metric based on the exponenential decay of the request rate.
-``When policy" is used by Mantle to set a conditional for kicking off migration;
-in this case, the cluster migrates load if the current server, ``whoami", has
-load (\texttt{>0.1}) and the next server (physical servers have ``ranks") has a
-small amount load (\texttt{<0.1}). Finally, the ``where policy" is used by
-Mantle to determine how much load to shed. The policy fills in a ``targets"
-array to tell the migrator how much load to send to each server; in this case,
-we shed half of the current load to the next server. The unit of load is a
-single sequencer for a ZLog log.
+Figure~\ref{listing:greedy-spill} for the reader's convenience. An MDS is a
+physical server capable of housing a sequencer.
 
@@
 \begin{lstlisting}
-  -- Server load
-  load = server[i]["req_rate"]
+  -- Metadata server load
+  mdsload = MDSs[i]["all"]
 
   -- When policy
-  if server[whoami]["load"]>.01 and 
-    server[whoami+1]["load"]<.01 then
+  if MDSs[whoami]["load"]>.01 and 
+    MDSs[whoami+1]["load"]<.01 then
 
   -- Where policy
-  targets[whoami+1]=load/2
+  targets[whoami+1]=allmetaload/2
 \end{lstlisting}
@@
 cluster to alleviate load.  Each curve shows the throughput (\(y\) axis) over
-time (\(x\) axis).  ``Underloaded (Baseline)'' is the performance of one sequencer
+time (\(x\) axis).  ``1 Sequencer Baseline'' is the performance of one sequencer
 on one physical server serving 7 clients - there are no other sequencers or
 logs in the system and the curve represents the maximum throughput of a
-sequencer on an underutilized physical server.  ``Overloaded: No Balancing" adds a
+sequencer on an underutilized physical server.  ``No Balancing" adds a
 second sequencer of equal load to the same physical server with the intent of
@@
 The stability of throughput curves (i.e. the variation on the \(y\) axis) indicate which sequencers
-are on overloaded physical servers.  ``Underloaded (Baseline)" is the best
+are on overloaded physical servers.  ``1 Sequencer Baseline" is the best
 performance for this particular workload because there is no contention.
 Performance is stable and the MDS is under-loaded because we do not see the
-spike at the end like the other curves.  ``Overloaded: No Balancing" shows an
+spike at the end like the other curves.  ``No Balancing" shows an
 overloaded sequencer because the performance of the monitored sequencer is low
@@
 The last two curves of Figure~\ref{fig:mantle-seq-thruput} show the performance
-improvements when sequencers are load balanced across the cluster.
-The load balancing policies differ by the threshold at which they start balancing:
-``Overloaded: Sequencer Policy" starts balancing immediately and ``Overloaded: Conservative
+improvements when sequencers are load balanced across the cluster. Both curves
+show the throughput of one of the sequencers when a second sequencer is
+running on the same physical server. The load balancing policies differ by the threshold at which they start balancing:
+``Sequencer Policy" starts balancing immediately and ``Conservative
 Policy" waits for 120 seconds before migrating sequencers.

### 2016-10-15 paper/paper.tex

--- 2016-10-14 paper/paper.tex
@@
 subsystems and leverage their optimizations, established correctness,
-robustness, and efficiency. We will refer to this programmability as
-\emph{programmable storage}, which differs from \emph{active storage} (the
-injection and execution of arbitrary code in a storage system) and
-\emph{software-defined storage} (the control of thin-provisioning of storage).
+robustness, and efficiency. 
+%We will refer to this programmability as
+%\emph{programmable storage}, which differs from \emph{active storage} (the
+%injection and execution of arbitrary code in a storage system) 
+%and
+%\emph{software-defined storage} (the control of thin-provisioning of storage).
+
+{\it \textbf{We define a programmable storage system}} to be a storage system that facilitates the 
+re-use and extension of existing storage abstractions and services provided by the underlying software stack,
+to enable the creation of new storage abstractions via composition.
+A programmable storage system is realized by exposing existing functionality (such as 
+metadata services and synchronization capabilities) as interfaces that can be ``glued together''
+using a high-level language.  Programmable storage differs from \emph{active storage},
+or the injection and execution of arbitrary code in a storage system.
 
@@
 
-{\it \textbf{The definition a programmable storage system}} to be ``a storage
-system that allows existing abstractions to be re-used and composed to form
-new data and metadata management services whose operation is orothogonal to
-system correctness".  Our prototype includes (but is not limited to) the
-following building blocks:
+Our prototype includes a non-exhaustive set of interfaces that can be used as building blocks
+for constructing novel storage abstractions, including:
 

### 2016-10-15 paper/paper.tex
tweak
tweaks

@@
 \usepackage{pifont}
+\usepackage{color}
 \newcommand{\cmark}{\ding{51}}%
+
+\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
 %\newcommand{\xmark}{\ding{55}}%
@@
 Storage systems are caught between rapidly changing data
-processing systems and the increasing speed of storage devices. This puts
-tremendous pressure on storage systems to adapt both in terms of their
+processing applications and the increasing speed of storage devices. This puts
+tremendous pressure on storage systems to support rapid evolution both in terms of their
 interfaces and their performance. But adapting storage systems can be difficult
@@
 demands~\cite{apache_contributors_parquet_2014}. 
+\paa{the previous sentence doesn't scan well. do we just want to say that 
+emerging ``big data'' applications demand interface evolution to support flexible
+consistency as well as flexible structured data representation?}
 Finally, production-quality
@@
 
-These three trends put evolutionary pressure on storage systems and raise the
-question of whether there are principles that storage systems designers can follow
+%These three trends put evolutionary pressure on storage systems and raise the
+%question of whether 
+The evolutionary pressure placed on storage systems by these trends raises the
+question of whether
+there are principles that storage systems designers can follow
 to evolve storage systems efficiently, without jeopardizing years of
@@
 
-In this paper we investigate an approach that focuses on generalizing and
-composing existing
+In this paper we investigate an approach that focuses on 
+%generalizing 
+%and
+%composing 
+identifying and exposing existing
 storage system resources, services, and abstractions that in a generalized form
@@
 {\it \textbf{We define a programmable storage system}} to be a storage system that facilitates the 
-re-use and extension of existing storage abstractions and services provided by the underlying software stack,
-to enable the creation of new storage abstractions via composition.
-A programmable storage system is realized by exposing existing functionality (such as 
-metadata services and synchronization capabilities) as interfaces that can be ``glued together''
+re-use and extension of existing storage abstractions provided by the underlying software stack,
+to enable the creation of new services via composition.
+A programmable storage system can be realized by exposing existing functionality (such as 
+metadata services and synchronization and monitoring capabilities) as interfaces that can be ``glued together''
+in a variety of ways
 using a high-level language.  Programmable storage differs from \emph{active storage},
@@
 To illustrate the benefits and challenges of this approach we have designed and
-evaluated Malacology, a programmable storage system for constructing new
+evaluated Malacology, a programmable storage system 
+that facilitates the construction of
+new
 services by re-purposing existing subsystem abstractions of the storage stack.
-We build the Malacology in Ceph, a popular open source software storage stack.
+We build Malacology in Ceph, a popular open source software storage stack.
 We choose Ceph to demonstrate the concept of programmable storage because it
@@
 
-{\it \textbf{We implement two distributed services}} using programmable storage
-to demonstrate its feasability:
+{\it \textbf{We implement two distributed services}} using 
+%programmable storage
+%to demonstrate its feasability:
+Malacology to demonstrate the feasibility of the programmable storage approach:
 
@@
 the best performance. For example, HDFS was designed specifically to serve
-Hadoop's jobs, and uses techniques like exposing data locality and relaxing POSIX constraints to achieve
+Hadoop's jobs~\paa{lol not sure if I buy that hadoop achieves ``best performance''}, and uses techniques like exposing data locality and relaxing POSIX constraints to achieve
 the best performance~\cite{CITEME}. Alternatively, general-purpose storage

### 2016-10-15 experiments/mds-zlog-seq-migrate-master/run.sh
mantle: experiments for proxy vs. client balancing mode
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -4,7 +4,7 @@ set -ex
 
 cd site
 ANSIBLE_LOG_PATH="logs/baseliner.log" \
-ansible-playbook --forks 50 --skip-tags "with_pkg"
+ansible-playbook --forks 50 --skip-tags "with_pkg" -e "site=mantle" \
   ceph.yml \
   ../workloads/zeqr-isolated.yml \
   ../workloads/zeqr-nobalancing.yml \

### 2016-10-17 paper/paper.tex
paper: apply peter's changes for services section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 authors showed how using Mantle, one can implement a single node metadata
-service, a distributed metadata services with hashing, and a distributed
+service, a distributed metadata service with hashing, and a distributed
 metadata service with dynamic subtree partitioning.
@@
 The original implementation was ``hard-coded" into Ceph and lacked robustness
-(lacking things such as versioning, durability, or reliable policy
-distribution).  Re-implemented using Malacology, Mantle now enjoys (1) the
-versioning of MONs and (2) the durability and distribution provided by RADOS.
-The version of the load balancer, which is stored in the MDS map, corresponds
-to an object name in RADOS that houses the load balancer code.
+(no versioning, durability, or policy distribution).  Re-implemented using
+Malacology, Mantle now enjoys (1) the versioning provided by the monitor
+daemons and (2) the durability and distribution provided by Ceph's reliable
+object store.
 
@@
 Ensuring that the version of the current load balancer is consistent across the
-MDS cluster was ignored in the original implementation. The user had to set the
-version on each individual MDS and it was trivial to make the version
-inconsistent across MDS nodes.
+physical servers in the metadata cluster was ignored in the original
+implementation. The user had to set the version on each individual server and
+it was trivial to make the version inconsistent.
 
 With Malacology, Mantle stores the version of the current load balancer in the
-MDS map.  Similar to the PG map or MON maps, the MDS map has all the
-information about the topology and state of the MDS cluster.  Recall that the
-MONs propogate these maps across the cluster and ensure that the map versions
-are consistent with PAXOS.  While storing the code snippets that define the
-policies in the MDS map is possible, it is not recommended as it is an
-expensive operation (a PAXOS voting round).
+Ceph map using the service metadata interface. The version of the load balancer
+corresponds to an object name in the balancing policy. Using the service
+metadata interface ensures that all metadata servers use the same version of
+the load balancer. As a result, the policy version inherits the consensus and
+consistency of Ceph's internal monitor daemons.
 
@@
 
-\subsection{ZLog: A Fast Distributed Shared-Log}
+\subsection{ZLog: A Fast Distributed Shared Log}
 \label{sec:zlog}
@@
 distributed shared-log that is based on the CORFU
-protocol~\cite{balakrishnan_corfu_2012}. A
-shared-log is a general, yet powerful abstraction used to build many
-distributed systems. However, existing implementations that rely on consensus
-algorithms such as PAXOS funnel I/O through a single point introducing a
-bottleneck that restricts throughput. In contrast, the CORFU protocol is able
-to achieve high throughput using a network counter service, called a 
-\emph{sequencer}, that avoids all persistent I/O in the common case.
+protocol~\cite{balakrishnan_corfu_2012}. A shared-log is a general, yet
+powerful abstraction used to build many distributed systems, such as
+cloud-based metadata management~\cite{balakrishnan:sosp13} and elastic database
+storage engines~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
+However, existing implementations that rely on consensus algorithms such as
+PAXOS funnel I/O through a single point introducing a bottleneck that restricts
+throughput.  In contrast, the CORFU protocol is able to achieve high throughput
+using a network counter service, called a \emph{sequencer}, for obtaining the
+entry at the tail of log. This metadata operation that avoids all persistent
+I/O in the common case.
 
 While a full description of the CORFU system is beyond the scope of this
-paper, we'll briefly describe the custom storage device interface, sequencer
+paper, we briefly describe the custom storage device interface, sequencer
 service, and recovery protocol, and how these services are instantiated in the
@@
 log positions to clients by reading from a volatile, in-memory counter which
-can run at a very high throughput and at low-latency. And since the sequencer
+can run at a very high throughput and at low latency. Since the sequencer
 is centralized, ensuring serializability in the common case is trivial.  The
@@
 latency for an approach based on a round-robin batching mode, and compare this
-mode to one in which the metadata server mediates access to the sequencer
-state when in it is being shared among multiple clients.
+mode to one in which the metadata server mediates access to the sequencer state
+when in it is being shared among multiple clients. Quantifying these trade-offs
+should provide administrators with guidelines for setting the tunables for
+different ``caching" modes of the sequencer.
 
@@
 process itself may be handled in many ways, such as leader election using an
-auxiliary service such as PAXOS.
+auxiliary service such as PAXOS. In our implementation, the recovery is the
+same (and is inherited from) the Ceph metadata service.
 
 \section{Evaluation}
-\label{evaluation}
+\label{sec:evaluation}
 
@@
 \subsection{Mapping ZLog onto Malacology}
-\label{mapping-zlog-onto-malacology}
+\label{sec:mapping-zlog-onto-malacology}
 We evaluate the mapping of ZLog onto Malacology by examining the costs of

### 2016-10-17 paper/paper.tex
tweaks

@@
 
-Building storage stacks from the ground-up for a specific purpose results in
+Building storage stacks from the ground up for a specific purpose results in
 the best performance. For example, HDFS was designed specifically to serve
@@
 (OSDs) that provide Ceph with data durability and integrity using replication,
-erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already has some
+erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already 
+%has some
+provides some degree of
 programmability; the OSDs support domain-specific object interfaces that are
@@
 
-The Ceph community provides empirical evidence that developers embrace
+The Ceph community provides empirical evidence that developers are already beginning to embrace
 programmable storage. Figure~\ref{fig:obj-int-dev-growth} shows a dramatic
@@
 that API changes are a burden for users.  Rather it appears that a gap in
-existing interfaces are being addressed through programmability. In fact,
+existing interfaces are being addressed through ad-hoc approaches to programmability. In fact,
 Table~\ref{table:objclasses} categorizes existing interfaces and we clearly see

### 2016-10-17 experiments/zlog-seqr-redux/run.sh
seq: add test
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -1,0 +1,16 @@
+#!/bin/bash
+
+set -e
+set -x
+
+CEPH_BIN=/home/nwatkins/ceph/build/bin
+export CEPH_CONF=/home/nwatkins/ceph/build/ceph.conf
+
+runtime=30
+capdelay="0.0"
+quota=0
+for nclients in 1 2 4 8 16; do
+  perf_file="seq-client.nc-${nclients}.cd-${capdelay}.qa-${quota}"
+  $CEPH_BIN/seq-client --runtime=$runtime --perf_file=${perf_file} \
+    --instances=$nclients --capdelay=$capdelay --quota=${quota}
+done

### 2016-10-17 experiments/mds-zlog-seq-migrate-master/run.sh
mantle: more baseline benchmarks
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -2,14 +2,22 @@
 
 set -ex
 
+ANSIBLE="ansible-playbook --forks 50 --skip-tags with_pkg"
+
 cd site
-ANSIBLE_LOG_PATH="logs/baseliner.log" \
-ansible-playbook --forks 50 --skip-tags "with_pkg" -e "site=mantle" \
-  ceph.yml \
-  ../workloads/zeqr-isolated.yml \
-  ../workloads/zeqr-nobalancing.yml \
-  mantle.yml \
-  ../workloads/zeqr-mantle.yml
+
+export ANSIBLE_LOG_PATH="../logs/setup.log"
+$ANSIBLE ceph.yml ../workloads/zeqr-setup.yml
+
+for j in 0 1 2; do
+  for i in `seq 1 6`; do
+    site="${i}seq"
+    export ANSIBLE_LOG_PATH="../logs/${site}.log"
+    $ANSIBLE -i "inventory/hosts-${site}" -e "site=${site}-baseline" ../workloads/zeqr-multiclient.yml
+  done
+  mv ../results ../results-baseline-run${j}; mkdir ../results
+done
+
 cd -
 
 tar czf ../mds-zlog-seq-migrate-master.tar.gz ../mds-zlog-seq-migrate-master

### 2016-10-17 paper/paper.tex
staged
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-10-17 paper/paper.tex
@@
 Malacology our evaluation focuses on the performance of the internal
-abstractions that support the Malacogy interfaces. First, we benchmark
-scenarios where contention is at the sequencer by examining the intefaces used
-to map ZLog onto Malacology; specifically, we describe the sequencer
-implementation and the propogation logical device interfaces. Next, we
-benchmark scenarios when there are multiple logs by using Mantle to (1)
-balance sequencers across a cluster and (2) recover from failure.
+abstractions when used for to construct the Mantle and ZLog services. First, we
+benchmark scenarios where contention is at the sequencer by examining the
+intefaces used to map ZLog onto Malacology; specifically, we describe the
+sequencer implementation and the propogation of object \& data interfaces
+interfaces. Next, we benchmark scenarios when there are multiple logs by using
+Mantle to (1) balance sequencers across a cluster and (2) recover from failure.
 
@@
 We evaluate the load balancing interface by showing how the cluster of servers
-housing sequencers alleviate load.  The experiments use our internal cluster
-which uses 10 nodes to store objects, one node to monitor the cluster, and 3
-nodes that can accomodate sequencers. The workload is two logs, one with 7
-clients and the other with 8 clients; we chose this workload because this load
-is the point at which a single physical server gets overloaded.
+housing sequencers alleviates load.  The experiments are run on a cluster with
+10 nodes to store objects, one node to monitor the cluster, and 3 nodes that
+can accomodate sequencers. The workload has up to 6 logs each with 2 clients.
+Figure~\ref{CITEME} shows the throughput per sequencer when all seqeuencers are
+co-located of one physical server. Two sequencers on the same server show no
+performance degradation but co-locating 3, 4, 5, or 6 sequencers leads to a
+2x, 3x, 4x, and 5x slowdown, respectively. For the rest of the experiments we
+load balance 3, 4, 5, or 6 sequencers across the cluxter because this is hte
+point at which a single physical server gets overloaded.
+
+Figure~\ref{CITEME} shows the performance profiles of the sequencers as they
+migrate around the cluster. Each graph in that figure shows different ``waves"
+of sequencers, where a ``wave" is a burst of clients trying to append to the
+end of the log. These waves are examples of hotspots and flash crowds; they are
+dynamic and resource constrained because clients access a shared resource. Wave
+1, policy 1 shows how the sequencers initially get distributed poorly. From x
+to x seconds, all seqeuencers are co-located. At x seconds, 3 of the sequencers
+(1, 2, 3) migrate off the first server and onto the second one; this is what we
+want.
 

### 2016-10-18 experiments/mds-zlog-seq-migrate-redux-waves/run.sh
mantle: add performance profiles
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,23 @@
+#!/bin/bash
+
+set -ex
+
+ANSIBLE="ansible-playbook --forks 50 --skip-tags with_pkg"
+
+cd site
+
+export ANSIBLE_LOG_PATH="../logs/setup.log"
+$ANSIBLE ceph.yml ../workloads/zeqr-setup.yml
+
+for j in 0 1 2; do
+  for i in `seq 1 6`; do
+    site="${i}seq"
+    export ANSIBLE_LOG_PATH="../logs/${site}.log"
+    $ANSIBLE -i "inventory/hosts-${site}" -e "site=${site}-baseline" ../workloads/zeqr-multiclient.yml
+  done
+  mv ../results ../results-baseline-run${j}; mkdir ../results
+done
+
+cd -
+
+tar czf ../mds-zlog-seq-migrate-master.tar.gz ../mds-zlog-seq-migrate-master

### 2016-10-18 paper/paper.tex
mantle: add benchmarks to paper
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
+\begin{figure}[t!]
+\centering
+\includegraphics{figures/mantle-balancer-profiles.png}
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
+load balancing interface alleviates load for multiple sequencers in
+resource-constrained environments.  ``Underloaded (Baseline)'' is the
+throughput of one sequencer running alone on a cluster; the other curves show
+how load balancing affects the throughput of one sequencer when sequencers for
+other logs are initially co-located on the same physical server (i.e. an
+"Overloaded" environment).}\label{fig:mantle-seq-thruput}
+\end{figure}
+
+\begin{figure}[t!]
+\centering
+\includegraphics{figures/mantle-modes.png}
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
+load balancing interface alleviates load for multiple sequencers in
+resource-constrained environments.  ``Underloaded (Baseline)'' is the
+throughput of one sequencer running alone on a cluster; the other curves show
+how load balancing affects the throughput of one sequencer when sequencers for
+other logs are initially co-located on the same physical server (i.e. an
+"Overloaded" environment).}\label{fig:mantle-seq-thruput}
+\end{figure}
+
 We use a balancer policy specifically designed for multiple sequencers. The

### 2016-10-18 paper/paper.tex
paper updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 %users to continue to entrust their data to the storage system.
-
 In this paper we investigate an approach that focuses on 
@@
 in a variety of ways
-using a high-level language.  Programmable storage differs from \emph{active storage},
-or the injection and execution of arbitrary code in a storage system.
+using a high-level language. In contrast to programmable storage, 
+\emph{active storage} differs through its focus on the injection and
+execution of code within a storage system or storage device.
 
@@
 
-Our prototype includes a non-exhaustive set of interfaces that can be used as building blocks
-for constructing novel storage abstractions, including:
+A prototype programmable storage system that includes a non-exhaustive set of
+interfaces that can be used as building blocks for constructing novel storage
+abstractions, including:
 
@@
 the best performance. For example, HDFS was designed specifically to serve
-Hadoop's jobs~\paa{lol not sure if I buy that hadoop achieves ``best performance''}, and uses techniques like exposing data locality and relaxing POSIX constraints to achieve
-the best performance~\cite{CITEME}. Alternatively, general-purpose storage
+Hadoop's jobs, and uses techniques like exposing data locality and relaxing POSIX constraints to
+to achieve application-specific I/O optimizations.~\cite{CITEME}. Alternatively, general-purpose storage
 stacks are built with the flexiblity to serve many applications by providing a
@@
 community since 2010. In that figure, classes are functional groupings of methods on storage
-objects.  What is most remarkable is that this trend contradicts the notion
+objects (e.g. remotely computing and caching the checksum of an object extent).  What is most remarkable is that this trend contradicts the notion
 that API changes are a burden for users.  Rather it appears that a gap in
@@
 interface facility of Ceph could be due to a number of reasons, such as the
-default algorithms/tunables of the storage system are insufficient for the
-application's performance goals, programmers want to exploit
-application-specific semantics, and/or programmers know how to manage resources
-to improve performance. Their solution is a way to work around the
+default algorithms/tunables of the storage system being insufficient for the
+application's performance goals, programmers wanting to exploit
+application-specific semantics, and/or programmers knowing how to manage resources
+to improve performance. A solution based on application-specific object
+interfaces is a way to work around the
 traditionally rigid storage APIs because custom object interfaces give
@@
 making programmability a `feature', rather than a `hack' or `workaround', we
-help standardize a development process that is largely ad-hoc.
+help standardize a development process that now is largely ad-hoc.
 
@@
 Ceph. We choose Ceph because it is a production quality system and because it
-is open source. The large developer community ensures that code is relatively
+is open source. The large developer community ensures that code is
 robust and the visibility of the code lets us expose any interface we want. In
@@
 and off \cite{sevilla:sc15-mantle} the performance critical path. Finally, the
-flexibility of the runtime allows to sandbox applications in order to address
+flexibility of the runtime allows execution sandboxing in order to address
 security and performance concerns.
@@
 and clients is critical to provide strong consistency guarantees to clients.
-Ceph maintains cluster state information in per-subsystem ``maps'' that record
-membership and status information.  A PAXOS-based monitoring service is
+Ceph maintains cluster state information in per-subsystem data structures
+called ``maps'' that record
+membership and status information.  A Paxos-based monitoring service is
 responsible for integrating state changes into cluster maps, responding to
@@
 state. As a fundamental building block of many system designs, consensus
-abstractions such as PAXOS are a common technique for maintaining consistent
+abstractions such as Paxos are a common technique for maintaining consistent
 data versions, and are a useful system to expose.
 
-The default behavior of the monitor can be seen as a PAXOS-based notification
+The default behavior of the monitor can be seen as a Paxos-based notification
 system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing
@@
 However, existing implementations that rely on consensus algorithms such as
-PAXOS funnel I/O through a single point introducing a bottleneck that restricts
+Paxos funnel I/O through a single point introducing a bottleneck that restricts
 throughput.  In contrast, the CORFU protocol is able to achieve high throughput
@@
 process itself may be handled in many ways, such as leader election using an
-auxiliary service such as PAXOS. In our implementation, the recovery is the
+auxiliary service such as Paxos. In our implementation, the recovery is the
 same (and is inherited from) the Ceph metadata service.
@@
 costs of such a service by embedding interface definitions in a Ceph cluster
-membership data structure that is managed by a PAXOS instance, and distributed
+membership data structure that is managed by a Paxos instance, and distributed
 using a scalable, peer-to-peer gossip protocol that is built into Ceph. We
@@
 distributing an interface update on all OSDs in the cluster. The latency is
-defined as the elapsed time following the PAXOS proposal for an interface
-update until each OSD makes the update live (the cost of the PAXOS proposal is
+defined as the elapsed time following the Paxos proposal for an interface
+update until each OSD makes the update live (the cost of the Paxos proposal is
 configurable and is discussed below). The latency measurements were taken on
@@
 cluster-wide propagation of interface updates, includes the network round-trip
-to the interface management service, the PAXOS commit protocol, and other
-factors such as system load. By default PAXOS proposals occur periodically with
+to the interface management service, the Paxos commit protocol, and other
+factors such as system load. By default Paxos proposals occur periodically with
 a 1 second interval in order to accumulate updates. In a minimum, realistic

### 2016-10-18 paper/paper.tex
paper updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 optimizations~\cite{gray_tape_2007,gray_flash_2008}. 
-Second, the demand for
-managing structured data and flexible consistency semantics at scale pressure
-big data processing systems to use storage abstractions that can meet these
-demands~\cite{apache_contributors_parquet_2014}. 
-\paa{the previous sentence doesn't scan well. do we just want to say that 
-emerging ``big data'' applications demand interface evolution to support flexible
-consistency as well as flexible structured data representation?}
+Second, emerging ``big data'' applications demand interface evolution to
+support flexible consistency as well as flexible structured data
+representations.  ~\cite{apache_contributors_parquet_2014}. 
 Finally, production-quality
@@
 abstractions nor do we contend that the abstractions we choose provide the best
-trade-offs for all applications. For example, if consensus is correctly
-exposed, one could implement versioning, serialization, or capabilities on top;
-but perhaps versioning, serialization, and capabilities would be better
-interfaces for a given set of applications. These questions are not answered in
+trade-offs for all applications.
+For example, if consensus is correctly exposed one could implement
+high-level features like versioning, serialization, or various flavors
+of strongly consistent data management on top; but perhaps a low-level
+consensus interface is suited well for a particular set of applications.
+These questions are not answered in
 this paper and instead we focus on showing the feasibility of building such a
@@
 clients to identify when new values (termed epochs in Ceph) are associated to
-given maps. While Ceph does not expose this service directly, it does allow
-limited use by clients through a key-value service intended for low-volume,
-small size read-only configuration information. While a key-value service is
-generally useful, it doesn't provide many of the useful services hidden within
-the monitoring framework that would be required for applications with more
-demanding requirements, such as creating arbitrary maps or associating
+given maps.  While Ceph does not expose this service directly, a key-value
+service designed for managing configuration metadata that is built on top of
+the consensus engine is an example of a high-level service within Ceph that
+re-uses existing abstractions.  While a key-value service is generally useful,
+it doesn't provide many of the useful services hidden within the monitoring
+framework that would be required for applications with more demanding
+requirements, such as creating arbitrary maps or associating
 application-specific logic that can be executed for particular maps (or values

### 2016-10-18 paper/paper.tex
updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 \newcommand{\cmark}{\ding{51}}%
+\usepackage[font=footnotesize,labelfont=bf]{caption}
 
@@
 {\it Service Metadata}
-  & Zookeeper coordination~\cite{hunt_zookeeper_2010}, Chubby~\cite{website:chubby}
+  & Zookeeper, Chubby coordination~\cite{hunt_zookeeper_2010,website:chubby}
   & cluster state management~\cite{website:ceph-mon}
@@
   \\
-Beacon
-  & Apache RabbitMQ~\cite{CITEME}
-  & inter-daemon messages, centralized logging
-  & debugging \& monitoring
-  \\
 \end{tabular}
@@
 framework that would be required for applications with more demanding
-requirements, such as creating arbitrary maps or associating
-application-specific logic that can be executed for particular maps (or values
-in a map).  Since the monitor is intended to be out of high-performance I/O
+requirements, such as creating arbitrary maps with application-specific
+semantics and interfaces. Since the monitor is intended to be out of high-performance I/O
 paths, a general guideline is to make use of this functionality infrequently
@@
 functionality to register, version and propagate dynamic code (Lua scripts) for
-new object interfaces defined in storage daemons (\S\ref{active-storage}) and
+new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
 policies in the load balancer \S\ref{malacology:mds}).  Using this service
 guarantees that interface definitions are not only made durable, but are
-transparently and consistently propagated throughput the cluster and that
+transparently and consistently propagated throughout the cluster so that
 clients are properly synchronized with the latest interfaces.
@@
 
-\subsection{Object \& Data Interface}
+\subsection{Object I/O Interfaces}
 \label{object-data-interface}
@@
 interfaces or other custom object interfaces, and handle serialization of
-function input and output. A wide range of native interfaces are available to
-developers, such as reading and writing to a byte stream, and controlling
-snapshots and clones. These object interfaces may be transactionally composed
-with access to a sorted key-value service allowing applications to create rich
-semantics. An example is using a read-modify-write guard to protect a shared
-index defined over content stored in a byte stream.
+function input and output.
+A wide range of native interfaces are available to developers such as reading
+and write to a byte stream, controlling object snapshots and clones,
+and accessing a sorted key-value database. These native object interfaces
+may be transactionally composed along with application specific logic to
+create interfaces with a rich set of semantics.
+An example is the use of a read-modify-write operation to perform consistent
+object updates across the byte stream and key-value database interfaces
+without relying on complex client-driven concurrency control.
 
@@
 are expressed in thousands of lines of code can be implemented in approximately
-an order of magnitude less code~\cite{geambasu_comet_2010}.
+an order of magnitude less code~\cite{geambasu_comet_2010}. While the use of
+Lua does not prevent deployment of malicious code, certain types of coding
+mistakes can handled gracefully, and access policies are used to restrict the
+feature to trusted users.
 

### 2016-10-18 paper/paper.tex
updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 metadata cluster. This cluster uses load balancing policies to migrate
-directory inodes around the cluster to alleviate load. The policies use metrics
+directory inodes around the cluster to alleviate overload at single nodes. The policies use metrics
 based on system state (e.g.  CPU and memory utilization) and statistics
 collected by the cluster (e.g. the popularity of an inode). Ceph uses dynamic
-subtree partitioning to move variable sized namespace subtrees. These units can
+subtree partitioning to move variable sized namespace
+subtrees~\cite{weil:sc2004-dyn-metadata}. These units can
 be shipped anywhere (i.e., to any metadata server of any capacity) at any time
@@
 
-{\bf Shared Resource Interface.} The metadata service also manages sessions
-with its clients. This gives both subsystems (metadata servers and clients) the
-ability to cache data and obtain locks. The metadata service is responsible for
-responding to client requests for capabilities by first revoking caps from
-other clients.  Currently, clients voluntarily release their capabilities, and
-the metadata server maintains a queue of requests.  Both clients and metadata
-servers use a best-effort approach to revoking or releasing capabilities.
-
-The clients retain the best-effort, voluntary release of capabilities, but in
-Malacology, the queuing policies are generalized within the MDS to allow for
-centralized policies such as fairness or priority.
+{\bf Shared resource interface.}
+The metadata service manages client sessions, allowing clients to obtain locks
+(e.g. file byte ranges), and capabilities (e.g. to cache file data). Clients
+and metadata servers use a cooperative protocol in which clients voluntarily
+release resources back to the metadata service in order to implement sharing
+policies. While the current policy for sharing access and voluntarily
+releasing resources is largely best-effort, Malacology supports generalized
+policies between metadata servers and clients that can be used to implement 
+fairness or priority.
 
@@
 safely load, manage, and persist its load balancing policies.  Composing Mantle
-using the service metadata, durability and beacon interfaces, Mantle can safely
-version balancer policies, save balancer policies in the back end object store
+using the service metadata and durability interfaces, Mantle can safely
+version balancer policies, save balancer policies in the backend object store
 and centralize warnings/errors.

### 2016-10-18 paper/paper.tex
updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 uses many techniques to ensure that data is not corrupted or lost, such as
-erasure coding, replication, and data scrubbing. Furthermore, many of these
-techniques try to be autonomous so that work is distributed across the
-cluster. For example, when placement groups change, the OSDs re-balance and
-re-shard data in the background in a process called placement group splitting.
-
+erasure coding, replication, and data scrubbing.
+Furthermore, many of these features are implemented using a peer-to-peer
+protocol that allows object storage devices to operate autonomously without
+a centralized coordinator.
+For example, when the number of placement groups change, the OSDs re-balance and
+re-shard data in the background in a process called placement group splitting
+in which OSDs communicate directly with each other to coverage on a new state.
 In order to reduce load on the monitoring service, Ceph OSDs use a gossip
@@
 (no versioning, durability, or policy distribution).  Re-implemented using
-Malacology, Mantle now enjoys (1) the versioning provided by the monitor
-daemons and (2) the durability and distribution provided by Ceph's reliable
-object store.
+Malacology, Mantle now enjoys (1) the versioning provided by the metadata
+service interface provided by the monitor daemons and (2) the durability and
+distribution provided by Ceph's reliable object store.
 
@@
 implementation. The user had to set the version on each individual server and
-it was trivial to make the version inconsistent.
+it was trivial to make the version inconsistent. Maintaining a consistent
+version is important for balancing policies that are cooperative in which
+local decisions are made assuming properties about other instances of the same
+policy.
 
 With Malacology, Mantle stores the version of the current load balancer in the
-Ceph map using the service metadata interface. The version of the load balancer
+Ceph service metadata interface. The version of the load balancer
 corresponds to an object name in the balancing policy. Using the service
 metadata interface ensures that all metadata servers use the same version of
-the load balancer. As a result, the policy version inherits the consensus and
+the load balancer. As a result, the policy version inherits the
 consistency of Ceph's internal monitor daemons.
 
-The user changes the version of the load balancer using a new CLI command:
-
-\noindent \texttt{ fs\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}
+The user changes the version of the load balancer using a new CLI command: \\
+
+\noindent \texttt{\$> ceph fs\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}
 
@@
 The load balancer version described above corresponds to the name of an object
-in RADOS that holds the actual Lua balancing code. The objects containing load
-balancer code are stored in a system-maintained group of objects (i.e. a pool).
+in RADOS that holds the actual Lua balancing code.
 When MDS nodes start balancing load, they first check the latest version from
@@
 which stored load balancer code on the local file system -- a technique which
-is unreliable and results in inconsistencies if any MDS accidently modifies the
-code.
+is unreliable and may result in silent corruption.
 
@@
 In the original implementation, Mantle would log all errors, warnings, and
-debug messages to the local log on each MDS. To get the simplest status
-messages or to debug problems, the user would have to log into each MDS
-individually, look at the logs, and reason about their causality.
-
-With Malacology, Mantle re-usues the centralized logging features of the MON
-subsystem. Important errors, warnings, and info messages are sent with the MON
-beacon subsystem and appear in the monitor cluster log so instead of users
+debug messages to a log stored locally on each MDS server. To get the simplest
+status messages or to debug problems, the user would have to log into each MDS
+individually, look at the logs, and reason about causality and ordering.
+
+With Malacology, Mantle re-usues the centralized logging features of the
+monitoring service. Important errors, warnings, and info messages are
+collected
+by the monitoring subsystem and appear in the monitor cluster log so instead of users
 going to each node, they can watch messages appear at the monitor. Messages are
-used discretly, so as not to spam the monitor with frivolous debugging but the
-large events, like balancer version changes or failed subsystems show up in the
+logged sparingly, so as not to spam the monitor with frivolous debugging but
+important events, like balancer version changes or failed subsystems show up in the
 centralized log.

### 2016-10-18 paper/paper.tex
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 \paragraph*{Malacology:} interfaces are added in strategic spots in the
-metadata service for guarding shared resources and load balancing.
+metadata service for guarding shared resources, defining new file types,
+and load balancing.
 
@@
 
+{\bf File types.} The metadata service manages a POSIX file system hierarchy
+where files and directories are represented as inode data structures and
+expose a POSIX file interface. Application that manage large amounts of
+metadata (e.g. users or database snapshots) often require a naming service. We
+allow new inode types to be defined such that applications can create
+domain-specific interfaces to inodes that may modify locking and capability
+policies. We will show this is used in Section~\ref{sec:seq} when we discuss a
+distributed shared-log built on Malacology.
+
 {\bf Load balancing.} Mantle~\cite{sevilla:sc15-mantle}, a programmable
@@
 version balancer policies, save balancer policies in the backend object store
-and centralize warnings/errors.
+and centralize warnings/errors. When combined with file types, load balancing
+policies can express a variety of rules such as multi-tentant and
+multi-application balancing.
 
@@
 protocol~\cite{balakrishnan_corfu_2012}. A shared-log is a general, yet
-powerful abstraction used to build many distributed systems, such as
+powerful abstraction used to construct distributed systems, such as
 cloud-based metadata management~\cite{balakrishnan:sosp13} and elastic database
@@
 throughput.  In contrast, the CORFU protocol is able to achieve high throughput
-using a network counter service, called a \emph{sequencer}, for obtaining the
-entry at the tail of log. This metadata operation that avoids all persistent
-I/O in the common case.
+using a network counter service, called a \emph{sequencer}, that is used to
+decouple log position assignment from log I/O.
 
@@
 \subsubsection{Sequencer}
+\label{sec:seq}
 
@@
 
+{\bf Sequencer interface.} The sequencer resource supports the ability to
+\emph{read()} the current tail value as well as get the \emph{next()} position in
+the log which also atomically increments the tail position.
 We implement the sequencer service in Malacology as an application-specific
@@
 representing each log service instance within the standard POSIX hierarchical
-namespace.
-
-{\bf Sequencer interface.} The sequencer resource supports the ability to
-\emph{read()} the current tail value as well get the \emph{next()} position in
-the log which also atomically increments the tail position. We implement this
-functionality as methods on ZLog inode resources. The primary challenge in
-mapping the sequencer resource to the metadata service is handling
-serialization correctly.
+namespace. The primary challenge in mapping the sequencer resource to the
+metadata service is handling serialization correctly to maintain the global
+ordering provided by the CORFU protocol.
 
@@
 inode.  Interestingly, we found that the capability system in Ceph uses a
-strategy to reduce metadata service load by allowing multiple clients that
-open a shared file to temporarily cache the inode resource, resulting in a
-round-robin, best-effort batching behavior. When a single client is accessing
-the sequencer resource it is able to cache state locally without interruption.
+strategy to reduce metadata service load by allowing clients that open a
+shared file to temporarily obtain an exclusive cached copy of the resource,
+resulting in a round-robin, best-effort batching behavior. When a single
+client is accessing the sequencer resource it is able to increment the
+sequencer locally. Any competing client cannot query the sequencer until the
+metadata service has granted it exclusive access.
 
@@
 
-The storage interface is a critical component in the CORFU design. Clients
-independently map log positions that they have obtained from a sequencer
+The storage interface is a critical component in the CORFU protocol. Clients
+independently map log positions that they have obtained from the sequencer
 service (described in detail in the next section) onto storage devices, while
@@
 
-In order to repopulate the cached tail value during recovery of a sequencer,
+In order to repopulate the sequencer state (i.e. the cached, current tail of
+the log) during recovery of a sequencer,
 the maximum position in the log must be obtained. To do this, the storage
@@
 new epoch value and returns the maximum log position that has been written.
-Since the recovery protocol does not inform clients of a new epoch value until
-the end of the process, the validity of the tail can be guaranteed despite the
-iterative process of contacting storage devices.  Recovery of the sequencer
-process itself may be handled in many ways, such as leader election using an
-auxiliary service such as Paxos. In our implementation, the recovery is the
-same (and is inherited from) the Ceph metadata service.
+
+Since the sequencer servivce does not resume until the recovery process has
+completed, there cannot be a race with clients appending to the log, and the
+immutability of the log allow reads to never block during a sequencer failure.
+Recovery of the sequencer process itself may be handled in many ways, such as
+leader election using an auxiliary service such as Paxos. In our
+implementation, the recovery is the same (and is inherited from) the Ceph
+metadata service. Handling the failure of a client that holds the sequencer
+state is similar, although a timeout is used to determine when a client should
+be considered unavailable.
 

### 2016-10-18 paper/paper.tex
Adds discussion on SDS and Active Storage
Fixes #93

@@
 robustness, and efficiency.
-%We will refer to this programmability as
-%\emph{programmable storage}, which differs from \emph{active storage} (the
-%injection and execution of arbitrary code in a storage system) 
-%and
-%\emph{software-defined storage} (the control of thin-provisioning of storage).
 
@@
 \emph{active storage} differs through its focus on the injection and
-execution of code within a storage system or storage device.
+execution of code within a storage system or storage device. Programmable 
+storage differs from \emph{active storage}~\cite{riedel:vldb98} in that the 
+former is applicable to any component of the storage system, while the latter 
+focuses at the data access level (i.e. inject and execute arbitrary code at the 
+I/O access layers). Given this contrast, we can say that active storage is an 
+example of how one internal component (the storage layer) is exposed in a 
+programmable storage system.
 
@@
 notion of control/data separation was also applied in software-defined storage
-(SDS)~\cite{arpaci:sosp01,thereska:sosp13,stefanovici:fast16}. Similarly,
+(SDS)~\cite{thereska:sosp13,stefanovici:fast16}. Similarly,
 IOStack~\cite{gracia:internet16} is providing policy-based provisioning and
-filtering in OpenStack Swift.
+filtering in OpenStack Swift. According to a SNIA white 
+paper~\cite{carlson:snia2014}, the primary goal of SDS is to control and 
+facilitate flexible and dynamic provisioning of storage resources of different 
+kinds, including flash memory and disk drives, to create a virtualized mapping 
+between common storage abstractions (e.g. files, objects, and blocks) and 
+storage devices taking data service objectives in terms of protection, 
+availability, performance, and security into account. A programmable storage 
+system exposes internal abstractions so that end users (not necessarily 
+operators) can create new services on top of the storage stack. Thus, our notion 
+of programmable storage differs from ``software-defined storage'' (SDS) in terms 
+of goals and scope, although definitions of SDS are still in flux.
 

### 2016-10-19 experiments/zlog-seqr-redux/run.sh
results
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -3,14 +3,15 @@
 set -e
 set -x
 
-CEPH_BIN=/home/nwatkins/ceph/build/bin
-export CEPH_CONF=/home/nwatkins/ceph/build/ceph.conf
+CEPH_BIN=/home/nwatkins/src/ceph/build/bin
+export CEPH_CONF=/home/nwatkins/src/ceph/build/ceph.conf
 
-runtime=30
+prefix=$1
+runtime=60
 capdelay="0.0"
 quota=0
 for nclients in 1 2 4 8 16; do
-  perf_file="seq-client.nc-${nclients}.cd-${capdelay}.qa-${quota}"
+  perf_file="${prefix}.nc-${nclients}.cd-${capdelay}.qa-${quota}"
   $CEPH_BIN/seq-client --runtime=$runtime --perf_file=${perf_file} \
     --instances=$nclients --capdelay=$capdelay --quota=${quota}
 done

### 2016-10-19 experiments/zlog-seqr-redux/run.sh
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -8,9 +8,10 @@ export CEPH_CONF=/home/nwatkins/src/ceph/build/ceph.conf
 
 prefix=$1
 runtime=60
-capdelay="0.0"
+capdelay="1.0"
 quota=0
-for nclients in 1 2 4 8 16; do
+nclients=4
+for capdelay in 1.0 0.5 0.25 0.125 .0625 0.03125; do
   perf_file="${prefix}.nc-${nclients}.cd-${capdelay}.qa-${quota}"
   $CEPH_BIN/seq-client --runtime=$runtime --perf_file=${perf_file} \
     --instances=$nclients --capdelay=$capdelay --quota=${quota}

### 2016-10-19 paper/paper.tex
paper: generalize cls table, re-align text blocks, fix service table (#99)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-%\documentclass[10pt,twocolumn]{article}
 \documentclass[preprint]{sigplanconf-eurosys}
-
-%\usepackage{fullpage}
-%\usepackage{times}
 \usepackage{multirow}
 \usepackage{arydshln}
-%\usepackage{amssymb,amsmath}
-%\usepackage{ifxetex,ifluatex}
 \usepackage{amsmath}
-%\usepackage{mathptmx}
 \usepackage{graphicx}
 \usepackage{listings}
-%\usepackage{amssymb}
 \usepackage{pifont}
@@
 \newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
-%\newcommand{\xmark}{\ding{55}}%
-%
-%
-%
-%% use upquote if available, for straight quotes in verbatim environments
-%\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-%
-%\usepackage[font=footnotesize,labelfont=bf]{caption}
 \usepackage{pstricks}
@@
 \makeatother
-%% Scale images if necessary, so that they will not overflow the page
-%% margins by default, and it is still possible to overwrite the defaults
-%% using explicit options in \includegraphics[width, height, ...]{}
 \setkeys{Gin}{width=\maxwidth,keepaspectratio}
@@
 \urlstyle{same}  % don't use monospace font for urls
-%%\setlength{\parindent}{0pt}
-%%\setlength{\parskip}{6pt plus 2pt minus 1pt}
-%%\setlength{\emergencystretch}{3em}  % prevent overfull lines
-%\setcounter{secnumdepth}{5}
 
@@
 \begin{abstract}
-Storage systems are caught between rapidly changing data
-processing applications and the increasing speed of storage devices. This puts
-tremendous pressure on storage systems to support rapid evolution both in terms of their
+Storage systems are caught between rapidly changing data processing
+applications and the increasing speed of storage devices. This puts tremendous
+pressure on storage systems to support rapid evolution both in terms of their
 interfaces and their performance. But adapting storage systems can be difficult
@@
 programmability by composing existing primitives into two new higher-level
-services: a file system metadata load balancer and a high-performance distributed
-shared-log that leverages flash devices. The evaluation demonstrates that our
-services inherit desirable qualities of the back-end storage system, including
-the ability to balance load, efficiently propagate cluster metadata, recover
-from failure, and to make trade-offs for latency/throughput using leases.
+services: a file system metadata load balancer and a high-performance
+distributed shared-log that leverages flash devices. The evaluation
+demonstrates that our services inherit desirable qualities of the back-end
+storage system, including the ability to balance load, efficiently propagate
+cluster metadata, recover from failure, and to make trade-offs for
+latency/throughput using leases.
 \end{abstract}
@@
 code paths and incorporate new
-optimizations~\cite{gray_tape_2007,gray_flash_2008}. 
-Second, emerging ``big data'' applications demand interface evolution to
-support flexible consistency as well as flexible structured data
-representations.  ~\cite{apache_contributors_parquet_2014}. 
-Finally, production-quality
-scalable storage systems available as open source software have established and
-are continuing to establish new, \emph{de-facto} API standards at a faster pace
+optimizations~\cite{gray_tape_2007,gray_flash_2008}.  Second, emerging ``big
+data'' applications demand interface evolution to support flexible consistency
+as well as flexible structured data representations.
+~\cite{apache_contributors_parquet_2014}.  Finally, production-quality scalable
+storage systems available as open source software have established and are
+continuing to establish new, \emph{de-facto} API standards at a faster pace
 than traditional standards
@@
 
-%These three trends put evolutionary pressure on storage systems and raise the
-%question of whether 
 The evolutionary pressure placed on storage systems by these trends raises the
-question of whether
-there are principles that storage systems designers can follow
-to evolve storage systems efficiently, without jeopardizing years of
-code-hardening and performance optimization efforts.
-% that are important for
-%users to continue to entrust their data to the storage system.
-In this paper we investigate an approach that focuses on 
-%generalizing 
-%and
-%composing 
-identifying and exposing existing
+question of whether there are principles that storage systems designers can
+follow to evolve storage systems efficiently, without jeopardizing years of
+code-hardening and performance optimization efforts.  In this paper we
+investigate an approach that focuses on identifying and exposing existing
 storage system resources, services, and abstractions that in a generalized form
 can be used to \emph{program} new services. By composing higher-level services
-over existing abstraction one can reuse
-subsystems and leverage their optimizations, established correctness,
-robustness, and efficiency.
-
-{\it \textbf{We define a programmable storage system}} to be a storage system that facilitates the 
-re-use and extension of existing storage abstractions provided by the underlying software stack,
-to enable the creation of new services via composition.
-A programmable storage system can be realized by exposing existing functionality (such as 
-metadata services and synchronization and monitoring capabilities) as interfaces that can be ``glued together''
-in a variety of ways
-using a high-level language. In contrast to programmable storage, 
-\emph{active storage} differs through its focus on the injection and
+over existing abstraction one can reuse subsystems and leverage their
+optimizations, established correctness, robustness, and efficiency.
+
+{\it \textbf{We define a programmable storage system}} to be a storage system
+that facilitates the re-use and extension of existing storage abstractions
+provided by the underlying software stack, to enable the creation of new
+services via composition.  A programmable storage system can be realized by
+exposing existing functionality (such as metadata services and synchronization
+and monitoring capabilities) as interfaces that can be ``glued together'' in a
+variety of ways using a high-level language. In contrast to programmable
+storage, \emph{active storage} differs through its focus on the injection and
 execution of code within a storage system or storage device. Programmable
@@
 To illustrate the benefits and challenges of this approach we have designed and
-evaluated Malacology, a programmable storage system 
-that facilitates the construction of
-new
-services by re-purposing existing subsystem abstractions of the storage stack.
-We build Malacology in Ceph, a popular open source software storage stack.
-We choose Ceph to demonstrate the concept of programmable storage because it
-offers a broad spectrum of existing services, including distributed locking and
-caching services provided by metadata servers, durability and object interfaces
-provided by the backend object store, and propagation of consistent
-cluster state provided by the monitoring service (see
+evaluated Malacology, a programmable storage system that facilitates the
+construction of new services by re-purposing existing subsystem abstractions of
+the storage stack.  We build Malacology in Ceph, a popular open source software
+storage stack.  We choose Ceph to demonstrate the concept of programmable
+storage because it offers a broad spectrum of existing services, including
+distributed locking and caching services provided by metadata servers,
+durability and object interfaces provided by the backend object store, and
+propagation of consistent cluster state provided by the monitoring service (see
 Figure~\ref{fig:overview}). As we will show in this paper, Malacology is
@@
 
-{\it \textbf{We implement two distributed services}} using 
-%programmable storage
-%to demonstrate its feasability:
-Malacology to demonstrate the feasibility of the programmable storage approach:
+{\it \textbf{We implement two distributed services}} using Malacology to
+demonstrate the feasibility of the programmable storage approach:
 
@@
 framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
-implementation (\S\ref{evaluation}).  Finally we discuss related work and
+implementation (\S\ref{sec:evaluation}).  Finally we discuss related work and
 conclude.
@@
 the best performance. For example, HDFS was designed specifically to serve
-Hadoop's jobs, and uses techniques like exposing data locality and relaxing POSIX constraints to
-to achieve application-specific I/O optimizations.~\cite{CITEME}. Alternatively, general-purpose storage
-stacks are built with the flexiblity to serve many applications by providing a
-variety of interfaces and tunable parameters. Unfortunately, managing competing
-forces in these systems is difficult and users
-want more control from the general-purpose storage stacks
-without going as far as building their storage system from the ground up.
-
-% does this belong?
-%Ceph~\cite{weil:osdi2006-ceph,ceph}, the increasingly
-%popular, production-level open-source distributed storage system, was
-%originally built to be an HPC file system. But demands for storage systems that
-%could slide in behind cloud infrastructures, like OpenStack, shifted the focus
-%to Ceph's internal object store as a backend for large cloud block devices,
-%S3-like objects, and file system workloads.
+Hadoop's jobs, and uses techniques like exposing data locality and relaxing
+POSIX constraints to to achieve application-specific I/O
+optimizations.~\cite{CITEME}. Alternatively, general-purpose storage stacks are
+built with the flexiblity to serve many applications by providing a variety of
+interfaces and tunable parameters. Unfortunately, managing competing forces in
+these systems is difficult and users want more control from the general-purpose
+storage stacks without going as far as building their storage system from the
+ground up.
 
 To demonstrate this trend in storage systems we examine the state of
-programmability in Ceph. Something of a storage swiss army knife, Ceph simultaneously supports file,
-block, and object interfaces on a single
+programmability in Ceph. Something of a storage swiss army knife, Ceph
+simultaneously supports file, block, and object interfaces on a single
 cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
@@
 (OSDs) that provide Ceph with data durability and integrity using replication,
-erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already 
-%has some
-provides some degree of
-programmability; the OSDs support domain-specific object interfaces that are
-implemented by composing existing low-level storage interfaces that execute
-atomically. These interfaces are
-written in C++ and are statically loaded into the system.
-
-% By introducing programmability concepts into Ceph, we can build new services
-% by carefully exposing internal storage services to applications.
-
-The Ceph community provides empirical evidence that developers are already beginning to embrace
-programmable storage. Figure~\ref{fig:obj-int-dev-growth} shows a dramatic
-growth in the production use of domain-specific interfaces in the Ceph
-community since 2010. In that figure, classes are functional groupings of methods on storage
-objects (e.g. remotely computing and caching the checksum of an object extent).  What is most remarkable is that this trend contradicts the notion
-that API changes are a burden for users.  Rather it appears that a gap in
-existing interfaces are being addressed through ad-hoc approaches to programmability. In fact,
-Table~\ref{table:objclasses} categorizes existing interfaces and we clearly see
-a trend towards reusable services.
+erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already provides
+some degree of programmability; the OSDs support domain-specific object
+interfaces that are implemented by composing existing low-level storage
+interfaces that execute atomically. These interfaces are written in C++ and are
+statically loaded into the system.
+
+The Ceph community provides empirical evidence that developers are already
+beginning to embrace programmable storage. Figure~\ref{fig:obj-int-dev-growth}
+shows a dramatic growth in the production use of domain-specific interfaces in
+the Ceph community since 2010. In that figure, classes are functional groupings
+of methods on storage objects (e.g. remotely computing and caching the checksum
+of an object extent).  What is most remarkable is that this trend contradicts
+the notion that API changes are a burden for users.  Rather it appears that a
+gap in existing interfaces are being addressed through ad-hoc approaches to
+programmability. In fact, Table~\ref{table:objclasses} categorizes existing
+interfaces and we clearly see a trend towards reusable services.
 
@@
   \begin{tabular}{l|l|l}
-    Category & Specialization& \# \\ \hline
-    \multirow{2}{*}{Locking} & Shared & \multirow{2}{*}{6} \\
-                             & Exclusive & \\ \hdashline
-    \multirow{3}{*}{Logging} & Replica & 3 \\
-                             & State & 4 \\
-                             & Timestamped & 4 \\ \hdashline
-    \multirow{4}{*}{Metadata Managment} 
-                             & RADOS Block Device  & 37 \\
-                             & RADOS Gatway & 27 \\
-                             & User & 5 \\
-                             & Version & 5 \\ \hdashline
-    Garbage Collection       & Reference Counting & 4 \\
+    Category & Example & \# \\ \hline
+    Logging  & Geographically distribute replicas & 11 \\ \hdashline
+    Metadata & Snapshots in the block device OR  & \multirow{2}{*}{74} \\
+    Management & Scan extents for file system repair & \\ \hdashline
+    Locking  & Grants clients exclusive access & 6 \\ \hdashline
+    Other & Garbage collection, reference counting  & 4\\
 \end{tabular}
@@
 related to performance, availability, consistency, convenience, etc., are not
-satisfied by the existing default set of interfaces. The popularity of the custom object
-interface facility of Ceph could be due to a number of reasons, such as the
-default algorithms/tunables of the storage system being insufficient for the
-application's performance goals, programmers wanting to exploit
-application-specific semantics, and/or programmers knowing how to manage resources
-to improve performance. A solution based on application-specific object
-interfaces is a way to work around the
+satisfied by the existing default set of interfaces. The popularity of the
+custom object interface facility of Ceph could be due to a number of reasons,
+such as the default algorithms/tunables of the storage system being
+insufficient for the application's performance goals, programmers wanting to
+exploit application-specific semantics, and/or programmers knowing how to
+manage resources to improve performance. A solution based on
+application-specific object interfaces is a way to work around the
 traditionally rigid storage APIs because custom object interfaces give
@@
 
-%\section{Challenges}
-%\label{sec:challenges}
-%
-%The challenges of introducing programmability into existing storage systems:
-%(1) Programmability while keeping storage system available, (2) Finding and
-%factoring out hard-wired policies and optimizations, (3) Bugs that were hidden
-%because mechanisms were only exercised according to hard-wired policies and
-%not in their full generality they were thought of -- programmability into a
-%storage system turned out to be a great debugging tool, (4) version control
-%and propagation of updates throughout the system without interrupting
-%production workloads, (5) ...
-
 \section{Challenges}
@@
 
-\item Mechanisms that are often only exercised according to hard-wired policies and
-not in their full generality have hidden bugs that are revealed as soon as
+\item Mechanisms that are often only exercised according to hard-wired policies
+and not in their full generality have hidden bugs that are revealed as soon as
 those mechanisms are governed by different policies. In our experience
@@
 storage systems efficiently and without jeopardizing years of code-hardening
-and performance optimization efforts.  Although Malacology uses
-the internal abstractions of the underlying storage system, including its
-subsystems, components, and implementations, we emphasize that our system
-still addresses the general challenges outlined above.
+and performance optimization efforts.  Although Malacology uses the internal
+abstractions of the underlying storage system, including its subsystems,
+components, and implementations, we emphasize that our system still addresses
+the general challenges outlined above.
 
@@
 abstractions nor do we contend that the abstractions we choose provide the best
-trade-offs for all applications.
-For example, if consensus is correctly exposed one could implement
-high-level features like versioning, serialization, or various flavors
-of strongly consistent data management on top; but perhaps a low-level
-consensus interface is suited well for a particular set of applications.
-These questions are not answered in
-this paper and instead we focus on showing the feasibility of building such a
-system, given advances in the quality and robustness of today's storage stacks.
+trade-offs for all applications.  For example, if consensus is correctly
+exposed one could implement high-level features like versioning, serialization,
+or various flavors of strongly consistent data management on top; but perhaps a
+low-level consensus interface is suited well for a particular set of
+applications.  These questions are not answered in this paper and instead we
+focus on showing the feasibility of building such a system, given advances in
+the quality and robustness of today's storage stacks.
 
@@
 {\it Service Metadata}
-  & Zookeeper, Chubby coordination~\cite{hunt_zookeeper_2010,website:chubby}
+  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,website:chubby}
   & cluster state management~\cite{website:ceph-mon}
@@
   & MPI collective IO, burst
-  & POSIX metadata protocols~\cite{CITEME}
+  & POSIX metadata protocols
   & serialization \& batching
@@
   & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
-  & migration in metadata cluster~\cite{weil:sc2004-dyn-metadata}
+  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
   & migration \& sampling
@@
 
-%Thus, system
-%services realized through extensibility should be versioned together as a
-%group, exposing the version to allow introspection and domain-specific
-%treatment of version mismatch scenarios.
-
 In the case of Ceph, a consistent view of cluster state among server daemons
@@
 Ceph maintains cluster state information in per-subsystem data structures
-called ``maps'' that record
-membership and status information.  A Paxos-based monitoring service is
-responsible for integrating state changes into cluster maps, responding to
-requests from out-of-date clients and synchronizing members of the cluster
-whenever there is a change in a map so that they all observe the same system
-state. As a fundamental building block of many system designs, consensus
-abstractions such as Paxos are a common technique for maintaining consistent
-data versions, and are a useful system to expose.
+called ``maps'' that record membership and status information.  A Paxos-based
+monitoring service is responsible for integrating state changes into cluster
+maps, responding to requests from out-of-date clients and synchronizing members
+of the cluster whenever there is a change in a map so that they all observe the
+same system state. As a fundamental building block of many system designs,
+consensus abstractions such as Paxos are a common technique for maintaining
+consistent data versions, and are a useful system to expose.
 
@@
 requirements, such as creating arbitrary maps with application-specific
-semantics and interfaces. Since the monitor is intended to be out of high-performance I/O
-paths, a general guideline is to make use of this functionality infrequently
-and to assign small values to maps.
+semantics and interfaces. Since the monitor is intended to be out of
+high-performance I/O paths, a general guideline is to make use of this
+functionality infrequently and to assign small values to maps.
 
@@
 interfaces or other custom object interfaces, and handle serialization of
-function input and output.
-A wide range of native interfaces are available to developers such as reading
-and write to a byte stream, controlling object snapshots and clones,
-and accessing a sorted key-value database. These native object interfaces
-may be transactionally composed along with application specific logic to
-create interfaces with a rich set of semantics.
-An example is the use of a read-modify-write operation to perform consistent
-object updates across the byte stream and key-value database interfaces
-without relying on complex client-driven concurrency control.
+function input and output.  A wide range of native interfaces are available to
+developers such as reading and write to a byte stream, controlling object
+snapshots and clones, and accessing a sorted key-value database. These native
+object interfaces may be transactionally composed along with application
+specific logic to create interfaces with a rich set of semantics.  An example
+is the use of a read-modify-write operation to perform consistent object
+updates across the byte stream and key-value database interfaces without
+relying on complex client-driven concurrency control.
 
@@
 metadata cluster. This cluster uses load balancing policies to migrate
-directory inodes around the cluster to alleviate overload at single nodes. The policies use metrics
-based on system state (e.g.  CPU and memory utilization) and statistics
-collected by the cluster (e.g. the popularity of an inode). Ceph uses dynamic
-subtree partitioning to move variable sized namespace
-subtrees~\cite{weil:sc2004-dyn-metadata}. These units can
-be shipped anywhere (i.e., to any metadata server of any capacity) at any time
-for any reason. The original balancer was designed with hard-coded policies and
-tunables.
+directory inodes around the cluster to alleviate overload at single nodes. The
+policies use metrics based on system state (e.g.  CPU and memory utilization)
+and statistics collected by the cluster (e.g. the popularity of an inode). Ceph
+uses dynamic subtree partitioning to move variable sized namespace
+subtrees~\cite{weil:sc2004-dyn-metadata}. These units can be shipped anywhere
+(i.e., to any metadata server of any capacity) at any time for any reason. The
+original balancer was designed with hard-coded policies and tunables.
 
@@
 
-{\bf File types.} The metadata service manages a POSIX file system hierarchy
-where files and directories are represented as inode data structures and
-expose a POSIX file interface. Application that manage large amounts of
+{\bf File type interface.} The metadata service manages a POSIX file system
+hierarchy where files and directories are represented as inode data structures
+and expose a POSIX file interface. Application that manage large amounts of
 metadata (e.g. users or database snapshots) often require a naming service. We
@@
 
-{\bf Load balancing.} Mantle~\cite{sevilla:sc15-mantle}, a programmable
-metadata load balancer, has been implemented on Malacology so that its
-infrastructure can enjoy all the properties of existing internal abstractions.
-The load balancing logic is intact but the infrastructure has been improved to
-safely load, manage, and persist its load balancing policies.  Composing Mantle
-using the service metadata and durability interfaces, Mantle can safely
-version balancer policies, save balancer policies in the backend object store
-and centralize warnings/errors. When combined with file types, load balancing
-policies can express a variety of rules such as multi-tentant and
-multi-application balancing.
+{\bf Load balancing interface.} Mantle~\cite{sevilla:sc15-mantle}, a
+programmable metadata load balancer, has been implemented on Malacology so that
+its infrastructure can enjoy all the properties of existing internal
+abstractions.  The load balancing logic is intact but the infrastructure has
+been improved to safely load, manage, and persist its load balancing policies.
+Composing Mantle using the service metadata and durability interfaces, Mantle
+can safely version balancer policies, save balancer policies in the backend
+object store and centralize warnings/errors. When combined with file types,
+load balancing policies can express a variety of rules such as multi-tentant
+and multi-application balancing.
 
@@
 Ceph provides storage by striping and replicating data across
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-uses many techniques to ensure that data is not corrupted or lost, such as
-erasure coding, replication, and data scrubbing.
-Furthermore, many of these features are implemented using a peer-to-peer
-protocol that allows object storage devices to operate autonomously without
-a centralized coordinator.
-For example, when the number of placement groups change, the OSDs re-balance and
+RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS uses
+many techniques to ensure that data is not corrupted or lost, such as erasure
+coding, replication, and data scrubbing.  Furthermore, many of these features
+are implemented using a peer-to-peer protocol that allows object storage
+devices to operate autonomously without a centralized coordinator.  For
+example, when the number of placement groups change, the OSDs re-balance and
 re-shard data in the background in a process called placement group splitting
@@
 Mantle & load balance metadata     & Load Balancing  \\
-       & agree on current balancer & Versioning      \\ 
+       & agree on current balancer & Service Metadata\\
        & control client write lock & Shared Resource \\
        & persist policy logic      & Durability      \\ \hline
-ZLog   & implement CORFU           & Logical Device  \\ 
-       & install ZLog interfaces   & Versioning      \\
+ZLog   & implement CORFU           & Object \& Data  \\ 
+       & install ZLog interfaces   & Service Metadata\\
        & persist log entries       & Durability      \\ \hline
 ZLog   & load balance sequencers   & Load Balancing  \\
-Seqr.  & guard shared resource     & Shared Resource \\ 
+Seqr.  & manage access to log tail & Shared Resource \\ 
 \end{tabular}
@@
 it was trivial to make the version inconsistent. Maintaining a consistent
-version is important for balancing policies that are cooperative in which
-local decisions are made assuming properties about other instances of the same
+version is important for balancing policies that are cooperative in which local
+decisions are made assuming properties about other instances of the same
 policy.
@@
 With Malacology, Mantle stores the version of the current load balancer in the
-Ceph service metadata interface. The version of the load balancer
-corresponds to an object name in the balancing policy. Using the service
-metadata interface ensures that all metadata servers use the same version of
-the load balancer. As a result, the policy version inherits the
-consistency of Ceph's internal monitor daemons.
+Ceph service metadata interface. The version of the load balancer corresponds
+to an object name in the balancing policy. Using the service metadata interface
+ensures that all metadata servers use the same version of the load balancer. As
+a result, the policy version inherits the consistency of Ceph's internal
+monitor daemons.
 
@@
 The load balancer version described above corresponds to the name of an object
-in RADOS that holds the actual Lua balancing code.
-When MDS nodes start balancing load, they first check the latest version from
-the MDS map and compare it to the balancer they have loaded. If the version has
-changed, they dereference the pointer to the balancer version by reading the
-corresponding object in RADOS. This is in contrast to the original balancer
-which stored load balancer code on the local file system -- a technique which
-is unreliable and may result in silent corruption.
+in RADOS that holds the actual Lua balancing code.  When MDS nodes start
+balancing load, they first check the latest version from the MDS map and
+compare it to the balancer they have loaded. If the version has changed, they
+dereference the pointer to the balancer version by reading the corresponding
+object in RADOS. This is in contrast to the original balancer which stored load
+balancer code on the local file system -- a technique which is unreliable and
+may result in silent corruption.
 
@@
 With Malacology, Mantle re-usues the centralized logging features of the
-monitoring service. Important errors, warnings, and info messages are
-collected
-by the monitoring subsystem and appear in the monitor cluster log so instead of users
-going to each node, they can watch messages appear at the monitor. Messages are
-logged sparingly, so as not to spam the monitor with frivolous debugging but
-important events, like balancer version changes or failed subsystems show up in the
-centralized log.
+monitoring service. Important errors, warnings, and info messages are collected
+by the monitoring subsystem and appear in the monitor cluster log so instead of
+users going to each node, they can watch messages appear at the monitor.
+Messages are logged sparingly, so as not to spam the monitor with frivolous
+debugging but important events, like balancer version changes or failed
+subsystems show up in the centralized log.
 
@@
 strategy that we had not previously considered. In particular, for bursty
-clients, and clients that can tolerate increased latency, this mode of operation may
-allow a system to achieve much higher throughput than a system with a
-centralized sequencer service.
-We utilize the programmability of the metadata service to define a new policy
-for handling capabilities that controls the amount of time that clients are
-able to cache the sequencer resource. This allows an administrator or
-application to control the trade-off between latency and throughput beyond the
-standard best-effort policy that is present in Ceph by default.
+clients, and clients that can tolerate increased latency, this mode of
+operation may allow a system to achieve much higher throughput than a system
+with a centralized sequencer service.  We utilize the programmability of the
+metadata service to define a new policy for handling capabilities that controls
+the amount of time that clients are able to cache the sequencer resource. This
+allows an administrator or application to control the trade-off between latency
+and throughput beyond the standard best-effort policy that is present in Ceph
+by default.
 
@@
 In order to repopulate the sequencer state (i.e. the cached, current tail of
-the log) during recovery of a sequencer,
-the maximum position in the log must be obtained. To do this, the storage
-interface exposes an additional \emph{seal} method that atomically installs a
-new epoch value and returns the maximum log position that has been written.
+the log) during recovery of a sequencer, the maximum position in the log must
+be obtained. To do this, the storage interface exposes an additional
+\emph{seal} method that atomically installs a new epoch value and returns the
+maximum log position that has been written.
 
@@
 
-\section{Evaluation}
+\section{DON'T READ THIS!!! Evaluation}
 \label{sec:evaluation}
@@
 
-\textbf{The Malacology Approach:} Re-use the capability management features of the MDS.
-While the batching mechanism can achieve high-throughput, it is not appropriate
-for applications that require predictable latency. The throughput in
-Figure~\ref{fig:captp} labeled $2C,ZLog$ corresponds to a centralized
+\textbf{The Malacology Approach:} Re-use the capability management features of
+the MDS.  While the batching mechanism can achieve high-throughput, it is not
+appropriate for applications that require predictable latency. The throughput
+in Figure~\ref{fig:captp} labeled $2C,ZLog$ corresponds to a centralized
 sequencer implementation. While throughput is lower for two clients (because of
@@
 ``req\_rate" is a metric based on the exponenential decay of the request rate.
-``When policy" is used by Mantle to set a conditional for kicking off migration;
-in this case, the cluster migrates load if the current server, ``whoami", has
-load (\texttt{>0.1}) and the next server (physical servers have ``ranks") has a
-small amount load (\texttt{<0.1}). Finally, the ``where policy" is used by
-Mantle to determine how much load to shed. The policy fills in a ``targets"
-array to tell the migrator how much load to send to each server; in this case,
-we shed half of the current load to the next server. The unit of load is a
-single sequencer for a ZLog log.
+``When policy" is used by Mantle to set a conditional for kicking off
+migration; in this case, the cluster migrates load if the current server,
+``whoami", has load (\texttt{>0.1}) and the next server (physical servers have
+``ranks") has a small amount load (\texttt{<0.1}). Finally, the ``where policy"
+is used by Mantle to determine how much load to shed. The policy fills in a
+``targets" array to tell the migrator how much load to send to each server; in
+this case, we shed half of the current load to the next server. The unit of
+load is a single sequencer for a ZLog log.
 
@@
 cluster to alleviate load.  Each curve shows the throughput (\(y\) axis) over
-time (\(x\) axis).  ``Underloaded (Baseline)'' is the performance of one sequencer
-on one physical server serving 7 clients - there are no other sequencers or
-logs in the system and the curve represents the maximum throughput of a
-sequencer on an underutilized physical server.  ``Overloaded: No Balancing" adds a
-second sequencer of equal load to the same physical server with the intent of
-overloading the single physical server; recall that the throughput for this
-second sequencer is omitted. Compared to the baseline, the overloaded physical
-server with 2 sequencers has more than a 2\(\times\) throughput degradation.
-The stability of throughput curves (i.e. the variation on the \(y\) axis) indicate which sequencers
-are on overloaded physical servers.  ``Underloaded (Baseline)" is the best
-performance for this particular workload because there is no contention.
-Performance is stable and the MDS is under-loaded because we do not see the
-spike at the end like the other curves.  ``Overloaded: No Balancing" shows an
-overloaded sequencer because the performance of the monitored sequencer is low
-and unstable (high variation in performance).
+time (\(x\) axis).  ``Underloaded (Baseline)'' is the performance of one
+sequencer on one physical server serving 7 clients - there are no other
+sequencers or logs in the system and the curve represents the maximum
+throughput of a sequencer on an underutilized physical server.  ``Overloaded:
+No Balancing" adds a second sequencer of equal load to the same physical server
+with the intent of overloading the single physical server; recall that the
+throughput for this second sequencer is omitted. Compared to the baseline, the
+overloaded physical server with 2 sequencers has more than a 2\(\times\)
+throughput degradation.  The stability of throughput curves (i.e. the variation
+on the \(y\) axis) indicate which sequencers are on overloaded physical
+servers.  ``Underloaded (Baseline)" is the best performance for this particular
+workload because there is no contention.  Performance is stable and the MDS is
+under-loaded because we do not see the spike at the end like the other curves.
+``Overloaded: No Balancing" shows an overloaded sequencer because the
+performance of the monitored sequencer is low and unstable (high variation in
+performance).
 
 The last two curves of Figure~\ref{fig:mantle-seq-thruput} show the performance
-improvements when sequencers are load balanced across the cluster.
-The load balancing policies differ by the threshold at which they start balancing:
-``Overloaded: Sequencer Policy" starts balancing immediately and ``Overloaded: Conservative
-Policy" waits for 120 seconds before migrating sequencers.
+improvements when sequencers are load balanced across the cluster.  The load
+balancing policies differ by the threshold at which they start balancing:
+``Overloaded: Sequencer Policy" starts balancing immediately and ``Overloaded:
+Conservative Policy" waits for 120 seconds before migrating sequencers.
 
@@
 %
-%\begin{figure}[htbp]
-%\centering
-%\includegraphics{figures/caps-delay-thruput.png}
-%\caption{Forcing the client to drop their capabilities later (delay)
-%improves throughput}
-%\end{figure}
+%\begin{figure}[htbp] \centering
+%\includegraphics{figures/caps-delay-thruput.png} \caption{Forcing the client
+%to drop their capabilities later (delay) improves throughput} \end{figure}
 
@@
 system exposes internal abstractions so that end users (not necessarily
-operators) can create new services on top of the storage stack. Thus, our notion 
-of programmable storage differs from ``software-defined storage'' (SDS) in terms 
-of goals and scope, although definitions of SDS are still in flux.
+operators) can create new services on top of the storage stack. Thus, our
+notion of programmable storage differs from ``software-defined storage'' (SDS)
+in terms of goals and scope, although definitions of SDS are still in flux.
 
@@
 Malacology uses the same Active and Typed Storage module presented in
-DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File
-Manifolds can be implemented with small changes to the Malacology
-framework, namely asynchronous object calls and Lua stubs in the inode,
-respectively.
+DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File Manifolds
+can be implemented with small changes to the Malacology framework, namely
+asynchronous object calls and Lua stubs in the inode, respectively.
 

### 2016-10-19 experiments/zlog-seqr-redux/run.sh
updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -7,11 +7,12 @@ CEPH_BIN=/home/nwatkins/src/ceph/build/bin
 export CEPH_CONF=/home/nwatkins/src/ceph/build/ceph.conf
 
 prefix=$1
-runtime=60
-capdelay="1.0"
+runtime=120
 quota=0
-nclients=4
-for capdelay in 1.0 0.5 0.25 0.125 .0625 0.03125; do
+nclients=2
+#capdelay=1.0
+
+for capdelay in 1.0 0.5 0.25 0.125 0.0625 0.03125 0.015625 0.0078125 0.00390625 0.001953125 0.0009765625 0.00048828125; do
   perf_file="${prefix}.nc-${nclients}.cd-${capdelay}.qa-${quota}"
   $CEPH_BIN/seq-client --runtime=$runtime --perf_file=${perf_file} \
     --instances=$nclients --capdelay=$capdelay --quota=${quota}

### 2016-10-19 experiments/zlog-seqr-redux/run.sh
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -9,7 +9,7 @@ export CEPH_CONF=/home/nwatkins/src/ceph/build/ceph.conf
 prefix=$1
 runtime=120
 quota=0
-nclients=2
+nclients=4
 #capdelay=1.0
 
 for capdelay in 1.0 0.5 0.25 0.125 0.0625 0.03125 0.015625 0.0078125 0.00390625 0.001953125 0.0009765625 0.00048828125; do

### 2016-10-19 experiments/zlog-seqr-redux/run.sh
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -9,7 +9,7 @@ export CEPH_CONF=/home/nwatkins/src/ceph/build/ceph.conf
 prefix=$1
 runtime=120
 quota=0
-nclients=4
+nclients=8
 #capdelay=1.0
 
 for capdelay in 1.0 0.5 0.25 0.125 0.0625 0.03125 0.015625 0.0078125 0.00390625 0.001953125 0.0009765625 0.00048828125; do

### 2016-10-19 experiments/zlog-seqr-redux/run.sh
updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -6,14 +6,17 @@ set -x
 CEPH_BIN=/home/nwatkins/src/ceph/build/bin
 export CEPH_CONF=/home/nwatkins/src/ceph/build/ceph.conf
 
-prefix=$1
 runtime=120
 quota=0
 nclients=8
+#prefix=$1
 #capdelay=1.0
 
-for capdelay in 1.0 0.5 0.25 0.125 0.0625 0.03125 0.015625 0.0078125 0.00390625 0.001953125 0.0009765625 0.00048828125; do
-  perf_file="${prefix}.nc-${nclients}.cd-${capdelay}.qa-${quota}"
-  $CEPH_BIN/seq-client --runtime=$runtime --perf_file=${perf_file} \
-    --instances=$nclients --capdelay=$capdelay --quota=${quota}
+for nclients in 2 4 8; do
+  prefix="${nclients}c_shrink_capdelay"
+  for capdelay in 1.0 0.5 0.25 0.125 0.0625 0.03125 0.015625 0.0078125 0.00390625 0.001953125 0.0009765625 0.00048828125; do
+    perf_file="${prefix}.nc-${nclients}.cd-${capdelay}.qa-${quota}"
+    $CEPH_BIN/seq-client --runtime=$runtime --perf_file=${perf_file} \
+      --instances=$nclients --capdelay=$capdelay --quota=${quota}
+  done
 done

### 2016-10-19 experiments/zlog-seqr-redux/run.sh
results
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -7,16 +7,21 @@ CEPH_BIN=/home/nwatkins/src/ceph/build/bin
 export CEPH_CONF=/home/nwatkins/src/ceph/build/ceph.conf
 
 runtime=120
-quota=0
-nclients=8
-#prefix=$1
-#capdelay=1.0
 
-for nclients in 2 4 8; do
-  prefix="${nclients}c_shrink_capdelay"
-  for capdelay in 1.0 0.5 0.25 0.125 0.0625 0.03125 0.015625 0.0078125 0.00390625 0.001953125 0.0009765625 0.00048828125; do
-    perf_file="${prefix}.nc-${nclients}.cd-${capdelay}.qa-${quota}"
-    $CEPH_BIN/seq-client --runtime=$runtime --perf_file=${perf_file} \
-      --instances=$nclients --capdelay=$capdelay --quota=${quota}
-  done
+capdelay="0.125"
+nclients=2
+for quota in 1 10 100 1000 10000 100000 1000000; do
+  prefix="2c_0.125s_increase_quota"
+  perf_file="${prefix}.nc-${nclients}.cd-${capdelay}.qa-${quota}"
+  $CEPH_BIN/seq-client --runtime=$runtime --perf_file=${perf_file} \
+    --instances=$nclients --capdelay=$capdelay --quota=${quota}
 done
+
+#for nclients in 2 4 8; do
+#  prefix="${nclients}c_shrink_capdelay"
+#  for capdelay in 1.0 0.5 0.25 0.125 0.0625 0.03125 0.015625 0.0078125 0.00390625 0.001953125 0.0009765625 0.00048828125; do
+#    perf_file="${prefix}.nc-${nclients}.cd-${capdelay}.qa-${quota}"
+#    $CEPH_BIN/seq-client --runtime=$runtime --perf_file=${perf_file} \
+#      --instances=$nclients --capdelay=$capdelay --quota=${quota}
+#  done
+#done

### 2016-10-20 paper/paper.tex
tweaks to abstract

@@
 data to the storage system.  We introduce  Malacology, a prototype programmable
-storage system to explore how existing abstractions of common services found in
-storage systems can be leveraged to address new data processing systems and the
+storage system, to explore how existing abstractions of common storage system services 
+%found in
+%storage systems 
+can be leveraged to 
+%address 
+adapt to the needs of
+new data processing systems and the
 increasing speed of storage devices. This approach allows a large degree of
@@
 storage system, including the ability to balance load, efficiently propagate
-cluster metadata, recover from failure, and to make trade-offs for
-latency/throughput using leases.
+cluster metadata, recover from failure, and to navigate trade-offs between
+latency and throughput using leases.
 \end{abstract}

### 2016-10-20 paper/paper.tex
intro tweaks

@@
 data'' applications demand interface evolution to support flexible consistency
-as well as flexible structured data representations.
-~\cite{apache_contributors_parquet_2014}.  Finally, production-quality scalable
+as well as flexible structured data 
+representations.~\cite{apache_contributors_parquet_2014}.  Finally, production-quality scalable
 storage systems available as open source software have established and are
@@
 execution of code within a storage system or storage device. Programmable
-storage differs from \emph{active storage}~\cite{riedel:vldb98} in that the
+storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection and
+execution of code within a storage system or storage device---in
+that the
 former is applicable to any component of the storage system, while the latter
@@
 expressive enough to provide the functionality necessary for implementing new
-services. Our contributions are:
-
-A prototype programmable storage system that includes a non-exhaustive set of
+services. 
+%Our contributions are:
+
+% prototype programmable storage system that 
+Malacology
+includes a non-exhaustive set of
 interfaces that can be used as building blocks for constructing novel storage
@@
 \item A high-performance distributed shared log service called ZLog, that is an
-implementation of CORFU~\cite{balakrishnan_corfu_2012}
+implementation of CORFU.~\cite{balakrishnan_corfu_2012}
 
@@
 framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
-implementation (\S\ref{sec:evaluation}).  Finally we discuss related work and
-conclude.
+implementation (\S\ref{sec:evaluation}).  We conclude by discussing related and future work.
+% and
+%conclude.
 

### 2016-10-20 paper/paper.tex
tweak

@@
 that the source code of the storage system and the necessary expertise for
-understanding it is available:
+understanding it is available.  Some challenges include:
 

### 2016-10-20 experiments/zlog-seqr-redux/run.sh
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -8,13 +8,13 @@ export CEPH_CONF=/home/nwatkins/src/ceph/build/ceph.conf
 
 runtime=120
 
-capdelay="0.125"
+capdelay="0.25"
 nclients=2
 for quota in 1 10 100 1000 10000 100000 1000000; do
-  prefix="2c_0.125s_increase_quota"
+  prefix="2c_0.25s_increase_quota2"
   perf_file="${prefix}.nc-${nclients}.cd-${capdelay}.qa-${quota}"
   $CEPH_BIN/seq-client --runtime=$runtime --perf_file=${perf_file} \
-    --instances=$nclients --capdelay=$capdelay --quota=${quota}
+    --instances=$nclients --capdelay=$capdelay --quota=${quota} --latcdf
 done
 
 #for nclients in 2 4 8; do

### 2016-10-20 paper/paper.tex
paper: mantle eval outline, much tired, so schleepy (#101)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\section{DON'T READ THIS!!! Evaluation}
+\section{Evaluation}
 \label{sec:evaluation}
@@
 Malacology our evaluation focuses on the performance of the internal
-abstractions when used for to construct the Mantle and ZLog services. First, we
+abstractions when used to construct the Mantle and ZLog services. First, we
 benchmark scenarios where contention is at the sequencer by examining the
-intefaces used to map ZLog onto Malacology; specifically, we describe the
-sequencer implementation and the propogation of object \& data interfaces
+interfaces used to map ZLog onto Malacology; specifically, we describe the
+sequencer implementation and the propagation of object \& data interfaces
 interfaces. Next, we benchmark scenarios when there are multiple logs by using
@@
 both Mantle and ZLog, as we try to find the best metadata balancing policies
-and physical design implementation trade-offs respectively.  Since this work
+and physical design implementation trade-offs.  Since this work
 focuses on the programmabilty of Malacology, the goal of the benchmarks is to
@@
 \subsubsection{Sequencer Implementation}
-
-We evaluated the performance of the ZLog Sequencer implemented in several
+\label{sec:sequencer-implementation}
+
+We evaluate the performance of the ZLog Sequencer implemented in several
 different configurations by showing a trade-off between aggregate throughput
@@
 
-\textbf{The Malacology Approach:} Re-use the capability management features of
-the MDS.  While the batching mechanism can achieve high-throughput, it is not
-appropriate for applications that require predictable latency. The throughput
-in Figure~\ref{fig:captp} labeled $2C,ZLog$ corresponds to a centralized
-sequencer implementation. While throughput is lower for two clients (because of
-the synchronous network round-trip required for each request),
+\noindent\textbf{The Malacology Approach:} Re-use the capability management
+features of the MDS.  While the batching mechanism can achieve high-throughput,
+it is not appropriate for applications that require predictable latency. The
+throughput in Figure~\ref{fig:captp} labeled $2C,ZLog$ corresponds to a
+centralized sequencer implementation. While throughput is lower for two clients
+(because of the synchronous network round-trip required for each request),
 Figure~\ref{fig:capcdf} shows that this approach achieves more predicable
 latency compared to all other approaches except for the single client locally
-cached mode.
+cached mode.\\
+
+\noindent\textbf{Features:} delay and quota parameters control throughput vs.
+latency
 
@@
 
-\subsection{ZLog and Mantle in a Multi-tenant Environment}
+\subsection{Load Balancing Multiple ZLog Sequencers with Mantle}
 
@@
 resources, monitoring the workloads, monitoring physical nodes, partitioning
-resources, and managing multiple sequencers.
+resources, cache roherence, and managing multiple sequencers.
 
@@
 of the Malacology load balancing interface (e.g., migrating, monitoring, and
-partitioning load) to (1) alleviate load when multiple sequencers overwhelm a
-physical servers and (2) recover from failure.
+partitioning load) to (1) alleviate load from overloaded servers with multiple
+sequencers and (2) recover from failure.
 
@@
 \centering
-\includegraphics{figures/mantle-seq-thruput.png}
+\includegraphics{figures/mantle-balancer-behaviors.png}
 \caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
-load balancing interface alleviates load for multiple sequencers in
-resource-constrained environments.  ``Underloaded (Baseline)'' is the
-throughput of one sequencer running alone on a cluster; the other curves show
-how load balancing affects the throughput of one sequencer when sequencers for
-other logs are initially co-located on the same physical server (i.e. an
-"Overloaded" environment).}\label{fig:mantle-seq-thruput}
+load balancing interface alleviates overloaded servers in environments with multiple sequencers in
+}\label{fig:mantle-balancer-behaviors}
 \end{figure}
-
-\subsubsection{Load Balancing Sequencers}
-
-We evaluate the load balancing interface by showing how the cluster of servers
-housing sequencers alleviates load.  The experiments are run on a cluster with
-10 nodes to store objects, one node to monitor the cluster, and 3 nodes that
-can accomodate sequencers. The workload has up to 6 logs each with 2 clients.
-Figure~\ref{CITEME} shows the throughput per sequencer when all seqeuencers are
-co-located of one physical server. Two sequencers on the same server show no
-performance degradation but co-locating 3, 4, 5, or 6 sequencers leads to a
-2x, 3x, 4x, and 5x slowdown, respectively. For the rest of the experiments we
-load balance 3, 4, 5, or 6 sequencers across the cluxter because this is hte
-point at which a single physical server gets overloaded.
-
-Figure~\ref{CITEME} shows the performance profiles of the sequencers as they
-migrate around the cluster. Each graph in that figure shows different ``waves"
-of sequencers, where a ``wave" is a burst of clients trying to append to the
-end of the log. These waves are examples of hotspots and flash crowds; they are
-dynamic and resource constrained because clients access a shared resource. Wave
-1, policy 1 shows how the sequencers initially get distributed poorly. From x
-to x seconds, all seqeuencers are co-located. At x seconds, 3 of the sequencers
-(1, 2, 3) migrate off the first server and onto the second one; this is what we
-want.
 
@@
 \centering
-\includegraphics{figures/mantle-balancer-profiles.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
-load balancing interface alleviates load for multiple sequencers in
-resource-constrained environments.  ``Underloaded (Baseline)'' is the
-throughput of one sequencer running alone on a cluster; the other curves show
-how load balancing affects the throughput of one sequencer when sequencers for
-other logs are initially co-located on the same physical server (i.e. an
-"Overloaded" environment).}\label{fig:mantle-seq-thruput}
+\includegraphics{figures/mantle-balancer-performance.png}
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
+}\label{fig:mantle-balancer-performance}
 \end{figure}
+
+The experiments are run on a cluster with 10 nodes to store objects, one node
+to monitor the cluster, and 3 nodes that can accomodate sequencers. Our first
+experiment co-locates 3 sequencers each with 4 clients on the same physical
+server.  Instead of measuring contention at the clients like we did in
+Section~\ref{sec:sequencer-implementation}, these experiments create contention
+at the sequencers by setting the capability infrastructure to ``round trip
+mode". In this mode clients go to the sequencer for every request. We
+implement this by piggybacking on \texttt{fstat()} and using the shared
+resource interface to force round trips.  Because the sequencer hands out
+positions for the tail of the log, the workload is read-heavy.
+
+Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput (\(y\) axis)
+over time (\(x\) axis) of different load balancers that migrate ZLog sequencers
+acround the cluster; ``No Balancing" keeps all sequencers on one server,
+``CephFS" migrates sequencers using the original algorithms packaged with Ceph,
+and the ``Mantle" curve shows the throughput of a custom load balancer we wrote
+specifically for sequencers. The increased throughput for the ``CephFS" and
+``Mantle" curves between 0 and 60 seconds are a result of migrating the
+sequencer(s) off of overloadeed servers. In addition to showing that migrating
+sequencers improves performance, Figure~\ref{fig:mantle-balancer-behaviors}
+also demonstrates features that we will explore in the rest of this section.
+Section~\ref{sec:feature-balancing-modes} and~\ref{sec:feature-migration-units}
+shows why the performance of Mantle is better than CephFS when the cluster
+stabilizes at time 100 seconds and Section~\ref{sec:feature-backoff} explains
+the slope and location of the rebalancing phase between 0 and 60 seconds by
+comparing the aggressiveness of the CephFS and Mantle balancers.
+
+\subsubsection{Feature: Balancing Modes}
+\label{sec:feature-balancing-modes}
+
+The load balancing service can operate in different balancing modes that stress
+the subsystems that handle client requests or forwarded requests in different
+ways. In Figure~\ref{fig:mantle-balancer-behaviors}, the ``CephFS" curve shows
+the performance of the balancing mode that CephFS uses {\it most of the time}.
+CephFS has 3 modes for balancing load: CPU mode, workload mode, and hybrid
+mode. All three have the same structure for making migration decisions but vary
+based on the metric used to calculate load. For this sequencer workload, the 3
+different modes all have the same performance, as shown in
+Figure~\ref{fig:mantle-balancer-performance}, because the load balancer falls
+into the same mode a majority of the time.  The high variation in performance
+for the ``CephFS CPU Mode" bar reflects the uncertainty of using something as
+dynamic and unpredictable as CPU utilization to make migration decisions. In
+addition to the suboptimal performance and unpredictability, the fact that the
+CephFS balancers all fall into the same mode despite using different balancers
+is a problem as it prevents administrators from properly exploring the
+balancing state space.
+
+Mantle gives the administrator more control over the balancing policies; for
+the ``Mantle" bar in Figure~\ref{fig:mantle-balancer-performance} we use the
+load balancing interface to program logic for balancing read-heavy workload
+into the load balancing service. The resulting performance is more stable and
+performs better because we created our own mode for balancing sequencers.
+
+We have identified two balancing modes that are relevant for making migration
+decisions for sequencers. Using Mantle, the administrator can put the load
+balancing service into ``proxy mode" or ``sequencer mode". In ``proxy mode" one
+server receives all requests and farms off the requests to slave servers.  In
+``sequencer mode" clients go directly to the server that has their sequencer.
+These modes are illustrated in Figure~\ref{fig:mantle-modes}. ``None" is when
+all sequencers are co-located on one physical server -- performance for that
+mode is shown by the ``No Balancing" curve in
+Figure~\ref{fig:mantle-balancer-behaviors}. ``Sequencer" mode in the lower left
+shows how clients for sequencer 2 are re-directed to server B. The two ``Proxy"
+mode diagrams on the right show how clients continue sending reqeusts to server
+A even though sequencer 2 has migrated to server B (the difference between
+``Proxy (Half)" and ``Proxy (Full)" is discussed in the next section).
+
+\begin{figure}[t!]
+\centering
+\includegraphics{figures/mantle-mode-behavior.png}
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
+}\label{fig:mantle-mode-behavior}
+\end{figure}
+
+Figure~\ref{fig:mantle-mode-performance} shows the throughput (\(y\) axis) over
+time (\(x\) axis) of the two different modes for an environment with only 2
+sequencers (again 4 clients each) and 2 servers. The curves for both sequencers
+in Figure~\ref{fig:mantle-mode-performance}(a) start at less than 1000
+ops/second and at time 60 seconds Mantle migrates Sequencer 1 to the slave
+server. Performance of Sequencer 2 decreases because it has stayed on the proxy
+which now handles all client requests, processes requests for Sequencer 2, and
+forwards requests for Sequencer 1 to the slave server. Sequencer 1 performance
+improves dramatically because lookups for that sequencer are parallelized with
+the operations on server 1. Cluster throughput improves at the cost of
+decreased throughput and unfair throughput shares for Sequencer 2.
+Figure~\ref{fig:mantle-mode-performance}(b) is set to sequencer mode manually
+(no balancing phase) and shows that performance is worse than proxy mode {\it
+and} unfair amongst sequencers. In this case, the scatter-gather process used
+for cache coherence in the metadata protocols causes strain on the server
+housing Sequencer 2 resulting in uneven performance.
+
+Intuitively, sequencer mode is a superior balancing mode and should perform
+better; after all, this is the best strategy for distributing POSIX metadata in
+write-heavy workloads~\cite{ren:sc2014-indexfs,
+patil:fast2011-giga,sevilla:sc15-mantle}. But our experiments show that proxy
+mode is superior for the sequencer's read-heavy workload.  In this case,
+separating the handling of the client request and doing the actual operation
+improves performance. Processing client requests and doing the actual operation
+is too heavy weight for one server and sequencers on slave nodes can go faster
+because work is split up; the proxy processes the client request and the slave
+finds the tail of the log and delivers the answer.  This phenomenon is not
+uncommon and has been observed in chain replication~\cite{CITEME}.
+
+\subsubsection{Feature: Migration Units}
+\label{sec:feature-migration-units}
+
+How much load to try and migrate is not a tunable parameter in the original
+CephFS balancers. The most useful way to express how much load to send is as a
+fraction of some other metric such as total load. Expressing this heuristic is
+not easily achievable with outward facing tunable parameters (i.e. system
+knobs) but can be expressed with programmability quite easily. To force the
+load balancer into ``Proxy (Half)" scenario in Figure~\ref{fig:mantle-modes},
+which sets the balancer to proxy mode with migration units of half the load on
+the current server we can use the Mantle policy:\\
+
+\noindent \texttt{ targets[\#mds+1] = mds[whoami]["load"]/2 }\\
+
+This code snippet uses globally defined variables and tables from the Mantle
+API to send half of the load on the current server (whoami) to the next ranked
+server (\#mds + 1); the \texttt{targets} array is a globally defined table that
+the balancer uses to do the migrations. Alternatively, to set the mode to the
+``Proxy (Full)"` setup in Figure~\ref{fig:mantle-modes}, one would use:\\
+
+\noindent \texttt{ targets[\#mds+1] = mds[whoami]["load"] }\\
 
@@
 other logs are initially co-located on the same physical server (i.e. an
-"Overloaded" environment).}\label{fig:mantle-seq-thruput}
+"Overloaded" environment).}\label{fig:mantle-modes}
 \end{figure}
 
-We use a balancer policy specifically designed for multiple sequencers. The
-balancer aggressively sheds half its load to the next server as soon as it can.
-It is based off the ``Greedy Spill Even" approach in the
-Mantle~\cite{sevilla:sc15-mantle} paper but is reproduced in
-Figure~\ref{listing:greedy-spill} for the reader's convenience. Each line of
-code uses variables exposed by the Mantle API and is executed as a callback by
-Mantle.  ``Server load" lets the administrator control what metric constitutes
-load on a server; ``i" is a variable used by Mantle to iterate over servers and
-``req\_rate" is a metric based on the exponenential decay of the request rate.
-``When policy" is used by Mantle to set a conditional for kicking off
-migration; in this case, the cluster migrates load if the current server,
-``whoami", has load (\texttt{>0.1}) and the next server (physical servers have
-``ranks") has a small amount load (\texttt{<0.1}). Finally, the ``where policy"
-is used by Mantle to determine how much load to shed. The policy fills in a
-``targets" array to tell the migrator how much load to send to each server; in
-this case, we shed half of the current load to the next server. The unit of
-load is a single sequencer for a ZLog log.
+\begin{figure}[t!]
+\centering
+\includegraphics{figures/mantle-mode-performance.png}
+\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
+}\label{fig:mantle-mode-performance}
+\end{figure}
+
+Figure~\ref{fig:mantle-mode-performance} shows the performance of the modes
+using different migration units. Recall that this setup only has 2 sequencers
+and 2 servers, so performance may be different at scale. Even so, it is clear
+sequencer mode does not perform as well for read-heavy workloads and we even
+see a throughput improvement when migrating all load off the first server,
+leaving the first server to do administrative tasks (this is actually common in
+the metadata cluster because the ``first" server does a lot of the cache
+coherence work) while the second server does all the client work. Proxy Mode
+does the best in both cases and shows large performance gains when completely
+decoupling client request handling and operation processing in Proxy Mode
+(Full).
+
+The parameter that controls the migration units helps the administrator control
+the sequencer or co-location or distribution across the cluster. This trade-off
+was explored extensively in the Mantle paper and warrants further exploration
+as all our experiments indicate a rich set of states to explore.
+
+\subsubsection{Feature: Backoff Paramters}
+\label{sec:feature-backoff}
+
+The balancing phase from 0 to 60 seconds in
+Figure~\ref{mantle-balancer-behaviors} shows different degrees of
+aggressiveness in making migration decisions; CephFS makes a decision quickly
+and throughput jumps 2.5K ops/second within 10 seconds while Mantle takes a
+more time to stabilize. This conservative behavior is controlled by programming
+the balancer to (1) use different conditions for when to migrate and (2) use a
+threshold for sustained overload. \texttt{when()} is part the Mantle API and
+the callback expects to return a true or false, which corresponds to whether to
+migrate or not. For the Mantle curve in Figure~\ref{mantle-balancer-behaviors}
+we set the \texttt{when()} policy to wait for load on the receiving server to
+reach 0.1. At the beginning of the run the load of the receiving server spikes
+because requests for cache-coherence start as the sequencers are created. It
+takes 60 seconds for this load to drop to an acceptable level. Also, the
+migration phase is a little more conservative in sending load; the Mantle curve
+in Figure~\ref{mantle-balancer-behaviors} takes longer to reach peak throughput
+because we do the first migration and {\it wait to see how it affects the
+system}; seeing that after migration the third server is still underloaded
+allows the balancer to further improve performance by migrating a sequencer to
+the laster server.
+
+The other way to change aggressiveness of the decision making is to program
+into the balancer a policy that waits a certain number of iterations after a
+migration before proceeding. In Mantle, the policy would use the save state
+function to do a countdown after a migation:
 
@@
 \begin{lstlisting}
-  -- Server load
-  load = server[i]["req_rate"]
-
-  -- When policy
-  if server[whoami]["load"]>.01 and 
-    server[whoami+1]["load"]<.01 then
-
-  -- Where policy
-  targets[whoami+1]=load/2
+  countdown = GET_STATE()
+  if countdown > 0 then
+    -- OK to migrate
+  else
+    PUT_STATE(countdown - 1)
+  end
 \end{lstlisting}
@@
 
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-cpu-utilization.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
-CPU utilization for a sequencer in an environment with the ``Conservative
-Policy" load balancer shows that the load balancing service is not
-multi-threaded. The load balancing service is designed this way because the
-performance for a guarded resource will not improve with more cores.}
-\label{fig:mantle-cpu-utilization}
-\end{figure}
-
-Figure~\ref{fig:mantle-seq-thruput} shows a performance profile of one of the
-sequencers as the load balancing interface migrates sequencers around the
-cluster to alleviate load.  Each curve shows the throughput (\(y\) axis) over
-time (\(x\) axis).  ``Underloaded (Baseline)'' is the performance of one
-sequencer on one physical server serving 7 clients - there are no other
-sequencers or logs in the system and the curve represents the maximum
-throughput of a sequencer on an underutilized physical server.  ``Overloaded:
-No Balancing" adds a second sequencer of equal load to the same physical server
-with the intent of overloading the single physical server; recall that the
-throughput for this second sequencer is omitted. Compared to the baseline, the
-overloaded physical server with 2 sequencers has more than a 2\(\times\)
-throughput degradation.  The stability of throughput curves (i.e. the variation
-on the \(y\) axis) indicate which sequencers are on overloaded physical
-servers.  ``Underloaded (Baseline)" is the best performance for this particular
-workload because there is no contention.  Performance is stable and the MDS is
-under-loaded because we do not see the spike at the end like the other curves.
-``Overloaded: No Balancing" shows an overloaded sequencer because the
-performance of the monitored sequencer is low and unstable (high variation in
-performance).
-
-The last two curves of Figure~\ref{fig:mantle-seq-thruput} show the performance
-improvements when sequencers are load balanced across the cluster.  The load
-balancing policies differ by the threshold at which they start balancing:
-``Overloaded: Sequencer Policy" starts balancing immediately and ``Overloaded:
-Conservative Policy" waits for 120 seconds before migrating sequencers.
-
-%Overall, the clients achieve lower performance than the capability experiments
-%#in Figure~\ref{fig:captp} because the sequencer uses a slower metadata
-%operation to manage the capability.  It appears the bottleneck is somewhere
-%with the clients -- perhaps it is our 1GBit/s network link.
-
-Figure~\ref{fig:mantle-cpu-utilization} shows how load balancing effects the
-physical servers by plotting the CPU utilization average over 1 minute (\(y\)
-axis) over time (\(x\) axis).  It shows the ``Conservative Policy" load
-balancer from the run in Figure~\ref{fig:mantle-seq-thruput}, so recall that
-resources are constrained during the job by co-locating a second sequencer on
-the same physical server.  CPU utilization spikes when physical servers receive
-load (at 2 minutes physical server 1 receives a sequencer from physical server
-0) but the CPU utilization never eclipses 50\%. The utilization is low because
-the load balancer interface is not multi-threaded.  Recall the the load
-balancing interface re-uses the POSIX metadata cluster migration mechanisms;
-serving file system metadata is single-threaded because it has been shown that
-metadata performance does not improve with multiple
-threads~\cite{konstantinos:pdsw2014-lustre-metadata}.
-
-\textbf{The Malacology Approach:} Re-use the Ceph metadata migration, load
-balancing, cache coherence, and capabilities subsystems of the metadata
-cluster. Applying Mantle's load balancing policy shows 2.02\(\times\)
-performance improvement and only a 16\% degradation in performance over the
-baseline, even though the baseline has half as many clients.
-
-\subsubsection{Recovery}
-
-\begin{figure}[t]
-\centering
-\includegraphics{figures/recovery-seq-thruput.png}
-\caption{The sequencer using Mantle also inherits all the recovery protocols of
-the Ceph metadata cluster. The throughput of the sequencer drops when we
-terminal MDS 0. The performance quickly resumes and the sequencer is only down
-for 15 seconds.}
-\label{fig:recovery-seq-thruput}
-\end{figure}
-
-This experiment shows how a sequencer handing out tokens can use the Ceph MDS
-recovery protocols to recover. For this workload, two sequencers are running on
-a metadata cluster {\it without} multiple active MDS nodes. In this cluster,
-MDS 1 and MDS 2 are in ``standby" mode, meaning one of them will step in should
-the head MDS 0 fail. At the 60 second mark, we manually kill MDS 0.
-
-Figure~\ref{fig:recovery-seq-thruput} shows the sequencer throughput (\(y\)
-axis) over time (\(x\) axis). When MDS 0 dies, the MON cluster notices and
-quickly assigns a new MDS to handle the namespace. This is when MDS 1 takes
-over the sequencer duties and starts handing out tokens.
-
-\textbf{The Malacology Approach:} Re-use the Ceph metadata protocols for
-failure detection and recovery. The sequencer recovers in 15 seconds on a new
-MDS and it maintains the same throughput that it had before the failure.
+Behavior graphs and performance numbers for this feature is omitted, as the
+throughput graphs are as expected: the more conservative the approach the less
+overall throughput.
+
+%We use a balancer policy specifically designed for multiple sequencers. The
+%balancer aggressively sheds half its load to the next server as soon as it can.
+%It is based off the ``Greedy Spill Even" approach in the
+%Mantle~\cite{sevilla:sc15-mantle} paper but is reproduced in
+%Figure~\ref{listing:greedy-spill} for the reader's convenience. Each line of
+%code uses variables exposed by the Mantle API and is executed as a callback by
+%Mantle.  ``Server load" lets the administrator control what metric constitutes
+%load on a server; ``i" is a variable used by Mantle to iterate over servers and
+%``req\_rate" is a metric based on the exponenential decay of the request rate.
+%``When policy" is used by Mantle to set a conditional for kicking off
+%migration; in this case, the cluster migrates load if the current server,
+%``whoami", has load (\texttt{>0.1}) and the next server (physical servers have
+%``ranks") has a small amount load (\texttt{<0.1}). Finally, the ``where policy"
+%is used by Mantle to determine how much load to shed. The policy fills in a
+%``targets" array to tell the migrator how much load to send to each server; in
+%this case, we shed half of the current load to the next server. The unit of
+%load is a single sequencer for a ZLog log.
+%
+%\begin{figure}[]
+%\begin{lstlisting}
+%  -- Server load
+%  load = server[i]["req_rate"]
+%  
+%  -- When policy
+%  if server[whoami]["load"]>.01 and 
+%    server[whoami+1]["load"]<.01 then
+%    
+%  -- Where policy
+%  targets[whoami+1]=load/2
+%\end{lstlisting}
+%\caption{The sequencer balancing policy that works best for bursty workloads. 
+%\label{listing:greedy-spill}}
+%\end{figure}
+%
+%Figure~\ref{fig:mantle-seq-thruput} shows a performance profile of one of the
+%sequencers as the load balancing interface migrates sequencers around the
+%cluster to alleviate load.  Each curve shows the throughput (\(y\) axis) over
+%time (\(x\) axis).  ``Underloaded (Baseline)'' is the performance of one
+%sequencer on one physical server serving 7 clients - there are no other
+%sequencers or logs in the system and the curve represents the maximum
+%throughput of a sequencer on an underutilized physical server.  ``Overloaded:
+%No Balancing" adds a second sequencer of equal load to the same physical server
+%with the intent of overloading the single physical server; recall that the
+%throughput for this second sequencer is omitted. Compared to the baseline, the
+%overloaded physical server with 2 sequencers has more than a 2\(\times\)
+%throughput degradation.  The stability of throughput curves (i.e. the variation
+%on the \(y\) axis) indicate which sequencers are on overloaded physical
+%servers.  ``Underloaded (Baseline)" is the best performance for this particular
+%workload because there is no contention.  Performance is stable and the MDS is
+%under-loaded because we do not see the spike at the end like the other curves.
+%``Overloaded: No Balancing" shows an overloaded sequencer because the
+%performance of the monitored sequencer is low and unstable (high variation in
+%performance). 
+%
+%The last two curves of Figure~\ref{fig:mantle-seq-thruput} show the performance
+%improvements when sequencers are load balanced across the cluster.  The load
+%balancing policies differ by the threshold at which they start balancing:
+%``Overloaded: Sequencer Policy" starts balancing immediately and ``Overloaded:
+%Conservative Policy" waits for 120 seconds before migrating sequencers.
+%
+%%Overall, the clients achieve lower performance than the capability experiments
+%%#in Figure~\ref{fig:captp} because the sequencer uses a slower metadata
+%%operation to manage the capability.  It appears the bottleneck is somewhere
+%%with the clients -- perhaps it is our 1GBit/s network link.
+%
+%Figure~\ref{fig:mantle-cpu-utilization} shows how load balancing effects the
+%physical servers by plotting the CPU utilization average over 1 minute (\(y\)
+%axis) over time (\(x\) axis).  It shows the ``Conservative Policy" load
+%balancer from the run in Figure~\ref{fig:mantle-seq-thruput}, so recall that
+%resources are constrained during the job by co-locating a second sequencer on
+%the same physical server.  CPU utilization spikes when physical servers receive
+%load (at 2 minutes physical server 1 receives a sequencer from physical server
+%0) but the CPU utilization never eclipses 50\%. The utilization is low because
+%the load balancer interface is not multi-threaded.  Recall the the load
+%balancing interface re-uses the POSIX metadata cluster migration mechanisms;
+%serving file system metadata is single-threaded because it has been shown that
+%metadata performance does not improve with multiple
+%threads~\cite{konstantinos:pdsw2014-lustre-metadata}.
+
+\noindent\textbf{The Malacology Approach:} pull load balancing out of the
+storage system to balance sequencers across a cluster. Compared to putting all
+sequencers on the same server, load balancing a sequencer per physical server
+shows a 1.8\(\times\) throughput improvement.\\
+
+\noindent\textbf{Features:} different load balancing modes control forwarding
+vs. client redirection, tuning the backoff parameter controls conservative vs.
+aggressive decision making, and tuning load migration unit parameters controls
+sequencer distribution vs. co-location.\\
 

### 2016-10-20 experiments/jupyter.py

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+docker run -d -v `pwd`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook

### 2016-10-20 paper/paper.tex
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>
paper: finish mantle evaluation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % prototype programmable storage system that 
-Malacology
-includes a non-exhaustive set of
-interfaces that can be used as building blocks for constructing novel storage
-abstractions, including:
+Malacology includes a non-exhaustive set of interfaces that can be used as
+building blocks for constructing novel storage abstractions, including:
 
@@
 Malacology our evaluation focuses on the performance of the internal
-abstractions when used to construct the Mantle and ZLog services. First, we
-benchmark scenarios where contention is at the sequencer by examining the
-interfaces used to map ZLog onto Malacology; specifically, we describe the
-sequencer implementation and the propagation of object \& data interfaces
-interfaces. Next, we benchmark scenarios when there are multiple logs by using
-Mantle to (1) balance sequencers across a cluster and (2) recover from failure.
+abstractions used to construct the Mantle and ZLog services. We also discuss
+latent capabilities we discovered in this process that lets us navigate
+different trade-offs within the services themselves. First, we benchmark
+scenarios where contention is at the sequencer by examining the interfaces used
+to map ZLog onto Malacology; specifically, we describe the sequencer
+implementation and the propagation of object \& data interfaces interfaces.
+Next, we benchmark scenarios when there are multiple logs by using Mantle to
+balance sequencers across a cluster.
 
@@
 provide reasonable performance.
-
-We have conducted our evaluation on CloudLab as well as our own 20 node
-cluster. We follow the ``Popper Convention"~\cite{anonymous} so as to try and
-make our experiments as reproducible as possible. We have a GitHub repository
-with our source code, experimental environment, configuration parameters, and
-workloads but due to the double-blind policy, we have temporarily disabled the
-links to our graphs and experimental details.
 
@@
 low as 10ms throughput achieve a significant proportion of the locally cached
-throughput.
+throughput.\\
 
@@
 
-\subsection{Load Balancing Multiple ZLog Sequencers with Mantle}
+\subsection{Load Balancing ZLog Sequencers with Mantle}
 
@@
 load balancer that can migrate the shared resource (in this case, the resource
-that guards the current token) would require building subsystems for migrating
-resources, monitoring the workloads, monitoring physical nodes, partitioning
-resources, cache roherence, and managing multiple sequencers.
-
-The following experiments demonstrate the feasibility of using the mechanisms
-of the Malacology load balancing interface (e.g., migrating, monitoring, and
-partitioning load) to (1) alleviate load from overloaded servers with multiple
-sequencers and (2) recover from failure.
+that mediates access to the tail of the log) would require building subsystems
+for migrating resources, monitoring the workloads, collecting metrics that
+describe the utilization on the physical nodes, partitioning resources,
+maintaining cache coroherence, and managing multiple sequencers.  The following
+experiments demonstrate the feasibility of using the mechanisms of the
+Malacology load balancing interface to inherit all these features and to
+alleviate load from overloaded servers with multiple sequencers.
 
@@
 \includegraphics{figures/mantle-balancer-behaviors.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
-load balancing interface alleviates overloaded servers in environments with multiple sequencers in
+\caption{The load balancing interface lets administrators control the migration
+of sequencers in environments with multiple ZLog logs. For this setup (3
+sequencers and 3 servers), CephFS and Mantle migrate load and show improved
+throughput over the setup where all sequencers are co-located on the same
+server (``No Balancing").  Sections~\ref{sec:feature-balancing-modes}
+and~\ref{sec:feature-migration-units} quantify the performance improvement of
+the stable phase at time 100 seconds and Section~\ref{sec:feature-backoff}
+examines the growth phase at time 0 to 60 seconds.
 }\label{fig:mantle-balancer-behaviors}
@@
 \includegraphics{figures/mantle-balancer-performance.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
-}\label{fig:mantle-balancer-performance}
+\caption{Despite using different balancing modes, the CephFS balancers have the
+same performance because they end up migrating the sequencers in the same way.
+The Mantle bar programs a balancer designed for sequencers into the service
+using the load balancing interface. This is the same workload/setup as
+Figure~\ref{fig:mantle-balancer-behaviors}.}\label{fig:mantle-balancer-performance}
 \end{figure}
@@
 experiment co-locates 3 sequencers each with 4 clients on the same physical
-server.  Instead of measuring contention at the clients like we did in
-Section~\ref{sec:sequencer-implementation}, these experiments create contention
-at the sequencers by setting the capability infrastructure to ``round trip
-mode". In this mode clients go to the sequencer for every request. We
-implement this by piggybacking on \texttt{fstat()} and using the shared
-resource interface to force round trips.  Because the sequencer hands out
-positions for the tail of the log, the workload is read-heavy.
+server.  Instead of measuring contention at the clients like
+Section~\ref{sec:sequencer-implementation}, these experiments measure
+contention at the sequencers by forcing clients to make round-trips every
+request. We implement this using the shared resource interface to piggyback on
+\texttt{fstat()} and to program a policy into \texttt{getattr()} that forces
+round-trips.  Because the sequencer's only function is to  hand out positions
+for the tail of the log, the workload is read-heavy.
 
@@
 ``CephFS" migrates sequencers using the original algorithms packaged with Ceph,
-and the ``Mantle" curve shows the throughput of a custom load balancer we wrote
-specifically for sequencers. The increased throughput for the ``CephFS" and
-``Mantle" curves between 0 and 60 seconds are a result of migrating the
-sequencer(s) off of overloadeed servers. In addition to showing that migrating
-sequencers improves performance, Figure~\ref{fig:mantle-balancer-behaviors}
-also demonstrates features that we will explore in the rest of this section.
-Section~\ref{sec:feature-balancing-modes} and~\ref{sec:feature-migration-units}
-shows why the performance of Mantle is better than CephFS when the cluster
-stabilizes at time 100 seconds and Section~\ref{sec:feature-backoff} explains
-the slope and location of the rebalancing phase between 0 and 60 seconds by
-comparing the aggressiveness of the CephFS and Mantle balancers.
+and the ``Mantle" uses a custom load balancer we wrote specifically for
+sequencers. The increased throughput for the CephFS and Mantle curves between 0
+and 60 seconds are a result of migrating the sequencer(s) off overloadeed
+servers. In addition to showing that migrating sequencers improves performance,
+Figure~\ref{fig:mantle-balancer-behaviors} also demonstrates features that we
+will explore in the rest of this section.
+Sections~\ref{sec:feature-balancing-modes}
+and~\ref{sec:feature-migration-units} quantify the differences in performance
+when the cluster stabilizes at time 100 seconds and
+Section~\ref{sec:feature-backoff} examines the slope and start time of the
+rebalancing phase between 0 and 60 seconds by comparing the aggressiveness of
+the balancers.
 
@@
 The load balancing service can operate in different balancing modes that stress
-the subsystems that handle client requests or forwarded requests in different
-ways. In Figure~\ref{fig:mantle-balancer-behaviors}, the ``CephFS" curve shows
-the performance of the balancing mode that CephFS uses {\it most of the time}.
+the subsystems that receive and forward client requests in different ways. In
+Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
+performance of the balancing mode that CephFS uses {\it most of the time}.
 CephFS has 3 modes for balancing load: CPU mode, workload mode, and hybrid
@@
 into the same mode a majority of the time.  The high variation in performance
-for the ``CephFS CPU Mode" bar reflects the uncertainty of using something as
+for the CephFS CPU Mode bar reflects the uncertainty of using something as
 dynamic and unpredictable as CPU utilization to make migration decisions. In
@@
 CephFS balancers all fall into the same mode despite using different balancers
-is a problem as it prevents administrators from properly exploring the
+is a problem becuase it prevents administrators from properly exploring the
 balancing state space.
@@
 Mantle gives the administrator more control over the balancing policies; for
-the ``Mantle" bar in Figure~\ref{fig:mantle-balancer-performance} we use the
-load balancing interface to program logic for balancing read-heavy workload
-into the load balancing service. The resulting performance is more stable and
-performs better because we created our own mode for balancing sequencers.
-
-We have identified two balancing modes that are relevant for making migration
-decisions for sequencers. Using Mantle, the administrator can put the load
-balancing service into ``proxy mode" or ``sequencer mode". In ``proxy mode" one
-server receives all requests and farms off the requests to slave servers.  In
-``sequencer mode" clients go directly to the server that has their sequencer.
-These modes are illustrated in Figure~\ref{fig:mantle-modes}. ``None" is when
-all sequencers are co-located on one physical server -- performance for that
-mode is shown by the ``No Balancing" curve in
-Figure~\ref{fig:mantle-balancer-behaviors}. ``Sequencer" mode in the lower left
-shows how clients for sequencer 2 are re-directed to server B. The two ``Proxy"
-mode diagrams on the right show how clients continue sending reqeusts to server
-A even though sequencer 2 has migrated to server B (the difference between
-``Proxy (Half)" and ``Proxy (Full)" is discussed in the next section).
+the Mantle bar in Figure~\ref{fig:mantle-balancer-performance} we use the load
+balancing interface to program logic for balancing read-heavy workloads,
+resulting in higher throughput and better stability.  When we did this we also
+identified two balancing modes that are relevant for making migration decisions
+for sequencers.
+
+Using Mantle, the administrator can put the load balancing service into ``proxy
+mode" or ``client mode". In proxy mode one server receives all requests and
+farms off the requests to slave servers; the slaves servers do the actual tail
+finding operation. In client mode, clients interact directly with the server
+that has their sequencer.  These modes are illustrated in
+Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
+co-located on one physical server -- performance for that mode is shown by the
+No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
+Mode" in the lower left shows how clients for sequencer 2 are re-directed to
+server B. The two ``Proxy mode" diagrams on the right show how clients continue
+sending reqeusts to server A even though sequencer 2 has migrated to server B
+(the difference between ``Proxy Mode (Half)" and ``Proxy Mode (Full)" is
+discussed in the next section).
+
+\begin{figure}[t!]
+\centering
+\includegraphics{figures/mantle-modes.png}
+\caption{The client and proxy balancing modes work well for migrating
+sequencers to alleviate overloaded servers. Client mode migrates the sequencer
+and clients start sending requests to the new server. Proxy mode migrates the
+sequencer but clients continue sending their requests to the first server.
+``Half"/``Full" indicate the size of the migration units (discussed in
+Section~\ref{sec:feature-migration-units}).}\label{fig:mantle-modes}
+\end{figure}
 
@@
 \includegraphics{figures/mantle-mode-behavior.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
+\caption{For this 2 sequencer/2 server setup, the performance of proxy mode
+achieves the highest throughput but at the cost of lower throughput for one of
+the sequencers. Client mode is more fair but results in lower cluster
+throughput; the difference in throughput amongst sequencers can be attributed
+to one of the servers doing cache coherence.
 }\label{fig:mantle-mode-behavior}
@@
 
-Figure~\ref{fig:mantle-mode-performance} shows the throughput (\(y\) axis) over
+Figure~\ref{fig:mantle-mode-behavior} shows the throughput (\(y\) axis) over
 time (\(x\) axis) of the two different modes for an environment with only 2
 sequencers (again 4 clients each) and 2 servers. The curves for both sequencers
-in Figure~\ref{fig:mantle-mode-performance}(a) start at less than 1000
-ops/second and at time 60 seconds Mantle migrates Sequencer 1 to the slave
-server. Performance of Sequencer 2 decreases because it has stayed on the proxy
-which now handles all client requests, processes requests for Sequencer 2, and
-forwards requests for Sequencer 1 to the slave server. Sequencer 1 performance
-improves dramatically because lookups for that sequencer are parallelized with
-the operations on server 1. Cluster throughput improves at the cost of
-decreased throughput and unfair throughput shares for Sequencer 2.
-Figure~\ref{fig:mantle-mode-performance}(b) is set to sequencer mode manually
-(no balancing phase) and shows that performance is worse than proxy mode {\it
-and} unfair amongst sequencers. In this case, the scatter-gather process used
-for cache coherence in the metadata protocols causes strain on the server
-housing Sequencer 2 resulting in uneven performance.
-
-Intuitively, sequencer mode is a superior balancing mode and should perform
+in Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second
+and at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
+Performance of Sequencer 2 decreases because it stayed on the proxy which now
+handles all client requests, processes requests for Sequencer 2, and forwards
+requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
+because lookups for that sequencer are parallelized with the operations on the
+other server. Cluster throughput improves at the cost of decreased throughput
+for Sequencer 2.  Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer
+mode manually (no balancing phase) and shows that the cluster throughput is
+worse than the cluster throughput of proxy mode. That graph also shows that
+Sequencer 2 has less throughput than Sequencer 2. In this case, the
+scatter-gather process used for cache coherence in the metadata protocols
+causes strain on the server housing Sequencer 2 resulting in this uneven
+performance.
+
+Intuitively, client mode is a superior balancing mode and should perform
 better; after all, this is the best strategy for distributing POSIX metadata in
-write-heavy workloads~\cite{ren:sc2014-indexfs,
-patil:fast2011-giga,sevilla:sc15-mantle}. But our experiments show that proxy
-mode is superior for the sequencer's read-heavy workload.  In this case,
-separating the handling of the client request and doing the actual operation
-improves performance. Processing client requests and doing the actual operation
-is too heavy weight for one server and sequencers on slave nodes can go faster
-because work is split up; the proxy processes the client request and the slave
-finds the tail of the log and delivers the answer.  This phenomenon is not
+write-heavy workloads~\cite{ren:sc2014-indexfs, patil:fast2011-giga,
+sevilla:sc15-mantle}.  But our experiments show that using proxy mode improves
+performance for this smale scale setup because it separates (1) the handling of
+the client requests and (2) finding the tail of the log and responding to
+clients.  Doing both steps is too heavy weight for one server and sequencers on
+slave nodes can go faster because work is split up; this phenomenon is not
 uncommon and has been observed in chain replication~\cite{CITEME}.
@@
 
-How much load to try and migrate is not a tunable parameter in the original
-CephFS balancers. The most useful way to express how much load to send is as a
-fraction of some other metric such as total load. Expressing this heuristic is
-not easily achievable with outward facing tunable parameters (i.e. system
-knobs) but can be expressed with programmability quite easily. To force the
-load balancer into ``Proxy (Half)" scenario in Figure~\ref{fig:mantle-modes},
-which sets the balancer to proxy mode with migration units of half the load on
-the current server we can use the Mantle policy:\\
+We define a ``migration unit" as the amount of load to try and migrate.  For
+example, it is useful to express migration units as a fraction of the total
+load.  Expressing this heuristic is not easily achievable with outward facing
+tunable parameters (i.e. system knobs) and so the original CephFS balancers do
+not have a migration unit tunable.  With programmability, migration units can
+be expressed quite easily. To force the load balancer into the Proxy Mode (Half)
+scenario in Figure~\ref{fig:mantle-modes}, which sets the balancer to proxy
+mode with migration units of half the load on the current server we can use the
+Mantle policy:\\
 
@@
 the balancer uses to do the migrations. Alternatively, to set the mode to the
-``Proxy (Full)"` setup in Figure~\ref{fig:mantle-modes}, one would use:\\
+``Proxy (Full)"` setup in Figure~\ref{fig:mantle-modes} we use:\\
 
 \noindent \texttt{ targets[\#mds+1] = mds[whoami]["load"] }\\
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-modes.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
-load balancing interface alleviates load for multiple sequencers in
-resource-constrained environments.  ``Underloaded (Baseline)'' is the
-throughput of one sequencer running alone on a cluster; the other curves show
-how load balancing affects the throughput of one sequencer when sequencers for
-other logs are initially co-located on the same physical server (i.e. an
-"Overloaded" environment).}\label{fig:mantle-modes}
-\end{figure}
 
@@
 \includegraphics{figures/mantle-mode-performance.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
-}\label{fig:mantle-mode-performance}
+\caption{The performance of the different modes using various migration units,
+indicated by parentheses, shows almost a 2\(\times\) improvement in the best
+case.  }\label{fig:mantle-mode-performance}
 \end{figure}
@@
 and 2 servers, so performance may be different at scale. Even so, it is clear
-sequencer mode does not perform as well for read-heavy workloads and we even
-see a throughput improvement when migrating all load off the first server,
-leaving the first server to do administrative tasks (this is actually common in
-the metadata cluster because the ``first" server does a lot of the cache
-coherence work) while the second server does all the client work. Proxy Mode
-does the best in both cases and shows large performance gains when completely
-decoupling client request handling and operation processing in Proxy Mode
-(Full).
+that client mode does not perform as well for read-heavy workloads. We
+even see a throughput improvement when migrating all load off the first server,
+leaving the first server to do administrative tasks (this is common in the
+metadata cluster because the first server does a lot of the cache coherence
+work) while the second server does all the client work. Proxy mode does the
+best in both cases and shows large performance gains when completely decoupling
+client request handling and operation processing in Proxy Mode (Full).
 
 The parameter that controls the migration units helps the administrator control
-the sequencer or co-location or distribution across the cluster. This trade-off
-was explored extensively in the Mantle paper and warrants further exploration
-as all our experiments indicate a rich set of states to explore.
-
-\subsubsection{Feature: Backoff Paramters}
+the sequencer co-location or distribution across the cluster. This trade-off
+was explored extensively in the Mantle paper but the experiments we present
+here are indicative of an even richer set of states to explore.
+
+\subsubsection{Feature: Backoff}
 \label{sec:feature-backoff}
@@
 The balancing phase from 0 to 60 seconds in
-Figure~\ref{mantle-balancer-behaviors} shows different degrees of
+Figure~\ref{fig:mantle-balancer-behaviors} shows different degrees of
 aggressiveness in making migration decisions; CephFS makes a decision quickly
-and throughput jumps 2.5K ops/second within 10 seconds while Mantle takes a
+and throughput jumps to 2500 ops/second within 10 seconds while Mantle takes
 more time to stabilize. This conservative behavior is controlled by programming
-the balancer to (1) use different conditions for when to migrate and (2) use a
-threshold for sustained overload. \texttt{when()} is part the Mantle API and
-the callback expects to return a true or false, which corresponds to whether to
-migrate or not. For the Mantle curve in Figure~\ref{mantle-balancer-behaviors}
-we set the \texttt{when()} policy to wait for load on the receiving server to
-reach 0.1. At the beginning of the run the load of the receiving server spikes
-because requests for cache-coherence start as the sequencers are created. It
-takes 60 seconds for this load to drop to an acceptable level. Also, the
-migration phase is a little more conservative in sending load; the Mantle curve
-in Figure~\ref{mantle-balancer-behaviors} takes longer to reach peak throughput
-because we do the first migration and {\it wait to see how it affects the
-system}; seeing that after migration the third server is still underloaded
-allows the balancer to further improve performance by migrating a sequencer to
-the laster server.
+the balancer to (1) use different conditions for when to migrate and (2) using
+a threshold for sustained overload.
+
+We control the conditions for migration using \texttt{when()}. \texttt{when()}
+is part the Mantle API and the callback expects a return value of true or
+false, which corresponds to whether to migrate or not. For the Mantle curve in
+Figure~\ref{fig:mantle-balancer-behaviors} we set the \texttt{when()} policy to
+wait for load on the receiving server to fall below a threshold. At the
+beginning of the run the load of the receiving server spikes because requests
+for cache-coherence start as the sequencers are created. It takes 60 seconds
+for this load to drop to an acceptable level and it is at this point that the
+migration starts.
+
+The Mantle curve in Figure~\ref{fig:mantle-balancer-behaviors} also takes longer
+to reach peak throughput.  This conservative decision making is intentional
+because we want the policy wait to see how migrations affect the system before
+proceeding. The Mantle curve in Figure~\ref{fig:mantle-balancer-behaviors} does
+a migration right before 50 seconds, realizes that there is a third underloaded
+server, and does another migration.
 
 The other way to change aggressiveness of the decision making is to program
-into the balancer a policy that waits a certain number of iterations after a
-migration before proceeding. In Mantle, the policy would use the save state
-function to do a countdown after a migation:
-
-\begin{figure}[]
-\begin{lstlisting}
-  countdown = GET_STATE()
-  if countdown > 0 then
-    -- OK to migrate
-  else
-    PUT_STATE(countdown - 1)
-  end
-\end{lstlisting}
-\caption{The sequencer balancing policy that works best for bursty workloads. 
-\label{listing:greedy-spill}}
-\end{figure}
-
-Behavior graphs and performance numbers for this feature is omitted, as the
-throughput graphs are as expected: the more conservative the approach the less
-overall throughput.
+into the balancer a threshold for sustatined overload. This forces the balancer
+to wait a certain number of iterations after a migration before proceeding. In
+Mantle, the policy would use the save state function to do a countdown after a
+migation.  Behavior graphs and performance numbers for this backoff feature is
+omitted for space considerations, but the throughput graphs are as expected:
+the more conservative the approach the less overall throughput.\\
 
@@
 storage system to balance sequencers across a cluster. Compared to putting all
-sequencers on the same server, load balancing a sequencer per physical server
+sequencers on the same server, load balancing a sequencer using proxy mode
 shows a 1.8\(\times\) throughput improvement.\\
@@
 \noindent\textbf{Features:} different load balancing modes control forwarding
-vs. client redirection, tuning the backoff parameter controls conservative vs.
-aggressive decision making, and tuning load migration unit parameters controls
-sequencer distribution vs. co-location.\\
+vs. client redirection, tuning the load migration unit parameter controls
+sequencer distribution vs. co-location, and tuning the backoff parameter
+controls conservative vs.  aggressive decision making.
 

### 2016-10-20 paper/paper.tex
wordsmithing

@@
 
-We wanted to implement CORFU and we had two choices: build it from the ground
-up or do it on top of Malacology. To show the feasability of building on
-Malacology our evaluation focuses on the performance of the internal
-abstractions used to construct the Mantle and ZLog services. We also discuss
-latent capabilities we discovered in this process that lets us navigate
+%We wanted to implement CORFU and we had two choices: build it from the ground
+%up or do it on top of Malacology. To show the feasability of building on
+%Malacology our evaluation focuses on the
+
+Our evaluation demonstrates the feasibility of building new service abstractions
+atop programmable storage, focusing on the
+performance of the internal
+abstractions exposed by Malacology and
+used to construct the Mantle and ZLog services. We also discuss
+latent capabilities we discovered in this process that let us navigate
 different trade-offs within the services themselves. First, we benchmark
-scenarios where contention is at the sequencer by examining the interfaces used
+scenarios 
+%where contention is at the sequencer 
+with high sequencer contention
+by examining the interfaces used
 to map ZLog onto Malacology; specifically, we describe the sequencer
-implementation and the propagation of object \& data interfaces interfaces.
-Next, we benchmark scenarios when there are multiple logs by using Mantle to
+implementation and the propagation of object and data interfaces interfaces.
+Next, we benchmark scenarios in which the storage system manages multiple logs by using Mantle to
 balance sequencers across a cluster.
 
-Detailed system-wide performance measurements and analysis is forthcoming for
-both Mantle and ZLog, as we try to find the best metadata balancing policies
-and physical design implementation trade-offs.  Since this work
+% should save this for future work, otherwise it undercuts the eval.
+%Detailed system-wide performance measurements and analysis is forthcoming for
+%both Mantle and ZLog, as we try to find the best metadata balancing policies
+%and physical design implementation trade-offs.
+
+Since this work
 focuses on the programmabilty of Malacology, the goal of the benchmarks is to
 show that the components and subsystems that support the Malacology interfaces
-provide reasonable performance.
+provide reasonable performance, as well as to give examples of the flexibility
+that Malacology provides to programmers.
+%As we search for the best metadata balancing policies
+%and physical design implementation trade-offs,
+%we plan to rel
 
@@
 We evaluate the mapping of ZLog onto Malacology by examining the costs of
-utilizing the metadata service as a sequencer, and propagating interface
+utilizing the metadata service as a sequencer, and of propagating interface
 changes throughout the cluster using the monitoring service. We have omitted
@@
 production with acceptable performance, introducing a small constant overhead
-per operation but otherwise does not affect scalability.
+per operation without affecting scalability.
 
@@
 \centering
-\includegraphics{figures/caps-delay-thruput.png}
+\includegraphics{figures/tradeoff.png}
 \caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]

### 2016-10-20 paper/paper.tex
comment

@@
 \noindent\textbf{The Malacology Approach:} Re-use the capability management
-features of the MDS.  While the batching mechanism can achieve high-throughput,
+features of the MDS.  While the batching mechanism can achieve high throughput,
 it is not appropriate for applications that require predictable latency. The
@@
 latency
+\paa{this should probably be a complete sentence or sentences.  also, I don't recall
+seeing any discussion of quotas}
 

### 2016-10-20 paper/paper.tex
comment
comments

@@
 
+\paa{meta point: I am not sure the "The Malacology Approach" thing works, because it
+makes the reader think that we are contrasting with a strawman. I think that if we replace this with (eg)
+``Malacology reuses ....'', ``Malacology pulls'', etc, it reads more smoothly.}
 \noindent\textbf{The Malacology Approach:} Re-use the capability management
@@
 
+\paa{there should probably be a sum-up sentence here; eg Malacology exposes knobs that
+allow application programmers to tune the storage system to the needs of their application.}
+
 \noindent\textbf{Features:} delay and quota parameters control throughput vs.
@@
 the semantics of interface management resulting in one or more interface
-updates that must be consistent across the cluster. The following experiments
-focus on the cost of achieving distribution of consistent a interface version.
+updates that must be consistent across the cluster. \paa{the word ``interface''
+occurs too many times in the last sentence, but I'm not sure how to fix it.}
+The following experiments
+focus on the cost of distributing a consistent interface version.
 
@@
 store were deployed, showing a latency of less than 54 ms with a probability
-of 90\% and a worst case latency of 194 ms. The curve labeled "3 OSD (HDD)" is
+of 90\% and a worst case latency of 194 ms. The curve labeled "3 OSD (HDD)" 
+\paa{DANGER -- didn't we delete this?}
+is
 a smaller cluster where each of the three OSDs uses a spinning disk as the

### 2016-10-20 paper/paper.tex
comments

@@
 For scenarios with multiple shared logs and many clients, co-locating
+\paa{you might want a sentence or two of high-level explanation of this use case.
+in practice, a storage system implementing CORFU will support a multiplicity of independent
+totally-ordered logs (eg, one for each application).}
 sequencers on the same physical node is not ideal. Unfortunately, building a

### 2016-10-21 paper/paper.tex
Minor mods 6 intro and 6.2 intro

--- 2016-10-20 paper/paper.tex
@@
 Since this work
-focuses on the programmabilty of Malacology, the goal of the benchmarks is to
+focuses on the programmabilty of Malacology, the goal of this section is to
 show that the components and subsystems that support the Malacology interfaces
-provide reasonable performance, as well as to give examples of the flexibility
+provide reasonable relative performance, as well as to give examples of the flexibility
 that Malacology provides to programmers.
@@
 For scenarios with multiple shared logs and many clients, co-locating
-sequencers on the same physical node is not ideal. Unfortunately, building a
+sequencers on the same physical node is not ideal. Building a
 load balancer that can migrate the shared resource (in this case, the resource
-that mediates access to the tail of the log) would require building subsystems
+that mediates access to the tail of the log) is a time-consuming, non-trivial task that requires building subsystems
 for migrating resources, monitoring the workloads, collecting metrics that
 describe the utilization on the physical nodes, partitioning resources,
-maintaining cache coroherence, and managing multiple sequencers.  The following
+maintaining cache coherence, and managing multiple sequencers. The following
 experiments demonstrate the feasibility of using the mechanisms of the
@@
 Section~\ref{sec:sequencer-implementation}, these experiments measure
-contention at the sequencers by forcing clients to make round-trips every
-request. We implement this using the shared resource interface to piggyback on
-\texttt{fstat()} and to program a policy into \texttt{getattr()} that forces
+contention at the sequencers by forcing clients to make round-trips for every
+request. We implement this using the shared resource interface that forces
 round-trips.  Because the sequencer's only function is to  hand out positions
@@
 ``CephFS" migrates sequencers using the original algorithms packaged with Ceph,
-and the ``Mantle" uses a custom load balancer we wrote specifically for
+and ``Mantle" uses a custom load balancer we wrote specifically for
 sequencers. The increased throughput for the CephFS and Mantle curves between 0

### 2016-10-21 paper/paper.tex
paper: add capdelay/quota figure
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \title{Malacology: A Programmable Storage System}
-\authorinfo{Paper ID: 123}
-           {Total Page: 12}
+\authorinfo{Paper ID: 165}
+           {Total Page: 14}
 \date{}
@@
 \caption{Scalable storage systems consist of storage daemons which store data
-fragments, monitor daemons that maintain cluster state, and service-specific
+fragments, monitor daemons (M) that maintain cluster state, and service-specific
 daemons such as file system metadata servers. Malacology enables the
@@
 \end{tabular}
-\caption{A variety of RADOS object storage classes exist to expose interfaces
+\caption{A variety of object storage classes exist to expose interfaces
     to applications. \# is the number of methods that use these categories.
@@
 
-\section{Malacology}
+\section{Malacology: A Programmable Storage System}
 \label{sec:malacology}
@@
 \centering
+\includegraphics{figures/capdelay-quota-behavior-a.png}
+\includegraphics{figures/capdelay-quota-behavior-b.png}
+\includegraphics{figures/capdelay-quota-behavior-c.png}
+\caption{ Each dot in these graphs is an individual request and they are spread
+randomly along the \(y\) axis. The default behavior (``None") shows
+unpredictable performance, ``delay" lets each client hold the lease longer, and
+``quota" gives each client the lease for a certain number of operations. Delay
+improves latency and fairness while quota maximizes throughput.}
+\label{fig:capdelay-quota-behavior}
+\end{figure}
+
+\begin{figure}[h]
+\centering
 \includegraphics{figures/tradeoff.png}
@@
 sequencer but clients continue sending their requests to the first server.
-``Half"/``Full" indicate the size of the migration units (discussed in
+``Half" is the size of the migration units (discussed in
 Section~\ref{sec:feature-migration-units}).}\label{fig:mantle-modes}
@@
 
-\noindent\textbf{The Malacology Approach:} pull load balancing out of the
-storage system to balance sequencers across a cluster. Compared to putting all
-sequencers on the same server, load balancing a sequencer using proxy mode
-shows a 1.8\(\times\) throughput improvement.\\
+\noindent\textbf{Malacology pulls} the load balancing out of the storage system
+to balance sequencers across a cluster. Compared to putting all sequencers on
+the same server, load balancing a sequencer using the proxy mode shows a
+1.8\(\times\) throughput improvement.\\
 
@@
 deficiencies. However, this duplication has real-world problems related to
-reliability. We propose that system expose their services in a safe way
+reliability. We propose that systems expose their services in a safe way
 allowing application developers to customize system behavior to meet their
-needs while not sacrificing correctness.
-
-We are intending to pursue this work towards the goal of constructing a set of
-customization points that allow a wide variety of storage system services to be
-configured on-the-fly in existing systems. This work is one point along that
-path in which we have looked at targeting special-purpose storage systems.
-Ultimately we want to utilize declarative methods for expressing new services.
+needs while not sacrificing correctness. To illustrate the benefits of this
+approach we presented Malacology, a programmable storage system that
+facilitates the construction of new services by re-purposing existing subsystem
+abstractions of the storage stack.
+
+Future work will focus on constructing more interfaces to support a wide
+variety of storage system services that can be configured on-the-fly in existing
+systems. This work is one point along that path to producing general-purpose
+storage systems that can
+target  special-purpose applications.  Ultimately we want to utilize
+declarative methods for expressing new services.
 

### 2016-10-21 paper/paper.tex

@@
 For scenarios with multiple shared logs and many clients, co-locating
+\paa{you might want a sentence or two of high-level explanation of this use case.
+in practice, a storage system implementing CORFU will support a multiplicity of independent
+totally-ordered logs (eg, one for each application).}
 sequencers on the same physical node is not ideal. Building a

### 2016-10-21 paper/paper.tex
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-10-21 paper/paper.tex
@@
 
-In Section~\ref{consistencyversioning-of-cluster-state} we described an
-interface management service that is responsible for the durable and
-consistent cluster-wide distribution of interface versions. We evaluate the
-costs of such a service by embedding interface definitions in a Ceph cluster
-membership data structure that is managed by a Paxos instance, and distributed
-using a scalable, peer-to-peer gossip protocol that is built into Ceph. We
-assume an architecture that consists of a light-weight interface that enforces
-the semantics of interface management resulting in one or more interface
-updates that must be consistent across the cluster. \paa{the word ``interface''
-occurs too many times in the last sentence, but I'm not sure how to fix it.}
-The following experiments
-focus on the cost of distributing a consistent interface version.
+Domain-specific data interfaces (Section~\ref{object-data-interfaces}) allow
+co-design between applications and the storage system. Malacology supports
+custom object interfaces in RADOS that require interface implementations to
+installed on the storage devices in the system, supports the evolution of
+interfaces through automatic system-wide versioning and installation through
+the service metdata interface (Section~\ref{service-metadata}). We evaluate
+the performance of installing a new version of an interface in the cluster,
+which is an important metric for applications that are designed with
+frequently evolving interfaces.
+
+We demonstrate the feasibility of utilizing the Ceph monitoring sub-system by
+evaluating the performance of installing and distributing interface updates.
+Figure~\ref{fig:propdelay} shows the CDF of the latency of interface updates.
+The interfaces are Lua scripts embedded in the cluster map and distributed
+using a peer-to-peer gossip protocol.  The latency is defined as the elapsed
+time following the Paxos proposal for an interface update until each OSD makes
+the update live (the cost of the Paxos proposal is configurable and is
+discussed below). The latency measurements were taken on the nodes running
+Ceph server daemons, and thus exclude the client round-trip cost. In each of
+the experiments 1000 interface updates were observed.
+
+Figure~\ref{fig:propdelay} shows the lower bound cost for updates in a large
+cluster. In the experiment labeled "120 OSD (RAM)" a cluster of 120 OSDs (10
+OSDs x 12 servers) using an in-memory data store were deployed, showing a
+latency of less than 54 ms with a probability of 90\% and a worst case latency
+of 194 ms. These costs demonstrate the penality of distributing the interface
+in a large cluster. In In practice the costs are, in addition to cluster-wide
+propagation of interface updates, includes the network round-trip to the
+interface management service, the Paxos commit protocol, and other factors
+such as system load. By default Paxos proposals occur periodically with a 1
+second interval in order to accumulate updates. In a minimum, realistic quorum
+of 3 monitors using HDD-based storage, we were able to decrease this interval
+to an average of 222 ms.
 
@@
 \includegraphics[trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
-Cluster-wide interface update latency, excluding the Paxos proposal cost for
-interface commit. Each of the 120 OSDs used a RAM-based object store, while the
-3 OSDs used an HDD-backed object store.}
+\caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
+interface commit.}
 \label{fig:propdelay}
 \end{figure}
-
-Figure~\ref{fig:propdelay} shows the CDF of the latency of installing and
-distributing an interface update on all OSDs in the cluster. The latency is
-defined as the elapsed time following the Paxos proposal for an interface
-update until each OSD makes the update live (the cost of the Paxos proposal is
-configurable and is discussed below). The latency measurements were taken on
-the nodes running Ceph server daemons, and thus exclude the client round-trip
-cost. In each of the experiments 1000 interface updates were observed.
-
-The first experiment shown in Figure~\ref{fig:propdelay} illustrates a lower
-bound cost for updates in a large cluster by avoiding disk I/O for storing
-interface updates locally within each OSD.  In the experiment labeled "120 OSD
-(RAM)" a cluster of 120 OSDs (10 OSDs x 12 servers) using an in-memory data
-store were deployed, showing a latency of less than 54 ms with a probability
-of 90\% and a worst case latency of 194 ms. The curve labeled "3 OSD (HDD)" 
-\paa{DANGER -- didn't we delete this?}
-is
-a smaller cluster where each of the three OSDs uses a spinning disk as the
-object store that saves a copy of the cluster map, and shows a latency of 37
-ms with a 90\% probability and a worst case latency of 255 ms. Since the cost
-of a disk I/O is larger than the same I/O executed against the in-memory
-device, the predominant cost in a lightly loaded cluster appears to be the
-number of OSDs.
-
-\textbf{The Malacology Approach:} Re-use Ceph's Monitor functionality to
-propagate application-specific state that needs to be highly-available and
-strongly consistent.  The real cost of interface management, in addition to
-cluster-wide propagation of interface updates, includes the network round-trip
-to the interface management service, the Paxos commit protocol, and other
-factors such as system load. By default Paxos proposals occur periodically with
-a 1 second interval in order to accumulate updates. In a minimum, realistic
-quorum of 3 monitors using HDD-based storage, we were able to decrease this
-interval to an average of 222 ms.
 

### 2016-10-21 paper/paper.tex
tweak

--- 2016-10-21 paper/paper.tex
@@
 used to construct the Mantle and ZLog services. We also discuss
-latent capabilities we discovered in this process that let us navigate
+latent capabilities we discovered in this process that allow us---as well as future
+storage system programmers---to navigate
 different trade-offs within the services themselves. First, we benchmark

### 2016-10-21 paper/paper.tex
tweaks

@@
 custom object interfaces in RADOS that require interface implementations to
-installed on the storage devices in the system, supports the evolution of
+be installed on the storage devices in the system, supporting the evolution of
 interfaces through automatic system-wide versioning and installation through
@@
 of 194 ms. These costs demonstrate the penality of distributing the interface
-in a large cluster. In In practice the costs are, in addition to cluster-wide
-propagation of interface updates, includes the network round-trip to the
-interface management service, the Paxos commit protocol, and other factors
+in a large cluster. In practice the costs include, in addition to cluster-wide
+propagation of interface updates, the network round-trip to the
+interface management service, the Paxos commit protocol itself, and other factors
 such as system load. By default Paxos proposals occur periodically with a 1

### 2016-10-21 paper/paper.tex
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 \label{sec:mapping-zlog-onto-malacology}
-We evaluate the mapping of ZLog onto Malacology by examining the costs of
-utilizing the metadata service as a sequencer, and of propagating interface
-changes throughout the cluster using the monitoring service. We have omitted
-evaluation of the overhead costs of adding object interfaces themselves due to
-lack of space, and the fact that they are heavily tested and used already in
-production with acceptable performance, introducing a small constant overhead
-per operation without affecting scalability.
+
+We evaluate Malacology by exploring one possible mapping of the ZLog
+implementation of CORFU onto Ceph in which we re-use (1) the metadata service
+to manage naming and synchronization of the sequencer resource by treating the
+resource as an inode, and (2) the monitoring sub-subsystem to distribute and
+install application-specific I/O interface required of the CORFU protcool. In
+Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the inode
+abstraction for implementing the sequencer resource enables load balancing policies
+to migrate the sequencer resource in heavy-load situation.
 
@@
 \subsection{Load Balancing ZLog Sequencers with Mantle}
+\label{sec:zlog-balancing}
 

### 2016-10-21 paper/paper.tex
paper: transitions for mantle eval
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-10-21 paper/paper.tex
@@
 
-Our evaluation demonstrates the feasibility of building new service abstractions
-atop programmable storage, focusing on the
-performance of the internal
-abstractions exposed by Malacology and
-used to construct the Mantle and ZLog services. We also discuss
-latent capabilities we discovered in this process that allow us---as well as future
-storage system programmers---to navigate
-different trade-offs within the services themselves. First, we benchmark
-scenarios 
-%where contention is at the sequencer 
-with high sequencer contention
-by examining the interfaces used
-to map ZLog onto Malacology; specifically, we describe the sequencer
-implementation and the propagation of object and data interfaces interfaces.
-Next, we benchmark scenarios in which the storage system manages multiple logs by using Mantle to
-balance sequencers across a cluster.
+Our evaluation demonstrates the feasibility of building new service
+abstractions atop programmable storage, focusing on the performance of the
+internal abstractions exposed by Malacology and used to construct the Mantle
+and ZLog services. We also discuss latent capabilities we discovered in this
+process that let us navigate different trade-offs within the services
+themselves. First, we benchmark scenarios with high sequencer contention by
+examining the interfaces used to map ZLog onto Malacology; specifically, we
+describe the sequencer implementation and the propagation of object and data
+interfaces interfaces.  Next, we benchmark scenarios in which the storage
+system manages multiple logs by using Mantle to balance sequencers across a
+cluster.
 
@@
 
-Since this work
-focuses on the programmabilty of Malacology, the goal of this section is to
-show that the components and subsystems that support the Malacology interfaces
-provide reasonable relative performance, as well as to give examples of the flexibility
-that Malacology provides to programmers.
+Since this work focuses on the programmabilty of Malacology, the goal of this
+section is to show that the components and subsystems that support the
+Malacology interfaces provide reasonable relative performance, as well as to
+give examples of the flexibility that Malacology provides to programmers.
+
 %As we search for the best metadata balancing policies
@@
 
-For scenarios with multiple shared logs and many clients, co-locating
-\paa{you might want a sentence or two of high-level explanation of this use case.
-in practice, a storage system implementing CORFU will support a multiplicity of independent
-totally-ordered logs (eg, one for each application).}
-sequencers on the same physical node is not ideal. Building a
-load balancer that can migrate the shared resource (in this case, the resource
-that mediates access to the tail of the log) is a time-consuming, non-trivial task that requires building subsystems
-for migrating resources, monitoring the workloads, collecting metrics that
-describe the utilization on the physical nodes, partitioning resources,
-maintaining cache coherence, and managing multiple sequencers. The following
-experiments demonstrate the feasibility of using the mechanisms of the
-Malacology load balancing interface to inherit all these features and to
-alleviate load from overloaded servers with multiple sequencers.
+In practice, a storage system implementing CORFU will support a multiplicity of
+independent totally-ordered logs for each application.  For this scenario
+co-locating sequencers on the same physical node is not ideal. Building a load
+balancer that can migrate the shared resource (in this case, the resource that
+mediates access to the tail of the log) is a time-consuming, non-trivial task
+that requires building subsystems for migrating resources, monitoring the
+workloads, collecting metrics that describe the utilization on the physical
+nodes, partitioning resources, maintaining cache coherence, and managing
+multiple sequencers. The following experiments demonstrate the feasibility of
+using the mechanisms of the Malacology load balancing interface to inherit all
+these features and to alleviate load from overloaded servers housing multiple
+sequencers.
 
@@
 
+First, we show how the ZLog service can orchestrate multiple sequencers in a
+cluster using the Malacology load balancing interface.
 Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput (\(y\) axis)
@@
 acround the cluster; ``No Balancing" keeps all sequencers on one server,
-``CephFS" migrates sequencers using the original algorithms packaged with Ceph,
-and ``Mantle" uses a custom load balancer we wrote specifically for
-sequencers. The increased throughput for the CephFS and Mantle curves between 0
-and 60 seconds are a result of migrating the sequencer(s) off overloadeed
-servers. In addition to showing that migrating sequencers improves performance,
+``CephFS" migrates sequencers using the CephFS load balancers, and ``Mantle"
+uses a custom load balancer we wrote specifically for sequencers. The CephFS
+load balancers are hard-coded into Ceph while Mantle is an API for
+balancing load presented in~Mantle~\cite{sevilla:sc15-mantle} and
+re-implemented on top of Malacology in this paper.  The increased throughput
+for the CephFS and Mantle curves between 0 and 60 seconds are a result of
+migrating the sequencer(s) off overloadeed servers.
+
+In addition to showing that migrating sequencers improves performance,
 Figure~\ref{fig:mantle-balancer-behaviors} also demonstrates features that we
@@
 
-The load balancing service can operate in different balancing modes that stress
-the subsystems that receive and forward client requests in different ways. In
+Next, we quantify the performance benefits shown in
+Figure~\ref{fig:mantle-balancer-behaviors}.  To understand why load balancers
+perform differently we need to explain the different balancing modes that the
+load balancer service uses and how they stress the subsystems that receive and
+forward client requests in different ways. In
 Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
-performance of the balancing mode that CephFS uses {\it most of the time}.
+performance of the balancing mode that CephFS falls into {\it most of the time}.
 CephFS has 3 modes for balancing load: CPU mode, workload mode, and hybrid
 mode. All three have the same structure for making migration decisions but vary
-based on the metric used to calculate load. For this sequencer workload, the 3
-different modes all have the same performance, as shown in
+based on the metric used to calculate load. For this sequencer workload the 3
+different modes all have the same performance, shown in
 Figure~\ref{fig:mantle-balancer-performance}, because the load balancer falls
@@
 
-Figure~\ref{fig:mantle-mode-behavior} shows the throughput (\(y\) axis) over
-time (\(x\) axis) of the two different modes for an environment with only 2
-sequencers (again 4 clients each) and 2 servers. The curves for both sequencers
-in Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second
-and at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
+Figure~\ref{fig:mantle-mode-behavior} shows the throughput over time of the two
+different modes for an environment with only 2 sequencers (again 4 clients
+each) and 2 servers. The curves for both sequencers in
+Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second and
+at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
 Performance of Sequencer 2 decreases because it stayed on the proxy which now
@@
 requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
-because lookups for that sequencer are parallelized with the operations on the
-other server. Cluster throughput improves at the cost of decreased throughput
-for Sequencer 2.  Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer
-mode manually (no balancing phase) and shows that the cluster throughput is
-worse than the cluster throughput of proxy mode. That graph also shows that
-Sequencer 2 has less throughput than Sequencer 2. In this case, the
-scatter-gather process used for cache coherence in the metadata protocols
-causes strain on the server housing Sequencer 2 resulting in this uneven
-performance.
-
-Intuitively, client mode is a superior balancing mode and should perform
-better; after all, this is the best strategy for distributing POSIX metadata in
-write-heavy workloads~\cite{ren:sc2014-indexfs, patil:fast2011-giga,
-sevilla:sc15-mantle}.  But our experiments show that using proxy mode improves
-performance for this smale scale setup because it separates (1) the handling of
+because distributing the sequencers in this way separates (1) the handling of
 the client requests and (2) finding the tail of the log and responding to
@@
 slave nodes can go faster because work is split up; this phenomenon is not
-uncommon and has been observed in chain replication~\cite{CITEME}.
+uncommon and has been observed in chain replication~\cite{CITEME}.  Cluster
+throughput improves at the cost of decreased throughput for Sequencer 2.
+Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer mode manually (no
+balancing phase) and shows that the cluster throughput is worse than the
+cluster throughput of proxy mode. That graph also shows that Sequencer 2 has
+less throughput than Sequencer 2. In this case, the scatter-gather process used
+for cache coherence in the metadata protocols causes strain on the server
+housing Sequencer 2 resulting in this uneven performance.
 
@@
 
-We define a ``migration unit" as the amount of load to try and migrate.  For
-example, it is useful to express migration units as a fraction of the total
+Another factor that affects performance in this environment is ``how much" load
+is on each server; these experiments quantify that effect by programming
+policies into the load balancing interface to control the amount of load to
+migrate. We defined a  migration unit as the amount of load to try and migrate.
+For example, it is useful to express migration units as a fraction of the total
 load.  Expressing this heuristic is not easily achievable with outward facing
@@
 not have a migration unit tunable.  With programmability, migration units can
-be expressed quite easily. To force the load balancer into the Proxy Mode (Half)
-scenario in Figure~\ref{fig:mantle-modes}, which sets the balancer to proxy
-mode with migration units of half the load on the current server we can use the
-Mantle policy:\\
+be expressed quite easily. To force the load balancer into the Proxy Mode
+(Half) scenario in Figure~\ref{fig:mantle-modes}, which sets the balancer to
+proxy mode with migration units of half the load on the current server we can
+use the Mantle policy:\\
 
@@
 
-The balancing phase from 0 to 60 seconds in
-Figure~\ref{fig:mantle-balancer-behaviors} shows different degrees of
-aggressiveness in making migration decisions; CephFS makes a decision quickly
-and throughput jumps to 2500 ops/second within 10 seconds while Mantle takes
-more time to stabilize. This conservative behavior is controlled by programming
-the balancer to (1) use different conditions for when to migrate and (2) using
-a threshold for sustained overload.
+Tuning the aggressiveness of the load balancer decision making is also a
+trade-off that administrators can control and explore. The balancing phase from
+0 to 60 seconds in Figure~\ref{fig:mantle-balancer-behaviors} shows different
+degrees of aggressiveness in making migration decisions; CephFS makes a
+decision quickly and throughput jumps to 2500 ops/second within 10 seconds
+while Mantle takes more time to stabilize. This conservative behavior is
+controlled by programming the balancer to (1) use different conditions for when
+to migrate and (2) using a threshold for sustained overload.
 
@@
 
-%We use a balancer policy specifically designed for multiple sequencers. The
-%balancer aggressively sheds half its load to the next server as soon as it can.
-%It is based off the ``Greedy Spill Even" approach in the
-%Mantle~\cite{sevilla:sc15-mantle} paper but is reproduced in
-%Figure~\ref{listing:greedy-spill} for the reader's convenience. Each line of
-%code uses variables exposed by the Mantle API and is executed as a callback by
-%Mantle.  ``Server load" lets the administrator control what metric constitutes
-%load on a server; ``i" is a variable used by Mantle to iterate over servers and
-%``req\_rate" is a metric based on the exponenential decay of the request rate.
-%``When policy" is used by Mantle to set a conditional for kicking off
-%migration; in this case, the cluster migrates load if the current server,
-%``whoami", has load (\texttt{>0.1}) and the next server (physical servers have
-%``ranks") has a small amount load (\texttt{<0.1}). Finally, the ``where policy"
-%is used by Mantle to determine how much load to shed. The policy fills in a
-%``targets" array to tell the migrator how much load to send to each server; in
-%this case, we shed half of the current load to the next server. The unit of
-%load is a single sequencer for a ZLog log.
-%
-%\begin{figure}[]
-%\begin{lstlisting}
-%  -- Server load
-%  load = server[i]["req_rate"]
-%  
-%  -- When policy
-%  if server[whoami]["load"]>.01 and 
-%    server[whoami+1]["load"]<.01 then
-%    
-%  -- Where policy
-%  targets[whoami+1]=load/2
-%\end{lstlisting}
-%\caption{The sequencer balancing policy that works best for bursty workloads. 
-%\label{listing:greedy-spill}}
-%\end{figure}
-%
-%Figure~\ref{fig:mantle-seq-thruput} shows a performance profile of one of the
-%sequencers as the load balancing interface migrates sequencers around the
-%cluster to alleviate load.  Each curve shows the throughput (\(y\) axis) over
-%time (\(x\) axis).  ``Underloaded (Baseline)'' is the performance of one
-%sequencer on one physical server serving 7 clients - there are no other
-%sequencers or logs in the system and the curve represents the maximum
-%throughput of a sequencer on an underutilized physical server.  ``Overloaded:
-%No Balancing" adds a second sequencer of equal load to the same physical server
-%with the intent of overloading the single physical server; recall that the
-%throughput for this second sequencer is omitted. Compared to the baseline, the
-%overloaded physical server with 2 sequencers has more than a 2\(\times\)
-%throughput degradation.  The stability of throughput curves (i.e. the variation
-%on the \(y\) axis) indicate which sequencers are on overloaded physical
-%servers.  ``Underloaded (Baseline)" is the best performance for this particular
-%workload because there is no contention.  Performance is stable and the MDS is
-%under-loaded because we do not see the spike at the end like the other curves.
-%``Overloaded: No Balancing" shows an overloaded sequencer because the
-%performance of the monitored sequencer is low and unstable (high variation in
-%performance). 
-%
-%The last two curves of Figure~\ref{fig:mantle-seq-thruput} show the performance
-%improvements when sequencers are load balanced across the cluster.  The load
-%balancing policies differ by the threshold at which they start balancing:
-%``Overloaded: Sequencer Policy" starts balancing immediately and ``Overloaded:
-%Conservative Policy" waits for 120 seconds before migrating sequencers.
-%
-%%Overall, the clients achieve lower performance than the capability experiments
-%%#in Figure~\ref{fig:captp} because the sequencer uses a slower metadata
-%%operation to manage the capability.  It appears the bottleneck is somewhere
-%%with the clients -- perhaps it is our 1GBit/s network link.
-%
-%Figure~\ref{fig:mantle-cpu-utilization} shows how load balancing effects the
-%physical servers by plotting the CPU utilization average over 1 minute (\(y\)
-%axis) over time (\(x\) axis).  It shows the ``Conservative Policy" load
-%balancer from the run in Figure~\ref{fig:mantle-seq-thruput}, so recall that
-%resources are constrained during the job by co-locating a second sequencer on
-%the same physical server.  CPU utilization spikes when physical servers receive
-%load (at 2 minutes physical server 1 receives a sequencer from physical server
-%0) but the CPU utilization never eclipses 50\%. The utilization is low because
-%the load balancer interface is not multi-threaded.  Recall the the load
-%balancing interface re-uses the POSIX metadata cluster migration mechanisms;
-%serving file system metadata is single-threaded because it has been shown that
-%metadata performance does not improve with multiple
-%threads~\cite{konstantinos:pdsw2014-lustre-metadata}.
-
 \noindent\textbf{Malacology pulls} the load balancing out of the storage system
@@
 the same server, load balancing a sequencer using the proxy mode shows a
-1.8\(\times\) throughput improvement.\\
-
-\noindent\textbf{Features:} different load balancing modes control forwarding
-vs. client redirection, tuning the load migration unit parameter controls
-sequencer distribution vs. co-location, and tuning the backoff parameter
-controls conservative vs.  aggressive decision making.
+1.8\(\times\) throughput improvement. This latent capability also gives future
+programmers the ability to explore the different load balancing trade-offs
+including: load balancing modes to control forwarding vs. client redirection,
+load migration units to control sequencer distribution vs. co-location, and
+backoffs to control conservative vs. aggressive decision making.
 

### 2016-10-21 paper/paper.tex
paper: reduce captions and cut figures in half
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics{figures/overview.png}
-\caption{Scalable storage systems consist of storage daemons which store data
-fragments, monitor daemons (M) that maintain cluster state, and service-specific
-daemons such as file system metadata servers. Malacology enables the
-programmability of internal abstractions (indicated by bold arrows) in order to
-re-use and compose existing subsystems.  With Malacology, we add two new
-services, ZLog and Mantle, that sit alongside the traditional user-facing APIs
-(file, block, and object) in the storage stack.  \label{fig:overview}}
+\caption{Scalable storage systems have storage daemons which store data,
+monitor daemons (M) that maintain cluster state, and service-specific daemons
+({\it e.g.}, file system metadata servers). Malacology enables the programmability of
+internal abstractions (bold arrows) to re-use and compose
+existing subsystems.  With Malacology, we add new services, ZLog and
+Mantle, that sit alongside traditional user-facing APIs (file, block,
+object).}\label{fig:overview}
 \end{figure}
@@
 \centering
-\includegraphics[trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
+\includegraphics[width=75mm,trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
 \caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
@@
 \includegraphics{figures/mantle-balancer-behaviors.png}
-\caption{The load balancing interface lets administrators control the migration
-of sequencers in environments with multiple ZLog logs. For this setup (3
-sequencers and 3 servers), CephFS and Mantle migrate load and show improved
-throughput over the setup where all sequencers are co-located on the same
-server (``No Balancing").  Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify the performance improvement of
-the stable phase at time 100 seconds and Section~\ref{sec:feature-backoff}
-examines the growth phase at time 0 to 60 seconds.
+\caption{CephFS/Mantle load balancing have better throughput than co-locating
+all sequencers on the same server.  Sections~\ref{sec:feature-balancing-modes}
+and~\ref{sec:feature-migration-units} quantify this improvement;
+Section~\ref{sec:feature-backoff} examines the migration at 0-60 seconds.
 }\label{fig:mantle-balancer-behaviors}
@@
 \includegraphics{figures/mantle-balancer-performance.png}
-\caption{Despite using different balancing modes, the CephFS balancers have the
-same performance because they end up migrating the sequencers in the same way.
-The Mantle bar programs a balancer designed for sequencers into the service
-using the load balancing interface. This is the same workload/setup as
-Figure~\ref{fig:mantle-balancer-behaviors}.}\label{fig:mantle-balancer-performance}
+\caption{All CephFS balancing modes have the same performance Mantle uses a
+balancer designed for sequencers.}\label{fig:mantle-balancer-performance}
 \end{figure}
@@
 sending reqeusts to server A even though sequencer 2 has migrated to server B
-(the difference between ``Proxy Mode (Half)" and ``Proxy Mode (Full)" is
-discussed in the next section).
+(the meaning of ``(Half)" is discussed in the next section).
 
@@
 \includegraphics{figures/mantle-modes.png}
-\caption{The client and proxy balancing modes work well for migrating
-sequencers to alleviate overloaded servers. Client mode migrates the sequencer
-and clients start sending requests to the new server. Proxy mode migrates the
-sequencer but clients continue sending their requests to the first server.
-``Half" is the size of the migration units (discussed in
-Section~\ref{sec:feature-migration-units}).}\label{fig:mantle-modes}
+\caption{ In client mode clients sending requests to the server that houses
+their sequencer. In proxy mode clients continue sending their requests to the
+first server.  }\label{fig:mantle-modes}
 \end{figure}
@@
 \includegraphics{figures/mantle-mode-behavior.png}
-\caption{For this 2 sequencer/2 server setup, the performance of proxy mode
-achieves the highest throughput but at the cost of lower throughput for one of
-the sequencers. Client mode is more fair but results in lower cluster
-throughput; the difference in throughput amongst sequencers can be attributed
-to one of the servers doing cache coherence.
+\caption{The performance of proxy mode achieves the highest throughput but at
+the cost of lower throughput for one of the sequencers. Client mode is more
+fair but results in lower cluster throughput.
 }\label{fig:mantle-mode-behavior}
@@
 policies into the load balancing interface to control the amount of load to
-migrate. We defined a  migration unit as the amount of load to try and migrate.
+migrate. We define migration unit as the amount of load to try and migrate.
 For example, it is useful to express migration units as a fraction of the total
@@
 tunable parameters (i.e. system knobs) and so the original CephFS balancers do
-not have a migration unit tunable.  With programmability, migration units can
-be expressed quite easily. To force the load balancer into the Proxy Mode
-(Half) scenario in Figure~\ref{fig:mantle-modes}, which sets the balancer to
-proxy mode with migration units of half the load on the current server we can
-use the Mantle policy:\\
-
-\noindent \texttt{ targets[\#mds+1] = mds[whoami]["load"]/2 }\\
+not have a migration unit tunable.  With Mantle's programmable interface, we
+force the load balancer into the Proxy Mode (Full) scenario in
+Figure~\ref{fig:mantle-modes}, which sets the balancer to proxy mode with
+migration units equal to the entire load on the current server we can use the Mantle
+policy:\\
+
+\noindent \texttt{ targets[\#mds+1] = mds[whoami]["load"] }\\
 
@@
 the balancer uses to do the migrations. Alternatively, to set the mode to the
-``Proxy (Full)"` setup in Figure~\ref{fig:mantle-modes} we use:\\
-
-\noindent \texttt{ targets[\#mds+1] = mds[whoami]["load"] }\\
+``Proxy (Half)"` setup in Figure~\ref{fig:mantle-modes} we can divide the right
+hand side by 2.
 
@@
 \includegraphics{figures/mantle-mode-performance.png}
-\caption{The performance of the different modes using various migration units,
-indicated by parentheses, shows almost a 2\(\times\) improvement in the best
-case.  }\label{fig:mantle-mode-performance}
+\caption{The performance of the different modes using various migration units
+shows almost a 2\(\times\) improvement in the best case.
+}\label{fig:mantle-mode-performance}
 \end{figure}
@@
 and 2 servers, so performance may be different at scale. Even so, it is clear
-that client mode does not perform as well for read-heavy workloads. We
-even see a throughput improvement when migrating all load off the first server,
-leaving the first server to do administrative tasks (this is common in the
-metadata cluster because the first server does a lot of the cache coherence
-work) while the second server does all the client work. Proxy mode does the
-best in both cases and shows large performance gains when completely decoupling
-client request handling and operation processing in Proxy Mode (Full).
-
-The parameter that controls the migration units helps the administrator control
-the sequencer co-location or distribution across the cluster. This trade-off
-was explored extensively in the Mantle paper but the experiments we present
-here are indicative of an even richer set of states to explore.
+that client mode does not perform as well for read-heavy workloads. We even see
+a throughput improvement when migrating all load off the first server, leaving
+the first server to do administrative tasks (this is common in the metadata
+cluster because the first server does a lot of the cache coherence work) while
+the second server does all the client work. Proxy mode does the best in both
+cases and shows large performance gains when completely decoupling client
+request handling and operation processing in Proxy Mode (Full).  The parameter
+that controls the migration units helps the administrator control the sequencer
+co-location or distribution across the cluster. This trade-off was explored
+extensively in the Mantle paper but the experiments we present here are
+indicative of an even richer set of states to explore.
 
@@
 for this load to drop to an acceptable level and it is at this point that the
-migration starts.
-
-The Mantle curve in Figure~\ref{fig:mantle-balancer-behaviors} also takes longer
-to reach peak throughput.  This conservative decision making is intentional
-because we want the policy wait to see how migrations affect the system before
-proceeding. The Mantle curve in Figure~\ref{fig:mantle-balancer-behaviors} does
-a migration right before 50 seconds, realizes that there is a third underloaded
-server, and does another migration.
+migration starts.  The Mantle curve in
+Figure~\ref{fig:mantle-balancer-behaviors} also takes longer to reach peak
+throughput.  This conservative decision making is intentional because we want
+the policy wait to see how migrations affect the system before proceeding. The
+Mantle curve in Figure~\ref{fig:mantle-balancer-behaviors} does a migration
+right before 50 seconds, realizes that there is a third underloaded server, and
+does another migration.
 

### 2016-10-21 paper/paper.tex
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-We evaluate the performance of the ZLog Sequencer implemented in several
-different configurations by showing a trade-off between aggregate throughput
-achieved (shown in Figure~\ref{fig:captp}), and latency distribution (shown in
-Figure~\ref{fig:capcdf}).
-
-In Figure~\ref{fig:captp} the highest throughput labeled $1C$ shows that a
-single client caching the sequencer resource may obtain new sequences at a rate
-of nearly 6 million positions per second. This is not surprising given that all
-state is cached locally in the client. When this resource is shared between two
-clients, the MDS uses a best-effort approach to granting the sequencer resource
-to each client (value $2C,BE$). This sharing has a large impact on performance,
-as shown by the nearly three orders of magnitude drop off in throughput.
+We evaluate the feasibility of using the metadata service to implement a
+seqeuencer resource that is responsible for maintaining a total ordering of
+the log. Clients contact the sequencer to obtain the tail of the log and then
+independently initiate I/O, thus we measure both the throughput and latency of
+obtaining new tail positions which inherently bound client append performance.
+
+The sequencer is implemented as a custom file type in which the sequencer
+state (a 64-bit integer) is embedded in the inode of a file representing the
+sequencer. A total ordering of the log is imposed by the re-use of the
+capability service that can be used to grant exclusive access of inode state
+to clients. The metadata service is responsible for maintaining exclusivity
+and granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
+behavior of the system in which a best-effort policy is used.  The two colors
+represent points in time that the clients were able to access the resource.
+The best-effort policy provides interleaving a high degree of interleaving
+between clients but the system spends a large portion of time re-distributing
+the capability, reducing overall throughput.
+
+In order to control the performance of the system we implement a
+policy that (1) restricts the length of time that a client may maintain
+exclusive access and (2) limits the number of log positions that a client may
+generate without yielding to other clients waiting for access. The behavior of
+these two modes is illustrated in Figures~\ref{fig:capdelay-quota-behavior}
+(b) and (c), respectively.
 
@@
 \includegraphics{figures/capdelay-quota-behavior-c.png}
-\caption{ Each dot in these graphs is an individual request and they are spread
+\caption{Each dot in these graphs is an individual request and they are spread
 randomly along the \(y\) axis. The default behavior (``None") shows
@@
 \end{figure}
+
+Figure~\ref{fig:captp} demonstrates a configurable trade-off between
+throughput and latency. In the experiment two clients are run each with a
+fixed 0.25 second maximum reservation on the capability, and we vary the size
+of the log position quota running each configuration for two minutes. The
+total operations per second is the combined throughput of the two clients, and
+the average latency is number of microseconds required to obtain a new log
+position. With a small quota more time is spent exchanging exclusive access,
+while a large quota reservation allows clients to experience a much lower
+latency because they uncontended access for a longer period of time.
 
@@
 
+To get a better picture of latency, Figure~\ref{fig:capcdf} shows the CDF of
+latency for each client in all experiment configurations. At the 99th
+percentile clients accessed the sequencer is less than a millisecond. The CDF
+is cropped at the 99.999th percentile due to large outliers that we believe
+occur in instances in which the MDS is performing I/O while it is in the
+process of re-distributing the capability to another client.
+
 \begin{figure}[h]
@@
 \includegraphics{figures/caps-delay-latency.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
-latency distribution shows that if the clients hold their capabilities longer,
+\caption{
+The latency distribution shows that if the clients hold their capabilities longer,
 performance decreases. Changing the delay via policy can help the sequencer and
@@
 
-While a best-effort approach to sharing capabilities makes sense for a file
-system client, throughput drops significantly with two clients as the majority
-of the time neither client has the capability as it ping-pongs between
-clients. This observation reveals a tuning parameter: how long a client
-maintains its capability lease. In throughput labeled $2C,<time>$ we show
-aggregate throughput as we decrease the amount of time that a client holds its
-lease before responding to a request for the capability.  At lease times as
-low as 10ms throughput achieve a significant proportion of the locally cached
-throughput.\\
-
-\paa{meta point: I am not sure the "The Malacology Approach" thing works, because it
-makes the reader think that we are contrasting with a strawman. I think that if we replace this with (eg)
-``Malacology reuses ....'', ``Malacology pulls'', etc, it reads more smoothly.}
-\noindent\textbf{The Malacology Approach:} Re-use the capability management
-features of the MDS.  While the batching mechanism can achieve high throughput,
-it is not appropriate for applications that require predictable latency. The
-throughput in Figure~\ref{fig:captp} labeled $2C,ZLog$ corresponds to a
-centralized sequencer implementation. While throughput is lower for two clients
-(because of the synchronous network round-trip required for each request),
-Figure~\ref{fig:capcdf} shows that this approach achieves more predicable
-latency compared to all other approaches except for the single client locally
-cached mode.\\
-
-\paa{there should probably be a sum-up sentence here; eg Malacology exposes knobs that
-allow application programmers to tune the storage system to the needs of their application.}
-
-\noindent\textbf{Features:} delay and quota parameters control throughput vs.
-latency
-\paa{this should probably be a complete sentence or sentences.  also, I don't recall
-seeing any discussion of quotas}
+We have shown how exposing the internal capability management service allows
+users to navigate latency and throughput tradeoffs.  Other approaches to
+designing the sequencer service also exist, such as using a centralized
+service in which each access is a network round-trip. In contrast to the
+mechanism we explored which is appropriate for clients with bursty workloads,
+it may be easier to provide predictable performance using a centralized
+service, and we will be exploring in the future how this can be achieved using
+the capability system.
 

### 2016-10-21 paper/paper.tex
paper: prune backoff evaluation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-10-21 paper/paper.tex
@@
 
-We control the conditions for migration using \texttt{when()}. \texttt{when()}
-is part the Mantle API and the callback expects a return value of true or
-false, which corresponds to whether to migrate or not. For the Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} we set the \texttt{when()} policy to
-wait for load on the receiving server to fall below a threshold. At the
-beginning of the run the load of the receiving server spikes because requests
-for cache-coherence start as the sequencers are created. It takes 60 seconds
-for this load to drop to an acceptable level and it is at this point that the
-migration starts.  The Mantle curve in
+We control the conditions for when to migrate using \texttt{when()}, a callback
+in the Mantle API.  For the Mantle curve in
+Figure~\ref{fig:mantle-balancer-behaviors} we program \texttt{when()} to wait
+for load on the receiving server to fall below a threshold. This makes the
+balancer more conservative because it takes 60 seconds for cache coherence
+messages to settle.  The Mantle curve in
 Figure~\ref{fig:mantle-balancer-behaviors} also takes longer to reach peak
-throughput.  This conservative decision making is intentional because we want
-the policy wait to see how migrations affect the system before proceeding. The
-Mantle curve in Figure~\ref{fig:mantle-balancer-behaviors} does a migration
-right before 50 seconds, realizes that there is a third underloaded server, and
-does another migration.
+throughput because we want the policy to wait to see how migrations affect the
+system before proceeding; the balancer does a migration right before 50
+seconds, realizes that there is a third underloaded server, and does another
+migration.
 
@@
 migation.  Behavior graphs and performance numbers for this backoff feature is
-omitted for space considerations, but the throughput graphs are as expected:
-the more conservative the approach the less overall throughput.\\
+omitted for space considerations, but our experiments confirm that the more
+conservative the approach the less overall throughput.\\
 

### 2016-10-21 paper/paper.tex
Small mods to abstract

@@
 \begin{abstract}
-Storage systems are caught between rapidly changing data processing
-applications and the increasing speed of storage devices. This puts tremendous
-pressure on storage systems to support rapid evolution both in terms of their
+Storage systems are caught between the need to evolve data processing applications efficiently and quickly,
+and the increasing velocity with which storage device technology evolves. This puts tremendous
+pressure on storage systems to support rapid change both in terms of their
 interfaces and their performance. But adapting storage systems can be difficult
@@
 performance optimization efforts that were necessary for users to entrust their
-data to the storage system.  We introduce  Malacology, a prototype programmable
-storage system, to explore how existing abstractions of common storage system services 
-%found in
-%storage systems 
+data to the storage system.
+In this work, we introduce the concept of a programmable storage system, a system that exposes internal services and abstractions with the goal of re-using and building new services on top of the storage stack.
+We present Malacology, a prototype implementation of a programmable storage system based on Ceph
+to explore how existing abstractions of common storage system services 
 can be leveraged to 
-%address 
 adapt to the needs of
 new data processing systems and the
-increasing speed of storage devices. This approach allows a large degree of
+increasing speed of storage devices. We show how this approach allows a large degree of
 flexibility for storage systems to evolve without sacrificing the robustness of
@@
 \end{abstract}
+
+\begin{figure}[tb]
+\centering
+\includegraphics{figures/overview.png}
+\caption{Scalable storage systems have storage daemons which store data,
+monitor daemons (M) that maintain cluster state, and service-specific daemons
+({\it e.g.}, file system metadata servers). Malacology enables the programmability of
+internal abstractions (bold arrows) to re-use and compose
+existing subsystems.  With Malacology, we add new services, ZLog and
+Mantle, that sit alongside traditional user-facing APIs (file, block,
+object).}\label{fig:overview}
+\end{figure}
 
@@
 bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
-
-\begin{figure}[tb]
-\centering
-\includegraphics{figures/overview.png}
-\caption{Scalable storage systems have storage daemons which store data,
-monitor daemons (M) that maintain cluster state, and service-specific daemons
-({\it e.g.}, file system metadata servers). Malacology enables the programmability of
-internal abstractions (bold arrows) to re-use and compose
-existing subsystems.  With Malacology, we add new services, ZLog and
-Mantle, that sit alongside traditional user-facing APIs (file, block,
-object).}\label{fig:overview}
-\end{figure}
 
@@
 
-%{\bf Policy management.} A new interface is added to the metadata service
-%subsystem using the monitoring framework to support management of interfaces.
-%This ensures that the correct balancer version is consistently distributed and
-%verified on all MDS nodes. Policies are also durably saved with Ceph's
-%replication and redundancy.
-
-%These problems are general, distributed systems load balancing problems and
-%share many of the same properties as the POSIX metadata
-%wall~\cite{alam:pdsw2011-metadata-scaling,ghemawat:sosp2003-gfs,hildebrand:msst2005-pnfs,weil_ceph_2006,welch:fast2008-panasas,shvachko:login2012-hdfs-scalability}.
-%Many distributed file systems decouple metadata and data I/O so that these
-%services can scale independently. Despite this optimization, scaling the
-%metadata services is still difficult because metadata accesses impose small
-%and frequent requests on the underlying storage
-%system~\cite{roselli:atec2000-FS-workloads}. Many techniques for designing the
-%metadata services have been proposed to accommodate this workload:
-%Lustre~\cite{konstantinos:pdsw2014-lustre-metadata},
-%GFS~\cite{ghemawat:sosp2003-gfs}, and
-%HDFS~\cite{shvachko:login2012-hdfs-scalability} keep all metadata on one
-%server; GIGA+~\cite{patil:fast2011-giga}, IndexFS~\cite{ren:sc2014-indexfs},
-%Lazy Hybrid~\cite{brandt:msst2003-lh}, GPFS~\cite{schmuck:fast2002-gpfs}, and
-%pNFS~\cite{hildebrand:supercomputing2006-pNFS} hash the file system namespace
-%across a dedicated cluster and cache metadata;
-%Panasas~\cite{welch:fast2008-panasas} and
-%CephFS~\cite{weil:sc2004-dyn-metadata} use subtree partitioning to carve up
-%the namespace and distribute it across a cluster.
-%
-%Both the object and balancer interfaces define functions and attach them
-%to the core Lua sandbox. For example, the Lua object storage interface
-%class attaches data, extended attribute, object map, and object version
-%functions, to the Lua core functions, as shown in Figure
-%\ref{fig:cls-osd-mds}. The advantages of this is that we avoid
-%duplicating code, we provide a framework for putting Lua code in other
-%parts of the system, and we remove components and APIs that are too
-%integrated with the OSD.
-
-%To facilitate the use of Lua in both the MDS and OSD, we modularize the LuaVM
-%in a core sandbox wrapper and link against it, as shown in
-%Figure~\ref{fig:cls-osd-mds}. This core Lua wrapper contains just the
-%dependencies, symbols, and functions, needed to run the Lua VM and strips away
-%all the functionality inherent to object classes (like placement group
-%filters, cryptography functions, object metadata,  object data, object
-%extended attributes, and object versions). Some functions of Ceph's more
-%helpful subsystems, like the logging facilities for the daemons, were re-used
-%in the MDS.  With this scheme, The OSD will dynamically open the shared
-%library created by Lua object interface shared library while the MDS will
-%dynamically open the Lua balancer interface shared library; both shared
-%libraries will the Lua core.
-
-%{\bf File Types.} We leverage the same active and typed storage module
-%presented in DataMods~\cite{watkins_datamods_2012} to assign state and a set of
-%functions to a file (stored in the inode). This is valuable for
-%metadata load balancing so that certain types of files can have a higher
-%priority when making balancing decisions.
-
-%\begin{figure}[h]
-%\centering
-%\includegraphics{figures/cls-osd-mds.png}
-%\caption{To generalize the LuaVM for use in both the OSD and MDS,
-%Malacology detaches the core functionality into a Lua core module. The
-%VM is dynamically loaded first (1) and then interfaces can execute (2).
-%\label{fig:cls-osd-mds}}
-%\end{figure}
-
-% Straw man example
-%Migrating resources is an important part of load balancing but understanding
-%the effects of different configurations and parameters is difficult. Systems
-%already virtualize and migrate memory; soon systems will be able to migrate
-%other resources, like CPU, disk, and memory. Figuring out how, when, and where
-%to move these resources depends on complex metrics like utilization,
-%configuration, and workload.
-
-%These hooks embed Lua code in the MDS using the OSD subsystems. By re-using
-%the OSD interface classes, the MDS gets the safety and robustness of loading
-%dynamic code (including graceful failures and helpful errors), the ease of
-%transferring state between object interfaces and the OSD internals,
-%integration with testing and correctness suites, and \texttt{struct}s for
-%interface data and handlers.
-
 \subsection{Durability Interface}
@@
 these data structures, we use this service to automatically install interfaces
-in OSDs, and install policies within the MDS such that clients and daemons are
+in OSDs, and install policies within the metadata server daemon (MDS) such that clients and daemons are
 synchronized on correct implementations without restarting.
@@
 \noindent \texttt{\$> ceph fs\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}
-
-%There are Ceph components that are obstacles to this approach, namely the size
-%of the  MDS map, API changes, and security. The Ceph engineers want to avoid
-%using the maps to exchange information; the maps should be lightweight and fast
-%because the MONs need to be able to version MDS cluster state quickly. API
-%changes can also be a burden, especially for the MDS map, since all clients
-%need to have the same version of the CephFS library in order to properly
-%interact with CephFS. Finally, Ceph tries to avoid injecting strings into the
-%cluster because they are insecure and buggy.
 

### 2016-10-21 paper/paper.tex
paper: combine mantle bar graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-10-21 paper/paper.tex
@@
 \centering
+\includegraphics{figures/mantle-mode-performance.png}
 \includegraphics{figures/mantle-balancer-performance.png}
@@
 Another factor that affects performance in this environment is ``how much" load
-is on each server; these experiments quantify that effect by programming
-policies into the load balancing interface to control the amount of load to
-migrate. We define migration unit as the amount of load to try and migrate.
-For example, it is useful to express migration units as a fraction of the total
-load.  Expressing this heuristic is not easily achievable with outward facing
-tunable parameters (i.e. system knobs) and so the original CephFS balancers do
-not have a migration unit tunable.  With Mantle's programmable interface, we
-force the load balancer into the Proxy Mode (Full) scenario in
-Figure~\ref{fig:mantle-modes}, which sets the balancer to proxy mode with
-migration units equal to the entire load on the current server we can use the Mantle
-policy:\\
-
-\noindent \texttt{ targets[\#mds+1] = mds[whoami]["load"] }\\
+is on each server; these experiments quantify that effect by programming the
+load balancing interface to control the amount of load to migrate. We call this
+metric a ``migration unit".  Expressing this heuristic is not easily achievable
+with outward facing tunable parameters (i.e. system knobs) but with Mantle's
+programmable interface, we can force the load balancer to change its migration
+units. To force the balancer into the Proxy Mode (Half) scenario in
+Figure~\ref{fig:mantle-modes}, which uses migration units that are half the
+load on the current server:\\
+
+\noindent \texttt{ targets[whoami+1] = mds[whoami]["load"]/2 }\\
 
@@
 API to send half of the load on the current server (whoami) to the next ranked
-server (\#mds + 1); the \texttt{targets} array is a globally defined table that
-the balancer uses to do the migrations. Alternatively, to set the mode to the
-``Proxy (Half)"` setup in Figure~\ref{fig:mantle-modes} we can divide the right
-hand side by 2.
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-mode-performance.png}
-\caption{The performance of the different modes using various migration units
-shows almost a 2\(\times\) improvement in the best case.
-}\label{fig:mantle-mode-performance}
-\end{figure}
+server (whoami + 1); the \texttt{targets} array is a globally defined table that
+the balancer uses to do the migrations. Alternatively, to migrate all load a
+time step, we can remove the division by 2.
+
+%\begin{figure}[t!]
+%\centering
+%\includegraphics{figures/mantle-mode-performance.png}
+%\caption{The performance of the different modes using various migration units
+%shows almost a 2\(\times\) improvement in the best case.
+%}\label{fig:mantle-mode-performance}
+%\end{figure}
 

### 2016-10-21 paper/paper.tex
paper: combine graphs and captions
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \centering
-\includegraphics[width=75mm,trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
+\includegraphics[width=60mm,trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
 \caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
@@
 \centering
+\includegraphics{figures/mantle-balancer-performance.png}
 \includegraphics{figures/mantle-mode-performance.png}
-\includegraphics{figures/mantle-balancer-performance.png}
-\caption{All CephFS balancing modes have the same performance Mantle uses a
-balancer designed for sequencers.}\label{fig:mantle-balancer-performance}
+\caption{In (a) all CephFS balancing modes have the same performance; Mantle uses a
+balancer designed for sequencers. In (b) the best combination of mode and
+migration units can have up to a 2\(\times\) improvement.}\label{fig:mantle-balancer-performance}
 \end{figure}

### 2016-10-21 paper/paper.tex
paper: general shrinkage - squash deadline day commits - touch up 6.1 and fix captions - remove awkward duplicate active storage discussion - remove table I spent 4 months on - add citations for Lua, Paxos, chain replication - add carlos 2am feedback - consumed Domino's pizza with pineapple, pepperoni, cheese
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 \usepackage[font=footnotesize,labelfont=bf]{caption}
+
+\setlength{\belowcaptionskip}{-10pt}
 
@@
 \authorinfo{Paper ID: 165}
-           {Total Page: 14}
+           {Total Pages: 12}
 \date{}
@@
 
-\begin{abstract}
-Storage systems are caught between the need to evolve data processing applications efficiently and quickly,
-and the increasing velocity with which storage device technology evolves. This puts tremendous
-pressure on storage systems to support rapid change both in terms of their
-interfaces and their performance. But adapting storage systems can be difficult
-because unprincipled changes might jeopardize years of code-hardening and
-performance optimization efforts that were necessary for users to entrust their
-data to the storage system.
-In this work, we introduce the concept of a programmable storage system, a system that exposes internal services and abstractions with the goal of re-using and building new services on top of the storage stack.
-We present Malacology, a prototype implementation of a programmable storage system based on Ceph
-to explore how existing abstractions of common storage system services 
-can be leveraged to 
-adapt to the needs of
-new data processing systems and the
-increasing speed of storage devices. We show how this approach allows a large degree of
-flexibility for storage systems to evolve without sacrificing the robustness of
-their code-hardened subsystems. We illustrate the advantages and challenges of
-programmability by composing existing primitives into two new higher-level
-services: a file system metadata load balancer and a high-performance
-distributed shared-log that leverages flash devices. The evaluation
-demonstrates that our services inherit desirable qualities of the back-end
-storage system, including the ability to balance load, efficiently propagate
-cluster metadata, recover from failure, and to navigate trade-offs between
-latency and throughput using leases.
+\begin{abstract} Storage systems are caught between the need to evolve data
+processing applications efficiently and quickly, and the increasing velocity
+with which storage device technology evolves. This puts tremendous pressure on
+storage systems to support rapid change both in terms of their interfaces and
+their performance. But adapting storage systems can be difficult because
+unprincipled changes might jeopardize years of code-hardening and performance
+optimization efforts that were necessary for users to entrust their data to the
+storage system. We introduce the concept of a programmable storage system, a
+system that exposes internal services and abstractions with the goal of
+re-using and building new services on top of the storage stack. We present
+Malacology, a prototype implementation of a programmable storage system to
+explore how existing abstractions of common storage system services can be
+leveraged to adapt to the needs of new data processing systems and the
+increasing variety of storage devices. We show how this approach allows a large
+degree of flexibility for storage systems to evolve without sacrificing the
+robustness of their code-hardened subsystems. We illustrate the advantages and
+challenges of programmability by composing existing primitives into two new
+higher-level services: a metadata load balancer and a high-performance
+distributed shared-log. The evaluation demonstrates that our services inherit
+desirable qualities of the back-end storage system, including the ability to
+balance load, efficiently propagate cluster metadata, recover from failure, and
+to navigate trade-offs between latency and throughput using leases.
 \end{abstract}
@@
 can be used to \emph{program} new services. By composing higher-level services
-over existing abstraction one can reuse subsystems and leverage their
+over existing abstractions one can reuse subsystems and leverage their
 optimizations, established correctness, robustness, and efficiency.
 
-{\it \textbf{We define a programmable storage system}} to be a storage system
-that facilitates the re-use and extension of existing storage abstractions
-provided by the underlying software stack, to enable the creation of new
-services via composition.  A programmable storage system can be realized by
-exposing existing functionality (such as metadata services and synchronization
-and monitoring capabilities) as interfaces that can be ``glued together'' in a
-variety of ways using a high-level language. In contrast to programmable
-storage, \emph{active storage} differs through its focus on the injection and
-execution of code within a storage system or storage device. Programmable
-storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection and
-execution of code within a storage system or storage device---in
-that the
+{\it \textbf{Contribution 1:}} We define a programmable storage system to be a
+storage system that facilitates the re-use and extension of existing storage
+abstractions provided by the underlying software stack, to enable the creation
+of new services via composition.  A programmable storage system can be realized
+by exposing existing functionality (such as metadata services and
+synchronization and monitoring capabilities) as interfaces that can be ``glued
+together'' in a variety of ways using a high-level language. Programmable
+storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection
+and execution of code within a storage system or storage device---in that the
 former is applicable to any component of the storage system, while the latter
-focuses at the data access level (i.e. inject and execute arbitrary code at the
-I/O access layers). Given this contrast, we can say that active storage is an
-example of how one internal component (the storage layer) is exposed in a
-programmable storage system.
+focuses at the data access level. Given this contrast, we can say that active
+storage is an example of how one internal component (the storage layer) is
+exposed in a programmable storage system.
 
@@
 propagation of consistent cluster state provided by the monitoring service (see
-Figure~\ref{fig:overview}). As we will show in this paper, Malacology is
-expressive enough to provide the functionality necessary for implementing new
-services. 
-%Our contributions are:
-
-% prototype programmable storage system that 
+Figure~\ref{fig:overview}).  Malacology is expressive enough to provide the
+functionality necessary for implementing new services.
+
 Malacology includes a non-exhaustive set of interfaces that can be used as
@@
 \item An interface for \textbf{durability} that persists policies using the
-underyling storage stack's object store.
+underlying storage stack's object store.
 
@@
 
-{\it \textbf{We implement two distributed services}} using Malacology to
+{\it \textbf{Contribution 2:}} We implement two distributed services using Malacology to
 demonstrate the feasibility of the programmable storage approach:
@@
 motivate the need for programmable storage by describing current practices in
-the open source community. Next we describe Malacology by presenting the
+the open source software community. Next we describe Malacology by presenting the
 subsystems within the underlying storage system that we re-purpose, and briefly
@@
 Building storage stacks from the ground up for a specific purpose results in
-the best performance. For example, HDFS was designed specifically to serve
-Hadoop's jobs, and uses techniques like exposing data locality and relaxing
-POSIX constraints to to achieve application-specific I/O
-optimizations.~\cite{CITEME}. Alternatively, general-purpose storage stacks are
-built with the flexiblity to serve many applications by providing a variety of
-interfaces and tunable parameters. Unfortunately, managing competing forces in
-these systems is difficult and users want more control from the general-purpose
-storage stacks without going as far as building their storage system from the
-ground up.
-
-To demonstrate this trend in storage systems we examine the state of
-programmability in Ceph. Something of a storage swiss army knife, Ceph
-simultaneously supports file, block, and object interfaces on a single
-cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
-Distributed Object Storage (RADOS) system is a cluster of storage devices
-(OSDs) that provide Ceph with data durability and integrity using replication,
-erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already provides
-some degree of programmability; the OSDs support domain-specific object
-interfaces that are implemented by composing existing low-level storage
+the best performance. For example, GFS~\cite{ghemawat:sosp03} and
+HDFS~\cite{shvachko:msst10} were designed specifically to serve Map/Reduce and
+Hadoop jobs, and use techniques like exposing data locality and relaxing POSIX
+constraints to achieve application-specific I/O optimizations. Another example
+is Boxwood~\cite{maccormick:osdi04}, which experimented with B-trees
+and chunk stores as storage abstractions to simplify application building.
+Alternatively, general-purpose storage stacks are built with the flexibility to
+serve many applications by providing a variety of interfaces and tunable
+parameters.  Unfortunately, managing competing forces in these systems is
+difficult and users want more control from the general-purpose storage stacks
+without going as far as building their storage system from the ground up.
+
+To demonstrate a recent trend towards more application-specific storage systems
+we examine the state of programmability in Ceph. Something of a storage Swiss
+army knife, Ceph simultaneously supports file, block, and object interfaces on
+a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
+Distributed Object Storage (RADOS) system is a cluster of object storage
+devices (OSDs) that provide Ceph with data durability and integrity using
+replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already
+provides some degree of programmability; the OSDs support domain-specific
+object interfaces that are implemented by composing existing low-level storage
 interfaces that execute atomically. These interfaces are written in C++ and are
@@
 of an object extent).  What is most remarkable is that this trend contradicts
-the notion that API changes are a burden for users.  Rather it appears that a
-gap in existing interfaces are being addressed through ad-hoc approaches to
+the notion that API changes are a burden for users.  Rather it appears that
+gaps in existing interfaces are being addressed through ad-hoc approaches to
 programmability. In fact, Table~\ref{table:objclasses} categorizes existing
@@
 \includegraphics{figures/obj-int-dev-growth.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
-Since 2010, the growth in the number of co-designed object storage interfaces
+\caption{Since 2010, the growth in the number of co-designed object storage interfaces
 in Ceph has been accelerating. This plot is the number of object classes (a
@@
 \caption{A variety of object storage classes exist to expose interfaces
-    to applications. \# is the number of methods that use these categories.
+    to applications. \# is the number of methods that implement these categories.
 }
@@
 programmers the ability to tell the storage system about their application: if
-it is CPU or IO bound, if it has locality, if its size has the potential to
-overload a single proxy node, etc.  The programmers know what the problem is
+the application is CPU or IO bound, if it has locality, if its size has the potential to
+overload a single node, etc.  Programmers often know what the problem is
 and how to solve it, but until the ability to modify object interfaces, they
@@
 \item Policies and optimizations are usually hard-wired into the services and
-one has to be careful when factoring them out not to introduce additional bugs.
-These policies and optimiations are usually cross-cutting solutions to concerns
-or trade-offs that cannot be fully explored yet (as they relate to workload or
-hardware). Given these policies and optimizations, decomposition of otherwise
-orthogonal internal abstractions can be difficult or dangerous.
+one has to be careful when factoring them to avoid introducing additional bugs.
+These policies and optimizations are usually cross-cutting solutions to
+concerns or trade-offs that cannot be fully explored at the time the code is
+written (as they relate to workload or hardware). Given these policies and
+optimizations, decomposition of otherwise orthogonal internal abstractions can
+be difficult or dangerous.
 
@@
 
-Malacology implements and demonstrates the ideas of programmable storage in
-Ceph. We choose Ceph because it is a production quality system and because it
+The Malacology prototype we present has been implemented on Ceph.
+While there are other systems on top of which Malacology could be implemented (see Table~\ref{table:examples}), we choose Ceph because it is a production quality system and because it
 is open source. The large developer community ensures that code is
@@
 \begin{table*}
-\centering
+\centering\small
 \begin{tabular}{  l | l | l | l }
-\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
+%\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
 \multicolumn{4}{c}{} \\
@@
 {\it Service Metadata}
-  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,website:chubby}
+  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
   & cluster state management~\cite{website:ceph-mon}
@@
 \end{tabular}
-\caption{The same ``internal abstractions" are common across large-scale
+\caption{Common internal abstractions. The same ``internal abstractions" are common across large-scale
 systems because they provide primitives that solve general distributed systems
@@
 abstractions as interfaces that higher level applications can use; {\it
-italicized abstractions} are interfaces contributed in this paper.}
+italicized abstractions} are interfaces discussed in this paper.}
 \label{table:examples}
@@
 programming can occur without having to restart the storage system (see also
-below,~\S\ref{sec:durability}).
-
-There are multiple reasons that make Lua an attractive runtime for implementing
-Malacology. Lua is a portable, embedded scripting language that offers superior
-performance and productivity trade-offs, including a JIT-based implementation
-that is well known for near native performance. Additionally, Lua has been used
-extensively in game engines, and systems research \cite{neto:dls14-luaos},
-including storage systems where it has been effectively used both on
+below,~\S\ref{sec:durability}). Table~\ref{table:examples} shows the internal services from Ceph that we expose in the Malacology prototype via Lua~\cite{ierusalimschy1996lua} bindings.
+
+Lua is a portable embedded scripting language and we choose it as the
+interpreted language for Malacology because it offers superior performance and
+productivity trade-offs, including a JIT-based implementation that is well
+known for near native performance. Additionally, Lua has been used extensively
+in game engines, and systems research \cite{neto:dls14-luaos}, including
+storage systems where it has been effectively used both on
 \cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010}
@@
 flexibility of the runtime allows execution sandboxing in order to address
-security and performance concerns.
-
-We will now discuss the common subsystems used to manage storage system and how
-Malacology makes them programmable.
+security and performance concerns. We will now discuss the common subsystems
+used to manage storage system and how Malacology makes them programmable.
 
@@
 % straw man example
-Keeping track of the state of a distributed system is an essential part of any
+Keeping track of state in a distributed system is an essential part of any
 successful service and a necessary component in order to diagnose and detect
 failures, when they occur. This is further complicated by variable propagation
-delays and heterogenous hardware in highly dynamic environments.
+delays and heterogeneous hardware in dynamic environments.
 
@@
 and clients is critical to provide strong consistency guarantees to clients.
-Ceph maintains cluster state information in per-subsystem data structures
-called ``maps'' that record membership and status information.  A Paxos-based
-monitoring service is responsible for integrating state changes into cluster
-maps, responding to requests from out-of-date clients and synchronizing members
-of the cluster whenever there is a change in a map so that they all observe the
-same system state. As a fundamental building block of many system designs,
-consensus abstractions such as Paxos are a common technique for maintaining
-consistent data versions, and are a useful system to expose.
+Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
+metadata server) data structures called ``maps'' that record membership and
+status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
+responsible for integrating state changes into cluster maps, responding to
+requests from out-of-date clients and synchronizing members of the cluster
+whenever there is a change in a map so that they all observe the same system
+state. As a fundamental building block of many system designs, consensus
+abstractions such as Paxos are a common technique for maintaining consistent
+data versions, and are a useful system to expose.
 
@@
 clients to identify when new values (termed epochs in Ceph) are associated to
-given maps.  While Ceph does not expose this service directly, a key-value
+given maps.  Since Ceph does not expose this service directly, as part of our Malacology implementation, we expose a key-value
 service designed for managing configuration metadata that is built on top of
-the consensus engine is an example of a high-level service within Ceph that
-re-uses existing abstractions.  While a key-value service is generally useful,
-it doesn't provide many of the useful services hidden within the monitoring
-framework that would be required for applications with more demanding
-requirements, such as creating arbitrary maps with application-specific
-semantics and interfaces. Since the monitor is intended to be out of
-high-performance I/O paths, a general guideline is to make use of this
+the consensus engine. Since the monitor is intended to be out of the
+high-performance I/O path, a general guideline is to make use of this
 functionality infrequently and to assign small values to maps.
@@
 interface, such as authorization control (just specific clients can write new
-values) or to trigger actions based on specific values (e.g. sanitize values).
+values) or triggering actions based on specific values (e.g. sanitize values).
 The higher level services we implement in \S\ref{sec:services} make use of this
@@
 Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
-supports the creation of application-specific object
+supports application-specific object
 interfaces~\cite{weil_rados_2007}. The ability to offload computation can
@@
 
-An object interface is a plugin structured in a similar way to that of an RPC
+An object interface is a plugin structured similarly to that of an RPC
 in which a developer creates a named function within the cluster that clients
@@
 function input and output.  A wide range of native interfaces are available to
-developers such as reading and write to a byte stream, controlling object
+developers such as reading and writing to a byte stream, controlling object
 snapshots and clones, and accessing a sorted key-value database. These native
-object interfaces may be transactionally composed along with application
-specific logic to create interfaces with a rich set of semantics.  An example
-is the use of a read-modify-write operation to perform consistent object
-updates across the byte stream and key-value database interfaces without
-relying on complex client-driven concurrency control.
-
-\begin{figure}[htbp]
+interfaces may be transactionally composed along with application
+specific logic to create semantically rich interfaces. An example would be
+an interface that atomically updates a matrix stored in the bytestream and an
+index of the matrix stored in the key-value database.
+
+\begin{figure}[tbp]
 \centering
 \includegraphics{figures/implementation.png}
-\caption{Malacology re-uses Ceph subsystems to allow users to dynamically
-define new object/data interfaces and load balancing policies. It achieves this
-by extending the object storage daemon (OSD) and metadata server daemon (MDS)
-subsystems with an embedded Lua VM.  It uses the of the strongly-consistent
-service metadata interface to propogate the object interfaces and versions
-across the cluster.
-\label{fig:implementation}} \end{figure}
+\caption{Malacology allows users to dynamically define object/metadata
+interfaces by extending the object storage daemon (OSD) and metadata server
+(MDS) subsystems with an embedded Lua VM.  It uses the service metadata
+interface (M) to propagate interfaces/versions across the cluster
+\label{fig:implementation}}
+\end{figure}
 
@@
 Lua does not prevent deployment of malicious code, certain types of coding
-mistakes can handled gracefully, and access policies are used to restrict the
-feature to trusted users.
+mistakes can be handled gracefully, and access policies are used to limit access
+to trusted users~\cite{ierusalimschy1996lua}.
 
@@
 allowed to access as well as what state they may cache and modify locally.
-While designed for a fixed file abstraction, indexing, locking, and caching are
+While designed for the file abstraction, indexing, locking, and caching are
 all common services that are useful to a broad spectrum of applications.
@@
 metadata cluster. This cluster uses load balancing policies to migrate
-directory inodes around the cluster to alleviate overload at single nodes. The
-policies use metrics based on system state (e.g.  CPU and memory utilization)
-and statistics collected by the cluster (e.g. the popularity of an inode). Ceph
-uses dynamic subtree partitioning to move variable sized namespace
-subtrees~\cite{weil:sc2004-dyn-metadata}. These units can be shipped anywhere
-(i.e., to any metadata server of any capacity) at any time for any reason. The
-original balancer was designed with hard-coded policies and tunables.
+directory inodes around the cluster to alleviate overload at single
+nodes~\cite{weil:sc2004-dyn-metadata}. The policies use metrics based on system
+state (e.g.  CPU and memory utilization) and statistics collected by the
+cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
+partitioning to move variable sized namespace subtrees. These units can be
+shipped anywhere (i.e., to any metadata server of any capacity) at any time for
+any reason. The original balancer was designed with hard-coded policies and
+tunables.
 
@@
 
-\paragraph*{Malacology:} interfaces are added in strategic spots in the
+\paragraph*{Malacology:} Interfaces are added in strategic spots in the
 metadata service for guarding shared resources, defining new file types,
@@
 hierarchy where files and directories are represented as inode data structures
-and expose a POSIX file interface. Application that manage large amounts of
+and expose a POSIX file interface. Applications that manage large amounts of
 metadata (e.g. users or database snapshots) often require a naming service. We
@@
 domain-specific interfaces to inodes that may modify locking and capability
-policies. We will show this is used in Section~\ref{sec:seq} when we discuss a
+policies. We will show how this is used in Section~\ref{sec:seq} when we discuss a
 distributed shared-log built on Malacology.
@@
 been improved to safely load, manage, and persist its load balancing policies.
-Composing Mantle using the service metadata and durability interfaces, Mantle
+Composing Mantle with the service metadata and durability interfaces, Mantle
 can safely version balancer policies, save balancer policies in the backend
 object store and centralize warnings/errors. When combined with file types,
-load balancing policies can express a variety of rules such as multi-tentant
+load balancing policies can express a variety of rules such as multi-tenant
 and multi-application balancing.
@@
 Ceph provides storage by striping and replicating data across
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS uses
-many techniques to ensure that data is not corrupted or lost, such as erasure
-coding, replication, and data scrubbing.  Furthermore, many of these features
+RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
+protects data using common techniques such as erasure
+coding, replication, and scrubbing.  Furthermore, many of these features
 are implemented using a peer-to-peer protocol that allows object storage
@@
 re-shard data in the background in a process called placement group splitting
-in which OSDs communicate directly with each other to coverage on a new state.
+in which OSDs communicate directly with each other to cover a new state.
 In order to reduce load on the monitoring service, Ceph OSDs use a gossip
@@
 \label{services}
-\begin{table}
-\centering
-\begin{tabular}{  l | l | l }
-\textbf{Service}              &
-\textbf{Used to...}           &
-\textbf{Malacology Interface} \\ \hline
-Mantle & load balance metadata     & Load Balancing  \\
-       & agree on current balancer & Service Metadata\\
-       & control client write lock & Shared Resource \\
-       & persist policy logic      & Durability      \\ \hline
-ZLog   & implement CORFU           & Object \& Data  \\ 
-       & install ZLog interfaces   & Service Metadata\\
-       & persist log entries       & Durability      \\ \hline
-ZLog   & load balance sequencers   & Load Balancing  \\
-Seqr.  & manage access to log tail & Shared Resource \\ 
-\end{tabular}
-\caption{Mantle and ZLog are built using Malacology interfaces so they inherit the
-robustness of the subsystems of the underlying stack.}
-\label{table:implementation}
-\end{table}
+%\begin{table}
+%\centering\small
+%\begin{tabular}{  l | l | l }
+%\textbf{Service}              &
+%\textbf{Used to...}           &
+%\textbf{Malacology Interface} \\ \hline
+%Mantle & load balance metadata     & Load Balancing  \\
+%       & agree on current balancer & Service Metadata\\
+%       & control client write lock & Shared Resource \\
+%       & persist policy logic      & Durability      \\ \hline
+%ZLog   & implement CORFU           & Object \& Data  \\ 
+%       & install ZLog interfaces   & Service Metadata\\
+%       & persist log entries       & Durability      \\ \hline
+%ZLog   & load balance sequencers   & Load Balancing  \\
+%Seqr.  & manage access to log tail & Shared Resource \\ 
+%\end{tabular}
+%\caption{Mantle and ZLog are built using Malacology interfaces so they inherit the
+%robustness of the subsystems of the underlying stack.}
+%\label{table:implementation}
+%\end{table}
 
@@
 service interface provided by the monitor daemons and (2) the durability and
-distribution provided by Ceph's reliable object store.
+distribution provided by Ceph's reliable object store.  Re-using the internal
+abstractions with Malacology resulted in a 2\(\times\) reduction in source code
+compared to the original implementation.
 
@@
 Ensuring that the version of the current load balancer is consistent across the
-physical servers in the metadata cluster was ignored in the original
+physical servers in the metadata cluster was not addressed in the original
 implementation. The user had to set the version on each individual server and
-it was trivial to make the version inconsistent. Maintaining a consistent
-version is important for balancing policies that are cooperative in which local
+it was trivial to make the versions inconsistent. Maintaining consistent
+versions is important for balancing policies that are cooperative in which local
 decisions are made assuming properties about other instances of the same
@@
 a result, the policy version inherits the consistency of Ceph's internal
-monitor daemons.
-
-The user changes the version of the load balancer using a new CLI command: \\
-
-\noindent \texttt{\$> ceph fs\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}
+monitor daemons. The user changes the version of the load balancer using a new
+CLI command.
 
@@
 with a timeout: if the asynchronous read does not come back within half the
-balancing tick interval the operation is cancelled and a Connection Timeout
+balancing tick interval the operation is canceled and a Connection Timeout
 error is returned. By default, the balancing tick interval is 10 seconds, so
@@
 
-Storing the balancers in RADOS is only possible because we use Lua as the
+Storing the balancers in RADOS is simplified by the use of an interpreted
 language for writing balancer code. If we used a language that needs to be
@@
 
-With Malacology, Mantle re-usues the centralized logging features of the
+With Malacology, Mantle re-uses the centralized logging features of the
 monitoring service. Important errors, warnings, and info messages are collected
@@
 distributed shared-log that is based on the CORFU
-protocol~\cite{balakrishnan_corfu_2012}. A shared-log is a general, yet
+protocol~\cite{balakrishnan_corfu_2012}. The shared-log is a
 powerful abstraction used to construct distributed systems, such as
-cloud-based metadata management~\cite{balakrishnan:sosp13} and elastic database
-storage engines~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
+metadata management~\cite{balakrishnan:sosp13} and elastic database
+systems~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
 However, existing implementations that rely on consensus algorithms such as
@@
 throughput. In contrast, the CORFU protocol is able to achieve high throughput
-using a network counter service, called a \emph{sequencer}, that is used to
-decouple log position assignment from log I/O.
+using a network counter called a \emph{sequencer}, that
+decouples log position assignment from log I/O.
 
@@
 log positions to clients by reading from a volatile, in-memory counter which
-can run at a very high throughput and at low latency. Since the sequencer
-is centralized, ensuring serializability in the common case is trivial.  The
+can run at a very high throughput and at low latency. Since the sequencer is
+centralized, ensuring serializability in the common case is trivial.  The
 primary challenge in CORFU is handling the failure of the sequencer in a way
 that preserves correctness. Failure of the sequencer service in CORFU is
-handled by a recovery algorithm that recomputes the new sequencer state using
-an application-specific custom storage interface to discover the tail of the
-log, while simultaneously invalidating stale client requests using an
-epoch-based protocol.
+handled by a recovery algorithm that recomputes the new sequencer state using a
+CORFU-specific custom storage interface to discover the tail of the log, while
+simultaneously invalidating stale client requests using an epoch-based
+protocol.
 
 {\bf Sequencer interface.} The sequencer resource supports the ability to
-\emph{read()} the current tail value as well as get the \emph{next()} position in
+\emph{read()} the current tail value and get the \emph{next()} position in
 the log which also atomically increments the tail position.
 We implement the sequencer service in Malacology as an application-specific
-file type that associates a small amount of metadata, including the 64-bit log
-tail position, to the inode state managed by the Ceph metadata service. Since
-a sequencer is associated with a unique log, this approach has the added
-benefit of allowing the metadata service to also handle naming, by
-representing each log service instance within the standard POSIX hierarchical
+file type that associates a small amount of metadata
+with the inode state managed by the Ceph metadata service.
+This approach has the added
+benefit of allowing the metadata service to handle naming, by
+representing each sequencer instance in the standard POSIX hierarchical
 namespace. The primary challenge in mapping the sequencer resource to the
@@
 sequencer locally. Any competing client cannot query the sequencer until the
-metadata service has granted it exclusive access.
+metadata service has granted it access.
 
@@
 
-In Section~\ref{evaluation} we quantify the trade-offs of throughput and
+In Section~\ref{sec:evaluation} we quantify the trade-offs of throughput and
 latency for an approach based on a round-robin batching mode, and compare this
 mode to one in which the metadata server mediates access to the sequencer state
-when in it is being shared among multiple clients. Quantifying these trade-offs
+when it is being shared among multiple clients. Quantifying these trade-offs
 should provide administrators with guidelines for setting the tunables for
@@
 
-{\bf Balancing policies.}
-As opposed to the batching mode for controlling access to the sequencer
+{\bf Balancing policies.} As opposed to the batching mode for controlling access to the sequencer
 resource, more predictable latency can be achieved by treating the sequencer
@@
 metadata service load balancing to construct a service-specific load balancing
-policy. As opposed to a generic balancing policy that may strive for uniform
+policy. As opposed to a balancing policy that strives for uniform
 load distribution, a ZLog-specific policy may utilize knowledge of inode types
@@
 
-Since the sequencer servivce does not resume until the recovery process has
+Since the sequencer service does not resume until the recovery process has
 completed, there cannot be a race with clients appending to the log, and the
@@
 leader election using an auxiliary service such as Paxos. In our
-implementation, the recovery is the same (and is inherited from) the Ceph
+implementation, the recovery is the same as (and is inherited from) the Ceph
 metadata service. Handling the failure of a client that holds the sequencer
@@
 
-Since this work focuses on the programmabilty of Malacology, the goal of this
+Since this work focuses on the programmability of Malacology, the goal of this
 section is to show that the components and subsystems that support the
@@
 resource as an inode, and (2) the monitoring sub-subsystem to distribute and
-install application-specific I/O interface required of the CORFU protcool. In
+install application-specific I/O interface required of the CORFU protocol. In
 Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the inode
@@
 We evaluate the feasibility of using the metadata service to implement a
-seqeuencer resource that is responsible for maintaining a total ordering of
+sequencer resource that is responsible for maintaining a total ordering of
 the log. Clients contact the sequencer to obtain the tail of the log and then
 independently initiate I/O, thus we measure both the throughput and latency of
-obtaining new tail positions which inherently bound client append performance.
+obtaining new tail positions which bounds client append performance.
 
@@
 represent points in time that the clients were able to access the resource.
-The best-effort policy provides interleaving a high degree of interleaving
+The best-effort policy shows a high degree of interleaving
 between clients but the system spends a large portion of time re-distributing
@@
 
-\begin{figure}[h]
+\begin{figure}[tbp]
 \centering
@@
 \includegraphics{figures/capdelay-quota-behavior-c.png}
-\caption{Each dot in these graphs is an individual request and they are spread
-randomly along the \(y\) axis. The default behavior (``None") shows
-unpredictable performance, ``delay" lets each client hold the lease longer, and
-``quota" gives each client the lease for a certain number of operations. Delay
-improves latency and fairness while quota maximizes throughput.}
+\caption{Each dot is an individual request, spread randomly along the \(y\)
+axis. The default behavior is unpredictable, ``delay" lets clients hold the
+lease longer, and ``quota" gives clients the lease for a number of operations.}
 \label{fig:capdelay-quota-behavior}
@@
 while a large quota reservation allows clients to experience a much lower
-latency because they uncontended access for a longer period of time.
-
-\begin{figure}[h]
+latency because they experience uncontended access for a longer period of time.
+
+\begin{figure}[tbp]
 \centering
 \includegraphics{figures/tradeoff.png}
-\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
-Sequencer throughput by re-using various services.  The highest performance is
+\caption{Sequencer throughput by re-using various services.  The highest performance is
 achieved using a single client with exclusive, cacheable privilege. Round-robin
@@
 The latency distribution shows that if the clients hold their capabilities longer,
-performance decreases. Changing the delay via policy can help the sequencer and
+performance decreases. Changing the delay can help the sequencer and
 clients strike a balance between throughput and latency.}
@@
 
-We have shown how exposing the internal capability management service allows
-users to navigate latency and throughput tradeoffs.  Other approaches to
-designing the sequencer service also exist, such as using a centralized
-service in which each access is a network round-trip. In contrast to the
-mechanism we explored which is appropriate for clients with bursty workloads,
-it may be easier to provide predictable performance using a centralized
-service, and we will be exploring in the future how this can be achieved using
-the capability system.
+\textbf{Malacology exposes} the internal capability management service and
+allows users to navigate latency and throughput trade-offs.  Other approaches to
+designing the sequencer service also exist, such as using a centralized service
+in which each access is a network round-trip. In contrast to the mechanism we
+explored which is appropriate for clients with bursty workloads, it may be
+easier to provide predictable performance using a centralized service, and we
+will be exploring in the future how this can be achieved using the capability
+system.
 
@@
 
-Domain-specific data interfaces (Section~\ref{object-data-interfaces}) allow
+Domain-specific data interfaces (Section~\ref{sec:application-specifc-storage-stacks}) allow
 co-design between applications and the storage system. Malacology supports
@@
 interfaces through automatic system-wide versioning and installation through
-the service metdata interface (Section~\ref{service-metadata}). We evaluate
-the performance of installing a new version of an interface in the cluster,
-which is an important metric for applications that are designed with
-frequently evolving interfaces.
+the service metadata interface (Section~\ref{service-metadata}). We evaluate
+the performance of installing a new interface version in the cluster,
+which is an important metric for applications that frequently evolve
+interfaces.
 
@@
 Figure~\ref{fig:propdelay} shows the lower bound cost for updates in a large
-cluster. In the experiment labeled "120 OSD (RAM)" a cluster of 120 OSDs (10
+cluster. In the experiment labeled ``120 OSD (RAM)" a cluster of 120 OSDs (10
 OSDs x 12 servers) using an in-memory data store were deployed, showing a
 latency of less than 54 ms with a probability of 90\% and a worst case latency
-of 194 ms. These costs demonstrate the penality of distributing the interface
+of 194 ms. These costs demonstrate the penalty of distributing the interface
 in a large cluster. In practice the costs include, in addition to cluster-wide
@@
 independent totally-ordered logs for each application.  For this scenario
-co-locating sequencers on the same physical node is not ideal. Building a load
-balancer that can migrate the shared resource (in this case, the resource that
-mediates access to the tail of the log) is a time-consuming, non-trivial task
-that requires building subsystems for migrating resources, monitoring the
+co-locating sequencers on the same physical node is not ideal but building a load
+balancer that can migrate the shared resource ({\it e.g.,}, the resource that
+mediates access to the tail of the log) is a time-consuming, non-trivial task.
+It requires building subsystems for migrating resources, monitoring the
 workloads, collecting metrics that describe the utilization on the physical
@@
 multiple sequencers. The following experiments demonstrate the feasibility of
-using the mechanisms of the Malacology load balancing interface to inherit all
-these features and to alleviate load from overloaded servers housing multiple
-sequencers.
+using the mechanisms of the Malacology load balancing interface to inherit
+these features and to alleviate load from overloaded servers.
 
@@
 The experiments are run on a cluster with 10 nodes to store objects, one node
-to monitor the cluster, and 3 nodes that can accomodate sequencers. Our first
-experiment co-locates 3 sequencers each with 4 clients on the same physical
-server.  Instead of measuring contention at the clients like
+to monitor the cluster, and 3 nodes that can accommodate sequencers. 
+Instead of measuring contention at the clients like
 Section~\ref{sec:sequencer-implementation}, these experiments measure
@@
 
-First, we show how the ZLog service can orchestrate multiple sequencers in a
-cluster using the Malacology load balancing interface.
-Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput (\(y\) axis)
-over time (\(x\) axis) of different load balancers that migrate ZLog sequencers
-acround the cluster; ``No Balancing" keeps all sequencers on one server,
-``CephFS" migrates sequencers using the CephFS load balancers, and ``Mantle"
-uses a custom load balancer we wrote specifically for sequencers. The CephFS
-load balancers are hard-coded into Ceph while Mantle is an API for
-balancing load presented in~Mantle~\cite{sevilla:sc15-mantle} and
-re-implemented on top of Malacology in this paper.  The increased throughput
-for the CephFS and Mantle curves between 0 and 60 seconds are a result of
-migrating the sequencer(s) off overloadeed servers.
+First, we show how the ZLog service can orchestrate multiple sequencers using
+the Malacology load balancing interface.
+Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput over time of
+different load balancers as they migrate 3 sequencers (with 4 clients) around
+the cluster; ``No Balancing" keeps all sequencers on one server, ``CephFS"
+migrates sequencers using the CephFS load balancers, and ``Mantle" uses a
+custom load balancer we wrote specifically for sequencers. The CephFS load
+balancers are hard-coded into Ceph while Mantle is an API for balancing load
+presented in~Mantle~\cite{sevilla:sc15-mantle} and re-implemented on top of
+Malacology in this paper.  The increased throughput for the CephFS and Mantle
+curves between 0 and 60 seconds are a result of migrating the sequencer(s) off
+overloaded servers.
 
@@
 different modes all have the same performance, shown in
-Figure~\ref{fig:mantle-balancer-performance}, because the load balancer falls
+Figure~\ref{fig:mantle-balancer-performance} (a), because the load balancer falls
 into the same mode a majority of the time.  The high variation in performance
@@
 dynamic and unpredictable as CPU utilization to make migration decisions. In
-addition to the suboptimal performance and unpredictability, the fact that the
+addition to the suboptimal performance and unpredictability, because
 CephFS balancers all fall into the same mode despite using different balancers
-is a problem becuase it prevents administrators from properly exploring the
+is a problem because it prevents administrators from properly exploring the
 balancing state space.
 
-Mantle gives the administrator more control over the balancing policies; for
-the Mantle bar in Figure~\ref{fig:mantle-balancer-performance} we use the load
+Mantle gives the administrator more control over balancing policies; for
+the Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the load
 balancing interface to program logic for balancing read-heavy workloads,
-resulting in higher throughput and better stability.  When we did this we also
-identified two balancing modes that are relevant for making migration decisions
+resulting in better throughput and stability.  When we did this we also
+identified two balancing modes relevant for making migration decisions
 for sequencers.
@@
 server B. The two ``Proxy mode" diagrams on the right show how clients continue
-sending reqeusts to server A even though sequencer 2 has migrated to server B
+sending requests to server A even though sequencer 2 has migrated to server B
 (the meaning of ``(Half)" is discussed in the next section).
@@
 clients.  Doing both steps is too heavy weight for one server and sequencers on
-slave nodes can go faster because work is split up; this phenomenon is not
-uncommon and has been observed in chain replication~\cite{CITEME}.  Cluster
+slave nodes can go faster if work is split up; this phenomenon is not
+uncommon and has been observed in chain replication~\cite{chain_rep}.
+
+Cluster
 throughput improves at the cost of decreased throughput for Sequencer 2.
@@
 
-Another factor that affects performance in this environment is ``how much" load
+Another factor that affects performance in this environment is how much load
 is on each server; these experiments quantify that effect by programming the
@@
 units. To force the balancer into the Proxy Mode (Half) scenario in
-Figure~\ref{fig:mantle-modes}, which uses migration units that are half the
-load on the current server:\\
-
-\noindent \texttt{ targets[whoami+1] = mds[whoami]["load"]/2 }\\
+Figure~\ref{fig:mantle-modes}, which uses migration units equal to half the
+load on the current server, we can use: \texttt{ targets[whoami+1] = mds[whoami]["load"]/2
+}.
 
@@
 
-Figure~\ref{fig:mantle-mode-performance} shows the performance of the modes
+Figure~\ref{fig:mantle-balancer-performance} (b) shows the performance of the modes
 using different migration units. Recall that this setup only has 2 sequencers
@@
 The other way to change aggressiveness of the decision making is to program
-into the balancer a threshold for sustatined overload. This forces the balancer
+into the balancer a threshold for sustained overload. This forces the balancer
 to wait a certain number of iterations after a migration before proceeding. In
 Mantle, the policy would use the save state function to do a countdown after a
-migation.  Behavior graphs and performance numbers for this backoff feature is
+migration.  Behavior graphs and performance numbers for this backoff feature is
 omitted for space considerations, but our experiments confirm that the more
-conservative the approach the less overall throughput.\\
-
-\noindent\textbf{Malacology pulls} the load balancing out of the storage system
-to balance sequencers across a cluster. Compared to putting all sequencers on
-the same server, load balancing a sequencer using the proxy mode shows a
-1.8\(\times\) throughput improvement. This latent capability also gives future
-programmers the ability to explore the different load balancing trade-offs
-including: load balancing modes to control forwarding vs. client redirection,
-load migration units to control sequencer distribution vs. co-location, and
-backoffs to control conservative vs. aggressive decision making.
+conservative the approach the less overall throughput.
+
+\textbf{Malacology pulls} the load balancing service out of the storage system
+to balance sequencers across a cluster. This latent capability also gives
+future programmers the ability to explore the different load balancing
+trade-offs including: load balancing modes to control forwarding vs. client
+redirection, load migration units to control sequencer distribution vs.
+co-location, and backoffs to control conservative vs. aggressive decision
+making.
 
@@
 distributed storage systems is not new, but we are not aware of work that makes
-generalization of existing services into programmable resourses a key principle
+generalization of existing services into programmable resources a key principle
 in storage systems design.
@@
 Another view of programmable storage systems is one of tailoring systems
-resources to applications~\cite{arpaci:sosp01}. Related work includes work
-around the Exokernel~\cite{engler:sosp95} and SPIN~\cite{bershad:sosp95} and
-Vino~\cite{seltzer:osdi96}, the latter two addressed the ability of injecting
+resources to applications~\cite{arpaci:sosp01}. Related efforts include the
+Exokernel~\cite{engler:sosp95}, SPIN~\cite{bershad:sosp95} and
+Vino~\cite{seltzer:osdi96} projects; the latter two addressed the ability of injecting
 code into the kernel to specialize resource management. Another approach is to
@@
 
-\section{Conclusion and Future Work}\label{conclusion-and-future-work}
+\section{Conclusion and Future Work}
+\label{conclusion-and-future-work}
 
 Programmable storage is a viable method for eliminating duplication of complex
-error prone software that are used as workarounds for storage system
-deficiencies. However, this duplication has real-world problems related to
-reliability. We propose that systems expose their services in a safe way
+error-prone software used as workarounds for storage system
+deficiencies. We propose that systems expose their services in a safe way
 allowing application developers to customize system behavior to meet their

### 2017-01-30 paper/paper.tex
paper: add implementation overview figure
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-The guiding principle
-is to re-use existing services and extend them so that these services can be
-\emph{programmed}. We accomplish programmability of a service by exporting
-bindings (or ``hooks'') for an interpreted programming language so that
-programming can occur without having to restart the storage system (see also
-below,~\S\ref{sec:durability}). Table~\ref{table:examples} shows the internal services from Ceph that we expose in the Malacology prototype via Lua~\cite{ierusalimschy1996lua} bindings.
+\begin{figure}[tbp]
+\centering
+\includegraphics{figures/implementation-overview.png}
+\caption{Malacology is implemented on the daemons and clients that run in a
+Ceph cluster. The gray callout bubbles show where each interface is in the
+storage stack and the Ceph modules, functions, and/or variables we modified.
+\label{fig:implementation-overview}}
+\end{figure}
+
+The guiding principle is to re-use existing services and extend them so that
+these services can be \emph{programmed}. We accomplish programmability of a
+service by exporting bindings (or ``hooks'') for an interpreted programming
+language so that programming can occur without having to restart the storage
+system (see also below,~\S\ref{sec:durability}). Table~\ref{table:examples}
+shows the internal services from Ceph that we expose in the Malacology
+prototype via Lua~\cite{ierusalimschy1996lua} bindings and
+Figure~\ref{fig:implementation-overview} shows where each interface lives in
+the storage stack.
 

### 2017-02-07 paper/paper.tex
paper: integrate design and implementation sections
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-10-21 paper/paper.tex
@@
 \usepackage{color}
+\usepackage{comment}
+\includecomment{newcomment}
+\includecomment{oldcomment}
 \newcommand{\cmark}{\ding{51}}%
@@
 \newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
+\usepackage[normalem]{ulem}
+\newcommand{\nc}[1]{{
+  \textcolor{red}{\textbf{#1}}
+}}
+\newcommand{\oc}[1]{{
+  \textcolor{blue}{\textbf{\sout{#1}}}
+}}
+
 \usepackage{pstricks}
@@
 
+\begin{newcomment}
+
+\nc{New changes are in red}
+
+\oc{Old prose is in blue}
+
+%\nc{
+%\begin{enumerate}
+%  \item clarify what was implemented: diagram of interfaces and prose specifying the difference between Malacology and prior work (reviewers B, C, D, E)
+%  \item describe what makes a programmable storage system (reviewers C, D)
+%  \item organize motivation: group interfaces into categories (e.g., core functionality, features, and performance optimizations) and describe other applications we can build with Malacology (reviewers B, C)
+%  \item clarify terminology: service vs. file system metadata, storage vs. file system, dirty vs. clean slate approaches, soft-state caching (reviewers B, C)
+%  \item address security, access control, and how it affects composability/multi-tenancy on correctness (reviewer D)
+%\end{enumerate}
+%}
+\end{newcomment}
+
 A storage system implements abstractions designed to persistently store data
@@
 
-To address these challenges we present Malacology. Malacology is both a
-prototype for a programmable storage system and a design approach to evolve
+To address these challenges we present Malacology
+\begin{oldcomment}
+\oc{. Malacology is both a prototype for a programmable storage system and a}
+\end{oldcomment}
+\begin{newcomment}
+\nc{, our prototype programmable storage system. It uses the programmable storage}
+\end{newcomment}
+design approach to evolve
 storage systems efficiently and without jeopardizing years of code-hardening
@@
 \multicolumn{4}{c}{} \\
-\textbf{Abstraction}                    &
+\textbf{Interface}                      &
 \textbf{Example in Production Systems}  &
@@
 
-The guiding principle
-is to re-use existing services and extend them so that these services can be
-\emph{programmed}. We accomplish programmability of a service by exporting
-bindings (or ``hooks'') for an interpreted programming language so that
-programming can occur without having to restart the storage system (see also
-below,~\S\ref{sec:durability}). Table~\ref{table:examples} shows the internal services from Ceph that we expose in the Malacology prototype via Lua~\cite{ierusalimschy1996lua} bindings.
+The guiding principle is to re-use existing services and extend them so that
+these services can be \emph{programmed}. We accomplish programmability of a
+service by exporting bindings (or ``hooks'') for an interpreted programming
+language so that programming can occur without having to restart the storage
+system (see also below,~\S\ref{sec:durability}). Table~\ref{table:examples}
+shows the internal services from Ceph that we expose in the Malacology
+prototype via Lua~\cite{ierusalimschy1996lua} bindings
+\begin{newcomment}
+\nc{and Figure~\ref{fig:implementation-overview} compares what was already
+present in Ceph (gray boxes) to the Malacology interfaces we added (blue
+trapezoids); Section~\ref{sec:services} will describe the higher-level
+services (pink arrows) we built with these interfaces.}
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=100mm]{figures/implementation-overview.png}
+\caption{\nc{Malacology is implemented on the daemons and clients that run in a
+Ceph cluster. The interfaces expose internal subsystems of the Ceph modules and
+are used as building blocks for higher-level services.}
+\label{fig:implementation-overview}}
+\end{figure}
+\end{newcomment}
 
@@
 
+\begin{newcomment}
+  \nc{\\ \noindent\it{\textbf{Ceph Implementation: }}}
+\end{newcomment}
 In the case of Ceph, a consistent view of cluster state among server daemons
@@
 
+\begin{newcomment}
+  \nc{\\ \noindent\it{\textbf{Ceph Implementation: }}}
+\end{newcomment}
 An object interface is a plugin structured similarly to that of an RPC
@@
 
-The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction~\cite{weil:sc2004-dyn-metadata}. In general, distributed
-file systems protect resources by providing hierarchical indexing and
-distributed locking services.  In Ceph, the locking service implements a
+\begin{oldcomment}
+\oc{The distributed metadata service in Ceph provides clients with a POSIX file
+system abstraction~\cite{weil:sc2004-dyn-metadata}.}
+\end{oldcomment}
+\begin{newcomment}
+\nc{File systems provide clients with the familiar POSIX file abstraction.
+While this guarantees strong consistency it comes at the cost of scalability
+and lower performance.}
+\end{newcomment}
+In general, distributed file systems protect resources by providing
+hierarchical indexing and distributed locking services.
+
+\begin{newcomment}
+\nc{\\ \noindent\it{\textbf{Ceph Implementation: }}}
+\end{newcomment}
+In Ceph, the locking service implements a
 capability-based system that expresses what data and metadata clients are

### 2017-02-09 paper/paper.tex
paper: address issues 1, 2, 3 from reviewers
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-10-21 paper/paper.tex
@@
 \usepackage{color}
+\usepackage{comment}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Uncomment for diff
+\includecomment{notes}
+\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
+\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
+\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
+
+% Uncomment for camera-ready
+%\excludecomment{notes}
+%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
+%\newcommand{\oldcomment}[1]{}
+%\newcommand{\addressesissue}[1]{}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \newcommand{\cmark}{\ding{51}}%
@@
 \newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
+\usepackage[normalem]{ulem}
 \usepackage{pstricks}
@@
             bookmarks=true,
-            pdfauthor={},
+            pdfauthor={Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, Peter Alvaro, Shel Finkelstein, Jeff LeFevre},
             pdftitle={Malacology: A Programmable Storage System},
@@
 \title{Malacology: A Programmable Storage System}
-\authorinfo{Paper ID: 165}
-           {Total Pages: 12}
+%\authorinfo{Paper ID: 165}          {Total Pages: 12}
+\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre}
+           {\{msevilla, jayhawk, ivo, cmaltzahn\}@soe.ucsc.edu, \{palvaro, finkelstein, jlefevre\}@ucsc.edu}
 \date{}
@@
 \maketitle
+
+\begin{notes}
+\textcolor{red}{
+\noindent Issues raised by reviewers:
+\begin{enumerate}
+  \item clarify what was implemented: diagram of interfaces and prose
+  specifying the difference between Malacology and prior work (reviewers B, C,
+  D, E)
+  \item describe what makes a programmable storage system (reviewers C, D)
+  \item organize motivation: group interfaces into categories (e.g., core
+  functionality, features, and performance optimizations) and describe other
+  applications we can build with Malacology (reviewers B, C)
+  \item clarify terminology: service vs. file system metadata, storage vs.
+  file system, dirty vs. clean slate approaches, soft-state caching (reviewers
+  B, C)
+  \item address security, access control, and the safety; specify how it
+  affects composability, performance, and multi-tenancy (reviewer D)
+\end{enumerate}
+}
+\end{notes}
 
@@
 
-To address these challenges we present Malacology. Malacology is both a
-prototype for a programmable storage system and a design approach to evolve
-storage systems efficiently and without jeopardizing years of code-hardening
-and performance optimization efforts.  Although Malacology uses the internal
-abstractions of the underlying storage system, including its subsystems,
-components, and implementations, we emphasize that our system still addresses
-the general challenges outlined above.
+To address these challenges we present Malacology
+\oldcomment{. Malacology is both a prototype for a programmable storage system
+and a}
+\newcomment{, our prototype programmable storage system. It uses the
+programmable storage}
+\addressesissue{2}
+design approach to evolve storage systems efficiently and
+without jeopardizing years of code-hardening and performance optimization
+efforts.  Although Malacology uses the internal abstractions of the underlying
+storage system, including its subsystems, components, and implementations, we
+emphasize that our system still addresses the general challenges outlined
+above.
 
@@
 right internal abstractions and picking the correct layers for exposing them.
-In this paper, we do not present an exhaustive list of possible internal
-abstractions nor do we contend that the abstractions we choose provide the best
-trade-offs for all applications.  For example, if consensus is correctly
-exposed one could implement high-level features like versioning, serialization,
-or various flavors of strongly consistent data management on top; but perhaps a
-low-level consensus interface is suited well for a particular set of
-applications.  These questions are not answered in this paper and instead we
-focus on showing the feasibility of building such a system, given advances in
-the quality and robustness of today's storage stacks.
+\oldcomment{In this paper, we do not present an exhaustive list of possible
+internal abstractions nor do we contend that the abstractions we choose provide
+the best trade-offs for all applications.} 
+\newcomment{A programmable storage is not defined by what abstractions are
+exposed, rather a programmable storage system adheres to the design approach of
+choosing interfaces so administrators can have better control of the storage
+system. The interfaces presented in this paper are abstractions that we found
+useful for building ZLog and Mantle, yet they may not provide the best
+trade-offs for all higher-level services.}
+\addressesissue{2}
+For example, if consensus is correctly exposed one could implement
+high-level features like versioning, serialization, or various flavors of
+strongly consistent data management on top; but perhaps a low-level consensus
+interface is suited well for a particular set of applications.  These questions
+are not answered in this paper and instead we focus on showing the feasibility
+of building such a system, given advances in the quality and robustness of
+today's storage stacks.
 
@@
 \centering\small
-\begin{tabular}{  l | l | l | l }
+\begin{tabular}{ l | c | l | l | l }
 %\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
 \multicolumn{4}{c}{} \\
-\textbf{Abstraction}                    &
+\textbf{Interface}                      &
+\textbf{Sec}                            &
 \textbf{Example in Production Systems}  &
@@
 \textbf{Provided Primitives}            \\ \hline
-{\it Service Metadata}
+Service Metadata
+  & \S\ref{sec:service-metadata-interface}
   & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
   & cluster state management~\cite{website:ceph-mon}
-  & consensus \& consistency
+  & consensus/consistency
   \\
-{\it Shared Resource}
+Data I/O
+  & \S\ref{sec:data-io-interface}
+  & Swift in situ storage/compute~\cite{website:zerocloud}
+  & object interface classes~\cite{website:cls-lua}
+  & transaction/atomicity
+  \\
+Shared Resource
+  & \S\ref{sec:distributed-metadata-interfaces}
   & MPI collective IO, burst
   & POSIX metadata protocols
-  & serialization \& batching
+  & serialization/batching
   \\
-{\it Object \& Data}
-  & Swift in situ storage/compute~\cite{website:zerocloud}
-  & object interface classes~\cite{website:cls-lua}
-  & transaction \& atomicity
+File Type
+  & \S\ref{sec:distributed-metadata-interfaces}
+  & ?????
+  & file striping strategy
+  & data/metadata access
+  \\
+Load Balancing
+  & \S\ref{sec:distributed-metadata-interfaces}
+  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
+  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
+  & migration/sampling
   \\
 Durability
+  & \S\ref{sec:durability}
   & S3/Swift interfaces (RESTful API)
   & object store library~\cite{weil_rados_2007}
-  & interface to persistence
-  \\
-Load Balancing
-  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
-  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
-  & migration \& sampling
+  & persistence/safety
   \\
 \end{tabular}
-\caption{Common internal abstractions. The same ``internal abstractions" are common across large-scale
-systems because they provide primitives that solve general distributed systems
-problems.  Here we list examples of what these internal abstractions are used
-for in ``production systems" and in Ceph.  Malacology provides these internal
-abstractions as interfaces that higher level applications can use; {\it
-italicized abstractions} are interfaces discussed in this paper.}
+\caption{Common internal abstractions. The same ``internal abstractions" are
+common across large-scale systems because they provide primitives that solve
+general distributed systems problems.  Here we list examples of what these
+internal abstractions are used for in ``production systems" and in Ceph.
+Malacology provides these internal abstractions as interfaces
+\newcomment{(Section~\ref{sec:malacology})}
+that higher level
+\oldcomment{applications}
+\newcomment{services (Section~\ref{sec:services})}
+can use.
+\oldcomment{; {\it italicized abstractions} are interfaces discussed in this
+paper.}}
+\addressesissue{1}
 \label{table:examples}
@@
 
-The guiding principle
-is to re-use existing services and extend them so that these services can be
-\emph{programmed}. We accomplish programmability of a service by exporting
-bindings (or ``hooks'') for an interpreted programming language so that
-programming can occur without having to restart the storage system (see also
-below,~\S\ref{sec:durability}). Table~\ref{table:examples} shows the internal services from Ceph that we expose in the Malacology prototype via Lua~\cite{ierusalimschy1996lua} bindings.
+The guiding principle is to re-use existing services and extend them so that
+these services can be \emph{programmed}. We accomplish programmability of a
+service by exporting bindings (or ``hooks'') for an interpreted programming
+language so that programming can occur without having to restart the storage
+system (see also below,~\S\ref{sec:durability}). Table~\ref{table:examples}
+shows the internal services from Ceph that we expose in the Malacology
+prototype via Lua~\cite{ierusalimschy1996lua} bindings
+\oldcomment{.}
+\newcomment{and Figure~\ref{fig:implementation-overview} compares what was
+already present in Ceph (gray boxes) to the Malacology interfaces we added.
+Section~\ref{sec:services} will describe the higher-level services we built
+with these interfaces.}
+\addressesissue{1}
+
+\begin{figure}[tbp]
+\centering
+\includegraphics{figures/implementation-overview.png}
+\caption{\newcomment{Malacology is implemented on the daemons and clients that
+run in a Ceph cluster. Interfaces expose internal subsystems and are used as
+building blocks for higher-level services.}\addressesissue{1}
+\label{fig:implementation-overview}}
+\end{figure}
 
@@
 \label{sec:mon}
+\label{sec:service-metadata-interface}
 \label{service-metadata}
@@
 
+\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
 In the case of Ceph, a consistent view of cluster state among server daemons
@@
 high-performance I/O path, a general guideline is to make use of this
-functionality infrequently and to assign small values to maps.
-
-\paragraph*{Malacology:} Malacology exposes a strongly-consistent view of
+functionality infrequently and to assign small values to maps.\\
+
+\noindent{\it{\textbf{Malacology: }}} Malacology exposes a strongly-consistent view of
 time-varying service metadata as a service rather than a hidden internal
@@
 new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
-policies in the load balancer \S\ref{malacology:mds}).  Using this service
+policies in the load balancer \S\ref{malacology:mds}.  Using this service
 guarantees that interface definitions are not only made durable, but are
@@
 clients are properly synchronized with the latest interfaces.
+
+\newcomment{\\ \noindent \it{\textbf{Impact: }} provides core functionality
+because it lets daemons come to consensus on system critical state. Bugs in the
+internal subsystems or ommitting this from services that need this type of
+consistency affects correctness.}
+\addressesissue{3, 5}
 
@@
 
-\subsection{Object I/O Interfaces}
+\oldcomment{\noindent 4.2 Object I/O Intefrace}
+\subsection{Data I/O Interface}
+\label{sec:data-io-interface}
 \label{object-data-interface}
-
 Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
@@
 
+\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
 An object interface is a plugin structured similarly to that of an RPC
@@
 an interface that atomically updates a matrix stored in the bytestream and an
-index of the matrix stored in the key-value database.
+index of the matrix stored in the key-value database.\\
 
@@
 
-\paragraph*{Malacology:} The implementation of Ceph's object abstraction,
+\noindent {\it{\textbf{Malacology: }}} The implementation of Ceph's object abstraction,
 although powerful, does not readily support programmability. Supporting only
@@
 
-All of these modifications work in tandem to implement the desired behavior,
+\oldcomment{All of these modifications work in tandem to implement the desired behavior,
 and are typically co-designed together such that each depends on the other to
-behave as expected.
+behave as expected.}
+\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
+performance by pushing behavior to lower parts of the storage stack, thereby
+minimizing hops and distributing computation.}
+\addressesissue{3, 5}
 
 \subsection{Distributed Metadata Interfaces}
+\begin{notes}
+\textcolor{red}{This section has been re-organized}
+\end{notes}
+\label{sec:distributed-metadata-interfaces}
 \label{malacology:mds}
 
-The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction~\cite{weil:sc2004-dyn-metadata}. In general, distributed
-file systems protect resources by providing hierarchical indexing and
-distributed locking services.  In Ceph, the locking service implements a
-capability-based system that expresses what data and metadata clients are
-allowed to access as well as what state they may cache and modify locally.
-While designed for the file abstraction, indexing, locking, and caching are
-all common services that are useful to a broad spectrum of applications.
-Distributed applications that share centralized resources (e.g. a database or
-directory) face similar challenges which are often solved using
-application-specific sharding.
-
-Ceph also addresses the challenge of balancing metadata load with a separate
-metadata cluster. This cluster uses load balancing policies to migrate
-directory inodes around the cluster to alleviate overload at single
-nodes~\cite{weil:sc2004-dyn-metadata}. The policies use metrics based on system
-state (e.g.  CPU and memory utilization) and statistics collected by the
-cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
-partitioning to move variable sized namespace subtrees. These units can be
-shipped anywhere (i.e., to any metadata server of any capacity) at any time for
-any reason. The original balancer was designed with hard-coded policies and
-tunables.
+\oldcomment{The distributed metadata service in Ceph provides clients with a POSIX file
+system abstraction~\cite{weil:sc2004-dyn-metadata}.}
+\newcomment{File systems provide clients with the familiar POSIX file abstraction.
+While this guarantees strong consistency it comes at the cost of scalability,
+increased complexity, and lower performance.}
+\addressesissue{1}
+In general, distributed file systems protect resources by providing
+hierarchical indexing and distributed locking services.
 
@@
 
-\paragraph*{Malacology:} Interfaces are added in strategic spots in the
+\oldcomment{\it{\textbf{Malacology:}} Interfaces are added in strategic spots in the
 metadata service for guarding shared resources, defining new file types,
-and load balancing.
-
-{\bf Shared resource interface.}
-The metadata service manages client sessions, allowing clients to obtain locks
-(e.g. file byte ranges), and capabilities (e.g. to cache file data). Clients
-and metadata servers use a cooperative protocol in which clients voluntarily
-release resources back to the metadata service in order to implement sharing
-policies. While the current policy for sharing access and voluntarily
-releasing resources is largely best-effort, Malacology supports generalized
-policies between metadata servers and clients that can be used to implement 
-fairness or priority.
-
-{\bf File type interface.} The metadata service manages a POSIX file system
-hierarchy where files and directories are represented as inode data structures
-and expose a POSIX file interface. Applications that manage large amounts of
-metadata (e.g. users or database snapshots) often require a naming service. We
-allow new inode types to be defined such that applications can create
-domain-specific interfaces to inodes that may modify locking and capability
-policies. We will show how this is used in Section~\ref{sec:seq} when we discuss a
-distributed shared-log built on Malacology.
-
-{\bf Load balancing interface.} Mantle~\cite{sevilla:sc15-mantle}, a
-programmable metadata load balancer, has been implemented on Malacology so that
-its infrastructure can enjoy all the properties of existing internal
-abstractions.  The load balancing logic is intact but the infrastructure has
-been improved to safely load, manage, and persist its load balancing policies.
-Composing Mantle with the service metadata and durability interfaces, Mantle
-can safely version balancer policies, save balancer policies in the backend
-object store and centralize warnings/errors. When combined with file types,
-load balancing policies can express a variety of rules such as multi-tenant
-and multi-application balancing.
+and load balancing.}
+
+\oldcomment{\noindent Shared resource interface.}
+\subsubsection{Shared Resource Interface}
+
+In general, file system metadata servers manages client sessions, allowing
+clients to obtain locks (e.g. file byte ranges), and capabilities (e.g. to
+cache file data). Clients and metadata servers use a cooperative protocol in
+which clients voluntarily release resources back to the metadata service in
+order to implement sharing policies.
+
+\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
+In Ceph, the locking service implements a capability-based system that
+expresses what data and metadata clients are allowed to access as well as what
+state they may cache and modify locally.  While designed for the file
+abstraction, indexing, locking, and caching are all common services that are
+useful to a broad spectrum of applications.  Distributed applications that
+share centralized resources (e.g. a database or directory) face similar
+challenges which are often solved using application-specific sharding.
+
+\newcomment{\\ \noindent\it{\textbf{Malacology: }}}
+While the current policy for sharing access and voluntarily releasing resources
+is largely best-effort, Malacology supports generalized policies between
+metadata servers and clients that can be used to implement fairness or
+priority.
+
+\newcomment{\\ \noindent {\it{\textbf{Impact: }}} provides core functionality
+to protect and provide exclusive access for any shared resource. May hurt
+performance if the resource in question does not require strong consistency.}
+\addressesissue{3}
+
+\oldcomment{\noindent File type interface.}
+\subsubsection{File Type Interface}
+The metadata service manages a POSIX file system hierarchy where files and
+directories are represented as inode data structures and expose a POSIX file
+interface. Applications that manage large amounts of metadata (e.g. users or
+database snapshots) often require a naming service. We allow new inode types to
+be defined such that applications can create domain-specific interfaces to
+inodes that may modify locking and capability policies. We will show how this
+is used in Section~\ref{sec:seq} when we discuss a distributed shared-log built
+on Malacology.
+
+\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this interface is both a feature
+and a performance optimization. It is a feature because it allows developers to
+add support for different  storage types, such as how to read new file formats
+or what consistency semantics to use for a specific subtree in the hierarchical
+namespace. It is also a performance optimization because future programmers can
+add optimizations for processing specific types of files into the inode itself.}
+\addressesissue{3}
+
+%%% TODO: striping strategy is stored with the inode
+
+\oldcomment{\noindent \bf Load balancing interface.} 
+\subsubsection{Load Balancing Interface}
+
+Ceph also addresses the challenge of balancing metadata load with a separate
+metadata cluster. This cluster uses load balancing policies to migrate
+directory inodes around the cluster to alleviate overload at single
+nodes~\cite{weil:sc2004-dyn-metadata}. The policies use metrics based on system
+state (e.g.  CPU and memory utilization) and statistics collected by the
+cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
+partitioning to move variable sized namespace subtrees. These units can be
+shipped anywhere (i.e., to any metadata server of any capacity) at any time for
+any reason. The original balancer was designed with hard-coded policies and
+tunables.\\
+
+\noindent {\it{\textbf{Malacology: }}}
+Mantle~\cite{sevilla:sc15-mantle}, a programmable metadata load balancer, has
+been implemented on Malacology so that its infrastructure can enjoy all the
+properties of existing internal abstractions.  The load balancing logic is
+intact but the infrastructure has been improved to safely load, manage, and
+persist its load balancing policies.  Composing Mantle with the service
+metadata and durability interfaces, Mantle can safely version balancer
+policies, save balancer policies in the backend object store and centralize
+warnings/errors. When combined with file types, load balancing policies can
+express a variety of rules such as multi-tenant and multi-application
+balancing.
+
+\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
+performance by allowing them to specify how to partition, replicate, and
+distribute metadata in response to overloaded servers.}
+\addressesissue{3}
 
@@
 system, and autonomously initiate recovery mechanisms when failures are
-discovered.
-
-\paragraph*{Malacology:} Metadata service policies and object storage
+discovered.\\
+
+\noindent {\it{\textbf{Malacology: }}} Metadata service policies and object storage
 interfaces are stored durability within RADOS and managed by storing references
@@
 synchronized on correct implementations without restarting.
+
+\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this is a feature because it
+adds data safety and persistence to system metadata; while nice to have it does
+effect correctness.}
+\addressesissue{3}
 

### 2017-02-13 paper/paper.tex
paper: version 0; addresses reviewer confusion - clarify what was implemented - describe what makes a programmable storage system - organize motivation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % Uncomment for diff
-\includecomment{notes}
-\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
-\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
-\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
+%\includecomment{notes}
+%\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
+%\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
+%\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
 
 % Uncomment for camera-ready
-%\excludecomment{notes}
-%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
-%\newcommand{\oldcomment}[1]{}
-%\newcommand{\addressesissue}[1]{}
-
+\excludecomment{notes}
+\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
+\newcommand{\oldcomment}[1]{}
+\newcommand{\addressesissue}[1]{}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\authorinfo{Paper ID: 165}          {Total Pages: 12}
-\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre}
-           {\{msevilla, jayhawk, ivo, cmaltzahn\}@soe.ucsc.edu, \{palvaro, finkelstein, jlefevre\}@ucsc.edu}
+\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre, Carlos Maltzahn}
+           {\{msevilla, jayhawk, ivo\}@soe.ucsc.edu, \{palvaro, shel, jlefevre, carlosm\}@ucsc.edu}
 \date{}
@@
 \textcolor{red}{
-\noindent Issues raised by reviewers:
+\noindent Issues raised by reviewers, paper augmented as noted inline (red/blue text):
 \begin{enumerate}
@@
 system. The interfaces presented in this paper are abstractions that we found
-useful for building ZLog and Mantle, yet they may not provide the best
+useful for building our prototype services ZLog and Mantle, yet they may not provide the best
 trade-offs for all higher-level services.}
@@
 \textbf{Interface}                      &
-\textbf{Sec}                            &
+\textbf{Section}                        &
 \textbf{Example in Production Systems}  &
@@
 Shared Resource
-  & \S\ref{sec:distributed-metadata-interfaces}
+  & \S\ref{sec:shared-resource-interface}
   & MPI collective IO, burst
@@
 File Type
-  & \S\ref{sec:distributed-metadata-interfaces}
+  & \S\ref{sec:file-type-interface}
   & ?????
@@
 Load Balancing
-  & \S\ref{sec:distributed-metadata-interfaces}
+  & \S\ref{sec:load-balancing-interface}
   & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
@@
 
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
+\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}}
 In the case of Ceph, a consistent view of cluster state among server daemons
@@
 transparently and consistently propagated throughout the cluster so that
-clients are properly synchronized with the latest interfaces.
-
-\newcomment{\\ \noindent \it{\textbf{Impact: }} provides core functionality
+clients are properly synchronized with the latest interfaces.\\
+
+\noindent\newcomment{{\it{\textbf{Impact: }}} provides core functionality
 because it lets daemons come to consensus on system critical state. Bugs in the
@@
 
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
+\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}}
 An object interface is a plugin structured similarly to that of an RPC
@@
 \subsubsection{Shared Resource Interface}
+\label{sec:shared-resource-interface}
 
@@
 
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
+\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}}
 In Ceph, the locking service implements a capability-based system that
@@
 \oldcomment{\noindent File type interface.}
-\subsubsection{File Type Interface}
-The metadata service manages a POSIX file system hierarchy where files and
-directories are represented as inode data structures and expose a POSIX file
-interface. Applications that manage large amounts of metadata (e.g. users or
-database snapshots) often require a naming service. We allow new inode types to
+\subsubsection{Malacology: File Type Interface}
+\label{sec:file-type-interface}
+
+Applications that manage large amounts of metadata (e.g. users or database
+snapshots) often require a naming service.  The metadata service manages a
+POSIX file system hierarchy where files and directories are represented as
+inode data structures and expose a POSIX file interface.
+
+\newcomment{\\ \noindent {\it{\textbf{Existing Ceph Implementation: }}} CephFS
+is the POSIX compliant file system that uses Ceph.  Inodes are quite large (1KB
+for an inode, 400 bytes for a directory entry, and 700 bytes for a directory)
+and contain CephFS-specific policies like how to stripe data across RADOS.}
+
+\newcomment{\\ \noindent {\it{\textbf{Malacology: }}}} We allow new inode types to
 be defined such that applications can create domain-specific interfaces to
@@
 \subsubsection{Load Balancing Interface}
-
-Ceph also addresses the challenge of balancing metadata load with a separate
+\label{sec:load-balancing-interface}
+
+\newcomment{Many large scale storage systems separate metadata and data IO so
+that the corresponding services can scale independently. Metadata requests
+transfer small amounts of data and they happen relatively frequently so many
+systems employ separate metadata clusters.}
+
+\newcomment{\\ \noindent {\it{\textbf{Existing Ceph Implementation: }}}} Ceph
+also addresses the challenge of balancing metadata load with a separate
 metadata cluster. This cluster uses load balancing policies to migrate
 directory inodes around the cluster to alleviate overload at single
-nodes~\cite{weil:sc2004-dyn-metadata}. The policies use metrics based on system
-state (e.g.  CPU and memory utilization) and statistics collected by the
+nodes~\cite{weil:sc2004-dyn-metadata}.  The policies use metrics based on
+system state (e.g.  CPU and memory utilization) and statistics collected by the
 cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
@@
 
+\newcomment{Object stores protect data using techniques like erasure coding,
+replication, and data scrubbing. For scalability, many of these features are
+implemented using a peer-to-peer protocol that allows object storage devices to
+operate autonomously without a centralized coordinator.\\}
+
+\noindent{\it{\textbf{Existing Ceph Implementation: }}}
 Ceph provides storage by striping and replicating data across
 RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-protects data using common techniques such as erasure
-coding, replication, and scrubbing.  Furthermore, many of these features
-are implemented using a peer-to-peer protocol that allows object storage
-devices to operate autonomously without a centralized coordinator.  For
-example, when the number of placement groups change, the OSDs re-balance and
-re-shard data in the background in a process called placement group splitting
-in which OSDs communicate directly with each other to cover a new state.
-In order to reduce load on the monitoring service, Ceph OSDs use a gossip
-protocol to efficiently propagate changes to cluster maps throughout the
+protects data using common techniques such as erasure coding, replication, and
+scrubbing.  example, when the number of placement groups change, the OSDs
+re-balance and re-shard data in the background in a process called placement
+group splitting in which OSDs communicate directly with each other to cover a
+new state.  In order to reduce load on the monitoring service, Ceph OSDs use a
+gossip protocol to efficiently propagate changes to cluster maps throughout the
 system, and autonomously initiate recovery mechanisms when failures are
@@
 adds data safety and persistence to system metadata; while nice to have it does
-effect correctness.}
+not necessarily effect correctness.}
 \addressesissue{3}

### 2017-02-22 paper/paper.tex
paper: version 1 - adds future work and clarifies metadata - brings in more citations and all our projects - still under the page limit (we get 14 + refs)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Uncomment for diff
+% Uncomment for diff 0
 %\includecomment{notes}
@@
 
-% Uncomment for camera-ready
-\excludecomment{notes}
+% Uncomment for diff 1
+\includecomment{notes}
+\newcommand{\newcommentone}[1]{{\textcolor{red}{\textbf{#1}}}}
+\newcommand{\oldcommentone}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
 \newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
 \newcommand{\oldcomment}[1]{}
-\newcommand{\addressesissue}[1]{}
+\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
+
+% Uncomment for camera-ready
+%\excludecomment{notes}
+%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
+%\newcommand{\oldcomment}[1]{}
+%\newcommand{\newcommentone}[1]{{\textcolor{black}{#1}}}
+%\newcommand{\oldcommentone}[1]{}
+%\newcommand{\addressesissue}[1]{}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 challenges of programmability by composing existing primitives into two new
-higher-level services: a metadata load balancer and a high-performance
+higher-level services: a \newcommentone{file system} metadata load balancer and a high-performance
 distributed shared-log. The evaluation demonstrates that our services inherit
@@
 to navigate trade-offs between latency and throughput using leases.
+\addressesissue{4}
 \end{abstract}
@@
 of new services via composition.  A programmable storage system can be realized
-by exposing existing functionality (such as metadata services and
+by exposing existing functionality (such as \newcommentone{file system and cluster} metadata services and
 synchronization and monitoring capabilities) as interfaces that can be ``glued
@@
 exposed in a programmable storage system.
+\addressesissue{4}
 
@@
 storage because it offers a broad spectrum of existing services, including
-distributed locking and caching services provided by metadata servers,
+distributed locking and caching services provided by \newcommentone{file system} metadata servers,
 durability and object interfaces provided by the backend object store, and
@@
 functionality necessary for implementing new services.
+\addressesissue{4}
 
@@
 delays and heterogeneous hardware in dynamic environments.
-
-\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}}
-In the case of Ceph, a consistent view of cluster state among server daemons
-and clients is critical to provide strong consistency guarantees to clients.
-Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
-metadata server) data structures called ``maps'' that record membership and
-status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
-responsible for integrating state changes into cluster maps, responding to
-requests from out-of-date clients and synchronizing members of the cluster
-whenever there is a change in a map so that they all observe the same system
-state. As a fundamental building block of many system designs, consensus
-abstractions such as Paxos are a common technique for maintaining consistent
-data versions, and are a useful system to expose.
+\newcommentone{Service metadata is information about the daemons in the system
+and includes membership details, hardware layout ({\it e.g.}, racks, power
+supplies, etc.), data layout, and daemon state/configuration. It differs from
+traditional file system metadata which is information about files. For the rest
+of the paper when we refer to metadata server or service, we mean a daemon that manages file
+system metadata.}
+\addressesissue{4}
+
+\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}} In the
+case of Ceph, a consistent view of cluster state among server daemons and
+clients is critical to provide strong consistency guarantees to clients.  Ceph
+maintains cluster state information in per-subsystem  data structures called
+``maps'' that record membership and status information.  A
+Paxos~\cite{lamport_parttime_1998} monitoring service is responsible for
+integrating state changes into cluster maps, responding to requests from
+out-of-date clients and synchronizing members of the cluster whenever there is
+a change in a map so that they all observe the same system state. As a
+fundamental building block of many system designs, consensus abstractions such
+as Paxos are a common technique for maintaining consistent data versions, and
+are a useful system to expose.
 
@@
 given maps.  Since Ceph does not expose this service directly, as part of our Malacology implementation, we expose a key-value
-service designed for managing configuration metadata that is built on top of
+service designed for managing \oldcommentone{configuration} \newcommentone{service} metadata that is built on top of
 the consensus engine. Since the monitor is intended to be out of the
@@
 values) or triggering actions based on specific values (e.g. sanitize values).
-The higher level services we implement in \S\ref{sec:services} make use of this
+The higher-level services we implement in \S\ref{sec:services} make use of this
 functionality to register, version and propagate dynamic code (Lua scripts) for
@@
 \includegraphics{figures/implementation.png}
-\caption{Malacology allows users to dynamically define object/metadata
+\caption{Malacology allows users to dynamically define object/\newcommentone{file system} metadata
 interfaces by extending the object storage daemon (OSD) and metadata server
@@
 \oldcomment{\it{\textbf{Malacology:}} Interfaces are added in strategic spots in the
-metadata service for guarding shared resources, defining new file types,
+\newcommentone{file system} metadata servers for guarding shared resources, defining new file types,
 and load balancing.}
@@
 cache file data). Clients and metadata servers use a cooperative protocol in
-which clients voluntarily release resources back to the metadata service in
+which clients voluntarily release resources back to the \newcommentone{file system} metadata service in
 order to implement sharing policies.
@@
 In Ceph, the locking service implements a capability-based system that
-expresses what data and metadata clients are allowed to access as well as what
+expresses what data and \newcommentone{file system} metadata clients are allowed to access as well as what
 state they may cache and modify locally.  While designed for the file
@@
 
-Applications that manage large amounts of metadata (e.g. users or database
+Applications that manage large amounts of \newcommentone{file system} metadata (e.g. users or database
 snapshots) often require a naming service.  The metadata service manages a
@@
 
-\newcomment{Many large scale storage systems separate metadata and data IO so
+\newcomment{Many large scale storage systems separate file system metadata and data IO so
 that the corresponding services can scale independently. Metadata requests
 transfer small amounts of data and they happen relatively frequently so many
-systems employ separate metadata clusters.}
+systems employ separate file system metadata clusters.}
 
 \newcomment{\\ \noindent {\it{\textbf{Existing Ceph Implementation: }}}} Ceph
-also addresses the challenge of balancing metadata load with a separate
+also addresses the challenge of balancing \newcommentone{file system} metadata load with a separate
 metadata cluster. This cluster uses load balancing policies to migrate
@@
 \noindent {\it{\textbf{Malacology: }}}
-Mantle~\cite{sevilla:sc15-mantle}, a programmable metadata load balancer, has
+Mantle~\cite{sevilla:sc15-mantle}, a programmable \newcomment{file system}metadata load balancer, has
 been implemented on Malacology so that its infrastructure can enjoy all the
@@
 these data structures, we use this service to automatically install interfaces
-in OSDs, and install policies within the metadata server daemon (MDS) such that clients and daemons are
+in OSDs, and install policies within the metadata server daemon \oldcommentone{(MDS)} such that clients and daemons are
 synchronized on correct implementations without restarting.
@@
 
+\section{Future Work}
+
+\newcommentone{This work is one point along the path towards producing
+general-purpose storage systems that can target special-purpose applications.
+Continuing to expore existing storage system functionality as reusable building
+blocks will enable the proliferation of storage-centric APIs, allowing
+developers to custom-fit their applications to take advantage of the
+capabilities of the underlying system. When adding new interfaces and composing
+them together, careful consideration should be made to address security, access
+control, and safety. Correctness and performance will surely be affected if
+interfaces are combined poorly and these drawbacks will be quantified.}
+\addressesissue{5}
+
+\newcommentone{The next higher-level services we plan to build using the
+interfaces in Table~\ref{table:examples} are: an elastic cloud database, a data
+layout manager, and a data processing engine. Some cloud databases aggressively
+cache data to hide the overheads and weaker consistency semantics of the
+distributed file systems they sit on~\cite{pavlo:sigmod16} but using Malacology
+we could build a database that works well because of the storage system not in
+spite of it. We plan to use the data IO interface for push down predicates, the
+file type interface for maintaining indices and stats, and the durability
+interface internals for ingest and data movement. The data layout manager would
+be similar to DRepl~\cite{ionkov:msst2013} and would use the file type,
+durability, and data I/O interface to manage different views of data. Finally,
+the data processing engine would be similar to MapReduce and would use the data
+I/O, shared resource, and load balancing interface to implement the resource
+manager, computation engine, and storage platform in the Hadoop ecosystem.}
+
+\newcommentone{ Our experience with Zlog and Mantle illustrates that the labor
+of wrapping existing services in interfaces is justified by the power and
+flexibility they afford to programmers.  In exchange for this flexibility,
+however, programmable storage systems force programmers to forfeit the
+\emph{protection from change} afforded by narrow storage interfaces like the
+POSIX API.  To implement applications built atop systems such as Malacology,
+programmers must navigate a complex design space, simultaneously addressing
+often orthogonal concerns including functional correctness, performance and
+fault tolerance. Worse still, these compositions can be sensitive to changes in
+the underlying environment (such as hardware upgrades, software version changes
+and evolving workloads).  For example, a major version change in Ceph required
+us to rewrite significant parts of Zlog to maintain acceptable performance --
+each such evolution costs developer time and risks introducing bugs.}
+
+\newcommentone{We are actively exploring the use of high-level
+\emph{declarative} languages based on Datalog~\cite{alvaro:cidr11} to program
+storage APIs.  In this way, a programmer can specify the functional behavior of
+a storage interface once in a relational (or algebraic) language, allowing an
+optimizer to model search the space of functionally equivalent physical
+implementations, and re-optimizing when characteristics change.  Much like
+query planning and optimization in database systems~\cite{hellerstein:cidr15},
+this approach will separate the concerns of correctness and performance, and
+protect applications (which evolve slowly) against changes in highly dynamic
+storage system.}
+ \addressesissue{3}
+
 \section{Related Work}
@@
 gap between applications and
-storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}).
+storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}.
 
@@
 
-\section{Conclusion and Future Work}
+\section{Conclusion}
 \label{conclusion-and-future-work}
@@
 
-Future work will focus on constructing more interfaces to support a wide
+\oldcommentone{Future work will focus on constructing more interfaces to support a wide
 variety of storage system services that can be configured on-the-fly in existing
@@
 target  special-purpose applications.  Ultimately we want to utilize
-declarative methods for expressing new services.
+declarative methods for expressing new services.}
 

### 2017-02-23 paper/0_abstract.tex
paper: cleanup line breaks and break into separate files
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-02-09 paper/paper.tex
@@
-\documentclass[preprint]{sigplanconf-eurosys}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage{pifont}
-\usepackage{color}
-\usepackage{comment}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Uncomment for diff
-\includecomment{notes}
-\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
-\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
-\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
-
-% Uncomment for camera-ready
-%\excludecomment{notes}
-%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
-%\newcommand{\oldcomment}[1]{}
-%\newcommand{\addressesissue}[1]{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\cmark}{\ding{51}}%
-\usepackage[font=footnotesize,labelfont=bf]{caption}
-
-\setlength{\belowcaptionskip}{-10pt}
-
-\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
-\usepackage[normalem]{ulem}
-\usepackage{pstricks}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-\setkeys{Gin}{width=\maxwidth,keepaspectratio}
-
-\usepackage[unicode=true]{hyperref}
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, Peter Alvaro, Shel Finkelstein, Jeff LeFevre},
-            pdftitle={Malacology: A Programmable Storage System},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=blue,
-            linkcolor=black,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-
-\title{Malacology: A Programmable Storage System}
-%\authorinfo{Paper ID: 165}          {Total Pages: 12}
-\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre}
-           {\{msevilla, jayhawk, ivo, cmaltzahn\}@soe.ucsc.edu, \{palvaro, finkelstein, jlefevre\}@ucsc.edu}
-\date{}
-\begin{document}
-
-\maketitle
-
 \begin{notes}
 \textcolor{red}{
-\noindent Issues raised by reviewers:
+\noindent Issues raised by reviewers, paper augmented as noted inline (red/blue text):
 \begin{enumerate}
@@
 challenges of programmability by composing existing primitives into two new
-higher-level services: a metadata load balancer and a high-performance
-distributed shared-log. The evaluation demonstrates that our services inherit
-desirable qualities of the back-end storage system, including the ability to
-balance load, efficiently propagate cluster metadata, recover from failure, and
-to navigate trade-offs between latency and throughput using leases.
+higher-level services: a \newcommentone{file system} metadata load balancer and
+a high-performance distributed shared-log. The evaluation demonstrates that our
+services inherit desirable qualities of the back-end storage system, including
+the ability to balance load, efficiently propagate cluster metadata, recover
+from failure, and to navigate trade-offs between latency and throughput using
+leases.  \addressesissue{4}
 \end{abstract}
-
-\begin{figure}[tb]
-\centering
-\includegraphics{figures/overview.png}
-\caption{Scalable storage systems have storage daemons which store data,
-monitor daemons (M) that maintain cluster state, and service-specific daemons
-({\it e.g.}, file system metadata servers). Malacology enables the programmability of
-internal abstractions (bold arrows) to re-use and compose
-existing subsystems.  With Malacology, we add new services, ZLog and
-Mantle, that sit alongside traditional user-facing APIs (file, block,
-object).}\label{fig:overview}
-\end{figure}
-
-\section{Introduction}
-\label{introduction}
-\label{sec:intro}
-
-A storage system implements abstractions designed to persistently store data
-and must exhibit a high level of correctness to prevent data loss.  Storage
-systems have evolved around storage devices that often were orders of magnitude
-slower than CPU and memory, and therefore could dominate overall performance if
-not used carefully. Over the last few decades members of the storage systems
-community have developed clever strategies to meet correctness requirements
-while somewhat hiding the latency of traditional storage
-media~\cite{brewer_disks_2016}. To avoid lock-in by a particular vendor, users
-of storage systems have preferred systems with highly standardized APIs and
-lowest common denominator abstract data types such as blocks of bytes and byte
-stream files~\cite{armbrust_view_2010}.
-
-A number of recent developments have disrupted traditional storage systems.
-First, the falling prices of flash storage and the availability of new types of
-non-volatile memory that are orders of magnitude faster than traditional
-spinning media are moving overall performance bottlenecks away from storage
-devices to CPUs and networking, and pressure storage systems to shorten their
-code paths and incorporate new
-optimizations~\cite{gray_tape_2007,gray_flash_2008}.  Second, emerging ``big
-data'' applications demand interface evolution to support flexible consistency
-as well as flexible structured data 
-representations.~\cite{apache_contributors_parquet_2014}.  Finally, production-quality scalable
-storage systems available as open source software have established and are
-continuing to establish new, \emph{de-facto} API standards at a faster pace
-than traditional standards
-bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
-
-The evolutionary pressure placed on storage systems by these trends raises the
-question of whether there are principles that storage systems designers can
-follow to evolve storage systems efficiently, without jeopardizing years of
-code-hardening and performance optimization efforts.  In this paper we
-investigate an approach that focuses on identifying and exposing existing
-storage system resources, services, and abstractions that in a generalized form
-can be used to \emph{program} new services. By composing higher-level services
-over existing abstractions one can reuse subsystems and leverage their
-optimizations, established correctness, robustness, and efficiency.
-
-{\it \textbf{Contribution 1:}} We define a programmable storage system to be a
-storage system that facilitates the re-use and extension of existing storage
-abstractions provided by the underlying software stack, to enable the creation
-of new services via composition.  A programmable storage system can be realized
-by exposing existing functionality (such as metadata services and
-synchronization and monitoring capabilities) as interfaces that can be ``glued
-together'' in a variety of ways using a high-level language. Programmable
-storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection
-and execution of code within a storage system or storage device---in that the
-former is applicable to any component of the storage system, while the latter
-focuses at the data access level. Given this contrast, we can say that active
-storage is an example of how one internal component (the storage layer) is
-exposed in a programmable storage system.
-
-To illustrate the benefits and challenges of this approach we have designed and
-evaluated Malacology, a programmable storage system that facilitates the
-construction of new services by re-purposing existing subsystem abstractions of
-the storage stack.  We build Malacology in Ceph, a popular open source software
-storage stack.  We choose Ceph to demonstrate the concept of programmable
-storage because it offers a broad spectrum of existing services, including
-distributed locking and caching services provided by metadata servers,
-durability and object interfaces provided by the backend object store, and
-propagation of consistent cluster state provided by the monitoring service (see
-Figure~\ref{fig:overview}).  Malacology is expressive enough to provide the
-functionality necessary for implementing new services.
-
-Malacology includes a non-exhaustive set of interfaces that can be used as
-building blocks for constructing novel storage abstractions, including:
-
-\begin{enumerate}
-
-\item An interface for managing strongly-consistent time-varying \textbf{service
-metadata}.
-
-\item An interface for installing and evolving domain-specific, cluster-wide
-\textbf{data interfaces}.
-
-\item An interface for managing access to \textbf{shared resources} using a
-variety of optimization strategies.
-
-\item An interface for \textbf{load balancing} resources across the cluster.
-
-\item An interface for \textbf{durability} that persists policies using the
-underlying storage stack's object store.
-
-\end{enumerate}
-
-{\it \textbf{Contribution 2:}} We implement two distributed services using Malacology to
-demonstrate the feasibility of the programmable storage approach:
-
-\begin{enumerate}
-
-\item A high-performance distributed shared log service called ZLog, that is an
-implementation of CORFU.~\cite{balakrishnan_corfu_2012}
-
-\item An implementation of Mantle, the programmable load balancing
-service~\cite{sevilla:sc15-mantle}
-
-\end{enumerate}
-
-The remainder of this paper is structured as follows. First, we describe and
-motivate the need for programmable storage by describing current practices in
-the open source software community. Next we describe Malacology by presenting the
-subsystems within the underlying storage system that we re-purpose, and briefly
-describe how those system are used within Malacology (\S\ref{sec:malacology}).
-Then we describe the services that we have constructed within the Malacology
-framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
-implementation (\S\ref{sec:evaluation}).  We conclude by discussing related and future work.
-% and
-%conclude.
-
-\section{Application-Specific Storage Stacks}
-\label{sec:application-specifc-storage-stacks}
-
-Building storage stacks from the ground up for a specific purpose results in
-the best performance. For example, GFS~\cite{ghemawat:sosp03} and
-HDFS~\cite{shvachko:msst10} were designed specifically to serve Map/Reduce and
-Hadoop jobs, and use techniques like exposing data locality and relaxing POSIX
-constraints to achieve application-specific I/O optimizations. Another example
-is Boxwood~\cite{maccormick:osdi04}, which experimented with B-trees
-and chunk stores as storage abstractions to simplify application building.
-Alternatively, general-purpose storage stacks are built with the flexibility to
-serve many applications by providing a variety of interfaces and tunable
-parameters.  Unfortunately, managing competing forces in these systems is
-difficult and users want more control from the general-purpose storage stacks
-without going as far as building their storage system from the ground up.
-
-To demonstrate a recent trend towards more application-specific storage systems
-we examine the state of programmability in Ceph. Something of a storage Swiss
-army knife, Ceph simultaneously supports file, block, and object interfaces on
-a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
-Distributed Object Storage (RADOS) system is a cluster of object storage
-devices (OSDs) that provide Ceph with data durability and integrity using
-replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already
-provides some degree of programmability; the OSDs support domain-specific
-object interfaces that are implemented by composing existing low-level storage
-interfaces that execute atomically. These interfaces are written in C++ and are
-statically loaded into the system.
-
-The Ceph community provides empirical evidence that developers are already
-beginning to embrace programmable storage. Figure~\ref{fig:obj-int-dev-growth}
-shows a dramatic growth in the production use of domain-specific interfaces in
-the Ceph community since 2010. In that figure, classes are functional groupings
-of methods on storage objects (e.g. remotely computing and caching the checksum
-of an object extent).  What is most remarkable is that this trend contradicts
-the notion that API changes are a burden for users.  Rather it appears that
-gaps in existing interfaces are being addressed through ad-hoc approaches to
-programmability. In fact, Table~\ref{table:objclasses} categorizes existing
-interfaces and we clearly see a trend towards reusable services.
-
-\begin{figure}[ht]
-\centering
-\includegraphics{figures/obj-int-dev-growth.png}
-\caption{Since 2010, the growth in the number of co-designed object storage interfaces
-in Ceph has been accelerating. This plot is the number of object classes (a
-group of interfaces), and the total number of methods (the actual API
-end-points).}
-\label{fig:obj-int-dev-growth}
-\end{figure}
-
-\begin{table}[ht]
-\centering
-  \begin{tabular}{l|l|l}
-    Category & Example & \# \\ \hline
-    Logging  & Geographically distribute replicas & 11 \\ \hdashline
-    Metadata & Snapshots in the block device OR  & \multirow{2}{*}{74} \\
-    Management & Scan extents for file system repair & \\ \hdashline
-    Locking  & Grants clients exclusive access & 6 \\ \hdashline
-    Other & Garbage collection, reference counting  & 4\\
-\end{tabular}
-\caption{A variety of object storage classes exist to expose interfaces
-    to applications. \# is the number of methods that implement these categories.
-}
-\label{table:objclasses}
-\end{table}
-
-The takeaway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
-already trying to use programmability because their needs, whether they be
-related to performance, availability, consistency, convenience, etc., are not
-satisfied by the existing default set of interfaces. The popularity of the
-custom object interface facility of Ceph could be due to a number of reasons,
-such as the default algorithms/tunables of the storage system being
-insufficient for the application's performance goals, programmers wanting to
-exploit application-specific semantics, and/or programmers knowing how to
-manage resources to improve performance. A solution based on
-application-specific object interfaces is a way to work around the
-traditionally rigid storage APIs because custom object interfaces give
-programmers the ability to tell the storage system about their application: if
-the application is CPU or IO bound, if it has locality, if its size has the potential to
-overload a single node, etc.  Programmers often know what the problem is
-and how to solve it, but until the ability to modify object interfaces, they
-had no way to express to the storage system how to handle their data.
-
-Our approach is to expose more of the commonly used, code-hardened subsystems
-of the underlying storage system as interfaces. The intent is that these
-interfaces, which can be as simple as a redirection to the persistent data
-store or as complicated as a strongly consistent directory service, should be
-used and re-used in many contexts to implement a wide range of services. By
-making programmability a `feature', rather than a `hack' or `workaround', we
-help standardize a development process that now is largely ad-hoc.
-
-\section{Challenges}
-\label{sec:challenges}
-
-Implementing the infrastructure for programmability into existing services and
-abstractions of distributed storage systems is challenging, even if one assumes
-that the source code of the storage system and the necessary expertise for
-understanding it is available.  Some challenges include:
-
-\begin{itemize}
-
-\item Storage systems are generally required to be highly available so that any
-complete restarts of the storage system to reprogram them is usually
-unacceptable.
-
-\item Policies and optimizations are usually hard-wired into the services and
-one has to be careful when factoring them to avoid introducing additional bugs.
-These policies and optimizations are usually cross-cutting solutions to
-concerns or trade-offs that cannot be fully explored at the time the code is
-written (as they relate to workload or hardware). Given these policies and
-optimizations, decomposition of otherwise orthogonal internal abstractions can
-be difficult or dangerous.
-
-\item Mechanisms that are often only exercised according to hard-wired policies
-and not in their full generality have hidden bugs that are revealed as soon as
-those mechanisms are governed by different policies. In our experience
-introducing programmability into a storage system proved to be a great
-debugging tool.
-
-\item Programmability, especially in live systems, implies changes that need to
-be carefully managed by the system itself, including versioning and propagation
-of those changes without affecting correctness.
-
-\end{itemize}
-
-To address these challenges we present Malacology
-\oldcomment{. Malacology is both a prototype for a programmable storage system
-and a}
-\newcomment{, our prototype programmable storage system. It uses the
-programmable storage}
-\addressesissue{2}
-design approach to evolve storage systems efficiently and
-without jeopardizing years of code-hardening and performance optimization
-efforts.  Although Malacology uses the internal abstractions of the underlying
-storage system, including its subsystems, components, and implementations, we
-emphasize that our system still addresses the general challenges outlined
-above.
-
-The main challenge of designing a programmable storage system is choosing the
-right internal abstractions and picking the correct layers for exposing them.
-\oldcomment{In this paper, we do not present an exhaustive list of possible
-internal abstractions nor do we contend that the abstractions we choose provide
-the best trade-offs for all applications.} 
-\newcomment{A programmable storage is not defined by what abstractions are
-exposed, rather a programmable storage system adheres to the design approach of
-choosing interfaces so administrators can have better control of the storage
-system. The interfaces presented in this paper are abstractions that we found
-useful for building ZLog and Mantle, yet they may not provide the best
-trade-offs for all higher-level services.}
-\addressesissue{2}
-For example, if consensus is correctly exposed one could implement
-high-level features like versioning, serialization, or various flavors of
-strongly consistent data management on top; but perhaps a low-level consensus
-interface is suited well for a particular set of applications.  These questions
-are not answered in this paper and instead we focus on showing the feasibility
-of building such a system, given advances in the quality and robustness of
-today's storage stacks.
-
-The Malacology prototype we present has been implemented on Ceph.
-While there are other systems on top of which Malacology could be implemented (see Table~\ref{table:examples}), we choose Ceph because it is a production quality system and because it
-is open source. The large developer community ensures that code is
-robust and the visibility of the code lets us expose any interface we want. In
-the next section we describe the Ceph components that we expose as Malacology
-interfaces.
-
-\section{Malacology: A Programmable Storage System}
-\label{sec:malacology}
-
-\begin{table*}
-\centering\small
-\begin{tabular}{ l | c | l | l | l }
-%\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
-\multicolumn{4}{c}{} \\
-\textbf{Interface}                      &
-\textbf{Sec}                            &
-\textbf{Example in Production Systems}  &
-\textbf{Example in Ceph}                &
-\textbf{Provided Primitives}            \\ \hline
-Service Metadata
-  & \S\ref{sec:service-metadata-interface}
-  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
-  & cluster state management~\cite{website:ceph-mon}
-  & consensus/consistency
-  \\
-Data I/O
-  & \S\ref{sec:data-io-interface}
-  & Swift in situ storage/compute~\cite{website:zerocloud}
-  & object interface classes~\cite{website:cls-lua}
-  & transaction/atomicity
-  \\
-Shared Resource
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & MPI collective IO, burst
-  & POSIX metadata protocols
-  & serialization/batching
-  \\
-File Type
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & ?????
-  & file striping strategy
-  & data/metadata access
-  \\
-Load Balancing
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
-  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
-  & migration/sampling
-  \\
-Durability
-  & \S\ref{sec:durability}
-  & S3/Swift interfaces (RESTful API)
-  & object store library~\cite{weil_rados_2007}
-  & persistence/safety
-  \\
-\end{tabular}
-\caption{Common internal abstractions. The same ``internal abstractions" are
-common across large-scale systems because they provide primitives that solve
-general distributed systems problems.  Here we list examples of what these
-internal abstractions are used for in ``production systems" and in Ceph.
-Malacology provides these internal abstractions as interfaces
-\newcomment{(Section~\ref{sec:malacology})}
-that higher level
-\oldcomment{applications}
-\newcomment{services (Section~\ref{sec:services})}
-can use.
-\oldcomment{; {\it italicized abstractions} are interfaces discussed in this
-paper.}}
-\addressesissue{1}
-\label{table:examples}
-\end{table*}
-
-The guiding principle is to re-use existing services and extend them so that
-these services can be \emph{programmed}. We accomplish programmability of a
-service by exporting bindings (or ``hooks'') for an interpreted programming
-language so that programming can occur without having to restart the storage
-system (see also below,~\S\ref{sec:durability}). Table~\ref{table:examples}
-shows the internal services from Ceph that we expose in the Malacology
-prototype via Lua~\cite{ierusalimschy1996lua} bindings
-\oldcomment{.}
-\newcomment{and Figure~\ref{fig:implementation-overview} compares what was
-already present in Ceph (gray boxes) to the Malacology interfaces we added.
-Section~\ref{sec:services} will describe the higher-level services we built
-with these interfaces.}
-\addressesissue{1}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation-overview.png}
-\caption{\newcomment{Malacology is implemented on the daemons and clients that
-run in a Ceph cluster. Interfaces expose internal subsystems and are used as
-building blocks for higher-level services.}\addressesissue{1}
-\label{fig:implementation-overview}}
-\end{figure}
-
-Lua is a portable embedded scripting language and we choose it as the
-interpreted language for Malacology because it offers superior performance and
-productivity trade-offs, including a JIT-based implementation that is well
-known for near native performance. Additionally, Lua has been used extensively
-in game engines, and systems research \cite{neto:dls14-luaos}, including
-storage systems where it has been effectively used both on
-\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010}
-and off \cite{sevilla:sc15-mantle} the performance critical path. Finally, the
-flexibility of the runtime allows execution sandboxing in order to address
-security and performance concerns. We will now discuss the common subsystems
-used to manage storage system and how Malacology makes them programmable.
-
-\subsection{Service Metadata Interface}
-\label{sec:mon}
-\label{sec:service-metadata-interface}
-\label{service-metadata}
-
-% straw man example
-Keeping track of state in a distributed system is an essential part of any
-successful service and a necessary component in order to diagnose and detect
-failures, when they occur. This is further complicated by variable propagation
-delays and heterogeneous hardware in dynamic environments.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In the case of Ceph, a consistent view of cluster state among server daemons
-and clients is critical to provide strong consistency guarantees to clients.
-Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
-metadata server) data structures called ``maps'' that record membership and
-status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
-responsible for integrating state changes into cluster maps, responding to
-requests from out-of-date clients and synchronizing members of the cluster
-whenever there is a change in a map so that they all observe the same system
-state. As a fundamental building block of many system designs, consensus
-abstractions such as Paxos are a common technique for maintaining consistent
-data versions, and are a useful system to expose.
-
-The default behavior of the monitor can be seen as a Paxos-based notification
-system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing
-clients to identify when new values (termed epochs in Ceph) are associated to
-given maps.  Since Ceph does not expose this service directly, as part of our Malacology implementation, we expose a key-value
-service designed for managing configuration metadata that is built on top of
-the consensus engine. Since the monitor is intended to be out of the
-high-performance I/O path, a general guideline is to make use of this
-functionality infrequently and to assign small values to maps.\\
-
-\noindent{\it{\textbf{Malacology: }}} Malacology exposes a strongly-consistent view of
-time-varying service metadata as a service rather than a hidden internal
-component. Malacology provides a generic API for adding arbitrary values to
-existing subsystem cluster maps. As a consequence of this, applications can
-define simple but useful service-specific logic to the strongly-consistent
-interface, such as authorization control (just specific clients can write new
-values) or triggering actions based on specific values (e.g. sanitize values).
-The higher level services we implement in \S\ref{sec:services} make use of this
-functionality to register, version and propagate dynamic code (Lua scripts) for
-new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
-policies in the load balancer \S\ref{malacology:mds}.  Using this service
-guarantees that interface definitions are not only made durable, but are
-transparently and consistently propagated throughout the cluster so that
-clients are properly synchronized with the latest interfaces.
-
-\newcomment{\\ \noindent \it{\textbf{Impact: }} provides core functionality
-because it lets daemons come to consensus on system critical state. Bugs in the
-internal subsystems or ommitting this from services that need this type of
-consistency affects correctness.}
-\addressesissue{3, 5}
-
-%And finally, it is important to not underestimate the importance of the
-%preservation of the modifications that enable a new system service. In many
-%(if not all) cases, interfaces defining access to data are just as important
-%as the data itself by virtue of inherently providing structural context. Thus,
-%all components of a system service must be kept to the same standards of
-%protection as data in the system itself.
-
-\oldcomment{\noindent 4.2 Object I/O Intefrace}
-\subsection{Data I/O Interface}
-\label{sec:data-io-interface}
-\label{object-data-interface}
-Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
-supports application-specific object
-interfaces~\cite{weil_rados_2007}. The ability to offload computation can
-reduce data movement, and transactional interfaces can significantly simplify
-construction of complex storage interfaces that require uncoordinated parallel
-access.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-An object interface is a plugin structured similarly to that of an RPC
-in which a developer creates a named function within the cluster that clients
-may invoke. In the case of Ceph each function is implicitly run within the
-context of an object specified when the function is called. Developers of
-object interfaces express behavior by creating a composition of native
-interfaces or other custom object interfaces, and handle serialization of
-function input and output.  A wide range of native interfaces are available to
-developers such as reading and writing to a byte stream, controlling object
-snapshots and clones, and accessing a sorted key-value database. These native
-interfaces may be transactionally composed along with application
-specific logic to create semantically rich interfaces. An example would be
-an interface that atomically updates a matrix stored in the bytestream and an
-index of the matrix stored in the key-value database.\\
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation.png}
-\caption{Malacology allows users to dynamically define object/metadata
-interfaces by extending the object storage daemon (OSD) and metadata server
-(MDS) subsystems with an embedded Lua VM.  It uses the service metadata
-interface (M) to propagate interfaces/versions across the cluster
-\label{fig:implementation}}
-\end{figure}
-
-\noindent {\it{\textbf{Malacology: }}} The implementation of Ceph's object abstraction,
-although powerful, does not readily support programmability. Supporting only
-C/C++ for object interface developers, Ceph requires distribution of compiled
-binaries for the correct architecture, adding a large barrier of entry for
-developers and system administrators. Second, having no way to dynamically
-unload modules, any changes require a full restart of a storage daemon which
-may have serious performance impacts due to loss of cached data. And finally,
-the security limitations of the framework limit the use of object interfaces to
-all but those with administrative level access and deep technical expertise.
-
-To address these concerns, Malacology takes advantage of Lua extensions
-contributed by the Ceph community. This allows new object interfaces to be
-dynamically loaded into the system and modified at runtime, resulting in a
-object storage API with economy of expression, which at the same time provides
-the full set of features of the base object class. New object interfaces that
-are expressed in thousands of lines of code can be implemented in approximately
-an order of magnitude less code~\cite{geambasu_comet_2010}. While the use of
-Lua does not prevent deployment of malicious code, certain types of coding
-mistakes can be handled gracefully, and access policies are used to limit access
-to trusted users~\cite{ierusalimschy1996lua}.
-
-\oldcomment{All of these modifications work in tandem to implement the desired behavior,
-and are typically co-designed together such that each depends on the other to
-behave as expected.}
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by pushing behavior to lower parts of the storage stack, thereby
-minimizing hops and distributing computation.}
-\addressesissue{3, 5}
-
-\subsection{Distributed Metadata Interfaces}
-\begin{notes}
-\textcolor{red}{This section has been re-organized}
-\end{notes}
-\label{sec:distributed-metadata-interfaces}
-\label{malacology:mds}
-
-\oldcomment{The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction~\cite{weil:sc2004-dyn-metadata}.}
-\newcomment{File systems provide clients with the familiar POSIX file abstraction.
-While this guarantees strong consistency it comes at the cost of scalability,
-increased complexity, and lower performance.}
-\addressesissue{1}
-In general, distributed file systems protect resources by providing
-hierarchical indexing and distributed locking services.
-
-%The metadata cluster moves units of the namespace called directory fragments
-%and assigns them to servers using its own metadata load balancer. It then uses
-%a hard-coded policy to balance these fragments using a scalarized load metric
-%based on CPU, workload, and file system operation metrics. Although the CephFS
-%balancer has been shown to be complicated, its complexity is justifiable.
-%Having the flexibility to choose what to move, where to move it, and how much
-%to move is very powerful. The problem is that it is difficult to pick a
-%one-size-fits-all policy, system administrators may want to partition load
-%based on SLAs, or applications are better equipped to make these decisions,
-%etc. -- but we choose instead to expose a balancing API at strategic points in
-%the balancing logic using Lua hooks. ~\cite{sevilla:sc15-mantle}.
-
-\oldcomment{\it{\textbf{Malacology:}} Interfaces are added in strategic spots in the
-metadata service for guarding shared resources, defining new file types,
-and load balancing.}
-
-\oldcomment{\noindent Shared resource interface.}
-\subsubsection{Shared Resource Interface}
-
-In general, file system metadata servers manages client sessions, allowing
-clients to obtain locks (e.g. file byte ranges), and capabilities (e.g. to
-cache file data). Clients and metadata servers use a cooperative protocol in
-which clients voluntarily release resources back to the metadata service in
-order to implement sharing policies.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In Ceph, the locking service implements a capability-based system that
-expresses what data and metadata clients are allowed to access as well as what
-state they may cache and modify locally.  While designed for the file
-abstraction, indexing, locking, and caching are all common services that are
-useful to a broad spectrum of applications.  Distributed applications that
-share centralized resources (e.g. a database or directory) face similar
-challenges which are often solved using application-specific sharding.
-
-\newcomment{\\ \noindent\it{\textbf{Malacology: }}}
-While the current policy for sharing access and voluntarily releasing resources
-is largely best-effort, Malacology supports generalized policies between
-metadata servers and clients that can be used to implement fairness or
-priority.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} provides core functionality
-to protect and provide exclusive access for any shared resource. May hurt
-performance if the resource in question does not require strong consistency.}
-\addressesissue{3}
-
-\oldcomment{\noindent File type interface.}
-\subsubsection{File Type Interface}
-The metadata service manages a POSIX file system hierarchy where files and
-directories are represented as inode data structures and expose a POSIX file
-interface. Applications that manage large amounts of metadata (e.g. users or
-database snapshots) often require a naming service. We allow new inode types to
-be defined such that applications can create domain-specific interfaces to
-inodes that may modify locking and capability policies. We will show how this
-is used in Section~\ref{sec:seq} when we discuss a distributed shared-log built
-on Malacology.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this interface is both a feature
-and a performance optimization. It is a feature because it allows developers to
-add support for different  storage types, such as how to read new file formats
-or what consistency semantics to use for a specific subtree in the hierarchical
-namespace. It is also a performance optimization because future programmers can
-add optimizations for processing specific types of files into the inode itself.}
-\addressesissue{3}
-
-%%% TODO: striping strategy is stored with the inode
-
-\oldcomment{\noindent \bf Load balancing interface.} 
-\subsubsection{Load Balancing Interface}
-
-Ceph also addresses the challenge of balancing metadata load with a separate
-metadata cluster. This cluster uses load balancing policies to migrate
-directory inodes around the cluster to alleviate overload at single
-nodes~\cite{weil:sc2004-dyn-metadata}. The policies use metrics based on system
-state (e.g.  CPU and memory utilization) and statistics collected by the
-cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
-partitioning to move variable sized namespace subtrees. These units can be
-shipped anywhere (i.e., to any metadata server of any capacity) at any time for
-any reason. The original balancer was designed with hard-coded policies and
-tunables.\\
-
-\noindent {\it{\textbf{Malacology: }}}
-Mantle~\cite{sevilla:sc15-mantle}, a programmable metadata load balancer, has
-been implemented on Malacology so that its infrastructure can enjoy all the
-properties of existing internal abstractions.  The load balancing logic is
-intact but the infrastructure has been improved to safely load, manage, and
-persist its load balancing policies.  Composing Mantle with the service
-metadata and durability interfaces, Mantle can safely version balancer
-policies, save balancer policies in the backend object store and centralize
-warnings/errors. When combined with file types, load balancing policies can
-express a variety of rules such as multi-tenant and multi-application
-balancing.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by allowing them to specify how to partition, replicate, and
-distribute metadata in response to overloaded servers.}
-\addressesissue{3}
-
-\subsection{Durability Interface}
-\label{sec:durability}
-
-Ceph provides storage by striping and replicating data across
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-protects data using common techniques such as erasure
-coding, replication, and scrubbing.  Furthermore, many of these features
-are implemented using a peer-to-peer protocol that allows object storage
-devices to operate autonomously without a centralized coordinator.  For
-example, when the number of placement groups change, the OSDs re-balance and
-re-shard data in the background in a process called placement group splitting
-in which OSDs communicate directly with each other to cover a new state.
-In order to reduce load on the monitoring service, Ceph OSDs use a gossip
-protocol to efficiently propagate changes to cluster maps throughout the
-system, and autonomously initiate recovery mechanisms when failures are
-discovered.\\
-
-\noindent {\it{\textbf{Malacology: }}} Metadata service policies and object storage
-interfaces are stored durability within RADOS and managed by storing references
-with the object maps. Since the cluster already propagates a consistent view of
-these data structures, we use this service to automatically install interfaces
-in OSDs, and install policies within the metadata server daemon (MDS) such that clients and daemons are
-synchronized on correct implementations without restarting.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this is a feature because it
-adds data safety and persistence to system metadata; while nice to have it does
-effect correctness.}
-\addressesissue{3}
-
-\section{Services Built on Malacology}
-\label{sec:services}
-\label{services-built-on-malacology}
-
-\label{services}
-%\begin{table}
-%\centering\small
-%\begin{tabular}{  l | l | l }
-%\textbf{Service}              &
-%\textbf{Used to...}           &
-%\textbf{Malacology Interface} \\ \hline
-%Mantle & load balance metadata     & Load Balancing  \\
-%       & agree on current balancer & Service Metadata\\
-%       & control client write lock & Shared Resource \\
-%       & persist policy logic      & Durability      \\ \hline
-%ZLog   & implement CORFU           & Object \& Data  \\ 
-%       & install ZLog interfaces   & Service Metadata\\
-%       & persist log entries       & Durability      \\ \hline
-%ZLog   & load balance sequencers   & Load Balancing  \\
-%Seqr.  & manage access to log tail & Shared Resource \\ 
-%\end{tabular}
-%\caption{Mantle and ZLog are built using Malacology interfaces so they inherit the
-%robustness of the subsystems of the underlying stack.}
-%\label{table:implementation}
-%\end{table}
-
-In this section we describe two services built on top of Malacology. The first
-is Mantle, a framework for dynamically specifying metadata load balancing
-policies. The second system, ZLog, is a high-performance distributed shared-log.
-In addition to these services, we'll demonstrate how we combine ZLog and Mantle
-to implement service-aware metadata load balancing policies.
-
-\subsection{Mantle: Programmable Load Balancer}
-\label{sec:mantle}
-
-Mantle is a programmable load balancer that separates the metadata balancing
-policies from their mechanisms. Administrators inject code to change how the
-metadata cluster distributes metadata. In ~\cite{sevilla:sc15-mantle} the
-authors showed how using Mantle, one can implement a single node metadata
-service, a distributed metadata service with hashing, and a distributed
-metadata service with dynamic subtree partitioning.
-
-The original implementation was ``hard-coded" into Ceph and lacked robustness
-(no versioning, durability, or policy distribution).  Re-implemented using
-Malacology, Mantle now enjoys (1) the versioning provided by the metadata
-service interface provided by the monitor daemons and (2) the durability and
-distribution provided by Ceph's reliable object store.  Re-using the internal
-abstractions with Malacology resulted in a 2\(\times\) reduction in source code
-compared to the original implementation.
-
-\subsubsection{Versioning Balancer Policies}
-
-Ensuring that the version of the current load balancer is consistent across the
-physical servers in the metadata cluster was not addressed in the original
-implementation. The user had to set the version on each individual server and
-it was trivial to make the versions inconsistent. Maintaining consistent
-versions is important for balancing policies that are cooperative in which local
-decisions are made assuming properties about other instances of the same
-policy.
-
-With Malacology, Mantle stores the version of the current load balancer in the
-Ceph service metadata interface. The version of the load balancer corresponds
-to an object name in the balancing policy. Using the service metadata interface
-ensures that all metadata servers use the same version of the load balancer. As
-a result, the policy version inherits the consistency of Ceph's internal
-monitor daemons. The user changes the version of the load balancer using a new
-CLI command.
-
-\subsubsection{Making Balancer Policies Durable}
-
-The load balancer version described above corresponds to the name of an object
-in RADOS that holds the actual Lua balancing code.  When MDS nodes start
-balancing load, they first check the latest version from the MDS map and
-compare it to the balancer they have loaded. If the version has changed, they
-dereference the pointer to the balancer version by reading the corresponding
-object in RADOS. This is in contrast to the original balancer which stored load
-balancer code on the local file system -- a technique which is unreliable and
-may result in silent corruption.
-
-The balancer pulls the Lua code from RADOS synchronously; asynchronous reads
-are not possible because of the architecture of the MDS. The synchronous
-behavior is not the default behavior for RADOS operations, so we achieve this
-with a timeout: if the asynchronous read does not come back within half the
-balancing tick interval the operation is canceled and a Connection Timeout
-error is returned. By default, the balancing tick interval is 10 seconds, so
-Mantle will use a 5 second second timeout.
-
-This design allows Mantle to immediately return an error if anything
-RADOS-related goes wrong.  We use this implementation because we do not want to
-do a blocking OSD read from inside the global MDS lock. Doing so would bring
-down the MDS cluster if any of the OSDs are not responsive.
-
-Storing the balancers in RADOS is simplified by the use of an interpreted
-language for writing balancer code. If we used a language that needs to be
-compiled, like the C++ object classes in the OSD, we would need to ensure
-binary compatibility, which is complicated by different operating systems,
-distributions, and compilers.
-
-\subsubsection{Logging, Debugging, and Warnings}
-
-In the original implementation, Mantle would log all errors, warnings, and
-debug messages to a log stored locally on each MDS server. To get the simplest
-status messages or to debug problems, the user would have to log into each MDS
-individually, look at the logs, and reason about causality and ordering.
-
-With Malacology, Mantle re-uses the centralized logging features of the
-monitoring service. Important errors, warnings, and info messages are collected
-by the monitoring subsystem and appear in the monitor cluster log so instead of
-users going to each node, they can watch messages appear at the monitor.
-Messages are logged sparingly, so as not to spam the monitor with frivolous
-debugging but important events, like balancer version changes or failed
-subsystems show up in the centralized log.
-
-\subsection{ZLog: A Fast Distributed Shared Log}
-\label{sec:zlog}
-
-The second service implemented on Malacology is ZLog, a high-performance
-distributed shared-log that is based on the CORFU
-protocol~\cite{balakrishnan_corfu_2012}. The shared-log is a
-powerful abstraction used to construct distributed systems, such as
-metadata management~\cite{balakrishnan:sosp13} and elastic database
-systems~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
-However, existing implementations that rely on consensus algorithms such as
-Paxos funnel I/O through a single point introducing a bottleneck that restricts
-throughput. In contrast, the CORFU protocol is able to achieve high throughput
-using a network counter called a \emph{sequencer}, that
-decouples log position assignment from log I/O.
-
-While a full description of the CORFU system is beyond the scope of this
-paper, we briefly describe the custom storage device interface, sequencer
-service, and recovery protocol, and how these services are instantiated in the
-Malacology framework.
-
-\subsubsection{Sequencer}
-\label{sec:seq}
-
-High-performance in CORFU is achieved using a sequencer service that assigns
-log positions to clients by reading from a volatile, in-memory counter which
-can run at a very high throughput and at low latency. Since the sequencer is
-centralized, ensuring serializability in the common case is trivial.  The
-primary challenge in CORFU is handling the failure of the sequencer in a way
-that preserves correctness. Failure of the sequencer service in CORFU is
-handled by a recovery algorithm that recomputes the new sequencer state using a
-CORFU-specific custom storage interface to discover the tail of the log, while
-simultaneously invalidating stale client requests using an epoch-based
-protocol.
-
-{\bf Sequencer interface.} The sequencer resource supports the ability to
-\emph{read()} the current tail value and get the \emph{next()} position in
-the log which also atomically increments the tail position.
-We implement the sequencer service in Malacology as an application-specific
-file type that associates a small amount of metadata
-with the inode state managed by the Ceph metadata service.
-This approach has the added
-benefit of allowing the metadata service to handle naming, by
-representing each sequencer instance in the standard POSIX hierarchical
-namespace. The primary challenge in mapping the sequencer resource to the
-metadata service is handling serialization correctly to maintain the global
-ordering provided by the CORFU protocol.
-
-Initially we sought to directly model the sequencer service in Ceph as a
-non-exclusive, non-cacheable resource, forcing clients to perform a round-trip
-access to the resource at the authoritative metadata server for the sequencer
-inode.  Interestingly, we found that the capability system in Ceph uses a
-strategy to reduce metadata service load by allowing clients that open a
-shared file to temporarily obtain an exclusive cached copy of the resource,
-resulting in a round-robin, best-effort batching behavior. When a single
-client is accessing the sequencer resource it is able to increment the
-sequencer locally. Any competing client cannot query the sequencer until the
-metadata service has granted it access.
-
-While unexpected, this discovery allowed us to explore an implementation
-strategy that we had not previously considered. In particular, for bursty
-clients, and clients that can tolerate increased latency, this mode of
-operation may allow a system to achieve much higher throughput than a system
-with a centralized sequencer service.  We utilize the programmability of the
-metadata service to define a new policy for handling capabilities that controls
-the amount of time that clients are able to cache the sequencer resource. This
-allows an administrator or application to control the trade-off between latency
-and throughput beyond the standard best-effort policy that is present in Ceph
-by default.
-
-In Section~\ref{sec:evaluation} we quantify the trade-offs of throughput and
-latency for an approach based on a round-robin batching mode, and compare this
-mode to one in which the metadata server mediates access to the sequencer state
-when it is being shared among multiple clients. Quantifying these trade-offs
-should provide administrators with guidelines for setting the tunables for
-different ``caching" modes of the sequencer.
-
-{\bf Balancing policies.} As opposed to the batching mode for controlling access to the sequencer
-resource, more predictable latency can be achieved by treating the sequencer
-inode as a shared non-cacheable resource, forcing clients to make a round-trip
-to the metadata service. However, the shared nature of the metadata service
-may prevent the sequencer from achieving maximum throughput. To provide a
-solution to this issue we have taken advantage of the programmability of the
-metadata service load balancing to construct a service-specific load balancing
-policy. As opposed to a balancing policy that strives for uniform
-load distribution, a ZLog-specific policy may utilize knowledge of inode types
-to migrate the sequencer service to provisioned hardware during periods of
-contention or high demand.
-
-\subsubsection{Storage Interface}
-
-The storage interface is a critical component in the CORFU protocol. Clients
-independently map log positions that they have obtained from the sequencer
-service (described in detail in the next section) onto storage devices, while
-storage devices provide an intelligent \emph{write}-once, random \emph{read}
-interface for accessing log entries. The key to correctness in CORFU lies with
-the enforcement of up-to-date epoch tags on client requests; requests tagged
-with out-of-date epoch values are rejected, and clients are expected to
-request a new tail from the sequencer after refreshing state from an auxiliary
-service.  This mechanism forms the basis for sequencer recovery.
-
-In order to repopulate the sequencer state (i.e. the cached, current tail of
-the log) during recovery of a sequencer, the maximum position in the log must
-be obtained. To do this, the storage interface exposes an additional
-\emph{seal} method that atomically installs a new epoch value and returns the
-maximum log position that has been written.
-
-Since the sequencer service does not resume until the recovery process has
-completed, there cannot be a race with clients appending to the log, and the
-immutability of the log allow reads to never block during a sequencer failure.
-Recovery of the sequencer process itself may be handled in many ways, such as
-leader election using an auxiliary service such as Paxos. In our
-implementation, the recovery is the same as (and is inherited from) the Ceph
-metadata service. Handling the failure of a client that holds the sequencer
-state is similar, although a timeout is used to determine when a client should
-be considered unavailable.
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-%We wanted to implement CORFU and we had two choices: build it from the ground
-%up or do it on top of Malacology. To show the feasability of building on
-%Malacology our evaluation focuses on the
-
-Our evaluation demonstrates the feasibility of building new service
-abstractions atop programmable storage, focusing on the performance of the
-internal abstractions exposed by Malacology and used to construct the Mantle
-and ZLog services. We also discuss latent capabilities we discovered in this
-process that let us navigate different trade-offs within the services
-themselves. First, we benchmark scenarios with high sequencer contention by
-examining the interfaces used to map ZLog onto Malacology; specifically, we
-describe the sequencer implementation and the propagation of object and data
-interfaces interfaces.  Next, we benchmark scenarios in which the storage
-system manages multiple logs by using Mantle to balance sequencers across a
-cluster.
-
-% should save this for future work, otherwise it undercuts the eval.
-%Detailed system-wide performance measurements and analysis is forthcoming for
-%both Mantle and ZLog, as we try to find the best metadata balancing policies
-%and physical design implementation trade-offs.
-
-Since this work focuses on the programmability of Malacology, the goal of this
-section is to show that the components and subsystems that support the
-Malacology interfaces provide reasonable relative performance, as well as to
-give examples of the flexibility that Malacology provides to programmers.
-
-%As we search for the best metadata balancing policies
-%and physical design implementation trade-offs,
-%we plan to rel
-
-\subsection{Mapping ZLog onto Malacology}
-\label{sec:mapping-zlog-onto-malacology}
-
-We evaluate Malacology by exploring one possible mapping of the ZLog
-implementation of CORFU onto Ceph in which we re-use (1) the metadata service
-to manage naming and synchronization of the sequencer resource by treating the
-resource as an inode, and (2) the monitoring sub-subsystem to distribute and
-install application-specific I/O interface required of the CORFU protocol. In
-Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the inode
-abstraction for implementing the sequencer resource enables load balancing policies
-to migrate the sequencer resource in heavy-load situation.
-
-\subsubsection{Sequencer Implementation}
-\label{sec:sequencer-implementation}
-
-We evaluate the feasibility of using the metadata service to implement a
-sequencer resource that is responsible for maintaining a total ordering of
-the log. Clients contact the sequencer to obtain the tail of the log and then
-independently initiate I/O, thus we measure both the throughput and latency of
-obtaining new tail positions which bounds client append performance.
-
-The sequencer is implemented as a custom file type in which the sequencer
-state (a 64-bit integer) is embedded in the inode of a file representing the
-sequencer. A total ordering of the log is imposed by the re-use of the
-capability service that can be used to grant exclusive access of inode state
-to clients. The metadata service is responsible for maintaining exclusivity
-and granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
-behavior of the system in which a best-effort policy is used.  The two colors
-represent points in time that the clients were able to access the resource.
-The best-effort policy shows a high degree of interleaving
-between clients but the system spends a large portion of time re-distributing
-the capability, reducing overall throughput.
-
-In order to control the performance of the system we implement a
-policy that (1) restricts the length of time that a client may maintain
-exclusive access and (2) limits the number of log positions that a client may
-generate without yielding to other clients waiting for access. The behavior of
-these two modes is illustrated in Figures~\ref{fig:capdelay-quota-behavior}
-(b) and (c), respectively.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/capdelay-quota-behavior-a.png}
-\includegraphics{figures/capdelay-quota-behavior-b.png}
-\includegraphics{figures/capdelay-quota-behavior-c.png}
-\caption{Each dot is an individual request, spread randomly along the \(y\)
-axis. The default behavior is unpredictable, ``delay" lets clients hold the
-lease longer, and ``quota" gives clients the lease for a number of operations.}
-\label{fig:capdelay-quota-behavior}
-\end{figure}
-
-Figure~\ref{fig:captp} demonstrates a configurable trade-off between
-throughput and latency. In the experiment two clients are run each with a
-fixed 0.25 second maximum reservation on the capability, and we vary the size
-of the log position quota running each configuration for two minutes. The
-total operations per second is the combined throughput of the two clients, and
-the average latency is number of microseconds required to obtain a new log
-position. With a small quota more time is spent exchanging exclusive access,
-while a large quota reservation allows clients to experience a much lower
-latency because they experience uncontended access for a longer period of time.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/tradeoff.png}
-\caption{Sequencer throughput by re-using various services.  The highest performance is
-achieved using a single client with exclusive, cacheable privilege. Round-robin
-sharing of the sequencer resource is affected by the amount of time the
-resource is held, with best-effort performing the worst.}
-\label{fig:captp}
-\end{figure}
-
-To get a better picture of latency, Figure~\ref{fig:capcdf} shows the CDF of
-latency for each client in all experiment configurations. At the 99th
-percentile clients accessed the sequencer is less than a millisecond. The CDF
-is cropped at the 99.999th percentile due to large outliers that we believe
-occur in instances in which the MDS is performing I/O while it is in the
-process of re-distributing the capability to another client.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/caps-delay-latency.png}
-\caption{
-The latency distribution shows that if the clients hold their capabilities longer,
-performance decreases. Changing the delay can help the sequencer and
-clients strike a balance between throughput and latency.}
-\label{fig:capcdf}
-\end{figure}
-
-\textbf{Malacology exposes} the internal capability management service and
-allows users to navigate latency and throughput trade-offs.  Other approaches to
-designing the sequencer service also exist, such as using a centralized service
-in which each access is a network round-trip. In contrast to the mechanism we
-explored which is appropriate for clients with bursty workloads, it may be
-easier to provide predictable performance using a centralized service, and we
-will be exploring in the future how this can be achieved using the capability
-system.
-
-\subsubsection{Interface Propagation}
-
-Domain-specific data interfaces (Section~\ref{sec:application-specifc-storage-stacks}) allow
-co-design between applications and the storage system. Malacology supports
-custom object interfaces in RADOS that require interface implementations to
-be installed on the storage devices in the system, supporting the evolution of
-interfaces through automatic system-wide versioning and installation through
-the service metadata interface (Section~\ref{service-metadata}). We evaluate
-the performance of installing a new interface version in the cluster,
-which is an important metric for applications that frequently evolve
-interfaces.
-
-We demonstrate the feasibility of utilizing the Ceph monitoring sub-system by
-evaluating the performance of installing and distributing interface updates.
-Figure~\ref{fig:propdelay} shows the CDF of the latency of interface updates.
-The interfaces are Lua scripts embedded in the cluster map and distributed
-using a peer-to-peer gossip protocol.  The latency is defined as the elapsed
-time following the Paxos proposal for an interface update until each OSD makes
-the update live (the cost of the Paxos proposal is configurable and is
-discussed below). The latency measurements were taken on the nodes running
-Ceph server daemons, and thus exclude the client round-trip cost. In each of
-the experiments 1000 interface updates were observed.
-
-Figure~\ref{fig:propdelay} shows the lower bound cost for updates in a large
-cluster. In the experiment labeled ``120 OSD (RAM)" a cluster of 120 OSDs (10
-OSDs x 12 servers) using an in-memory data store were deployed, showing a
-latency of less than 54 ms with a probability of 90\% and a worst case latency
-of 194 ms. These costs demonstrate the penalty of distributing the interface
-in a large cluster. In practice the costs include, in addition to cluster-wide
-propagation of interface updates, the network round-trip to the
-interface management service, the Paxos commit protocol itself, and other factors
-such as system load. By default Paxos proposals occur periodically with a 1
-second interval in order to accumulate updates. In a minimum, realistic quorum
-of 3 monitors using HDD-based storage, we were able to decrease this interval
-to an average of 222 ms.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=60mm,trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
-\caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
-interface commit.}
-\label{fig:propdelay}
-\end{figure}
-
-\subsection{Load Balancing ZLog Sequencers with Mantle}
-\label{sec:zlog-balancing}
-
-In practice, a storage system implementing CORFU will support a multiplicity of
-independent totally-ordered logs for each application.  For this scenario
-co-locating sequencers on the same physical node is not ideal but building a load
-balancer that can migrate the shared resource ({\it e.g.,}, the resource that
-mediates access to the tail of the log) is a time-consuming, non-trivial task.
-It requires building subsystems for migrating resources, monitoring the
-workloads, collecting metrics that describe the utilization on the physical
-nodes, partitioning resources, maintaining cache coherence, and managing
-multiple sequencers. The following experiments demonstrate the feasibility of
-using the mechanisms of the Malacology load balancing interface to inherit
-these features and to alleviate load from overloaded servers.
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-behaviors.png}
-\caption{CephFS/Mantle load balancing have better throughput than co-locating
-all sequencers on the same server.  Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify this improvement;
-Section~\ref{sec:feature-backoff} examines the migration at 0-60 seconds.
-}\label{fig:mantle-balancer-behaviors}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-performance.png}
-\includegraphics{figures/mantle-mode-performance.png}
-\caption{In (a) all CephFS balancing modes have the same performance; Mantle uses a
-balancer designed for sequencers. In (b) the best combination of mode and
-migration units can have up to a 2\(\times\) improvement.}\label{fig:mantle-balancer-performance}
-\end{figure}
-
-The experiments are run on a cluster with 10 nodes to store objects, one node
-to monitor the cluster, and 3 nodes that can accommodate sequencers. 
-Instead of measuring contention at the clients like
-Section~\ref{sec:sequencer-implementation}, these experiments measure
-contention at the sequencers by forcing clients to make round-trips for every
-request. We implement this using the shared resource interface that forces
-round-trips.  Because the sequencer's only function is to  hand out positions
-for the tail of the log, the workload is read-heavy.
-
-First, we show how the ZLog service can orchestrate multiple sequencers using
-the Malacology load balancing interface.
-Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput over time of
-different load balancers as they migrate 3 sequencers (with 4 clients) around
-the cluster; ``No Balancing" keeps all sequencers on one server, ``CephFS"
-migrates sequencers using the CephFS load balancers, and ``Mantle" uses a
-custom load balancer we wrote specifically for sequencers. The CephFS load
-balancers are hard-coded into Ceph while Mantle is an API for balancing load
-presented in~Mantle~\cite{sevilla:sc15-mantle} and re-implemented on top of
-Malacology in this paper.  The increased throughput for the CephFS and Mantle
-curves between 0 and 60 seconds are a result of migrating the sequencer(s) off
-overloaded servers.
-
-In addition to showing that migrating sequencers improves performance,
-Figure~\ref{fig:mantle-balancer-behaviors} also demonstrates features that we
-will explore in the rest of this section.
-Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify the differences in performance
-when the cluster stabilizes at time 100 seconds and
-Section~\ref{sec:feature-backoff} examines the slope and start time of the
-rebalancing phase between 0 and 60 seconds by comparing the aggressiveness of
-the balancers.
-
-\subsubsection{Feature: Balancing Modes}
-\label{sec:feature-balancing-modes}
-
-Next, we quantify the performance benefits shown in
-Figure~\ref{fig:mantle-balancer-behaviors}.  To understand why load balancers
-perform differently we need to explain the different balancing modes that the
-load balancer service uses and how they stress the subsystems that receive and
-forward client requests in different ways. In
-Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
-performance of the balancing mode that CephFS falls into {\it most of the time}.
-CephFS has 3 modes for balancing load: CPU mode, workload mode, and hybrid
-mode. All three have the same structure for making migration decisions but vary
-based on the metric used to calculate load. For this sequencer workload the 3
-different modes all have the same performance, shown in
-Figure~\ref{fig:mantle-balancer-performance} (a), because the load balancer falls
-into the same mode a majority of the time.  The high variation in performance
-for the CephFS CPU Mode bar reflects the uncertainty of using something as
-dynamic and unpredictable as CPU utilization to make migration decisions. In
-addition to the suboptimal performance and unpredictability, because
-CephFS balancers all fall into the same mode despite using different balancers
-is a problem because it prevents administrators from properly exploring the
-balancing state space.
-
-Mantle gives the administrator more control over balancing policies; for
-the Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the load
-balancing interface to program logic for balancing read-heavy workloads,
-resulting in better throughput and stability.  When we did this we also
-identified two balancing modes relevant for making migration decisions
-for sequencers.
-
-Using Mantle, the administrator can put the load balancing service into ``proxy
-mode" or ``client mode". In proxy mode one server receives all requests and
-farms off the requests to slave servers; the slaves servers do the actual tail
-finding operation. In client mode, clients interact directly with the server
-that has their sequencer.  These modes are illustrated in
-Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
-co-located on one physical server -- performance for that mode is shown by the
-No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
-Mode" in the lower left shows how clients for sequencer 2 are re-directed to
-server B. The two ``Proxy mode" diagrams on the right show how clients continue
-sending requests to server A even though sequencer 2 has migrated to server B
-(the meaning of ``(Half)" is discussed in the next section).
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-modes.png}
-\caption{ In client mode clients sending requests to the server that houses
-their sequencer. In proxy mode clients continue sending their requests to the
-first server.  }\label{fig:mantle-modes}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-mode-behavior.png}
-\caption{The performance of proxy mode achieves the highest throughput but at
-the cost of lower throughput for one of the sequencers. Client mode is more
-fair but results in lower cluster throughput.
-}\label{fig:mantle-mode-behavior}
-\end{figure}
-
-Figure~\ref{fig:mantle-mode-behavior} shows the throughput over time of the two
-different modes for an environment with only 2 sequencers (again 4 clients
-each) and 2 servers. The curves for both sequencers in
-Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second and
-at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
-Performance of Sequencer 2 decreases because it stayed on the proxy which now
-handles all client requests, processes requests for Sequencer 2, and forwards
-requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
-because distributing the sequencers in this way separates (1) the handling of
-the client requests and (2) finding the tail of the log and responding to
-clients.  Doing both steps is too heavy weight for one server and sequencers on
-slave nodes can go faster if work is split up; this phenomenon is not
-uncommon and has been observed in chain replication~\cite{chain_rep}.
-
-Cluster
-throughput improves at the cost of decreased throughput for Sequencer 2.
-Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer mode manually (no
-balancing phase) and shows that the cluster throughput is worse than the
-cluster throughput of proxy mode. That graph also shows that Sequencer 2 has
-less throughput than Sequencer 2. In this case, the scatter-gather process used
-for cache coherence in the metadata protocols causes strain on the server
-housing Sequencer 2 resulting in this uneven performance.
-
-\subsubsection{Feature: Migration Units}
-\label{sec:feature-migration-units}
-
-Another factor that affects performance in this environment is how much load
-is on each server; these experiments quantify that effect by programming the
-load balancing interface to control the amount of load to migrate. We call this
-metric a ``migration unit".  Expressing this heuristic is not easily achievable
-with outward facing tunable parameters (i.e. system knobs) but with Mantle's
-programmable interface, we can force the load balancer to change its migration
-units. To force the balancer into the Proxy Mode (Half) scenario in
-Figure~\ref{fig:mantle-modes}, which uses migration units equal to half the
-load on the current server, we can use: \texttt{ targets[whoami+1] = mds[whoami]["load"]/2
-}.
-
-This code snippet uses globally defined variables and tables from the Mantle
-API to send half of the load on the current server (whoami) to the next ranked
-server (whoami + 1); the \texttt{targets} array is a globally defined table that
-the balancer uses to do the migrations. Alternatively, to migrate all load a
-time step, we can remove the division by 2.
-
-%\begin{figure}[t!]
-%\centering
-%\includegraphics{figures/mantle-mode-performance.png}
-%\caption{The performance of the different modes using various migration units
-%shows almost a 2\(\times\) improvement in the best case.
-%}\label{fig:mantle-mode-performance}
-%\end{figure}
-
-Figure~\ref{fig:mantle-balancer-performance} (b) shows the performance of the modes
-using different migration units. Recall that this setup only has 2 sequencers
-and 2 servers, so performance may be different at scale. Even so, it is clear
-that client mode does not perform as well for read-heavy workloads. We even see
-a throughput improvement when migrating all load off the first server, leaving
-the first server to do administrative tasks (this is common in the metadata
-cluster because the first server does a lot of the cache coherence work) while
-the second server does all the client work. Proxy mode does the best in both
-cases and shows large performance gains when completely decoupling client
-request handling and operation processing in Proxy Mode (Full).  The parameter
-that controls the migration units helps the administrator control the sequencer
-co-location or distribution across the cluster. This trade-off was explored
-extensively in the Mantle paper but the experiments we present here are
-indicative of an even richer set of states to explore.
-
-\subsubsection{Feature: Backoff}
-\label{sec:feature-backoff}
-
-Tuning the aggressiveness of the load balancer decision making is also a
-trade-off that administrators can control and explore. The balancing phase from
-0 to 60 seconds in Figure~\ref{fig:mantle-balancer-behaviors} shows different
-degrees of aggressiveness in making migration decisions; CephFS makes a
-decision quickly and throughput jumps to 2500 ops/second within 10 seconds
-while Mantle takes more time to stabilize. This conservative behavior is
-controlled by programming the balancer to (1) use different conditions for when
-to migrate and (2) using a threshold for sustained overload.
-
-We control the conditions for when to migrate using \texttt{when()}, a callback
-in the Mantle API.  For the Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} we program \texttt{when()} to wait
-for load on the receiving server to fall below a threshold. This makes the
-balancer more conservative because it takes 60 seconds for cache coherence
-messages to settle.  The Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} also takes longer to reach peak
-throughput because we want the policy to wait to see how migrations affect the
-system before proceeding; the balancer does a migration right before 50
-seconds, realizes that there is a third underloaded server, and does another
-migration.
-
-The other way to change aggressiveness of the decision making is to program
-into the balancer a threshold for sustained overload. This forces the balancer
-to wait a certain number of iterations after a migration before proceeding. In
-Mantle, the policy would use the save state function to do a countdown after a
-migration.  Behavior graphs and performance numbers for this backoff feature is
-omitted for space considerations, but our experiments confirm that the more
-conservative the approach the less overall throughput.
-
-\textbf{Malacology pulls} the load balancing service out of the storage system
-to balance sequencers across a cluster. This latent capability also gives
-future programmers the ability to explore the different load balancing
-trade-offs including: load balancing modes to control forwarding vs. client
-redirection, load migration units to control sequencer distribution vs.
-co-location, and backoffs to control conservative vs. aggressive decision
-making.
-
-%\subsubsection{Experiment: Object Class
-%Programmability}\label{experiment-object-class-programmability}
-%
-%Add quote from CORFU paper that talks about what they had to hack
-%
-%\subsubsection{Experiment: Multi-Client
-%Burstiness}\label{experiment-multi-client-burstiness}
-%
-%\begin{figure}[htbp] \centering
-%\includegraphics{figures/caps-delay-thruput.png} \caption{Forcing the client
-%to drop their capabilities later (delay) improves throughput} \end{figure}
-
-\section{Related Work}
-
-% IOStack: Software-Defined Object Storage
-% sRoute: Treating the Storage Stack Like a Network
-   % - lots of good refs to check out
-% Information and control in gray-box systems
-
-Programmability of operating systems and networking resources, including
-distributed storage systems is not new, but we are not aware of work that makes
-generalization of existing services into programmable resources a key principle
-in storage systems design.
-
-Programmable storage systems can be viewed as an infrastructure for creating
-abstractions to better separate policies from mechanisms. This idea is not new.
-Software-defined networks (SDNs) create such an abstraction by separating the
-control plane from the data plane (see for example~\cite{jain:sigcomm13}). This
-notion of control/data separation was also applied in software-defined storage
-(SDS)~\cite{thereska:sosp13,stefanovici:fast16}. Similarly,
-IOStack~\cite{gracia:internet16} is providing policy-based provisioning and
-filtering in OpenStack Swift. According to a SNIA white
-paper~\cite{carlson:snia2014}, the primary goal of SDS is to control and
-facilitate flexible and dynamic provisioning of storage resources of different
-kinds, including flash memory and disk drives, to create a virtualized mapping
-between common storage abstractions (e.g. files, objects, and blocks) and
-storage devices taking data service objectives in terms of protection,
-availability, performance, and security into account. A programmable storage
-system exposes internal abstractions so that end users (not necessarily
-operators) can create new services on top of the storage stack. Thus, our
-notion of programmable storage differs from ``software-defined storage'' (SDS)
-in terms of goals and scope, although definitions of SDS are still in flux.
-
-Another view of programmable storage systems is one of tailoring systems
-resources to applications~\cite{arpaci:sosp01}. Related efforts include the
-Exokernel~\cite{engler:sosp95}, SPIN~\cite{bershad:sosp95} and
-Vino~\cite{seltzer:osdi96} projects; the latter two addressed the ability of injecting
-code into the kernel to specialize resource management. Another approach is to
-pass hints between the different layers of the IO stack to bridge the semantic
-gap between applications and
-storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}).
-
-Malacology uses the same Active and Typed Storage module presented in
-DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File Manifolds
-can be implemented with small changes to the Malacology framework, namely
-asynchronous object calls and Lua stubs in the inode, respectively.
-
-\section{Conclusion and Future Work}
-\label{conclusion-and-future-work}
-
-Programmable storage is a viable method for eliminating duplication of complex
-error-prone software used as workarounds for storage system
-deficiencies. We propose that systems expose their services in a safe way
-allowing application developers to customize system behavior to meet their
-needs while not sacrificing correctness. To illustrate the benefits of this
-approach we presented Malacology, a programmable storage system that
-facilitates the construction of new services by re-purposing existing subsystem
-abstractions of the storage stack.
-
-Future work will focus on constructing more interfaces to support a wide
-variety of storage system services that can be configured on-the-fly in existing
-systems. This work is one point along that path to producing general-purpose
-storage systems that can
-target  special-purpose applications.  Ultimately we want to utilize
-declarative methods for expressing new services.
-
-\bibliography{paper}
-\bibliographystyle{plain}
-%\printbibliography
-
-\end{document}

### 2017-02-23 paper/1_intro.tex
paper: cleanup line breaks and break into separate files
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-02-09 paper/paper.tex
@@
-\documentclass[preprint]{sigplanconf-eurosys}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage{pifont}
-\usepackage{color}
-\usepackage{comment}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Uncomment for diff
-\includecomment{notes}
-\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
-\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
-\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
-
-% Uncomment for camera-ready
-%\excludecomment{notes}
-%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
-%\newcommand{\oldcomment}[1]{}
-%\newcommand{\addressesissue}[1]{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\cmark}{\ding{51}}%
-\usepackage[font=footnotesize,labelfont=bf]{caption}
-
-\setlength{\belowcaptionskip}{-10pt}
-
-\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
-\usepackage[normalem]{ulem}
-\usepackage{pstricks}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-\setkeys{Gin}{width=\maxwidth,keepaspectratio}
-
-\usepackage[unicode=true]{hyperref}
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, Peter Alvaro, Shel Finkelstein, Jeff LeFevre},
-            pdftitle={Malacology: A Programmable Storage System},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=blue,
-            linkcolor=black,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-
-\title{Malacology: A Programmable Storage System}
-%\authorinfo{Paper ID: 165}          {Total Pages: 12}
-\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre}
-           {\{msevilla, jayhawk, ivo, cmaltzahn\}@soe.ucsc.edu, \{palvaro, finkelstein, jlefevre\}@ucsc.edu}
-\date{}
-\begin{document}
-
-\maketitle
-
-\begin{notes}
-\textcolor{red}{
-\noindent Issues raised by reviewers:
-\begin{enumerate}
-  \item clarify what was implemented: diagram of interfaces and prose
-  specifying the difference between Malacology and prior work (reviewers B, C,
-  D, E)
-  \item describe what makes a programmable storage system (reviewers C, D)
-  \item organize motivation: group interfaces into categories (e.g., core
-  functionality, features, and performance optimizations) and describe other
-  applications we can build with Malacology (reviewers B, C)
-  \item clarify terminology: service vs. file system metadata, storage vs.
-  file system, dirty vs. clean slate approaches, soft-state caching (reviewers
-  B, C)
-  \item address security, access control, and the safety; specify how it
-  affects composability, performance, and multi-tenancy (reviewer D)
-\end{enumerate}
-}
-\end{notes}
-
-\begin{abstract} Storage systems are caught between the need to evolve data
-processing applications efficiently and quickly, and the increasing velocity
-with which storage device technology evolves. This puts tremendous pressure on
-storage systems to support rapid change both in terms of their interfaces and
-their performance. But adapting storage systems can be difficult because
-unprincipled changes might jeopardize years of code-hardening and performance
-optimization efforts that were necessary for users to entrust their data to the
-storage system. We introduce the concept of a programmable storage system, a
-system that exposes internal services and abstractions with the goal of
-re-using and building new services on top of the storage stack. We present
-Malacology, a prototype implementation of a programmable storage system to
-explore how existing abstractions of common storage system services can be
-leveraged to adapt to the needs of new data processing systems and the
-increasing variety of storage devices. We show how this approach allows a large
-degree of flexibility for storage systems to evolve without sacrificing the
-robustness of their code-hardened subsystems. We illustrate the advantages and
-challenges of programmability by composing existing primitives into two new
-higher-level services: a metadata load balancer and a high-performance
-distributed shared-log. The evaluation demonstrates that our services inherit
-desirable qualities of the back-end storage system, including the ability to
-balance load, efficiently propagate cluster metadata, recover from failure, and
-to navigate trade-offs between latency and throughput using leases.
-\end{abstract}
+\section{Introduction}
+\label{introduction}
+\label{sec:intro}
 
@@
 \end{figure}
-
-\section{Introduction}
-\label{introduction}
-\label{sec:intro}
 
@@
 as well as flexible structured data
-representations.~\cite{apache_contributors_parquet_2014}.  Finally, production-quality scalable
-storage systems available as open source software have established and are
-continuing to establish new, \emph{de-facto} API standards at a faster pace
-than traditional standards
+representations.~\cite{apache_contributors_parquet_2014}.  Finally,
+production-quality scalable storage systems available as open source software
+have established and are continuing to establish new, \emph{de-facto} API
+standards at a faster pace than traditional standards
 bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
@@
 of new services via composition.  A programmable storage system can be realized
-by exposing existing functionality (such as metadata services and
-synchronization and monitoring capabilities) as interfaces that can be ``glued
-together'' in a variety of ways using a high-level language. Programmable
-storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection
-and execution of code within a storage system or storage device---in that the
-former is applicable to any component of the storage system, while the latter
-focuses at the data access level. Given this contrast, we can say that active
-storage is an example of how one internal component (the storage layer) is
-exposed in a programmable storage system.
+by exposing existing functionality (such as \newcommentone{file system and
+cluster} metadata services and synchronization and monitoring capabilities) as
+interfaces that can be ``glued together'' in a variety of ways using a
+high-level language. Programmable storage differs from \emph{active
+storage}~\cite{riedel:vldb98}---the injection and execution of code within a
+storage system or storage device---in that the former is applicable to any
+component of the storage system, while the latter focuses at the data access
+level. Given this contrast, we can say that active storage is an example of how
+one internal component (the storage layer) is exposed in a programmable storage
+system.  \addressesissue{4}
 
@@
 storage because it offers a broad spectrum of existing services, including
-distributed locking and caching services provided by metadata servers,
-durability and object interfaces provided by the backend object store, and
-propagation of consistent cluster state provided by the monitoring service (see
-Figure~\ref{fig:overview}).  Malacology is expressive enough to provide the
-functionality necessary for implementing new services.
+distributed locking and caching services provided by \newcommentone{file
+system} metadata servers, durability and object interfaces provided by the
+backend object store, and propagation of consistent cluster state provided by
+the monitoring service (see Figure~\ref{fig:overview}).  Malacology is
+expressive enough to provide the functionality necessary for implementing new
+services.  \addressesissue{4}
 
@@
 
-{\it \textbf{Contribution 2:}} We implement two distributed services using Malacology to
-demonstrate the feasibility of the programmable storage approach:
+{\it \textbf{Contribution 2:}} We implement two distributed services using
+Malacology to demonstrate the feasibility of the programmable storage approach:
 
@@
 motivate the need for programmable storage by describing current practices in
-the open source software community. Next we describe Malacology by presenting the
-subsystems within the underlying storage system that we re-purpose, and briefly
-describe how those system are used within Malacology (\S\ref{sec:malacology}).
-Then we describe the services that we have constructed within the Malacology
-framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
-implementation (\S\ref{sec:evaluation}).  We conclude by discussing related and future work.
-% and
-%conclude.
-
-\section{Application-Specific Storage Stacks}
-\label{sec:application-specifc-storage-stacks}
-
-Building storage stacks from the ground up for a specific purpose results in
-the best performance. For example, GFS~\cite{ghemawat:sosp03} and
-HDFS~\cite{shvachko:msst10} were designed specifically to serve Map/Reduce and
-Hadoop jobs, and use techniques like exposing data locality and relaxing POSIX
-constraints to achieve application-specific I/O optimizations. Another example
-is Boxwood~\cite{maccormick:osdi04}, which experimented with B-trees
-and chunk stores as storage abstractions to simplify application building.
-Alternatively, general-purpose storage stacks are built with the flexibility to
-serve many applications by providing a variety of interfaces and tunable
-parameters.  Unfortunately, managing competing forces in these systems is
-difficult and users want more control from the general-purpose storage stacks
-without going as far as building their storage system from the ground up.
-
-To demonstrate a recent trend towards more application-specific storage systems
-we examine the state of programmability in Ceph. Something of a storage Swiss
-army knife, Ceph simultaneously supports file, block, and object interfaces on
-a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
-Distributed Object Storage (RADOS) system is a cluster of object storage
-devices (OSDs) that provide Ceph with data durability and integrity using
-replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already
-provides some degree of programmability; the OSDs support domain-specific
-object interfaces that are implemented by composing existing low-level storage
-interfaces that execute atomically. These interfaces are written in C++ and are
-statically loaded into the system.
-
-The Ceph community provides empirical evidence that developers are already
-beginning to embrace programmable storage. Figure~\ref{fig:obj-int-dev-growth}
-shows a dramatic growth in the production use of domain-specific interfaces in
-the Ceph community since 2010. In that figure, classes are functional groupings
-of methods on storage objects (e.g. remotely computing and caching the checksum
-of an object extent).  What is most remarkable is that this trend contradicts
-the notion that API changes are a burden for users.  Rather it appears that
-gaps in existing interfaces are being addressed through ad-hoc approaches to
-programmability. In fact, Table~\ref{table:objclasses} categorizes existing
-interfaces and we clearly see a trend towards reusable services.
-
-\begin{figure}[ht]
-\centering
-\includegraphics{figures/obj-int-dev-growth.png}
-\caption{Since 2010, the growth in the number of co-designed object storage interfaces
-in Ceph has been accelerating. This plot is the number of object classes (a
-group of interfaces), and the total number of methods (the actual API
-end-points).}
-\label{fig:obj-int-dev-growth}
-\end{figure}
-
-\begin{table}[ht]
-\centering
-  \begin{tabular}{l|l|l}
-    Category & Example & \# \\ \hline
-    Logging  & Geographically distribute replicas & 11 \\ \hdashline
-    Metadata & Snapshots in the block device OR  & \multirow{2}{*}{74} \\
-    Management & Scan extents for file system repair & \\ \hdashline
-    Locking  & Grants clients exclusive access & 6 \\ \hdashline
-    Other & Garbage collection, reference counting  & 4\\
-\end{tabular}
-\caption{A variety of object storage classes exist to expose interfaces
-    to applications. \# is the number of methods that implement these categories.
-}
-\label{table:objclasses}
-\end{table}
-
-The takeaway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
-already trying to use programmability because their needs, whether they be
-related to performance, availability, consistency, convenience, etc., are not
-satisfied by the existing default set of interfaces. The popularity of the
-custom object interface facility of Ceph could be due to a number of reasons,
-such as the default algorithms/tunables of the storage system being
-insufficient for the application's performance goals, programmers wanting to
-exploit application-specific semantics, and/or programmers knowing how to
-manage resources to improve performance. A solution based on
-application-specific object interfaces is a way to work around the
-traditionally rigid storage APIs because custom object interfaces give
-programmers the ability to tell the storage system about their application: if
-the application is CPU or IO bound, if it has locality, if its size has the potential to
-overload a single node, etc.  Programmers often know what the problem is
-and how to solve it, but until the ability to modify object interfaces, they
-had no way to express to the storage system how to handle their data.
-
-Our approach is to expose more of the commonly used, code-hardened subsystems
-of the underlying storage system as interfaces. The intent is that these
-interfaces, which can be as simple as a redirection to the persistent data
-store or as complicated as a strongly consistent directory service, should be
-used and re-used in many contexts to implement a wide range of services. By
-making programmability a `feature', rather than a `hack' or `workaround', we
-help standardize a development process that now is largely ad-hoc.
-
-\section{Challenges}
-\label{sec:challenges}
-
-Implementing the infrastructure for programmability into existing services and
-abstractions of distributed storage systems is challenging, even if one assumes
-that the source code of the storage system and the necessary expertise for
-understanding it is available.  Some challenges include:
-
-\begin{itemize}
-
-\item Storage systems are generally required to be highly available so that any
-complete restarts of the storage system to reprogram them is usually
-unacceptable.
-
-\item Policies and optimizations are usually hard-wired into the services and
-one has to be careful when factoring them to avoid introducing additional bugs.
-These policies and optimizations are usually cross-cutting solutions to
-concerns or trade-offs that cannot be fully explored at the time the code is
-written (as they relate to workload or hardware). Given these policies and
-optimizations, decomposition of otherwise orthogonal internal abstractions can
-be difficult or dangerous.
-
-\item Mechanisms that are often only exercised according to hard-wired policies
-and not in their full generality have hidden bugs that are revealed as soon as
-those mechanisms are governed by different policies. In our experience
-introducing programmability into a storage system proved to be a great
-debugging tool.
-
-\item Programmability, especially in live systems, implies changes that need to
-be carefully managed by the system itself, including versioning and propagation
-of those changes without affecting correctness.
-
-\end{itemize}
-
-To address these challenges we present Malacology
-\oldcomment{. Malacology is both a prototype for a programmable storage system
-and a}
-\newcomment{, our prototype programmable storage system. It uses the
-programmable storage}
-\addressesissue{2}
-design approach to evolve storage systems efficiently and
-without jeopardizing years of code-hardening and performance optimization
-efforts.  Although Malacology uses the internal abstractions of the underlying
-storage system, including its subsystems, components, and implementations, we
-emphasize that our system still addresses the general challenges outlined
-above.
-
-The main challenge of designing a programmable storage system is choosing the
-right internal abstractions and picking the correct layers for exposing them.
-\oldcomment{In this paper, we do not present an exhaustive list of possible
-internal abstractions nor do we contend that the abstractions we choose provide
-the best trade-offs for all applications.} 
-\newcomment{A programmable storage is not defined by what abstractions are
-exposed, rather a programmable storage system adheres to the design approach of
-choosing interfaces so administrators can have better control of the storage
-system. The interfaces presented in this paper are abstractions that we found
-useful for building ZLog and Mantle, yet they may not provide the best
-trade-offs for all higher-level services.}
-\addressesissue{2}
-For example, if consensus is correctly exposed one could implement
-high-level features like versioning, serialization, or various flavors of
-strongly consistent data management on top; but perhaps a low-level consensus
-interface is suited well for a particular set of applications.  These questions
-are not answered in this paper and instead we focus on showing the feasibility
-of building such a system, given advances in the quality and robustness of
-today's storage stacks.
-
-The Malacology prototype we present has been implemented on Ceph.
-While there are other systems on top of which Malacology could be implemented (see Table~\ref{table:examples}), we choose Ceph because it is a production quality system and because it
-is open source. The large developer community ensures that code is
-robust and the visibility of the code lets us expose any interface we want. In
-the next section we describe the Ceph components that we expose as Malacology
-interfaces.
-
-\section{Malacology: A Programmable Storage System}
-\label{sec:malacology}
-
-\begin{table*}
-\centering\small
-\begin{tabular}{ l | c | l | l | l }
-%\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
-\multicolumn{4}{c}{} \\
-\textbf{Interface}                      &
-\textbf{Sec}                            &
-\textbf{Example in Production Systems}  &
-\textbf{Example in Ceph}                &
-\textbf{Provided Primitives}            \\ \hline
-Service Metadata
-  & \S\ref{sec:service-metadata-interface}
-  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
-  & cluster state management~\cite{website:ceph-mon}
-  & consensus/consistency
-  \\
-Data I/O
-  & \S\ref{sec:data-io-interface}
-  & Swift in situ storage/compute~\cite{website:zerocloud}
-  & object interface classes~\cite{website:cls-lua}
-  & transaction/atomicity
-  \\
-Shared Resource
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & MPI collective IO, burst
-  & POSIX metadata protocols
-  & serialization/batching
-  \\
-File Type
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & ?????
-  & file striping strategy
-  & data/metadata access
-  \\
-Load Balancing
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
-  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
-  & migration/sampling
-  \\
-Durability
-  & \S\ref{sec:durability}
-  & S3/Swift interfaces (RESTful API)
-  & object store library~\cite{weil_rados_2007}
-  & persistence/safety
-  \\
-\end{tabular}
-\caption{Common internal abstractions. The same ``internal abstractions" are
-common across large-scale systems because they provide primitives that solve
-general distributed systems problems.  Here we list examples of what these
-internal abstractions are used for in ``production systems" and in Ceph.
-Malacology provides these internal abstractions as interfaces
-\newcomment{(Section~\ref{sec:malacology})}
-that higher level
-\oldcomment{applications}
-\newcomment{services (Section~\ref{sec:services})}
-can use.
-\oldcomment{; {\it italicized abstractions} are interfaces discussed in this
-paper.}}
-\addressesissue{1}
-\label{table:examples}
-\end{table*}
-
-The guiding principle is to re-use existing services and extend them so that
-these services can be \emph{programmed}. We accomplish programmability of a
-service by exporting bindings (or ``hooks'') for an interpreted programming
-language so that programming can occur without having to restart the storage
-system (see also below,~\S\ref{sec:durability}). Table~\ref{table:examples}
-shows the internal services from Ceph that we expose in the Malacology
-prototype via Lua~\cite{ierusalimschy1996lua} bindings
-\oldcomment{.}
-\newcomment{and Figure~\ref{fig:implementation-overview} compares what was
-already present in Ceph (gray boxes) to the Malacology interfaces we added.
-Section~\ref{sec:services} will describe the higher-level services we built
-with these interfaces.}
-\addressesissue{1}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation-overview.png}
-\caption{\newcomment{Malacology is implemented on the daemons and clients that
-run in a Ceph cluster. Interfaces expose internal subsystems and are used as
-building blocks for higher-level services.}\addressesissue{1}
-\label{fig:implementation-overview}}
-\end{figure}
-
-Lua is a portable embedded scripting language and we choose it as the
-interpreted language for Malacology because it offers superior performance and
-productivity trade-offs, including a JIT-based implementation that is well
-known for near native performance. Additionally, Lua has been used extensively
-in game engines, and systems research \cite{neto:dls14-luaos}, including
-storage systems where it has been effectively used both on
-\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010}
-and off \cite{sevilla:sc15-mantle} the performance critical path. Finally, the
-flexibility of the runtime allows execution sandboxing in order to address
-security and performance concerns. We will now discuss the common subsystems
-used to manage storage system and how Malacology makes them programmable.
-
-\subsection{Service Metadata Interface}
-\label{sec:mon}
-\label{sec:service-metadata-interface}
-\label{service-metadata}
-
-% straw man example
-Keeping track of state in a distributed system is an essential part of any
-successful service and a necessary component in order to diagnose and detect
-failures, when they occur. This is further complicated by variable propagation
-delays and heterogeneous hardware in dynamic environments.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In the case of Ceph, a consistent view of cluster state among server daemons
-and clients is critical to provide strong consistency guarantees to clients.
-Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
-metadata server) data structures called ``maps'' that record membership and
-status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
-responsible for integrating state changes into cluster maps, responding to
-requests from out-of-date clients and synchronizing members of the cluster
-whenever there is a change in a map so that they all observe the same system
-state. As a fundamental building block of many system designs, consensus
-abstractions such as Paxos are a common technique for maintaining consistent
-data versions, and are a useful system to expose.
-
-The default behavior of the monitor can be seen as a Paxos-based notification
-system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing
-clients to identify when new values (termed epochs in Ceph) are associated to
-given maps.  Since Ceph does not expose this service directly, as part of our Malacology implementation, we expose a key-value
-service designed for managing configuration metadata that is built on top of
-the consensus engine. Since the monitor is intended to be out of the
-high-performance I/O path, a general guideline is to make use of this
-functionality infrequently and to assign small values to maps.\\
-
-\noindent{\it{\textbf{Malacology: }}} Malacology exposes a strongly-consistent view of
-time-varying service metadata as a service rather than a hidden internal
-component. Malacology provides a generic API for adding arbitrary values to
-existing subsystem cluster maps. As a consequence of this, applications can
-define simple but useful service-specific logic to the strongly-consistent
-interface, such as authorization control (just specific clients can write new
-values) or triggering actions based on specific values (e.g. sanitize values).
-The higher level services we implement in \S\ref{sec:services} make use of this
-functionality to register, version and propagate dynamic code (Lua scripts) for
-new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
-policies in the load balancer \S\ref{malacology:mds}.  Using this service
-guarantees that interface definitions are not only made durable, but are
-transparently and consistently propagated throughout the cluster so that
-clients are properly synchronized with the latest interfaces.
-
-\newcomment{\\ \noindent \it{\textbf{Impact: }} provides core functionality
-because it lets daemons come to consensus on system critical state. Bugs in the
-internal subsystems or ommitting this from services that need this type of
-consistency affects correctness.}
-\addressesissue{3, 5}
-
-%And finally, it is important to not underestimate the importance of the
-%preservation of the modifications that enable a new system service. In many
-%(if not all) cases, interfaces defining access to data are just as important
-%as the data itself by virtue of inherently providing structural context. Thus,
-%all components of a system service must be kept to the same standards of
-%protection as data in the system itself.
-
-\oldcomment{\noindent 4.2 Object I/O Intefrace}
-\subsection{Data I/O Interface}
-\label{sec:data-io-interface}
-\label{object-data-interface}
-Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
-supports application-specific object
-interfaces~\cite{weil_rados_2007}. The ability to offload computation can
-reduce data movement, and transactional interfaces can significantly simplify
-construction of complex storage interfaces that require uncoordinated parallel
-access.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-An object interface is a plugin structured similarly to that of an RPC
-in which a developer creates a named function within the cluster that clients
-may invoke. In the case of Ceph each function is implicitly run within the
-context of an object specified when the function is called. Developers of
-object interfaces express behavior by creating a composition of native
-interfaces or other custom object interfaces, and handle serialization of
-function input and output.  A wide range of native interfaces are available to
-developers such as reading and writing to a byte stream, controlling object
-snapshots and clones, and accessing a sorted key-value database. These native
-interfaces may be transactionally composed along with application
-specific logic to create semantically rich interfaces. An example would be
-an interface that atomically updates a matrix stored in the bytestream and an
-index of the matrix stored in the key-value database.\\
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation.png}
-\caption{Malacology allows users to dynamically define object/metadata
-interfaces by extending the object storage daemon (OSD) and metadata server
-(MDS) subsystems with an embedded Lua VM.  It uses the service metadata
-interface (M) to propagate interfaces/versions across the cluster
-\label{fig:implementation}}
-\end{figure}
-
-\noindent {\it{\textbf{Malacology: }}} The implementation of Ceph's object abstraction,
-although powerful, does not readily support programmability. Supporting only
-C/C++ for object interface developers, Ceph requires distribution of compiled
-binaries for the correct architecture, adding a large barrier of entry for
-developers and system administrators. Second, having no way to dynamically
-unload modules, any changes require a full restart of a storage daemon which
-may have serious performance impacts due to loss of cached data. And finally,
-the security limitations of the framework limit the use of object interfaces to
-all but those with administrative level access and deep technical expertise.
-
-To address these concerns, Malacology takes advantage of Lua extensions
-contributed by the Ceph community. This allows new object interfaces to be
-dynamically loaded into the system and modified at runtime, resulting in a
-object storage API with economy of expression, which at the same time provides
-the full set of features of the base object class. New object interfaces that
-are expressed in thousands of lines of code can be implemented in approximately
-an order of magnitude less code~\cite{geambasu_comet_2010}. While the use of
-Lua does not prevent deployment of malicious code, certain types of coding
-mistakes can be handled gracefully, and access policies are used to limit access
-to trusted users~\cite{ierusalimschy1996lua}.
-
-\oldcomment{All of these modifications work in tandem to implement the desired behavior,
-and are typically co-designed together such that each depends on the other to
-behave as expected.}
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by pushing behavior to lower parts of the storage stack, thereby
-minimizing hops and distributing computation.}
-\addressesissue{3, 5}
-
-\subsection{Distributed Metadata Interfaces}
-\begin{notes}
-\textcolor{red}{This section has been re-organized}
-\end{notes}
-\label{sec:distributed-metadata-interfaces}
-\label{malacology:mds}
-
-\oldcomment{The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction~\cite{weil:sc2004-dyn-metadata}.}
-\newcomment{File systems provide clients with the familiar POSIX file abstraction.
-While this guarantees strong consistency it comes at the cost of scalability,
-increased complexity, and lower performance.}
-\addressesissue{1}
-In general, distributed file systems protect resources by providing
-hierarchical indexing and distributed locking services.
-
-%The metadata cluster moves units of the namespace called directory fragments
-%and assigns them to servers using its own metadata load balancer. It then uses
-%a hard-coded policy to balance these fragments using a scalarized load metric
-%based on CPU, workload, and file system operation metrics. Although the CephFS
-%balancer has been shown to be complicated, its complexity is justifiable.
-%Having the flexibility to choose what to move, where to move it, and how much
-%to move is very powerful. The problem is that it is difficult to pick a
-%one-size-fits-all policy, system administrators may want to partition load
-%based on SLAs, or applications are better equipped to make these decisions,
-%etc. -- but we choose instead to expose a balancing API at strategic points in
-%the balancing logic using Lua hooks. ~\cite{sevilla:sc15-mantle}.
-
-\oldcomment{\it{\textbf{Malacology:}} Interfaces are added in strategic spots in the
-metadata service for guarding shared resources, defining new file types,
-and load balancing.}
-
-\oldcomment{\noindent Shared resource interface.}
-\subsubsection{Shared Resource Interface}
-
-In general, file system metadata servers manages client sessions, allowing
-clients to obtain locks (e.g. file byte ranges), and capabilities (e.g. to
-cache file data). Clients and metadata servers use a cooperative protocol in
-which clients voluntarily release resources back to the metadata service in
-order to implement sharing policies.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In Ceph, the locking service implements a capability-based system that
-expresses what data and metadata clients are allowed to access as well as what
-state they may cache and modify locally.  While designed for the file
-abstraction, indexing, locking, and caching are all common services that are
-useful to a broad spectrum of applications.  Distributed applications that
-share centralized resources (e.g. a database or directory) face similar
-challenges which are often solved using application-specific sharding.
-
-\newcomment{\\ \noindent\it{\textbf{Malacology: }}}
-While the current policy for sharing access and voluntarily releasing resources
-is largely best-effort, Malacology supports generalized policies between
-metadata servers and clients that can be used to implement fairness or
-priority.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} provides core functionality
-to protect and provide exclusive access for any shared resource. May hurt
-performance if the resource in question does not require strong consistency.}
-\addressesissue{3}
-
-\oldcomment{\noindent File type interface.}
-\subsubsection{File Type Interface}
-The metadata service manages a POSIX file system hierarchy where files and
-directories are represented as inode data structures and expose a POSIX file
-interface. Applications that manage large amounts of metadata (e.g. users or
-database snapshots) often require a naming service. We allow new inode types to
-be defined such that applications can create domain-specific interfaces to
-inodes that may modify locking and capability policies. We will show how this
-is used in Section~\ref{sec:seq} when we discuss a distributed shared-log built
-on Malacology.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this interface is both a feature
-and a performance optimization. It is a feature because it allows developers to
-add support for different  storage types, such as how to read new file formats
-or what consistency semantics to use for a specific subtree in the hierarchical
-namespace. It is also a performance optimization because future programmers can
-add optimizations for processing specific types of files into the inode itself.}
-\addressesissue{3}
-
-%%% TODO: striping strategy is stored with the inode
-
-\oldcomment{\noindent \bf Load balancing interface.} 
-\subsubsection{Load Balancing Interface}
-
-Ceph also addresses the challenge of balancing metadata load with a separate
-metadata cluster. This cluster uses load balancing policies to migrate
-directory inodes around the cluster to alleviate overload at single
-nodes~\cite{weil:sc2004-dyn-metadata}. The policies use metrics based on system
-state (e.g.  CPU and memory utilization) and statistics collected by the
-cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
-partitioning to move variable sized namespace subtrees. These units can be
-shipped anywhere (i.e., to any metadata server of any capacity) at any time for
-any reason. The original balancer was designed with hard-coded policies and
-tunables.\\
-
-\noindent {\it{\textbf{Malacology: }}}
-Mantle~\cite{sevilla:sc15-mantle}, a programmable metadata load balancer, has
-been implemented on Malacology so that its infrastructure can enjoy all the
-properties of existing internal abstractions.  The load balancing logic is
-intact but the infrastructure has been improved to safely load, manage, and
-persist its load balancing policies.  Composing Mantle with the service
-metadata and durability interfaces, Mantle can safely version balancer
-policies, save balancer policies in the backend object store and centralize
-warnings/errors. When combined with file types, load balancing policies can
-express a variety of rules such as multi-tenant and multi-application
-balancing.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by allowing them to specify how to partition, replicate, and
-distribute metadata in response to overloaded servers.}
-\addressesissue{3}
-
-\subsection{Durability Interface}
-\label{sec:durability}
-
-Ceph provides storage by striping and replicating data across
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-protects data using common techniques such as erasure
-coding, replication, and scrubbing.  Furthermore, many of these features
-are implemented using a peer-to-peer protocol that allows object storage
-devices to operate autonomously without a centralized coordinator.  For
-example, when the number of placement groups change, the OSDs re-balance and
-re-shard data in the background in a process called placement group splitting
-in which OSDs communicate directly with each other to cover a new state.
-In order to reduce load on the monitoring service, Ceph OSDs use a gossip
-protocol to efficiently propagate changes to cluster maps throughout the
-system, and autonomously initiate recovery mechanisms when failures are
-discovered.\\
-
-\noindent {\it{\textbf{Malacology: }}} Metadata service policies and object storage
-interfaces are stored durability within RADOS and managed by storing references
-with the object maps. Since the cluster already propagates a consistent view of
-these data structures, we use this service to automatically install interfaces
-in OSDs, and install policies within the metadata server daemon (MDS) such that clients and daemons are
-synchronized on correct implementations without restarting.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this is a feature because it
-adds data safety and persistence to system metadata; while nice to have it does
-effect correctness.}
-\addressesissue{3}
-
-\section{Services Built on Malacology}
-\label{sec:services}
-\label{services-built-on-malacology}
-
-\label{services}
-%\begin{table}
-%\centering\small
-%\begin{tabular}{  l | l | l }
-%\textbf{Service}              &
-%\textbf{Used to...}           &
-%\textbf{Malacology Interface} \\ \hline
-%Mantle & load balance metadata     & Load Balancing  \\
-%       & agree on current balancer & Service Metadata\\
-%       & control client write lock & Shared Resource \\
-%       & persist policy logic      & Durability      \\ \hline
-%ZLog   & implement CORFU           & Object \& Data  \\ 
-%       & install ZLog interfaces   & Service Metadata\\
-%       & persist log entries       & Durability      \\ \hline
-%ZLog   & load balance sequencers   & Load Balancing  \\
-%Seqr.  & manage access to log tail & Shared Resource \\ 
-%\end{tabular}
-%\caption{Mantle and ZLog are built using Malacology interfaces so they inherit the
-%robustness of the subsystems of the underlying stack.}
-%\label{table:implementation}
-%\end{table}
-
-In this section we describe two services built on top of Malacology. The first
-is Mantle, a framework for dynamically specifying metadata load balancing
-policies. The second system, ZLog, is a high-performance distributed shared-log.
-In addition to these services, we'll demonstrate how we combine ZLog and Mantle
-to implement service-aware metadata load balancing policies.
-
-\subsection{Mantle: Programmable Load Balancer}
-\label{sec:mantle}
-
-Mantle is a programmable load balancer that separates the metadata balancing
-policies from their mechanisms. Administrators inject code to change how the
-metadata cluster distributes metadata. In ~\cite{sevilla:sc15-mantle} the
-authors showed how using Mantle, one can implement a single node metadata
-service, a distributed metadata service with hashing, and a distributed
-metadata service with dynamic subtree partitioning.
-
-The original implementation was ``hard-coded" into Ceph and lacked robustness
-(no versioning, durability, or policy distribution).  Re-implemented using
-Malacology, Mantle now enjoys (1) the versioning provided by the metadata
-service interface provided by the monitor daemons and (2) the durability and
-distribution provided by Ceph's reliable object store.  Re-using the internal
-abstractions with Malacology resulted in a 2\(\times\) reduction in source code
-compared to the original implementation.
-
-\subsubsection{Versioning Balancer Policies}
-
-Ensuring that the version of the current load balancer is consistent across the
-physical servers in the metadata cluster was not addressed in the original
-implementation. The user had to set the version on each individual server and
-it was trivial to make the versions inconsistent. Maintaining consistent
-versions is important for balancing policies that are cooperative in which local
-decisions are made assuming properties about other instances of the same
-policy.
-
-With Malacology, Mantle stores the version of the current load balancer in the
-Ceph service metadata interface. The version of the load balancer corresponds
-to an object name in the balancing policy. Using the service metadata interface
-ensures that all metadata servers use the same version of the load balancer. As
-a result, the policy version inherits the consistency of Ceph's internal
-monitor daemons. The user changes the version of the load balancer using a new
-CLI command.
-
-\subsubsection{Making Balancer Policies Durable}
-
-The load balancer version described above corresponds to the name of an object
-in RADOS that holds the actual Lua balancing code.  When MDS nodes start
-balancing load, they first check the latest version from the MDS map and
-compare it to the balancer they have loaded. If the version has changed, they
-dereference the pointer to the balancer version by reading the corresponding
-object in RADOS. This is in contrast to the original balancer which stored load
-balancer code on the local file system -- a technique which is unreliable and
-may result in silent corruption.
-
-The balancer pulls the Lua code from RADOS synchronously; asynchronous reads
-are not possible because of the architecture of the MDS. The synchronous
-behavior is not the default behavior for RADOS operations, so we achieve this
-with a timeout: if the asynchronous read does not come back within half the
-balancing tick interval the operation is canceled and a Connection Timeout
-error is returned. By default, the balancing tick interval is 10 seconds, so
-Mantle will use a 5 second second timeout.
-
-This design allows Mantle to immediately return an error if anything
-RADOS-related goes wrong.  We use this implementation because we do not want to
-do a blocking OSD read from inside the global MDS lock. Doing so would bring
-down the MDS cluster if any of the OSDs are not responsive.
-
-Storing the balancers in RADOS is simplified by the use of an interpreted
-language for writing balancer code. If we used a language that needs to be
-compiled, like the C++ object classes in the OSD, we would need to ensure
-binary compatibility, which is complicated by different operating systems,
-distributions, and compilers.
-
-\subsubsection{Logging, Debugging, and Warnings}
-
-In the original implementation, Mantle would log all errors, warnings, and
-debug messages to a log stored locally on each MDS server. To get the simplest
-status messages or to debug problems, the user would have to log into each MDS
-individually, look at the logs, and reason about causality and ordering.
-
-With Malacology, Mantle re-uses the centralized logging features of the
-monitoring service. Important errors, warnings, and info messages are collected
-by the monitoring subsystem and appear in the monitor cluster log so instead of
-users going to each node, they can watch messages appear at the monitor.
-Messages are logged sparingly, so as not to spam the monitor with frivolous
-debugging but important events, like balancer version changes or failed
-subsystems show up in the centralized log.
-
-\subsection{ZLog: A Fast Distributed Shared Log}
-\label{sec:zlog}
-
-The second service implemented on Malacology is ZLog, a high-performance
-distributed shared-log that is based on the CORFU
-protocol~\cite{balakrishnan_corfu_2012}. The shared-log is a
-powerful abstraction used to construct distributed systems, such as
-metadata management~\cite{balakrishnan:sosp13} and elastic database
-systems~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
-However, existing implementations that rely on consensus algorithms such as
-Paxos funnel I/O through a single point introducing a bottleneck that restricts
-throughput. In contrast, the CORFU protocol is able to achieve high throughput
-using a network counter called a \emph{sequencer}, that
-decouples log position assignment from log I/O.
-
-While a full description of the CORFU system is beyond the scope of this
-paper, we briefly describe the custom storage device interface, sequencer
-service, and recovery protocol, and how these services are instantiated in the
-Malacology framework.
-
-\subsubsection{Sequencer}
-\label{sec:seq}
-
-High-performance in CORFU is achieved using a sequencer service that assigns
-log positions to clients by reading from a volatile, in-memory counter which
-can run at a very high throughput and at low latency. Since the sequencer is
-centralized, ensuring serializability in the common case is trivial.  The
-primary challenge in CORFU is handling the failure of the sequencer in a way
-that preserves correctness. Failure of the sequencer service in CORFU is
-handled by a recovery algorithm that recomputes the new sequencer state using a
-CORFU-specific custom storage interface to discover the tail of the log, while
-simultaneously invalidating stale client requests using an epoch-based
-protocol.
-
-{\bf Sequencer interface.} The sequencer resource supports the ability to
-\emph{read()} the current tail value and get the \emph{next()} position in
-the log which also atomically increments the tail position.
-We implement the sequencer service in Malacology as an application-specific
-file type that associates a small amount of metadata
-with the inode state managed by the Ceph metadata service.
-This approach has the added
-benefit of allowing the metadata service to handle naming, by
-representing each sequencer instance in the standard POSIX hierarchical
-namespace. The primary challenge in mapping the sequencer resource to the
-metadata service is handling serialization correctly to maintain the global
-ordering provided by the CORFU protocol.
-
-Initially we sought to directly model the sequencer service in Ceph as a
-non-exclusive, non-cacheable resource, forcing clients to perform a round-trip
-access to the resource at the authoritative metadata server for the sequencer
-inode.  Interestingly, we found that the capability system in Ceph uses a
-strategy to reduce metadata service load by allowing clients that open a
-shared file to temporarily obtain an exclusive cached copy of the resource,
-resulting in a round-robin, best-effort batching behavior. When a single
-client is accessing the sequencer resource it is able to increment the
-sequencer locally. Any competing client cannot query the sequencer until the
-metadata service has granted it access.
-
-While unexpected, this discovery allowed us to explore an implementation
-strategy that we had not previously considered. In particular, for bursty
-clients, and clients that can tolerate increased latency, this mode of
-operation may allow a system to achieve much higher throughput than a system
-with a centralized sequencer service.  We utilize the programmability of the
-metadata service to define a new policy for handling capabilities that controls
-the amount of time that clients are able to cache the sequencer resource. This
-allows an administrator or application to control the trade-off between latency
-and throughput beyond the standard best-effort policy that is present in Ceph
-by default.
-
-In Section~\ref{sec:evaluation} we quantify the trade-offs of throughput and
-latency for an approach based on a round-robin batching mode, and compare this
-mode to one in which the metadata server mediates access to the sequencer state
-when it is being shared among multiple clients. Quantifying these trade-offs
-should provide administrators with guidelines for setting the tunables for
-different ``caching" modes of the sequencer.
-
-{\bf Balancing policies.} As opposed to the batching mode for controlling access to the sequencer
-resource, more predictable latency can be achieved by treating the sequencer
-inode as a shared non-cacheable resource, forcing clients to make a round-trip
-to the metadata service. However, the shared nature of the metadata service
-may prevent the sequencer from achieving maximum throughput. To provide a
-solution to this issue we have taken advantage of the programmability of the
-metadata service load balancing to construct a service-specific load balancing
-policy. As opposed to a balancing policy that strives for uniform
-load distribution, a ZLog-specific policy may utilize knowledge of inode types
-to migrate the sequencer service to provisioned hardware during periods of
-contention or high demand.
-
-\subsubsection{Storage Interface}
-
-The storage interface is a critical component in the CORFU protocol. Clients
-independently map log positions that they have obtained from the sequencer
-service (described in detail in the next section) onto storage devices, while
-storage devices provide an intelligent \emph{write}-once, random \emph{read}
-interface for accessing log entries. The key to correctness in CORFU lies with
-the enforcement of up-to-date epoch tags on client requests; requests tagged
-with out-of-date epoch values are rejected, and clients are expected to
-request a new tail from the sequencer after refreshing state from an auxiliary
-service.  This mechanism forms the basis for sequencer recovery.
-
-In order to repopulate the sequencer state (i.e. the cached, current tail of
-the log) during recovery of a sequencer, the maximum position in the log must
-be obtained. To do this, the storage interface exposes an additional
-\emph{seal} method that atomically installs a new epoch value and returns the
-maximum log position that has been written.
-
-Since the sequencer service does not resume until the recovery process has
-completed, there cannot be a race with clients appending to the log, and the
-immutability of the log allow reads to never block during a sequencer failure.
-Recovery of the sequencer process itself may be handled in many ways, such as
-leader election using an auxiliary service such as Paxos. In our
-implementation, the recovery is the same as (and is inherited from) the Ceph
-metadata service. Handling the failure of a client that holds the sequencer
-state is similar, although a timeout is used to determine when a client should
-be considered unavailable.
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-%We wanted to implement CORFU and we had two choices: build it from the ground
-%up or do it on top of Malacology. To show the feasability of building on
-%Malacology our evaluation focuses on the
-
-Our evaluation demonstrates the feasibility of building new service
-abstractions atop programmable storage, focusing on the performance of the
-internal abstractions exposed by Malacology and used to construct the Mantle
-and ZLog services. We also discuss latent capabilities we discovered in this
-process that let us navigate different trade-offs within the services
-themselves. First, we benchmark scenarios with high sequencer contention by
-examining the interfaces used to map ZLog onto Malacology; specifically, we
-describe the sequencer implementation and the propagation of object and data
-interfaces interfaces.  Next, we benchmark scenarios in which the storage
-system manages multiple logs by using Mantle to balance sequencers across a
-cluster.
-
-% should save this for future work, otherwise it undercuts the eval.
-%Detailed system-wide performance measurements and analysis is forthcoming for
-%both Mantle and ZLog, as we try to find the best metadata balancing policies
-%and physical design implementation trade-offs.
-
-Since this work focuses on the programmability of Malacology, the goal of this
-section is to show that the components and subsystems that support the
-Malacology interfaces provide reasonable relative performance, as well as to
-give examples of the flexibility that Malacology provides to programmers.
-
-%As we search for the best metadata balancing policies
-%and physical design implementation trade-offs,
-%we plan to rel
-
-\subsection{Mapping ZLog onto Malacology}
-\label{sec:mapping-zlog-onto-malacology}
-
-We evaluate Malacology by exploring one possible mapping of the ZLog
-implementation of CORFU onto Ceph in which we re-use (1) the metadata service
-to manage naming and synchronization of the sequencer resource by treating the
-resource as an inode, and (2) the monitoring sub-subsystem to distribute and
-install application-specific I/O interface required of the CORFU protocol. In
-Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the inode
-abstraction for implementing the sequencer resource enables load balancing policies
-to migrate the sequencer resource in heavy-load situation.
-
-\subsubsection{Sequencer Implementation}
-\label{sec:sequencer-implementation}
-
-We evaluate the feasibility of using the metadata service to implement a
-sequencer resource that is responsible for maintaining a total ordering of
-the log. Clients contact the sequencer to obtain the tail of the log and then
-independently initiate I/O, thus we measure both the throughput and latency of
-obtaining new tail positions which bounds client append performance.
-
-The sequencer is implemented as a custom file type in which the sequencer
-state (a 64-bit integer) is embedded in the inode of a file representing the
-sequencer. A total ordering of the log is imposed by the re-use of the
-capability service that can be used to grant exclusive access of inode state
-to clients. The metadata service is responsible for maintaining exclusivity
-and granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
-behavior of the system in which a best-effort policy is used.  The two colors
-represent points in time that the clients were able to access the resource.
-The best-effort policy shows a high degree of interleaving
-between clients but the system spends a large portion of time re-distributing
-the capability, reducing overall throughput.
-
-In order to control the performance of the system we implement a
-policy that (1) restricts the length of time that a client may maintain
-exclusive access and (2) limits the number of log positions that a client may
-generate without yielding to other clients waiting for access. The behavior of
-these two modes is illustrated in Figures~\ref{fig:capdelay-quota-behavior}
-(b) and (c), respectively.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/capdelay-quota-behavior-a.png}
-\includegraphics{figures/capdelay-quota-behavior-b.png}
-\includegraphics{figures/capdelay-quota-behavior-c.png}
-\caption{Each dot is an individual request, spread randomly along the \(y\)
-axis. The default behavior is unpredictable, ``delay" lets clients hold the
-lease longer, and ``quota" gives clients the lease for a number of operations.}
-\label{fig:capdelay-quota-behavior}
-\end{figure}
-
-Figure~\ref{fig:captp} demonstrates a configurable trade-off between
-throughput and latency. In the experiment two clients are run each with a
-fixed 0.25 second maximum reservation on the capability, and we vary the size
-of the log position quota running each configuration for two minutes. The
-total operations per second is the combined throughput of the two clients, and
-the average latency is number of microseconds required to obtain a new log
-position. With a small quota more time is spent exchanging exclusive access,
-while a large quota reservation allows clients to experience a much lower
-latency because they experience uncontended access for a longer period of time.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/tradeoff.png}
-\caption{Sequencer throughput by re-using various services.  The highest performance is
-achieved using a single client with exclusive, cacheable privilege. Round-robin
-sharing of the sequencer resource is affected by the amount of time the
-resource is held, with best-effort performing the worst.}
-\label{fig:captp}
-\end{figure}
-
-To get a better picture of latency, Figure~\ref{fig:capcdf} shows the CDF of
-latency for each client in all experiment configurations. At the 99th
-percentile clients accessed the sequencer is less than a millisecond. The CDF
-is cropped at the 99.999th percentile due to large outliers that we believe
-occur in instances in which the MDS is performing I/O while it is in the
-process of re-distributing the capability to another client.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/caps-delay-latency.png}
-\caption{
-The latency distribution shows that if the clients hold their capabilities longer,
-performance decreases. Changing the delay can help the sequencer and
-clients strike a balance between throughput and latency.}
-\label{fig:capcdf}
-\end{figure}
-
-\textbf{Malacology exposes} the internal capability management service and
-allows users to navigate latency and throughput trade-offs.  Other approaches to
-designing the sequencer service also exist, such as using a centralized service
-in which each access is a network round-trip. In contrast to the mechanism we
-explored which is appropriate for clients with bursty workloads, it may be
-easier to provide predictable performance using a centralized service, and we
-will be exploring in the future how this can be achieved using the capability
-system.
-
-\subsubsection{Interface Propagation}
-
-Domain-specific data interfaces (Section~\ref{sec:application-specifc-storage-stacks}) allow
-co-design between applications and the storage system. Malacology supports
-custom object interfaces in RADOS that require interface implementations to
-be installed on the storage devices in the system, supporting the evolution of
-interfaces through automatic system-wide versioning and installation through
-the service metadata interface (Section~\ref{service-metadata}). We evaluate
-the performance of installing a new interface version in the cluster,
-which is an important metric for applications that frequently evolve
-interfaces.
-
-We demonstrate the feasibility of utilizing the Ceph monitoring sub-system by
-evaluating the performance of installing and distributing interface updates.
-Figure~\ref{fig:propdelay} shows the CDF of the latency of interface updates.
-The interfaces are Lua scripts embedded in the cluster map and distributed
-using a peer-to-peer gossip protocol.  The latency is defined as the elapsed
-time following the Paxos proposal for an interface update until each OSD makes
-the update live (the cost of the Paxos proposal is configurable and is
-discussed below). The latency measurements were taken on the nodes running
-Ceph server daemons, and thus exclude the client round-trip cost. In each of
-the experiments 1000 interface updates were observed.
-
-Figure~\ref{fig:propdelay} shows the lower bound cost for updates in a large
-cluster. In the experiment labeled ``120 OSD (RAM)" a cluster of 120 OSDs (10
-OSDs x 12 servers) using an in-memory data store were deployed, showing a
-latency of less than 54 ms with a probability of 90\% and a worst case latency
-of 194 ms. These costs demonstrate the penalty of distributing the interface
-in a large cluster. In practice the costs include, in addition to cluster-wide
-propagation of interface updates, the network round-trip to the
-interface management service, the Paxos commit protocol itself, and other factors
-such as system load. By default Paxos proposals occur periodically with a 1
-second interval in order to accumulate updates. In a minimum, realistic quorum
-of 3 monitors using HDD-based storage, we were able to decrease this interval
-to an average of 222 ms.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=60mm,trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
-\caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
-interface commit.}
-\label{fig:propdelay}
-\end{figure}
-
-\subsection{Load Balancing ZLog Sequencers with Mantle}
-\label{sec:zlog-balancing}
-
-In practice, a storage system implementing CORFU will support a multiplicity of
-independent totally-ordered logs for each application.  For this scenario
-co-locating sequencers on the same physical node is not ideal but building a load
-balancer that can migrate the shared resource ({\it e.g.,}, the resource that
-mediates access to the tail of the log) is a time-consuming, non-trivial task.
-It requires building subsystems for migrating resources, monitoring the
-workloads, collecting metrics that describe the utilization on the physical
-nodes, partitioning resources, maintaining cache coherence, and managing
-multiple sequencers. The following experiments demonstrate the feasibility of
-using the mechanisms of the Malacology load balancing interface to inherit
-these features and to alleviate load from overloaded servers.
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-behaviors.png}
-\caption{CephFS/Mantle load balancing have better throughput than co-locating
-all sequencers on the same server.  Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify this improvement;
-Section~\ref{sec:feature-backoff} examines the migration at 0-60 seconds.
-}\label{fig:mantle-balancer-behaviors}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-performance.png}
-\includegraphics{figures/mantle-mode-performance.png}
-\caption{In (a) all CephFS balancing modes have the same performance; Mantle uses a
-balancer designed for sequencers. In (b) the best combination of mode and
-migration units can have up to a 2\(\times\) improvement.}\label{fig:mantle-balancer-performance}
-\end{figure}
-
-The experiments are run on a cluster with 10 nodes to store objects, one node
-to monitor the cluster, and 3 nodes that can accommodate sequencers. 
-Instead of measuring contention at the clients like
-Section~\ref{sec:sequencer-implementation}, these experiments measure
-contention at the sequencers by forcing clients to make round-trips for every
-request. We implement this using the shared resource interface that forces
-round-trips.  Because the sequencer's only function is to  hand out positions
-for the tail of the log, the workload is read-heavy.
-
-First, we show how the ZLog service can orchestrate multiple sequencers using
-the Malacology load balancing interface.
-Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput over time of
-different load balancers as they migrate 3 sequencers (with 4 clients) around
-the cluster; ``No Balancing" keeps all sequencers on one server, ``CephFS"
-migrates sequencers using the CephFS load balancers, and ``Mantle" uses a
-custom load balancer we wrote specifically for sequencers. The CephFS load
-balancers are hard-coded into Ceph while Mantle is an API for balancing load
-presented in~Mantle~\cite{sevilla:sc15-mantle} and re-implemented on top of
-Malacology in this paper.  The increased throughput for the CephFS and Mantle
-curves between 0 and 60 seconds are a result of migrating the sequencer(s) off
-overloaded servers.
-
-In addition to showing that migrating sequencers improves performance,
-Figure~\ref{fig:mantle-balancer-behaviors} also demonstrates features that we
-will explore in the rest of this section.
-Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify the differences in performance
-when the cluster stabilizes at time 100 seconds and
-Section~\ref{sec:feature-backoff} examines the slope and start time of the
-rebalancing phase between 0 and 60 seconds by comparing the aggressiveness of
-the balancers.
-
-\subsubsection{Feature: Balancing Modes}
-\label{sec:feature-balancing-modes}
-
-Next, we quantify the performance benefits shown in
-Figure~\ref{fig:mantle-balancer-behaviors}.  To understand why load balancers
-perform differently we need to explain the different balancing modes that the
-load balancer service uses and how they stress the subsystems that receive and
-forward client requests in different ways. In
-Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
-performance of the balancing mode that CephFS falls into {\it most of the time}.
-CephFS has 3 modes for balancing load: CPU mode, workload mode, and hybrid
-mode. All three have the same structure for making migration decisions but vary
-based on the metric used to calculate load. For this sequencer workload the 3
-different modes all have the same performance, shown in
-Figure~\ref{fig:mantle-balancer-performance} (a), because the load balancer falls
-into the same mode a majority of the time.  The high variation in performance
-for the CephFS CPU Mode bar reflects the uncertainty of using something as
-dynamic and unpredictable as CPU utilization to make migration decisions. In
-addition to the suboptimal performance and unpredictability, because
-CephFS balancers all fall into the same mode despite using different balancers
-is a problem because it prevents administrators from properly exploring the
-balancing state space.
-
-Mantle gives the administrator more control over balancing policies; for
-the Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the load
-balancing interface to program logic for balancing read-heavy workloads,
-resulting in better throughput and stability.  When we did this we also
-identified two balancing modes relevant for making migration decisions
-for sequencers.
-
-Using Mantle, the administrator can put the load balancing service into ``proxy
-mode" or ``client mode". In proxy mode one server receives all requests and
-farms off the requests to slave servers; the slaves servers do the actual tail
-finding operation. In client mode, clients interact directly with the server
-that has their sequencer.  These modes are illustrated in
-Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
-co-located on one physical server -- performance for that mode is shown by the
-No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
-Mode" in the lower left shows how clients for sequencer 2 are re-directed to
-server B. The two ``Proxy mode" diagrams on the right show how clients continue
-sending requests to server A even though sequencer 2 has migrated to server B
-(the meaning of ``(Half)" is discussed in the next section).
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-modes.png}
-\caption{ In client mode clients sending requests to the server that houses
-their sequencer. In proxy mode clients continue sending their requests to the
-first server.  }\label{fig:mantle-modes}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-mode-behavior.png}
-\caption{The performance of proxy mode achieves the highest throughput but at
-the cost of lower throughput for one of the sequencers. Client mode is more
-fair but results in lower cluster throughput.
-}\label{fig:mantle-mode-behavior}
-\end{figure}
-
-Figure~\ref{fig:mantle-mode-behavior} shows the throughput over time of the two
-different modes for an environment with only 2 sequencers (again 4 clients
-each) and 2 servers. The curves for both sequencers in
-Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second and
-at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
-Performance of Sequencer 2 decreases because it stayed on the proxy which now
-handles all client requests, processes requests for Sequencer 2, and forwards
-requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
-because distributing the sequencers in this way separates (1) the handling of
-the client requests and (2) finding the tail of the log and responding to
-clients.  Doing both steps is too heavy weight for one server and sequencers on
-slave nodes can go faster if work is split up; this phenomenon is not
-uncommon and has been observed in chain replication~\cite{chain_rep}.
-
-Cluster
-throughput improves at the cost of decreased throughput for Sequencer 2.
-Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer mode manually (no
-balancing phase) and shows that the cluster throughput is worse than the
-cluster throughput of proxy mode. That graph also shows that Sequencer 2 has
-less throughput than Sequencer 2. In this case, the scatter-gather process used
-for cache coherence in the metadata protocols causes strain on the server
-housing Sequencer 2 resulting in this uneven performance.
-
-\subsubsection{Feature: Migration Units}
-\label{sec:feature-migration-units}
-
-Another factor that affects performance in this environment is how much load
-is on each server; these experiments quantify that effect by programming the
-load balancing interface to control the amount of load to migrate. We call this
-metric a ``migration unit".  Expressing this heuristic is not easily achievable
-with outward facing tunable parameters (i.e. system knobs) but with Mantle's
-programmable interface, we can force the load balancer to change its migration
-units. To force the balancer into the Proxy Mode (Half) scenario in
-Figure~\ref{fig:mantle-modes}, which uses migration units equal to half the
-load on the current server, we can use: \texttt{ targets[whoami+1] = mds[whoami]["load"]/2
-}.
-
-This code snippet uses globally defined variables and tables from the Mantle
-API to send half of the load on the current server (whoami) to the next ranked
-server (whoami + 1); the \texttt{targets} array is a globally defined table that
-the balancer uses to do the migrations. Alternatively, to migrate all load a
-time step, we can remove the division by 2.
-
-%\begin{figure}[t!]
-%\centering
-%\includegraphics{figures/mantle-mode-performance.png}
-%\caption{The performance of the different modes using various migration units
-%shows almost a 2\(\times\) improvement in the best case.
-%}\label{fig:mantle-mode-performance}
-%\end{figure}
-
-Figure~\ref{fig:mantle-balancer-performance} (b) shows the performance of the modes
-using different migration units. Recall that this setup only has 2 sequencers
-and 2 servers, so performance may be different at scale. Even so, it is clear
-that client mode does not perform as well for read-heavy workloads. We even see
-a throughput improvement when migrating all load off the first server, leaving
-the first server to do administrative tasks (this is common in the metadata
-cluster because the first server does a lot of the cache coherence work) while
-the second server does all the client work. Proxy mode does the best in both
-cases and shows large performance gains when completely decoupling client
-request handling and operation processing in Proxy Mode (Full).  The parameter
-that controls the migration units helps the administrator control the sequencer
-co-location or distribution across the cluster. This trade-off was explored
-extensively in the Mantle paper but the experiments we present here are
-indicative of an even richer set of states to explore.
-
-\subsubsection{Feature: Backoff}
-\label{sec:feature-backoff}
-
-Tuning the aggressiveness of the load balancer decision making is also a
-trade-off that administrators can control and explore. The balancing phase from
-0 to 60 seconds in Figure~\ref{fig:mantle-balancer-behaviors} shows different
-degrees of aggressiveness in making migration decisions; CephFS makes a
-decision quickly and throughput jumps to 2500 ops/second within 10 seconds
-while Mantle takes more time to stabilize. This conservative behavior is
-controlled by programming the balancer to (1) use different conditions for when
-to migrate and (2) using a threshold for sustained overload.
-
-We control the conditions for when to migrate using \texttt{when()}, a callback
-in the Mantle API.  For the Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} we program \texttt{when()} to wait
-for load on the receiving server to fall below a threshold. This makes the
-balancer more conservative because it takes 60 seconds for cache coherence
-messages to settle.  The Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} also takes longer to reach peak
-throughput because we want the policy to wait to see how migrations affect the
-system before proceeding; the balancer does a migration right before 50
-seconds, realizes that there is a third underloaded server, and does another
-migration.
-
-The other way to change aggressiveness of the decision making is to program
-into the balancer a threshold for sustained overload. This forces the balancer
-to wait a certain number of iterations after a migration before proceeding. In
-Mantle, the policy would use the save state function to do a countdown after a
-migration.  Behavior graphs and performance numbers for this backoff feature is
-omitted for space considerations, but our experiments confirm that the more
-conservative the approach the less overall throughput.
-
-\textbf{Malacology pulls} the load balancing service out of the storage system
-to balance sequencers across a cluster. This latent capability also gives
-future programmers the ability to explore the different load balancing
-trade-offs including: load balancing modes to control forwarding vs. client
-redirection, load migration units to control sequencer distribution vs.
-co-location, and backoffs to control conservative vs. aggressive decision
-making.
-
-%\subsubsection{Experiment: Object Class
-%Programmability}\label{experiment-object-class-programmability}
-%
-%Add quote from CORFU paper that talks about what they had to hack
-%
-%\subsubsection{Experiment: Multi-Client
-%Burstiness}\label{experiment-multi-client-burstiness}
-%
-%\begin{figure}[htbp] \centering
-%\includegraphics{figures/caps-delay-thruput.png} \caption{Forcing the client
-%to drop their capabilities later (delay) improves throughput} \end{figure}
-
-\section{Related Work}
-
-% IOStack: Software-Defined Object Storage
-% sRoute: Treating the Storage Stack Like a Network
-   % - lots of good refs to check out
-% Information and control in gray-box systems
-
-Programmability of operating systems and networking resources, including
-distributed storage systems is not new, but we are not aware of work that makes
-generalization of existing services into programmable resources a key principle
-in storage systems design.
-
-Programmable storage systems can be viewed as an infrastructure for creating
-abstractions to better separate policies from mechanisms. This idea is not new.
-Software-defined networks (SDNs) create such an abstraction by separating the
-control plane from the data plane (see for example~\cite{jain:sigcomm13}). This
-notion of control/data separation was also applied in software-defined storage
-(SDS)~\cite{thereska:sosp13,stefanovici:fast16}. Similarly,
-IOStack~\cite{gracia:internet16} is providing policy-based provisioning and
-filtering in OpenStack Swift. According to a SNIA white
-paper~\cite{carlson:snia2014}, the primary goal of SDS is to control and
-facilitate flexible and dynamic provisioning of storage resources of different
-kinds, including flash memory and disk drives, to create a virtualized mapping
-between common storage abstractions (e.g. files, objects, and blocks) and
-storage devices taking data service objectives in terms of protection,
-availability, performance, and security into account. A programmable storage
-system exposes internal abstractions so that end users (not necessarily
-operators) can create new services on top of the storage stack. Thus, our
-notion of programmable storage differs from ``software-defined storage'' (SDS)
-in terms of goals and scope, although definitions of SDS are still in flux.
-
-Another view of programmable storage systems is one of tailoring systems
-resources to applications~\cite{arpaci:sosp01}. Related efforts include the
-Exokernel~\cite{engler:sosp95}, SPIN~\cite{bershad:sosp95} and
-Vino~\cite{seltzer:osdi96} projects; the latter two addressed the ability of injecting
-code into the kernel to specialize resource management. Another approach is to
-pass hints between the different layers of the IO stack to bridge the semantic
-gap between applications and
-storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}).
-
-Malacology uses the same Active and Typed Storage module presented in
-DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File Manifolds
-can be implemented with small changes to the Malacology framework, namely
-asynchronous object calls and Lua stubs in the inode, respectively.
-
-\section{Conclusion and Future Work}
-\label{conclusion-and-future-work}
-
-Programmable storage is a viable method for eliminating duplication of complex
-error-prone software used as workarounds for storage system
-deficiencies. We propose that systems expose their services in a safe way
-allowing application developers to customize system behavior to meet their
-needs while not sacrificing correctness. To illustrate the benefits of this
-approach we presented Malacology, a programmable storage system that
-facilitates the construction of new services by re-purposing existing subsystem
-abstractions of the storage stack.
-
-Future work will focus on constructing more interfaces to support a wide
-variety of storage system services that can be configured on-the-fly in existing
-systems. This work is one point along that path to producing general-purpose
-storage systems that can
-target  special-purpose applications.  Ultimately we want to utilize
-declarative methods for expressing new services.
-
-\bibliography{paper}
-\bibliographystyle{plain}
-%\printbibliography
-
-\end{document}
+the open source software community. Next we describe Malacology by presenting
+the subsystems within the underlying storage system that we re-purpose, and
+briefly describe how those system are used within Malacology
+(\S\ref{sec:malacology}).  Then we describe the services that we have
+constructed within the Malacology framework (\S\ref{sec:services}), and
+evaluate our ideas within our prototype implementation
+(\S\ref{sec:evaluation}).  We conclude by discussing related and future work.

### 2017-02-23 paper/3_challenges.tex
paper: cleanup line breaks and break into separate files
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-02-09 paper/paper.tex
@@
-\documentclass[preprint]{sigplanconf-eurosys}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage{pifont}
-\usepackage{color}
-\usepackage{comment}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Uncomment for diff
-\includecomment{notes}
-\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
-\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
-\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
-
-% Uncomment for camera-ready
-%\excludecomment{notes}
-%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
-%\newcommand{\oldcomment}[1]{}
-%\newcommand{\addressesissue}[1]{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\cmark}{\ding{51}}%
-\usepackage[font=footnotesize,labelfont=bf]{caption}
-
-\setlength{\belowcaptionskip}{-10pt}
-
-\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
-\usepackage[normalem]{ulem}
-\usepackage{pstricks}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-\setkeys{Gin}{width=\maxwidth,keepaspectratio}
-
-\usepackage[unicode=true]{hyperref}
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, Peter Alvaro, Shel Finkelstein, Jeff LeFevre},
-            pdftitle={Malacology: A Programmable Storage System},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=blue,
-            linkcolor=black,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-
-\title{Malacology: A Programmable Storage System}
-%\authorinfo{Paper ID: 165}          {Total Pages: 12}
-\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre}
-           {\{msevilla, jayhawk, ivo, cmaltzahn\}@soe.ucsc.edu, \{palvaro, finkelstein, jlefevre\}@ucsc.edu}
-\date{}
-\begin{document}
-
-\maketitle
-
-\begin{notes}
-\textcolor{red}{
-\noindent Issues raised by reviewers:
-\begin{enumerate}
-  \item clarify what was implemented: diagram of interfaces and prose
-  specifying the difference between Malacology and prior work (reviewers B, C,
-  D, E)
-  \item describe what makes a programmable storage system (reviewers C, D)
-  \item organize motivation: group interfaces into categories (e.g., core
-  functionality, features, and performance optimizations) and describe other
-  applications we can build with Malacology (reviewers B, C)
-  \item clarify terminology: service vs. file system metadata, storage vs.
-  file system, dirty vs. clean slate approaches, soft-state caching (reviewers
-  B, C)
-  \item address security, access control, and the safety; specify how it
-  affects composability, performance, and multi-tenancy (reviewer D)
-\end{enumerate}
-}
-\end{notes}
-
-\begin{abstract} Storage systems are caught between the need to evolve data
-processing applications efficiently and quickly, and the increasing velocity
-with which storage device technology evolves. This puts tremendous pressure on
-storage systems to support rapid change both in terms of their interfaces and
-their performance. But adapting storage systems can be difficult because
-unprincipled changes might jeopardize years of code-hardening and performance
-optimization efforts that were necessary for users to entrust their data to the
-storage system. We introduce the concept of a programmable storage system, a
-system that exposes internal services and abstractions with the goal of
-re-using and building new services on top of the storage stack. We present
-Malacology, a prototype implementation of a programmable storage system to
-explore how existing abstractions of common storage system services can be
-leveraged to adapt to the needs of new data processing systems and the
-increasing variety of storage devices. We show how this approach allows a large
-degree of flexibility for storage systems to evolve without sacrificing the
-robustness of their code-hardened subsystems. We illustrate the advantages and
-challenges of programmability by composing existing primitives into two new
-higher-level services: a metadata load balancer and a high-performance
-distributed shared-log. The evaluation demonstrates that our services inherit
-desirable qualities of the back-end storage system, including the ability to
-balance load, efficiently propagate cluster metadata, recover from failure, and
-to navigate trade-offs between latency and throughput using leases.
-\end{abstract}
-
-\begin{figure}[tb]
-\centering
-\includegraphics{figures/overview.png}
-\caption{Scalable storage systems have storage daemons which store data,
-monitor daemons (M) that maintain cluster state, and service-specific daemons
-({\it e.g.}, file system metadata servers). Malacology enables the programmability of
-internal abstractions (bold arrows) to re-use and compose
-existing subsystems.  With Malacology, we add new services, ZLog and
-Mantle, that sit alongside traditional user-facing APIs (file, block,
-object).}\label{fig:overview}
-\end{figure}
-
-\section{Introduction}
-\label{introduction}
-\label{sec:intro}
-
-A storage system implements abstractions designed to persistently store data
-and must exhibit a high level of correctness to prevent data loss.  Storage
-systems have evolved around storage devices that often were orders of magnitude
-slower than CPU and memory, and therefore could dominate overall performance if
-not used carefully. Over the last few decades members of the storage systems
-community have developed clever strategies to meet correctness requirements
-while somewhat hiding the latency of traditional storage
-media~\cite{brewer_disks_2016}. To avoid lock-in by a particular vendor, users
-of storage systems have preferred systems with highly standardized APIs and
-lowest common denominator abstract data types such as blocks of bytes and byte
-stream files~\cite{armbrust_view_2010}.
-
-A number of recent developments have disrupted traditional storage systems.
-First, the falling prices of flash storage and the availability of new types of
-non-volatile memory that are orders of magnitude faster than traditional
-spinning media are moving overall performance bottlenecks away from storage
-devices to CPUs and networking, and pressure storage systems to shorten their
-code paths and incorporate new
-optimizations~\cite{gray_tape_2007,gray_flash_2008}.  Second, emerging ``big
-data'' applications demand interface evolution to support flexible consistency
-as well as flexible structured data 
-representations.~\cite{apache_contributors_parquet_2014}.  Finally, production-quality scalable
-storage systems available as open source software have established and are
-continuing to establish new, \emph{de-facto} API standards at a faster pace
-than traditional standards
-bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
-
-The evolutionary pressure placed on storage systems by these trends raises the
-question of whether there are principles that storage systems designers can
-follow to evolve storage systems efficiently, without jeopardizing years of
-code-hardening and performance optimization efforts.  In this paper we
-investigate an approach that focuses on identifying and exposing existing
-storage system resources, services, and abstractions that in a generalized form
-can be used to \emph{program} new services. By composing higher-level services
-over existing abstractions one can reuse subsystems and leverage their
-optimizations, established correctness, robustness, and efficiency.
-
-{\it \textbf{Contribution 1:}} We define a programmable storage system to be a
-storage system that facilitates the re-use and extension of existing storage
-abstractions provided by the underlying software stack, to enable the creation
-of new services via composition.  A programmable storage system can be realized
-by exposing existing functionality (such as metadata services and
-synchronization and monitoring capabilities) as interfaces that can be ``glued
-together'' in a variety of ways using a high-level language. Programmable
-storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection
-and execution of code within a storage system or storage device---in that the
-former is applicable to any component of the storage system, while the latter
-focuses at the data access level. Given this contrast, we can say that active
-storage is an example of how one internal component (the storage layer) is
-exposed in a programmable storage system.
-
-To illustrate the benefits and challenges of this approach we have designed and
-evaluated Malacology, a programmable storage system that facilitates the
-construction of new services by re-purposing existing subsystem abstractions of
-the storage stack.  We build Malacology in Ceph, a popular open source software
-storage stack.  We choose Ceph to demonstrate the concept of programmable
-storage because it offers a broad spectrum of existing services, including
-distributed locking and caching services provided by metadata servers,
-durability and object interfaces provided by the backend object store, and
-propagation of consistent cluster state provided by the monitoring service (see
-Figure~\ref{fig:overview}).  Malacology is expressive enough to provide the
-functionality necessary for implementing new services.
-
-Malacology includes a non-exhaustive set of interfaces that can be used as
-building blocks for constructing novel storage abstractions, including:
-
-\begin{enumerate}
-
-\item An interface for managing strongly-consistent time-varying \textbf{service
-metadata}.
-
-\item An interface for installing and evolving domain-specific, cluster-wide
-\textbf{data interfaces}.
-
-\item An interface for managing access to \textbf{shared resources} using a
-variety of optimization strategies.
-
-\item An interface for \textbf{load balancing} resources across the cluster.
-
-\item An interface for \textbf{durability} that persists policies using the
-underlying storage stack's object store.
-
-\end{enumerate}
-
-{\it \textbf{Contribution 2:}} We implement two distributed services using Malacology to
-demonstrate the feasibility of the programmable storage approach:
-
-\begin{enumerate}
-
-\item A high-performance distributed shared log service called ZLog, that is an
-implementation of CORFU.~\cite{balakrishnan_corfu_2012}
-
-\item An implementation of Mantle, the programmable load balancing
-service~\cite{sevilla:sc15-mantle}
-
-\end{enumerate}
-
-The remainder of this paper is structured as follows. First, we describe and
-motivate the need for programmable storage by describing current practices in
-the open source software community. Next we describe Malacology by presenting the
-subsystems within the underlying storage system that we re-purpose, and briefly
-describe how those system are used within Malacology (\S\ref{sec:malacology}).
-Then we describe the services that we have constructed within the Malacology
-framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
-implementation (\S\ref{sec:evaluation}).  We conclude by discussing related and future work.
-% and
-%conclude.
-
-\section{Application-Specific Storage Stacks}
-\label{sec:application-specifc-storage-stacks}
-
-Building storage stacks from the ground up for a specific purpose results in
-the best performance. For example, GFS~\cite{ghemawat:sosp03} and
-HDFS~\cite{shvachko:msst10} were designed specifically to serve Map/Reduce and
-Hadoop jobs, and use techniques like exposing data locality and relaxing POSIX
-constraints to achieve application-specific I/O optimizations. Another example
-is Boxwood~\cite{maccormick:osdi04}, which experimented with B-trees
-and chunk stores as storage abstractions to simplify application building.
-Alternatively, general-purpose storage stacks are built with the flexibility to
-serve many applications by providing a variety of interfaces and tunable
-parameters.  Unfortunately, managing competing forces in these systems is
-difficult and users want more control from the general-purpose storage stacks
-without going as far as building their storage system from the ground up.
-
-To demonstrate a recent trend towards more application-specific storage systems
-we examine the state of programmability in Ceph. Something of a storage Swiss
-army knife, Ceph simultaneously supports file, block, and object interfaces on
-a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
-Distributed Object Storage (RADOS) system is a cluster of object storage
-devices (OSDs) that provide Ceph with data durability and integrity using
-replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already
-provides some degree of programmability; the OSDs support domain-specific
-object interfaces that are implemented by composing existing low-level storage
-interfaces that execute atomically. These interfaces are written in C++ and are
-statically loaded into the system.
-
-The Ceph community provides empirical evidence that developers are already
-beginning to embrace programmable storage. Figure~\ref{fig:obj-int-dev-growth}
-shows a dramatic growth in the production use of domain-specific interfaces in
-the Ceph community since 2010. In that figure, classes are functional groupings
-of methods on storage objects (e.g. remotely computing and caching the checksum
-of an object extent).  What is most remarkable is that this trend contradicts
-the notion that API changes are a burden for users.  Rather it appears that
-gaps in existing interfaces are being addressed through ad-hoc approaches to
-programmability. In fact, Table~\ref{table:objclasses} categorizes existing
-interfaces and we clearly see a trend towards reusable services.
-
-\begin{figure}[ht]
-\centering
-\includegraphics{figures/obj-int-dev-growth.png}
-\caption{Since 2010, the growth in the number of co-designed object storage interfaces
-in Ceph has been accelerating. This plot is the number of object classes (a
-group of interfaces), and the total number of methods (the actual API
-end-points).}
-\label{fig:obj-int-dev-growth}
-\end{figure}
-
-\begin{table}[ht]
-\centering
-  \begin{tabular}{l|l|l}
-    Category & Example & \# \\ \hline
-    Logging  & Geographically distribute replicas & 11 \\ \hdashline
-    Metadata & Snapshots in the block device OR  & \multirow{2}{*}{74} \\
-    Management & Scan extents for file system repair & \\ \hdashline
-    Locking  & Grants clients exclusive access & 6 \\ \hdashline
-    Other & Garbage collection, reference counting  & 4\\
-\end{tabular}
-\caption{A variety of object storage classes exist to expose interfaces
-    to applications. \# is the number of methods that implement these categories.
-}
-\label{table:objclasses}
-\end{table}
-
-The takeaway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
-already trying to use programmability because their needs, whether they be
-related to performance, availability, consistency, convenience, etc., are not
-satisfied by the existing default set of interfaces. The popularity of the
-custom object interface facility of Ceph could be due to a number of reasons,
-such as the default algorithms/tunables of the storage system being
-insufficient for the application's performance goals, programmers wanting to
-exploit application-specific semantics, and/or programmers knowing how to
-manage resources to improve performance. A solution based on
-application-specific object interfaces is a way to work around the
-traditionally rigid storage APIs because custom object interfaces give
-programmers the ability to tell the storage system about their application: if
-the application is CPU or IO bound, if it has locality, if its size has the potential to
-overload a single node, etc.  Programmers often know what the problem is
-and how to solve it, but until the ability to modify object interfaces, they
-had no way to express to the storage system how to handle their data.
-
-Our approach is to expose more of the commonly used, code-hardened subsystems
-of the underlying storage system as interfaces. The intent is that these
-interfaces, which can be as simple as a redirection to the persistent data
-store or as complicated as a strongly consistent directory service, should be
-used and re-used in many contexts to implement a wide range of services. By
-making programmability a `feature', rather than a `hack' or `workaround', we
-help standardize a development process that now is largely ad-hoc.
-
 \section{Challenges}
@@
 
-To address these challenges we present Malacology
-\oldcomment{. Malacology is both a prototype for a programmable storage system
-and a}
-\newcomment{, our prototype programmable storage system. It uses the
-programmable storage}
-\addressesissue{2}
-design approach to evolve storage systems efficiently and
+To address these challenges we present Malacology \oldcomment{. Malacology is
+both a prototype for a programmable storage system and a} \newcomment{, our
+prototype programmable storage system. It uses the programmable storage}
+\addressesissue{2} design approach to evolve storage systems efficiently and
 without jeopardizing years of code-hardening and performance optimization
@@
 internal abstractions nor do we contend that the abstractions we choose provide
-the best trade-offs for all applications.} 
-\newcomment{A programmable storage is not defined by what abstractions are
-exposed, rather a programmable storage system adheres to the design approach of
-choosing interfaces so administrators can have better control of the storage
-system. The interfaces presented in this paper are abstractions that we found
-useful for building ZLog and Mantle, yet they may not provide the best
-trade-offs for all higher-level services.}
-\addressesissue{2}
-For example, if consensus is correctly exposed one could implement
-high-level features like versioning, serialization, or various flavors of
-strongly consistent data management on top; but perhaps a low-level consensus
-interface is suited well for a particular set of applications.  These questions
-are not answered in this paper and instead we focus on showing the feasibility
-of building such a system, given advances in the quality and robustness of
-today's storage stacks.
+the best trade-offs for all applications.} \newcomment{A programmable storage
+is not defined by what abstractions are exposed, rather a programmable storage
+system adheres to the design approach of choosing interfaces so administrators
+can have better control of the storage system. The interfaces presented in this
+paper are abstractions that we found useful for building our prototype services
+ZLog and Mantle, yet they may not provide the best trade-offs for all
+higher-level services.} \addressesissue{2} For example, if consensus is
+correctly exposed one could implement high-level features like versioning,
+serialization, or various flavors of strongly consistent data management on
+top; but perhaps a low-level consensus interface is suited well for a
+particular set of applications.  These questions are not answered in this paper
+and instead we focus on showing the feasibility of building such a system,
+given advances in the quality and robustness of today's storage stacks.
 
-The Malacology prototype we present has been implemented on Ceph.
-While there are other systems on top of which Malacology could be implemented (see Table~\ref{table:examples}), we choose Ceph because it is a production quality system and because it
-is open source. The large developer community ensures that code is
-robust and the visibility of the code lets us expose any interface we want. In
-the next section we describe the Ceph components that we expose as Malacology
-interfaces.
-
-\section{Malacology: A Programmable Storage System}
-\label{sec:malacology}
-
-\begin{table*}
-\centering\small
-\begin{tabular}{ l | c | l | l | l }
-%\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
-\multicolumn{4}{c}{} \\
-\textbf{Interface}                      &
-\textbf{Sec}                            &
-\textbf{Example in Production Systems}  &
-\textbf{Example in Ceph}                &
-\textbf{Provided Primitives}            \\ \hline
-Service Metadata
-  & \S\ref{sec:service-metadata-interface}
-  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
-  & cluster state management~\cite{website:ceph-mon}
-  & consensus/consistency
-  \\
-Data I/O
-  & \S\ref{sec:data-io-interface}
-  & Swift in situ storage/compute~\cite{website:zerocloud}
-  & object interface classes~\cite{website:cls-lua}
-  & transaction/atomicity
-  \\
-Shared Resource
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & MPI collective IO, burst
-  & POSIX metadata protocols
-  & serialization/batching
-  \\
-File Type
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & ?????
-  & file striping strategy
-  & data/metadata access
-  \\
-Load Balancing
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
-  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
-  & migration/sampling
-  \\
-Durability
-  & \S\ref{sec:durability}
-  & S3/Swift interfaces (RESTful API)
-  & object store library~\cite{weil_rados_2007}
-  & persistence/safety
-  \\
-\end{tabular}
-\caption{Common internal abstractions. The same ``internal abstractions" are
-common across large-scale systems because they provide primitives that solve
-general distributed systems problems.  Here we list examples of what these
-internal abstractions are used for in ``production systems" and in Ceph.
-Malacology provides these internal abstractions as interfaces
-\newcomment{(Section~\ref{sec:malacology})}
-that higher level
-\oldcomment{applications}
-\newcomment{services (Section~\ref{sec:services})}
-can use.
-\oldcomment{; {\it italicized abstractions} are interfaces discussed in this
-paper.}}
-\addressesissue{1}
-\label{table:examples}
-\end{table*}
-
-The guiding principle is to re-use existing services and extend them so that
-these services can be \emph{programmed}. We accomplish programmability of a
-service by exporting bindings (or ``hooks'') for an interpreted programming
-language so that programming can occur without having to restart the storage
-system (see also below,~\S\ref{sec:durability}). Table~\ref{table:examples}
-shows the internal services from Ceph that we expose in the Malacology
-prototype via Lua~\cite{ierusalimschy1996lua} bindings
-\oldcomment{.}
-\newcomment{and Figure~\ref{fig:implementation-overview} compares what was
-already present in Ceph (gray boxes) to the Malacology interfaces we added.
-Section~\ref{sec:services} will describe the higher-level services we built
-with these interfaces.}
-\addressesissue{1}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation-overview.png}
-\caption{\newcomment{Malacology is implemented on the daemons and clients that
-run in a Ceph cluster. Interfaces expose internal subsystems and are used as
-building blocks for higher-level services.}\addressesissue{1}
-\label{fig:implementation-overview}}
-\end{figure}
-
-Lua is a portable embedded scripting language and we choose it as the
-interpreted language for Malacology because it offers superior performance and
-productivity trade-offs, including a JIT-based implementation that is well
-known for near native performance. Additionally, Lua has been used extensively
-in game engines, and systems research \cite{neto:dls14-luaos}, including
-storage systems where it has been effectively used both on
-\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010}
-and off \cite{sevilla:sc15-mantle} the performance critical path. Finally, the
-flexibility of the runtime allows execution sandboxing in order to address
-security and performance concerns. We will now discuss the common subsystems
-used to manage storage system and how Malacology makes them programmable.
-
-\subsection{Service Metadata Interface}
-\label{sec:mon}
-\label{sec:service-metadata-interface}
-\label{service-metadata}
-
-% straw man example
-Keeping track of state in a distributed system is an essential part of any
-successful service and a necessary component in order to diagnose and detect
-failures, when they occur. This is further complicated by variable propagation
-delays and heterogeneous hardware in dynamic environments.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In the case of Ceph, a consistent view of cluster state among server daemons
-and clients is critical to provide strong consistency guarantees to clients.
-Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
-metadata server) data structures called ``maps'' that record membership and
-status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
-responsible for integrating state changes into cluster maps, responding to
-requests from out-of-date clients and synchronizing members of the cluster
-whenever there is a change in a map so that they all observe the same system
-state. As a fundamental building block of many system designs, consensus
-abstractions such as Paxos are a common technique for maintaining consistent
-data versions, and are a useful system to expose.
-
-The default behavior of the monitor can be seen as a Paxos-based notification
-system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing
-clients to identify when new values (termed epochs in Ceph) are associated to
-given maps.  Since Ceph does not expose this service directly, as part of our Malacology implementation, we expose a key-value
-service designed for managing configuration metadata that is built on top of
-the consensus engine. Since the monitor is intended to be out of the
-high-performance I/O path, a general guideline is to make use of this
-functionality infrequently and to assign small values to maps.\\
-
-\noindent{\it{\textbf{Malacology: }}} Malacology exposes a strongly-consistent view of
-time-varying service metadata as a service rather than a hidden internal
-component. Malacology provides a generic API for adding arbitrary values to
-existing subsystem cluster maps. As a consequence of this, applications can
-define simple but useful service-specific logic to the strongly-consistent
-interface, such as authorization control (just specific clients can write new
-values) or triggering actions based on specific values (e.g. sanitize values).
-The higher level services we implement in \S\ref{sec:services} make use of this
-functionality to register, version and propagate dynamic code (Lua scripts) for
-new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
-policies in the load balancer \S\ref{malacology:mds}.  Using this service
-guarantees that interface definitions are not only made durable, but are
-transparently and consistently propagated throughout the cluster so that
-clients are properly synchronized with the latest interfaces.
-
-\newcomment{\\ \noindent \it{\textbf{Impact: }} provides core functionality
-because it lets daemons come to consensus on system critical state. Bugs in the
-internal subsystems or ommitting this from services that need this type of
-consistency affects correctness.}
-\addressesissue{3, 5}
-
-%And finally, it is important to not underestimate the importance of the
-%preservation of the modifications that enable a new system service. In many
-%(if not all) cases, interfaces defining access to data are just as important
-%as the data itself by virtue of inherently providing structural context. Thus,
-%all components of a system service must be kept to the same standards of
-%protection as data in the system itself.
-
-\oldcomment{\noindent 4.2 Object I/O Intefrace}
-\subsection{Data I/O Interface}
-\label{sec:data-io-interface}
-\label{object-data-interface}
-Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
-supports application-specific object
-interfaces~\cite{weil_rados_2007}. The ability to offload computation can
-reduce data movement, and transactional interfaces can significantly simplify
-construction of complex storage interfaces that require uncoordinated parallel
-access.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-An object interface is a plugin structured similarly to that of an RPC
-in which a developer creates a named function within the cluster that clients
-may invoke. In the case of Ceph each function is implicitly run within the
-context of an object specified when the function is called. Developers of
-object interfaces express behavior by creating a composition of native
-interfaces or other custom object interfaces, and handle serialization of
-function input and output.  A wide range of native interfaces are available to
-developers such as reading and writing to a byte stream, controlling object
-snapshots and clones, and accessing a sorted key-value database. These native
-interfaces may be transactionally composed along with application
-specific logic to create semantically rich interfaces. An example would be
-an interface that atomically updates a matrix stored in the bytestream and an
-index of the matrix stored in the key-value database.\\
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation.png}
-\caption{Malacology allows users to dynamically define object/metadata
-interfaces by extending the object storage daemon (OSD) and metadata server
-(MDS) subsystems with an embedded Lua VM.  It uses the service metadata
-interface (M) to propagate interfaces/versions across the cluster
-\label{fig:implementation}}
-\end{figure}
-
-\noindent {\it{\textbf{Malacology: }}} The implementation of Ceph's object abstraction,
-although powerful, does not readily support programmability. Supporting only
-C/C++ for object interface developers, Ceph requires distribution of compiled
-binaries for the correct architecture, adding a large barrier of entry for
-developers and system administrators. Second, having no way to dynamically
-unload modules, any changes require a full restart of a storage daemon which
-may have serious performance impacts due to loss of cached data. And finally,
-the security limitations of the framework limit the use of object interfaces to
-all but those with administrative level access and deep technical expertise.
-
-To address these concerns, Malacology takes advantage of Lua extensions
-contributed by the Ceph community. This allows new object interfaces to be
-dynamically loaded into the system and modified at runtime, resulting in a
-object storage API with economy of expression, which at the same time provides
-the full set of features of the base object class. New object interfaces that
-are expressed in thousands of lines of code can be implemented in approximately
-an order of magnitude less code~\cite{geambasu_comet_2010}. While the use of
-Lua does not prevent deployment of malicious code, certain types of coding
-mistakes can be handled gracefully, and access policies are used to limit access
-to trusted users~\cite{ierusalimschy1996lua}.
-
-\oldcomment{All of these modifications work in tandem to implement the desired behavior,
-and are typically co-designed together such that each depends on the other to
-behave as expected.}
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by pushing behavior to lower parts of the storage stack, thereby
-minimizing hops and distributing computation.}
-\addressesissue{3, 5}
-
-\subsection{Distributed Metadata Interfaces}
-\begin{notes}
-\textcolor{red}{This section has been re-organized}
-\end{notes}
-\label{sec:distributed-metadata-interfaces}
-\label{malacology:mds}
-
-\oldcomment{The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction~\cite{weil:sc2004-dyn-metadata}.}
-\newcomment{File systems provide clients with the familiar POSIX file abstraction.
-While this guarantees strong consistency it comes at the cost of scalability,
-increased complexity, and lower performance.}
-\addressesissue{1}
-In general, distributed file systems protect resources by providing
-hierarchical indexing and distributed locking services.
-
-%The metadata cluster moves units of the namespace called directory fragments
-%and assigns them to servers using its own metadata load balancer. It then uses
-%a hard-coded policy to balance these fragments using a scalarized load metric
-%based on CPU, workload, and file system operation metrics. Although the CephFS
-%balancer has been shown to be complicated, its complexity is justifiable.
-%Having the flexibility to choose what to move, where to move it, and how much
-%to move is very powerful. The problem is that it is difficult to pick a
-%one-size-fits-all policy, system administrators may want to partition load
-%based on SLAs, or applications are better equipped to make these decisions,
-%etc. -- but we choose instead to expose a balancing API at strategic points in
-%the balancing logic using Lua hooks. ~\cite{sevilla:sc15-mantle}.
-
-\oldcomment{\it{\textbf{Malacology:}} Interfaces are added in strategic spots in the
-metadata service for guarding shared resources, defining new file types,
-and load balancing.}
-
-\oldcomment{\noindent Shared resource interface.}
-\subsubsection{Shared Resource Interface}
-
-In general, file system metadata servers manages client sessions, allowing
-clients to obtain locks (e.g. file byte ranges), and capabilities (e.g. to
-cache file data). Clients and metadata servers use a cooperative protocol in
-which clients voluntarily release resources back to the metadata service in
-order to implement sharing policies.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In Ceph, the locking service implements a capability-based system that
-expresses what data and metadata clients are allowed to access as well as what
-state they may cache and modify locally.  While designed for the file
-abstraction, indexing, locking, and caching are all common services that are
-useful to a broad spectrum of applications.  Distributed applications that
-share centralized resources (e.g. a database or directory) face similar
-challenges which are often solved using application-specific sharding.
-
-\newcomment{\\ \noindent\it{\textbf{Malacology: }}}
-While the current policy for sharing access and voluntarily releasing resources
-is largely best-effort, Malacology supports generalized policies between
-metadata servers and clients that can be used to implement fairness or
-priority.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} provides core functionality
-to protect and provide exclusive access for any shared resource. May hurt
-performance if the resource in question does not require strong consistency.}
-\addressesissue{3}
-
-\oldcomment{\noindent File type interface.}
-\subsubsection{File Type Interface}
-The metadata service manages a POSIX file system hierarchy where files and
-directories are represented as inode data structures and expose a POSIX file
-interface. Applications that manage large amounts of metadata (e.g. users or
-database snapshots) often require a naming service. We allow new inode types to
-be defined such that applications can create domain-specific interfaces to
-inodes that may modify locking and capability policies. We will show how this
-is used in Section~\ref{sec:seq} when we discuss a distributed shared-log built
-on Malacology.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this interface is both a feature
-and a performance optimization. It is a feature because it allows developers to
-add support for different  storage types, such as how to read new file formats
-or what consistency semantics to use for a specific subtree in the hierarchical
-namespace. It is also a performance optimization because future programmers can
-add optimizations for processing specific types of files into the inode itself.}
-\addressesissue{3}
-
-%%% TODO: striping strategy is stored with the inode
-
-\oldcomment{\noindent \bf Load balancing interface.} 
-\subsubsection{Load Balancing Interface}
-
-Ceph also addresses the challenge of balancing metadata load with a separate
-metadata cluster. This cluster uses load balancing policies to migrate
-directory inodes around the cluster to alleviate overload at single
-nodes~\cite{weil:sc2004-dyn-metadata}. The policies use metrics based on system
-state (e.g.  CPU and memory utilization) and statistics collected by the
-cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
-partitioning to move variable sized namespace subtrees. These units can be
-shipped anywhere (i.e., to any metadata server of any capacity) at any time for
-any reason. The original balancer was designed with hard-coded policies and
-tunables.\\
-
-\noindent {\it{\textbf{Malacology: }}}
-Mantle~\cite{sevilla:sc15-mantle}, a programmable metadata load balancer, has
-been implemented on Malacology so that its infrastructure can enjoy all the
-properties of existing internal abstractions.  The load balancing logic is
-intact but the infrastructure has been improved to safely load, manage, and
-persist its load balancing policies.  Composing Mantle with the service
-metadata and durability interfaces, Mantle can safely version balancer
-policies, save balancer policies in the backend object store and centralize
-warnings/errors. When combined with file types, load balancing policies can
-express a variety of rules such as multi-tenant and multi-application
-balancing.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by allowing them to specify how to partition, replicate, and
-distribute metadata in response to overloaded servers.}
-\addressesissue{3}
-
-\subsection{Durability Interface}
-\label{sec:durability}
-
-Ceph provides storage by striping and replicating data across
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-protects data using common techniques such as erasure
-coding, replication, and scrubbing.  Furthermore, many of these features
-are implemented using a peer-to-peer protocol that allows object storage
-devices to operate autonomously without a centralized coordinator.  For
-example, when the number of placement groups change, the OSDs re-balance and
-re-shard data in the background in a process called placement group splitting
-in which OSDs communicate directly with each other to cover a new state.
-In order to reduce load on the monitoring service, Ceph OSDs use a gossip
-protocol to efficiently propagate changes to cluster maps throughout the
-system, and autonomously initiate recovery mechanisms when failures are
-discovered.\\
-
-\noindent {\it{\textbf{Malacology: }}} Metadata service policies and object storage
-interfaces are stored durability within RADOS and managed by storing references
-with the object maps. Since the cluster already propagates a consistent view of
-these data structures, we use this service to automatically install interfaces
-in OSDs, and install policies within the metadata server daemon (MDS) such that clients and daemons are
-synchronized on correct implementations without restarting.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this is a feature because it
-adds data safety and persistence to system metadata; while nice to have it does
-effect correctness.}
-\addressesissue{3}
-
-\section{Services Built on Malacology}
-\label{sec:services}
-\label{services-built-on-malacology}
-
-\label{services}
-%\begin{table}
-%\centering\small
-%\begin{tabular}{  l | l | l }
-%\textbf{Service}              &
-%\textbf{Used to...}           &
-%\textbf{Malacology Interface} \\ \hline
-%Mantle & load balance metadata     & Load Balancing  \\
-%       & agree on current balancer & Service Metadata\\
-%       & control client write lock & Shared Resource \\
-%       & persist policy logic      & Durability      \\ \hline
-%ZLog   & implement CORFU           & Object \& Data  \\ 
-%       & install ZLog interfaces   & Service Metadata\\
-%       & persist log entries       & Durability      \\ \hline
-%ZLog   & load balance sequencers   & Load Balancing  \\
-%Seqr.  & manage access to log tail & Shared Resource \\ 
-%\end{tabular}
-%\caption{Mantle and ZLog are built using Malacology interfaces so they inherit the
-%robustness of the subsystems of the underlying stack.}
-%\label{table:implementation}
-%\end{table}
-
-In this section we describe two services built on top of Malacology. The first
-is Mantle, a framework for dynamically specifying metadata load balancing
-policies. The second system, ZLog, is a high-performance distributed shared-log.
-In addition to these services, we'll demonstrate how we combine ZLog and Mantle
-to implement service-aware metadata load balancing policies.
-
-\subsection{Mantle: Programmable Load Balancer}
-\label{sec:mantle}
-
-Mantle is a programmable load balancer that separates the metadata balancing
-policies from their mechanisms. Administrators inject code to change how the
-metadata cluster distributes metadata. In ~\cite{sevilla:sc15-mantle} the
-authors showed how using Mantle, one can implement a single node metadata
-service, a distributed metadata service with hashing, and a distributed
-metadata service with dynamic subtree partitioning.
-
-The original implementation was ``hard-coded" into Ceph and lacked robustness
-(no versioning, durability, or policy distribution).  Re-implemented using
-Malacology, Mantle now enjoys (1) the versioning provided by the metadata
-service interface provided by the monitor daemons and (2) the durability and
-distribution provided by Ceph's reliable object store.  Re-using the internal
-abstractions with Malacology resulted in a 2\(\times\) reduction in source code
-compared to the original implementation.
-
-\subsubsection{Versioning Balancer Policies}
-
-Ensuring that the version of the current load balancer is consistent across the
-physical servers in the metadata cluster was not addressed in the original
-implementation. The user had to set the version on each individual server and
-it was trivial to make the versions inconsistent. Maintaining consistent
-versions is important for balancing policies that are cooperative in which local
-decisions are made assuming properties about other instances of the same
-policy.
-
-With Malacology, Mantle stores the version of the current load balancer in the
-Ceph service metadata interface. The version of the load balancer corresponds
-to an object name in the balancing policy. Using the service metadata interface
-ensures that all metadata servers use the same version of the load balancer. As
-a result, the policy version inherits the consistency of Ceph's internal
-monitor daemons. The user changes the version of the load balancer using a new
-CLI command.
-
-\subsubsection{Making Balancer Policies Durable}
-
-The load balancer version described above corresponds to the name of an object
-in RADOS that holds the actual Lua balancing code.  When MDS nodes start
-balancing load, they first check the latest version from the MDS map and
-compare it to the balancer they have loaded. If the version has changed, they
-dereference the pointer to the balancer version by reading the corresponding
-object in RADOS. This is in contrast to the original balancer which stored load
-balancer code on the local file system -- a technique which is unreliable and
-may result in silent corruption.
-
-The balancer pulls the Lua code from RADOS synchronously; asynchronous reads
-are not possible because of the architecture of the MDS. The synchronous
-behavior is not the default behavior for RADOS operations, so we achieve this
-with a timeout: if the asynchronous read does not come back within half the
-balancing tick interval the operation is canceled and a Connection Timeout
-error is returned. By default, the balancing tick interval is 10 seconds, so
-Mantle will use a 5 second second timeout.
-
-This design allows Mantle to immediately return an error if anything
-RADOS-related goes wrong.  We use this implementation because we do not want to
-do a blocking OSD read from inside the global MDS lock. Doing so would bring
-down the MDS cluster if any of the OSDs are not responsive.
-
-Storing the balancers in RADOS is simplified by the use of an interpreted
-language for writing balancer code. If we used a language that needs to be
-compiled, like the C++ object classes in the OSD, we would need to ensure
-binary compatibility, which is complicated by different operating systems,
-distributions, and compilers.
-
-\subsubsection{Logging, Debugging, and Warnings}
-
-In the original implementation, Mantle would log all errors, warnings, and
-debug messages to a log stored locally on each MDS server. To get the simplest
-status messages or to debug problems, the user would have to log into each MDS
-individually, look at the logs, and reason about causality and ordering.
-
-With Malacology, Mantle re-uses the centralized logging features of the
-monitoring service. Important errors, warnings, and info messages are collected
-by the monitoring subsystem and appear in the monitor cluster log so instead of
-users going to each node, they can watch messages appear at the monitor.
-Messages are logged sparingly, so as not to spam the monitor with frivolous
-debugging but important events, like balancer version changes or failed
-subsystems show up in the centralized log.
-
-\subsection{ZLog: A Fast Distributed Shared Log}
-\label{sec:zlog}
-
-The second service implemented on Malacology is ZLog, a high-performance
-distributed shared-log that is based on the CORFU
-protocol~\cite{balakrishnan_corfu_2012}. The shared-log is a
-powerful abstraction used to construct distributed systems, such as
-metadata management~\cite{balakrishnan:sosp13} and elastic database
-systems~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
-However, existing implementations that rely on consensus algorithms such as
-Paxos funnel I/O through a single point introducing a bottleneck that restricts
-throughput. In contrast, the CORFU protocol is able to achieve high throughput
-using a network counter called a \emph{sequencer}, that
-decouples log position assignment from log I/O.
-
-While a full description of the CORFU system is beyond the scope of this
-paper, we briefly describe the custom storage device interface, sequencer
-service, and recovery protocol, and how these services are instantiated in the
-Malacology framework.
-
-\subsubsection{Sequencer}
-\label{sec:seq}
-
-High-performance in CORFU is achieved using a sequencer service that assigns
-log positions to clients by reading from a volatile, in-memory counter which
-can run at a very high throughput and at low latency. Since the sequencer is
-centralized, ensuring serializability in the common case is trivial.  The
-primary challenge in CORFU is handling the failure of the sequencer in a way
-that preserves correctness. Failure of the sequencer service in CORFU is
-handled by a recovery algorithm that recomputes the new sequencer state using a
-CORFU-specific custom storage interface to discover the tail of the log, while
-simultaneously invalidating stale client requests using an epoch-based
-protocol.
-
-{\bf Sequencer interface.} The sequencer resource supports the ability to
-\emph{read()} the current tail value and get the \emph{next()} position in
-the log which also atomically increments the tail position.
-We implement the sequencer service in Malacology as an application-specific
-file type that associates a small amount of metadata
-with the inode state managed by the Ceph metadata service.
-This approach has the added
-benefit of allowing the metadata service to handle naming, by
-representing each sequencer instance in the standard POSIX hierarchical
-namespace. The primary challenge in mapping the sequencer resource to the
-metadata service is handling serialization correctly to maintain the global
-ordering provided by the CORFU protocol.
-
-Initially we sought to directly model the sequencer service in Ceph as a
-non-exclusive, non-cacheable resource, forcing clients to perform a round-trip
-access to the resource at the authoritative metadata server for the sequencer
-inode.  Interestingly, we found that the capability system in Ceph uses a
-strategy to reduce metadata service load by allowing clients that open a
-shared file to temporarily obtain an exclusive cached copy of the resource,
-resulting in a round-robin, best-effort batching behavior. When a single
-client is accessing the sequencer resource it is able to increment the
-sequencer locally. Any competing client cannot query the sequencer until the
-metadata service has granted it access.
-
-While unexpected, this discovery allowed us to explore an implementation
-strategy that we had not previously considered. In particular, for bursty
-clients, and clients that can tolerate increased latency, this mode of
-operation may allow a system to achieve much higher throughput than a system
-with a centralized sequencer service.  We utilize the programmability of the
-metadata service to define a new policy for handling capabilities that controls
-the amount of time that clients are able to cache the sequencer resource. This
-allows an administrator or application to control the trade-off between latency
-and throughput beyond the standard best-effort policy that is present in Ceph
-by default.
-
-In Section~\ref{sec:evaluation} we quantify the trade-offs of throughput and
-latency for an approach based on a round-robin batching mode, and compare this
-mode to one in which the metadata server mediates access to the sequencer state
-when it is being shared among multiple clients. Quantifying these trade-offs
-should provide administrators with guidelines for setting the tunables for
-different ``caching" modes of the sequencer.
-
-{\bf Balancing policies.} As opposed to the batching mode for controlling access to the sequencer
-resource, more predictable latency can be achieved by treating the sequencer
-inode as a shared non-cacheable resource, forcing clients to make a round-trip
-to the metadata service. However, the shared nature of the metadata service
-may prevent the sequencer from achieving maximum throughput. To provide a
-solution to this issue we have taken advantage of the programmability of the
-metadata service load balancing to construct a service-specific load balancing
-policy. As opposed to a balancing policy that strives for uniform
-load distribution, a ZLog-specific policy may utilize knowledge of inode types
-to migrate the sequencer service to provisioned hardware during periods of
-contention or high demand.
-
-\subsubsection{Storage Interface}
-
-The storage interface is a critical component in the CORFU protocol. Clients
-independently map log positions that they have obtained from the sequencer
-service (described in detail in the next section) onto storage devices, while
-storage devices provide an intelligent \emph{write}-once, random \emph{read}
-interface for accessing log entries. The key to correctness in CORFU lies with
-the enforcement of up-to-date epoch tags on client requests; requests tagged
-with out-of-date epoch values are rejected, and clients are expected to
-request a new tail from the sequencer after refreshing state from an auxiliary
-service.  This mechanism forms the basis for sequencer recovery.
-
-In order to repopulate the sequencer state (i.e. the cached, current tail of
-the log) during recovery of a sequencer, the maximum position in the log must
-be obtained. To do this, the storage interface exposes an additional
-\emph{seal} method that atomically installs a new epoch value and returns the
-maximum log position that has been written.
-
-Since the sequencer service does not resume until the recovery process has
-completed, there cannot be a race with clients appending to the log, and the
-immutability of the log allow reads to never block during a sequencer failure.
-Recovery of the sequencer process itself may be handled in many ways, such as
-leader election using an auxiliary service such as Paxos. In our
-implementation, the recovery is the same as (and is inherited from) the Ceph
-metadata service. Handling the failure of a client that holds the sequencer
-state is similar, although a timeout is used to determine when a client should
-be considered unavailable.
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-%We wanted to implement CORFU and we had two choices: build it from the ground
-%up or do it on top of Malacology. To show the feasability of building on
-%Malacology our evaluation focuses on the
-
-Our evaluation demonstrates the feasibility of building new service
-abstractions atop programmable storage, focusing on the performance of the
-internal abstractions exposed by Malacology and used to construct the Mantle
-and ZLog services. We also discuss latent capabilities we discovered in this
-process that let us navigate different trade-offs within the services
-themselves. First, we benchmark scenarios with high sequencer contention by
-examining the interfaces used to map ZLog onto Malacology; specifically, we
-describe the sequencer implementation and the propagation of object and data
-interfaces interfaces.  Next, we benchmark scenarios in which the storage
-system manages multiple logs by using Mantle to balance sequencers across a
-cluster.
-
-% should save this for future work, otherwise it undercuts the eval.
-%Detailed system-wide performance measurements and analysis is forthcoming for
-%both Mantle and ZLog, as we try to find the best metadata balancing policies
-%and physical design implementation trade-offs.
-
-Since this work focuses on the programmability of Malacology, the goal of this
-section is to show that the components and subsystems that support the
-Malacology interfaces provide reasonable relative performance, as well as to
-give examples of the flexibility that Malacology provides to programmers.
-
-%As we search for the best metadata balancing policies
-%and physical design implementation trade-offs,
-%we plan to rel
-
-\subsection{Mapping ZLog onto Malacology}
-\label{sec:mapping-zlog-onto-malacology}
-
-We evaluate Malacology by exploring one possible mapping of the ZLog
-implementation of CORFU onto Ceph in which we re-use (1) the metadata service
-to manage naming and synchronization of the sequencer resource by treating the
-resource as an inode, and (2) the monitoring sub-subsystem to distribute and
-install application-specific I/O interface required of the CORFU protocol. In
-Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the inode
-abstraction for implementing the sequencer resource enables load balancing policies
-to migrate the sequencer resource in heavy-load situation.
-
-\subsubsection{Sequencer Implementation}
-\label{sec:sequencer-implementation}
-
-We evaluate the feasibility of using the metadata service to implement a
-sequencer resource that is responsible for maintaining a total ordering of
-the log. Clients contact the sequencer to obtain the tail of the log and then
-independently initiate I/O, thus we measure both the throughput and latency of
-obtaining new tail positions which bounds client append performance.
-
-The sequencer is implemented as a custom file type in which the sequencer
-state (a 64-bit integer) is embedded in the inode of a file representing the
-sequencer. A total ordering of the log is imposed by the re-use of the
-capability service that can be used to grant exclusive access of inode state
-to clients. The metadata service is responsible for maintaining exclusivity
-and granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
-behavior of the system in which a best-effort policy is used.  The two colors
-represent points in time that the clients were able to access the resource.
-The best-effort policy shows a high degree of interleaving
-between clients but the system spends a large portion of time re-distributing
-the capability, reducing overall throughput.
-
-In order to control the performance of the system we implement a
-policy that (1) restricts the length of time that a client may maintain
-exclusive access and (2) limits the number of log positions that a client may
-generate without yielding to other clients waiting for access. The behavior of
-these two modes is illustrated in Figures~\ref{fig:capdelay-quota-behavior}
-(b) and (c), respectively.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/capdelay-quota-behavior-a.png}
-\includegraphics{figures/capdelay-quota-behavior-b.png}
-\includegraphics{figures/capdelay-quota-behavior-c.png}
-\caption{Each dot is an individual request, spread randomly along the \(y\)
-axis. The default behavior is unpredictable, ``delay" lets clients hold the
-lease longer, and ``quota" gives clients the lease for a number of operations.}
-\label{fig:capdelay-quota-behavior}
-\end{figure}
-
-Figure~\ref{fig:captp} demonstrates a configurable trade-off between
-throughput and latency. In the experiment two clients are run each with a
-fixed 0.25 second maximum reservation on the capability, and we vary the size
-of the log position quota running each configuration for two minutes. The
-total operations per second is the combined throughput of the two clients, and
-the average latency is number of microseconds required to obtain a new log
-position. With a small quota more time is spent exchanging exclusive access,
-while a large quota reservation allows clients to experience a much lower
-latency because they experience uncontended access for a longer period of time.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/tradeoff.png}
-\caption{Sequencer throughput by re-using various services.  The highest performance is
-achieved using a single client with exclusive, cacheable privilege. Round-robin
-sharing of the sequencer resource is affected by the amount of time the
-resource is held, with best-effort performing the worst.}
-\label{fig:captp}
-\end{figure}
-
-To get a better picture of latency, Figure~\ref{fig:capcdf} shows the CDF of
-latency for each client in all experiment configurations. At the 99th
-percentile clients accessed the sequencer is less than a millisecond. The CDF
-is cropped at the 99.999th percentile due to large outliers that we believe
-occur in instances in which the MDS is performing I/O while it is in the
-process of re-distributing the capability to another client.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/caps-delay-latency.png}
-\caption{
-The latency distribution shows that if the clients hold their capabilities longer,
-performance decreases. Changing the delay can help the sequencer and
-clients strike a balance between throughput and latency.}
-\label{fig:capcdf}
-\end{figure}
-
-\textbf{Malacology exposes} the internal capability management service and
-allows users to navigate latency and throughput trade-offs.  Other approaches to
-designing the sequencer service also exist, such as using a centralized service
-in which each access is a network round-trip. In contrast to the mechanism we
-explored which is appropriate for clients with bursty workloads, it may be
-easier to provide predictable performance using a centralized service, and we
-will be exploring in the future how this can be achieved using the capability
-system.
-
-\subsubsection{Interface Propagation}
-
-Domain-specific data interfaces (Section~\ref{sec:application-specifc-storage-stacks}) allow
-co-design between applications and the storage system. Malacology supports
-custom object interfaces in RADOS that require interface implementations to
-be installed on the storage devices in the system, supporting the evolution of
-interfaces through automatic system-wide versioning and installation through
-the service metadata interface (Section~\ref{service-metadata}). We evaluate
-the performance of installing a new interface version in the cluster,
-which is an important metric for applications that frequently evolve
-interfaces.
-
-We demonstrate the feasibility of utilizing the Ceph monitoring sub-system by
-evaluating the performance of installing and distributing interface updates.
-Figure~\ref{fig:propdelay} shows the CDF of the latency of interface updates.
-The interfaces are Lua scripts embedded in the cluster map and distributed
-using a peer-to-peer gossip protocol.  The latency is defined as the elapsed
-time following the Paxos proposal for an interface update until each OSD makes
-the update live (the cost of the Paxos proposal is configurable and is
-discussed below). The latency measurements were taken on the nodes running
-Ceph server daemons, and thus exclude the client round-trip cost. In each of
-the experiments 1000 interface updates were observed.
-
-Figure~\ref{fig:propdelay} shows the lower bound cost for updates in a large
-cluster. In the experiment labeled ``120 OSD (RAM)" a cluster of 120 OSDs (10
-OSDs x 12 servers) using an in-memory data store were deployed, showing a
-latency of less than 54 ms with a probability of 90\% and a worst case latency
-of 194 ms. These costs demonstrate the penalty of distributing the interface
-in a large cluster. In practice the costs include, in addition to cluster-wide
-propagation of interface updates, the network round-trip to the
-interface management service, the Paxos commit protocol itself, and other factors
-such as system load. By default Paxos proposals occur periodically with a 1
-second interval in order to accumulate updates. In a minimum, realistic quorum
-of 3 monitors using HDD-based storage, we were able to decrease this interval
-to an average of 222 ms.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=60mm,trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
-\caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
-interface commit.}
-\label{fig:propdelay}
-\end{figure}
-
-\subsection{Load Balancing ZLog Sequencers with Mantle}
-\label{sec:zlog-balancing}
-
-In practice, a storage system implementing CORFU will support a multiplicity of
-independent totally-ordered logs for each application.  For this scenario
-co-locating sequencers on the same physical node is not ideal but building a load
-balancer that can migrate the shared resource ({\it e.g.,}, the resource that
-mediates access to the tail of the log) is a time-consuming, non-trivial task.
-It requires building subsystems for migrating resources, monitoring the
-workloads, collecting metrics that describe the utilization on the physical
-nodes, partitioning resources, maintaining cache coherence, and managing
-multiple sequencers. The following experiments demonstrate the feasibility of
-using the mechanisms of the Malacology load balancing interface to inherit
-these features and to alleviate load from overloaded servers.
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-behaviors.png}
-\caption{CephFS/Mantle load balancing have better throughput than co-locating
-all sequencers on the same server.  Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify this improvement;
-Section~\ref{sec:feature-backoff} examines the migration at 0-60 seconds.
-}\label{fig:mantle-balancer-behaviors}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-performance.png}
-\includegraphics{figures/mantle-mode-performance.png}
-\caption{In (a) all CephFS balancing modes have the same performance; Mantle uses a
-balancer designed for sequencers. In (b) the best combination of mode and
-migration units can have up to a 2\(\times\) improvement.}\label{fig:mantle-balancer-performance}
-\end{figure}
-
-The experiments are run on a cluster with 10 nodes to store objects, one node
-to monitor the cluster, and 3 nodes that can accommodate sequencers. 
-Instead of measuring contention at the clients like
-Section~\ref{sec:sequencer-implementation}, these experiments measure
-contention at the sequencers by forcing clients to make round-trips for every
-request. We implement this using the shared resource interface that forces
-round-trips.  Because the sequencer's only function is to  hand out positions
-for the tail of the log, the workload is read-heavy.
-
-First, we show how the ZLog service can orchestrate multiple sequencers using
-the Malacology load balancing interface.
-Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput over time of
-different load balancers as they migrate 3 sequencers (with 4 clients) around
-the cluster; ``No Balancing" keeps all sequencers on one server, ``CephFS"
-migrates sequencers using the CephFS load balancers, and ``Mantle" uses a
-custom load balancer we wrote specifically for sequencers. The CephFS load
-balancers are hard-coded into Ceph while Mantle is an API for balancing load
-presented in~Mantle~\cite{sevilla:sc15-mantle} and re-implemented on top of
-Malacology in this paper.  The increased throughput for the CephFS and Mantle
-curves between 0 and 60 seconds are a result of migrating the sequencer(s) off
-overloaded servers.
-
-In addition to showing that migrating sequencers improves performance,
-Figure~\ref{fig:mantle-balancer-behaviors} also demonstrates features that we
-will explore in the rest of this section.
-Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify the differences in performance
-when the cluster stabilizes at time 100 seconds and
-Section~\ref{sec:feature-backoff} examines the slope and start time of the
-rebalancing phase between 0 and 60 seconds by comparing the aggressiveness of
-the balancers.
-
-\subsubsection{Feature: Balancing Modes}
-\label{sec:feature-balancing-modes}
-
-Next, we quantify the performance benefits shown in
-Figure~\ref{fig:mantle-balancer-behaviors}.  To understand why load balancers
-perform differently we need to explain the different balancing modes that the
-load balancer service uses and how they stress the subsystems that receive and
-forward client requests in different ways. In
-Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
-performance of the balancing mode that CephFS falls into {\it most of the time}.
-CephFS has 3 modes for balancing load: CPU mode, workload mode, and hybrid
-mode. All three have the same structure for making migration decisions but vary
-based on the metric used to calculate load. For this sequencer workload the 3
-different modes all have the same performance, shown in
-Figure~\ref{fig:mantle-balancer-performance} (a), because the load balancer falls
-into the same mode a majority of the time.  The high variation in performance
-for the CephFS CPU Mode bar reflects the uncertainty of using something as
-dynamic and unpredictable as CPU utilization to make migration decisions. In
-addition to the suboptimal performance and unpredictability, because
-CephFS balancers all fall into the same mode despite using different balancers
-is a problem because it prevents administrators from properly exploring the
-balancing state space.
-
-Mantle gives the administrator more control over balancing policies; for
-the Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the load
-balancing interface to program logic for balancing read-heavy workloads,
-resulting in better throughput and stability.  When we did this we also
-identified two balancing modes relevant for making migration decisions
-for sequencers.
-
-Using Mantle, the administrator can put the load balancing service into ``proxy
-mode" or ``client mode". In proxy mode one server receives all requests and
-farms off the requests to slave servers; the slaves servers do the actual tail
-finding operation. In client mode, clients interact directly with the server
-that has their sequencer.  These modes are illustrated in
-Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
-co-located on one physical server -- performance for that mode is shown by the
-No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
-Mode" in the lower left shows how clients for sequencer 2 are re-directed to
-server B. The two ``Proxy mode" diagrams on the right show how clients continue
-sending requests to server A even though sequencer 2 has migrated to server B
-(the meaning of ``(Half)" is discussed in the next section).
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-modes.png}
-\caption{ In client mode clients sending requests to the server that houses
-their sequencer. In proxy mode clients continue sending their requests to the
-first server.  }\label{fig:mantle-modes}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-mode-behavior.png}
-\caption{The performance of proxy mode achieves the highest throughput but at
-the cost of lower throughput for one of the sequencers. Client mode is more
-fair but results in lower cluster throughput.
-}\label{fig:mantle-mode-behavior}
-\end{figure}
-
-Figure~\ref{fig:mantle-mode-behavior} shows the throughput over time of the two
-different modes for an environment with only 2 sequencers (again 4 clients
-each) and 2 servers. The curves for both sequencers in
-Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second and
-at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
-Performance of Sequencer 2 decreases because it stayed on the proxy which now
-handles all client requests, processes requests for Sequencer 2, and forwards
-requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
-because distributing the sequencers in this way separates (1) the handling of
-the client requests and (2) finding the tail of the log and responding to
-clients.  Doing both steps is too heavy weight for one server and sequencers on
-slave nodes can go faster if work is split up; this phenomenon is not
-uncommon and has been observed in chain replication~\cite{chain_rep}.
-
-Cluster
-throughput improves at the cost of decreased throughput for Sequencer 2.
-Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer mode manually (no
-balancing phase) and shows that the cluster throughput is worse than the
-cluster throughput of proxy mode. That graph also shows that Sequencer 2 has
-less throughput than Sequencer 2. In this case, the scatter-gather process used
-for cache coherence in the metadata protocols causes strain on the server
-housing Sequencer 2 resulting in this uneven performance.
-
-\subsubsection{Feature: Migration Units}
-\label{sec:feature-migration-units}
-
-Another factor that affects performance in this environment is how much load
-is on each server; these experiments quantify that effect by programming the
-load balancing interface to control the amount of load to migrate. We call this
-metric a ``migration unit".  Expressing this heuristic is not easily achievable
-with outward facing tunable parameters (i.e. system knobs) but with Mantle's
-programmable interface, we can force the load balancer to change its migration
-units. To force the balancer into the Proxy Mode (Half) scenario in
-Figure~\ref{fig:mantle-modes}, which uses migration units equal to half the
-load on the current server, we can use: \texttt{ targets[whoami+1] = mds[whoami]["load"]/2
-}.
-
-This code snippet uses globally defined variables and tables from the Mantle
-API to send half of the load on the current server (whoami) to the next ranked
-server (whoami + 1); the \texttt{targets} array is a globally defined table that
-the balancer uses to do the migrations. Alternatively, to migrate all load a
-time step, we can remove the division by 2.
-
-%\begin{figure}[t!]
-%\centering
-%\includegraphics{figures/mantle-mode-performance.png}
-%\caption{The performance of the different modes using various migration units
-%shows almost a 2\(\times\) improvement in the best case.
-%}\label{fig:mantle-mode-performance}
-%\end{figure}
-
-Figure~\ref{fig:mantle-balancer-performance} (b) shows the performance of the modes
-using different migration units. Recall that this setup only has 2 sequencers
-and 2 servers, so performance may be different at scale. Even so, it is clear
-that client mode does not perform as well for read-heavy workloads. We even see
-a throughput improvement when migrating all load off the first server, leaving
-the first server to do administrative tasks (this is common in the metadata
-cluster because the first server does a lot of the cache coherence work) while
-the second server does all the client work. Proxy mode does the best in both
-cases and shows large performance gains when completely decoupling client
-request handling and operation processing in Proxy Mode (Full).  The parameter
-that controls the migration units helps the administrator control the sequencer
-co-location or distribution across the cluster. This trade-off was explored
-extensively in the Mantle paper but the experiments we present here are
-indicative of an even richer set of states to explore.
-
-\subsubsection{Feature: Backoff}
-\label{sec:feature-backoff}
-
-Tuning the aggressiveness of the load balancer decision making is also a
-trade-off that administrators can control and explore. The balancing phase from
-0 to 60 seconds in Figure~\ref{fig:mantle-balancer-behaviors} shows different
-degrees of aggressiveness in making migration decisions; CephFS makes a
-decision quickly and throughput jumps to 2500 ops/second within 10 seconds
-while Mantle takes more time to stabilize. This conservative behavior is
-controlled by programming the balancer to (1) use different conditions for when
-to migrate and (2) using a threshold for sustained overload.
-
-We control the conditions for when to migrate using \texttt{when()}, a callback
-in the Mantle API.  For the Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} we program \texttt{when()} to wait
-for load on the receiving server to fall below a threshold. This makes the
-balancer more conservative because it takes 60 seconds for cache coherence
-messages to settle.  The Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} also takes longer to reach peak
-throughput because we want the policy to wait to see how migrations affect the
-system before proceeding; the balancer does a migration right before 50
-seconds, realizes that there is a third underloaded server, and does another
-migration.
-
-The other way to change aggressiveness of the decision making is to program
-into the balancer a threshold for sustained overload. This forces the balancer
-to wait a certain number of iterations after a migration before proceeding. In
-Mantle, the policy would use the save state function to do a countdown after a
-migration.  Behavior graphs and performance numbers for this backoff feature is
-omitted for space considerations, but our experiments confirm that the more
-conservative the approach the less overall throughput.
-
-\textbf{Malacology pulls} the load balancing service out of the storage system
-to balance sequencers across a cluster. This latent capability also gives
-future programmers the ability to explore the different load balancing
-trade-offs including: load balancing modes to control forwarding vs. client
-redirection, load migration units to control sequencer distribution vs.
-co-location, and backoffs to control conservative vs. aggressive decision
-making.
-
-%\subsubsection{Experiment: Object Class
-%Programmability}\label{experiment-object-class-programmability}
-%
-%Add quote from CORFU paper that talks about what they had to hack
-%
-%\subsubsection{Experiment: Multi-Client
-%Burstiness}\label{experiment-multi-client-burstiness}
-%
-%\begin{figure}[htbp] \centering
-%\includegraphics{figures/caps-delay-thruput.png} \caption{Forcing the client
-%to drop their capabilities later (delay) improves throughput} \end{figure}
-
-\section{Related Work}
-
-% IOStack: Software-Defined Object Storage
-% sRoute: Treating the Storage Stack Like a Network
-   % - lots of good refs to check out
-% Information and control in gray-box systems
-
-Programmability of operating systems and networking resources, including
-distributed storage systems is not new, but we are not aware of work that makes
-generalization of existing services into programmable resources a key principle
-in storage systems design.
-
-Programmable storage systems can be viewed as an infrastructure for creating
-abstractions to better separate policies from mechanisms. This idea is not new.
-Software-defined networks (SDNs) create such an abstraction by separating the
-control plane from the data plane (see for example~\cite{jain:sigcomm13}). This
-notion of control/data separation was also applied in software-defined storage
-(SDS)~\cite{thereska:sosp13,stefanovici:fast16}. Similarly,
-IOStack~\cite{gracia:internet16} is providing policy-based provisioning and
-filtering in OpenStack Swift. According to a SNIA white
-paper~\cite{carlson:snia2014}, the primary goal of SDS is to control and
-facilitate flexible and dynamic provisioning of storage resources of different
-kinds, including flash memory and disk drives, to create a virtualized mapping
-between common storage abstractions (e.g. files, objects, and blocks) and
-storage devices taking data service objectives in terms of protection,
-availability, performance, and security into account. A programmable storage
-system exposes internal abstractions so that end users (not necessarily
-operators) can create new services on top of the storage stack. Thus, our
-notion of programmable storage differs from ``software-defined storage'' (SDS)
-in terms of goals and scope, although definitions of SDS are still in flux.
-
-Another view of programmable storage systems is one of tailoring systems
-resources to applications~\cite{arpaci:sosp01}. Related efforts include the
-Exokernel~\cite{engler:sosp95}, SPIN~\cite{bershad:sosp95} and
-Vino~\cite{seltzer:osdi96} projects; the latter two addressed the ability of injecting
-code into the kernel to specialize resource management. Another approach is to
-pass hints between the different layers of the IO stack to bridge the semantic
-gap between applications and
-storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}).
-
-Malacology uses the same Active and Typed Storage module presented in
-DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File Manifolds
-can be implemented with small changes to the Malacology framework, namely
-asynchronous object calls and Lua stubs in the inode, respectively.
-
-\section{Conclusion and Future Work}
-\label{conclusion-and-future-work}
-
-Programmable storage is a viable method for eliminating duplication of complex
-error-prone software used as workarounds for storage system
-deficiencies. We propose that systems expose their services in a safe way
-allowing application developers to customize system behavior to meet their
-needs while not sacrificing correctness. To illustrate the benefits of this
-approach we presented Malacology, a programmable storage system that
-facilitates the construction of new services by re-purposing existing subsystem
-abstractions of the storage stack.
-
-Future work will focus on constructing more interfaces to support a wide
-variety of storage system services that can be configured on-the-fly in existing
-systems. This work is one point along that path to producing general-purpose
-storage systems that can
-target  special-purpose applications.  Ultimately we want to utilize
-declarative methods for expressing new services.
-
-\bibliography{paper}
-\bibliographystyle{plain}
-%\printbibliography
-
-\end{document}
+The Malacology prototype we present has been implemented on Ceph.  While there
+are other systems on top of which Malacology could be implemented (see
+Table~\ref{table:examples}), we choose Ceph because it is a production quality
+system and because it is open source. The large developer community ensures
+that code is robust and the visibility of the code lets us expose any interface
+we want. In the next section we describe the Ceph components that we expose as
+Malacology interfaces.

### 2017-02-23 paper/4_malacology_a_programmable_storage_system.tex
paper: cleanup line breaks and break into separate files
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-02-13 paper/paper.tex
@@
-\documentclass[preprint]{sigplanconf-eurosys}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage{pifont}
-\usepackage{color}
-\usepackage{comment}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Uncomment for diff
-%\includecomment{notes}
-%\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
-%\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
-%\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
-
-% Uncomment for camera-ready
-\excludecomment{notes}
-\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
-\newcommand{\oldcomment}[1]{}
-\newcommand{\addressesissue}[1]{}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\cmark}{\ding{51}}%
-\usepackage[font=footnotesize,labelfont=bf]{caption}
-
-\setlength{\belowcaptionskip}{-10pt}
-
-\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
-\usepackage[normalem]{ulem}
-\usepackage{pstricks}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-\setkeys{Gin}{width=\maxwidth,keepaspectratio}
-
-\usepackage[unicode=true]{hyperref}
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, Peter Alvaro, Shel Finkelstein, Jeff LeFevre},
-            pdftitle={Malacology: A Programmable Storage System},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=blue,
-            linkcolor=black,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-
-\title{Malacology: A Programmable Storage System}
-%\authorinfo{Paper ID: 165}          {Total Pages: 12}
-\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre, Carlos Maltzahn}
-           {\{msevilla, jayhawk, ivo\}@soe.ucsc.edu, \{palvaro, shel, jlefevre, carlosm\}@ucsc.edu}
-\date{}
-\begin{document}
-
-\maketitle
-
-\begin{notes}
-\textcolor{red}{
-\noindent Issues raised by reviewers, paper augmented as noted inline (red/blue text):
-\begin{enumerate}
-  \item clarify what was implemented: diagram of interfaces and prose
-  specifying the difference between Malacology and prior work (reviewers B, C,
-  D, E)
-  \item describe what makes a programmable storage system (reviewers C, D)
-  \item organize motivation: group interfaces into categories (e.g., core
-  functionality, features, and performance optimizations) and describe other
-  applications we can build with Malacology (reviewers B, C)
-  \item clarify terminology: service vs. file system metadata, storage vs.
-  file system, dirty vs. clean slate approaches, soft-state caching (reviewers
-  B, C)
-  \item address security, access control, and the safety; specify how it
-  affects composability, performance, and multi-tenancy (reviewer D)
-\end{enumerate}
-}
-\end{notes}
-
-\begin{abstract} Storage systems are caught between the need to evolve data
-processing applications efficiently and quickly, and the increasing velocity
-with which storage device technology evolves. This puts tremendous pressure on
-storage systems to support rapid change both in terms of their interfaces and
-their performance. But adapting storage systems can be difficult because
-unprincipled changes might jeopardize years of code-hardening and performance
-optimization efforts that were necessary for users to entrust their data to the
-storage system. We introduce the concept of a programmable storage system, a
-system that exposes internal services and abstractions with the goal of
-re-using and building new services on top of the storage stack. We present
-Malacology, a prototype implementation of a programmable storage system to
-explore how existing abstractions of common storage system services can be
-leveraged to adapt to the needs of new data processing systems and the
-increasing variety of storage devices. We show how this approach allows a large
-degree of flexibility for storage systems to evolve without sacrificing the
-robustness of their code-hardened subsystems. We illustrate the advantages and
-challenges of programmability by composing existing primitives into two new
-higher-level services: a metadata load balancer and a high-performance
-distributed shared-log. The evaluation demonstrates that our services inherit
-desirable qualities of the back-end storage system, including the ability to
-balance load, efficiently propagate cluster metadata, recover from failure, and
-to navigate trade-offs between latency and throughput using leases.
-\end{abstract}
-
-\begin{figure}[tb]
-\centering
-\includegraphics{figures/overview.png}
-\caption{Scalable storage systems have storage daemons which store data,
-monitor daemons (M) that maintain cluster state, and service-specific daemons
-({\it e.g.}, file system metadata servers). Malacology enables the programmability of
-internal abstractions (bold arrows) to re-use and compose
-existing subsystems.  With Malacology, we add new services, ZLog and
-Mantle, that sit alongside traditional user-facing APIs (file, block,
-object).}\label{fig:overview}
-\end{figure}
-
-\section{Introduction}
-\label{introduction}
-\label{sec:intro}
-
-A storage system implements abstractions designed to persistently store data
-and must exhibit a high level of correctness to prevent data loss.  Storage
-systems have evolved around storage devices that often were orders of magnitude
-slower than CPU and memory, and therefore could dominate overall performance if
-not used carefully. Over the last few decades members of the storage systems
-community have developed clever strategies to meet correctness requirements
-while somewhat hiding the latency of traditional storage
-media~\cite{brewer_disks_2016}. To avoid lock-in by a particular vendor, users
-of storage systems have preferred systems with highly standardized APIs and
-lowest common denominator abstract data types such as blocks of bytes and byte
-stream files~\cite{armbrust_view_2010}.
-
-A number of recent developments have disrupted traditional storage systems.
-First, the falling prices of flash storage and the availability of new types of
-non-volatile memory that are orders of magnitude faster than traditional
-spinning media are moving overall performance bottlenecks away from storage
-devices to CPUs and networking, and pressure storage systems to shorten their
-code paths and incorporate new
-optimizations~\cite{gray_tape_2007,gray_flash_2008}.  Second, emerging ``big
-data'' applications demand interface evolution to support flexible consistency
-as well as flexible structured data 
-representations.~\cite{apache_contributors_parquet_2014}.  Finally, production-quality scalable
-storage systems available as open source software have established and are
-continuing to establish new, \emph{de-facto} API standards at a faster pace
-than traditional standards
-bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
-
-The evolutionary pressure placed on storage systems by these trends raises the
-question of whether there are principles that storage systems designers can
-follow to evolve storage systems efficiently, without jeopardizing years of
-code-hardening and performance optimization efforts.  In this paper we
-investigate an approach that focuses on identifying and exposing existing
-storage system resources, services, and abstractions that in a generalized form
-can be used to \emph{program} new services. By composing higher-level services
-over existing abstractions one can reuse subsystems and leverage their
-optimizations, established correctness, robustness, and efficiency.
-
-{\it \textbf{Contribution 1:}} We define a programmable storage system to be a
-storage system that facilitates the re-use and extension of existing storage
-abstractions provided by the underlying software stack, to enable the creation
-of new services via composition.  A programmable storage system can be realized
-by exposing existing functionality (such as metadata services and
-synchronization and monitoring capabilities) as interfaces that can be ``glued
-together'' in a variety of ways using a high-level language. Programmable
-storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection
-and execution of code within a storage system or storage device---in that the
-former is applicable to any component of the storage system, while the latter
-focuses at the data access level. Given this contrast, we can say that active
-storage is an example of how one internal component (the storage layer) is
-exposed in a programmable storage system.
-
-To illustrate the benefits and challenges of this approach we have designed and
-evaluated Malacology, a programmable storage system that facilitates the
-construction of new services by re-purposing existing subsystem abstractions of
-the storage stack.  We build Malacology in Ceph, a popular open source software
-storage stack.  We choose Ceph to demonstrate the concept of programmable
-storage because it offers a broad spectrum of existing services, including
-distributed locking and caching services provided by metadata servers,
-durability and object interfaces provided by the backend object store, and
-propagation of consistent cluster state provided by the monitoring service (see
-Figure~\ref{fig:overview}).  Malacology is expressive enough to provide the
-functionality necessary for implementing new services.
-
-Malacology includes a non-exhaustive set of interfaces that can be used as
-building blocks for constructing novel storage abstractions, including:
-
-\begin{enumerate}
-
-\item An interface for managing strongly-consistent time-varying \textbf{service
-metadata}.
-
-\item An interface for installing and evolving domain-specific, cluster-wide
-\textbf{data interfaces}.
-
-\item An interface for managing access to \textbf{shared resources} using a
-variety of optimization strategies.
-
-\item An interface for \textbf{load balancing} resources across the cluster.
-
-\item An interface for \textbf{durability} that persists policies using the
-underlying storage stack's object store.
-
-\end{enumerate}
-
-{\it \textbf{Contribution 2:}} We implement two distributed services using Malacology to
-demonstrate the feasibility of the programmable storage approach:
-
-\begin{enumerate}
-
-\item A high-performance distributed shared log service called ZLog, that is an
-implementation of CORFU.~\cite{balakrishnan_corfu_2012}
-
-\item An implementation of Mantle, the programmable load balancing
-service~\cite{sevilla:sc15-mantle}
-
-\end{enumerate}
-
-The remainder of this paper is structured as follows. First, we describe and
-motivate the need for programmable storage by describing current practices in
-the open source software community. Next we describe Malacology by presenting the
-subsystems within the underlying storage system that we re-purpose, and briefly
-describe how those system are used within Malacology (\S\ref{sec:malacology}).
-Then we describe the services that we have constructed within the Malacology
-framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
-implementation (\S\ref{sec:evaluation}).  We conclude by discussing related and future work.
-% and
-%conclude.
-
-\section{Application-Specific Storage Stacks}
-\label{sec:application-specifc-storage-stacks}
-
-Building storage stacks from the ground up for a specific purpose results in
-the best performance. For example, GFS~\cite{ghemawat:sosp03} and
-HDFS~\cite{shvachko:msst10} were designed specifically to serve Map/Reduce and
-Hadoop jobs, and use techniques like exposing data locality and relaxing POSIX
-constraints to achieve application-specific I/O optimizations. Another example
-is Boxwood~\cite{maccormick:osdi04}, which experimented with B-trees
-and chunk stores as storage abstractions to simplify application building.
-Alternatively, general-purpose storage stacks are built with the flexibility to
-serve many applications by providing a variety of interfaces and tunable
-parameters.  Unfortunately, managing competing forces in these systems is
-difficult and users want more control from the general-purpose storage stacks
-without going as far as building their storage system from the ground up.
-
-To demonstrate a recent trend towards more application-specific storage systems
-we examine the state of programmability in Ceph. Something of a storage Swiss
-army knife, Ceph simultaneously supports file, block, and object interfaces on
-a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
-Distributed Object Storage (RADOS) system is a cluster of object storage
-devices (OSDs) that provide Ceph with data durability and integrity using
-replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already
-provides some degree of programmability; the OSDs support domain-specific
-object interfaces that are implemented by composing existing low-level storage
-interfaces that execute atomically. These interfaces are written in C++ and are
-statically loaded into the system.
-
-The Ceph community provides empirical evidence that developers are already
-beginning to embrace programmable storage. Figure~\ref{fig:obj-int-dev-growth}
-shows a dramatic growth in the production use of domain-specific interfaces in
-the Ceph community since 2010. In that figure, classes are functional groupings
-of methods on storage objects (e.g. remotely computing and caching the checksum
-of an object extent).  What is most remarkable is that this trend contradicts
-the notion that API changes are a burden for users.  Rather it appears that
-gaps in existing interfaces are being addressed through ad-hoc approaches to
-programmability. In fact, Table~\ref{table:objclasses} categorizes existing
-interfaces and we clearly see a trend towards reusable services.
-
-\begin{figure}[ht]
-\centering
-\includegraphics{figures/obj-int-dev-growth.png}
-\caption{Since 2010, the growth in the number of co-designed object storage interfaces
-in Ceph has been accelerating. This plot is the number of object classes (a
-group of interfaces), and the total number of methods (the actual API
-end-points).}
-\label{fig:obj-int-dev-growth}
-\end{figure}
-
-\begin{table}[ht]
-\centering
-  \begin{tabular}{l|l|l}
-    Category & Example & \# \\ \hline
-    Logging  & Geographically distribute replicas & 11 \\ \hdashline
-    Metadata & Snapshots in the block device OR  & \multirow{2}{*}{74} \\
-    Management & Scan extents for file system repair & \\ \hdashline
-    Locking  & Grants clients exclusive access & 6 \\ \hdashline
-    Other & Garbage collection, reference counting  & 4\\
-\end{tabular}
-\caption{A variety of object storage classes exist to expose interfaces
-    to applications. \# is the number of methods that implement these categories.
-}
-\label{table:objclasses}
-\end{table}
-
-The takeaway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
-already trying to use programmability because their needs, whether they be
-related to performance, availability, consistency, convenience, etc., are not
-satisfied by the existing default set of interfaces. The popularity of the
-custom object interface facility of Ceph could be due to a number of reasons,
-such as the default algorithms/tunables of the storage system being
-insufficient for the application's performance goals, programmers wanting to
-exploit application-specific semantics, and/or programmers knowing how to
-manage resources to improve performance. A solution based on
-application-specific object interfaces is a way to work around the
-traditionally rigid storage APIs because custom object interfaces give
-programmers the ability to tell the storage system about their application: if
-the application is CPU or IO bound, if it has locality, if its size has the potential to
-overload a single node, etc.  Programmers often know what the problem is
-and how to solve it, but until the ability to modify object interfaces, they
-had no way to express to the storage system how to handle their data.
-
-Our approach is to expose more of the commonly used, code-hardened subsystems
-of the underlying storage system as interfaces. The intent is that these
-interfaces, which can be as simple as a redirection to the persistent data
-store or as complicated as a strongly consistent directory service, should be
-used and re-used in many contexts to implement a wide range of services. By
-making programmability a `feature', rather than a `hack' or `workaround', we
-help standardize a development process that now is largely ad-hoc.
-
-\section{Challenges}
-\label{sec:challenges}
-
-Implementing the infrastructure for programmability into existing services and
-abstractions of distributed storage systems is challenging, even if one assumes
-that the source code of the storage system and the necessary expertise for
-understanding it is available.  Some challenges include:
-
-\begin{itemize}
-
-\item Storage systems are generally required to be highly available so that any
-complete restarts of the storage system to reprogram them is usually
-unacceptable.
-
-\item Policies and optimizations are usually hard-wired into the services and
-one has to be careful when factoring them to avoid introducing additional bugs.
-These policies and optimizations are usually cross-cutting solutions to
-concerns or trade-offs that cannot be fully explored at the time the code is
-written (as they relate to workload or hardware). Given these policies and
-optimizations, decomposition of otherwise orthogonal internal abstractions can
-be difficult or dangerous.
-
-\item Mechanisms that are often only exercised according to hard-wired policies
-and not in their full generality have hidden bugs that are revealed as soon as
-those mechanisms are governed by different policies. In our experience
-introducing programmability into a storage system proved to be a great
-debugging tool.
-
-\item Programmability, especially in live systems, implies changes that need to
-be carefully managed by the system itself, including versioning and propagation
-of those changes without affecting correctness.
-
-\end{itemize}
-
-To address these challenges we present Malacology
-\oldcomment{. Malacology is both a prototype for a programmable storage system
-and a}
-\newcomment{, our prototype programmable storage system. It uses the
-programmable storage}
-\addressesissue{2}
-design approach to evolve storage systems efficiently and
-without jeopardizing years of code-hardening and performance optimization
-efforts.  Although Malacology uses the internal abstractions of the underlying
-storage system, including its subsystems, components, and implementations, we
-emphasize that our system still addresses the general challenges outlined
-above.
-
-The main challenge of designing a programmable storage system is choosing the
-right internal abstractions and picking the correct layers for exposing them.
-\oldcomment{In this paper, we do not present an exhaustive list of possible
-internal abstractions nor do we contend that the abstractions we choose provide
-the best trade-offs for all applications.} 
-\newcomment{A programmable storage is not defined by what abstractions are
-exposed, rather a programmable storage system adheres to the design approach of
-choosing interfaces so administrators can have better control of the storage
-system. The interfaces presented in this paper are abstractions that we found
-useful for building our prototype services ZLog and Mantle, yet they may not provide the best
-trade-offs for all higher-level services.}
-\addressesissue{2}
-For example, if consensus is correctly exposed one could implement
-high-level features like versioning, serialization, or various flavors of
-strongly consistent data management on top; but perhaps a low-level consensus
-interface is suited well for a particular set of applications.  These questions
-are not answered in this paper and instead we focus on showing the feasibility
-of building such a system, given advances in the quality and robustness of
-today's storage stacks.
-
-The Malacology prototype we present has been implemented on Ceph.
-While there are other systems on top of which Malacology could be implemented (see Table~\ref{table:examples}), we choose Ceph because it is a production quality system and because it
-is open source. The large developer community ensures that code is
-robust and the visibility of the code lets us expose any interface we want. In
-the next section we describe the Ceph components that we expose as Malacology
-interfaces.
-
 \section{Malacology: A Programmable Storage System}
@@
 \end{tabular}
-\caption{Common internal abstractions. The same ``internal abstractions" are
-common across large-scale systems because they provide primitives that solve
-general distributed systems problems.  Here we list examples of what these
-internal abstractions are used for in ``production systems" and in Ceph.
-Malacology provides these internal abstractions as interfaces
-\newcomment{(Section~\ref{sec:malacology})}
-that higher level
-\oldcomment{applications}
-\newcomment{services (Section~\ref{sec:services})}
-can use.
-\oldcomment{; {\it italicized abstractions} are interfaces discussed in this
-paper.}}
+\caption{Common internal abstractions. The same ``internal abstractions" common
+across large-scale systems because they provide primitives that solve general
+distributed systems problems.  Here we list examples of what these internal
+abstractions are used for in ``production systems" and in Ceph.  Malacology
+provides these internal abstractions as interfaces
+\newcomment{(Section~\ref{sec:malacology})} that higher level
+\oldcomment{applications} \newcomment{services (Section~\ref{sec:services})}
+can use.  \oldcomment{; {\it italicized abstractions} are interfaces discussed
+in this paper.}}
 \addressesissue{1}
@@
 shows the internal services from Ceph that we expose in the Malacology
-prototype via Lua~\cite{ierusalimschy1996lua} bindings
-\oldcomment{.}
+prototype via Lua~\cite{ierusalimschy1996lua} bindings\oldcomment{.}
 \newcomment{and Figure~\ref{fig:implementation-overview} compares what was
@@
 Section~\ref{sec:services} will describe the higher-level services we built
-with these interfaces.}
-\addressesissue{1}
+with these interfaces.} \addressesissue{1}
 
@@
 delays and heterogeneous hardware in dynamic environments.
+\newcommentone{Service metadata is information about the daemons in the system
+and includes membership details, hardware layout ({\it e.g.}, racks, power
+supplies, etc.), data layout, and daemon state/configuration. It differs from
+traditional file system metadata which is information about files. For the rest
+of the paper when we refer to metadata server or service, we mean a daemon that
+manages file system metadata.} \addressesissue{4}
 
-\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}}
-In the case of Ceph, a consistent view of cluster state among server daemons
-and clients is critical to provide strong consistency guarantees to clients.
-Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
-metadata server) data structures called ``maps'' that record membership and
-status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
-responsible for integrating state changes into cluster maps, responding to
-requests from out-of-date clients and synchronizing members of the cluster
-whenever there is a change in a map so that they all observe the same system
-state. As a fundamental building block of many system designs, consensus
-abstractions such as Paxos are a common technique for maintaining consistent
-data versions, and are a useful system to expose.
+\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}} In the
+case of Ceph, a consistent view of cluster state among server daemons and
+clients is critical to provide strong consistency guarantees to clients.  Ceph
+maintains cluster state information in per-subsystem  data structures called
+``maps'' that record membership and status information.  A
+Paxos~\cite{lamport_parttime_1998} monitoring service is responsible for
+integrating state changes into cluster maps, responding to requests from
+out-of-date clients and synchronizing members of the cluster whenever there is
+a change in a map so that they all observe the same system state. As a
+fundamental building block of many system designs, consensus abstractions such
+as Paxos are a common technique for maintaining consistent data versions, and
+are a useful system to expose.
 
@@
 clients to identify when new values (termed epochs in Ceph) are associated to
-given maps.  Since Ceph does not expose this service directly, as part of our Malacology implementation, we expose a key-value
-service designed for managing configuration metadata that is built on top of
-the consensus engine. Since the monitor is intended to be out of the
+given maps.  Since Ceph does not expose this service directly, as part of our
+Malacology implementation, we expose a key-value service designed for managing
+\oldcommentone{configuration} \newcommentone{service} metadata that is built on
+top of the consensus engine. Since the monitor is intended to be out of the
 high-performance I/O path, a general guideline is to make use of this
@@
 
-\noindent{\it{\textbf{Malacology: }}} Malacology exposes a strongly-consistent view of
-time-varying service metadata as a service rather than a hidden internal
-component. Malacology provides a generic API for adding arbitrary values to
-existing subsystem cluster maps. As a consequence of this, applications can
-define simple but useful service-specific logic to the strongly-consistent
-interface, such as authorization control (just specific clients can write new
-values) or triggering actions based on specific values (e.g. sanitize values).
-The higher level services we implement in \S\ref{sec:services} make use of this
-functionality to register, version and propagate dynamic code (Lua scripts) for
-new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
-policies in the load balancer \S\ref{malacology:mds}.  Using this service
-guarantees that interface definitions are not only made durable, but are
-transparently and consistently propagated throughout the cluster so that
-clients are properly synchronized with the latest interfaces.\\
+\noindent{\it{\textbf{Malacology: }}} Malacology exposes a strongly-consistent
+view of time-varying service metadata as a service rather than a hidden
+internal component. Malacology provides a generic API for adding arbitrary
+values to existing subsystem cluster maps. As a consequence of this,
+applications can define simple but useful service-specific logic to the
+strongly-consistent interface, such as authorization control (just specific
+clients can write new values) or triggering actions based on specific values
+(e.g. sanitize values).  The higher-level services we implement in
+\S\ref{sec:services} make use of this functionality to register, version and
+propagate dynamic code (Lua scripts) for new object interfaces defined in
+storage daemons (\S\ref{object-data-interface}) and policies in the load
+balancer \S\ref{malacology:mds}.  Using this service guarantees that interface
+definitions are not only made durable, but are transparently and consistently
+propagated throughout the cluster so that clients are properly synchronized
+with the latest interfaces.\\
 
@@
 internal subsystems or ommitting this from services that need this type of
-consistency affects correctness.}
-\addressesissue{3, 5}
+consistency affects correctness.} \addressesissue{3, 5}
 
@@
 \label{object-data-interface}
+
 Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
-supports application-specific object
-interfaces~\cite{weil_rados_2007}. The ability to offload computation can
-reduce data movement, and transactional interfaces can significantly simplify
-construction of complex storage interfaces that require uncoordinated parallel
-access.
+supports application-specific object interfaces~\cite{weil_rados_2007}. The
+ability to offload computation can reduce data movement, and transactional
+interfaces can significantly simplify construction of complex storage
+interfaces that require uncoordinated parallel access.
 
-\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}}
-An object interface is a plugin structured similarly to that of an RPC
-in which a developer creates a named function within the cluster that clients
-may invoke. In the case of Ceph each function is implicitly run within the
-context of an object specified when the function is called. Developers of
-object interfaces express behavior by creating a composition of native
-interfaces or other custom object interfaces, and handle serialization of
-function input and output.  A wide range of native interfaces are available to
-developers such as reading and writing to a byte stream, controlling object
-snapshots and clones, and accessing a sorted key-value database. These native
-interfaces may be transactionally composed along with application
-specific logic to create semantically rich interfaces. An example would be
-an interface that atomically updates a matrix stored in the bytestream and an
-index of the matrix stored in the key-value database.\\
+\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}} An object
+interface is a plugin structured similarly to that of an RPC in which a
+developer creates a named function within the cluster that clients may invoke.
+In the case of Ceph each function is implicitly run within the context of an
+object specified when the function is called. Developers of object interfaces
+express behavior by creating a composition of native interfaces or other custom
+object interfaces, and handle serialization of function input and output.  A
+wide range of native interfaces are available to developers such as reading and
+writing to a byte stream, controlling object snapshots and clones, and
+accessing a sorted key-value database. These native interfaces may be
+transactionally composed along with application specific logic to create
+semantically rich interfaces. An example would be an interface that atomically
+updates a matrix stored in the bytestream and an index of the matrix stored in
+the key-value database.\\
 
@@
 \includegraphics{figures/implementation.png}
-\caption{Malacology allows users to dynamically define object/metadata
-interfaces by extending the object storage daemon (OSD) and metadata server
-(MDS) subsystems with an embedded Lua VM.  It uses the service metadata
-interface (M) to propagate interfaces/versions across the cluster
-\label{fig:implementation}}
+\caption{Malacology allows users to dynamically define
+object/\newcommentone{file system} metadata interfaces by extending the object
+storage daemon (OSD) and metadata server (MDS) subsystems with an embedded Lua
+VM.  It uses the service metadata interface (M) to propagate
+interfaces/versions across the cluster \label{fig:implementation}}
 \end{figure}
 
-\noindent {\it{\textbf{Malacology: }}} The implementation of Ceph's object abstraction,
-although powerful, does not readily support programmability. Supporting only
-C/C++ for object interface developers, Ceph requires distribution of compiled
-binaries for the correct architecture, adding a large barrier of entry for
-developers and system administrators. Second, having no way to dynamically
-unload modules, any changes require a full restart of a storage daemon which
-may have serious performance impacts due to loss of cached data. And finally,
-the security limitations of the framework limit the use of object interfaces to
-all but those with administrative level access and deep technical expertise.
+\noindent {\it{\textbf{Malacology: }}} The implementation of Ceph's object
+abstraction, although powerful, does not readily support programmability.
+Supporting only C/C++ for object interface developers, Ceph requires
+distribution of compiled binaries for the correct architecture, adding a large
+barrier of entry for developers and system administrators. Second, having no
+way to dynamically unload modules, any changes require a full restart of a
+storage daemon which may have serious performance impacts due to loss of cached
+data. And finally, the security limitations of the framework limit the use of
+object interfaces to all but those with administrative level access and deep
+technical expertise.
 
@@
 Lua does not prevent deployment of malicious code, certain types of coding
-mistakes can be handled gracefully, and access policies are used to limit access
-to trusted users~\cite{ierusalimschy1996lua}.
+mistakes can be handled gracefully, and access policies are used to limit
+access to trusted users~\cite{ierusalimschy1996lua}.
 
-\oldcomment{All of these modifications work in tandem to implement the desired behavior,
-and are typically co-designed together such that each depends on the other to
-behave as expected.}
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by pushing behavior to lower parts of the storage stack, thereby
-minimizing hops and distributing computation.}
+\oldcomment{All of these modifications work in tandem to implement the desired
+behavior, and are typically co-designed together such that each depends on the
+other to behave as expected.} \newcomment{\\ \noindent {\it{\textbf{Impact: }}}
+helps applications optimize performance by pushing behavior to lower parts of
+the storage stack, thereby minimizing hops and distributing computation.}
 \addressesissue{3, 5}
@@
 
-\oldcomment{The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction~\cite{weil:sc2004-dyn-metadata}.}
-\newcomment{File systems provide clients with the familiar POSIX file abstraction.
-While this guarantees strong consistency it comes at the cost of scalability,
-increased complexity, and lower performance.}
-\addressesissue{1}
+\oldcomment{The distributed metadata service in Ceph provides clients with a
+POSIX file system abstraction~\cite{weil:sc2004-dyn-metadata}.}
+\newcomment{File systems provide clients with the familiar POSIX file
+abstraction.  While this guarantees strong consistency it comes at the cost of
+scalability, increased complexity, and lower performance.} \addressesissue{1}
 In general, distributed file systems protect resources by providing
@@
 
-\oldcomment{\it{\textbf{Malacology:}} Interfaces are added in strategic spots in the
-metadata service for guarding shared resources, defining new file types,
-and load balancing.}
+\oldcomment{\it{\textbf{Malacology:}} Interfaces are added in strategic spots
+in the \newcommentone{file system} metadata servers for guarding shared
+resources, defining new file types, and load balancing.}
 
@@
 cache file data). Clients and metadata servers use a cooperative protocol in
-which clients voluntarily release resources back to the metadata service in
-order to implement sharing policies.
+which clients voluntarily release resources back to the \newcommentone{file
+system} metadata service in order to implement sharing policies.
 
-\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}}
-In Ceph, the locking service implements a capability-based system that
-expresses what data and metadata clients are allowed to access as well as what
-state they may cache and modify locally.  While designed for the file
-abstraction, indexing, locking, and caching are all common services that are
-useful to a broad spectrum of applications.  Distributed applications that
+\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}} In Ceph,
+the locking service implements a capability-based system that expresses what
+data and \newcommentone{file system} metadata clients are allowed to access as
+well as what state they may cache and modify locally.  While designed for the
+file abstraction, indexing, locking, and caching are all common services that
+are useful to a broad spectrum of applications.  Distributed applications that
 share centralized resources (e.g. a database or directory) face similar
@@
 
-\newcomment{\\ \noindent\it{\textbf{Malacology: }}}
-While the current policy for sharing access and voluntarily releasing resources
-is largely best-effort, Malacology supports generalized policies between
-metadata servers and clients that can be used to implement fairness or
-priority.
+\newcomment{\\ \noindent\it{\textbf{Malacology: }}} While the current policy
+for sharing access and voluntarily releasing resources is largely best-effort,
+Malacology supports generalized policies between metadata servers and clients
+that can be used to implement fairness or priority.
 
@@
 
-Applications that manage large amounts of metadata (e.g. users or database
-snapshots) often require a naming service.  The metadata service manages a
-POSIX file system hierarchy where files and directories are represented as
-inode data structures and expose a POSIX file interface.
+Applications that manage large amounts of \newcommentone{file system} metadata
+(e.g. users or database snapshots) often require a naming service.  The
+metadata service manages a POSIX file system hierarchy where files and
+directories are represented as inode data structures and expose a POSIX file
+interface.
 
@@
 
-\newcomment{\\ \noindent {\it{\textbf{Malacology: }}}} We allow new inode types to
-be defined such that applications can create domain-specific interfaces to
+\newcomment{\\ \noindent {\it{\textbf{Malacology: }}}} We allow new inode types
+to be defined such that applications can create domain-specific interfaces to
 inodes that may modify locking and capability policies. We will show how this
@@
 
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this interface is both a feature
-and a performance optimization. It is a feature because it allows developers to
-add support for different  storage types, such as how to read new file formats
-or what consistency semantics to use for a specific subtree in the hierarchical
-namespace. It is also a performance optimization because future programmers can
-add optimizations for processing specific types of files into the inode itself.}
-\addressesissue{3}
+\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this interface is both a
+feature and a performance optimization. It is a feature because it allows
+developers to add support for different  storage types, such as how to read new
+file formats or what consistency semantics to use for a specific subtree in the
+hierarchical namespace. It is also a performance optimization because future
+programmers can add optimizations for processing specific types of files into
+the inode itself.} \addressesissue{3}
 
@@
 
-\newcomment{Many large scale storage systems separate metadata and data IO so
-that the corresponding services can scale independently. Metadata requests
-transfer small amounts of data and they happen relatively frequently so many
-systems employ separate metadata clusters.}
+\newcomment{Many large scale storage systems separate file system metadata and
+data IO so that the corresponding services can scale independently. Metadata
+requests transfer small amounts of data and they happen relatively frequently
+so many systems employ separate file system metadata clusters.}
 
 \newcomment{\\ \noindent {\it{\textbf{Existing Ceph Implementation: }}}} Ceph
-also addresses the challenge of balancing metadata load with a separate
-metadata cluster. This cluster uses load balancing policies to migrate
-directory inodes around the cluster to alleviate overload at single
-nodes~\cite{weil:sc2004-dyn-metadata}.  The policies use metrics based on
-system state (e.g.  CPU and memory utilization) and statistics collected by the
-cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
-partitioning to move variable sized namespace subtrees. These units can be
-shipped anywhere (i.e., to any metadata server of any capacity) at any time for
-any reason. The original balancer was designed with hard-coded policies and
+also addresses the challenge of balancing \newcommentone{file system} metadata
+load with a separate metadata cluster. This cluster uses load balancing
+policies to migrate directory inodes around the cluster to alleviate overload
+at single nodes~\cite{weil:sc2004-dyn-metadata}.  The policies use metrics
+based on system state (e.g.  CPU and memory utilization) and statistics
+collected by the cluster (e.g. the popularity of an inode). Ceph uses dynamic
+subtree partitioning to move variable sized namespace subtrees. These units can
+be shipped anywhere (i.e., to any metadata server of any capacity) at any time
+for any reason. The original balancer was designed with hard-coded policies and
 tunables.\\
 
-\noindent {\it{\textbf{Malacology: }}}
-Mantle~\cite{sevilla:sc15-mantle}, a programmable metadata load balancer, has
-been implemented on Malacology so that its infrastructure can enjoy all the
+\noindent {\it{\textbf{Malacology: }}} Mantle~\cite{sevilla:sc15-mantle}, a
+programmable \newcomment{file system}metadata load balancer, has been
+implemented on Malacology so that its infrastructure can enjoy all the
 properties of existing internal abstractions.  The load balancing logic is
@@
 performance by allowing them to specify how to partition, replicate, and
-distribute metadata in response to overloaded servers.}
-\addressesissue{3}
+distribute metadata in response to overloaded servers.} \addressesissue{3}
 
@@
 
-\noindent{\it{\textbf{Existing Ceph Implementation: }}}
-Ceph provides storage by striping and replicating data across
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-protects data using common techniques such as erasure coding, replication, and
-scrubbing.  example, when the number of placement groups change, the OSDs
-re-balance and re-shard data in the background in a process called placement
-group splitting in which OSDs communicate directly with each other to cover a
-new state.  In order to reduce load on the monitoring service, Ceph OSDs use a
-gossip protocol to efficiently propagate changes to cluster maps throughout the
-system, and autonomously initiate recovery mechanisms when failures are
-discovered.\\
+\noindent{\it{\textbf{Existing Ceph Implementation: }}} Ceph provides storage
+by striping and replicating data across RADOS~\cite{weil_rados_2007}, the
+reliable distributed object store. RADOS protects data using common techniques
+such as erasure coding, replication, and scrubbing.  example, when the number
+of placement groups change, the OSDs re-balance and re-shard data in the
+background in a process called placement group splitting in which OSDs
+communicate directly with each other to cover a new state.  In order to reduce
+load on the monitoring service, Ceph OSDs use a gossip protocol to efficiently
+propagate changes to cluster maps throughout the system, and autonomously
+initiate recovery mechanisms when failures are discovered.\\
 
-\noindent {\it{\textbf{Malacology: }}} Metadata service policies and object storage
-interfaces are stored durability within RADOS and managed by storing references
-with the object maps. Since the cluster already propagates a consistent view of
-these data structures, we use this service to automatically install interfaces
-in OSDs, and install policies within the metadata server daemon (MDS) such that clients and daemons are
-synchronized on correct implementations without restarting.
+\noindent {\it{\textbf{Malacology: }}} Metadata service policies and object
+storage interfaces are stored durability within RADOS and managed by storing
+references with the object maps. Since the cluster already propagates a
+consistent view of these data structures, we use this service to automatically
+install interfaces in OSDs, and install policies within the metadata server
+daemon \oldcommentone{(MDS)} such that clients and daemons are synchronized on
+correct implementations without restarting.
 
@@
 adds data safety and persistence to system metadata; while nice to have it does
-not necessarily effect correctness.}
-\addressesissue{3}
-
-\section{Services Built on Malacology}
-\label{sec:services}
-\label{services-built-on-malacology}
-
-\label{services}
-%\begin{table}
-%\centering\small
-%\begin{tabular}{  l | l | l }
-%\textbf{Service}              &
-%\textbf{Used to...}           &
-%\textbf{Malacology Interface} \\ \hline
-%Mantle & load balance metadata     & Load Balancing  \\
-%       & agree on current balancer & Service Metadata\\
-%       & control client write lock & Shared Resource \\
-%       & persist policy logic      & Durability      \\ \hline
-%ZLog   & implement CORFU           & Object \& Data  \\ 
-%       & install ZLog interfaces   & Service Metadata\\
-%       & persist log entries       & Durability      \\ \hline
-%ZLog   & load balance sequencers   & Load Balancing  \\
-%Seqr.  & manage access to log tail & Shared Resource \\ 
-%\end{tabular}
-%\caption{Mantle and ZLog are built using Malacology interfaces so they inherit the
-%robustness of the subsystems of the underlying stack.}
-%\label{table:implementation}
-%\end{table}
-
-In this section we describe two services built on top of Malacology. The first
-is Mantle, a framework for dynamically specifying metadata load balancing
-policies. The second system, ZLog, is a high-performance distributed shared-log.
-In addition to these services, we'll demonstrate how we combine ZLog and Mantle
-to implement service-aware metadata load balancing policies.
-
-\subsection{Mantle: Programmable Load Balancer}
-\label{sec:mantle}
-
-Mantle is a programmable load balancer that separates the metadata balancing
-policies from their mechanisms. Administrators inject code to change how the
-metadata cluster distributes metadata. In ~\cite{sevilla:sc15-mantle} the
-authors showed how using Mantle, one can implement a single node metadata
-service, a distributed metadata service with hashing, and a distributed
-metadata service with dynamic subtree partitioning.
-
-The original implementation was ``hard-coded" into Ceph and lacked robustness
-(no versioning, durability, or policy distribution).  Re-implemented using
-Malacology, Mantle now enjoys (1) the versioning provided by the metadata
-service interface provided by the monitor daemons and (2) the durability and
-distribution provided by Ceph's reliable object store.  Re-using the internal
-abstractions with Malacology resulted in a 2\(\times\) reduction in source code
-compared to the original implementation.
-
-\subsubsection{Versioning Balancer Policies}
-
-Ensuring that the version of the current load balancer is consistent across the
-physical servers in the metadata cluster was not addressed in the original
-implementation. The user had to set the version on each individual server and
-it was trivial to make the versions inconsistent. Maintaining consistent
-versions is important for balancing policies that are cooperative in which local
-decisions are made assuming properties about other instances of the same
-policy.
-
-With Malacology, Mantle stores the version of the current load balancer in the
-Ceph service metadata interface. The version of the load balancer corresponds
-to an object name in the balancing policy. Using the service metadata interface
-ensures that all metadata servers use the same version of the load balancer. As
-a result, the policy version inherits the consistency of Ceph's internal
-monitor daemons. The user changes the version of the load balancer using a new
-CLI command.
-
-\subsubsection{Making Balancer Policies Durable}
-
-The load balancer version described above corresponds to the name of an object
-in RADOS that holds the actual Lua balancing code.  When MDS nodes start
-balancing load, they first check the latest version from the MDS map and
-compare it to the balancer they have loaded. If the version has changed, they
-dereference the pointer to the balancer version by reading the corresponding
-object in RADOS. This is in contrast to the original balancer which stored load
-balancer code on the local file system -- a technique which is unreliable and
-may result in silent corruption.
-
-The balancer pulls the Lua code from RADOS synchronously; asynchronous reads
-are not possible because of the architecture of the MDS. The synchronous
-behavior is not the default behavior for RADOS operations, so we achieve this
-with a timeout: if the asynchronous read does not come back within half the
-balancing tick interval the operation is canceled and a Connection Timeout
-error is returned. By default, the balancing tick interval is 10 seconds, so
-Mantle will use a 5 second second timeout.
-
-This design allows Mantle to immediately return an error if anything
-RADOS-related goes wrong.  We use this implementation because we do not want to
-do a blocking OSD read from inside the global MDS lock. Doing so would bring
-down the MDS cluster if any of the OSDs are not responsive.
-
-Storing the balancers in RADOS is simplified by the use of an interpreted
-language for writing balancer code. If we used a language that needs to be
-compiled, like the C++ object classes in the OSD, we would need to ensure
-binary compatibility, which is complicated by different operating systems,
-distributions, and compilers.
-
-\subsubsection{Logging, Debugging, and Warnings}
-
-In the original implementation, Mantle would log all errors, warnings, and
-debug messages to a log stored locally on each MDS server. To get the simplest
-status messages or to debug problems, the user would have to log into each MDS
-individually, look at the logs, and reason about causality and ordering.
-
-With Malacology, Mantle re-uses the centralized logging features of the
-monitoring service. Important errors, warnings, and info messages are collected
-by the monitoring subsystem and appear in the monitor cluster log so instead of
-users going to each node, they can watch messages appear at the monitor.
-Messages are logged sparingly, so as not to spam the monitor with frivolous
-debugging but important events, like balancer version changes or failed
-subsystems show up in the centralized log.
-
-\subsection{ZLog: A Fast Distributed Shared Log}
-\label{sec:zlog}
-
-The second service implemented on Malacology is ZLog, a high-performance
-distributed shared-log that is based on the CORFU
-protocol~\cite{balakrishnan_corfu_2012}. The shared-log is a
-powerful abstraction used to construct distributed systems, such as
-metadata management~\cite{balakrishnan:sosp13} and elastic database
-systems~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
-However, existing implementations that rely on consensus algorithms such as
-Paxos funnel I/O through a single point introducing a bottleneck that restricts
-throughput. In contrast, the CORFU protocol is able to achieve high throughput
-using a network counter called a \emph{sequencer}, that
-decouples log position assignment from log I/O.
-
-While a full description of the CORFU system is beyond the scope of this
-paper, we briefly describe the custom storage device interface, sequencer
-service, and recovery protocol, and how these services are instantiated in the
-Malacology framework.
-
-\subsubsection{Sequencer}
-\label{sec:seq}
-
-High-performance in CORFU is achieved using a sequencer service that assigns
-log positions to clients by reading from a volatile, in-memory counter which
-can run at a very high throughput and at low latency. Since the sequencer is
-centralized, ensuring serializability in the common case is trivial.  The
-primary challenge in CORFU is handling the failure of the sequencer in a way
-that preserves correctness. Failure of the sequencer service in CORFU is
-handled by a recovery algorithm that recomputes the new sequencer state using a
-CORFU-specific custom storage interface to discover the tail of the log, while
-simultaneously invalidating stale client requests using an epoch-based
-protocol.
-
-{\bf Sequencer interface.} The sequencer resource supports the ability to
-\emph{read()} the current tail value and get the \emph{next()} position in
-the log which also atomically increments the tail position.
-We implement the sequencer service in Malacology as an application-specific
-file type that associates a small amount of metadata
-with the inode state managed by the Ceph metadata service.
-This approach has the added
-benefit of allowing the metadata service to handle naming, by
-representing each sequencer instance in the standard POSIX hierarchical
-namespace. The primary challenge in mapping the sequencer resource to the
-metadata service is handling serialization correctly to maintain the global
-ordering provided by the CORFU protocol.
-
-Initially we sought to directly model the sequencer service in Ceph as a
-non-exclusive, non-cacheable resource, forcing clients to perform a round-trip
-access to the resource at the authoritative metadata server for the sequencer
-inode.  Interestingly, we found that the capability system in Ceph uses a
-strategy to reduce metadata service load by allowing clients that open a
-shared file to temporarily obtain an exclusive cached copy of the resource,
-resulting in a round-robin, best-effort batching behavior. When a single
-client is accessing the sequencer resource it is able to increment the
-sequencer locally. Any competing client cannot query the sequencer until the
-metadata service has granted it access.
-
-While unexpected, this discovery allowed us to explore an implementation
-strategy that we had not previously considered. In particular, for bursty
-clients, and clients that can tolerate increased latency, this mode of
-operation may allow a system to achieve much higher throughput than a system
-with a centralized sequencer service.  We utilize the programmability of the
-metadata service to define a new policy for handling capabilities that controls
-the amount of time that clients are able to cache the sequencer resource. This
-allows an administrator or application to control the trade-off between latency
-and throughput beyond the standard best-effort policy that is present in Ceph
-by default.
-
-In Section~\ref{sec:evaluation} we quantify the trade-offs of throughput and
-latency for an approach based on a round-robin batching mode, and compare this
-mode to one in which the metadata server mediates access to the sequencer state
-when it is being shared among multiple clients. Quantifying these trade-offs
-should provide administrators with guidelines for setting the tunables for
-different ``caching" modes of the sequencer.
-
-{\bf Balancing policies.} As opposed to the batching mode for controlling access to the sequencer
-resource, more predictable latency can be achieved by treating the sequencer
-inode as a shared non-cacheable resource, forcing clients to make a round-trip
-to the metadata service. However, the shared nature of the metadata service
-may prevent the sequencer from achieving maximum throughput. To provide a
-solution to this issue we have taken advantage of the programmability of the
-metadata service load balancing to construct a service-specific load balancing
-policy. As opposed to a balancing policy that strives for uniform
-load distribution, a ZLog-specific policy may utilize knowledge of inode types
-to migrate the sequencer service to provisioned hardware during periods of
-contention or high demand.
-
-\subsubsection{Storage Interface}
-
-The storage interface is a critical component in the CORFU protocol. Clients
-independently map log positions that they have obtained from the sequencer
-service (described in detail in the next section) onto storage devices, while
-storage devices provide an intelligent \emph{write}-once, random \emph{read}
-interface for accessing log entries. The key to correctness in CORFU lies with
-the enforcement of up-to-date epoch tags on client requests; requests tagged
-with out-of-date epoch values are rejected, and clients are expected to
-request a new tail from the sequencer after refreshing state from an auxiliary
-service.  This mechanism forms the basis for sequencer recovery.
-
-In order to repopulate the sequencer state (i.e. the cached, current tail of
-the log) during recovery of a sequencer, the maximum position in the log must
-be obtained. To do this, the storage interface exposes an additional
-\emph{seal} method that atomically installs a new epoch value and returns the
-maximum log position that has been written.
-
-Since the sequencer service does not resume until the recovery process has
-completed, there cannot be a race with clients appending to the log, and the
-immutability of the log allow reads to never block during a sequencer failure.
-Recovery of the sequencer process itself may be handled in many ways, such as
-leader election using an auxiliary service such as Paxos. In our
-implementation, the recovery is the same as (and is inherited from) the Ceph
-metadata service. Handling the failure of a client that holds the sequencer
-state is similar, although a timeout is used to determine when a client should
-be considered unavailable.
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-%We wanted to implement CORFU and we had two choices: build it from the ground
-%up or do it on top of Malacology. To show the feasability of building on
-%Malacology our evaluation focuses on the
-
-Our evaluation demonstrates the feasibility of building new service
-abstractions atop programmable storage, focusing on the performance of the
-internal abstractions exposed by Malacology and used to construct the Mantle
-and ZLog services. We also discuss latent capabilities we discovered in this
-process that let us navigate different trade-offs within the services
-themselves. First, we benchmark scenarios with high sequencer contention by
-examining the interfaces used to map ZLog onto Malacology; specifically, we
-describe the sequencer implementation and the propagation of object and data
-interfaces interfaces.  Next, we benchmark scenarios in which the storage
-system manages multiple logs by using Mantle to balance sequencers across a
-cluster.
-
-% should save this for future work, otherwise it undercuts the eval.
-%Detailed system-wide performance measurements and analysis is forthcoming for
-%both Mantle and ZLog, as we try to find the best metadata balancing policies
-%and physical design implementation trade-offs.
-
-Since this work focuses on the programmability of Malacology, the goal of this
-section is to show that the components and subsystems that support the
-Malacology interfaces provide reasonable relative performance, as well as to
-give examples of the flexibility that Malacology provides to programmers.
-
-%As we search for the best metadata balancing policies
-%and physical design implementation trade-offs,
-%we plan to rel
-
-\subsection{Mapping ZLog onto Malacology}
-\label{sec:mapping-zlog-onto-malacology}
-
-We evaluate Malacology by exploring one possible mapping of the ZLog
-implementation of CORFU onto Ceph in which we re-use (1) the metadata service
-to manage naming and synchronization of the sequencer resource by treating the
-resource as an inode, and (2) the monitoring sub-subsystem to distribute and
-install application-specific I/O interface required of the CORFU protocol. In
-Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the inode
-abstraction for implementing the sequencer resource enables load balancing policies
-to migrate the sequencer resource in heavy-load situation.
-
-\subsubsection{Sequencer Implementation}
-\label{sec:sequencer-implementation}
-
-We evaluate the feasibility of using the metadata service to implement a
-sequencer resource that is responsible for maintaining a total ordering of
-the log. Clients contact the sequencer to obtain the tail of the log and then
-independently initiate I/O, thus we measure both the throughput and latency of
-obtaining new tail positions which bounds client append performance.
-
-The sequencer is implemented as a custom file type in which the sequencer
-state (a 64-bit integer) is embedded in the inode of a file representing the
-sequencer. A total ordering of the log is imposed by the re-use of the
-capability service that can be used to grant exclusive access of inode state
-to clients. The metadata service is responsible for maintaining exclusivity
-and granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
-behavior of the system in which a best-effort policy is used.  The two colors
-represent points in time that the clients were able to access the resource.
-The best-effort policy shows a high degree of interleaving
-between clients but the system spends a large portion of time re-distributing
-the capability, reducing overall throughput.
-
-In order to control the performance of the system we implement a
-policy that (1) restricts the length of time that a client may maintain
-exclusive access and (2) limits the number of log positions that a client may
-generate without yielding to other clients waiting for access. The behavior of
-these two modes is illustrated in Figures~\ref{fig:capdelay-quota-behavior}
-(b) and (c), respectively.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/capdelay-quota-behavior-a.png}
-\includegraphics{figures/capdelay-quota-behavior-b.png}
-\includegraphics{figures/capdelay-quota-behavior-c.png}
-\caption{Each dot is an individual request, spread randomly along the \(y\)
-axis. The default behavior is unpredictable, ``delay" lets clients hold the
-lease longer, and ``quota" gives clients the lease for a number of operations.}
-\label{fig:capdelay-quota-behavior}
-\end{figure}
-
-Figure~\ref{fig:captp} demonstrates a configurable trade-off between
-throughput and latency. In the experiment two clients are run each with a
-fixed 0.25 second maximum reservation on the capability, and we vary the size
-of the log position quota running each configuration for two minutes. The
-total operations per second is the combined throughput of the two clients, and
-the average latency is number of microseconds required to obtain a new log
-position. With a small quota more time is spent exchanging exclusive access,
-while a large quota reservation allows clients to experience a much lower
-latency because they experience uncontended access for a longer period of time.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/tradeoff.png}
-\caption{Sequencer throughput by re-using various services.  The highest performance is
-achieved using a single client with exclusive, cacheable privilege. Round-robin
-sharing of the sequencer resource is affected by the amount of time the
-resource is held, with best-effort performing the worst.}
-\label{fig:captp}
-\end{figure}
-
-To get a better picture of latency, Figure~\ref{fig:capcdf} shows the CDF of
-latency for each client in all experiment configurations. At the 99th
-percentile clients accessed the sequencer is less than a millisecond. The CDF
-is cropped at the 99.999th percentile due to large outliers that we believe
-occur in instances in which the MDS is performing I/O while it is in the
-process of re-distributing the capability to another client.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/caps-delay-latency.png}
-\caption{
-The latency distribution shows that if the clients hold their capabilities longer,
-performance decreases. Changing the delay can help the sequencer and
-clients strike a balance between throughput and latency.}
-\label{fig:capcdf}
-\end{figure}
-
-\textbf{Malacology exposes} the internal capability management service and
-allows users to navigate latency and throughput trade-offs.  Other approaches to
-designing the sequencer service also exist, such as using a centralized service
-in which each access is a network round-trip. In contrast to the mechanism we
-explored which is appropriate for clients with bursty workloads, it may be
-easier to provide predictable performance using a centralized service, and we
-will be exploring in the future how this can be achieved using the capability
-system.
-
-\subsubsection{Interface Propagation}
-
-Domain-specific data interfaces (Section~\ref{sec:application-specifc-storage-stacks}) allow
-co-design between applications and the storage system. Malacology supports
-custom object interfaces in RADOS that require interface implementations to
-be installed on the storage devices in the system, supporting the evolution of
-interfaces through automatic system-wide versioning and installation through
-the service metadata interface (Section~\ref{service-metadata}). We evaluate
-the performance of installing a new interface version in the cluster,
-which is an important metric for applications that frequently evolve
-interfaces.
-
-We demonstrate the feasibility of utilizing the Ceph monitoring sub-system by
-evaluating the performance of installing and distributing interface updates.
-Figure~\ref{fig:propdelay} shows the CDF of the latency of interface updates.
-The interfaces are Lua scripts embedded in the cluster map and distributed
-using a peer-to-peer gossip protocol.  The latency is defined as the elapsed
-time following the Paxos proposal for an interface update until each OSD makes
-the update live (the cost of the Paxos proposal is configurable and is
-discussed below). The latency measurements were taken on the nodes running
-Ceph server daemons, and thus exclude the client round-trip cost. In each of
-the experiments 1000 interface updates were observed.
-
-Figure~\ref{fig:propdelay} shows the lower bound cost for updates in a large
-cluster. In the experiment labeled ``120 OSD (RAM)" a cluster of 120 OSDs (10
-OSDs x 12 servers) using an in-memory data store were deployed, showing a
-latency of less than 54 ms with a probability of 90\% and a worst case latency
-of 194 ms. These costs demonstrate the penalty of distributing the interface
-in a large cluster. In practice the costs include, in addition to cluster-wide
-propagation of interface updates, the network round-trip to the
-interface management service, the Paxos commit protocol itself, and other factors
-such as system load. By default Paxos proposals occur periodically with a 1
-second interval in order to accumulate updates. In a minimum, realistic quorum
-of 3 monitors using HDD-based storage, we were able to decrease this interval
-to an average of 222 ms.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=60mm,trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
-\caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
-interface commit.}
-\label{fig:propdelay}
-\end{figure}
-
-\subsection{Load Balancing ZLog Sequencers with Mantle}
-\label{sec:zlog-balancing}
-
-In practice, a storage system implementing CORFU will support a multiplicity of
-independent totally-ordered logs for each application.  For this scenario
-co-locating sequencers on the same physical node is not ideal but building a load
-balancer that can migrate the shared resource ({\it e.g.,}, the resource that
-mediates access to the tail of the log) is a time-consuming, non-trivial task.
-It requires building subsystems for migrating resources, monitoring the
-workloads, collecting metrics that describe the utilization on the physical
-nodes, partitioning resources, maintaining cache coherence, and managing
-multiple sequencers. The following experiments demonstrate the feasibility of
-using the mechanisms of the Malacology load balancing interface to inherit
-these features and to alleviate load from overloaded servers.
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-behaviors.png}
-\caption{CephFS/Mantle load balancing have better throughput than co-locating
-all sequencers on the same server.  Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify this improvement;
-Section~\ref{sec:feature-backoff} examines the migration at 0-60 seconds.
-}\label{fig:mantle-balancer-behaviors}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-performance.png}
-\includegraphics{figures/mantle-mode-performance.png}
-\caption{In (a) all CephFS balancing modes have the same performance; Mantle uses a
-balancer designed for sequencers. In (b) the best combination of mode and
-migration units can have up to a 2\(\times\) improvement.}\label{fig:mantle-balancer-performance}
-\end{figure}
-
-The experiments are run on a cluster with 10 nodes to store objects, one node
-to monitor the cluster, and 3 nodes that can accommodate sequencers. 
-Instead of measuring contention at the clients like
-Section~\ref{sec:sequencer-implementation}, these experiments measure
-contention at the sequencers by forcing clients to make round-trips for every
-request. We implement this using the shared resource interface that forces
-round-trips.  Because the sequencer's only function is to  hand out positions
-for the tail of the log, the workload is read-heavy.
-
-First, we show how the ZLog service can orchestrate multiple sequencers using
-the Malacology load balancing interface.
-Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput over time of
-different load balancers as they migrate 3 sequencers (with 4 clients) around
-the cluster; ``No Balancing" keeps all sequencers on one server, ``CephFS"
-migrates sequencers using the CephFS load balancers, and ``Mantle" uses a
-custom load balancer we wrote specifically for sequencers. The CephFS load
-balancers are hard-coded into Ceph while Mantle is an API for balancing load
-presented in~Mantle~\cite{sevilla:sc15-mantle} and re-implemented on top of
-Malacology in this paper.  The increased throughput for the CephFS and Mantle
-curves between 0 and 60 seconds are a result of migrating the sequencer(s) off
-overloaded servers.
-
-In addition to showing that migrating sequencers improves performance,
-Figure~\ref{fig:mantle-balancer-behaviors} also demonstrates features that we
-will explore in the rest of this section.
-Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify the differences in performance
-when the cluster stabilizes at time 100 seconds and
-Section~\ref{sec:feature-backoff} examines the slope and start time of the
-rebalancing phase between 0 and 60 seconds by comparing the aggressiveness of
-the balancers.
-
-\subsubsection{Feature: Balancing Modes}
-\label{sec:feature-balancing-modes}
-
-Next, we quantify the performance benefits shown in
-Figure~\ref{fig:mantle-balancer-behaviors}.  To understand why load balancers
-perform differently we need to explain the different balancing modes that the
-load balancer service uses and how they stress the subsystems that receive and
-forward client requests in different ways. In
-Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
-performance of the balancing mode that CephFS falls into {\it most of the time}.
-CephFS has 3 modes for balancing load: CPU mode, workload mode, and hybrid
-mode. All three have the same structure for making migration decisions but vary
-based on the metric used to calculate load. For this sequencer workload the 3
-different modes all have the same performance, shown in
-Figure~\ref{fig:mantle-balancer-performance} (a), because the load balancer falls
-into the same mode a majority of the time.  The high variation in performance
-for the CephFS CPU Mode bar reflects the uncertainty of using something as
-dynamic and unpredictable as CPU utilization to make migration decisions. In
-addition to the suboptimal performance and unpredictability, because
-CephFS balancers all fall into the same mode despite using different balancers
-is a problem because it prevents administrators from properly exploring the
-balancing state space.
-
-Mantle gives the administrator more control over balancing policies; for
-the Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the load
-balancing interface to program logic for balancing read-heavy workloads,
-resulting in better throughput and stability.  When we did this we also
-identified two balancing modes relevant for making migration decisions
-for sequencers.
-
-Using Mantle, the administrator can put the load balancing service into ``proxy
-mode" or ``client mode". In proxy mode one server receives all requests and
-farms off the requests to slave servers; the slaves servers do the actual tail
-finding operation. In client mode, clients interact directly with the server
-that has their sequencer.  These modes are illustrated in
-Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
-co-located on one physical server -- performance for that mode is shown by the
-No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
-Mode" in the lower left shows how clients for sequencer 2 are re-directed to
-server B. The two ``Proxy mode" diagrams on the right show how clients continue
-sending requests to server A even though sequencer 2 has migrated to server B
-(the meaning of ``(Half)" is discussed in the next section).
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-modes.png}
-\caption{ In client mode clients sending requests to the server that houses
-their sequencer. In proxy mode clients continue sending their requests to the
-first server.  }\label{fig:mantle-modes}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-mode-behavior.png}
-\caption{The performance of proxy mode achieves the highest throughput but at
-the cost of lower throughput for one of the sequencers. Client mode is more
-fair but results in lower cluster throughput.
-}\label{fig:mantle-mode-behavior}
-\end{figure}
-
-Figure~\ref{fig:mantle-mode-behavior} shows the throughput over time of the two
-different modes for an environment with only 2 sequencers (again 4 clients
-each) and 2 servers. The curves for both sequencers in
-Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second and
-at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
-Performance of Sequencer 2 decreases because it stayed on the proxy which now
-handles all client requests, processes requests for Sequencer 2, and forwards
-requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
-because distributing the sequencers in this way separates (1) the handling of
-the client requests and (2) finding the tail of the log and responding to
-clients.  Doing both steps is too heavy weight for one server and sequencers on
-slave nodes can go faster if work is split up; this phenomenon is not
-uncommon and has been observed in chain replication~\cite{chain_rep}.
-
-Cluster
-throughput improves at the cost of decreased throughput for Sequencer 2.
-Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer mode manually (no
-balancing phase) and shows that the cluster throughput is worse than the
-cluster throughput of proxy mode. That graph also shows that Sequencer 2 has
-less throughput than Sequencer 2. In this case, the scatter-gather process used
-for cache coherence in the metadata protocols causes strain on the server
-housing Sequencer 2 resulting in this uneven performance.
-
-\subsubsection{Feature: Migration Units}
-\label{sec:feature-migration-units}
-
-Another factor that affects performance in this environment is how much load
-is on each server; these experiments quantify that effect by programming the
-load balancing interface to control the amount of load to migrate. We call this
-metric a ``migration unit".  Expressing this heuristic is not easily achievable
-with outward facing tunable parameters (i.e. system knobs) but with Mantle's
-programmable interface, we can force the load balancer to change its migration
-units. To force the balancer into the Proxy Mode (Half) scenario in
-Figure~\ref{fig:mantle-modes}, which uses migration units equal to half the
-load on the current server, we can use: \texttt{ targets[whoami+1] = mds[whoami]["load"]/2
-}.
-
-This code snippet uses globally defined variables and tables from the Mantle
-API to send half of the load on the current server (whoami) to the next ranked
-server (whoami + 1); the \texttt{targets} array is a globally defined table that
-the balancer uses to do the migrations. Alternatively, to migrate all load a
-time step, we can remove the division by 2.
-
-%\begin{figure}[t!]
-%\centering
-%\includegraphics{figures/mantle-mode-performance.png}
-%\caption{The performance of the different modes using various migration units
-%shows almost a 2\(\times\) improvement in the best case.
-%}\label{fig:mantle-mode-performance}
-%\end{figure}
-
-Figure~\ref{fig:mantle-balancer-performance} (b) shows the performance of the modes
-using different migration units. Recall that this setup only has 2 sequencers
-and 2 servers, so performance may be different at scale. Even so, it is clear
-that client mode does not perform as well for read-heavy workloads. We even see
-a throughput improvement when migrating all load off the first server, leaving
-the first server to do administrative tasks (this is common in the metadata
-cluster because the first server does a lot of the cache coherence work) while
-the second server does all the client work. Proxy mode does the best in both
-cases and shows large performance gains when completely decoupling client
-request handling and operation processing in Proxy Mode (Full).  The parameter
-that controls the migration units helps the administrator control the sequencer
-co-location or distribution across the cluster. This trade-off was explored
-extensively in the Mantle paper but the experiments we present here are
-indicative of an even richer set of states to explore.
-
-\subsubsection{Feature: Backoff}
-\label{sec:feature-backoff}
-
-Tuning the aggressiveness of the load balancer decision making is also a
-trade-off that administrators can control and explore. The balancing phase from
-0 to 60 seconds in Figure~\ref{fig:mantle-balancer-behaviors} shows different
-degrees of aggressiveness in making migration decisions; CephFS makes a
-decision quickly and throughput jumps to 2500 ops/second within 10 seconds
-while Mantle takes more time to stabilize. This conservative behavior is
-controlled by programming the balancer to (1) use different conditions for when
-to migrate and (2) using a threshold for sustained overload.
-
-We control the conditions for when to migrate using \texttt{when()}, a callback
-in the Mantle API.  For the Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} we program \texttt{when()} to wait
-for load on the receiving server to fall below a threshold. This makes the
-balancer more conservative because it takes 60 seconds for cache coherence
-messages to settle.  The Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} also takes longer to reach peak
-throughput because we want the policy to wait to see how migrations affect the
-system before proceeding; the balancer does a migration right before 50
-seconds, realizes that there is a third underloaded server, and does another
-migration.
-
-The other way to change aggressiveness of the decision making is to program
-into the balancer a threshold for sustained overload. This forces the balancer
-to wait a certain number of iterations after a migration before proceeding. In
-Mantle, the policy would use the save state function to do a countdown after a
-migration.  Behavior graphs and performance numbers for this backoff feature is
-omitted for space considerations, but our experiments confirm that the more
-conservative the approach the less overall throughput.
-
-\textbf{Malacology pulls} the load balancing service out of the storage system
-to balance sequencers across a cluster. This latent capability also gives
-future programmers the ability to explore the different load balancing
-trade-offs including: load balancing modes to control forwarding vs. client
-redirection, load migration units to control sequencer distribution vs.
-co-location, and backoffs to control conservative vs. aggressive decision
-making.
-
-%\subsubsection{Experiment: Object Class
-%Programmability}\label{experiment-object-class-programmability}
-%
-%Add quote from CORFU paper that talks about what they had to hack
-%
-%\subsubsection{Experiment: Multi-Client
-%Burstiness}\label{experiment-multi-client-burstiness}
-%
-%\begin{figure}[htbp] \centering
-%\includegraphics{figures/caps-delay-thruput.png} \caption{Forcing the client
-%to drop their capabilities later (delay) improves throughput} \end{figure}
-
-\section{Related Work}
-
-% IOStack: Software-Defined Object Storage
-% sRoute: Treating the Storage Stack Like a Network
-   % - lots of good refs to check out
-% Information and control in gray-box systems
-
-Programmability of operating systems and networking resources, including
-distributed storage systems is not new, but we are not aware of work that makes
-generalization of existing services into programmable resources a key principle
-in storage systems design.
-
-Programmable storage systems can be viewed as an infrastructure for creating
-abstractions to better separate policies from mechanisms. This idea is not new.
-Software-defined networks (SDNs) create such an abstraction by separating the
-control plane from the data plane (see for example~\cite{jain:sigcomm13}). This
-notion of control/data separation was also applied in software-defined storage
-(SDS)~\cite{thereska:sosp13,stefanovici:fast16}. Similarly,
-IOStack~\cite{gracia:internet16} is providing policy-based provisioning and
-filtering in OpenStack Swift. According to a SNIA white
-paper~\cite{carlson:snia2014}, the primary goal of SDS is to control and
-facilitate flexible and dynamic provisioning of storage resources of different
-kinds, including flash memory and disk drives, to create a virtualized mapping
-between common storage abstractions (e.g. files, objects, and blocks) and
-storage devices taking data service objectives in terms of protection,
-availability, performance, and security into account. A programmable storage
-system exposes internal abstractions so that end users (not necessarily
-operators) can create new services on top of the storage stack. Thus, our
-notion of programmable storage differs from ``software-defined storage'' (SDS)
-in terms of goals and scope, although definitions of SDS are still in flux.
-
-Another view of programmable storage systems is one of tailoring systems
-resources to applications~\cite{arpaci:sosp01}. Related efforts include the
-Exokernel~\cite{engler:sosp95}, SPIN~\cite{bershad:sosp95} and
-Vino~\cite{seltzer:osdi96} projects; the latter two addressed the ability of injecting
-code into the kernel to specialize resource management. Another approach is to
-pass hints between the different layers of the IO stack to bridge the semantic
-gap between applications and
-storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}).
-
-Malacology uses the same Active and Typed Storage module presented in
-DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File Manifolds
-can be implemented with small changes to the Malacology framework, namely
-asynchronous object calls and Lua stubs in the inode, respectively.
-
-\section{Conclusion and Future Work}
-\label{conclusion-and-future-work}
-
-Programmable storage is a viable method for eliminating duplication of complex
-error-prone software used as workarounds for storage system
-deficiencies. We propose that systems expose their services in a safe way
-allowing application developers to customize system behavior to meet their
-needs while not sacrificing correctness. To illustrate the benefits of this
-approach we presented Malacology, a programmable storage system that
-facilitates the construction of new services by re-purposing existing subsystem
-abstractions of the storage stack.
-
-Future work will focus on constructing more interfaces to support a wide
-variety of storage system services that can be configured on-the-fly in existing
-systems. This work is one point along that path to producing general-purpose
-storage systems that can
-target  special-purpose applications.  Ultimately we want to utilize
-declarative methods for expressing new services.
-
-\bibliography{paper}
-\bibliographystyle{plain}
-%\printbibliography
-
-\end{document}
+not necessarily effect correctness.} \addressesissue{3}

### 2017-02-23 paper/5_services_built_on_malacology.tex
paper: cleanup line breaks and break into separate files
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-02-09 paper/paper.tex
@@
-\documentclass[preprint]{sigplanconf-eurosys}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage{pifont}
-\usepackage{color}
-\usepackage{comment}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Uncomment for diff
-\includecomment{notes}
-\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
-\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
-\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
-
-% Uncomment for camera-ready
-%\excludecomment{notes}
-%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
-%\newcommand{\oldcomment}[1]{}
-%\newcommand{\addressesissue}[1]{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\cmark}{\ding{51}}%
-\usepackage[font=footnotesize,labelfont=bf]{caption}
-
-\setlength{\belowcaptionskip}{-10pt}
-
-\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
-\usepackage[normalem]{ulem}
-\usepackage{pstricks}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-\setkeys{Gin}{width=\maxwidth,keepaspectratio}
-
-\usepackage[unicode=true]{hyperref}
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, Peter Alvaro, Shel Finkelstein, Jeff LeFevre},
-            pdftitle={Malacology: A Programmable Storage System},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=blue,
-            linkcolor=black,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-
-\title{Malacology: A Programmable Storage System}
-%\authorinfo{Paper ID: 165}          {Total Pages: 12}
-\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre}
-           {\{msevilla, jayhawk, ivo, cmaltzahn\}@soe.ucsc.edu, \{palvaro, finkelstein, jlefevre\}@ucsc.edu}
-\date{}
-\begin{document}
-
-\maketitle
-
-\begin{notes}
-\textcolor{red}{
-\noindent Issues raised by reviewers:
-\begin{enumerate}
-  \item clarify what was implemented: diagram of interfaces and prose
-  specifying the difference between Malacology and prior work (reviewers B, C,
-  D, E)
-  \item describe what makes a programmable storage system (reviewers C, D)
-  \item organize motivation: group interfaces into categories (e.g., core
-  functionality, features, and performance optimizations) and describe other
-  applications we can build with Malacology (reviewers B, C)
-  \item clarify terminology: service vs. file system metadata, storage vs.
-  file system, dirty vs. clean slate approaches, soft-state caching (reviewers
-  B, C)
-  \item address security, access control, and the safety; specify how it
-  affects composability, performance, and multi-tenancy (reviewer D)
-\end{enumerate}
-}
-\end{notes}
-
-\begin{abstract} Storage systems are caught between the need to evolve data
-processing applications efficiently and quickly, and the increasing velocity
-with which storage device technology evolves. This puts tremendous pressure on
-storage systems to support rapid change both in terms of their interfaces and
-their performance. But adapting storage systems can be difficult because
-unprincipled changes might jeopardize years of code-hardening and performance
-optimization efforts that were necessary for users to entrust their data to the
-storage system. We introduce the concept of a programmable storage system, a
-system that exposes internal services and abstractions with the goal of
-re-using and building new services on top of the storage stack. We present
-Malacology, a prototype implementation of a programmable storage system to
-explore how existing abstractions of common storage system services can be
-leveraged to adapt to the needs of new data processing systems and the
-increasing variety of storage devices. We show how this approach allows a large
-degree of flexibility for storage systems to evolve without sacrificing the
-robustness of their code-hardened subsystems. We illustrate the advantages and
-challenges of programmability by composing existing primitives into two new
-higher-level services: a metadata load balancer and a high-performance
-distributed shared-log. The evaluation demonstrates that our services inherit
-desirable qualities of the back-end storage system, including the ability to
-balance load, efficiently propagate cluster metadata, recover from failure, and
-to navigate trade-offs between latency and throughput using leases.
-\end{abstract}
-
-\begin{figure}[tb]
-\centering
-\includegraphics{figures/overview.png}
-\caption{Scalable storage systems have storage daemons which store data,
-monitor daemons (M) that maintain cluster state, and service-specific daemons
-({\it e.g.}, file system metadata servers). Malacology enables the programmability of
-internal abstractions (bold arrows) to re-use and compose
-existing subsystems.  With Malacology, we add new services, ZLog and
-Mantle, that sit alongside traditional user-facing APIs (file, block,
-object).}\label{fig:overview}
-\end{figure}
-
-\section{Introduction}
-\label{introduction}
-\label{sec:intro}
-
-A storage system implements abstractions designed to persistently store data
-and must exhibit a high level of correctness to prevent data loss.  Storage
-systems have evolved around storage devices that often were orders of magnitude
-slower than CPU and memory, and therefore could dominate overall performance if
-not used carefully. Over the last few decades members of the storage systems
-community have developed clever strategies to meet correctness requirements
-while somewhat hiding the latency of traditional storage
-media~\cite{brewer_disks_2016}. To avoid lock-in by a particular vendor, users
-of storage systems have preferred systems with highly standardized APIs and
-lowest common denominator abstract data types such as blocks of bytes and byte
-stream files~\cite{armbrust_view_2010}.
-
-A number of recent developments have disrupted traditional storage systems.
-First, the falling prices of flash storage and the availability of new types of
-non-volatile memory that are orders of magnitude faster than traditional
-spinning media are moving overall performance bottlenecks away from storage
-devices to CPUs and networking, and pressure storage systems to shorten their
-code paths and incorporate new
-optimizations~\cite{gray_tape_2007,gray_flash_2008}.  Second, emerging ``big
-data'' applications demand interface evolution to support flexible consistency
-as well as flexible structured data 
-representations.~\cite{apache_contributors_parquet_2014}.  Finally, production-quality scalable
-storage systems available as open source software have established and are
-continuing to establish new, \emph{de-facto} API standards at a faster pace
-than traditional standards
-bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
-
-The evolutionary pressure placed on storage systems by these trends raises the
-question of whether there are principles that storage systems designers can
-follow to evolve storage systems efficiently, without jeopardizing years of
-code-hardening and performance optimization efforts.  In this paper we
-investigate an approach that focuses on identifying and exposing existing
-storage system resources, services, and abstractions that in a generalized form
-can be used to \emph{program} new services. By composing higher-level services
-over existing abstractions one can reuse subsystems and leverage their
-optimizations, established correctness, robustness, and efficiency.
-
-{\it \textbf{Contribution 1:}} We define a programmable storage system to be a
-storage system that facilitates the re-use and extension of existing storage
-abstractions provided by the underlying software stack, to enable the creation
-of new services via composition.  A programmable storage system can be realized
-by exposing existing functionality (such as metadata services and
-synchronization and monitoring capabilities) as interfaces that can be ``glued
-together'' in a variety of ways using a high-level language. Programmable
-storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection
-and execution of code within a storage system or storage device---in that the
-former is applicable to any component of the storage system, while the latter
-focuses at the data access level. Given this contrast, we can say that active
-storage is an example of how one internal component (the storage layer) is
-exposed in a programmable storage system.
-
-To illustrate the benefits and challenges of this approach we have designed and
-evaluated Malacology, a programmable storage system that facilitates the
-construction of new services by re-purposing existing subsystem abstractions of
-the storage stack.  We build Malacology in Ceph, a popular open source software
-storage stack.  We choose Ceph to demonstrate the concept of programmable
-storage because it offers a broad spectrum of existing services, including
-distributed locking and caching services provided by metadata servers,
-durability and object interfaces provided by the backend object store, and
-propagation of consistent cluster state provided by the monitoring service (see
-Figure~\ref{fig:overview}).  Malacology is expressive enough to provide the
-functionality necessary for implementing new services.
-
-Malacology includes a non-exhaustive set of interfaces that can be used as
-building blocks for constructing novel storage abstractions, including:
-
-\begin{enumerate}
-
-\item An interface for managing strongly-consistent time-varying \textbf{service
-metadata}.
-
-\item An interface for installing and evolving domain-specific, cluster-wide
-\textbf{data interfaces}.
-
-\item An interface for managing access to \textbf{shared resources} using a
-variety of optimization strategies.
-
-\item An interface for \textbf{load balancing} resources across the cluster.
-
-\item An interface for \textbf{durability} that persists policies using the
-underlying storage stack's object store.
-
-\end{enumerate}
-
-{\it \textbf{Contribution 2:}} We implement two distributed services using Malacology to
-demonstrate the feasibility of the programmable storage approach:
-
-\begin{enumerate}
-
-\item A high-performance distributed shared log service called ZLog, that is an
-implementation of CORFU.~\cite{balakrishnan_corfu_2012}
-
-\item An implementation of Mantle, the programmable load balancing
-service~\cite{sevilla:sc15-mantle}
-
-\end{enumerate}
-
-The remainder of this paper is structured as follows. First, we describe and
-motivate the need for programmable storage by describing current practices in
-the open source software community. Next we describe Malacology by presenting the
-subsystems within the underlying storage system that we re-purpose, and briefly
-describe how those system are used within Malacology (\S\ref{sec:malacology}).
-Then we describe the services that we have constructed within the Malacology
-framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
-implementation (\S\ref{sec:evaluation}).  We conclude by discussing related and future work.
-% and
-%conclude.
-
-\section{Application-Specific Storage Stacks}
-\label{sec:application-specifc-storage-stacks}
-
-Building storage stacks from the ground up for a specific purpose results in
-the best performance. For example, GFS~\cite{ghemawat:sosp03} and
-HDFS~\cite{shvachko:msst10} were designed specifically to serve Map/Reduce and
-Hadoop jobs, and use techniques like exposing data locality and relaxing POSIX
-constraints to achieve application-specific I/O optimizations. Another example
-is Boxwood~\cite{maccormick:osdi04}, which experimented with B-trees
-and chunk stores as storage abstractions to simplify application building.
-Alternatively, general-purpose storage stacks are built with the flexibility to
-serve many applications by providing a variety of interfaces and tunable
-parameters.  Unfortunately, managing competing forces in these systems is
-difficult and users want more control from the general-purpose storage stacks
-without going as far as building their storage system from the ground up.
-
-To demonstrate a recent trend towards more application-specific storage systems
-we examine the state of programmability in Ceph. Something of a storage Swiss
-army knife, Ceph simultaneously supports file, block, and object interfaces on
-a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
-Distributed Object Storage (RADOS) system is a cluster of object storage
-devices (OSDs) that provide Ceph with data durability and integrity using
-replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already
-provides some degree of programmability; the OSDs support domain-specific
-object interfaces that are implemented by composing existing low-level storage
-interfaces that execute atomically. These interfaces are written in C++ and are
-statically loaded into the system.
-
-The Ceph community provides empirical evidence that developers are already
-beginning to embrace programmable storage. Figure~\ref{fig:obj-int-dev-growth}
-shows a dramatic growth in the production use of domain-specific interfaces in
-the Ceph community since 2010. In that figure, classes are functional groupings
-of methods on storage objects (e.g. remotely computing and caching the checksum
-of an object extent).  What is most remarkable is that this trend contradicts
-the notion that API changes are a burden for users.  Rather it appears that
-gaps in existing interfaces are being addressed through ad-hoc approaches to
-programmability. In fact, Table~\ref{table:objclasses} categorizes existing
-interfaces and we clearly see a trend towards reusable services.
-
-\begin{figure}[ht]
-\centering
-\includegraphics{figures/obj-int-dev-growth.png}
-\caption{Since 2010, the growth in the number of co-designed object storage interfaces
-in Ceph has been accelerating. This plot is the number of object classes (a
-group of interfaces), and the total number of methods (the actual API
-end-points).}
-\label{fig:obj-int-dev-growth}
-\end{figure}
-
-\begin{table}[ht]
-\centering
-  \begin{tabular}{l|l|l}
-    Category & Example & \# \\ \hline
-    Logging  & Geographically distribute replicas & 11 \\ \hdashline
-    Metadata & Snapshots in the block device OR  & \multirow{2}{*}{74} \\
-    Management & Scan extents for file system repair & \\ \hdashline
-    Locking  & Grants clients exclusive access & 6 \\ \hdashline
-    Other & Garbage collection, reference counting  & 4\\
-\end{tabular}
-\caption{A variety of object storage classes exist to expose interfaces
-    to applications. \# is the number of methods that implement these categories.
-}
-\label{table:objclasses}
-\end{table}
-
-The takeaway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
-already trying to use programmability because their needs, whether they be
-related to performance, availability, consistency, convenience, etc., are not
-satisfied by the existing default set of interfaces. The popularity of the
-custom object interface facility of Ceph could be due to a number of reasons,
-such as the default algorithms/tunables of the storage system being
-insufficient for the application's performance goals, programmers wanting to
-exploit application-specific semantics, and/or programmers knowing how to
-manage resources to improve performance. A solution based on
-application-specific object interfaces is a way to work around the
-traditionally rigid storage APIs because custom object interfaces give
-programmers the ability to tell the storage system about their application: if
-the application is CPU or IO bound, if it has locality, if its size has the potential to
-overload a single node, etc.  Programmers often know what the problem is
-and how to solve it, but until the ability to modify object interfaces, they
-had no way to express to the storage system how to handle their data.
-
-Our approach is to expose more of the commonly used, code-hardened subsystems
-of the underlying storage system as interfaces. The intent is that these
-interfaces, which can be as simple as a redirection to the persistent data
-store or as complicated as a strongly consistent directory service, should be
-used and re-used in many contexts to implement a wide range of services. By
-making programmability a `feature', rather than a `hack' or `workaround', we
-help standardize a development process that now is largely ad-hoc.
-
-\section{Challenges}
-\label{sec:challenges}
-
-Implementing the infrastructure for programmability into existing services and
-abstractions of distributed storage systems is challenging, even if one assumes
-that the source code of the storage system and the necessary expertise for
-understanding it is available.  Some challenges include:
-
-\begin{itemize}
-
-\item Storage systems are generally required to be highly available so that any
-complete restarts of the storage system to reprogram them is usually
-unacceptable.
-
-\item Policies and optimizations are usually hard-wired into the services and
-one has to be careful when factoring them to avoid introducing additional bugs.
-These policies and optimizations are usually cross-cutting solutions to
-concerns or trade-offs that cannot be fully explored at the time the code is
-written (as they relate to workload or hardware). Given these policies and
-optimizations, decomposition of otherwise orthogonal internal abstractions can
-be difficult or dangerous.
-
-\item Mechanisms that are often only exercised according to hard-wired policies
-and not in their full generality have hidden bugs that are revealed as soon as
-those mechanisms are governed by different policies. In our experience
-introducing programmability into a storage system proved to be a great
-debugging tool.
-
-\item Programmability, especially in live systems, implies changes that need to
-be carefully managed by the system itself, including versioning and propagation
-of those changes without affecting correctness.
-
-\end{itemize}
-
-To address these challenges we present Malacology
-\oldcomment{. Malacology is both a prototype for a programmable storage system
-and a}
-\newcomment{, our prototype programmable storage system. It uses the
-programmable storage}
-\addressesissue{2}
-design approach to evolve storage systems efficiently and
-without jeopardizing years of code-hardening and performance optimization
-efforts.  Although Malacology uses the internal abstractions of the underlying
-storage system, including its subsystems, components, and implementations, we
-emphasize that our system still addresses the general challenges outlined
-above.
-
-The main challenge of designing a programmable storage system is choosing the
-right internal abstractions and picking the correct layers for exposing them.
-\oldcomment{In this paper, we do not present an exhaustive list of possible
-internal abstractions nor do we contend that the abstractions we choose provide
-the best trade-offs for all applications.} 
-\newcomment{A programmable storage is not defined by what abstractions are
-exposed, rather a programmable storage system adheres to the design approach of
-choosing interfaces so administrators can have better control of the storage
-system. The interfaces presented in this paper are abstractions that we found
-useful for building ZLog and Mantle, yet they may not provide the best
-trade-offs for all higher-level services.}
-\addressesissue{2}
-For example, if consensus is correctly exposed one could implement
-high-level features like versioning, serialization, or various flavors of
-strongly consistent data management on top; but perhaps a low-level consensus
-interface is suited well for a particular set of applications.  These questions
-are not answered in this paper and instead we focus on showing the feasibility
-of building such a system, given advances in the quality and robustness of
-today's storage stacks.
-
-The Malacology prototype we present has been implemented on Ceph.
-While there are other systems on top of which Malacology could be implemented (see Table~\ref{table:examples}), we choose Ceph because it is a production quality system and because it
-is open source. The large developer community ensures that code is
-robust and the visibility of the code lets us expose any interface we want. In
-the next section we describe the Ceph components that we expose as Malacology
-interfaces.
-
-\section{Malacology: A Programmable Storage System}
-\label{sec:malacology}
-
-\begin{table*}
-\centering\small
-\begin{tabular}{ l | c | l | l | l }
-%\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
-\multicolumn{4}{c}{} \\
-\textbf{Interface}                      &
-\textbf{Sec}                            &
-\textbf{Example in Production Systems}  &
-\textbf{Example in Ceph}                &
-\textbf{Provided Primitives}            \\ \hline
-Service Metadata
-  & \S\ref{sec:service-metadata-interface}
-  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
-  & cluster state management~\cite{website:ceph-mon}
-  & consensus/consistency
-  \\
-Data I/O
-  & \S\ref{sec:data-io-interface}
-  & Swift in situ storage/compute~\cite{website:zerocloud}
-  & object interface classes~\cite{website:cls-lua}
-  & transaction/atomicity
-  \\
-Shared Resource
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & MPI collective IO, burst
-  & POSIX metadata protocols
-  & serialization/batching
-  \\
-File Type
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & ?????
-  & file striping strategy
-  & data/metadata access
-  \\
-Load Balancing
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
-  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
-  & migration/sampling
-  \\
-Durability
-  & \S\ref{sec:durability}
-  & S3/Swift interfaces (RESTful API)
-  & object store library~\cite{weil_rados_2007}
-  & persistence/safety
-  \\
-\end{tabular}
-\caption{Common internal abstractions. The same ``internal abstractions" are
-common across large-scale systems because they provide primitives that solve
-general distributed systems problems.  Here we list examples of what these
-internal abstractions are used for in ``production systems" and in Ceph.
-Malacology provides these internal abstractions as interfaces
-\newcomment{(Section~\ref{sec:malacology})}
-that higher level
-\oldcomment{applications}
-\newcomment{services (Section~\ref{sec:services})}
-can use.
-\oldcomment{; {\it italicized abstractions} are interfaces discussed in this
-paper.}}
-\addressesissue{1}
-\label{table:examples}
-\end{table*}
-
-The guiding principle is to re-use existing services and extend them so that
-these services can be \emph{programmed}. We accomplish programmability of a
-service by exporting bindings (or ``hooks'') for an interpreted programming
-language so that programming can occur without having to restart the storage
-system (see also below,~\S\ref{sec:durability}). Table~\ref{table:examples}
-shows the internal services from Ceph that we expose in the Malacology
-prototype via Lua~\cite{ierusalimschy1996lua} bindings
-\oldcomment{.}
-\newcomment{and Figure~\ref{fig:implementation-overview} compares what was
-already present in Ceph (gray boxes) to the Malacology interfaces we added.
-Section~\ref{sec:services} will describe the higher-level services we built
-with these interfaces.}
-\addressesissue{1}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation-overview.png}
-\caption{\newcomment{Malacology is implemented on the daemons and clients that
-run in a Ceph cluster. Interfaces expose internal subsystems and are used as
-building blocks for higher-level services.}\addressesissue{1}
-\label{fig:implementation-overview}}
-\end{figure}
-
-Lua is a portable embedded scripting language and we choose it as the
-interpreted language for Malacology because it offers superior performance and
-productivity trade-offs, including a JIT-based implementation that is well
-known for near native performance. Additionally, Lua has been used extensively
-in game engines, and systems research \cite{neto:dls14-luaos}, including
-storage systems where it has been effectively used both on
-\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010}
-and off \cite{sevilla:sc15-mantle} the performance critical path. Finally, the
-flexibility of the runtime allows execution sandboxing in order to address
-security and performance concerns. We will now discuss the common subsystems
-used to manage storage system and how Malacology makes them programmable.
-
-\subsection{Service Metadata Interface}
-\label{sec:mon}
-\label{sec:service-metadata-interface}
-\label{service-metadata}
-
-% straw man example
-Keeping track of state in a distributed system is an essential part of any
-successful service and a necessary component in order to diagnose and detect
-failures, when they occur. This is further complicated by variable propagation
-delays and heterogeneous hardware in dynamic environments.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In the case of Ceph, a consistent view of cluster state among server daemons
-and clients is critical to provide strong consistency guarantees to clients.
-Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
-metadata server) data structures called ``maps'' that record membership and
-status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
-responsible for integrating state changes into cluster maps, responding to
-requests from out-of-date clients and synchronizing members of the cluster
-whenever there is a change in a map so that they all observe the same system
-state. As a fundamental building block of many system designs, consensus
-abstractions such as Paxos are a common technique for maintaining consistent
-data versions, and are a useful system to expose.
-
-The default behavior of the monitor can be seen as a Paxos-based notification
-system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing
-clients to identify when new values (termed epochs in Ceph) are associated to
-given maps.  Since Ceph does not expose this service directly, as part of our Malacology implementation, we expose a key-value
-service designed for managing configuration metadata that is built on top of
-the consensus engine. Since the monitor is intended to be out of the
-high-performance I/O path, a general guideline is to make use of this
-functionality infrequently and to assign small values to maps.\\
-
-\noindent{\it{\textbf{Malacology: }}} Malacology exposes a strongly-consistent view of
-time-varying service metadata as a service rather than a hidden internal
-component. Malacology provides a generic API for adding arbitrary values to
-existing subsystem cluster maps. As a consequence of this, applications can
-define simple but useful service-specific logic to the strongly-consistent
-interface, such as authorization control (just specific clients can write new
-values) or triggering actions based on specific values (e.g. sanitize values).
-The higher level services we implement in \S\ref{sec:services} make use of this
-functionality to register, version and propagate dynamic code (Lua scripts) for
-new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
-policies in the load balancer \S\ref{malacology:mds}.  Using this service
-guarantees that interface definitions are not only made durable, but are
-transparently and consistently propagated throughout the cluster so that
-clients are properly synchronized with the latest interfaces.
-
-\newcomment{\\ \noindent \it{\textbf{Impact: }} provides core functionality
-because it lets daemons come to consensus on system critical state. Bugs in the
-internal subsystems or ommitting this from services that need this type of
-consistency affects correctness.}
-\addressesissue{3, 5}
-
-%And finally, it is important to not underestimate the importance of the
-%preservation of the modifications that enable a new system service. In many
-%(if not all) cases, interfaces defining access to data are just as important
-%as the data itself by virtue of inherently providing structural context. Thus,
-%all components of a system service must be kept to the same standards of
-%protection as data in the system itself.
-
-\oldcomment{\noindent 4.2 Object I/O Intefrace}
-\subsection{Data I/O Interface}
-\label{sec:data-io-interface}
-\label{object-data-interface}
-Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
-supports application-specific object
-interfaces~\cite{weil_rados_2007}. The ability to offload computation can
-reduce data movement, and transactional interfaces can significantly simplify
-construction of complex storage interfaces that require uncoordinated parallel
-access.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-An object interface is a plugin structured similarly to that of an RPC
-in which a developer creates a named function within the cluster that clients
-may invoke. In the case of Ceph each function is implicitly run within the
-context of an object specified when the function is called. Developers of
-object interfaces express behavior by creating a composition of native
-interfaces or other custom object interfaces, and handle serialization of
-function input and output.  A wide range of native interfaces are available to
-developers such as reading and writing to a byte stream, controlling object
-snapshots and clones, and accessing a sorted key-value database. These native
-interfaces may be transactionally composed along with application
-specific logic to create semantically rich interfaces. An example would be
-an interface that atomically updates a matrix stored in the bytestream and an
-index of the matrix stored in the key-value database.\\
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation.png}
-\caption{Malacology allows users to dynamically define object/metadata
-interfaces by extending the object storage daemon (OSD) and metadata server
-(MDS) subsystems with an embedded Lua VM.  It uses the service metadata
-interface (M) to propagate interfaces/versions across the cluster
-\label{fig:implementation}}
-\end{figure}
-
-\noindent {\it{\textbf{Malacology: }}} The implementation of Ceph's object abstraction,
-although powerful, does not readily support programmability. Supporting only
-C/C++ for object interface developers, Ceph requires distribution of compiled
-binaries for the correct architecture, adding a large barrier of entry for
-developers and system administrators. Second, having no way to dynamically
-unload modules, any changes require a full restart of a storage daemon which
-may have serious performance impacts due to loss of cached data. And finally,
-the security limitations of the framework limit the use of object interfaces to
-all but those with administrative level access and deep technical expertise.
-
-To address these concerns, Malacology takes advantage of Lua extensions
-contributed by the Ceph community. This allows new object interfaces to be
-dynamically loaded into the system and modified at runtime, resulting in a
-object storage API with economy of expression, which at the same time provides
-the full set of features of the base object class. New object interfaces that
-are expressed in thousands of lines of code can be implemented in approximately
-an order of magnitude less code~\cite{geambasu_comet_2010}. While the use of
-Lua does not prevent deployment of malicious code, certain types of coding
-mistakes can be handled gracefully, and access policies are used to limit access
-to trusted users~\cite{ierusalimschy1996lua}.
-
-\oldcomment{All of these modifications work in tandem to implement the desired behavior,
-and are typically co-designed together such that each depends on the other to
-behave as expected.}
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by pushing behavior to lower parts of the storage stack, thereby
-minimizing hops and distributing computation.}
-\addressesissue{3, 5}
-
-\subsection{Distributed Metadata Interfaces}
-\begin{notes}
-\textcolor{red}{This section has been re-organized}
-\end{notes}
-\label{sec:distributed-metadata-interfaces}
-\label{malacology:mds}
-
-\oldcomment{The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction~\cite{weil:sc2004-dyn-metadata}.}
-\newcomment{File systems provide clients with the familiar POSIX file abstraction.
-While this guarantees strong consistency it comes at the cost of scalability,
-increased complexity, and lower performance.}
-\addressesissue{1}
-In general, distributed file systems protect resources by providing
-hierarchical indexing and distributed locking services.
-
-%The metadata cluster moves units of the namespace called directory fragments
-%and assigns them to servers using its own metadata load balancer. It then uses
-%a hard-coded policy to balance these fragments using a scalarized load metric
-%based on CPU, workload, and file system operation metrics. Although the CephFS
-%balancer has been shown to be complicated, its complexity is justifiable.
-%Having the flexibility to choose what to move, where to move it, and how much
-%to move is very powerful. The problem is that it is difficult to pick a
-%one-size-fits-all policy, system administrators may want to partition load
-%based on SLAs, or applications are better equipped to make these decisions,
-%etc. -- but we choose instead to expose a balancing API at strategic points in
-%the balancing logic using Lua hooks. ~\cite{sevilla:sc15-mantle}.
-
-\oldcomment{\it{\textbf{Malacology:}} Interfaces are added in strategic spots in the
-metadata service for guarding shared resources, defining new file types,
-and load balancing.}
-
-\oldcomment{\noindent Shared resource interface.}
-\subsubsection{Shared Resource Interface}
-
-In general, file system metadata servers manages client sessions, allowing
-clients to obtain locks (e.g. file byte ranges), and capabilities (e.g. to
-cache file data). Clients and metadata servers use a cooperative protocol in
-which clients voluntarily release resources back to the metadata service in
-order to implement sharing policies.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In Ceph, the locking service implements a capability-based system that
-expresses what data and metadata clients are allowed to access as well as what
-state they may cache and modify locally.  While designed for the file
-abstraction, indexing, locking, and caching are all common services that are
-useful to a broad spectrum of applications.  Distributed applications that
-share centralized resources (e.g. a database or directory) face similar
-challenges which are often solved using application-specific sharding.
-
-\newcomment{\\ \noindent\it{\textbf{Malacology: }}}
-While the current policy for sharing access and voluntarily releasing resources
-is largely best-effort, Malacology supports generalized policies between
-metadata servers and clients that can be used to implement fairness or
-priority.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} provides core functionality
-to protect and provide exclusive access for any shared resource. May hurt
-performance if the resource in question does not require strong consistency.}
-\addressesissue{3}
-
-\oldcomment{\noindent File type interface.}
-\subsubsection{File Type Interface}
-The metadata service manages a POSIX file system hierarchy where files and
-directories are represented as inode data structures and expose a POSIX file
-interface. Applications that manage large amounts of metadata (e.g. users or
-database snapshots) often require a naming service. We allow new inode types to
-be defined such that applications can create domain-specific interfaces to
-inodes that may modify locking and capability policies. We will show how this
-is used in Section~\ref{sec:seq} when we discuss a distributed shared-log built
-on Malacology.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this interface is both a feature
-and a performance optimization. It is a feature because it allows developers to
-add support for different  storage types, such as how to read new file formats
-or what consistency semantics to use for a specific subtree in the hierarchical
-namespace. It is also a performance optimization because future programmers can
-add optimizations for processing specific types of files into the inode itself.}
-\addressesissue{3}
-
-%%% TODO: striping strategy is stored with the inode
-
-\oldcomment{\noindent \bf Load balancing interface.} 
-\subsubsection{Load Balancing Interface}
-
-Ceph also addresses the challenge of balancing metadata load with a separate
-metadata cluster. This cluster uses load balancing policies to migrate
-directory inodes around the cluster to alleviate overload at single
-nodes~\cite{weil:sc2004-dyn-metadata}. The policies use metrics based on system
-state (e.g.  CPU and memory utilization) and statistics collected by the
-cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
-partitioning to move variable sized namespace subtrees. These units can be
-shipped anywhere (i.e., to any metadata server of any capacity) at any time for
-any reason. The original balancer was designed with hard-coded policies and
-tunables.\\
-
-\noindent {\it{\textbf{Malacology: }}}
-Mantle~\cite{sevilla:sc15-mantle}, a programmable metadata load balancer, has
-been implemented on Malacology so that its infrastructure can enjoy all the
-properties of existing internal abstractions.  The load balancing logic is
-intact but the infrastructure has been improved to safely load, manage, and
-persist its load balancing policies.  Composing Mantle with the service
-metadata and durability interfaces, Mantle can safely version balancer
-policies, save balancer policies in the backend object store and centralize
-warnings/errors. When combined with file types, load balancing policies can
-express a variety of rules such as multi-tenant and multi-application
-balancing.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by allowing them to specify how to partition, replicate, and
-distribute metadata in response to overloaded servers.}
-\addressesissue{3}
-
-\subsection{Durability Interface}
-\label{sec:durability}
-
-Ceph provides storage by striping and replicating data across
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-protects data using common techniques such as erasure
-coding, replication, and scrubbing.  Furthermore, many of these features
-are implemented using a peer-to-peer protocol that allows object storage
-devices to operate autonomously without a centralized coordinator.  For
-example, when the number of placement groups change, the OSDs re-balance and
-re-shard data in the background in a process called placement group splitting
-in which OSDs communicate directly with each other to cover a new state.
-In order to reduce load on the monitoring service, Ceph OSDs use a gossip
-protocol to efficiently propagate changes to cluster maps throughout the
-system, and autonomously initiate recovery mechanisms when failures are
-discovered.\\
-
-\noindent {\it{\textbf{Malacology: }}} Metadata service policies and object storage
-interfaces are stored durability within RADOS and managed by storing references
-with the object maps. Since the cluster already propagates a consistent view of
-these data structures, we use this service to automatically install interfaces
-in OSDs, and install policies within the metadata server daemon (MDS) such that clients and daemons are
-synchronized on correct implementations without restarting.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this is a feature because it
-adds data safety and persistence to system metadata; while nice to have it does
-effect correctness.}
-\addressesissue{3}
-
 \section{Services Built on Malacology}
@@
 is Mantle, a framework for dynamically specifying metadata load balancing
-policies. The second system, ZLog, is a high-performance distributed shared-log.
-In addition to these services, we'll demonstrate how we combine ZLog and Mantle
-to implement service-aware metadata load balancing policies.
+policies. The second system, ZLog, is a high-performance distributed
+shared-log.  In addition to these services, we'll demonstrate how we combine
+ZLog and Mantle to implement service-aware metadata load balancing policies.
 
@@
 it was trivial to make the versions inconsistent. Maintaining consistent
-versions is important for balancing policies that are cooperative in which local
-decisions are made assuming properties about other instances of the same
+versions is important for balancing policies that are cooperative in which
+local decisions are made assuming properties about other instances of the same
 policy.
@@
 distributed shared-log that is based on the CORFU
-protocol~\cite{balakrishnan_corfu_2012}. The shared-log is a
-powerful abstraction used to construct distributed systems, such as
-metadata management~\cite{balakrishnan:sosp13} and elastic database
-systems~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
-However, existing implementations that rely on consensus algorithms such as
-Paxos funnel I/O through a single point introducing a bottleneck that restricts
-throughput. In contrast, the CORFU protocol is able to achieve high throughput
-using a network counter called a \emph{sequencer}, that
-decouples log position assignment from log I/O.
+protocol~\cite{balakrishnan_corfu_2012}. The shared-log is a powerful
+abstraction used to construct distributed systems, such as metadata
+management~\cite{balakrishnan:sosp13} and elastic database
+systems~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.  However,
+existing implementations that rely on consensus algorithms such as Paxos funnel
+I/O through a single point introducing a bottleneck that restricts throughput.
+In contrast, the CORFU protocol is able to achieve high throughput using a
+network counter called a \emph{sequencer}, that decouples log position
+assignment from log I/O.
 
-While a full description of the CORFU system is beyond the scope of this
-paper, we briefly describe the custom storage device interface, sequencer
-service, and recovery protocol, and how these services are instantiated in the
-Malacology framework.
+While a full description of the CORFU system is beyond the scope of this paper,
+we briefly describe the custom storage device interface, sequencer service, and
+recovery protocol, and how these services are instantiated in the Malacology
+framework.
 
@@
 {\bf Sequencer interface.} The sequencer resource supports the ability to
-\emph{read()} the current tail value and get the \emph{next()} position in
-the log which also atomically increments the tail position.
-We implement the sequencer service in Malacology as an application-specific
-file type that associates a small amount of metadata
-with the inode state managed by the Ceph metadata service.
-This approach has the added
-benefit of allowing the metadata service to handle naming, by
-representing each sequencer instance in the standard POSIX hierarchical
-namespace. The primary challenge in mapping the sequencer resource to the
-metadata service is handling serialization correctly to maintain the global
-ordering provided by the CORFU protocol.
+\emph{read()} the current tail value and get the \emph{next()} position in the
+log which also atomically increments the tail position.  We implement the
+sequencer service in Malacology as an application-specific file type that
+associates a small amount of metadata with the inode state managed by the Ceph
+metadata service.  This approach has the added benefit of allowing the metadata
+service to handle naming, by representing each sequencer instance in the
+standard POSIX hierarchical namespace. The primary challenge in mapping the
+sequencer resource to the metadata service is handling serialization correctly
+to maintain the global ordering provided by the CORFU protocol.
 
@@
 inode.  Interestingly, we found that the capability system in Ceph uses a
-strategy to reduce metadata service load by allowing clients that open a
-shared file to temporarily obtain an exclusive cached copy of the resource,
-resulting in a round-robin, best-effort batching behavior. When a single
-client is accessing the sequencer resource it is able to increment the
-sequencer locally. Any competing client cannot query the sequencer until the
-metadata service has granted it access.
+strategy to reduce metadata service load by allowing clients that open a shared
+file to temporarily obtain an exclusive cached copy of the resource, resulting
+in a round-robin, best-effort batching behavior. When a single client is
+accessing the sequencer resource it is able to increment the sequencer locally.
+Any competing client cannot query the sequencer until the metadata service has
+granted it access.
 
@@
 
-{\bf Balancing policies.} As opposed to the batching mode for controlling access to the sequencer
-resource, more predictable latency can be achieved by treating the sequencer
-inode as a shared non-cacheable resource, forcing clients to make a round-trip
-to the metadata service. However, the shared nature of the metadata service
-may prevent the sequencer from achieving maximum throughput. To provide a
-solution to this issue we have taken advantage of the programmability of the
-metadata service load balancing to construct a service-specific load balancing
-policy. As opposed to a balancing policy that strives for uniform
-load distribution, a ZLog-specific policy may utilize knowledge of inode types
-to migrate the sequencer service to provisioned hardware during periods of
-contention or high demand.
+{\bf Balancing policies.} As opposed to the batching mode for controlling
+access to the sequencer resource, more predictable latency can be achieved by
+treating the sequencer inode as a shared non-cacheable resource, forcing
+clients to make a round-trip to the metadata service. However, the shared
+nature of the metadata service may prevent the sequencer from achieving maximum
+throughput. To provide a solution to this issue we have taken advantage of the
+programmability of the metadata service load balancing to construct a
+service-specific load balancing policy. As opposed to a balancing policy that
+strives for uniform load distribution, a ZLog-specific policy may utilize
+knowledge of inode types to migrate the sequencer service to provisioned
+hardware during periods of contention or high demand.
 
@@
 the enforcement of up-to-date epoch tags on client requests; requests tagged
-with out-of-date epoch values are rejected, and clients are expected to
-request a new tail from the sequencer after refreshing state from an auxiliary
-service.  This mechanism forms the basis for sequencer recovery.
+with out-of-date epoch values are rejected, and clients are expected to request
+a new tail from the sequencer after refreshing state from an auxiliary service.
+This mechanism forms the basis for sequencer recovery.
 
@@
 be considered unavailable.
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-%We wanted to implement CORFU and we had two choices: build it from the ground
-%up or do it on top of Malacology. To show the feasability of building on
-%Malacology our evaluation focuses on the
-
-Our evaluation demonstrates the feasibility of building new service
-abstractions atop programmable storage, focusing on the performance of the
-internal abstractions exposed by Malacology and used to construct the Mantle
-and ZLog services. We also discuss latent capabilities we discovered in this
-process that let us navigate different trade-offs within the services
-themselves. First, we benchmark scenarios with high sequencer contention by
-examining the interfaces used to map ZLog onto Malacology; specifically, we
-describe the sequencer implementation and the propagation of object and data
-interfaces interfaces.  Next, we benchmark scenarios in which the storage
-system manages multiple logs by using Mantle to balance sequencers across a
-cluster.
-
-% should save this for future work, otherwise it undercuts the eval.
-%Detailed system-wide performance measurements and analysis is forthcoming for
-%both Mantle and ZLog, as we try to find the best metadata balancing policies
-%and physical design implementation trade-offs.
-
-Since this work focuses on the programmability of Malacology, the goal of this
-section is to show that the components and subsystems that support the
-Malacology interfaces provide reasonable relative performance, as well as to
-give examples of the flexibility that Malacology provides to programmers.
-
-%As we search for the best metadata balancing policies
-%and physical design implementation trade-offs,
-%we plan to rel
-
-\subsection{Mapping ZLog onto Malacology}
-\label{sec:mapping-zlog-onto-malacology}
-
-We evaluate Malacology by exploring one possible mapping of the ZLog
-implementation of CORFU onto Ceph in which we re-use (1) the metadata service
-to manage naming and synchronization of the sequencer resource by treating the
-resource as an inode, and (2) the monitoring sub-subsystem to distribute and
-install application-specific I/O interface required of the CORFU protocol. In
-Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the inode
-abstraction for implementing the sequencer resource enables load balancing policies
-to migrate the sequencer resource in heavy-load situation.
-
-\subsubsection{Sequencer Implementation}
-\label{sec:sequencer-implementation}
-
-We evaluate the feasibility of using the metadata service to implement a
-sequencer resource that is responsible for maintaining a total ordering of
-the log. Clients contact the sequencer to obtain the tail of the log and then
-independently initiate I/O, thus we measure both the throughput and latency of
-obtaining new tail positions which bounds client append performance.
-
-The sequencer is implemented as a custom file type in which the sequencer
-state (a 64-bit integer) is embedded in the inode of a file representing the
-sequencer. A total ordering of the log is imposed by the re-use of the
-capability service that can be used to grant exclusive access of inode state
-to clients. The metadata service is responsible for maintaining exclusivity
-and granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
-behavior of the system in which a best-effort policy is used.  The two colors
-represent points in time that the clients were able to access the resource.
-The best-effort policy shows a high degree of interleaving
-between clients but the system spends a large portion of time re-distributing
-the capability, reducing overall throughput.
-
-In order to control the performance of the system we implement a
-policy that (1) restricts the length of time that a client may maintain
-exclusive access and (2) limits the number of log positions that a client may
-generate without yielding to other clients waiting for access. The behavior of
-these two modes is illustrated in Figures~\ref{fig:capdelay-quota-behavior}
-(b) and (c), respectively.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/capdelay-quota-behavior-a.png}
-\includegraphics{figures/capdelay-quota-behavior-b.png}
-\includegraphics{figures/capdelay-quota-behavior-c.png}
-\caption{Each dot is an individual request, spread randomly along the \(y\)
-axis. The default behavior is unpredictable, ``delay" lets clients hold the
-lease longer, and ``quota" gives clients the lease for a number of operations.}
-\label{fig:capdelay-quota-behavior}
-\end{figure}
-
-Figure~\ref{fig:captp} demonstrates a configurable trade-off between
-throughput and latency. In the experiment two clients are run each with a
-fixed 0.25 second maximum reservation on the capability, and we vary the size
-of the log position quota running each configuration for two minutes. The
-total operations per second is the combined throughput of the two clients, and
-the average latency is number of microseconds required to obtain a new log
-position. With a small quota more time is spent exchanging exclusive access,
-while a large quota reservation allows clients to experience a much lower
-latency because they experience uncontended access for a longer period of time.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/tradeoff.png}
-\caption{Sequencer throughput by re-using various services.  The highest performance is
-achieved using a single client with exclusive, cacheable privilege. Round-robin
-sharing of the sequencer resource is affected by the amount of time the
-resource is held, with best-effort performing the worst.}
-\label{fig:captp}
-\end{figure}
-
-To get a better picture of latency, Figure~\ref{fig:capcdf} shows the CDF of
-latency for each client in all experiment configurations. At the 99th
-percentile clients accessed the sequencer is less than a millisecond. The CDF
-is cropped at the 99.999th percentile due to large outliers that we believe
-occur in instances in which the MDS is performing I/O while it is in the
-process of re-distributing the capability to another client.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/caps-delay-latency.png}
-\caption{
-The latency distribution shows that if the clients hold their capabilities longer,
-performance decreases. Changing the delay can help the sequencer and
-clients strike a balance between throughput and latency.}
-\label{fig:capcdf}
-\end{figure}
-
-\textbf{Malacology exposes} the internal capability management service and
-allows users to navigate latency and throughput trade-offs.  Other approaches to
-designing the sequencer service also exist, such as using a centralized service
-in which each access is a network round-trip. In contrast to the mechanism we
-explored which is appropriate for clients with bursty workloads, it may be
-easier to provide predictable performance using a centralized service, and we
-will be exploring in the future how this can be achieved using the capability
-system.
-
-\subsubsection{Interface Propagation}
-
-Domain-specific data interfaces (Section~\ref{sec:application-specifc-storage-stacks}) allow
-co-design between applications and the storage system. Malacology supports
-custom object interfaces in RADOS that require interface implementations to
-be installed on the storage devices in the system, supporting the evolution of
-interfaces through automatic system-wide versioning and installation through
-the service metadata interface (Section~\ref{service-metadata}). We evaluate
-the performance of installing a new interface version in the cluster,
-which is an important metric for applications that frequently evolve
-interfaces.
-
-We demonstrate the feasibility of utilizing the Ceph monitoring sub-system by
-evaluating the performance of installing and distributing interface updates.
-Figure~\ref{fig:propdelay} shows the CDF of the latency of interface updates.
-The interfaces are Lua scripts embedded in the cluster map and distributed
-using a peer-to-peer gossip protocol.  The latency is defined as the elapsed
-time following the Paxos proposal for an interface update until each OSD makes
-the update live (the cost of the Paxos proposal is configurable and is
-discussed below). The latency measurements were taken on the nodes running
-Ceph server daemons, and thus exclude the client round-trip cost. In each of
-the experiments 1000 interface updates were observed.
-
-Figure~\ref{fig:propdelay} shows the lower bound cost for updates in a large
-cluster. In the experiment labeled ``120 OSD (RAM)" a cluster of 120 OSDs (10
-OSDs x 12 servers) using an in-memory data store were deployed, showing a
-latency of less than 54 ms with a probability of 90\% and a worst case latency
-of 194 ms. These costs demonstrate the penalty of distributing the interface
-in a large cluster. In practice the costs include, in addition to cluster-wide
-propagation of interface updates, the network round-trip to the
-interface management service, the Paxos commit protocol itself, and other factors
-such as system load. By default Paxos proposals occur periodically with a 1
-second interval in order to accumulate updates. In a minimum, realistic quorum
-of 3 monitors using HDD-based storage, we were able to decrease this interval
-to an average of 222 ms.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=60mm,trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
-\caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
-interface commit.}
-\label{fig:propdelay}
-\end{figure}
-
-\subsection{Load Balancing ZLog Sequencers with Mantle}
-\label{sec:zlog-balancing}
-
-In practice, a storage system implementing CORFU will support a multiplicity of
-independent totally-ordered logs for each application.  For this scenario
-co-locating sequencers on the same physical node is not ideal but building a load
-balancer that can migrate the shared resource ({\it e.g.,}, the resource that
-mediates access to the tail of the log) is a time-consuming, non-trivial task.
-It requires building subsystems for migrating resources, monitoring the
-workloads, collecting metrics that describe the utilization on the physical
-nodes, partitioning resources, maintaining cache coherence, and managing
-multiple sequencers. The following experiments demonstrate the feasibility of
-using the mechanisms of the Malacology load balancing interface to inherit
-these features and to alleviate load from overloaded servers.
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-behaviors.png}
-\caption{CephFS/Mantle load balancing have better throughput than co-locating
-all sequencers on the same server.  Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify this improvement;
-Section~\ref{sec:feature-backoff} examines the migration at 0-60 seconds.
-}\label{fig:mantle-balancer-behaviors}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-performance.png}
-\includegraphics{figures/mantle-mode-performance.png}
-\caption{In (a) all CephFS balancing modes have the same performance; Mantle uses a
-balancer designed for sequencers. In (b) the best combination of mode and
-migration units can have up to a 2\(\times\) improvement.}\label{fig:mantle-balancer-performance}
-\end{figure}
-
-The experiments are run on a cluster with 10 nodes to store objects, one node
-to monitor the cluster, and 3 nodes that can accommodate sequencers. 
-Instead of measuring contention at the clients like
-Section~\ref{sec:sequencer-implementation}, these experiments measure
-contention at the sequencers by forcing clients to make round-trips for every
-request. We implement this using the shared resource interface that forces
-round-trips.  Because the sequencer's only function is to  hand out positions
-for the tail of the log, the workload is read-heavy.
-
-First, we show how the ZLog service can orchestrate multiple sequencers using
-the Malacology load balancing interface.
-Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput over time of
-different load balancers as they migrate 3 sequencers (with 4 clients) around
-the cluster; ``No Balancing" keeps all sequencers on one server, ``CephFS"
-migrates sequencers using the CephFS load balancers, and ``Mantle" uses a
-custom load balancer we wrote specifically for sequencers. The CephFS load
-balancers are hard-coded into Ceph while Mantle is an API for balancing load
-presented in~Mantle~\cite{sevilla:sc15-mantle} and re-implemented on top of
-Malacology in this paper.  The increased throughput for the CephFS and Mantle
-curves between 0 and 60 seconds are a result of migrating the sequencer(s) off
-overloaded servers.
-
-In addition to showing that migrating sequencers improves performance,
-Figure~\ref{fig:mantle-balancer-behaviors} also demonstrates features that we
-will explore in the rest of this section.
-Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify the differences in performance
-when the cluster stabilizes at time 100 seconds and
-Section~\ref{sec:feature-backoff} examines the slope and start time of the
-rebalancing phase between 0 and 60 seconds by comparing the aggressiveness of
-the balancers.
-
-\subsubsection{Feature: Balancing Modes}
-\label{sec:feature-balancing-modes}
-
-Next, we quantify the performance benefits shown in
-Figure~\ref{fig:mantle-balancer-behaviors}.  To understand why load balancers
-perform differently we need to explain the different balancing modes that the
-load balancer service uses and how they stress the subsystems that receive and
-forward client requests in different ways. In
-Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
-performance of the balancing mode that CephFS falls into {\it most of the time}.
-CephFS has 3 modes for balancing load: CPU mode, workload mode, and hybrid
-mode. All three have the same structure for making migration decisions but vary
-based on the metric used to calculate load. For this sequencer workload the 3
-different modes all have the same performance, shown in
-Figure~\ref{fig:mantle-balancer-performance} (a), because the load balancer falls
-into the same mode a majority of the time.  The high variation in performance
-for the CephFS CPU Mode bar reflects the uncertainty of using something as
-dynamic and unpredictable as CPU utilization to make migration decisions. In
-addition to the suboptimal performance and unpredictability, because
-CephFS balancers all fall into the same mode despite using different balancers
-is a problem because it prevents administrators from properly exploring the
-balancing state space.
-
-Mantle gives the administrator more control over balancing policies; for
-the Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the load
-balancing interface to program logic for balancing read-heavy workloads,
-resulting in better throughput and stability.  When we did this we also
-identified two balancing modes relevant for making migration decisions
-for sequencers.
-
-Using Mantle, the administrator can put the load balancing service into ``proxy
-mode" or ``client mode". In proxy mode one server receives all requests and
-farms off the requests to slave servers; the slaves servers do the actual tail
-finding operation. In client mode, clients interact directly with the server
-that has their sequencer.  These modes are illustrated in
-Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
-co-located on one physical server -- performance for that mode is shown by the
-No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
-Mode" in the lower left shows how clients for sequencer 2 are re-directed to
-server B. The two ``Proxy mode" diagrams on the right show how clients continue
-sending requests to server A even though sequencer 2 has migrated to server B
-(the meaning of ``(Half)" is discussed in the next section).
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-modes.png}
-\caption{ In client mode clients sending requests to the server that houses
-their sequencer. In proxy mode clients continue sending their requests to the
-first server.  }\label{fig:mantle-modes}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-mode-behavior.png}
-\caption{The performance of proxy mode achieves the highest throughput but at
-the cost of lower throughput for one of the sequencers. Client mode is more
-fair but results in lower cluster throughput.
-}\label{fig:mantle-mode-behavior}
-\end{figure}
-
-Figure~\ref{fig:mantle-mode-behavior} shows the throughput over time of the two
-different modes for an environment with only 2 sequencers (again 4 clients
-each) and 2 servers. The curves for both sequencers in
-Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second and
-at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
-Performance of Sequencer 2 decreases because it stayed on the proxy which now
-handles all client requests, processes requests for Sequencer 2, and forwards
-requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
-because distributing the sequencers in this way separates (1) the handling of
-the client requests and (2) finding the tail of the log and responding to
-clients.  Doing both steps is too heavy weight for one server and sequencers on
-slave nodes can go faster if work is split up; this phenomenon is not
-uncommon and has been observed in chain replication~\cite{chain_rep}.
-
-Cluster
-throughput improves at the cost of decreased throughput for Sequencer 2.
-Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer mode manually (no
-balancing phase) and shows that the cluster throughput is worse than the
-cluster throughput of proxy mode. That graph also shows that Sequencer 2 has
-less throughput than Sequencer 2. In this case, the scatter-gather process used
-for cache coherence in the metadata protocols causes strain on the server
-housing Sequencer 2 resulting in this uneven performance.
-
-\subsubsection{Feature: Migration Units}
-\label{sec:feature-migration-units}
-
-Another factor that affects performance in this environment is how much load
-is on each server; these experiments quantify that effect by programming the
-load balancing interface to control the amount of load to migrate. We call this
-metric a ``migration unit".  Expressing this heuristic is not easily achievable
-with outward facing tunable parameters (i.e. system knobs) but with Mantle's
-programmable interface, we can force the load balancer to change its migration
-units. To force the balancer into the Proxy Mode (Half) scenario in
-Figure~\ref{fig:mantle-modes}, which uses migration units equal to half the
-load on the current server, we can use: \texttt{ targets[whoami+1] = mds[whoami]["load"]/2
-}.
-
-This code snippet uses globally defined variables and tables from the Mantle
-API to send half of the load on the current server (whoami) to the next ranked
-server (whoami + 1); the \texttt{targets} array is a globally defined table that
-the balancer uses to do the migrations. Alternatively, to migrate all load a
-time step, we can remove the division by 2.
-
-%\begin{figure}[t!]
-%\centering
-%\includegraphics{figures/mantle-mode-performance.png}
-%\caption{The performance of the different modes using various migration units
-%shows almost a 2\(\times\) improvement in the best case.
-%}\label{fig:mantle-mode-performance}
-%\end{figure}
-
-Figure~\ref{fig:mantle-balancer-performance} (b) shows the performance of the modes
-using different migration units. Recall that this setup only has 2 sequencers
-and 2 servers, so performance may be different at scale. Even so, it is clear
-that client mode does not perform as well for read-heavy workloads. We even see
-a throughput improvement when migrating all load off the first server, leaving
-the first server to do administrative tasks (this is common in the metadata
-cluster because the first server does a lot of the cache coherence work) while
-the second server does all the client work. Proxy mode does the best in both
-cases and shows large performance gains when completely decoupling client
-request handling and operation processing in Proxy Mode (Full).  The parameter
-that controls the migration units helps the administrator control the sequencer
-co-location or distribution across the cluster. This trade-off was explored
-extensively in the Mantle paper but the experiments we present here are
-indicative of an even richer set of states to explore.
-
-\subsubsection{Feature: Backoff}
-\label{sec:feature-backoff}
-
-Tuning the aggressiveness of the load balancer decision making is also a
-trade-off that administrators can control and explore. The balancing phase from
-0 to 60 seconds in Figure~\ref{fig:mantle-balancer-behaviors} shows different
-degrees of aggressiveness in making migration decisions; CephFS makes a
-decision quickly and throughput jumps to 2500 ops/second within 10 seconds
-while Mantle takes more time to stabilize. This conservative behavior is
-controlled by programming the balancer to (1) use different conditions for when
-to migrate and (2) using a threshold for sustained overload.
-
-We control the conditions for when to migrate using \texttt{when()}, a callback
-in the Mantle API.  For the Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} we program \texttt{when()} to wait
-for load on the receiving server to fall below a threshold. This makes the
-balancer more conservative because it takes 60 seconds for cache coherence
-messages to settle.  The Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} also takes longer to reach peak
-throughput because we want the policy to wait to see how migrations affect the
-system before proceeding; the balancer does a migration right before 50
-seconds, realizes that there is a third underloaded server, and does another
-migration.
-
-The other way to change aggressiveness of the decision making is to program
-into the balancer a threshold for sustained overload. This forces the balancer
-to wait a certain number of iterations after a migration before proceeding. In
-Mantle, the policy would use the save state function to do a countdown after a
-migration.  Behavior graphs and performance numbers for this backoff feature is
-omitted for space considerations, but our experiments confirm that the more
-conservative the approach the less overall throughput.
-
-\textbf{Malacology pulls} the load balancing service out of the storage system
-to balance sequencers across a cluster. This latent capability also gives
-future programmers the ability to explore the different load balancing
-trade-offs including: load balancing modes to control forwarding vs. client
-redirection, load migration units to control sequencer distribution vs.
-co-location, and backoffs to control conservative vs. aggressive decision
-making.
-
-%\subsubsection{Experiment: Object Class
-%Programmability}\label{experiment-object-class-programmability}
-%
-%Add quote from CORFU paper that talks about what they had to hack
-%
-%\subsubsection{Experiment: Multi-Client
-%Burstiness}\label{experiment-multi-client-burstiness}
-%
-%\begin{figure}[htbp] \centering
-%\includegraphics{figures/caps-delay-thruput.png} \caption{Forcing the client
-%to drop their capabilities later (delay) improves throughput} \end{figure}
-
-\section{Related Work}
-
-% IOStack: Software-Defined Object Storage
-% sRoute: Treating the Storage Stack Like a Network
-   % - lots of good refs to check out
-% Information and control in gray-box systems
-
-Programmability of operating systems and networking resources, including
-distributed storage systems is not new, but we are not aware of work that makes
-generalization of existing services into programmable resources a key principle
-in storage systems design.
-
-Programmable storage systems can be viewed as an infrastructure for creating
-abstractions to better separate policies from mechanisms. This idea is not new.
-Software-defined networks (SDNs) create such an abstraction by separating the
-control plane from the data plane (see for example~\cite{jain:sigcomm13}). This
-notion of control/data separation was also applied in software-defined storage
-(SDS)~\cite{thereska:sosp13,stefanovici:fast16}. Similarly,
-IOStack~\cite{gracia:internet16} is providing policy-based provisioning and
-filtering in OpenStack Swift. According to a SNIA white
-paper~\cite{carlson:snia2014}, the primary goal of SDS is to control and
-facilitate flexible and dynamic provisioning of storage resources of different
-kinds, including flash memory and disk drives, to create a virtualized mapping
-between common storage abstractions (e.g. files, objects, and blocks) and
-storage devices taking data service objectives in terms of protection,
-availability, performance, and security into account. A programmable storage
-system exposes internal abstractions so that end users (not necessarily
-operators) can create new services on top of the storage stack. Thus, our
-notion of programmable storage differs from ``software-defined storage'' (SDS)
-in terms of goals and scope, although definitions of SDS are still in flux.
-
-Another view of programmable storage systems is one of tailoring systems
-resources to applications~\cite{arpaci:sosp01}. Related efforts include the
-Exokernel~\cite{engler:sosp95}, SPIN~\cite{bershad:sosp95} and
-Vino~\cite{seltzer:osdi96} projects; the latter two addressed the ability of injecting
-code into the kernel to specialize resource management. Another approach is to
-pass hints between the different layers of the IO stack to bridge the semantic
-gap between applications and
-storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}).
-
-Malacology uses the same Active and Typed Storage module presented in
-DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File Manifolds
-can be implemented with small changes to the Malacology framework, namely
-asynchronous object calls and Lua stubs in the inode, respectively.
-
-\section{Conclusion and Future Work}
-\label{conclusion-and-future-work}
-
-Programmable storage is a viable method for eliminating duplication of complex
-error-prone software used as workarounds for storage system
-deficiencies. We propose that systems expose their services in a safe way
-allowing application developers to customize system behavior to meet their
-needs while not sacrificing correctness. To illustrate the benefits of this
-approach we presented Malacology, a programmable storage system that
-facilitates the construction of new services by re-purposing existing subsystem
-abstractions of the storage stack.
-
-Future work will focus on constructing more interfaces to support a wide
-variety of storage system services that can be configured on-the-fly in existing
-systems. This work is one point along that path to producing general-purpose
-storage systems that can
-target  special-purpose applications.  Ultimately we want to utilize
-declarative methods for expressing new services.
-
-\bibliography{paper}
-\bibliographystyle{plain}
-%\printbibliography
-
-\end{document}

### 2017-02-23 paper/6_evaluation.tex
paper: cleanup line breaks and break into separate files
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-02-09 paper/paper.tex
@@
-\documentclass[preprint]{sigplanconf-eurosys}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage{pifont}
-\usepackage{color}
-\usepackage{comment}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Uncomment for diff
-\includecomment{notes}
-\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
-\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
-\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
-
-% Uncomment for camera-ready
-%\excludecomment{notes}
-%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
-%\newcommand{\oldcomment}[1]{}
-%\newcommand{\addressesissue}[1]{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\cmark}{\ding{51}}%
-\usepackage[font=footnotesize,labelfont=bf]{caption}
-
-\setlength{\belowcaptionskip}{-10pt}
-
-\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
-\usepackage[normalem]{ulem}
-\usepackage{pstricks}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-\setkeys{Gin}{width=\maxwidth,keepaspectratio}
-
-\usepackage[unicode=true]{hyperref}
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, Peter Alvaro, Shel Finkelstein, Jeff LeFevre},
-            pdftitle={Malacology: A Programmable Storage System},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=blue,
-            linkcolor=black,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-
-\title{Malacology: A Programmable Storage System}
-%\authorinfo{Paper ID: 165}          {Total Pages: 12}
-\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre}
-           {\{msevilla, jayhawk, ivo, cmaltzahn\}@soe.ucsc.edu, \{palvaro, finkelstein, jlefevre\}@ucsc.edu}
-\date{}
-\begin{document}
-
-\maketitle
-
-\begin{notes}
-\textcolor{red}{
-\noindent Issues raised by reviewers:
-\begin{enumerate}
-  \item clarify what was implemented: diagram of interfaces and prose
-  specifying the difference between Malacology and prior work (reviewers B, C,
-  D, E)
-  \item describe what makes a programmable storage system (reviewers C, D)
-  \item organize motivation: group interfaces into categories (e.g., core
-  functionality, features, and performance optimizations) and describe other
-  applications we can build with Malacology (reviewers B, C)
-  \item clarify terminology: service vs. file system metadata, storage vs.
-  file system, dirty vs. clean slate approaches, soft-state caching (reviewers
-  B, C)
-  \item address security, access control, and the safety; specify how it
-  affects composability, performance, and multi-tenancy (reviewer D)
-\end{enumerate}
-}
-\end{notes}
-
-\begin{abstract} Storage systems are caught between the need to evolve data
-processing applications efficiently and quickly, and the increasing velocity
-with which storage device technology evolves. This puts tremendous pressure on
-storage systems to support rapid change both in terms of their interfaces and
-their performance. But adapting storage systems can be difficult because
-unprincipled changes might jeopardize years of code-hardening and performance
-optimization efforts that were necessary for users to entrust their data to the
-storage system. We introduce the concept of a programmable storage system, a
-system that exposes internal services and abstractions with the goal of
-re-using and building new services on top of the storage stack. We present
-Malacology, a prototype implementation of a programmable storage system to
-explore how existing abstractions of common storage system services can be
-leveraged to adapt to the needs of new data processing systems and the
-increasing variety of storage devices. We show how this approach allows a large
-degree of flexibility for storage systems to evolve without sacrificing the
-robustness of their code-hardened subsystems. We illustrate the advantages and
-challenges of programmability by composing existing primitives into two new
-higher-level services: a metadata load balancer and a high-performance
-distributed shared-log. The evaluation demonstrates that our services inherit
-desirable qualities of the back-end storage system, including the ability to
-balance load, efficiently propagate cluster metadata, recover from failure, and
-to navigate trade-offs between latency and throughput using leases.
-\end{abstract}
-
-\begin{figure}[tb]
-\centering
-\includegraphics{figures/overview.png}
-\caption{Scalable storage systems have storage daemons which store data,
-monitor daemons (M) that maintain cluster state, and service-specific daemons
-({\it e.g.}, file system metadata servers). Malacology enables the programmability of
-internal abstractions (bold arrows) to re-use and compose
-existing subsystems.  With Malacology, we add new services, ZLog and
-Mantle, that sit alongside traditional user-facing APIs (file, block,
-object).}\label{fig:overview}
-\end{figure}
-
-\section{Introduction}
-\label{introduction}
-\label{sec:intro}
-
-A storage system implements abstractions designed to persistently store data
-and must exhibit a high level of correctness to prevent data loss.  Storage
-systems have evolved around storage devices that often were orders of magnitude
-slower than CPU and memory, and therefore could dominate overall performance if
-not used carefully. Over the last few decades members of the storage systems
-community have developed clever strategies to meet correctness requirements
-while somewhat hiding the latency of traditional storage
-media~\cite{brewer_disks_2016}. To avoid lock-in by a particular vendor, users
-of storage systems have preferred systems with highly standardized APIs and
-lowest common denominator abstract data types such as blocks of bytes and byte
-stream files~\cite{armbrust_view_2010}.
-
-A number of recent developments have disrupted traditional storage systems.
-First, the falling prices of flash storage and the availability of new types of
-non-volatile memory that are orders of magnitude faster than traditional
-spinning media are moving overall performance bottlenecks away from storage
-devices to CPUs and networking, and pressure storage systems to shorten their
-code paths and incorporate new
-optimizations~\cite{gray_tape_2007,gray_flash_2008}.  Second, emerging ``big
-data'' applications demand interface evolution to support flexible consistency
-as well as flexible structured data 
-representations.~\cite{apache_contributors_parquet_2014}.  Finally, production-quality scalable
-storage systems available as open source software have established and are
-continuing to establish new, \emph{de-facto} API standards at a faster pace
-than traditional standards
-bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
-
-The evolutionary pressure placed on storage systems by these trends raises the
-question of whether there are principles that storage systems designers can
-follow to evolve storage systems efficiently, without jeopardizing years of
-code-hardening and performance optimization efforts.  In this paper we
-investigate an approach that focuses on identifying and exposing existing
-storage system resources, services, and abstractions that in a generalized form
-can be used to \emph{program} new services. By composing higher-level services
-over existing abstractions one can reuse subsystems and leverage their
-optimizations, established correctness, robustness, and efficiency.
-
-{\it \textbf{Contribution 1:}} We define a programmable storage system to be a
-storage system that facilitates the re-use and extension of existing storage
-abstractions provided by the underlying software stack, to enable the creation
-of new services via composition.  A programmable storage system can be realized
-by exposing existing functionality (such as metadata services and
-synchronization and monitoring capabilities) as interfaces that can be ``glued
-together'' in a variety of ways using a high-level language. Programmable
-storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection
-and execution of code within a storage system or storage device---in that the
-former is applicable to any component of the storage system, while the latter
-focuses at the data access level. Given this contrast, we can say that active
-storage is an example of how one internal component (the storage layer) is
-exposed in a programmable storage system.
-
-To illustrate the benefits and challenges of this approach we have designed and
-evaluated Malacology, a programmable storage system that facilitates the
-construction of new services by re-purposing existing subsystem abstractions of
-the storage stack.  We build Malacology in Ceph, a popular open source software
-storage stack.  We choose Ceph to demonstrate the concept of programmable
-storage because it offers a broad spectrum of existing services, including
-distributed locking and caching services provided by metadata servers,
-durability and object interfaces provided by the backend object store, and
-propagation of consistent cluster state provided by the monitoring service (see
-Figure~\ref{fig:overview}).  Malacology is expressive enough to provide the
-functionality necessary for implementing new services.
-
-Malacology includes a non-exhaustive set of interfaces that can be used as
-building blocks for constructing novel storage abstractions, including:
-
-\begin{enumerate}
-
-\item An interface for managing strongly-consistent time-varying \textbf{service
-metadata}.
-
-\item An interface for installing and evolving domain-specific, cluster-wide
-\textbf{data interfaces}.
-
-\item An interface for managing access to \textbf{shared resources} using a
-variety of optimization strategies.
-
-\item An interface for \textbf{load balancing} resources across the cluster.
-
-\item An interface for \textbf{durability} that persists policies using the
-underlying storage stack's object store.
-
-\end{enumerate}
-
-{\it \textbf{Contribution 2:}} We implement two distributed services using Malacology to
-demonstrate the feasibility of the programmable storage approach:
-
-\begin{enumerate}
-
-\item A high-performance distributed shared log service called ZLog, that is an
-implementation of CORFU.~\cite{balakrishnan_corfu_2012}
-
-\item An implementation of Mantle, the programmable load balancing
-service~\cite{sevilla:sc15-mantle}
-
-\end{enumerate}
-
-The remainder of this paper is structured as follows. First, we describe and
-motivate the need for programmable storage by describing current practices in
-the open source software community. Next we describe Malacology by presenting the
-subsystems within the underlying storage system that we re-purpose, and briefly
-describe how those system are used within Malacology (\S\ref{sec:malacology}).
-Then we describe the services that we have constructed within the Malacology
-framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
-implementation (\S\ref{sec:evaluation}).  We conclude by discussing related and future work.
-% and
-%conclude.
-
-\section{Application-Specific Storage Stacks}
-\label{sec:application-specifc-storage-stacks}
-
-Building storage stacks from the ground up for a specific purpose results in
-the best performance. For example, GFS~\cite{ghemawat:sosp03} and
-HDFS~\cite{shvachko:msst10} were designed specifically to serve Map/Reduce and
-Hadoop jobs, and use techniques like exposing data locality and relaxing POSIX
-constraints to achieve application-specific I/O optimizations. Another example
-is Boxwood~\cite{maccormick:osdi04}, which experimented with B-trees
-and chunk stores as storage abstractions to simplify application building.
-Alternatively, general-purpose storage stacks are built with the flexibility to
-serve many applications by providing a variety of interfaces and tunable
-parameters.  Unfortunately, managing competing forces in these systems is
-difficult and users want more control from the general-purpose storage stacks
-without going as far as building their storage system from the ground up.
-
-To demonstrate a recent trend towards more application-specific storage systems
-we examine the state of programmability in Ceph. Something of a storage Swiss
-army knife, Ceph simultaneously supports file, block, and object interfaces on
-a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
-Distributed Object Storage (RADOS) system is a cluster of object storage
-devices (OSDs) that provide Ceph with data durability and integrity using
-replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already
-provides some degree of programmability; the OSDs support domain-specific
-object interfaces that are implemented by composing existing low-level storage
-interfaces that execute atomically. These interfaces are written in C++ and are
-statically loaded into the system.
-
-The Ceph community provides empirical evidence that developers are already
-beginning to embrace programmable storage. Figure~\ref{fig:obj-int-dev-growth}
-shows a dramatic growth in the production use of domain-specific interfaces in
-the Ceph community since 2010. In that figure, classes are functional groupings
-of methods on storage objects (e.g. remotely computing and caching the checksum
-of an object extent).  What is most remarkable is that this trend contradicts
-the notion that API changes are a burden for users.  Rather it appears that
-gaps in existing interfaces are being addressed through ad-hoc approaches to
-programmability. In fact, Table~\ref{table:objclasses} categorizes existing
-interfaces and we clearly see a trend towards reusable services.
-
-\begin{figure}[ht]
-\centering
-\includegraphics{figures/obj-int-dev-growth.png}
-\caption{Since 2010, the growth in the number of co-designed object storage interfaces
-in Ceph has been accelerating. This plot is the number of object classes (a
-group of interfaces), and the total number of methods (the actual API
-end-points).}
-\label{fig:obj-int-dev-growth}
-\end{figure}
-
-\begin{table}[ht]
-\centering
-  \begin{tabular}{l|l|l}
-    Category & Example & \# \\ \hline
-    Logging  & Geographically distribute replicas & 11 \\ \hdashline
-    Metadata & Snapshots in the block device OR  & \multirow{2}{*}{74} \\
-    Management & Scan extents for file system repair & \\ \hdashline
-    Locking  & Grants clients exclusive access & 6 \\ \hdashline
-    Other & Garbage collection, reference counting  & 4\\
-\end{tabular}
-\caption{A variety of object storage classes exist to expose interfaces
-    to applications. \# is the number of methods that implement these categories.
-}
-\label{table:objclasses}
-\end{table}
-
-The takeaway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
-already trying to use programmability because their needs, whether they be
-related to performance, availability, consistency, convenience, etc., are not
-satisfied by the existing default set of interfaces. The popularity of the
-custom object interface facility of Ceph could be due to a number of reasons,
-such as the default algorithms/tunables of the storage system being
-insufficient for the application's performance goals, programmers wanting to
-exploit application-specific semantics, and/or programmers knowing how to
-manage resources to improve performance. A solution based on
-application-specific object interfaces is a way to work around the
-traditionally rigid storage APIs because custom object interfaces give
-programmers the ability to tell the storage system about their application: if
-the application is CPU or IO bound, if it has locality, if its size has the potential to
-overload a single node, etc.  Programmers often know what the problem is
-and how to solve it, but until the ability to modify object interfaces, they
-had no way to express to the storage system how to handle their data.
-
-Our approach is to expose more of the commonly used, code-hardened subsystems
-of the underlying storage system as interfaces. The intent is that these
-interfaces, which can be as simple as a redirection to the persistent data
-store or as complicated as a strongly consistent directory service, should be
-used and re-used in many contexts to implement a wide range of services. By
-making programmability a `feature', rather than a `hack' or `workaround', we
-help standardize a development process that now is largely ad-hoc.
-
-\section{Challenges}
-\label{sec:challenges}
-
-Implementing the infrastructure for programmability into existing services and
-abstractions of distributed storage systems is challenging, even if one assumes
-that the source code of the storage system and the necessary expertise for
-understanding it is available.  Some challenges include:
-
-\begin{itemize}
-
-\item Storage systems are generally required to be highly available so that any
-complete restarts of the storage system to reprogram them is usually
-unacceptable.
-
-\item Policies and optimizations are usually hard-wired into the services and
-one has to be careful when factoring them to avoid introducing additional bugs.
-These policies and optimizations are usually cross-cutting solutions to
-concerns or trade-offs that cannot be fully explored at the time the code is
-written (as they relate to workload or hardware). Given these policies and
-optimizations, decomposition of otherwise orthogonal internal abstractions can
-be difficult or dangerous.
-
-\item Mechanisms that are often only exercised according to hard-wired policies
-and not in their full generality have hidden bugs that are revealed as soon as
-those mechanisms are governed by different policies. In our experience
-introducing programmability into a storage system proved to be a great
-debugging tool.
-
-\item Programmability, especially in live systems, implies changes that need to
-be carefully managed by the system itself, including versioning and propagation
-of those changes without affecting correctness.
-
-\end{itemize}
-
-To address these challenges we present Malacology
-\oldcomment{. Malacology is both a prototype for a programmable storage system
-and a}
-\newcomment{, our prototype programmable storage system. It uses the
-programmable storage}
-\addressesissue{2}
-design approach to evolve storage systems efficiently and
-without jeopardizing years of code-hardening and performance optimization
-efforts.  Although Malacology uses the internal abstractions of the underlying
-storage system, including its subsystems, components, and implementations, we
-emphasize that our system still addresses the general challenges outlined
-above.
-
-The main challenge of designing a programmable storage system is choosing the
-right internal abstractions and picking the correct layers for exposing them.
-\oldcomment{In this paper, we do not present an exhaustive list of possible
-internal abstractions nor do we contend that the abstractions we choose provide
-the best trade-offs for all applications.} 
-\newcomment{A programmable storage is not defined by what abstractions are
-exposed, rather a programmable storage system adheres to the design approach of
-choosing interfaces so administrators can have better control of the storage
-system. The interfaces presented in this paper are abstractions that we found
-useful for building ZLog and Mantle, yet they may not provide the best
-trade-offs for all higher-level services.}
-\addressesissue{2}
-For example, if consensus is correctly exposed one could implement
-high-level features like versioning, serialization, or various flavors of
-strongly consistent data management on top; but perhaps a low-level consensus
-interface is suited well for a particular set of applications.  These questions
-are not answered in this paper and instead we focus on showing the feasibility
-of building such a system, given advances in the quality and robustness of
-today's storage stacks.
-
-The Malacology prototype we present has been implemented on Ceph.
-While there are other systems on top of which Malacology could be implemented (see Table~\ref{table:examples}), we choose Ceph because it is a production quality system and because it
-is open source. The large developer community ensures that code is
-robust and the visibility of the code lets us expose any interface we want. In
-the next section we describe the Ceph components that we expose as Malacology
-interfaces.
-
-\section{Malacology: A Programmable Storage System}
-\label{sec:malacology}
-
-\begin{table*}
-\centering\small
-\begin{tabular}{ l | c | l | l | l }
-%\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
-\multicolumn{4}{c}{} \\
-\textbf{Interface}                      &
-\textbf{Sec}                            &
-\textbf{Example in Production Systems}  &
-\textbf{Example in Ceph}                &
-\textbf{Provided Primitives}            \\ \hline
-Service Metadata
-  & \S\ref{sec:service-metadata-interface}
-  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
-  & cluster state management~\cite{website:ceph-mon}
-  & consensus/consistency
-  \\
-Data I/O
-  & \S\ref{sec:data-io-interface}
-  & Swift in situ storage/compute~\cite{website:zerocloud}
-  & object interface classes~\cite{website:cls-lua}
-  & transaction/atomicity
-  \\
-Shared Resource
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & MPI collective IO, burst
-  & POSIX metadata protocols
-  & serialization/batching
-  \\
-File Type
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & ?????
-  & file striping strategy
-  & data/metadata access
-  \\
-Load Balancing
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
-  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
-  & migration/sampling
-  \\
-Durability
-  & \S\ref{sec:durability}
-  & S3/Swift interfaces (RESTful API)
-  & object store library~\cite{weil_rados_2007}
-  & persistence/safety
-  \\
-\end{tabular}
-\caption{Common internal abstractions. The same ``internal abstractions" are
-common across large-scale systems because they provide primitives that solve
-general distributed systems problems.  Here we list examples of what these
-internal abstractions are used for in ``production systems" and in Ceph.
-Malacology provides these internal abstractions as interfaces
-\newcomment{(Section~\ref{sec:malacology})}
-that higher level
-\oldcomment{applications}
-\newcomment{services (Section~\ref{sec:services})}
-can use.
-\oldcomment{; {\it italicized abstractions} are interfaces discussed in this
-paper.}}
-\addressesissue{1}
-\label{table:examples}
-\end{table*}
-
-The guiding principle is to re-use existing services and extend them so that
-these services can be \emph{programmed}. We accomplish programmability of a
-service by exporting bindings (or ``hooks'') for an interpreted programming
-language so that programming can occur without having to restart the storage
-system (see also below,~\S\ref{sec:durability}). Table~\ref{table:examples}
-shows the internal services from Ceph that we expose in the Malacology
-prototype via Lua~\cite{ierusalimschy1996lua} bindings
-\oldcomment{.}
-\newcomment{and Figure~\ref{fig:implementation-overview} compares what was
-already present in Ceph (gray boxes) to the Malacology interfaces we added.
-Section~\ref{sec:services} will describe the higher-level services we built
-with these interfaces.}
-\addressesissue{1}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation-overview.png}
-\caption{\newcomment{Malacology is implemented on the daemons and clients that
-run in a Ceph cluster. Interfaces expose internal subsystems and are used as
-building blocks for higher-level services.}\addressesissue{1}
-\label{fig:implementation-overview}}
-\end{figure}
-
-Lua is a portable embedded scripting language and we choose it as the
-interpreted language for Malacology because it offers superior performance and
-productivity trade-offs, including a JIT-based implementation that is well
-known for near native performance. Additionally, Lua has been used extensively
-in game engines, and systems research \cite{neto:dls14-luaos}, including
-storage systems where it has been effectively used both on
-\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010}
-and off \cite{sevilla:sc15-mantle} the performance critical path. Finally, the
-flexibility of the runtime allows execution sandboxing in order to address
-security and performance concerns. We will now discuss the common subsystems
-used to manage storage system and how Malacology makes them programmable.
-
-\subsection{Service Metadata Interface}
-\label{sec:mon}
-\label{sec:service-metadata-interface}
-\label{service-metadata}
-
-% straw man example
-Keeping track of state in a distributed system is an essential part of any
-successful service and a necessary component in order to diagnose and detect
-failures, when they occur. This is further complicated by variable propagation
-delays and heterogeneous hardware in dynamic environments.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In the case of Ceph, a consistent view of cluster state among server daemons
-and clients is critical to provide strong consistency guarantees to clients.
-Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
-metadata server) data structures called ``maps'' that record membership and
-status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
-responsible for integrating state changes into cluster maps, responding to
-requests from out-of-date clients and synchronizing members of the cluster
-whenever there is a change in a map so that they all observe the same system
-state. As a fundamental building block of many system designs, consensus
-abstractions such as Paxos are a common technique for maintaining consistent
-data versions, and are a useful system to expose.
-
-The default behavior of the monitor can be seen as a Paxos-based notification
-system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing
-clients to identify when new values (termed epochs in Ceph) are associated to
-given maps.  Since Ceph does not expose this service directly, as part of our Malacology implementation, we expose a key-value
-service designed for managing configuration metadata that is built on top of
-the consensus engine. Since the monitor is intended to be out of the
-high-performance I/O path, a general guideline is to make use of this
-functionality infrequently and to assign small values to maps.\\
-
-\noindent{\it{\textbf{Malacology: }}} Malacology exposes a strongly-consistent view of
-time-varying service metadata as a service rather than a hidden internal
-component. Malacology provides a generic API for adding arbitrary values to
-existing subsystem cluster maps. As a consequence of this, applications can
-define simple but useful service-specific logic to the strongly-consistent
-interface, such as authorization control (just specific clients can write new
-values) or triggering actions based on specific values (e.g. sanitize values).
-The higher level services we implement in \S\ref{sec:services} make use of this
-functionality to register, version and propagate dynamic code (Lua scripts) for
-new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
-policies in the load balancer \S\ref{malacology:mds}.  Using this service
-guarantees that interface definitions are not only made durable, but are
-transparently and consistently propagated throughout the cluster so that
-clients are properly synchronized with the latest interfaces.
-
-\newcomment{\\ \noindent \it{\textbf{Impact: }} provides core functionality
-because it lets daemons come to consensus on system critical state. Bugs in the
-internal subsystems or ommitting this from services that need this type of
-consistency affects correctness.}
-\addressesissue{3, 5}
-
-%And finally, it is important to not underestimate the importance of the
-%preservation of the modifications that enable a new system service. In many
-%(if not all) cases, interfaces defining access to data are just as important
-%as the data itself by virtue of inherently providing structural context. Thus,
-%all components of a system service must be kept to the same standards of
-%protection as data in the system itself.
-
-\oldcomment{\noindent 4.2 Object I/O Intefrace}
-\subsection{Data I/O Interface}
-\label{sec:data-io-interface}
-\label{object-data-interface}
-Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
-supports application-specific object
-interfaces~\cite{weil_rados_2007}. The ability to offload computation can
-reduce data movement, and transactional interfaces can significantly simplify
-construction of complex storage interfaces that require uncoordinated parallel
-access.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-An object interface is a plugin structured similarly to that of an RPC
-in which a developer creates a named function within the cluster that clients
-may invoke. In the case of Ceph each function is implicitly run within the
-context of an object specified when the function is called. Developers of
-object interfaces express behavior by creating a composition of native
-interfaces or other custom object interfaces, and handle serialization of
-function input and output.  A wide range of native interfaces are available to
-developers such as reading and writing to a byte stream, controlling object
-snapshots and clones, and accessing a sorted key-value database. These native
-interfaces may be transactionally composed along with application
-specific logic to create semantically rich interfaces. An example would be
-an interface that atomically updates a matrix stored in the bytestream and an
-index of the matrix stored in the key-value database.\\
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation.png}
-\caption{Malacology allows users to dynamically define object/metadata
-interfaces by extending the object storage daemon (OSD) and metadata server
-(MDS) subsystems with an embedded Lua VM.  It uses the service metadata
-interface (M) to propagate interfaces/versions across the cluster
-\label{fig:implementation}}
-\end{figure}
-
-\noindent {\it{\textbf{Malacology: }}} The implementation of Ceph's object abstraction,
-although powerful, does not readily support programmability. Supporting only
-C/C++ for object interface developers, Ceph requires distribution of compiled
-binaries for the correct architecture, adding a large barrier of entry for
-developers and system administrators. Second, having no way to dynamically
-unload modules, any changes require a full restart of a storage daemon which
-may have serious performance impacts due to loss of cached data. And finally,
-the security limitations of the framework limit the use of object interfaces to
-all but those with administrative level access and deep technical expertise.
-
-To address these concerns, Malacology takes advantage of Lua extensions
-contributed by the Ceph community. This allows new object interfaces to be
-dynamically loaded into the system and modified at runtime, resulting in a
-object storage API with economy of expression, which at the same time provides
-the full set of features of the base object class. New object interfaces that
-are expressed in thousands of lines of code can be implemented in approximately
-an order of magnitude less code~\cite{geambasu_comet_2010}. While the use of
-Lua does not prevent deployment of malicious code, certain types of coding
-mistakes can be handled gracefully, and access policies are used to limit access
-to trusted users~\cite{ierusalimschy1996lua}.
-
-\oldcomment{All of these modifications work in tandem to implement the desired behavior,
-and are typically co-designed together such that each depends on the other to
-behave as expected.}
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by pushing behavior to lower parts of the storage stack, thereby
-minimizing hops and distributing computation.}
-\addressesissue{3, 5}
-
-\subsection{Distributed Metadata Interfaces}
-\begin{notes}
-\textcolor{red}{This section has been re-organized}
-\end{notes}
-\label{sec:distributed-metadata-interfaces}
-\label{malacology:mds}
-
-\oldcomment{The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction~\cite{weil:sc2004-dyn-metadata}.}
-\newcomment{File systems provide clients with the familiar POSIX file abstraction.
-While this guarantees strong consistency it comes at the cost of scalability,
-increased complexity, and lower performance.}
-\addressesissue{1}
-In general, distributed file systems protect resources by providing
-hierarchical indexing and distributed locking services.
-
-%The metadata cluster moves units of the namespace called directory fragments
-%and assigns them to servers using its own metadata load balancer. It then uses
-%a hard-coded policy to balance these fragments using a scalarized load metric
-%based on CPU, workload, and file system operation metrics. Although the CephFS
-%balancer has been shown to be complicated, its complexity is justifiable.
-%Having the flexibility to choose what to move, where to move it, and how much
-%to move is very powerful. The problem is that it is difficult to pick a
-%one-size-fits-all policy, system administrators may want to partition load
-%based on SLAs, or applications are better equipped to make these decisions,
-%etc. -- but we choose instead to expose a balancing API at strategic points in
-%the balancing logic using Lua hooks. ~\cite{sevilla:sc15-mantle}.
-
-\oldcomment{\it{\textbf{Malacology:}} Interfaces are added in strategic spots in the
-metadata service for guarding shared resources, defining new file types,
-and load balancing.}
-
-\oldcomment{\noindent Shared resource interface.}
-\subsubsection{Shared Resource Interface}
-
-In general, file system metadata servers manages client sessions, allowing
-clients to obtain locks (e.g. file byte ranges), and capabilities (e.g. to
-cache file data). Clients and metadata servers use a cooperative protocol in
-which clients voluntarily release resources back to the metadata service in
-order to implement sharing policies.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In Ceph, the locking service implements a capability-based system that
-expresses what data and metadata clients are allowed to access as well as what
-state they may cache and modify locally.  While designed for the file
-abstraction, indexing, locking, and caching are all common services that are
-useful to a broad spectrum of applications.  Distributed applications that
-share centralized resources (e.g. a database or directory) face similar
-challenges which are often solved using application-specific sharding.
-
-\newcomment{\\ \noindent\it{\textbf{Malacology: }}}
-While the current policy for sharing access and voluntarily releasing resources
-is largely best-effort, Malacology supports generalized policies between
-metadata servers and clients that can be used to implement fairness or
-priority.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} provides core functionality
-to protect and provide exclusive access for any shared resource. May hurt
-performance if the resource in question does not require strong consistency.}
-\addressesissue{3}
-
-\oldcomment{\noindent File type interface.}
-\subsubsection{File Type Interface}
-The metadata service manages a POSIX file system hierarchy where files and
-directories are represented as inode data structures and expose a POSIX file
-interface. Applications that manage large amounts of metadata (e.g. users or
-database snapshots) often require a naming service. We allow new inode types to
-be defined such that applications can create domain-specific interfaces to
-inodes that may modify locking and capability policies. We will show how this
-is used in Section~\ref{sec:seq} when we discuss a distributed shared-log built
-on Malacology.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this interface is both a feature
-and a performance optimization. It is a feature because it allows developers to
-add support for different  storage types, such as how to read new file formats
-or what consistency semantics to use for a specific subtree in the hierarchical
-namespace. It is also a performance optimization because future programmers can
-add optimizations for processing specific types of files into the inode itself.}
-\addressesissue{3}
-
-%%% TODO: striping strategy is stored with the inode
-
-\oldcomment{\noindent \bf Load balancing interface.} 
-\subsubsection{Load Balancing Interface}
-
-Ceph also addresses the challenge of balancing metadata load with a separate
-metadata cluster. This cluster uses load balancing policies to migrate
-directory inodes around the cluster to alleviate overload at single
-nodes~\cite{weil:sc2004-dyn-metadata}. The policies use metrics based on system
-state (e.g.  CPU and memory utilization) and statistics collected by the
-cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
-partitioning to move variable sized namespace subtrees. These units can be
-shipped anywhere (i.e., to any metadata server of any capacity) at any time for
-any reason. The original balancer was designed with hard-coded policies and
-tunables.\\
-
-\noindent {\it{\textbf{Malacology: }}}
-Mantle~\cite{sevilla:sc15-mantle}, a programmable metadata load balancer, has
-been implemented on Malacology so that its infrastructure can enjoy all the
-properties of existing internal abstractions.  The load balancing logic is
-intact but the infrastructure has been improved to safely load, manage, and
-persist its load balancing policies.  Composing Mantle with the service
-metadata and durability interfaces, Mantle can safely version balancer
-policies, save balancer policies in the backend object store and centralize
-warnings/errors. When combined with file types, load balancing policies can
-express a variety of rules such as multi-tenant and multi-application
-balancing.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by allowing them to specify how to partition, replicate, and
-distribute metadata in response to overloaded servers.}
-\addressesissue{3}
-
-\subsection{Durability Interface}
-\label{sec:durability}
-
-Ceph provides storage by striping and replicating data across
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-protects data using common techniques such as erasure
-coding, replication, and scrubbing.  Furthermore, many of these features
-are implemented using a peer-to-peer protocol that allows object storage
-devices to operate autonomously without a centralized coordinator.  For
-example, when the number of placement groups change, the OSDs re-balance and
-re-shard data in the background in a process called placement group splitting
-in which OSDs communicate directly with each other to cover a new state.
-In order to reduce load on the monitoring service, Ceph OSDs use a gossip
-protocol to efficiently propagate changes to cluster maps throughout the
-system, and autonomously initiate recovery mechanisms when failures are
-discovered.\\
-
-\noindent {\it{\textbf{Malacology: }}} Metadata service policies and object storage
-interfaces are stored durability within RADOS and managed by storing references
-with the object maps. Since the cluster already propagates a consistent view of
-these data structures, we use this service to automatically install interfaces
-in OSDs, and install policies within the metadata server daemon (MDS) such that clients and daemons are
-synchronized on correct implementations without restarting.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this is a feature because it
-adds data safety and persistence to system metadata; while nice to have it does
-effect correctness.}
-\addressesissue{3}
-
-\section{Services Built on Malacology}
-\label{sec:services}
-\label{services-built-on-malacology}
-
-\label{services}
-%\begin{table}
-%\centering\small
-%\begin{tabular}{  l | l | l }
-%\textbf{Service}              &
-%\textbf{Used to...}           &
-%\textbf{Malacology Interface} \\ \hline
-%Mantle & load balance metadata     & Load Balancing  \\
-%       & agree on current balancer & Service Metadata\\
-%       & control client write lock & Shared Resource \\
-%       & persist policy logic      & Durability      \\ \hline
-%ZLog   & implement CORFU           & Object \& Data  \\ 
-%       & install ZLog interfaces   & Service Metadata\\
-%       & persist log entries       & Durability      \\ \hline
-%ZLog   & load balance sequencers   & Load Balancing  \\
-%Seqr.  & manage access to log tail & Shared Resource \\ 
-%\end{tabular}
-%\caption{Mantle and ZLog are built using Malacology interfaces so they inherit the
-%robustness of the subsystems of the underlying stack.}
-%\label{table:implementation}
-%\end{table}
-
-In this section we describe two services built on top of Malacology. The first
-is Mantle, a framework for dynamically specifying metadata load balancing
-policies. The second system, ZLog, is a high-performance distributed shared-log.
-In addition to these services, we'll demonstrate how we combine ZLog and Mantle
-to implement service-aware metadata load balancing policies.
-
-\subsection{Mantle: Programmable Load Balancer}
-\label{sec:mantle}
-
-Mantle is a programmable load balancer that separates the metadata balancing
-policies from their mechanisms. Administrators inject code to change how the
-metadata cluster distributes metadata. In ~\cite{sevilla:sc15-mantle} the
-authors showed how using Mantle, one can implement a single node metadata
-service, a distributed metadata service with hashing, and a distributed
-metadata service with dynamic subtree partitioning.
-
-The original implementation was ``hard-coded" into Ceph and lacked robustness
-(no versioning, durability, or policy distribution).  Re-implemented using
-Malacology, Mantle now enjoys (1) the versioning provided by the metadata
-service interface provided by the monitor daemons and (2) the durability and
-distribution provided by Ceph's reliable object store.  Re-using the internal
-abstractions with Malacology resulted in a 2\(\times\) reduction in source code
-compared to the original implementation.
-
-\subsubsection{Versioning Balancer Policies}
-
-Ensuring that the version of the current load balancer is consistent across the
-physical servers in the metadata cluster was not addressed in the original
-implementation. The user had to set the version on each individual server and
-it was trivial to make the versions inconsistent. Maintaining consistent
-versions is important for balancing policies that are cooperative in which local
-decisions are made assuming properties about other instances of the same
-policy.
-
-With Malacology, Mantle stores the version of the current load balancer in the
-Ceph service metadata interface. The version of the load balancer corresponds
-to an object name in the balancing policy. Using the service metadata interface
-ensures that all metadata servers use the same version of the load balancer. As
-a result, the policy version inherits the consistency of Ceph's internal
-monitor daemons. The user changes the version of the load balancer using a new
-CLI command.
-
-\subsubsection{Making Balancer Policies Durable}
-
-The load balancer version described above corresponds to the name of an object
-in RADOS that holds the actual Lua balancing code.  When MDS nodes start
-balancing load, they first check the latest version from the MDS map and
-compare it to the balancer they have loaded. If the version has changed, they
-dereference the pointer to the balancer version by reading the corresponding
-object in RADOS. This is in contrast to the original balancer which stored load
-balancer code on the local file system -- a technique which is unreliable and
-may result in silent corruption.
-
-The balancer pulls the Lua code from RADOS synchronously; asynchronous reads
-are not possible because of the architecture of the MDS. The synchronous
-behavior is not the default behavior for RADOS operations, so we achieve this
-with a timeout: if the asynchronous read does not come back within half the
-balancing tick interval the operation is canceled and a Connection Timeout
-error is returned. By default, the balancing tick interval is 10 seconds, so
-Mantle will use a 5 second second timeout.
-
-This design allows Mantle to immediately return an error if anything
-RADOS-related goes wrong.  We use this implementation because we do not want to
-do a blocking OSD read from inside the global MDS lock. Doing so would bring
-down the MDS cluster if any of the OSDs are not responsive.
-
-Storing the balancers in RADOS is simplified by the use of an interpreted
-language for writing balancer code. If we used a language that needs to be
-compiled, like the C++ object classes in the OSD, we would need to ensure
-binary compatibility, which is complicated by different operating systems,
-distributions, and compilers.
-
-\subsubsection{Logging, Debugging, and Warnings}
-
-In the original implementation, Mantle would log all errors, warnings, and
-debug messages to a log stored locally on each MDS server. To get the simplest
-status messages or to debug problems, the user would have to log into each MDS
-individually, look at the logs, and reason about causality and ordering.
-
-With Malacology, Mantle re-uses the centralized logging features of the
-monitoring service. Important errors, warnings, and info messages are collected
-by the monitoring subsystem and appear in the monitor cluster log so instead of
-users going to each node, they can watch messages appear at the monitor.
-Messages are logged sparingly, so as not to spam the monitor with frivolous
-debugging but important events, like balancer version changes or failed
-subsystems show up in the centralized log.
-
-\subsection{ZLog: A Fast Distributed Shared Log}
-\label{sec:zlog}
-
-The second service implemented on Malacology is ZLog, a high-performance
-distributed shared-log that is based on the CORFU
-protocol~\cite{balakrishnan_corfu_2012}. The shared-log is a
-powerful abstraction used to construct distributed systems, such as
-metadata management~\cite{balakrishnan:sosp13} and elastic database
-systems~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
-However, existing implementations that rely on consensus algorithms such as
-Paxos funnel I/O through a single point introducing a bottleneck that restricts
-throughput. In contrast, the CORFU protocol is able to achieve high throughput
-using a network counter called a \emph{sequencer}, that
-decouples log position assignment from log I/O.
-
-While a full description of the CORFU system is beyond the scope of this
-paper, we briefly describe the custom storage device interface, sequencer
-service, and recovery protocol, and how these services are instantiated in the
-Malacology framework.
-
-\subsubsection{Sequencer}
-\label{sec:seq}
-
-High-performance in CORFU is achieved using a sequencer service that assigns
-log positions to clients by reading from a volatile, in-memory counter which
-can run at a very high throughput and at low latency. Since the sequencer is
-centralized, ensuring serializability in the common case is trivial.  The
-primary challenge in CORFU is handling the failure of the sequencer in a way
-that preserves correctness. Failure of the sequencer service in CORFU is
-handled by a recovery algorithm that recomputes the new sequencer state using a
-CORFU-specific custom storage interface to discover the tail of the log, while
-simultaneously invalidating stale client requests using an epoch-based
-protocol.
-
-{\bf Sequencer interface.} The sequencer resource supports the ability to
-\emph{read()} the current tail value and get the \emph{next()} position in
-the log which also atomically increments the tail position.
-We implement the sequencer service in Malacology as an application-specific
-file type that associates a small amount of metadata
-with the inode state managed by the Ceph metadata service.
-This approach has the added
-benefit of allowing the metadata service to handle naming, by
-representing each sequencer instance in the standard POSIX hierarchical
-namespace. The primary challenge in mapping the sequencer resource to the
-metadata service is handling serialization correctly to maintain the global
-ordering provided by the CORFU protocol.
-
-Initially we sought to directly model the sequencer service in Ceph as a
-non-exclusive, non-cacheable resource, forcing clients to perform a round-trip
-access to the resource at the authoritative metadata server for the sequencer
-inode.  Interestingly, we found that the capability system in Ceph uses a
-strategy to reduce metadata service load by allowing clients that open a
-shared file to temporarily obtain an exclusive cached copy of the resource,
-resulting in a round-robin, best-effort batching behavior. When a single
-client is accessing the sequencer resource it is able to increment the
-sequencer locally. Any competing client cannot query the sequencer until the
-metadata service has granted it access.
-
-While unexpected, this discovery allowed us to explore an implementation
-strategy that we had not previously considered. In particular, for bursty
-clients, and clients that can tolerate increased latency, this mode of
-operation may allow a system to achieve much higher throughput than a system
-with a centralized sequencer service.  We utilize the programmability of the
-metadata service to define a new policy for handling capabilities that controls
-the amount of time that clients are able to cache the sequencer resource. This
-allows an administrator or application to control the trade-off between latency
-and throughput beyond the standard best-effort policy that is present in Ceph
-by default.
-
-In Section~\ref{sec:evaluation} we quantify the trade-offs of throughput and
-latency for an approach based on a round-robin batching mode, and compare this
-mode to one in which the metadata server mediates access to the sequencer state
-when it is being shared among multiple clients. Quantifying these trade-offs
-should provide administrators with guidelines for setting the tunables for
-different ``caching" modes of the sequencer.
-
-{\bf Balancing policies.} As opposed to the batching mode for controlling access to the sequencer
-resource, more predictable latency can be achieved by treating the sequencer
-inode as a shared non-cacheable resource, forcing clients to make a round-trip
-to the metadata service. However, the shared nature of the metadata service
-may prevent the sequencer from achieving maximum throughput. To provide a
-solution to this issue we have taken advantage of the programmability of the
-metadata service load balancing to construct a service-specific load balancing
-policy. As opposed to a balancing policy that strives for uniform
-load distribution, a ZLog-specific policy may utilize knowledge of inode types
-to migrate the sequencer service to provisioned hardware during periods of
-contention or high demand.
-
-\subsubsection{Storage Interface}
-
-The storage interface is a critical component in the CORFU protocol. Clients
-independently map log positions that they have obtained from the sequencer
-service (described in detail in the next section) onto storage devices, while
-storage devices provide an intelligent \emph{write}-once, random \emph{read}
-interface for accessing log entries. The key to correctness in CORFU lies with
-the enforcement of up-to-date epoch tags on client requests; requests tagged
-with out-of-date epoch values are rejected, and clients are expected to
-request a new tail from the sequencer after refreshing state from an auxiliary
-service.  This mechanism forms the basis for sequencer recovery.
-
-In order to repopulate the sequencer state (i.e. the cached, current tail of
-the log) during recovery of a sequencer, the maximum position in the log must
-be obtained. To do this, the storage interface exposes an additional
-\emph{seal} method that atomically installs a new epoch value and returns the
-maximum log position that has been written.
-
-Since the sequencer service does not resume until the recovery process has
-completed, there cannot be a race with clients appending to the log, and the
-immutability of the log allow reads to never block during a sequencer failure.
-Recovery of the sequencer process itself may be handled in many ways, such as
-leader election using an auxiliary service such as Paxos. In our
-implementation, the recovery is the same as (and is inherited from) the Ceph
-metadata service. Handling the failure of a client that holds the sequencer
-state is similar, although a timeout is used to determine when a client should
-be considered unavailable.
-
 \section{Evaluation}
@@
 
-%As we search for the best metadata balancing policies
-%and physical design implementation trade-offs,
-%we plan to rel
+%As we search for the best metadata balancing policies and physical design
+%implementation trade-offs, we plan to rel
 
@@
 install application-specific I/O interface required of the CORFU protocol. In
-Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the inode
-abstraction for implementing the sequencer resource enables load balancing policies
-to migrate the sequencer resource in heavy-load situation.
+Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the
+inode abstraction for implementing the sequencer resource enables load
+balancing policies to migrate the sequencer resource in heavy-load situation.
 
@@
 We evaluate the feasibility of using the metadata service to implement a
-sequencer resource that is responsible for maintaining a total ordering of
-the log. Clients contact the sequencer to obtain the tail of the log and then
+sequencer resource that is responsible for maintaining a total ordering of the
+log. Clients contact the sequencer to obtain the tail of the log and then
 independently initiate I/O, thus we measure both the throughput and latency of
@@
 
-The sequencer is implemented as a custom file type in which the sequencer
-state (a 64-bit integer) is embedded in the inode of a file representing the
+The sequencer is implemented as a custom file type in which the sequencer state
+(a 64-bit integer) is embedded in the inode of a file representing the
 sequencer. A total ordering of the log is imposed by the re-use of the
-capability service that can be used to grant exclusive access of inode state
-to clients. The metadata service is responsible for maintaining exclusivity
-and granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
+capability service that can be used to grant exclusive access of inode state to
+clients. The metadata service is responsible for maintaining exclusivity and
+granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
 behavior of the system in which a best-effort policy is used.  The two colors
 represent points in time that the clients were able to access the resource.
-The best-effort policy shows a high degree of interleaving
-between clients but the system spends a large portion of time re-distributing
-the capability, reducing overall throughput.
+The best-effort policy shows a high degree of interleaving between clients but
+the system spends a large portion of time re-distributing the capability,
+reducing overall throughput.
 
-In order to control the performance of the system we implement a
-policy that (1) restricts the length of time that a client may maintain
-exclusive access and (2) limits the number of log positions that a client may
-generate without yielding to other clients waiting for access. The behavior of
-these two modes is illustrated in Figures~\ref{fig:capdelay-quota-behavior}
-(b) and (c), respectively.
+In order to control the performance of the system we implement a policy that
+(1) restricts the length of time that a client may maintain exclusive access
+and (2) limits the number of log positions that a client may generate without
+yielding to other clients waiting for access. The behavior of these two modes
+is illustrated in Figures~\ref{fig:capdelay-quota-behavior} (b) and (c),
+respectively.
 
@@
 
-Figure~\ref{fig:captp} demonstrates a configurable trade-off between
-throughput and latency. In the experiment two clients are run each with a
-fixed 0.25 second maximum reservation on the capability, and we vary the size
-of the log position quota running each configuration for two minutes. The
-total operations per second is the combined throughput of the two clients, and
-the average latency is number of microseconds required to obtain a new log
-position. With a small quota more time is spent exchanging exclusive access,
-while a large quota reservation allows clients to experience a much lower
-latency because they experience uncontended access for a longer period of time.
+Figure~\ref{fig:captp} demonstrates a configurable trade-off between throughput
+and latency. In the experiment two clients are run each with a fixed 0.25
+second maximum reservation on the capability, and we vary the size of the log
+position quota running each configuration for two minutes. The total operations
+per second is the combined throughput of the two clients, and the average
+latency is number of microseconds required to obtain a new log position. With a
+small quota more time is spent exchanging exclusive access, while a large quota
+reservation allows clients to experience a much lower latency because they
+experience uncontended access for a longer period of time.
 
@@
 \includegraphics{figures/tradeoff.png}
-\caption{Sequencer throughput by re-using various services.  The highest performance is
-achieved using a single client with exclusive, cacheable privilege. Round-robin
-sharing of the sequencer resource is affected by the amount of time the
-resource is held, with best-effort performing the worst.}
+\caption{Sequencer throughput by re-using various services.  The highest
+performance is achieved using a single client with exclusive, cacheable
+privilege. Round-robin sharing of the sequencer resource is affected by the
+amount of time the resource is held, with best-effort performing the worst.}
 \label{fig:captp}
@@
 \includegraphics{figures/caps-delay-latency.png}
-\caption{
-The latency distribution shows that if the clients hold their capabilities longer,
-performance decreases. Changing the delay can help the sequencer and
-clients strike a balance between throughput and latency.}
+\caption{ The latency distribution shows that if the clients hold their
+capabilities longer, performance decreases. Changing the delay can help the
+sequencer and clients strike a balance between throughput and latency.}
 \label{fig:capcdf}
@@
 \textbf{Malacology exposes} the internal capability management service and
-allows users to navigate latency and throughput trade-offs.  Other approaches to
-designing the sequencer service also exist, such as using a centralized service
-in which each access is a network round-trip. In contrast to the mechanism we
-explored which is appropriate for clients with bursty workloads, it may be
-easier to provide predictable performance using a centralized service, and we
-will be exploring in the future how this can be achieved using the capability
-system.
+allows users to navigate latency and throughput trade-offs.  Other approaches
+to designing the sequencer service also exist, such as using a centralized
+service in which each access is a network round-trip. In contrast to the
+mechanism we explored which is appropriate for clients with bursty workloads,
+it may be easier to provide predictable performance using a centralized
+service, and we will be exploring in the future how this can be achieved using
+the capability system.
 
@@
 
-Domain-specific data interfaces (Section~\ref{sec:application-specifc-storage-stacks}) allow
-co-design between applications and the storage system. Malacology supports
-custom object interfaces in RADOS that require interface implementations to
-be installed on the storage devices in the system, supporting the evolution of
-interfaces through automatic system-wide versioning and installation through
-the service metadata interface (Section~\ref{service-metadata}). We evaluate
-the performance of installing a new interface version in the cluster,
-which is an important metric for applications that frequently evolve
-interfaces.
+Domain-specific data interfaces
+(Section~\ref{sec:application-specifc-storage-stacks}) allow co-design between
+applications and the storage system. Malacology supports custom object
+interfaces in RADOS that require interface implementations to be installed on
+the storage devices in the system, supporting the evolution of interfaces
+through automatic system-wide versioning and installation through the service
+metadata interface (Section~\ref{service-metadata}). We evaluate the
+performance of installing a new interface version in the cluster, which is an
+important metric for applications that frequently evolve interfaces.
 
@@
 the update live (the cost of the Paxos proposal is configurable and is
-discussed below). The latency measurements were taken on the nodes running
-Ceph server daemons, and thus exclude the client round-trip cost. In each of
-the experiments 1000 interface updates were observed.
+discussed below). The latency measurements were taken on the nodes running Ceph
+server daemons, and thus exclude the client round-trip cost. In each of the
+experiments 1000 interface updates were observed.
 
@@
 latency of less than 54 ms with a probability of 90\% and a worst case latency
-of 194 ms. These costs demonstrate the penalty of distributing the interface
-in a large cluster. In practice the costs include, in addition to cluster-wide
-propagation of interface updates, the network round-trip to the
-interface management service, the Paxos commit protocol itself, and other factors
-such as system load. By default Paxos proposals occur periodically with a 1
-second interval in order to accumulate updates. In a minimum, realistic quorum
-of 3 monitors using HDD-based storage, we were able to decrease this interval
-to an average of 222 ms.
+of 194 ms. These costs demonstrate the penalty of distributing the interface in
+a large cluster. In practice the costs include, in addition to cluster-wide
+propagation of interface updates, the network round-trip to the interface
+management service, the Paxos commit protocol itself, and other factors such as
+system load. By default Paxos proposals occur periodically with a 1 second
+interval in order to accumulate updates. In a minimum, realistic quorum of 3
+monitors using HDD-based storage, we were able to decrease this interval to an
+average of 222 ms.
 
@@
 independent totally-ordered logs for each application.  For this scenario
-co-locating sequencers on the same physical node is not ideal but building a load
-balancer that can migrate the shared resource ({\it e.g.,}, the resource that
-mediates access to the tail of the log) is a time-consuming, non-trivial task.
-It requires building subsystems for migrating resources, monitoring the
+co-locating sequencers on the same physical node is not ideal but building a
+load balancer that can migrate the shared resource ({\it e.g.,}, the resource
+that mediates access to the tail of the log) is a time-consuming, non-trivial
+task.  It requires building subsystems for migrating resources, monitoring the
 workloads, collecting metrics that describe the utilization on the physical
@@
 \includegraphics{figures/mantle-mode-performance.png}
-\caption{In (a) all CephFS balancing modes have the same performance; Mantle uses a
-balancer designed for sequencers. In (b) the best combination of mode and
-migration units can have up to a 2\(\times\) improvement.}\label{fig:mantle-balancer-performance}
+\caption{In (a) all CephFS balancing modes have the same performance; Mantle
+uses a balancer designed for sequencers. In (b) the best combination of mode
+and migration units can have up to a 2\(\times\)
+improvement.}\label{fig:mantle-balancer-performance}
 \end{figure}
@@
 The experiments are run on a cluster with 10 nodes to store objects, one node
-to monitor the cluster, and 3 nodes that can accommodate sequencers. 
-Instead of measuring contention at the clients like
+to monitor the cluster, and 3 nodes that can accommodate sequencers.  Instead
+of measuring contention at the clients like
 Section~\ref{sec:sequencer-implementation}, these experiments measure
@@
 Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
-performance of the balancing mode that CephFS falls into {\it most of the time}.
-CephFS has 3 modes for balancing load: CPU mode, workload mode, and hybrid
-mode. All three have the same structure for making migration decisions but vary
-based on the metric used to calculate load. For this sequencer workload the 3
-different modes all have the same performance, shown in
-Figure~\ref{fig:mantle-balancer-performance} (a), because the load balancer falls
-into the same mode a majority of the time.  The high variation in performance
-for the CephFS CPU Mode bar reflects the uncertainty of using something as
-dynamic and unpredictable as CPU utilization to make migration decisions. In
-addition to the suboptimal performance and unpredictability, because
-CephFS balancers all fall into the same mode despite using different balancers
-is a problem because it prevents administrators from properly exploring the
-balancing state space.
+performance of the balancing mode that CephFS falls into {\it most of the
+time}.  CephFS has 3 modes for balancing load: CPU mode, workload mode, and
+hybrid mode. All three have the same structure for making migration decisions
+but vary based on the metric used to calculate load. For this sequencer
+workload the 3 different modes all have the same performance, shown in
+Figure~\ref{fig:mantle-balancer-performance} (a), because the load balancer
+falls into the same mode a majority of the time.  The high variation in
+performance for the CephFS CPU Mode bar reflects the uncertainty of using
+something as dynamic and unpredictable as CPU utilization to make migration
+decisions. In addition to the suboptimal performance and unpredictability,
+because CephFS balancers all fall into the same mode despite using different
+balancers is a problem because it prevents administrators from properly
+exploring the balancing state space.
 
-Mantle gives the administrator more control over balancing policies; for
-the Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the load
+Mantle gives the administrator more control over balancing policies; for the
+Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the load
 balancing interface to program logic for balancing read-heavy workloads,
 resulting in better throughput and stability.  When we did this we also
-identified two balancing modes relevant for making migration decisions
-for sequencers.
+identified two balancing modes relevant for making migration decisions for
+sequencers.
 
@@
 clients.  Doing both steps is too heavy weight for one server and sequencers on
-slave nodes can go faster if work is split up; this phenomenon is not
-uncommon and has been observed in chain replication~\cite{chain_rep}.
+slave nodes can go faster if work is split up; this phenomenon is not uncommon
+and has been observed in chain replication~\cite{chain_rep}.
 
-Cluster
-throughput improves at the cost of decreased throughput for Sequencer 2.
-Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer mode manually (no
-balancing phase) and shows that the cluster throughput is worse than the
+Cluster throughput improves at the cost of decreased throughput for Sequencer
+2.  Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer mode manually
+(no balancing phase) and shows that the cluster throughput is worse than the
 cluster throughput of proxy mode. That graph also shows that Sequencer 2 has
@@
 
-Another factor that affects performance in this environment is how much load
-is on each server; these experiments quantify that effect by programming the
-load balancing interface to control the amount of load to migrate. We call this
+Another factor that affects performance in this environment is how much load is
+on each server; these experiments quantify that effect by programming the load
+balancing interface to control the amount of load to migrate. We call this
 metric a ``migration unit".  Expressing this heuristic is not easily achievable
@@
 Figure~\ref{fig:mantle-modes}, which uses migration units equal to half the
-load on the current server, we can use: \texttt{ targets[whoami+1] = mds[whoami]["load"]/2
-}.
+load on the current server, we can use: \texttt{ targets[whoami+1] =
+mds[whoami]["load"]/2 }.
 
@@
 API to send half of the load on the current server (whoami) to the next ranked
-server (whoami + 1); the \texttt{targets} array is a globally defined table that
-the balancer uses to do the migrations. Alternatively, to migrate all load a
-time step, we can remove the division by 2.
+server (whoami + 1); the \texttt{targets} array is a globally defined table
+that the balancer uses to do the migrations. Alternatively, to migrate all load
+a time step, we can remove the division by 2.
 
@@
 
-Figure~\ref{fig:mantle-balancer-performance} (b) shows the performance of the modes
-using different migration units. Recall that this setup only has 2 sequencers
-and 2 servers, so performance may be different at scale. Even so, it is clear
-that client mode does not perform as well for read-heavy workloads. We even see
-a throughput improvement when migrating all load off the first server, leaving
-the first server to do administrative tasks (this is common in the metadata
-cluster because the first server does a lot of the cache coherence work) while
-the second server does all the client work. Proxy mode does the best in both
-cases and shows large performance gains when completely decoupling client
-request handling and operation processing in Proxy Mode (Full).  The parameter
-that controls the migration units helps the administrator control the sequencer
-co-location or distribution across the cluster. This trade-off was explored
-extensively in the Mantle paper but the experiments we present here are
-indicative of an even richer set of states to explore.
+Figure~\ref{fig:mantle-balancer-performance} (b) shows the performance of the
+modes using different migration units. Recall that this setup only has 2
+sequencers and 2 servers, so performance may be different at scale. Even so, it
+is clear that client mode does not perform as well for read-heavy workloads. We
+even see a throughput improvement when migrating all load off the first server,
+leaving the first server to do administrative tasks (this is common in the
+metadata cluster because the first server does a lot of the cache coherence
+work) while the second server does all the client work. Proxy mode does the
+best in both cases and shows large performance gains when completely decoupling
+client request handling and operation processing in Proxy Mode (Full).  The
+parameter that controls the migration units helps the administrator control the
+sequencer co-location or distribution across the cluster. This trade-off was
+explored extensively in the Mantle paper but the experiments we present here
+are indicative of an even richer set of states to explore.
 
@@
 %to drop their capabilities later (delay) improves throughput} \end{figure}
-
-\section{Related Work}
-
-% IOStack: Software-Defined Object Storage
-% sRoute: Treating the Storage Stack Like a Network
-   % - lots of good refs to check out
-% Information and control in gray-box systems
-
-Programmability of operating systems and networking resources, including
-distributed storage systems is not new, but we are not aware of work that makes
-generalization of existing services into programmable resources a key principle
-in storage systems design.
-
-Programmable storage systems can be viewed as an infrastructure for creating
-abstractions to better separate policies from mechanisms. This idea is not new.
-Software-defined networks (SDNs) create such an abstraction by separating the
-control plane from the data plane (see for example~\cite{jain:sigcomm13}). This
-notion of control/data separation was also applied in software-defined storage
-(SDS)~\cite{thereska:sosp13,stefanovici:fast16}. Similarly,
-IOStack~\cite{gracia:internet16} is providing policy-based provisioning and
-filtering in OpenStack Swift. According to a SNIA white
-paper~\cite{carlson:snia2014}, the primary goal of SDS is to control and
-facilitate flexible and dynamic provisioning of storage resources of different
-kinds, including flash memory and disk drives, to create a virtualized mapping
-between common storage abstractions (e.g. files, objects, and blocks) and
-storage devices taking data service objectives in terms of protection,
-availability, performance, and security into account. A programmable storage
-system exposes internal abstractions so that end users (not necessarily
-operators) can create new services on top of the storage stack. Thus, our
-notion of programmable storage differs from ``software-defined storage'' (SDS)
-in terms of goals and scope, although definitions of SDS are still in flux.
-
-Another view of programmable storage systems is one of tailoring systems
-resources to applications~\cite{arpaci:sosp01}. Related efforts include the
-Exokernel~\cite{engler:sosp95}, SPIN~\cite{bershad:sosp95} and
-Vino~\cite{seltzer:osdi96} projects; the latter two addressed the ability of injecting
-code into the kernel to specialize resource management. Another approach is to
-pass hints between the different layers of the IO stack to bridge the semantic
-gap between applications and
-storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}).
-
-Malacology uses the same Active and Typed Storage module presented in
-DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File Manifolds
-can be implemented with small changes to the Malacology framework, namely
-asynchronous object calls and Lua stubs in the inode, respectively.
-
-\section{Conclusion and Future Work}
-\label{conclusion-and-future-work}
-
-Programmable storage is a viable method for eliminating duplication of complex
-error-prone software used as workarounds for storage system
-deficiencies. We propose that systems expose their services in a safe way
-allowing application developers to customize system behavior to meet their
-needs while not sacrificing correctness. To illustrate the benefits of this
-approach we presented Malacology, a programmable storage system that
-facilitates the construction of new services by re-purposing existing subsystem
-abstractions of the storage stack.
-
-Future work will focus on constructing more interfaces to support a wide
-variety of storage system services that can be configured on-the-fly in existing
-systems. This work is one point along that path to producing general-purpose
-storage systems that can
-target  special-purpose applications.  Ultimately we want to utilize
-declarative methods for expressing new services.
-
-\bibliography{paper}
-\bibliographystyle{plain}
-%\printbibliography
-
-\end{document}

### 2017-02-23 paper/7_future_work.tex
paper: cleanup line breaks and break into separate files
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-02-22 paper/paper.tex
@@
-\documentclass[preprint]{sigplanconf-eurosys}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage{pifont}
-\usepackage{color}
-\usepackage{comment}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Uncomment for diff 0
-%\includecomment{notes}
-%\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
-%\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
-%\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
-
-% Uncomment for diff 1
-\includecomment{notes}
-\newcommand{\newcommentone}[1]{{\textcolor{red}{\textbf{#1}}}}
-\newcommand{\oldcommentone}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
-\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
-\newcommand{\oldcomment}[1]{}
-\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
-
-% Uncomment for camera-ready
-%\excludecomment{notes}
-%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
-%\newcommand{\oldcomment}[1]{}
-%\newcommand{\newcommentone}[1]{{\textcolor{black}{#1}}}
-%\newcommand{\oldcommentone}[1]{}
-%\newcommand{\addressesissue}[1]{}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\cmark}{\ding{51}}%
-\usepackage[font=footnotesize,labelfont=bf]{caption}
-
-\setlength{\belowcaptionskip}{-10pt}
-
-\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
-\usepackage[normalem]{ulem}
-\usepackage{pstricks}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-\setkeys{Gin}{width=\maxwidth,keepaspectratio}
-
-\usepackage[unicode=true]{hyperref}
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, Peter Alvaro, Shel Finkelstein, Jeff LeFevre},
-            pdftitle={Malacology: A Programmable Storage System},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=blue,
-            linkcolor=black,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-
-\title{Malacology: A Programmable Storage System}
-%\authorinfo{Paper ID: 165}          {Total Pages: 12}
-\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre, Carlos Maltzahn}
-           {\{msevilla, jayhawk, ivo\}@soe.ucsc.edu, \{palvaro, shel, jlefevre, carlosm\}@ucsc.edu}
-\date{}
-\begin{document}
-
-\maketitle
-
-\begin{notes}
-\textcolor{red}{
-\noindent Issues raised by reviewers, paper augmented as noted inline (red/blue text):
-\begin{enumerate}
-  \item clarify what was implemented: diagram of interfaces and prose
-  specifying the difference between Malacology and prior work (reviewers B, C,
-  D, E)
-  \item describe what makes a programmable storage system (reviewers C, D)
-  \item organize motivation: group interfaces into categories (e.g., core
-  functionality, features, and performance optimizations) and describe other
-  applications we can build with Malacology (reviewers B, C)
-  \item clarify terminology: service vs. file system metadata, storage vs.
-  file system, dirty vs. clean slate approaches, soft-state caching (reviewers
-  B, C)
-  \item address security, access control, and the safety; specify how it
-  affects composability, performance, and multi-tenancy (reviewer D)
-\end{enumerate}
-}
-\end{notes}
-
-\begin{abstract} Storage systems are caught between the need to evolve data
-processing applications efficiently and quickly, and the increasing velocity
-with which storage device technology evolves. This puts tremendous pressure on
-storage systems to support rapid change both in terms of their interfaces and
-their performance. But adapting storage systems can be difficult because
-unprincipled changes might jeopardize years of code-hardening and performance
-optimization efforts that were necessary for users to entrust their data to the
-storage system. We introduce the concept of a programmable storage system, a
-system that exposes internal services and abstractions with the goal of
-re-using and building new services on top of the storage stack. We present
-Malacology, a prototype implementation of a programmable storage system to
-explore how existing abstractions of common storage system services can be
-leveraged to adapt to the needs of new data processing systems and the
-increasing variety of storage devices. We show how this approach allows a large
-degree of flexibility for storage systems to evolve without sacrificing the
-robustness of their code-hardened subsystems. We illustrate the advantages and
-challenges of programmability by composing existing primitives into two new
-higher-level services: a \newcommentone{file system} metadata load balancer and a high-performance
-distributed shared-log. The evaluation demonstrates that our services inherit
-desirable qualities of the back-end storage system, including the ability to
-balance load, efficiently propagate cluster metadata, recover from failure, and
-to navigate trade-offs between latency and throughput using leases.
-\addressesissue{4}
-\end{abstract}
-
-\begin{figure}[tb]
-\centering
-\includegraphics{figures/overview.png}
-\caption{Scalable storage systems have storage daemons which store data,
-monitor daemons (M) that maintain cluster state, and service-specific daemons
-({\it e.g.}, file system metadata servers). Malacology enables the programmability of
-internal abstractions (bold arrows) to re-use and compose
-existing subsystems.  With Malacology, we add new services, ZLog and
-Mantle, that sit alongside traditional user-facing APIs (file, block,
-object).}\label{fig:overview}
-\end{figure}
-
-\section{Introduction}
-\label{introduction}
-\label{sec:intro}
-
-A storage system implements abstractions designed to persistently store data
-and must exhibit a high level of correctness to prevent data loss.  Storage
-systems have evolved around storage devices that often were orders of magnitude
-slower than CPU and memory, and therefore could dominate overall performance if
-not used carefully. Over the last few decades members of the storage systems
-community have developed clever strategies to meet correctness requirements
-while somewhat hiding the latency of traditional storage
-media~\cite{brewer_disks_2016}. To avoid lock-in by a particular vendor, users
-of storage systems have preferred systems with highly standardized APIs and
-lowest common denominator abstract data types such as blocks of bytes and byte
-stream files~\cite{armbrust_view_2010}.
-
-A number of recent developments have disrupted traditional storage systems.
-First, the falling prices of flash storage and the availability of new types of
-non-volatile memory that are orders of magnitude faster than traditional
-spinning media are moving overall performance bottlenecks away from storage
-devices to CPUs and networking, and pressure storage systems to shorten their
-code paths and incorporate new
-optimizations~\cite{gray_tape_2007,gray_flash_2008}.  Second, emerging ``big
-data'' applications demand interface evolution to support flexible consistency
-as well as flexible structured data 
-representations.~\cite{apache_contributors_parquet_2014}.  Finally, production-quality scalable
-storage systems available as open source software have established and are
-continuing to establish new, \emph{de-facto} API standards at a faster pace
-than traditional standards
-bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
-
-The evolutionary pressure placed on storage systems by these trends raises the
-question of whether there are principles that storage systems designers can
-follow to evolve storage systems efficiently, without jeopardizing years of
-code-hardening and performance optimization efforts.  In this paper we
-investigate an approach that focuses on identifying and exposing existing
-storage system resources, services, and abstractions that in a generalized form
-can be used to \emph{program} new services. By composing higher-level services
-over existing abstractions one can reuse subsystems and leverage their
-optimizations, established correctness, robustness, and efficiency.
-
-{\it \textbf{Contribution 1:}} We define a programmable storage system to be a
-storage system that facilitates the re-use and extension of existing storage
-abstractions provided by the underlying software stack, to enable the creation
-of new services via composition.  A programmable storage system can be realized
-by exposing existing functionality (such as \newcommentone{file system and cluster} metadata services and
-synchronization and monitoring capabilities) as interfaces that can be ``glued
-together'' in a variety of ways using a high-level language. Programmable
-storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection
-and execution of code within a storage system or storage device---in that the
-former is applicable to any component of the storage system, while the latter
-focuses at the data access level. Given this contrast, we can say that active
-storage is an example of how one internal component (the storage layer) is
-exposed in a programmable storage system.
-\addressesissue{4}
-
-To illustrate the benefits and challenges of this approach we have designed and
-evaluated Malacology, a programmable storage system that facilitates the
-construction of new services by re-purposing existing subsystem abstractions of
-the storage stack.  We build Malacology in Ceph, a popular open source software
-storage stack.  We choose Ceph to demonstrate the concept of programmable
-storage because it offers a broad spectrum of existing services, including
-distributed locking and caching services provided by \newcommentone{file system} metadata servers,
-durability and object interfaces provided by the backend object store, and
-propagation of consistent cluster state provided by the monitoring service (see
-Figure~\ref{fig:overview}).  Malacology is expressive enough to provide the
-functionality necessary for implementing new services.
-\addressesissue{4}
-
-Malacology includes a non-exhaustive set of interfaces that can be used as
-building blocks for constructing novel storage abstractions, including:
-
-\begin{enumerate}
-
-\item An interface for managing strongly-consistent time-varying \textbf{service
-metadata}.
-
-\item An interface for installing and evolving domain-specific, cluster-wide
-\textbf{data interfaces}.
-
-\item An interface for managing access to \textbf{shared resources} using a
-variety of optimization strategies.
-
-\item An interface for \textbf{load balancing} resources across the cluster.
-
-\item An interface for \textbf{durability} that persists policies using the
-underlying storage stack's object store.
-
-\end{enumerate}
-
-{\it \textbf{Contribution 2:}} We implement two distributed services using Malacology to
-demonstrate the feasibility of the programmable storage approach:
-
-\begin{enumerate}
-
-\item A high-performance distributed shared log service called ZLog, that is an
-implementation of CORFU.~\cite{balakrishnan_corfu_2012}
-
-\item An implementation of Mantle, the programmable load balancing
-service~\cite{sevilla:sc15-mantle}
-
-\end{enumerate}
-
-The remainder of this paper is structured as follows. First, we describe and
-motivate the need for programmable storage by describing current practices in
-the open source software community. Next we describe Malacology by presenting the
-subsystems within the underlying storage system that we re-purpose, and briefly
-describe how those system are used within Malacology (\S\ref{sec:malacology}).
-Then we describe the services that we have constructed within the Malacology
-framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
-implementation (\S\ref{sec:evaluation}).  We conclude by discussing related and future work.
-% and
-%conclude.
-
-\section{Application-Specific Storage Stacks}
-\label{sec:application-specifc-storage-stacks}
-
-Building storage stacks from the ground up for a specific purpose results in
-the best performance. For example, GFS~\cite{ghemawat:sosp03} and
-HDFS~\cite{shvachko:msst10} were designed specifically to serve Map/Reduce and
-Hadoop jobs, and use techniques like exposing data locality and relaxing POSIX
-constraints to achieve application-specific I/O optimizations. Another example
-is Boxwood~\cite{maccormick:osdi04}, which experimented with B-trees
-and chunk stores as storage abstractions to simplify application building.
-Alternatively, general-purpose storage stacks are built with the flexibility to
-serve many applications by providing a variety of interfaces and tunable
-parameters.  Unfortunately, managing competing forces in these systems is
-difficult and users want more control from the general-purpose storage stacks
-without going as far as building their storage system from the ground up.
-
-To demonstrate a recent trend towards more application-specific storage systems
-we examine the state of programmability in Ceph. Something of a storage Swiss
-army knife, Ceph simultaneously supports file, block, and object interfaces on
-a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
-Distributed Object Storage (RADOS) system is a cluster of object storage
-devices (OSDs) that provide Ceph with data durability and integrity using
-replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already
-provides some degree of programmability; the OSDs support domain-specific
-object interfaces that are implemented by composing existing low-level storage
-interfaces that execute atomically. These interfaces are written in C++ and are
-statically loaded into the system.
-
-The Ceph community provides empirical evidence that developers are already
-beginning to embrace programmable storage. Figure~\ref{fig:obj-int-dev-growth}
-shows a dramatic growth in the production use of domain-specific interfaces in
-the Ceph community since 2010. In that figure, classes are functional groupings
-of methods on storage objects (e.g. remotely computing and caching the checksum
-of an object extent).  What is most remarkable is that this trend contradicts
-the notion that API changes are a burden for users.  Rather it appears that
-gaps in existing interfaces are being addressed through ad-hoc approaches to
-programmability. In fact, Table~\ref{table:objclasses} categorizes existing
-interfaces and we clearly see a trend towards reusable services.
-
-\begin{figure}[ht]
-\centering
-\includegraphics{figures/obj-int-dev-growth.png}
-\caption{Since 2010, the growth in the number of co-designed object storage interfaces
-in Ceph has been accelerating. This plot is the number of object classes (a
-group of interfaces), and the total number of methods (the actual API
-end-points).}
-\label{fig:obj-int-dev-growth}
-\end{figure}
-
-\begin{table}[ht]
-\centering
-  \begin{tabular}{l|l|l}
-    Category & Example & \# \\ \hline
-    Logging  & Geographically distribute replicas & 11 \\ \hdashline
-    Metadata & Snapshots in the block device OR  & \multirow{2}{*}{74} \\
-    Management & Scan extents for file system repair & \\ \hdashline
-    Locking  & Grants clients exclusive access & 6 \\ \hdashline
-    Other & Garbage collection, reference counting  & 4\\
-\end{tabular}
-\caption{A variety of object storage classes exist to expose interfaces
-    to applications. \# is the number of methods that implement these categories.
-}
-\label{table:objclasses}
-\end{table}
-
-The takeaway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
-already trying to use programmability because their needs, whether they be
-related to performance, availability, consistency, convenience, etc., are not
-satisfied by the existing default set of interfaces. The popularity of the
-custom object interface facility of Ceph could be due to a number of reasons,
-such as the default algorithms/tunables of the storage system being
-insufficient for the application's performance goals, programmers wanting to
-exploit application-specific semantics, and/or programmers knowing how to
-manage resources to improve performance. A solution based on
-application-specific object interfaces is a way to work around the
-traditionally rigid storage APIs because custom object interfaces give
-programmers the ability to tell the storage system about their application: if
-the application is CPU or IO bound, if it has locality, if its size has the potential to
-overload a single node, etc.  Programmers often know what the problem is
-and how to solve it, but until the ability to modify object interfaces, they
-had no way to express to the storage system how to handle their data.
-
-Our approach is to expose more of the commonly used, code-hardened subsystems
-of the underlying storage system as interfaces. The intent is that these
-interfaces, which can be as simple as a redirection to the persistent data
-store or as complicated as a strongly consistent directory service, should be
-used and re-used in many contexts to implement a wide range of services. By
-making programmability a `feature', rather than a `hack' or `workaround', we
-help standardize a development process that now is largely ad-hoc.
-
-\section{Challenges}
-\label{sec:challenges}
-
-Implementing the infrastructure for programmability into existing services and
-abstractions of distributed storage systems is challenging, even if one assumes
-that the source code of the storage system and the necessary expertise for
-understanding it is available.  Some challenges include:
-
-\begin{itemize}
-
-\item Storage systems are generally required to be highly available so that any
-complete restarts of the storage system to reprogram them is usually
-unacceptable.
-
-\item Policies and optimizations are usually hard-wired into the services and
-one has to be careful when factoring them to avoid introducing additional bugs.
-These policies and optimizations are usually cross-cutting solutions to
-concerns or trade-offs that cannot be fully explored at the time the code is
-written (as they relate to workload or hardware). Given these policies and
-optimizations, decomposition of otherwise orthogonal internal abstractions can
-be difficult or dangerous.
-
-\item Mechanisms that are often only exercised according to hard-wired policies
-and not in their full generality have hidden bugs that are revealed as soon as
-those mechanisms are governed by different policies. In our experience
-introducing programmability into a storage system proved to be a great
-debugging tool.
-
-\item Programmability, especially in live systems, implies changes that need to
-be carefully managed by the system itself, including versioning and propagation
-of those changes without affecting correctness.
-
-\end{itemize}
-
-To address these challenges we present Malacology
-\oldcomment{. Malacology is both a prototype for a programmable storage system
-and a}
-\newcomment{, our prototype programmable storage system. It uses the
-programmable storage}
-\addressesissue{2}
-design approach to evolve storage systems efficiently and
-without jeopardizing years of code-hardening and performance optimization
-efforts.  Although Malacology uses the internal abstractions of the underlying
-storage system, including its subsystems, components, and implementations, we
-emphasize that our system still addresses the general challenges outlined
-above.
-
-The main challenge of designing a programmable storage system is choosing the
-right internal abstractions and picking the correct layers for exposing them.
-\oldcomment{In this paper, we do not present an exhaustive list of possible
-internal abstractions nor do we contend that the abstractions we choose provide
-the best trade-offs for all applications.} 
-\newcomment{A programmable storage is not defined by what abstractions are
-exposed, rather a programmable storage system adheres to the design approach of
-choosing interfaces so administrators can have better control of the storage
-system. The interfaces presented in this paper are abstractions that we found
-useful for building our prototype services ZLog and Mantle, yet they may not provide the best
-trade-offs for all higher-level services.}
-\addressesissue{2}
-For example, if consensus is correctly exposed one could implement
-high-level features like versioning, serialization, or various flavors of
-strongly consistent data management on top; but perhaps a low-level consensus
-interface is suited well for a particular set of applications.  These questions
-are not answered in this paper and instead we focus on showing the feasibility
-of building such a system, given advances in the quality and robustness of
-today's storage stacks.
-
-The Malacology prototype we present has been implemented on Ceph.
-While there are other systems on top of which Malacology could be implemented (see Table~\ref{table:examples}), we choose Ceph because it is a production quality system and because it
-is open source. The large developer community ensures that code is
-robust and the visibility of the code lets us expose any interface we want. In
-the next section we describe the Ceph components that we expose as Malacology
-interfaces.
-
-\section{Malacology: A Programmable Storage System}
-\label{sec:malacology}
-
-\begin{table*}
-\centering\small
-\begin{tabular}{ l | c | l | l | l }
-%\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
-\multicolumn{4}{c}{} \\
-\textbf{Interface}                      &
-\textbf{Section}                        &
-\textbf{Example in Production Systems}  &
-\textbf{Example in Ceph}                &
-\textbf{Provided Primitives}            \\ \hline
-Service Metadata
-  & \S\ref{sec:service-metadata-interface}
-  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
-  & cluster state management~\cite{website:ceph-mon}
-  & consensus/consistency
-  \\
-Data I/O
-  & \S\ref{sec:data-io-interface}
-  & Swift in situ storage/compute~\cite{website:zerocloud}
-  & object interface classes~\cite{website:cls-lua}
-  & transaction/atomicity
-  \\
-Shared Resource
-  & \S\ref{sec:shared-resource-interface}
-  & MPI collective IO, burst
-  & POSIX metadata protocols
-  & serialization/batching
-  \\
-File Type
-  & \S\ref{sec:file-type-interface}
-  & ?????
-  & file striping strategy
-  & data/metadata access
-  \\
-Load Balancing
-  & \S\ref{sec:load-balancing-interface}
-  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
-  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
-  & migration/sampling
-  \\
-Durability
-  & \S\ref{sec:durability}
-  & S3/Swift interfaces (RESTful API)
-  & object store library~\cite{weil_rados_2007}
-  & persistence/safety
-  \\
-\end{tabular}
-\caption{Common internal abstractions. The same ``internal abstractions" are
-common across large-scale systems because they provide primitives that solve
-general distributed systems problems.  Here we list examples of what these
-internal abstractions are used for in ``production systems" and in Ceph.
-Malacology provides these internal abstractions as interfaces
-\newcomment{(Section~\ref{sec:malacology})}
-that higher level
-\oldcomment{applications}
-\newcomment{services (Section~\ref{sec:services})}
-can use.
-\oldcomment{; {\it italicized abstractions} are interfaces discussed in this
-paper.}}
-\addressesissue{1}
-\label{table:examples}
-\end{table*}
-
-The guiding principle is to re-use existing services and extend them so that
-these services can be \emph{programmed}. We accomplish programmability of a
-service by exporting bindings (or ``hooks'') for an interpreted programming
-language so that programming can occur without having to restart the storage
-system (see also below,~\S\ref{sec:durability}). Table~\ref{table:examples}
-shows the internal services from Ceph that we expose in the Malacology
-prototype via Lua~\cite{ierusalimschy1996lua} bindings
-\oldcomment{.}
-\newcomment{and Figure~\ref{fig:implementation-overview} compares what was
-already present in Ceph (gray boxes) to the Malacology interfaces we added.
-Section~\ref{sec:services} will describe the higher-level services we built
-with these interfaces.}
-\addressesissue{1}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation-overview.png}
-\caption{\newcomment{Malacology is implemented on the daemons and clients that
-run in a Ceph cluster. Interfaces expose internal subsystems and are used as
-building blocks for higher-level services.}\addressesissue{1}
-\label{fig:implementation-overview}}
-\end{figure}
-
-Lua is a portable embedded scripting language and we choose it as the
-interpreted language for Malacology because it offers superior performance and
-productivity trade-offs, including a JIT-based implementation that is well
-known for near native performance. Additionally, Lua has been used extensively
-in game engines, and systems research \cite{neto:dls14-luaos}, including
-storage systems where it has been effectively used both on
-\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010}
-and off \cite{sevilla:sc15-mantle} the performance critical path. Finally, the
-flexibility of the runtime allows execution sandboxing in order to address
-security and performance concerns. We will now discuss the common subsystems
-used to manage storage system and how Malacology makes them programmable.
-
-\subsection{Service Metadata Interface}
-\label{sec:mon}
-\label{sec:service-metadata-interface}
-\label{service-metadata}
-
-% straw man example
-Keeping track of state in a distributed system is an essential part of any
-successful service and a necessary component in order to diagnose and detect
-failures, when they occur. This is further complicated by variable propagation
-delays and heterogeneous hardware in dynamic environments.
-\newcommentone{Service metadata is information about the daemons in the system
-and includes membership details, hardware layout ({\it e.g.}, racks, power
-supplies, etc.), data layout, and daemon state/configuration. It differs from
-traditional file system metadata which is information about files. For the rest
-of the paper when we refer to metadata server or service, we mean a daemon that manages file
-system metadata.}
-\addressesissue{4}
-
-\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}} In the
-case of Ceph, a consistent view of cluster state among server daemons and
-clients is critical to provide strong consistency guarantees to clients.  Ceph
-maintains cluster state information in per-subsystem  data structures called
-``maps'' that record membership and status information.  A
-Paxos~\cite{lamport_parttime_1998} monitoring service is responsible for
-integrating state changes into cluster maps, responding to requests from
-out-of-date clients and synchronizing members of the cluster whenever there is
-a change in a map so that they all observe the same system state. As a
-fundamental building block of many system designs, consensus abstractions such
-as Paxos are a common technique for maintaining consistent data versions, and
-are a useful system to expose.
-
-The default behavior of the monitor can be seen as a Paxos-based notification
-system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing
-clients to identify when new values (termed epochs in Ceph) are associated to
-given maps.  Since Ceph does not expose this service directly, as part of our Malacology implementation, we expose a key-value
-service designed for managing \oldcommentone{configuration} \newcommentone{service} metadata that is built on top of
-the consensus engine. Since the monitor is intended to be out of the
-high-performance I/O path, a general guideline is to make use of this
-functionality infrequently and to assign small values to maps.\\
-
-\noindent{\it{\textbf{Malacology: }}} Malacology exposes a strongly-consistent view of
-time-varying service metadata as a service rather than a hidden internal
-component. Malacology provides a generic API for adding arbitrary values to
-existing subsystem cluster maps. As a consequence of this, applications can
-define simple but useful service-specific logic to the strongly-consistent
-interface, such as authorization control (just specific clients can write new
-values) or triggering actions based on specific values (e.g. sanitize values).
-The higher-level services we implement in \S\ref{sec:services} make use of this
-functionality to register, version and propagate dynamic code (Lua scripts) for
-new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
-policies in the load balancer \S\ref{malacology:mds}.  Using this service
-guarantees that interface definitions are not only made durable, but are
-transparently and consistently propagated throughout the cluster so that
-clients are properly synchronized with the latest interfaces.\\
-
-\noindent\newcomment{{\it{\textbf{Impact: }}} provides core functionality
-because it lets daemons come to consensus on system critical state. Bugs in the
-internal subsystems or ommitting this from services that need this type of
-consistency affects correctness.}
-\addressesissue{3, 5}
-
-%And finally, it is important to not underestimate the importance of the
-%preservation of the modifications that enable a new system service. In many
-%(if not all) cases, interfaces defining access to data are just as important
-%as the data itself by virtue of inherently providing structural context. Thus,
-%all components of a system service must be kept to the same standards of
-%protection as data in the system itself.
-
-\oldcomment{\noindent 4.2 Object I/O Intefrace}
-\subsection{Data I/O Interface}
-\label{sec:data-io-interface}
-\label{object-data-interface}
-Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
-supports application-specific object
-interfaces~\cite{weil_rados_2007}. The ability to offload computation can
-reduce data movement, and transactional interfaces can significantly simplify
-construction of complex storage interfaces that require uncoordinated parallel
-access.
-
-\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}}
-An object interface is a plugin structured similarly to that of an RPC
-in which a developer creates a named function within the cluster that clients
-may invoke. In the case of Ceph each function is implicitly run within the
-context of an object specified when the function is called. Developers of
-object interfaces express behavior by creating a composition of native
-interfaces or other custom object interfaces, and handle serialization of
-function input and output.  A wide range of native interfaces are available to
-developers such as reading and writing to a byte stream, controlling object
-snapshots and clones, and accessing a sorted key-value database. These native
-interfaces may be transactionally composed along with application
-specific logic to create semantically rich interfaces. An example would be
-an interface that atomically updates a matrix stored in the bytestream and an
-index of the matrix stored in the key-value database.\\
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation.png}
-\caption{Malacology allows users to dynamically define object/\newcommentone{file system} metadata
-interfaces by extending the object storage daemon (OSD) and metadata server
-(MDS) subsystems with an embedded Lua VM.  It uses the service metadata
-interface (M) to propagate interfaces/versions across the cluster
-\label{fig:implementation}}
-\end{figure}
-
-\noindent {\it{\textbf{Malacology: }}} The implementation of Ceph's object abstraction,
-although powerful, does not readily support programmability. Supporting only
-C/C++ for object interface developers, Ceph requires distribution of compiled
-binaries for the correct architecture, adding a large barrier of entry for
-developers and system administrators. Second, having no way to dynamically
-unload modules, any changes require a full restart of a storage daemon which
-may have serious performance impacts due to loss of cached data. And finally,
-the security limitations of the framework limit the use of object interfaces to
-all but those with administrative level access and deep technical expertise.
-
-To address these concerns, Malacology takes advantage of Lua extensions
-contributed by the Ceph community. This allows new object interfaces to be
-dynamically loaded into the system and modified at runtime, resulting in a
-object storage API with economy of expression, which at the same time provides
-the full set of features of the base object class. New object interfaces that
-are expressed in thousands of lines of code can be implemented in approximately
-an order of magnitude less code~\cite{geambasu_comet_2010}. While the use of
-Lua does not prevent deployment of malicious code, certain types of coding
-mistakes can be handled gracefully, and access policies are used to limit access
-to trusted users~\cite{ierusalimschy1996lua}.
-
-\oldcomment{All of these modifications work in tandem to implement the desired behavior,
-and are typically co-designed together such that each depends on the other to
-behave as expected.}
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by pushing behavior to lower parts of the storage stack, thereby
-minimizing hops and distributing computation.}
-\addressesissue{3, 5}
-
-\subsection{Distributed Metadata Interfaces}
-\begin{notes}
-\textcolor{red}{This section has been re-organized}
-\end{notes}
-\label{sec:distributed-metadata-interfaces}
-\label{malacology:mds}
-
-\oldcomment{The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction~\cite{weil:sc2004-dyn-metadata}.}
-\newcomment{File systems provide clients with the familiar POSIX file abstraction.
-While this guarantees strong consistency it comes at the cost of scalability,
-increased complexity, and lower performance.}
-\addressesissue{1}
-In general, distributed file systems protect resources by providing
-hierarchical indexing and distributed locking services.
-
-%The metadata cluster moves units of the namespace called directory fragments
-%and assigns them to servers using its own metadata load balancer. It then uses
-%a hard-coded policy to balance these fragments using a scalarized load metric
-%based on CPU, workload, and file system operation metrics. Although the CephFS
-%balancer has been shown to be complicated, its complexity is justifiable.
-%Having the flexibility to choose what to move, where to move it, and how much
-%to move is very powerful. The problem is that it is difficult to pick a
-%one-size-fits-all policy, system administrators may want to partition load
-%based on SLAs, or applications are better equipped to make these decisions,
-%etc. -- but we choose instead to expose a balancing API at strategic points in
-%the balancing logic using Lua hooks. ~\cite{sevilla:sc15-mantle}.
-
-\oldcomment{\it{\textbf{Malacology:}} Interfaces are added in strategic spots in the
-\newcommentone{file system} metadata servers for guarding shared resources, defining new file types,
-and load balancing.}
-
-\oldcomment{\noindent Shared resource interface.}
-\subsubsection{Shared Resource Interface}
-\label{sec:shared-resource-interface}
-
-In general, file system metadata servers manages client sessions, allowing
-clients to obtain locks (e.g. file byte ranges), and capabilities (e.g. to
-cache file data). Clients and metadata servers use a cooperative protocol in
-which clients voluntarily release resources back to the \newcommentone{file system} metadata service in
-order to implement sharing policies.
-
-\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}}
-In Ceph, the locking service implements a capability-based system that
-expresses what data and \newcommentone{file system} metadata clients are allowed to access as well as what
-state they may cache and modify locally.  While designed for the file
-abstraction, indexing, locking, and caching are all common services that are
-useful to a broad spectrum of applications.  Distributed applications that
-share centralized resources (e.g. a database or directory) face similar
-challenges which are often solved using application-specific sharding.
-
-\newcomment{\\ \noindent\it{\textbf{Malacology: }}}
-While the current policy for sharing access and voluntarily releasing resources
-is largely best-effort, Malacology supports generalized policies between
-metadata servers and clients that can be used to implement fairness or
-priority.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} provides core functionality
-to protect and provide exclusive access for any shared resource. May hurt
-performance if the resource in question does not require strong consistency.}
-\addressesissue{3}
-
-\oldcomment{\noindent File type interface.}
-\subsubsection{Malacology: File Type Interface}
-\label{sec:file-type-interface}
-
-Applications that manage large amounts of \newcommentone{file system} metadata (e.g. users or database
-snapshots) often require a naming service.  The metadata service manages a
-POSIX file system hierarchy where files and directories are represented as
-inode data structures and expose a POSIX file interface.
-
-\newcomment{\\ \noindent {\it{\textbf{Existing Ceph Implementation: }}} CephFS
-is the POSIX compliant file system that uses Ceph.  Inodes are quite large (1KB
-for an inode, 400 bytes for a directory entry, and 700 bytes for a directory)
-and contain CephFS-specific policies like how to stripe data across RADOS.}
-
-\newcomment{\\ \noindent {\it{\textbf{Malacology: }}}} We allow new inode types to
-be defined such that applications can create domain-specific interfaces to
-inodes that may modify locking and capability policies. We will show how this
-is used in Section~\ref{sec:seq} when we discuss a distributed shared-log built
-on Malacology.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this interface is both a feature
-and a performance optimization. It is a feature because it allows developers to
-add support for different  storage types, such as how to read new file formats
-or what consistency semantics to use for a specific subtree in the hierarchical
-namespace. It is also a performance optimization because future programmers can
-add optimizations for processing specific types of files into the inode itself.}
-\addressesissue{3}
-
-%%% TODO: striping strategy is stored with the inode
-
-\oldcomment{\noindent \bf Load balancing interface.} 
-\subsubsection{Load Balancing Interface}
-\label{sec:load-balancing-interface}
-
-\newcomment{Many large scale storage systems separate file system metadata and data IO so
-that the corresponding services can scale independently. Metadata requests
-transfer small amounts of data and they happen relatively frequently so many
-systems employ separate file system metadata clusters.}
-
-\newcomment{\\ \noindent {\it{\textbf{Existing Ceph Implementation: }}}} Ceph
-also addresses the challenge of balancing \newcommentone{file system} metadata load with a separate
-metadata cluster. This cluster uses load balancing policies to migrate
-directory inodes around the cluster to alleviate overload at single
-nodes~\cite{weil:sc2004-dyn-metadata}.  The policies use metrics based on
-system state (e.g.  CPU and memory utilization) and statistics collected by the
-cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
-partitioning to move variable sized namespace subtrees. These units can be
-shipped anywhere (i.e., to any metadata server of any capacity) at any time for
-any reason. The original balancer was designed with hard-coded policies and
-tunables.\\
-
-\noindent {\it{\textbf{Malacology: }}}
-Mantle~\cite{sevilla:sc15-mantle}, a programmable \newcomment{file system}metadata load balancer, has
-been implemented on Malacology so that its infrastructure can enjoy all the
-properties of existing internal abstractions.  The load balancing logic is
-intact but the infrastructure has been improved to safely load, manage, and
-persist its load balancing policies.  Composing Mantle with the service
-metadata and durability interfaces, Mantle can safely version balancer
-policies, save balancer policies in the backend object store and centralize
-warnings/errors. When combined with file types, load balancing policies can
-express a variety of rules such as multi-tenant and multi-application
-balancing.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by allowing them to specify how to partition, replicate, and
-distribute metadata in response to overloaded servers.}
-\addressesissue{3}
-
-\subsection{Durability Interface}
-\label{sec:durability}
-
-\newcomment{Object stores protect data using techniques like erasure coding,
-replication, and data scrubbing. For scalability, many of these features are
-implemented using a peer-to-peer protocol that allows object storage devices to
-operate autonomously without a centralized coordinator.\\}
-
-\noindent{\it{\textbf{Existing Ceph Implementation: }}}
-Ceph provides storage by striping and replicating data across
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-protects data using common techniques such as erasure coding, replication, and
-scrubbing.  example, when the number of placement groups change, the OSDs
-re-balance and re-shard data in the background in a process called placement
-group splitting in which OSDs communicate directly with each other to cover a
-new state.  In order to reduce load on the monitoring service, Ceph OSDs use a
-gossip protocol to efficiently propagate changes to cluster maps throughout the
-system, and autonomously initiate recovery mechanisms when failures are
-discovered.\\
-
-\noindent {\it{\textbf{Malacology: }}} Metadata service policies and object storage
-interfaces are stored durability within RADOS and managed by storing references
-with the object maps. Since the cluster already propagates a consistent view of
-these data structures, we use this service to automatically install interfaces
-in OSDs, and install policies within the metadata server daemon \oldcommentone{(MDS)} such that clients and daemons are
-synchronized on correct implementations without restarting.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this is a feature because it
-adds data safety and persistence to system metadata; while nice to have it does
-not necessarily effect correctness.}
-\addressesissue{3}
-
-\section{Services Built on Malacology}
-\label{sec:services}
-\label{services-built-on-malacology}
-
-\label{services}
-%\begin{table}
-%\centering\small
-%\begin{tabular}{  l | l | l }
-%\textbf{Service}              &
-%\textbf{Used to...}           &
-%\textbf{Malacology Interface} \\ \hline
-%Mantle & load balance metadata     & Load Balancing  \\
-%       & agree on current balancer & Service Metadata\\
-%       & control client write lock & Shared Resource \\
-%       & persist policy logic      & Durability      \\ \hline
-%ZLog   & implement CORFU           & Object \& Data  \\ 
-%       & install ZLog interfaces   & Service Metadata\\
-%       & persist log entries       & Durability      \\ \hline
-%ZLog   & load balance sequencers   & Load Balancing  \\
-%Seqr.  & manage access to log tail & Shared Resource \\ 
-%\end{tabular}
-%\caption{Mantle and ZLog are built using Malacology interfaces so they inherit the
-%robustness of the subsystems of the underlying stack.}
-%\label{table:implementation}
-%\end{table}
-
-In this section we describe two services built on top of Malacology. The first
-is Mantle, a framework for dynamically specifying metadata load balancing
-policies. The second system, ZLog, is a high-performance distributed shared-log.
-In addition to these services, we'll demonstrate how we combine ZLog and Mantle
-to implement service-aware metadata load balancing policies.
-
-\subsection{Mantle: Programmable Load Balancer}
-\label{sec:mantle}
-
-Mantle is a programmable load balancer that separates the metadata balancing
-policies from their mechanisms. Administrators inject code to change how the
-metadata cluster distributes metadata. In ~\cite{sevilla:sc15-mantle} the
-authors showed how using Mantle, one can implement a single node metadata
-service, a distributed metadata service with hashing, and a distributed
-metadata service with dynamic subtree partitioning.
-
-The original implementation was ``hard-coded" into Ceph and lacked robustness
-(no versioning, durability, or policy distribution).  Re-implemented using
-Malacology, Mantle now enjoys (1) the versioning provided by the metadata
-service interface provided by the monitor daemons and (2) the durability and
-distribution provided by Ceph's reliable object store.  Re-using the internal
-abstractions with Malacology resulted in a 2\(\times\) reduction in source code
-compared to the original implementation.
-
-\subsubsection{Versioning Balancer Policies}
-
-Ensuring that the version of the current load balancer is consistent across the
-physical servers in the metadata cluster was not addressed in the original
-implementation. The user had to set the version on each individual server and
-it was trivial to make the versions inconsistent. Maintaining consistent
-versions is important for balancing policies that are cooperative in which local
-decisions are made assuming properties about other instances of the same
-policy.
-
-With Malacology, Mantle stores the version of the current load balancer in the
-Ceph service metadata interface. The version of the load balancer corresponds
-to an object name in the balancing policy. Using the service metadata interface
-ensures that all metadata servers use the same version of the load balancer. As
-a result, the policy version inherits the consistency of Ceph's internal
-monitor daemons. The user changes the version of the load balancer using a new
-CLI command.
-
-\subsubsection{Making Balancer Policies Durable}
-
-The load balancer version described above corresponds to the name of an object
-in RADOS that holds the actual Lua balancing code.  When MDS nodes start
-balancing load, they first check the latest version from the MDS map and
-compare it to the balancer they have loaded. If the version has changed, they
-dereference the pointer to the balancer version by reading the corresponding
-object in RADOS. This is in contrast to the original balancer which stored load
-balancer code on the local file system -- a technique which is unreliable and
-may result in silent corruption.
-
-The balancer pulls the Lua code from RADOS synchronously; asynchronous reads
-are not possible because of the architecture of the MDS. The synchronous
-behavior is not the default behavior for RADOS operations, so we achieve this
-with a timeout: if the asynchronous read does not come back within half the
-balancing tick interval the operation is canceled and a Connection Timeout
-error is returned. By default, the balancing tick interval is 10 seconds, so
-Mantle will use a 5 second second timeout.
-
-This design allows Mantle to immediately return an error if anything
-RADOS-related goes wrong.  We use this implementation because we do not want to
-do a blocking OSD read from inside the global MDS lock. Doing so would bring
-down the MDS cluster if any of the OSDs are not responsive.
-
-Storing the balancers in RADOS is simplified by the use of an interpreted
-language for writing balancer code. If we used a language that needs to be
-compiled, like the C++ object classes in the OSD, we would need to ensure
-binary compatibility, which is complicated by different operating systems,
-distributions, and compilers.
-
-\subsubsection{Logging, Debugging, and Warnings}
-
-In the original implementation, Mantle would log all errors, warnings, and
-debug messages to a log stored locally on each MDS server. To get the simplest
-status messages or to debug problems, the user would have to log into each MDS
-individually, look at the logs, and reason about causality and ordering.
-
-With Malacology, Mantle re-uses the centralized logging features of the
-monitoring service. Important errors, warnings, and info messages are collected
-by the monitoring subsystem and appear in the monitor cluster log so instead of
-users going to each node, they can watch messages appear at the monitor.
-Messages are logged sparingly, so as not to spam the monitor with frivolous
-debugging but important events, like balancer version changes or failed
-subsystems show up in the centralized log.
-
-\subsection{ZLog: A Fast Distributed Shared Log}
-\label{sec:zlog}
-
-The second service implemented on Malacology is ZLog, a high-performance
-distributed shared-log that is based on the CORFU
-protocol~\cite{balakrishnan_corfu_2012}. The shared-log is a
-powerful abstraction used to construct distributed systems, such as
-metadata management~\cite{balakrishnan:sosp13} and elastic database
-systems~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
-However, existing implementations that rely on consensus algorithms such as
-Paxos funnel I/O through a single point introducing a bottleneck that restricts
-throughput. In contrast, the CORFU protocol is able to achieve high throughput
-using a network counter called a \emph{sequencer}, that
-decouples log position assignment from log I/O.
-
-While a full description of the CORFU system is beyond the scope of this
-paper, we briefly describe the custom storage device interface, sequencer
-service, and recovery protocol, and how these services are instantiated in the
-Malacology framework.
-
-\subsubsection{Sequencer}
-\label{sec:seq}
-
-High-performance in CORFU is achieved using a sequencer service that assigns
-log positions to clients by reading from a volatile, in-memory counter which
-can run at a very high throughput and at low latency. Since the sequencer is
-centralized, ensuring serializability in the common case is trivial.  The
-primary challenge in CORFU is handling the failure of the sequencer in a way
-that preserves correctness. Failure of the sequencer service in CORFU is
-handled by a recovery algorithm that recomputes the new sequencer state using a
-CORFU-specific custom storage interface to discover the tail of the log, while
-simultaneously invalidating stale client requests using an epoch-based
-protocol.
-
-{\bf Sequencer interface.} The sequencer resource supports the ability to
-\emph{read()} the current tail value and get the \emph{next()} position in
-the log which also atomically increments the tail position.
-We implement the sequencer service in Malacology as an application-specific
-file type that associates a small amount of metadata
-with the inode state managed by the Ceph metadata service.
-This approach has the added
-benefit of allowing the metadata service to handle naming, by
-representing each sequencer instance in the standard POSIX hierarchical
-namespace. The primary challenge in mapping the sequencer resource to the
-metadata service is handling serialization correctly to maintain the global
-ordering provided by the CORFU protocol.
-
-Initially we sought to directly model the sequencer service in Ceph as a
-non-exclusive, non-cacheable resource, forcing clients to perform a round-trip
-access to the resource at the authoritative metadata server for the sequencer
-inode.  Interestingly, we found that the capability system in Ceph uses a
-strategy to reduce metadata service load by allowing clients that open a
-shared file to temporarily obtain an exclusive cached copy of the resource,
-resulting in a round-robin, best-effort batching behavior. When a single
-client is accessing the sequencer resource it is able to increment the
-sequencer locally. Any competing client cannot query the sequencer until the
-metadata service has granted it access.
-
-While unexpected, this discovery allowed us to explore an implementation
-strategy that we had not previously considered. In particular, for bursty
-clients, and clients that can tolerate increased latency, this mode of
-operation may allow a system to achieve much higher throughput than a system
-with a centralized sequencer service.  We utilize the programmability of the
-metadata service to define a new policy for handling capabilities that controls
-the amount of time that clients are able to cache the sequencer resource. This
-allows an administrator or application to control the trade-off between latency
-and throughput beyond the standard best-effort policy that is present in Ceph
-by default.
-
-In Section~\ref{sec:evaluation} we quantify the trade-offs of throughput and
-latency for an approach based on a round-robin batching mode, and compare this
-mode to one in which the metadata server mediates access to the sequencer state
-when it is being shared among multiple clients. Quantifying these trade-offs
-should provide administrators with guidelines for setting the tunables for
-different ``caching" modes of the sequencer.
-
-{\bf Balancing policies.} As opposed to the batching mode for controlling access to the sequencer
-resource, more predictable latency can be achieved by treating the sequencer
-inode as a shared non-cacheable resource, forcing clients to make a round-trip
-to the metadata service. However, the shared nature of the metadata service
-may prevent the sequencer from achieving maximum throughput. To provide a
-solution to this issue we have taken advantage of the programmability of the
-metadata service load balancing to construct a service-specific load balancing
-policy. As opposed to a balancing policy that strives for uniform
-load distribution, a ZLog-specific policy may utilize knowledge of inode types
-to migrate the sequencer service to provisioned hardware during periods of
-contention or high demand.
-
-\subsubsection{Storage Interface}
-
-The storage interface is a critical component in the CORFU protocol. Clients
-independently map log positions that they have obtained from the sequencer
-service (described in detail in the next section) onto storage devices, while
-storage devices provide an intelligent \emph{write}-once, random \emph{read}
-interface for accessing log entries. The key to correctness in CORFU lies with
-the enforcement of up-to-date epoch tags on client requests; requests tagged
-with out-of-date epoch values are rejected, and clients are expected to
-request a new tail from the sequencer after refreshing state from an auxiliary
-service.  This mechanism forms the basis for sequencer recovery.
-
-In order to repopulate the sequencer state (i.e. the cached, current tail of
-the log) during recovery of a sequencer, the maximum position in the log must
-be obtained. To do this, the storage interface exposes an additional
-\emph{seal} method that atomically installs a new epoch value and returns the
-maximum log position that has been written.
-
-Since the sequencer service does not resume until the recovery process has
-completed, there cannot be a race with clients appending to the log, and the
-immutability of the log allow reads to never block during a sequencer failure.
-Recovery of the sequencer process itself may be handled in many ways, such as
-leader election using an auxiliary service such as Paxos. In our
-implementation, the recovery is the same as (and is inherited from) the Ceph
-metadata service. Handling the failure of a client that holds the sequencer
-state is similar, although a timeout is used to determine when a client should
-be considered unavailable.
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-%We wanted to implement CORFU and we had two choices: build it from the ground
-%up or do it on top of Malacology. To show the feasability of building on
-%Malacology our evaluation focuses on the
-
-Our evaluation demonstrates the feasibility of building new service
-abstractions atop programmable storage, focusing on the performance of the
-internal abstractions exposed by Malacology and used to construct the Mantle
-and ZLog services. We also discuss latent capabilities we discovered in this
-process that let us navigate different trade-offs within the services
-themselves. First, we benchmark scenarios with high sequencer contention by
-examining the interfaces used to map ZLog onto Malacology; specifically, we
-describe the sequencer implementation and the propagation of object and data
-interfaces interfaces.  Next, we benchmark scenarios in which the storage
-system manages multiple logs by using Mantle to balance sequencers across a
-cluster.
-
-% should save this for future work, otherwise it undercuts the eval.
-%Detailed system-wide performance measurements and analysis is forthcoming for
-%both Mantle and ZLog, as we try to find the best metadata balancing policies
-%and physical design implementation trade-offs.
-
-Since this work focuses on the programmability of Malacology, the goal of this
-section is to show that the components and subsystems that support the
-Malacology interfaces provide reasonable relative performance, as well as to
-give examples of the flexibility that Malacology provides to programmers.
-
-%As we search for the best metadata balancing policies
-%and physical design implementation trade-offs,
-%we plan to rel
-
-\subsection{Mapping ZLog onto Malacology}
-\label{sec:mapping-zlog-onto-malacology}
-
-We evaluate Malacology by exploring one possible mapping of the ZLog
-implementation of CORFU onto Ceph in which we re-use (1) the metadata service
-to manage naming and synchronization of the sequencer resource by treating the
-resource as an inode, and (2) the monitoring sub-subsystem to distribute and
-install application-specific I/O interface required of the CORFU protocol. In
-Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the inode
-abstraction for implementing the sequencer resource enables load balancing policies
-to migrate the sequencer resource in heavy-load situation.
-
-\subsubsection{Sequencer Implementation}
-\label{sec:sequencer-implementation}
-
-We evaluate the feasibility of using the metadata service to implement a
-sequencer resource that is responsible for maintaining a total ordering of
-the log. Clients contact the sequencer to obtain the tail of the log and then
-independently initiate I/O, thus we measure both the throughput and latency of
-obtaining new tail positions which bounds client append performance.
-
-The sequencer is implemented as a custom file type in which the sequencer
-state (a 64-bit integer) is embedded in the inode of a file representing the
-sequencer. A total ordering of the log is imposed by the re-use of the
-capability service that can be used to grant exclusive access of inode state
-to clients. The metadata service is responsible for maintaining exclusivity
-and granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
-behavior of the system in which a best-effort policy is used.  The two colors
-represent points in time that the clients were able to access the resource.
-The best-effort policy shows a high degree of interleaving
-between clients but the system spends a large portion of time re-distributing
-the capability, reducing overall throughput.
-
-In order to control the performance of the system we implement a
-policy that (1) restricts the length of time that a client may maintain
-exclusive access and (2) limits the number of log positions that a client may
-generate without yielding to other clients waiting for access. The behavior of
-these two modes is illustrated in Figures~\ref{fig:capdelay-quota-behavior}
-(b) and (c), respectively.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/capdelay-quota-behavior-a.png}
-\includegraphics{figures/capdelay-quota-behavior-b.png}
-\includegraphics{figures/capdelay-quota-behavior-c.png}
-\caption{Each dot is an individual request, spread randomly along the \(y\)
-axis. The default behavior is unpredictable, ``delay" lets clients hold the
-lease longer, and ``quota" gives clients the lease for a number of operations.}
-\label{fig:capdelay-quota-behavior}
-\end{figure}
-
-Figure~\ref{fig:captp} demonstrates a configurable trade-off between
-throughput and latency. In the experiment two clients are run each with a
-fixed 0.25 second maximum reservation on the capability, and we vary the size
-of the log position quota running each configuration for two minutes. The
-total operations per second is the combined throughput of the two clients, and
-the average latency is number of microseconds required to obtain a new log
-position. With a small quota more time is spent exchanging exclusive access,
-while a large quota reservation allows clients to experience a much lower
-latency because they experience uncontended access for a longer period of time.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/tradeoff.png}
-\caption{Sequencer throughput by re-using various services.  The highest performance is
-achieved using a single client with exclusive, cacheable privilege. Round-robin
-sharing of the sequencer resource is affected by the amount of time the
-resource is held, with best-effort performing the worst.}
-\label{fig:captp}
-\end{figure}
-
-To get a better picture of latency, Figure~\ref{fig:capcdf} shows the CDF of
-latency for each client in all experiment configurations. At the 99th
-percentile clients accessed the sequencer is less than a millisecond. The CDF
-is cropped at the 99.999th percentile due to large outliers that we believe
-occur in instances in which the MDS is performing I/O while it is in the
-process of re-distributing the capability to another client.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/caps-delay-latency.png}
-\caption{
-The latency distribution shows that if the clients hold their capabilities longer,
-performance decreases. Changing the delay can help the sequencer and
-clients strike a balance between throughput and latency.}
-\label{fig:capcdf}
-\end{figure}
-
-\textbf{Malacology exposes} the internal capability management service and
-allows users to navigate latency and throughput trade-offs.  Other approaches to
-designing the sequencer service also exist, such as using a centralized service
-in which each access is a network round-trip. In contrast to the mechanism we
-explored which is appropriate for clients with bursty workloads, it may be
-easier to provide predictable performance using a centralized service, and we
-will be exploring in the future how this can be achieved using the capability
-system.
-
-\subsubsection{Interface Propagation}
-
-Domain-specific data interfaces (Section~\ref{sec:application-specifc-storage-stacks}) allow
-co-design between applications and the storage system. Malacology supports
-custom object interfaces in RADOS that require interface implementations to
-be installed on the storage devices in the system, supporting the evolution of
-interfaces through automatic system-wide versioning and installation through
-the service metadata interface (Section~\ref{service-metadata}). We evaluate
-the performance of installing a new interface version in the cluster,
-which is an important metric for applications that frequently evolve
-interfaces.
-
-We demonstrate the feasibility of utilizing the Ceph monitoring sub-system by
-evaluating the performance of installing and distributing interface updates.
-Figure~\ref{fig:propdelay} shows the CDF of the latency of interface updates.
-The interfaces are Lua scripts embedded in the cluster map and distributed
-using a peer-to-peer gossip protocol.  The latency is defined as the elapsed
-time following the Paxos proposal for an interface update until each OSD makes
-the update live (the cost of the Paxos proposal is configurable and is
-discussed below). The latency measurements were taken on the nodes running
-Ceph server daemons, and thus exclude the client round-trip cost. In each of
-the experiments 1000 interface updates were observed.
-
-Figure~\ref{fig:propdelay} shows the lower bound cost for updates in a large
-cluster. In the experiment labeled ``120 OSD (RAM)" a cluster of 120 OSDs (10
-OSDs x 12 servers) using an in-memory data store were deployed, showing a
-latency of less than 54 ms with a probability of 90\% and a worst case latency
-of 194 ms. These costs demonstrate the penalty of distributing the interface
-in a large cluster. In practice the costs include, in addition to cluster-wide
-propagation of interface updates, the network round-trip to the
-interface management service, the Paxos commit protocol itself, and other factors
-such as system load. By default Paxos proposals occur periodically with a 1
-second interval in order to accumulate updates. In a minimum, realistic quorum
-of 3 monitors using HDD-based storage, we were able to decrease this interval
-to an average of 222 ms.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=60mm,trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
-\caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
-interface commit.}
-\label{fig:propdelay}
-\end{figure}
-
-\subsection{Load Balancing ZLog Sequencers with Mantle}
-\label{sec:zlog-balancing}
-
-In practice, a storage system implementing CORFU will support a multiplicity of
-independent totally-ordered logs for each application.  For this scenario
-co-locating sequencers on the same physical node is not ideal but building a load
-balancer that can migrate the shared resource ({\it e.g.,}, the resource that
-mediates access to the tail of the log) is a time-consuming, non-trivial task.
-It requires building subsystems for migrating resources, monitoring the
-workloads, collecting metrics that describe the utilization on the physical
-nodes, partitioning resources, maintaining cache coherence, and managing
-multiple sequencers. The following experiments demonstrate the feasibility of
-using the mechanisms of the Malacology load balancing interface to inherit
-these features and to alleviate load from overloaded servers.
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-behaviors.png}
-\caption{CephFS/Mantle load balancing have better throughput than co-locating
-all sequencers on the same server.  Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify this improvement;
-Section~\ref{sec:feature-backoff} examines the migration at 0-60 seconds.
-}\label{fig:mantle-balancer-behaviors}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-performance.png}
-\includegraphics{figures/mantle-mode-performance.png}
-\caption{In (a) all CephFS balancing modes have the same performance; Mantle uses a
-balancer designed for sequencers. In (b) the best combination of mode and
-migration units can have up to a 2\(\times\) improvement.}\label{fig:mantle-balancer-performance}
-\end{figure}
-
-The experiments are run on a cluster with 10 nodes to store objects, one node
-to monitor the cluster, and 3 nodes that can accommodate sequencers. 
-Instead of measuring contention at the clients like
-Section~\ref{sec:sequencer-implementation}, these experiments measure
-contention at the sequencers by forcing clients to make round-trips for every
-request. We implement this using the shared resource interface that forces
-round-trips.  Because the sequencer's only function is to  hand out positions
-for the tail of the log, the workload is read-heavy.
-
-First, we show how the ZLog service can orchestrate multiple sequencers using
-the Malacology load balancing interface.
-Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput over time of
-different load balancers as they migrate 3 sequencers (with 4 clients) around
-the cluster; ``No Balancing" keeps all sequencers on one server, ``CephFS"
-migrates sequencers using the CephFS load balancers, and ``Mantle" uses a
-custom load balancer we wrote specifically for sequencers. The CephFS load
-balancers are hard-coded into Ceph while Mantle is an API for balancing load
-presented in~Mantle~\cite{sevilla:sc15-mantle} and re-implemented on top of
-Malacology in this paper.  The increased throughput for the CephFS and Mantle
-curves between 0 and 60 seconds are a result of migrating the sequencer(s) off
-overloaded servers.
-
-In addition to showing that migrating sequencers improves performance,
-Figure~\ref{fig:mantle-balancer-behaviors} also demonstrates features that we
-will explore in the rest of this section.
-Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify the differences in performance
-when the cluster stabilizes at time 100 seconds and
-Section~\ref{sec:feature-backoff} examines the slope and start time of the
-rebalancing phase between 0 and 60 seconds by comparing the aggressiveness of
-the balancers.
-
-\subsubsection{Feature: Balancing Modes}
-\label{sec:feature-balancing-modes}
-
-Next, we quantify the performance benefits shown in
-Figure~\ref{fig:mantle-balancer-behaviors}.  To understand why load balancers
-perform differently we need to explain the different balancing modes that the
-load balancer service uses and how they stress the subsystems that receive and
-forward client requests in different ways. In
-Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
-performance of the balancing mode that CephFS falls into {\it most of the time}.
-CephFS has 3 modes for balancing load: CPU mode, workload mode, and hybrid
-mode. All three have the same structure for making migration decisions but vary
-based on the metric used to calculate load. For this sequencer workload the 3
-different modes all have the same performance, shown in
-Figure~\ref{fig:mantle-balancer-performance} (a), because the load balancer falls
-into the same mode a majority of the time.  The high variation in performance
-for the CephFS CPU Mode bar reflects the uncertainty of using something as
-dynamic and unpredictable as CPU utilization to make migration decisions. In
-addition to the suboptimal performance and unpredictability, because
-CephFS balancers all fall into the same mode despite using different balancers
-is a problem because it prevents administrators from properly exploring the
-balancing state space.
-
-Mantle gives the administrator more control over balancing policies; for
-the Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the load
-balancing interface to program logic for balancing read-heavy workloads,
-resulting in better throughput and stability.  When we did this we also
-identified two balancing modes relevant for making migration decisions
-for sequencers.
-
-Using Mantle, the administrator can put the load balancing service into ``proxy
-mode" or ``client mode". In proxy mode one server receives all requests and
-farms off the requests to slave servers; the slaves servers do the actual tail
-finding operation. In client mode, clients interact directly with the server
-that has their sequencer.  These modes are illustrated in
-Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
-co-located on one physical server -- performance for that mode is shown by the
-No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
-Mode" in the lower left shows how clients for sequencer 2 are re-directed to
-server B. The two ``Proxy mode" diagrams on the right show how clients continue
-sending requests to server A even though sequencer 2 has migrated to server B
-(the meaning of ``(Half)" is discussed in the next section).
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-modes.png}
-\caption{ In client mode clients sending requests to the server that houses
-their sequencer. In proxy mode clients continue sending their requests to the
-first server.  }\label{fig:mantle-modes}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-mode-behavior.png}
-\caption{The performance of proxy mode achieves the highest throughput but at
-the cost of lower throughput for one of the sequencers. Client mode is more
-fair but results in lower cluster throughput.
-}\label{fig:mantle-mode-behavior}
-\end{figure}
-
-Figure~\ref{fig:mantle-mode-behavior} shows the throughput over time of the two
-different modes for an environment with only 2 sequencers (again 4 clients
-each) and 2 servers. The curves for both sequencers in
-Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second and
-at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
-Performance of Sequencer 2 decreases because it stayed on the proxy which now
-handles all client requests, processes requests for Sequencer 2, and forwards
-requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
-because distributing the sequencers in this way separates (1) the handling of
-the client requests and (2) finding the tail of the log and responding to
-clients.  Doing both steps is too heavy weight for one server and sequencers on
-slave nodes can go faster if work is split up; this phenomenon is not
-uncommon and has been observed in chain replication~\cite{chain_rep}.
-
-Cluster
-throughput improves at the cost of decreased throughput for Sequencer 2.
-Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer mode manually (no
-balancing phase) and shows that the cluster throughput is worse than the
-cluster throughput of proxy mode. That graph also shows that Sequencer 2 has
-less throughput than Sequencer 2. In this case, the scatter-gather process used
-for cache coherence in the metadata protocols causes strain on the server
-housing Sequencer 2 resulting in this uneven performance.
-
-\subsubsection{Feature: Migration Units}
-\label{sec:feature-migration-units}
-
-Another factor that affects performance in this environment is how much load
-is on each server; these experiments quantify that effect by programming the
-load balancing interface to control the amount of load to migrate. We call this
-metric a ``migration unit".  Expressing this heuristic is not easily achievable
-with outward facing tunable parameters (i.e. system knobs) but with Mantle's
-programmable interface, we can force the load balancer to change its migration
-units. To force the balancer into the Proxy Mode (Half) scenario in
-Figure~\ref{fig:mantle-modes}, which uses migration units equal to half the
-load on the current server, we can use: \texttt{ targets[whoami+1] = mds[whoami]["load"]/2
-}.
-
-This code snippet uses globally defined variables and tables from the Mantle
-API to send half of the load on the current server (whoami) to the next ranked
-server (whoami + 1); the \texttt{targets} array is a globally defined table that
-the balancer uses to do the migrations. Alternatively, to migrate all load a
-time step, we can remove the division by 2.
-
-%\begin{figure}[t!]
-%\centering
-%\includegraphics{figures/mantle-mode-performance.png}
-%\caption{The performance of the different modes using various migration units
-%shows almost a 2\(\times\) improvement in the best case.
-%}\label{fig:mantle-mode-performance}
-%\end{figure}
-
-Figure~\ref{fig:mantle-balancer-performance} (b) shows the performance of the modes
-using different migration units. Recall that this setup only has 2 sequencers
-and 2 servers, so performance may be different at scale. Even so, it is clear
-that client mode does not perform as well for read-heavy workloads. We even see
-a throughput improvement when migrating all load off the first server, leaving
-the first server to do administrative tasks (this is common in the metadata
-cluster because the first server does a lot of the cache coherence work) while
-the second server does all the client work. Proxy mode does the best in both
-cases and shows large performance gains when completely decoupling client
-request handling and operation processing in Proxy Mode (Full).  The parameter
-that controls the migration units helps the administrator control the sequencer
-co-location or distribution across the cluster. This trade-off was explored
-extensively in the Mantle paper but the experiments we present here are
-indicative of an even richer set of states to explore.
-
-\subsubsection{Feature: Backoff}
-\label{sec:feature-backoff}
-
-Tuning the aggressiveness of the load balancer decision making is also a
-trade-off that administrators can control and explore. The balancing phase from
-0 to 60 seconds in Figure~\ref{fig:mantle-balancer-behaviors} shows different
-degrees of aggressiveness in making migration decisions; CephFS makes a
-decision quickly and throughput jumps to 2500 ops/second within 10 seconds
-while Mantle takes more time to stabilize. This conservative behavior is
-controlled by programming the balancer to (1) use different conditions for when
-to migrate and (2) using a threshold for sustained overload.
-
-We control the conditions for when to migrate using \texttt{when()}, a callback
-in the Mantle API.  For the Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} we program \texttt{when()} to wait
-for load on the receiving server to fall below a threshold. This makes the
-balancer more conservative because it takes 60 seconds for cache coherence
-messages to settle.  The Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} also takes longer to reach peak
-throughput because we want the policy to wait to see how migrations affect the
-system before proceeding; the balancer does a migration right before 50
-seconds, realizes that there is a third underloaded server, and does another
-migration.
-
-The other way to change aggressiveness of the decision making is to program
-into the balancer a threshold for sustained overload. This forces the balancer
-to wait a certain number of iterations after a migration before proceeding. In
-Mantle, the policy would use the save state function to do a countdown after a
-migration.  Behavior graphs and performance numbers for this backoff feature is
-omitted for space considerations, but our experiments confirm that the more
-conservative the approach the less overall throughput.
-
-\textbf{Malacology pulls} the load balancing service out of the storage system
-to balance sequencers across a cluster. This latent capability also gives
-future programmers the ability to explore the different load balancing
-trade-offs including: load balancing modes to control forwarding vs. client
-redirection, load migration units to control sequencer distribution vs.
-co-location, and backoffs to control conservative vs. aggressive decision
-making.
-
-%\subsubsection{Experiment: Object Class
-%Programmability}\label{experiment-object-class-programmability}
-%
-%Add quote from CORFU paper that talks about what they had to hack
-%
-%\subsubsection{Experiment: Multi-Client
-%Burstiness}\label{experiment-multi-client-burstiness}
-%
-%\begin{figure}[htbp] \centering
-%\includegraphics{figures/caps-delay-thruput.png} \caption{Forcing the client
-%to drop their capabilities later (delay) improves throughput} \end{figure}
-
 \section{Future Work}
@@
 protect applications (which evolve slowly) against changes in highly dynamic
-storage system.}
- \addressesissue{3}
-
-\section{Related Work}
-
-% IOStack: Software-Defined Object Storage
-% sRoute: Treating the Storage Stack Like a Network
-   % - lots of good refs to check out
-% Information and control in gray-box systems
-
-Programmability of operating systems and networking resources, including
-distributed storage systems is not new, but we are not aware of work that makes
-generalization of existing services into programmable resources a key principle
-in storage systems design.
-
-Programmable storage systems can be viewed as an infrastructure for creating
-abstractions to better separate policies from mechanisms. This idea is not new.
-Software-defined networks (SDNs) create such an abstraction by separating the
-control plane from the data plane (see for example~\cite{jain:sigcomm13}). This
-notion of control/data separation was also applied in software-defined storage
-(SDS)~\cite{thereska:sosp13,stefanovici:fast16}. Similarly,
-IOStack~\cite{gracia:internet16} is providing policy-based provisioning and
-filtering in OpenStack Swift. According to a SNIA white
-paper~\cite{carlson:snia2014}, the primary goal of SDS is to control and
-facilitate flexible and dynamic provisioning of storage resources of different
-kinds, including flash memory and disk drives, to create a virtualized mapping
-between common storage abstractions (e.g. files, objects, and blocks) and
-storage devices taking data service objectives in terms of protection,
-availability, performance, and security into account. A programmable storage
-system exposes internal abstractions so that end users (not necessarily
-operators) can create new services on top of the storage stack. Thus, our
-notion of programmable storage differs from ``software-defined storage'' (SDS)
-in terms of goals and scope, although definitions of SDS are still in flux.
-
-Another view of programmable storage systems is one of tailoring systems
-resources to applications~\cite{arpaci:sosp01}. Related efforts include the
-Exokernel~\cite{engler:sosp95}, SPIN~\cite{bershad:sosp95} and
-Vino~\cite{seltzer:osdi96} projects; the latter two addressed the ability of injecting
-code into the kernel to specialize resource management. Another approach is to
-pass hints between the different layers of the IO stack to bridge the semantic
-gap between applications and
-storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}.
-
-Malacology uses the same Active and Typed Storage module presented in
-DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File Manifolds
-can be implemented with small changes to the Malacology framework, namely
-asynchronous object calls and Lua stubs in the inode, respectively.
-
-\section{Conclusion}
-\label{conclusion-and-future-work}
-
-Programmable storage is a viable method for eliminating duplication of complex
-error-prone software used as workarounds for storage system
-deficiencies. We propose that systems expose their services in a safe way
-allowing application developers to customize system behavior to meet their
-needs while not sacrificing correctness. To illustrate the benefits of this
-approach we presented Malacology, a programmable storage system that
-facilitates the construction of new services by re-purposing existing subsystem
-abstractions of the storage stack.
-
-\oldcommentone{Future work will focus on constructing more interfaces to support a wide
-variety of storage system services that can be configured on-the-fly in existing
-systems. This work is one point along that path to producing general-purpose
-storage systems that can
-target  special-purpose applications.  Ultimately we want to utilize
-declarative methods for expressing new services.}
-
-\bibliography{paper}
-\bibliographystyle{plain}
-%\printbibliography
-
-\end{document}
+storage system.} \addressesissue{3}

### 2017-02-23 paper/8_related_work.tex
paper: cleanup line breaks and break into separate files
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-02-09 paper/paper.tex
@@
-\documentclass[preprint]{sigplanconf-eurosys}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage{pifont}
-\usepackage{color}
-\usepackage{comment}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Uncomment for diff
-\includecomment{notes}
-\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
-\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
-\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
-
-% Uncomment for camera-ready
-%\excludecomment{notes}
-%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
-%\newcommand{\oldcomment}[1]{}
-%\newcommand{\addressesissue}[1]{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\cmark}{\ding{51}}%
-\usepackage[font=footnotesize,labelfont=bf]{caption}
-
-\setlength{\belowcaptionskip}{-10pt}
-
-\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
-\usepackage[normalem]{ulem}
-\usepackage{pstricks}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-\setkeys{Gin}{width=\maxwidth,keepaspectratio}
-
-\usepackage[unicode=true]{hyperref}
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, Peter Alvaro, Shel Finkelstein, Jeff LeFevre},
-            pdftitle={Malacology: A Programmable Storage System},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=blue,
-            linkcolor=black,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-
-\title{Malacology: A Programmable Storage System}
-%\authorinfo{Paper ID: 165}          {Total Pages: 12}
-\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre}
-           {\{msevilla, jayhawk, ivo, cmaltzahn\}@soe.ucsc.edu, \{palvaro, finkelstein, jlefevre\}@ucsc.edu}
-\date{}
-\begin{document}
-
-\maketitle
-
-\begin{notes}
-\textcolor{red}{
-\noindent Issues raised by reviewers:
-\begin{enumerate}
-  \item clarify what was implemented: diagram of interfaces and prose
-  specifying the difference between Malacology and prior work (reviewers B, C,
-  D, E)
-  \item describe what makes a programmable storage system (reviewers C, D)
-  \item organize motivation: group interfaces into categories (e.g., core
-  functionality, features, and performance optimizations) and describe other
-  applications we can build with Malacology (reviewers B, C)
-  \item clarify terminology: service vs. file system metadata, storage vs.
-  file system, dirty vs. clean slate approaches, soft-state caching (reviewers
-  B, C)
-  \item address security, access control, and the safety; specify how it
-  affects composability, performance, and multi-tenancy (reviewer D)
-\end{enumerate}
-}
-\end{notes}
-
-\begin{abstract} Storage systems are caught between the need to evolve data
-processing applications efficiently and quickly, and the increasing velocity
-with which storage device technology evolves. This puts tremendous pressure on
-storage systems to support rapid change both in terms of their interfaces and
-their performance. But adapting storage systems can be difficult because
-unprincipled changes might jeopardize years of code-hardening and performance
-optimization efforts that were necessary for users to entrust their data to the
-storage system. We introduce the concept of a programmable storage system, a
-system that exposes internal services and abstractions with the goal of
-re-using and building new services on top of the storage stack. We present
-Malacology, a prototype implementation of a programmable storage system to
-explore how existing abstractions of common storage system services can be
-leveraged to adapt to the needs of new data processing systems and the
-increasing variety of storage devices. We show how this approach allows a large
-degree of flexibility for storage systems to evolve without sacrificing the
-robustness of their code-hardened subsystems. We illustrate the advantages and
-challenges of programmability by composing existing primitives into two new
-higher-level services: a metadata load balancer and a high-performance
-distributed shared-log. The evaluation demonstrates that our services inherit
-desirable qualities of the back-end storage system, including the ability to
-balance load, efficiently propagate cluster metadata, recover from failure, and
-to navigate trade-offs between latency and throughput using leases.
-\end{abstract}
-
-\begin{figure}[tb]
-\centering
-\includegraphics{figures/overview.png}
-\caption{Scalable storage systems have storage daemons which store data,
-monitor daemons (M) that maintain cluster state, and service-specific daemons
-({\it e.g.}, file system metadata servers). Malacology enables the programmability of
-internal abstractions (bold arrows) to re-use and compose
-existing subsystems.  With Malacology, we add new services, ZLog and
-Mantle, that sit alongside traditional user-facing APIs (file, block,
-object).}\label{fig:overview}
-\end{figure}
-
-\section{Introduction}
-\label{introduction}
-\label{sec:intro}
-
-A storage system implements abstractions designed to persistently store data
-and must exhibit a high level of correctness to prevent data loss.  Storage
-systems have evolved around storage devices that often were orders of magnitude
-slower than CPU and memory, and therefore could dominate overall performance if
-not used carefully. Over the last few decades members of the storage systems
-community have developed clever strategies to meet correctness requirements
-while somewhat hiding the latency of traditional storage
-media~\cite{brewer_disks_2016}. To avoid lock-in by a particular vendor, users
-of storage systems have preferred systems with highly standardized APIs and
-lowest common denominator abstract data types such as blocks of bytes and byte
-stream files~\cite{armbrust_view_2010}.
-
-A number of recent developments have disrupted traditional storage systems.
-First, the falling prices of flash storage and the availability of new types of
-non-volatile memory that are orders of magnitude faster than traditional
-spinning media are moving overall performance bottlenecks away from storage
-devices to CPUs and networking, and pressure storage systems to shorten their
-code paths and incorporate new
-optimizations~\cite{gray_tape_2007,gray_flash_2008}.  Second, emerging ``big
-data'' applications demand interface evolution to support flexible consistency
-as well as flexible structured data 
-representations.~\cite{apache_contributors_parquet_2014}.  Finally, production-quality scalable
-storage systems available as open source software have established and are
-continuing to establish new, \emph{de-facto} API standards at a faster pace
-than traditional standards
-bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
-
-The evolutionary pressure placed on storage systems by these trends raises the
-question of whether there are principles that storage systems designers can
-follow to evolve storage systems efficiently, without jeopardizing years of
-code-hardening and performance optimization efforts.  In this paper we
-investigate an approach that focuses on identifying and exposing existing
-storage system resources, services, and abstractions that in a generalized form
-can be used to \emph{program} new services. By composing higher-level services
-over existing abstractions one can reuse subsystems and leverage their
-optimizations, established correctness, robustness, and efficiency.
-
-{\it \textbf{Contribution 1:}} We define a programmable storage system to be a
-storage system that facilitates the re-use and extension of existing storage
-abstractions provided by the underlying software stack, to enable the creation
-of new services via composition.  A programmable storage system can be realized
-by exposing existing functionality (such as metadata services and
-synchronization and monitoring capabilities) as interfaces that can be ``glued
-together'' in a variety of ways using a high-level language. Programmable
-storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection
-and execution of code within a storage system or storage device---in that the
-former is applicable to any component of the storage system, while the latter
-focuses at the data access level. Given this contrast, we can say that active
-storage is an example of how one internal component (the storage layer) is
-exposed in a programmable storage system.
-
-To illustrate the benefits and challenges of this approach we have designed and
-evaluated Malacology, a programmable storage system that facilitates the
-construction of new services by re-purposing existing subsystem abstractions of
-the storage stack.  We build Malacology in Ceph, a popular open source software
-storage stack.  We choose Ceph to demonstrate the concept of programmable
-storage because it offers a broad spectrum of existing services, including
-distributed locking and caching services provided by metadata servers,
-durability and object interfaces provided by the backend object store, and
-propagation of consistent cluster state provided by the monitoring service (see
-Figure~\ref{fig:overview}).  Malacology is expressive enough to provide the
-functionality necessary for implementing new services.
-
-Malacology includes a non-exhaustive set of interfaces that can be used as
-building blocks for constructing novel storage abstractions, including:
-
-\begin{enumerate}
-
-\item An interface for managing strongly-consistent time-varying \textbf{service
-metadata}.
-
-\item An interface for installing and evolving domain-specific, cluster-wide
-\textbf{data interfaces}.
-
-\item An interface for managing access to \textbf{shared resources} using a
-variety of optimization strategies.
-
-\item An interface for \textbf{load balancing} resources across the cluster.
-
-\item An interface for \textbf{durability} that persists policies using the
-underlying storage stack's object store.
-
-\end{enumerate}
-
-{\it \textbf{Contribution 2:}} We implement two distributed services using Malacology to
-demonstrate the feasibility of the programmable storage approach:
-
-\begin{enumerate}
-
-\item A high-performance distributed shared log service called ZLog, that is an
-implementation of CORFU.~\cite{balakrishnan_corfu_2012}
-
-\item An implementation of Mantle, the programmable load balancing
-service~\cite{sevilla:sc15-mantle}
-
-\end{enumerate}
-
-The remainder of this paper is structured as follows. First, we describe and
-motivate the need for programmable storage by describing current practices in
-the open source software community. Next we describe Malacology by presenting the
-subsystems within the underlying storage system that we re-purpose, and briefly
-describe how those system are used within Malacology (\S\ref{sec:malacology}).
-Then we describe the services that we have constructed within the Malacology
-framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
-implementation (\S\ref{sec:evaluation}).  We conclude by discussing related and future work.
-% and
-%conclude.
-
-\section{Application-Specific Storage Stacks}
-\label{sec:application-specifc-storage-stacks}
-
-Building storage stacks from the ground up for a specific purpose results in
-the best performance. For example, GFS~\cite{ghemawat:sosp03} and
-HDFS~\cite{shvachko:msst10} were designed specifically to serve Map/Reduce and
-Hadoop jobs, and use techniques like exposing data locality and relaxing POSIX
-constraints to achieve application-specific I/O optimizations. Another example
-is Boxwood~\cite{maccormick:osdi04}, which experimented with B-trees
-and chunk stores as storage abstractions to simplify application building.
-Alternatively, general-purpose storage stacks are built with the flexibility to
-serve many applications by providing a variety of interfaces and tunable
-parameters.  Unfortunately, managing competing forces in these systems is
-difficult and users want more control from the general-purpose storage stacks
-without going as far as building their storage system from the ground up.
-
-To demonstrate a recent trend towards more application-specific storage systems
-we examine the state of programmability in Ceph. Something of a storage Swiss
-army knife, Ceph simultaneously supports file, block, and object interfaces on
-a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
-Distributed Object Storage (RADOS) system is a cluster of object storage
-devices (OSDs) that provide Ceph with data durability and integrity using
-replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already
-provides some degree of programmability; the OSDs support domain-specific
-object interfaces that are implemented by composing existing low-level storage
-interfaces that execute atomically. These interfaces are written in C++ and are
-statically loaded into the system.
-
-The Ceph community provides empirical evidence that developers are already
-beginning to embrace programmable storage. Figure~\ref{fig:obj-int-dev-growth}
-shows a dramatic growth in the production use of domain-specific interfaces in
-the Ceph community since 2010. In that figure, classes are functional groupings
-of methods on storage objects (e.g. remotely computing and caching the checksum
-of an object extent).  What is most remarkable is that this trend contradicts
-the notion that API changes are a burden for users.  Rather it appears that
-gaps in existing interfaces are being addressed through ad-hoc approaches to
-programmability. In fact, Table~\ref{table:objclasses} categorizes existing
-interfaces and we clearly see a trend towards reusable services.
-
-\begin{figure}[ht]
-\centering
-\includegraphics{figures/obj-int-dev-growth.png}
-\caption{Since 2010, the growth in the number of co-designed object storage interfaces
-in Ceph has been accelerating. This plot is the number of object classes (a
-group of interfaces), and the total number of methods (the actual API
-end-points).}
-\label{fig:obj-int-dev-growth}
-\end{figure}
-
-\begin{table}[ht]
-\centering
-  \begin{tabular}{l|l|l}
-    Category & Example & \# \\ \hline
-    Logging  & Geographically distribute replicas & 11 \\ \hdashline
-    Metadata & Snapshots in the block device OR  & \multirow{2}{*}{74} \\
-    Management & Scan extents for file system repair & \\ \hdashline
-    Locking  & Grants clients exclusive access & 6 \\ \hdashline
-    Other & Garbage collection, reference counting  & 4\\
-\end{tabular}
-\caption{A variety of object storage classes exist to expose interfaces
-    to applications. \# is the number of methods that implement these categories.
-}
-\label{table:objclasses}
-\end{table}
-
-The takeaway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
-already trying to use programmability because their needs, whether they be
-related to performance, availability, consistency, convenience, etc., are not
-satisfied by the existing default set of interfaces. The popularity of the
-custom object interface facility of Ceph could be due to a number of reasons,
-such as the default algorithms/tunables of the storage system being
-insufficient for the application's performance goals, programmers wanting to
-exploit application-specific semantics, and/or programmers knowing how to
-manage resources to improve performance. A solution based on
-application-specific object interfaces is a way to work around the
-traditionally rigid storage APIs because custom object interfaces give
-programmers the ability to tell the storage system about their application: if
-the application is CPU or IO bound, if it has locality, if its size has the potential to
-overload a single node, etc.  Programmers often know what the problem is
-and how to solve it, but until the ability to modify object interfaces, they
-had no way to express to the storage system how to handle their data.
-
-Our approach is to expose more of the commonly used, code-hardened subsystems
-of the underlying storage system as interfaces. The intent is that these
-interfaces, which can be as simple as a redirection to the persistent data
-store or as complicated as a strongly consistent directory service, should be
-used and re-used in many contexts to implement a wide range of services. By
-making programmability a `feature', rather than a `hack' or `workaround', we
-help standardize a development process that now is largely ad-hoc.
-
-\section{Challenges}
-\label{sec:challenges}
-
-Implementing the infrastructure for programmability into existing services and
-abstractions of distributed storage systems is challenging, even if one assumes
-that the source code of the storage system and the necessary expertise for
-understanding it is available.  Some challenges include:
-
-\begin{itemize}
-
-\item Storage systems are generally required to be highly available so that any
-complete restarts of the storage system to reprogram them is usually
-unacceptable.
-
-\item Policies and optimizations are usually hard-wired into the services and
-one has to be careful when factoring them to avoid introducing additional bugs.
-These policies and optimizations are usually cross-cutting solutions to
-concerns or trade-offs that cannot be fully explored at the time the code is
-written (as they relate to workload or hardware). Given these policies and
-optimizations, decomposition of otherwise orthogonal internal abstractions can
-be difficult or dangerous.
-
-\item Mechanisms that are often only exercised according to hard-wired policies
-and not in their full generality have hidden bugs that are revealed as soon as
-those mechanisms are governed by different policies. In our experience
-introducing programmability into a storage system proved to be a great
-debugging tool.
-
-\item Programmability, especially in live systems, implies changes that need to
-be carefully managed by the system itself, including versioning and propagation
-of those changes without affecting correctness.
-
-\end{itemize}
-
-To address these challenges we present Malacology
-\oldcomment{. Malacology is both a prototype for a programmable storage system
-and a}
-\newcomment{, our prototype programmable storage system. It uses the
-programmable storage}
-\addressesissue{2}
-design approach to evolve storage systems efficiently and
-without jeopardizing years of code-hardening and performance optimization
-efforts.  Although Malacology uses the internal abstractions of the underlying
-storage system, including its subsystems, components, and implementations, we
-emphasize that our system still addresses the general challenges outlined
-above.
-
-The main challenge of designing a programmable storage system is choosing the
-right internal abstractions and picking the correct layers for exposing them.
-\oldcomment{In this paper, we do not present an exhaustive list of possible
-internal abstractions nor do we contend that the abstractions we choose provide
-the best trade-offs for all applications.} 
-\newcomment{A programmable storage is not defined by what abstractions are
-exposed, rather a programmable storage system adheres to the design approach of
-choosing interfaces so administrators can have better control of the storage
-system. The interfaces presented in this paper are abstractions that we found
-useful for building ZLog and Mantle, yet they may not provide the best
-trade-offs for all higher-level services.}
-\addressesissue{2}
-For example, if consensus is correctly exposed one could implement
-high-level features like versioning, serialization, or various flavors of
-strongly consistent data management on top; but perhaps a low-level consensus
-interface is suited well for a particular set of applications.  These questions
-are not answered in this paper and instead we focus on showing the feasibility
-of building such a system, given advances in the quality and robustness of
-today's storage stacks.
-
-The Malacology prototype we present has been implemented on Ceph.
-While there are other systems on top of which Malacology could be implemented (see Table~\ref{table:examples}), we choose Ceph because it is a production quality system and because it
-is open source. The large developer community ensures that code is
-robust and the visibility of the code lets us expose any interface we want. In
-the next section we describe the Ceph components that we expose as Malacology
-interfaces.
-
-\section{Malacology: A Programmable Storage System}
-\label{sec:malacology}
-
-\begin{table*}
-\centering\small
-\begin{tabular}{ l | c | l | l | l }
-%\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
-\multicolumn{4}{c}{} \\
-\textbf{Interface}                      &
-\textbf{Sec}                            &
-\textbf{Example in Production Systems}  &
-\textbf{Example in Ceph}                &
-\textbf{Provided Primitives}            \\ \hline
-Service Metadata
-  & \S\ref{sec:service-metadata-interface}
-  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
-  & cluster state management~\cite{website:ceph-mon}
-  & consensus/consistency
-  \\
-Data I/O
-  & \S\ref{sec:data-io-interface}
-  & Swift in situ storage/compute~\cite{website:zerocloud}
-  & object interface classes~\cite{website:cls-lua}
-  & transaction/atomicity
-  \\
-Shared Resource
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & MPI collective IO, burst
-  & POSIX metadata protocols
-  & serialization/batching
-  \\
-File Type
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & ?????
-  & file striping strategy
-  & data/metadata access
-  \\
-Load Balancing
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
-  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
-  & migration/sampling
-  \\
-Durability
-  & \S\ref{sec:durability}
-  & S3/Swift interfaces (RESTful API)
-  & object store library~\cite{weil_rados_2007}
-  & persistence/safety
-  \\
-\end{tabular}
-\caption{Common internal abstractions. The same ``internal abstractions" are
-common across large-scale systems because they provide primitives that solve
-general distributed systems problems.  Here we list examples of what these
-internal abstractions are used for in ``production systems" and in Ceph.
-Malacology provides these internal abstractions as interfaces
-\newcomment{(Section~\ref{sec:malacology})}
-that higher level
-\oldcomment{applications}
-\newcomment{services (Section~\ref{sec:services})}
-can use.
-\oldcomment{; {\it italicized abstractions} are interfaces discussed in this
-paper.}}
-\addressesissue{1}
-\label{table:examples}
-\end{table*}
-
-The guiding principle is to re-use existing services and extend them so that
-these services can be \emph{programmed}. We accomplish programmability of a
-service by exporting bindings (or ``hooks'') for an interpreted programming
-language so that programming can occur without having to restart the storage
-system (see also below,~\S\ref{sec:durability}). Table~\ref{table:examples}
-shows the internal services from Ceph that we expose in the Malacology
-prototype via Lua~\cite{ierusalimschy1996lua} bindings
-\oldcomment{.}
-\newcomment{and Figure~\ref{fig:implementation-overview} compares what was
-already present in Ceph (gray boxes) to the Malacology interfaces we added.
-Section~\ref{sec:services} will describe the higher-level services we built
-with these interfaces.}
-\addressesissue{1}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation-overview.png}
-\caption{\newcomment{Malacology is implemented on the daemons and clients that
-run in a Ceph cluster. Interfaces expose internal subsystems and are used as
-building blocks for higher-level services.}\addressesissue{1}
-\label{fig:implementation-overview}}
-\end{figure}
-
-Lua is a portable embedded scripting language and we choose it as the
-interpreted language for Malacology because it offers superior performance and
-productivity trade-offs, including a JIT-based implementation that is well
-known for near native performance. Additionally, Lua has been used extensively
-in game engines, and systems research \cite{neto:dls14-luaos}, including
-storage systems where it has been effectively used both on
-\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010}
-and off \cite{sevilla:sc15-mantle} the performance critical path. Finally, the
-flexibility of the runtime allows execution sandboxing in order to address
-security and performance concerns. We will now discuss the common subsystems
-used to manage storage system and how Malacology makes them programmable.
-
-\subsection{Service Metadata Interface}
-\label{sec:mon}
-\label{sec:service-metadata-interface}
-\label{service-metadata}
-
-% straw man example
-Keeping track of state in a distributed system is an essential part of any
-successful service and a necessary component in order to diagnose and detect
-failures, when they occur. This is further complicated by variable propagation
-delays and heterogeneous hardware in dynamic environments.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In the case of Ceph, a consistent view of cluster state among server daemons
-and clients is critical to provide strong consistency guarantees to clients.
-Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
-metadata server) data structures called ``maps'' that record membership and
-status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
-responsible for integrating state changes into cluster maps, responding to
-requests from out-of-date clients and synchronizing members of the cluster
-whenever there is a change in a map so that they all observe the same system
-state. As a fundamental building block of many system designs, consensus
-abstractions such as Paxos are a common technique for maintaining consistent
-data versions, and are a useful system to expose.
-
-The default behavior of the monitor can be seen as a Paxos-based notification
-system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing
-clients to identify when new values (termed epochs in Ceph) are associated to
-given maps.  Since Ceph does not expose this service directly, as part of our Malacology implementation, we expose a key-value
-service designed for managing configuration metadata that is built on top of
-the consensus engine. Since the monitor is intended to be out of the
-high-performance I/O path, a general guideline is to make use of this
-functionality infrequently and to assign small values to maps.\\
-
-\noindent{\it{\textbf{Malacology: }}} Malacology exposes a strongly-consistent view of
-time-varying service metadata as a service rather than a hidden internal
-component. Malacology provides a generic API for adding arbitrary values to
-existing subsystem cluster maps. As a consequence of this, applications can
-define simple but useful service-specific logic to the strongly-consistent
-interface, such as authorization control (just specific clients can write new
-values) or triggering actions based on specific values (e.g. sanitize values).
-The higher level services we implement in \S\ref{sec:services} make use of this
-functionality to register, version and propagate dynamic code (Lua scripts) for
-new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
-policies in the load balancer \S\ref{malacology:mds}.  Using this service
-guarantees that interface definitions are not only made durable, but are
-transparently and consistently propagated throughout the cluster so that
-clients are properly synchronized with the latest interfaces.
-
-\newcomment{\\ \noindent \it{\textbf{Impact: }} provides core functionality
-because it lets daemons come to consensus on system critical state. Bugs in the
-internal subsystems or ommitting this from services that need this type of
-consistency affects correctness.}
-\addressesissue{3, 5}
-
-%And finally, it is important to not underestimate the importance of the
-%preservation of the modifications that enable a new system service. In many
-%(if not all) cases, interfaces defining access to data are just as important
-%as the data itself by virtue of inherently providing structural context. Thus,
-%all components of a system service must be kept to the same standards of
-%protection as data in the system itself.
-
-\oldcomment{\noindent 4.2 Object I/O Intefrace}
-\subsection{Data I/O Interface}
-\label{sec:data-io-interface}
-\label{object-data-interface}
-Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
-supports application-specific object
-interfaces~\cite{weil_rados_2007}. The ability to offload computation can
-reduce data movement, and transactional interfaces can significantly simplify
-construction of complex storage interfaces that require uncoordinated parallel
-access.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-An object interface is a plugin structured similarly to that of an RPC
-in which a developer creates a named function within the cluster that clients
-may invoke. In the case of Ceph each function is implicitly run within the
-context of an object specified when the function is called. Developers of
-object interfaces express behavior by creating a composition of native
-interfaces or other custom object interfaces, and handle serialization of
-function input and output.  A wide range of native interfaces are available to
-developers such as reading and writing to a byte stream, controlling object
-snapshots and clones, and accessing a sorted key-value database. These native
-interfaces may be transactionally composed along with application
-specific logic to create semantically rich interfaces. An example would be
-an interface that atomically updates a matrix stored in the bytestream and an
-index of the matrix stored in the key-value database.\\
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation.png}
-\caption{Malacology allows users to dynamically define object/metadata
-interfaces by extending the object storage daemon (OSD) and metadata server
-(MDS) subsystems with an embedded Lua VM.  It uses the service metadata
-interface (M) to propagate interfaces/versions across the cluster
-\label{fig:implementation}}
-\end{figure}
-
-\noindent {\it{\textbf{Malacology: }}} The implementation of Ceph's object abstraction,
-although powerful, does not readily support programmability. Supporting only
-C/C++ for object interface developers, Ceph requires distribution of compiled
-binaries for the correct architecture, adding a large barrier of entry for
-developers and system administrators. Second, having no way to dynamically
-unload modules, any changes require a full restart of a storage daemon which
-may have serious performance impacts due to loss of cached data. And finally,
-the security limitations of the framework limit the use of object interfaces to
-all but those with administrative level access and deep technical expertise.
-
-To address these concerns, Malacology takes advantage of Lua extensions
-contributed by the Ceph community. This allows new object interfaces to be
-dynamically loaded into the system and modified at runtime, resulting in a
-object storage API with economy of expression, which at the same time provides
-the full set of features of the base object class. New object interfaces that
-are expressed in thousands of lines of code can be implemented in approximately
-an order of magnitude less code~\cite{geambasu_comet_2010}. While the use of
-Lua does not prevent deployment of malicious code, certain types of coding
-mistakes can be handled gracefully, and access policies are used to limit access
-to trusted users~\cite{ierusalimschy1996lua}.
-
-\oldcomment{All of these modifications work in tandem to implement the desired behavior,
-and are typically co-designed together such that each depends on the other to
-behave as expected.}
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by pushing behavior to lower parts of the storage stack, thereby
-minimizing hops and distributing computation.}
-\addressesissue{3, 5}
-
-\subsection{Distributed Metadata Interfaces}
-\begin{notes}
-\textcolor{red}{This section has been re-organized}
-\end{notes}
-\label{sec:distributed-metadata-interfaces}
-\label{malacology:mds}
-
-\oldcomment{The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction~\cite{weil:sc2004-dyn-metadata}.}
-\newcomment{File systems provide clients with the familiar POSIX file abstraction.
-While this guarantees strong consistency it comes at the cost of scalability,
-increased complexity, and lower performance.}
-\addressesissue{1}
-In general, distributed file systems protect resources by providing
-hierarchical indexing and distributed locking services.
-
-%The metadata cluster moves units of the namespace called directory fragments
-%and assigns them to servers using its own metadata load balancer. It then uses
-%a hard-coded policy to balance these fragments using a scalarized load metric
-%based on CPU, workload, and file system operation metrics. Although the CephFS
-%balancer has been shown to be complicated, its complexity is justifiable.
-%Having the flexibility to choose what to move, where to move it, and how much
-%to move is very powerful. The problem is that it is difficult to pick a
-%one-size-fits-all policy, system administrators may want to partition load
-%based on SLAs, or applications are better equipped to make these decisions,
-%etc. -- but we choose instead to expose a balancing API at strategic points in
-%the balancing logic using Lua hooks. ~\cite{sevilla:sc15-mantle}.
-
-\oldcomment{\it{\textbf{Malacology:}} Interfaces are added in strategic spots in the
-metadata service for guarding shared resources, defining new file types,
-and load balancing.}
-
-\oldcomment{\noindent Shared resource interface.}
-\subsubsection{Shared Resource Interface}
-
-In general, file system metadata servers manages client sessions, allowing
-clients to obtain locks (e.g. file byte ranges), and capabilities (e.g. to
-cache file data). Clients and metadata servers use a cooperative protocol in
-which clients voluntarily release resources back to the metadata service in
-order to implement sharing policies.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In Ceph, the locking service implements a capability-based system that
-expresses what data and metadata clients are allowed to access as well as what
-state they may cache and modify locally.  While designed for the file
-abstraction, indexing, locking, and caching are all common services that are
-useful to a broad spectrum of applications.  Distributed applications that
-share centralized resources (e.g. a database or directory) face similar
-challenges which are often solved using application-specific sharding.
-
-\newcomment{\\ \noindent\it{\textbf{Malacology: }}}
-While the current policy for sharing access and voluntarily releasing resources
-is largely best-effort, Malacology supports generalized policies between
-metadata servers and clients that can be used to implement fairness or
-priority.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} provides core functionality
-to protect and provide exclusive access for any shared resource. May hurt
-performance if the resource in question does not require strong consistency.}
-\addressesissue{3}
-
-\oldcomment{\noindent File type interface.}
-\subsubsection{File Type Interface}
-The metadata service manages a POSIX file system hierarchy where files and
-directories are represented as inode data structures and expose a POSIX file
-interface. Applications that manage large amounts of metadata (e.g. users or
-database snapshots) often require a naming service. We allow new inode types to
-be defined such that applications can create domain-specific interfaces to
-inodes that may modify locking and capability policies. We will show how this
-is used in Section~\ref{sec:seq} when we discuss a distributed shared-log built
-on Malacology.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this interface is both a feature
-and a performance optimization. It is a feature because it allows developers to
-add support for different  storage types, such as how to read new file formats
-or what consistency semantics to use for a specific subtree in the hierarchical
-namespace. It is also a performance optimization because future programmers can
-add optimizations for processing specific types of files into the inode itself.}
-\addressesissue{3}
-
-%%% TODO: striping strategy is stored with the inode
-
-\oldcomment{\noindent \bf Load balancing interface.} 
-\subsubsection{Load Balancing Interface}
-
-Ceph also addresses the challenge of balancing metadata load with a separate
-metadata cluster. This cluster uses load balancing policies to migrate
-directory inodes around the cluster to alleviate overload at single
-nodes~\cite{weil:sc2004-dyn-metadata}. The policies use metrics based on system
-state (e.g.  CPU and memory utilization) and statistics collected by the
-cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
-partitioning to move variable sized namespace subtrees. These units can be
-shipped anywhere (i.e., to any metadata server of any capacity) at any time for
-any reason. The original balancer was designed with hard-coded policies and
-tunables.\\
-
-\noindent {\it{\textbf{Malacology: }}}
-Mantle~\cite{sevilla:sc15-mantle}, a programmable metadata load balancer, has
-been implemented on Malacology so that its infrastructure can enjoy all the
-properties of existing internal abstractions.  The load balancing logic is
-intact but the infrastructure has been improved to safely load, manage, and
-persist its load balancing policies.  Composing Mantle with the service
-metadata and durability interfaces, Mantle can safely version balancer
-policies, save balancer policies in the backend object store and centralize
-warnings/errors. When combined with file types, load balancing policies can
-express a variety of rules such as multi-tenant and multi-application
-balancing.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by allowing them to specify how to partition, replicate, and
-distribute metadata in response to overloaded servers.}
-\addressesissue{3}
-
-\subsection{Durability Interface}
-\label{sec:durability}
-
-Ceph provides storage by striping and replicating data across
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-protects data using common techniques such as erasure
-coding, replication, and scrubbing.  Furthermore, many of these features
-are implemented using a peer-to-peer protocol that allows object storage
-devices to operate autonomously without a centralized coordinator.  For
-example, when the number of placement groups change, the OSDs re-balance and
-re-shard data in the background in a process called placement group splitting
-in which OSDs communicate directly with each other to cover a new state.
-In order to reduce load on the monitoring service, Ceph OSDs use a gossip
-protocol to efficiently propagate changes to cluster maps throughout the
-system, and autonomously initiate recovery mechanisms when failures are
-discovered.\\
-
-\noindent {\it{\textbf{Malacology: }}} Metadata service policies and object storage
-interfaces are stored durability within RADOS and managed by storing references
-with the object maps. Since the cluster already propagates a consistent view of
-these data structures, we use this service to automatically install interfaces
-in OSDs, and install policies within the metadata server daemon (MDS) such that clients and daemons are
-synchronized on correct implementations without restarting.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this is a feature because it
-adds data safety and persistence to system metadata; while nice to have it does
-effect correctness.}
-\addressesissue{3}
-
-\section{Services Built on Malacology}
-\label{sec:services}
-\label{services-built-on-malacology}
-
-\label{services}
-%\begin{table}
-%\centering\small
-%\begin{tabular}{  l | l | l }
-%\textbf{Service}              &
-%\textbf{Used to...}           &
-%\textbf{Malacology Interface} \\ \hline
-%Mantle & load balance metadata     & Load Balancing  \\
-%       & agree on current balancer & Service Metadata\\
-%       & control client write lock & Shared Resource \\
-%       & persist policy logic      & Durability      \\ \hline
-%ZLog   & implement CORFU           & Object \& Data  \\ 
-%       & install ZLog interfaces   & Service Metadata\\
-%       & persist log entries       & Durability      \\ \hline
-%ZLog   & load balance sequencers   & Load Balancing  \\
-%Seqr.  & manage access to log tail & Shared Resource \\ 
-%\end{tabular}
-%\caption{Mantle and ZLog are built using Malacology interfaces so they inherit the
-%robustness of the subsystems of the underlying stack.}
-%\label{table:implementation}
-%\end{table}
-
-In this section we describe two services built on top of Malacology. The first
-is Mantle, a framework for dynamically specifying metadata load balancing
-policies. The second system, ZLog, is a high-performance distributed shared-log.
-In addition to these services, we'll demonstrate how we combine ZLog and Mantle
-to implement service-aware metadata load balancing policies.
-
-\subsection{Mantle: Programmable Load Balancer}
-\label{sec:mantle}
-
-Mantle is a programmable load balancer that separates the metadata balancing
-policies from their mechanisms. Administrators inject code to change how the
-metadata cluster distributes metadata. In ~\cite{sevilla:sc15-mantle} the
-authors showed how using Mantle, one can implement a single node metadata
-service, a distributed metadata service with hashing, and a distributed
-metadata service with dynamic subtree partitioning.
-
-The original implementation was ``hard-coded" into Ceph and lacked robustness
-(no versioning, durability, or policy distribution).  Re-implemented using
-Malacology, Mantle now enjoys (1) the versioning provided by the metadata
-service interface provided by the monitor daemons and (2) the durability and
-distribution provided by Ceph's reliable object store.  Re-using the internal
-abstractions with Malacology resulted in a 2\(\times\) reduction in source code
-compared to the original implementation.
-
-\subsubsection{Versioning Balancer Policies}
-
-Ensuring that the version of the current load balancer is consistent across the
-physical servers in the metadata cluster was not addressed in the original
-implementation. The user had to set the version on each individual server and
-it was trivial to make the versions inconsistent. Maintaining consistent
-versions is important for balancing policies that are cooperative in which local
-decisions are made assuming properties about other instances of the same
-policy.
-
-With Malacology, Mantle stores the version of the current load balancer in the
-Ceph service metadata interface. The version of the load balancer corresponds
-to an object name in the balancing policy. Using the service metadata interface
-ensures that all metadata servers use the same version of the load balancer. As
-a result, the policy version inherits the consistency of Ceph's internal
-monitor daemons. The user changes the version of the load balancer using a new
-CLI command.
-
-\subsubsection{Making Balancer Policies Durable}
-
-The load balancer version described above corresponds to the name of an object
-in RADOS that holds the actual Lua balancing code.  When MDS nodes start
-balancing load, they first check the latest version from the MDS map and
-compare it to the balancer they have loaded. If the version has changed, they
-dereference the pointer to the balancer version by reading the corresponding
-object in RADOS. This is in contrast to the original balancer which stored load
-balancer code on the local file system -- a technique which is unreliable and
-may result in silent corruption.
-
-The balancer pulls the Lua code from RADOS synchronously; asynchronous reads
-are not possible because of the architecture of the MDS. The synchronous
-behavior is not the default behavior for RADOS operations, so we achieve this
-with a timeout: if the asynchronous read does not come back within half the
-balancing tick interval the operation is canceled and a Connection Timeout
-error is returned. By default, the balancing tick interval is 10 seconds, so
-Mantle will use a 5 second second timeout.
-
-This design allows Mantle to immediately return an error if anything
-RADOS-related goes wrong.  We use this implementation because we do not want to
-do a blocking OSD read from inside the global MDS lock. Doing so would bring
-down the MDS cluster if any of the OSDs are not responsive.
-
-Storing the balancers in RADOS is simplified by the use of an interpreted
-language for writing balancer code. If we used a language that needs to be
-compiled, like the C++ object classes in the OSD, we would need to ensure
-binary compatibility, which is complicated by different operating systems,
-distributions, and compilers.
-
-\subsubsection{Logging, Debugging, and Warnings}
-
-In the original implementation, Mantle would log all errors, warnings, and
-debug messages to a log stored locally on each MDS server. To get the simplest
-status messages or to debug problems, the user would have to log into each MDS
-individually, look at the logs, and reason about causality and ordering.
-
-With Malacology, Mantle re-uses the centralized logging features of the
-monitoring service. Important errors, warnings, and info messages are collected
-by the monitoring subsystem and appear in the monitor cluster log so instead of
-users going to each node, they can watch messages appear at the monitor.
-Messages are logged sparingly, so as not to spam the monitor with frivolous
-debugging but important events, like balancer version changes or failed
-subsystems show up in the centralized log.
-
-\subsection{ZLog: A Fast Distributed Shared Log}
-\label{sec:zlog}
-
-The second service implemented on Malacology is ZLog, a high-performance
-distributed shared-log that is based on the CORFU
-protocol~\cite{balakrishnan_corfu_2012}. The shared-log is a
-powerful abstraction used to construct distributed systems, such as
-metadata management~\cite{balakrishnan:sosp13} and elastic database
-systems~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
-However, existing implementations that rely on consensus algorithms such as
-Paxos funnel I/O through a single point introducing a bottleneck that restricts
-throughput. In contrast, the CORFU protocol is able to achieve high throughput
-using a network counter called a \emph{sequencer}, that
-decouples log position assignment from log I/O.
-
-While a full description of the CORFU system is beyond the scope of this
-paper, we briefly describe the custom storage device interface, sequencer
-service, and recovery protocol, and how these services are instantiated in the
-Malacology framework.
-
-\subsubsection{Sequencer}
-\label{sec:seq}
-
-High-performance in CORFU is achieved using a sequencer service that assigns
-log positions to clients by reading from a volatile, in-memory counter which
-can run at a very high throughput and at low latency. Since the sequencer is
-centralized, ensuring serializability in the common case is trivial.  The
-primary challenge in CORFU is handling the failure of the sequencer in a way
-that preserves correctness. Failure of the sequencer service in CORFU is
-handled by a recovery algorithm that recomputes the new sequencer state using a
-CORFU-specific custom storage interface to discover the tail of the log, while
-simultaneously invalidating stale client requests using an epoch-based
-protocol.
-
-{\bf Sequencer interface.} The sequencer resource supports the ability to
-\emph{read()} the current tail value and get the \emph{next()} position in
-the log which also atomically increments the tail position.
-We implement the sequencer service in Malacology as an application-specific
-file type that associates a small amount of metadata
-with the inode state managed by the Ceph metadata service.
-This approach has the added
-benefit of allowing the metadata service to handle naming, by
-representing each sequencer instance in the standard POSIX hierarchical
-namespace. The primary challenge in mapping the sequencer resource to the
-metadata service is handling serialization correctly to maintain the global
-ordering provided by the CORFU protocol.
-
-Initially we sought to directly model the sequencer service in Ceph as a
-non-exclusive, non-cacheable resource, forcing clients to perform a round-trip
-access to the resource at the authoritative metadata server for the sequencer
-inode.  Interestingly, we found that the capability system in Ceph uses a
-strategy to reduce metadata service load by allowing clients that open a
-shared file to temporarily obtain an exclusive cached copy of the resource,
-resulting in a round-robin, best-effort batching behavior. When a single
-client is accessing the sequencer resource it is able to increment the
-sequencer locally. Any competing client cannot query the sequencer until the
-metadata service has granted it access.
-
-While unexpected, this discovery allowed us to explore an implementation
-strategy that we had not previously considered. In particular, for bursty
-clients, and clients that can tolerate increased latency, this mode of
-operation may allow a system to achieve much higher throughput than a system
-with a centralized sequencer service.  We utilize the programmability of the
-metadata service to define a new policy for handling capabilities that controls
-the amount of time that clients are able to cache the sequencer resource. This
-allows an administrator or application to control the trade-off between latency
-and throughput beyond the standard best-effort policy that is present in Ceph
-by default.
-
-In Section~\ref{sec:evaluation} we quantify the trade-offs of throughput and
-latency for an approach based on a round-robin batching mode, and compare this
-mode to one in which the metadata server mediates access to the sequencer state
-when it is being shared among multiple clients. Quantifying these trade-offs
-should provide administrators with guidelines for setting the tunables for
-different ``caching" modes of the sequencer.
-
-{\bf Balancing policies.} As opposed to the batching mode for controlling access to the sequencer
-resource, more predictable latency can be achieved by treating the sequencer
-inode as a shared non-cacheable resource, forcing clients to make a round-trip
-to the metadata service. However, the shared nature of the metadata service
-may prevent the sequencer from achieving maximum throughput. To provide a
-solution to this issue we have taken advantage of the programmability of the
-metadata service load balancing to construct a service-specific load balancing
-policy. As opposed to a balancing policy that strives for uniform
-load distribution, a ZLog-specific policy may utilize knowledge of inode types
-to migrate the sequencer service to provisioned hardware during periods of
-contention or high demand.
-
-\subsubsection{Storage Interface}
-
-The storage interface is a critical component in the CORFU protocol. Clients
-independently map log positions that they have obtained from the sequencer
-service (described in detail in the next section) onto storage devices, while
-storage devices provide an intelligent \emph{write}-once, random \emph{read}
-interface for accessing log entries. The key to correctness in CORFU lies with
-the enforcement of up-to-date epoch tags on client requests; requests tagged
-with out-of-date epoch values are rejected, and clients are expected to
-request a new tail from the sequencer after refreshing state from an auxiliary
-service.  This mechanism forms the basis for sequencer recovery.
-
-In order to repopulate the sequencer state (i.e. the cached, current tail of
-the log) during recovery of a sequencer, the maximum position in the log must
-be obtained. To do this, the storage interface exposes an additional
-\emph{seal} method that atomically installs a new epoch value and returns the
-maximum log position that has been written.
-
-Since the sequencer service does not resume until the recovery process has
-completed, there cannot be a race with clients appending to the log, and the
-immutability of the log allow reads to never block during a sequencer failure.
-Recovery of the sequencer process itself may be handled in many ways, such as
-leader election using an auxiliary service such as Paxos. In our
-implementation, the recovery is the same as (and is inherited from) the Ceph
-metadata service. Handling the failure of a client that holds the sequencer
-state is similar, although a timeout is used to determine when a client should
-be considered unavailable.
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-%We wanted to implement CORFU and we had two choices: build it from the ground
-%up or do it on top of Malacology. To show the feasability of building on
-%Malacology our evaluation focuses on the
-
-Our evaluation demonstrates the feasibility of building new service
-abstractions atop programmable storage, focusing on the performance of the
-internal abstractions exposed by Malacology and used to construct the Mantle
-and ZLog services. We also discuss latent capabilities we discovered in this
-process that let us navigate different trade-offs within the services
-themselves. First, we benchmark scenarios with high sequencer contention by
-examining the interfaces used to map ZLog onto Malacology; specifically, we
-describe the sequencer implementation and the propagation of object and data
-interfaces interfaces.  Next, we benchmark scenarios in which the storage
-system manages multiple logs by using Mantle to balance sequencers across a
-cluster.
-
-% should save this for future work, otherwise it undercuts the eval.
-%Detailed system-wide performance measurements and analysis is forthcoming for
-%both Mantle and ZLog, as we try to find the best metadata balancing policies
-%and physical design implementation trade-offs.
-
-Since this work focuses on the programmability of Malacology, the goal of this
-section is to show that the components and subsystems that support the
-Malacology interfaces provide reasonable relative performance, as well as to
-give examples of the flexibility that Malacology provides to programmers.
-
-%As we search for the best metadata balancing policies
-%and physical design implementation trade-offs,
-%we plan to rel
-
-\subsection{Mapping ZLog onto Malacology}
-\label{sec:mapping-zlog-onto-malacology}
-
-We evaluate Malacology by exploring one possible mapping of the ZLog
-implementation of CORFU onto Ceph in which we re-use (1) the metadata service
-to manage naming and synchronization of the sequencer resource by treating the
-resource as an inode, and (2) the monitoring sub-subsystem to distribute and
-install application-specific I/O interface required of the CORFU protocol. In
-Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the inode
-abstraction for implementing the sequencer resource enables load balancing policies
-to migrate the sequencer resource in heavy-load situation.
-
-\subsubsection{Sequencer Implementation}
-\label{sec:sequencer-implementation}
-
-We evaluate the feasibility of using the metadata service to implement a
-sequencer resource that is responsible for maintaining a total ordering of
-the log. Clients contact the sequencer to obtain the tail of the log and then
-independently initiate I/O, thus we measure both the throughput and latency of
-obtaining new tail positions which bounds client append performance.
-
-The sequencer is implemented as a custom file type in which the sequencer
-state (a 64-bit integer) is embedded in the inode of a file representing the
-sequencer. A total ordering of the log is imposed by the re-use of the
-capability service that can be used to grant exclusive access of inode state
-to clients. The metadata service is responsible for maintaining exclusivity
-and granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
-behavior of the system in which a best-effort policy is used.  The two colors
-represent points in time that the clients were able to access the resource.
-The best-effort policy shows a high degree of interleaving
-between clients but the system spends a large portion of time re-distributing
-the capability, reducing overall throughput.
-
-In order to control the performance of the system we implement a
-policy that (1) restricts the length of time that a client may maintain
-exclusive access and (2) limits the number of log positions that a client may
-generate without yielding to other clients waiting for access. The behavior of
-these two modes is illustrated in Figures~\ref{fig:capdelay-quota-behavior}
-(b) and (c), respectively.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/capdelay-quota-behavior-a.png}
-\includegraphics{figures/capdelay-quota-behavior-b.png}
-\includegraphics{figures/capdelay-quota-behavior-c.png}
-\caption{Each dot is an individual request, spread randomly along the \(y\)
-axis. The default behavior is unpredictable, ``delay" lets clients hold the
-lease longer, and ``quota" gives clients the lease for a number of operations.}
-\label{fig:capdelay-quota-behavior}
-\end{figure}
-
-Figure~\ref{fig:captp} demonstrates a configurable trade-off between
-throughput and latency. In the experiment two clients are run each with a
-fixed 0.25 second maximum reservation on the capability, and we vary the size
-of the log position quota running each configuration for two minutes. The
-total operations per second is the combined throughput of the two clients, and
-the average latency is number of microseconds required to obtain a new log
-position. With a small quota more time is spent exchanging exclusive access,
-while a large quota reservation allows clients to experience a much lower
-latency because they experience uncontended access for a longer period of time.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/tradeoff.png}
-\caption{Sequencer throughput by re-using various services.  The highest performance is
-achieved using a single client with exclusive, cacheable privilege. Round-robin
-sharing of the sequencer resource is affected by the amount of time the
-resource is held, with best-effort performing the worst.}
-\label{fig:captp}
-\end{figure}
-
-To get a better picture of latency, Figure~\ref{fig:capcdf} shows the CDF of
-latency for each client in all experiment configurations. At the 99th
-percentile clients accessed the sequencer is less than a millisecond. The CDF
-is cropped at the 99.999th percentile due to large outliers that we believe
-occur in instances in which the MDS is performing I/O while it is in the
-process of re-distributing the capability to another client.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/caps-delay-latency.png}
-\caption{
-The latency distribution shows that if the clients hold their capabilities longer,
-performance decreases. Changing the delay can help the sequencer and
-clients strike a balance between throughput and latency.}
-\label{fig:capcdf}
-\end{figure}
-
-\textbf{Malacology exposes} the internal capability management service and
-allows users to navigate latency and throughput trade-offs.  Other approaches to
-designing the sequencer service also exist, such as using a centralized service
-in which each access is a network round-trip. In contrast to the mechanism we
-explored which is appropriate for clients with bursty workloads, it may be
-easier to provide predictable performance using a centralized service, and we
-will be exploring in the future how this can be achieved using the capability
-system.
-
-\subsubsection{Interface Propagation}
-
-Domain-specific data interfaces (Section~\ref{sec:application-specifc-storage-stacks}) allow
-co-design between applications and the storage system. Malacology supports
-custom object interfaces in RADOS that require interface implementations to
-be installed on the storage devices in the system, supporting the evolution of
-interfaces through automatic system-wide versioning and installation through
-the service metadata interface (Section~\ref{service-metadata}). We evaluate
-the performance of installing a new interface version in the cluster,
-which is an important metric for applications that frequently evolve
-interfaces.
-
-We demonstrate the feasibility of utilizing the Ceph monitoring sub-system by
-evaluating the performance of installing and distributing interface updates.
-Figure~\ref{fig:propdelay} shows the CDF of the latency of interface updates.
-The interfaces are Lua scripts embedded in the cluster map and distributed
-using a peer-to-peer gossip protocol.  The latency is defined as the elapsed
-time following the Paxos proposal for an interface update until each OSD makes
-the update live (the cost of the Paxos proposal is configurable and is
-discussed below). The latency measurements were taken on the nodes running
-Ceph server daemons, and thus exclude the client round-trip cost. In each of
-the experiments 1000 interface updates were observed.
-
-Figure~\ref{fig:propdelay} shows the lower bound cost for updates in a large
-cluster. In the experiment labeled ``120 OSD (RAM)" a cluster of 120 OSDs (10
-OSDs x 12 servers) using an in-memory data store were deployed, showing a
-latency of less than 54 ms with a probability of 90\% and a worst case latency
-of 194 ms. These costs demonstrate the penalty of distributing the interface
-in a large cluster. In practice the costs include, in addition to cluster-wide
-propagation of interface updates, the network round-trip to the
-interface management service, the Paxos commit protocol itself, and other factors
-such as system load. By default Paxos proposals occur periodically with a 1
-second interval in order to accumulate updates. In a minimum, realistic quorum
-of 3 monitors using HDD-based storage, we were able to decrease this interval
-to an average of 222 ms.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=60mm,trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
-\caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
-interface commit.}
-\label{fig:propdelay}
-\end{figure}
-
-\subsection{Load Balancing ZLog Sequencers with Mantle}
-\label{sec:zlog-balancing}
-
-In practice, a storage system implementing CORFU will support a multiplicity of
-independent totally-ordered logs for each application.  For this scenario
-co-locating sequencers on the same physical node is not ideal but building a load
-balancer that can migrate the shared resource ({\it e.g.,}, the resource that
-mediates access to the tail of the log) is a time-consuming, non-trivial task.
-It requires building subsystems for migrating resources, monitoring the
-workloads, collecting metrics that describe the utilization on the physical
-nodes, partitioning resources, maintaining cache coherence, and managing
-multiple sequencers. The following experiments demonstrate the feasibility of
-using the mechanisms of the Malacology load balancing interface to inherit
-these features and to alleviate load from overloaded servers.
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-behaviors.png}
-\caption{CephFS/Mantle load balancing have better throughput than co-locating
-all sequencers on the same server.  Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify this improvement;
-Section~\ref{sec:feature-backoff} examines the migration at 0-60 seconds.
-}\label{fig:mantle-balancer-behaviors}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-performance.png}
-\includegraphics{figures/mantle-mode-performance.png}
-\caption{In (a) all CephFS balancing modes have the same performance; Mantle uses a
-balancer designed for sequencers. In (b) the best combination of mode and
-migration units can have up to a 2\(\times\) improvement.}\label{fig:mantle-balancer-performance}
-\end{figure}
-
-The experiments are run on a cluster with 10 nodes to store objects, one node
-to monitor the cluster, and 3 nodes that can accommodate sequencers. 
-Instead of measuring contention at the clients like
-Section~\ref{sec:sequencer-implementation}, these experiments measure
-contention at the sequencers by forcing clients to make round-trips for every
-request. We implement this using the shared resource interface that forces
-round-trips.  Because the sequencer's only function is to  hand out positions
-for the tail of the log, the workload is read-heavy.
-
-First, we show how the ZLog service can orchestrate multiple sequencers using
-the Malacology load balancing interface.
-Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput over time of
-different load balancers as they migrate 3 sequencers (with 4 clients) around
-the cluster; ``No Balancing" keeps all sequencers on one server, ``CephFS"
-migrates sequencers using the CephFS load balancers, and ``Mantle" uses a
-custom load balancer we wrote specifically for sequencers. The CephFS load
-balancers are hard-coded into Ceph while Mantle is an API for balancing load
-presented in~Mantle~\cite{sevilla:sc15-mantle} and re-implemented on top of
-Malacology in this paper.  The increased throughput for the CephFS and Mantle
-curves between 0 and 60 seconds are a result of migrating the sequencer(s) off
-overloaded servers.
-
-In addition to showing that migrating sequencers improves performance,
-Figure~\ref{fig:mantle-balancer-behaviors} also demonstrates features that we
-will explore in the rest of this section.
-Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify the differences in performance
-when the cluster stabilizes at time 100 seconds and
-Section~\ref{sec:feature-backoff} examines the slope and start time of the
-rebalancing phase between 0 and 60 seconds by comparing the aggressiveness of
-the balancers.
-
-\subsubsection{Feature: Balancing Modes}
-\label{sec:feature-balancing-modes}
-
-Next, we quantify the performance benefits shown in
-Figure~\ref{fig:mantle-balancer-behaviors}.  To understand why load balancers
-perform differently we need to explain the different balancing modes that the
-load balancer service uses and how they stress the subsystems that receive and
-forward client requests in different ways. In
-Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
-performance of the balancing mode that CephFS falls into {\it most of the time}.
-CephFS has 3 modes for balancing load: CPU mode, workload mode, and hybrid
-mode. All three have the same structure for making migration decisions but vary
-based on the metric used to calculate load. For this sequencer workload the 3
-different modes all have the same performance, shown in
-Figure~\ref{fig:mantle-balancer-performance} (a), because the load balancer falls
-into the same mode a majority of the time.  The high variation in performance
-for the CephFS CPU Mode bar reflects the uncertainty of using something as
-dynamic and unpredictable as CPU utilization to make migration decisions. In
-addition to the suboptimal performance and unpredictability, because
-CephFS balancers all fall into the same mode despite using different balancers
-is a problem because it prevents administrators from properly exploring the
-balancing state space.
-
-Mantle gives the administrator more control over balancing policies; for
-the Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the load
-balancing interface to program logic for balancing read-heavy workloads,
-resulting in better throughput and stability.  When we did this we also
-identified two balancing modes relevant for making migration decisions
-for sequencers.
-
-Using Mantle, the administrator can put the load balancing service into ``proxy
-mode" or ``client mode". In proxy mode one server receives all requests and
-farms off the requests to slave servers; the slaves servers do the actual tail
-finding operation. In client mode, clients interact directly with the server
-that has their sequencer.  These modes are illustrated in
-Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
-co-located on one physical server -- performance for that mode is shown by the
-No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
-Mode" in the lower left shows how clients for sequencer 2 are re-directed to
-server B. The two ``Proxy mode" diagrams on the right show how clients continue
-sending requests to server A even though sequencer 2 has migrated to server B
-(the meaning of ``(Half)" is discussed in the next section).
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-modes.png}
-\caption{ In client mode clients sending requests to the server that houses
-their sequencer. In proxy mode clients continue sending their requests to the
-first server.  }\label{fig:mantle-modes}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-mode-behavior.png}
-\caption{The performance of proxy mode achieves the highest throughput but at
-the cost of lower throughput for one of the sequencers. Client mode is more
-fair but results in lower cluster throughput.
-}\label{fig:mantle-mode-behavior}
-\end{figure}
-
-Figure~\ref{fig:mantle-mode-behavior} shows the throughput over time of the two
-different modes for an environment with only 2 sequencers (again 4 clients
-each) and 2 servers. The curves for both sequencers in
-Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second and
-at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
-Performance of Sequencer 2 decreases because it stayed on the proxy which now
-handles all client requests, processes requests for Sequencer 2, and forwards
-requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
-because distributing the sequencers in this way separates (1) the handling of
-the client requests and (2) finding the tail of the log and responding to
-clients.  Doing both steps is too heavy weight for one server and sequencers on
-slave nodes can go faster if work is split up; this phenomenon is not
-uncommon and has been observed in chain replication~\cite{chain_rep}.
-
-Cluster
-throughput improves at the cost of decreased throughput for Sequencer 2.
-Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer mode manually (no
-balancing phase) and shows that the cluster throughput is worse than the
-cluster throughput of proxy mode. That graph also shows that Sequencer 2 has
-less throughput than Sequencer 2. In this case, the scatter-gather process used
-for cache coherence in the metadata protocols causes strain on the server
-housing Sequencer 2 resulting in this uneven performance.
-
-\subsubsection{Feature: Migration Units}
-\label{sec:feature-migration-units}
-
-Another factor that affects performance in this environment is how much load
-is on each server; these experiments quantify that effect by programming the
-load balancing interface to control the amount of load to migrate. We call this
-metric a ``migration unit".  Expressing this heuristic is not easily achievable
-with outward facing tunable parameters (i.e. system knobs) but with Mantle's
-programmable interface, we can force the load balancer to change its migration
-units. To force the balancer into the Proxy Mode (Half) scenario in
-Figure~\ref{fig:mantle-modes}, which uses migration units equal to half the
-load on the current server, we can use: \texttt{ targets[whoami+1] = mds[whoami]["load"]/2
-}.
-
-This code snippet uses globally defined variables and tables from the Mantle
-API to send half of the load on the current server (whoami) to the next ranked
-server (whoami + 1); the \texttt{targets} array is a globally defined table that
-the balancer uses to do the migrations. Alternatively, to migrate all load a
-time step, we can remove the division by 2.
-
-%\begin{figure}[t!]
-%\centering
-%\includegraphics{figures/mantle-mode-performance.png}
-%\caption{The performance of the different modes using various migration units
-%shows almost a 2\(\times\) improvement in the best case.
-%}\label{fig:mantle-mode-performance}
-%\end{figure}
-
-Figure~\ref{fig:mantle-balancer-performance} (b) shows the performance of the modes
-using different migration units. Recall that this setup only has 2 sequencers
-and 2 servers, so performance may be different at scale. Even so, it is clear
-that client mode does not perform as well for read-heavy workloads. We even see
-a throughput improvement when migrating all load off the first server, leaving
-the first server to do administrative tasks (this is common in the metadata
-cluster because the first server does a lot of the cache coherence work) while
-the second server does all the client work. Proxy mode does the best in both
-cases and shows large performance gains when completely decoupling client
-request handling and operation processing in Proxy Mode (Full).  The parameter
-that controls the migration units helps the administrator control the sequencer
-co-location or distribution across the cluster. This trade-off was explored
-extensively in the Mantle paper but the experiments we present here are
-indicative of an even richer set of states to explore.
-
-\subsubsection{Feature: Backoff}
-\label{sec:feature-backoff}
-
-Tuning the aggressiveness of the load balancer decision making is also a
-trade-off that administrators can control and explore. The balancing phase from
-0 to 60 seconds in Figure~\ref{fig:mantle-balancer-behaviors} shows different
-degrees of aggressiveness in making migration decisions; CephFS makes a
-decision quickly and throughput jumps to 2500 ops/second within 10 seconds
-while Mantle takes more time to stabilize. This conservative behavior is
-controlled by programming the balancer to (1) use different conditions for when
-to migrate and (2) using a threshold for sustained overload.
-
-We control the conditions for when to migrate using \texttt{when()}, a callback
-in the Mantle API.  For the Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} we program \texttt{when()} to wait
-for load on the receiving server to fall below a threshold. This makes the
-balancer more conservative because it takes 60 seconds for cache coherence
-messages to settle.  The Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} also takes longer to reach peak
-throughput because we want the policy to wait to see how migrations affect the
-system before proceeding; the balancer does a migration right before 50
-seconds, realizes that there is a third underloaded server, and does another
-migration.
-
-The other way to change aggressiveness of the decision making is to program
-into the balancer a threshold for sustained overload. This forces the balancer
-to wait a certain number of iterations after a migration before proceeding. In
-Mantle, the policy would use the save state function to do a countdown after a
-migration.  Behavior graphs and performance numbers for this backoff feature is
-omitted for space considerations, but our experiments confirm that the more
-conservative the approach the less overall throughput.
-
-\textbf{Malacology pulls} the load balancing service out of the storage system
-to balance sequencers across a cluster. This latent capability also gives
-future programmers the ability to explore the different load balancing
-trade-offs including: load balancing modes to control forwarding vs. client
-redirection, load migration units to control sequencer distribution vs.
-co-location, and backoffs to control conservative vs. aggressive decision
-making.
-
-%\subsubsection{Experiment: Object Class
-%Programmability}\label{experiment-object-class-programmability}
-%
-%Add quote from CORFU paper that talks about what they had to hack
-%
-%\subsubsection{Experiment: Multi-Client
-%Burstiness}\label{experiment-multi-client-burstiness}
-%
-%\begin{figure}[htbp] \centering
-%\includegraphics{figures/caps-delay-thruput.png} \caption{Forcing the client
-%to drop their capabilities later (delay) improves throughput} \end{figure}
-
 \section{Related Work}
@@
 Exokernel~\cite{engler:sosp95}, SPIN~\cite{bershad:sosp95} and
-Vino~\cite{seltzer:osdi96} projects; the latter two addressed the ability of injecting
-code into the kernel to specialize resource management. Another approach is to
-pass hints between the different layers of the IO stack to bridge the semantic
-gap between applications and
-storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}).
+Vino~\cite{seltzer:osdi96} projects; the latter two addressed the ability of
+injecting code into the kernel to specialize resource management. Another
+approach is to pass hints between the different layers of the IO stack to
+bridge the semantic gap between applications and
+storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}.
 
@@
 asynchronous object calls and Lua stubs in the inode, respectively.
-
-\section{Conclusion and Future Work}
-\label{conclusion-and-future-work}
-
-Programmable storage is a viable method for eliminating duplication of complex
-error-prone software used as workarounds for storage system
-deficiencies. We propose that systems expose their services in a safe way
-allowing application developers to customize system behavior to meet their
-needs while not sacrificing correctness. To illustrate the benefits of this
-approach we presented Malacology, a programmable storage system that
-facilitates the construction of new services by re-purposing existing subsystem
-abstractions of the storage stack.
-
-Future work will focus on constructing more interfaces to support a wide
-variety of storage system services that can be configured on-the-fly in existing
-systems. This work is one point along that path to producing general-purpose
-storage systems that can
-target  special-purpose applications.  Ultimately we want to utilize
-declarative methods for expressing new services.
-
-\bibliography{paper}
-\bibliographystyle{plain}
-%\printbibliography
-
-\end{document}

### 2017-02-23 paper/9_conclusion.tex
paper: cleanup line breaks and break into separate files
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-02-09 paper/paper.tex
@@
-\documentclass[preprint]{sigplanconf-eurosys}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage{pifont}
-\usepackage{color}
-\usepackage{comment}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Uncomment for diff
-\includecomment{notes}
-\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
-\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
-\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
-
-% Uncomment for camera-ready
-%\excludecomment{notes}
-%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
-%\newcommand{\oldcomment}[1]{}
-%\newcommand{\addressesissue}[1]{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\cmark}{\ding{51}}%
-\usepackage[font=footnotesize,labelfont=bf]{caption}
-
-\setlength{\belowcaptionskip}{-10pt}
-
-\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
-\usepackage[normalem]{ulem}
-\usepackage{pstricks}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-\setkeys{Gin}{width=\maxwidth,keepaspectratio}
-
-\usepackage[unicode=true]{hyperref}
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, Peter Alvaro, Shel Finkelstein, Jeff LeFevre},
-            pdftitle={Malacology: A Programmable Storage System},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=blue,
-            linkcolor=black,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-
-\title{Malacology: A Programmable Storage System}
-%\authorinfo{Paper ID: 165}          {Total Pages: 12}
-\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre}
-           {\{msevilla, jayhawk, ivo, cmaltzahn\}@soe.ucsc.edu, \{palvaro, finkelstein, jlefevre\}@ucsc.edu}
-\date{}
-\begin{document}
-
-\maketitle
-
-\begin{notes}
-\textcolor{red}{
-\noindent Issues raised by reviewers:
-\begin{enumerate}
-  \item clarify what was implemented: diagram of interfaces and prose
-  specifying the difference between Malacology and prior work (reviewers B, C,
-  D, E)
-  \item describe what makes a programmable storage system (reviewers C, D)
-  \item organize motivation: group interfaces into categories (e.g., core
-  functionality, features, and performance optimizations) and describe other
-  applications we can build with Malacology (reviewers B, C)
-  \item clarify terminology: service vs. file system metadata, storage vs.
-  file system, dirty vs. clean slate approaches, soft-state caching (reviewers
-  B, C)
-  \item address security, access control, and the safety; specify how it
-  affects composability, performance, and multi-tenancy (reviewer D)
-\end{enumerate}
-}
-\end{notes}
-
-\begin{abstract} Storage systems are caught between the need to evolve data
-processing applications efficiently and quickly, and the increasing velocity
-with which storage device technology evolves. This puts tremendous pressure on
-storage systems to support rapid change both in terms of their interfaces and
-their performance. But adapting storage systems can be difficult because
-unprincipled changes might jeopardize years of code-hardening and performance
-optimization efforts that were necessary for users to entrust their data to the
-storage system. We introduce the concept of a programmable storage system, a
-system that exposes internal services and abstractions with the goal of
-re-using and building new services on top of the storage stack. We present
-Malacology, a prototype implementation of a programmable storage system to
-explore how existing abstractions of common storage system services can be
-leveraged to adapt to the needs of new data processing systems and the
-increasing variety of storage devices. We show how this approach allows a large
-degree of flexibility for storage systems to evolve without sacrificing the
-robustness of their code-hardened subsystems. We illustrate the advantages and
-challenges of programmability by composing existing primitives into two new
-higher-level services: a metadata load balancer and a high-performance
-distributed shared-log. The evaluation demonstrates that our services inherit
-desirable qualities of the back-end storage system, including the ability to
-balance load, efficiently propagate cluster metadata, recover from failure, and
-to navigate trade-offs between latency and throughput using leases.
-\end{abstract}
-
-\begin{figure}[tb]
-\centering
-\includegraphics{figures/overview.png}
-\caption{Scalable storage systems have storage daemons which store data,
-monitor daemons (M) that maintain cluster state, and service-specific daemons
-({\it e.g.}, file system metadata servers). Malacology enables the programmability of
-internal abstractions (bold arrows) to re-use and compose
-existing subsystems.  With Malacology, we add new services, ZLog and
-Mantle, that sit alongside traditional user-facing APIs (file, block,
-object).}\label{fig:overview}
-\end{figure}
-
-\section{Introduction}
-\label{introduction}
-\label{sec:intro}
-
-A storage system implements abstractions designed to persistently store data
-and must exhibit a high level of correctness to prevent data loss.  Storage
-systems have evolved around storage devices that often were orders of magnitude
-slower than CPU and memory, and therefore could dominate overall performance if
-not used carefully. Over the last few decades members of the storage systems
-community have developed clever strategies to meet correctness requirements
-while somewhat hiding the latency of traditional storage
-media~\cite{brewer_disks_2016}. To avoid lock-in by a particular vendor, users
-of storage systems have preferred systems with highly standardized APIs and
-lowest common denominator abstract data types such as blocks of bytes and byte
-stream files~\cite{armbrust_view_2010}.
-
-A number of recent developments have disrupted traditional storage systems.
-First, the falling prices of flash storage and the availability of new types of
-non-volatile memory that are orders of magnitude faster than traditional
-spinning media are moving overall performance bottlenecks away from storage
-devices to CPUs and networking, and pressure storage systems to shorten their
-code paths and incorporate new
-optimizations~\cite{gray_tape_2007,gray_flash_2008}.  Second, emerging ``big
-data'' applications demand interface evolution to support flexible consistency
-as well as flexible structured data 
-representations.~\cite{apache_contributors_parquet_2014}.  Finally, production-quality scalable
-storage systems available as open source software have established and are
-continuing to establish new, \emph{de-facto} API standards at a faster pace
-than traditional standards
-bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
-
-The evolutionary pressure placed on storage systems by these trends raises the
-question of whether there are principles that storage systems designers can
-follow to evolve storage systems efficiently, without jeopardizing years of
-code-hardening and performance optimization efforts.  In this paper we
-investigate an approach that focuses on identifying and exposing existing
-storage system resources, services, and abstractions that in a generalized form
-can be used to \emph{program} new services. By composing higher-level services
-over existing abstractions one can reuse subsystems and leverage their
-optimizations, established correctness, robustness, and efficiency.
-
-{\it \textbf{Contribution 1:}} We define a programmable storage system to be a
-storage system that facilitates the re-use and extension of existing storage
-abstractions provided by the underlying software stack, to enable the creation
-of new services via composition.  A programmable storage system can be realized
-by exposing existing functionality (such as metadata services and
-synchronization and monitoring capabilities) as interfaces that can be ``glued
-together'' in a variety of ways using a high-level language. Programmable
-storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection
-and execution of code within a storage system or storage device---in that the
-former is applicable to any component of the storage system, while the latter
-focuses at the data access level. Given this contrast, we can say that active
-storage is an example of how one internal component (the storage layer) is
-exposed in a programmable storage system.
-
-To illustrate the benefits and challenges of this approach we have designed and
-evaluated Malacology, a programmable storage system that facilitates the
-construction of new services by re-purposing existing subsystem abstractions of
-the storage stack.  We build Malacology in Ceph, a popular open source software
-storage stack.  We choose Ceph to demonstrate the concept of programmable
-storage because it offers a broad spectrum of existing services, including
-distributed locking and caching services provided by metadata servers,
-durability and object interfaces provided by the backend object store, and
-propagation of consistent cluster state provided by the monitoring service (see
-Figure~\ref{fig:overview}).  Malacology is expressive enough to provide the
-functionality necessary for implementing new services.
-
-Malacology includes a non-exhaustive set of interfaces that can be used as
-building blocks for constructing novel storage abstractions, including:
-
-\begin{enumerate}
-
-\item An interface for managing strongly-consistent time-varying \textbf{service
-metadata}.
-
-\item An interface for installing and evolving domain-specific, cluster-wide
-\textbf{data interfaces}.
-
-\item An interface for managing access to \textbf{shared resources} using a
-variety of optimization strategies.
-
-\item An interface for \textbf{load balancing} resources across the cluster.
-
-\item An interface for \textbf{durability} that persists policies using the
-underlying storage stack's object store.
-
-\end{enumerate}
-
-{\it \textbf{Contribution 2:}} We implement two distributed services using Malacology to
-demonstrate the feasibility of the programmable storage approach:
-
-\begin{enumerate}
-
-\item A high-performance distributed shared log service called ZLog, that is an
-implementation of CORFU.~\cite{balakrishnan_corfu_2012}
-
-\item An implementation of Mantle, the programmable load balancing
-service~\cite{sevilla:sc15-mantle}
-
-\end{enumerate}
-
-The remainder of this paper is structured as follows. First, we describe and
-motivate the need for programmable storage by describing current practices in
-the open source software community. Next we describe Malacology by presenting the
-subsystems within the underlying storage system that we re-purpose, and briefly
-describe how those system are used within Malacology (\S\ref{sec:malacology}).
-Then we describe the services that we have constructed within the Malacology
-framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
-implementation (\S\ref{sec:evaluation}).  We conclude by discussing related and future work.
-% and
-%conclude.
-
-\section{Application-Specific Storage Stacks}
-\label{sec:application-specifc-storage-stacks}
-
-Building storage stacks from the ground up for a specific purpose results in
-the best performance. For example, GFS~\cite{ghemawat:sosp03} and
-HDFS~\cite{shvachko:msst10} were designed specifically to serve Map/Reduce and
-Hadoop jobs, and use techniques like exposing data locality and relaxing POSIX
-constraints to achieve application-specific I/O optimizations. Another example
-is Boxwood~\cite{maccormick:osdi04}, which experimented with B-trees
-and chunk stores as storage abstractions to simplify application building.
-Alternatively, general-purpose storage stacks are built with the flexibility to
-serve many applications by providing a variety of interfaces and tunable
-parameters.  Unfortunately, managing competing forces in these systems is
-difficult and users want more control from the general-purpose storage stacks
-without going as far as building their storage system from the ground up.
-
-To demonstrate a recent trend towards more application-specific storage systems
-we examine the state of programmability in Ceph. Something of a storage Swiss
-army knife, Ceph simultaneously supports file, block, and object interfaces on
-a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
-Distributed Object Storage (RADOS) system is a cluster of object storage
-devices (OSDs) that provide Ceph with data durability and integrity using
-replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already
-provides some degree of programmability; the OSDs support domain-specific
-object interfaces that are implemented by composing existing low-level storage
-interfaces that execute atomically. These interfaces are written in C++ and are
-statically loaded into the system.
-
-The Ceph community provides empirical evidence that developers are already
-beginning to embrace programmable storage. Figure~\ref{fig:obj-int-dev-growth}
-shows a dramatic growth in the production use of domain-specific interfaces in
-the Ceph community since 2010. In that figure, classes are functional groupings
-of methods on storage objects (e.g. remotely computing and caching the checksum
-of an object extent).  What is most remarkable is that this trend contradicts
-the notion that API changes are a burden for users.  Rather it appears that
-gaps in existing interfaces are being addressed through ad-hoc approaches to
-programmability. In fact, Table~\ref{table:objclasses} categorizes existing
-interfaces and we clearly see a trend towards reusable services.
-
-\begin{figure}[ht]
-\centering
-\includegraphics{figures/obj-int-dev-growth.png}
-\caption{Since 2010, the growth in the number of co-designed object storage interfaces
-in Ceph has been accelerating. This plot is the number of object classes (a
-group of interfaces), and the total number of methods (the actual API
-end-points).}
-\label{fig:obj-int-dev-growth}
-\end{figure}
-
-\begin{table}[ht]
-\centering
-  \begin{tabular}{l|l|l}
-    Category & Example & \# \\ \hline
-    Logging  & Geographically distribute replicas & 11 \\ \hdashline
-    Metadata & Snapshots in the block device OR  & \multirow{2}{*}{74} \\
-    Management & Scan extents for file system repair & \\ \hdashline
-    Locking  & Grants clients exclusive access & 6 \\ \hdashline
-    Other & Garbage collection, reference counting  & 4\\
-\end{tabular}
-\caption{A variety of object storage classes exist to expose interfaces
-    to applications. \# is the number of methods that implement these categories.
-}
-\label{table:objclasses}
-\end{table}
-
-The takeaway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
-already trying to use programmability because their needs, whether they be
-related to performance, availability, consistency, convenience, etc., are not
-satisfied by the existing default set of interfaces. The popularity of the
-custom object interface facility of Ceph could be due to a number of reasons,
-such as the default algorithms/tunables of the storage system being
-insufficient for the application's performance goals, programmers wanting to
-exploit application-specific semantics, and/or programmers knowing how to
-manage resources to improve performance. A solution based on
-application-specific object interfaces is a way to work around the
-traditionally rigid storage APIs because custom object interfaces give
-programmers the ability to tell the storage system about their application: if
-the application is CPU or IO bound, if it has locality, if its size has the potential to
-overload a single node, etc.  Programmers often know what the problem is
-and how to solve it, but until the ability to modify object interfaces, they
-had no way to express to the storage system how to handle their data.
-
-Our approach is to expose more of the commonly used, code-hardened subsystems
-of the underlying storage system as interfaces. The intent is that these
-interfaces, which can be as simple as a redirection to the persistent data
-store or as complicated as a strongly consistent directory service, should be
-used and re-used in many contexts to implement a wide range of services. By
-making programmability a `feature', rather than a `hack' or `workaround', we
-help standardize a development process that now is largely ad-hoc.
-
-\section{Challenges}
-\label{sec:challenges}
-
-Implementing the infrastructure for programmability into existing services and
-abstractions of distributed storage systems is challenging, even if one assumes
-that the source code of the storage system and the necessary expertise for
-understanding it is available.  Some challenges include:
-
-\begin{itemize}
-
-\item Storage systems are generally required to be highly available so that any
-complete restarts of the storage system to reprogram them is usually
-unacceptable.
-
-\item Policies and optimizations are usually hard-wired into the services and
-one has to be careful when factoring them to avoid introducing additional bugs.
-These policies and optimizations are usually cross-cutting solutions to
-concerns or trade-offs that cannot be fully explored at the time the code is
-written (as they relate to workload or hardware). Given these policies and
-optimizations, decomposition of otherwise orthogonal internal abstractions can
-be difficult or dangerous.
-
-\item Mechanisms that are often only exercised according to hard-wired policies
-and not in their full generality have hidden bugs that are revealed as soon as
-those mechanisms are governed by different policies. In our experience
-introducing programmability into a storage system proved to be a great
-debugging tool.
-
-\item Programmability, especially in live systems, implies changes that need to
-be carefully managed by the system itself, including versioning and propagation
-of those changes without affecting correctness.
-
-\end{itemize}
-
-To address these challenges we present Malacology
-\oldcomment{. Malacology is both a prototype for a programmable storage system
-and a}
-\newcomment{, our prototype programmable storage system. It uses the
-programmable storage}
-\addressesissue{2}
-design approach to evolve storage systems efficiently and
-without jeopardizing years of code-hardening and performance optimization
-efforts.  Although Malacology uses the internal abstractions of the underlying
-storage system, including its subsystems, components, and implementations, we
-emphasize that our system still addresses the general challenges outlined
-above.
-
-The main challenge of designing a programmable storage system is choosing the
-right internal abstractions and picking the correct layers for exposing them.
-\oldcomment{In this paper, we do not present an exhaustive list of possible
-internal abstractions nor do we contend that the abstractions we choose provide
-the best trade-offs for all applications.} 
-\newcomment{A programmable storage is not defined by what abstractions are
-exposed, rather a programmable storage system adheres to the design approach of
-choosing interfaces so administrators can have better control of the storage
-system. The interfaces presented in this paper are abstractions that we found
-useful for building ZLog and Mantle, yet they may not provide the best
-trade-offs for all higher-level services.}
-\addressesissue{2}
-For example, if consensus is correctly exposed one could implement
-high-level features like versioning, serialization, or various flavors of
-strongly consistent data management on top; but perhaps a low-level consensus
-interface is suited well for a particular set of applications.  These questions
-are not answered in this paper and instead we focus on showing the feasibility
-of building such a system, given advances in the quality and robustness of
-today's storage stacks.
-
-The Malacology prototype we present has been implemented on Ceph.
-While there are other systems on top of which Malacology could be implemented (see Table~\ref{table:examples}), we choose Ceph because it is a production quality system and because it
-is open source. The large developer community ensures that code is
-robust and the visibility of the code lets us expose any interface we want. In
-the next section we describe the Ceph components that we expose as Malacology
-interfaces.
-
-\section{Malacology: A Programmable Storage System}
-\label{sec:malacology}
-
-\begin{table*}
-\centering\small
-\begin{tabular}{ l | c | l | l | l }
-%\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
-\multicolumn{4}{c}{} \\
-\textbf{Interface}                      &
-\textbf{Sec}                            &
-\textbf{Example in Production Systems}  &
-\textbf{Example in Ceph}                &
-\textbf{Provided Primitives}            \\ \hline
-Service Metadata
-  & \S\ref{sec:service-metadata-interface}
-  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
-  & cluster state management~\cite{website:ceph-mon}
-  & consensus/consistency
-  \\
-Data I/O
-  & \S\ref{sec:data-io-interface}
-  & Swift in situ storage/compute~\cite{website:zerocloud}
-  & object interface classes~\cite{website:cls-lua}
-  & transaction/atomicity
-  \\
-Shared Resource
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & MPI collective IO, burst
-  & POSIX metadata protocols
-  & serialization/batching
-  \\
-File Type
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & ?????
-  & file striping strategy
-  & data/metadata access
-  \\
-Load Balancing
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
-  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
-  & migration/sampling
-  \\
-Durability
-  & \S\ref{sec:durability}
-  & S3/Swift interfaces (RESTful API)
-  & object store library~\cite{weil_rados_2007}
-  & persistence/safety
-  \\
-\end{tabular}
-\caption{Common internal abstractions. The same ``internal abstractions" are
-common across large-scale systems because they provide primitives that solve
-general distributed systems problems.  Here we list examples of what these
-internal abstractions are used for in ``production systems" and in Ceph.
-Malacology provides these internal abstractions as interfaces
-\newcomment{(Section~\ref{sec:malacology})}
-that higher level
-\oldcomment{applications}
-\newcomment{services (Section~\ref{sec:services})}
-can use.
-\oldcomment{; {\it italicized abstractions} are interfaces discussed in this
-paper.}}
-\addressesissue{1}
-\label{table:examples}
-\end{table*}
-
-The guiding principle is to re-use existing services and extend them so that
-these services can be \emph{programmed}. We accomplish programmability of a
-service by exporting bindings (or ``hooks'') for an interpreted programming
-language so that programming can occur without having to restart the storage
-system (see also below,~\S\ref{sec:durability}). Table~\ref{table:examples}
-shows the internal services from Ceph that we expose in the Malacology
-prototype via Lua~\cite{ierusalimschy1996lua} bindings
-\oldcomment{.}
-\newcomment{and Figure~\ref{fig:implementation-overview} compares what was
-already present in Ceph (gray boxes) to the Malacology interfaces we added.
-Section~\ref{sec:services} will describe the higher-level services we built
-with these interfaces.}
-\addressesissue{1}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation-overview.png}
-\caption{\newcomment{Malacology is implemented on the daemons and clients that
-run in a Ceph cluster. Interfaces expose internal subsystems and are used as
-building blocks for higher-level services.}\addressesissue{1}
-\label{fig:implementation-overview}}
-\end{figure}
-
-Lua is a portable embedded scripting language and we choose it as the
-interpreted language for Malacology because it offers superior performance and
-productivity trade-offs, including a JIT-based implementation that is well
-known for near native performance. Additionally, Lua has been used extensively
-in game engines, and systems research \cite{neto:dls14-luaos}, including
-storage systems where it has been effectively used both on
-\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010}
-and off \cite{sevilla:sc15-mantle} the performance critical path. Finally, the
-flexibility of the runtime allows execution sandboxing in order to address
-security and performance concerns. We will now discuss the common subsystems
-used to manage storage system and how Malacology makes them programmable.
-
-\subsection{Service Metadata Interface}
-\label{sec:mon}
-\label{sec:service-metadata-interface}
-\label{service-metadata}
-
-% straw man example
-Keeping track of state in a distributed system is an essential part of any
-successful service and a necessary component in order to diagnose and detect
-failures, when they occur. This is further complicated by variable propagation
-delays and heterogeneous hardware in dynamic environments.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In the case of Ceph, a consistent view of cluster state among server daemons
-and clients is critical to provide strong consistency guarantees to clients.
-Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
-metadata server) data structures called ``maps'' that record membership and
-status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
-responsible for integrating state changes into cluster maps, responding to
-requests from out-of-date clients and synchronizing members of the cluster
-whenever there is a change in a map so that they all observe the same system
-state. As a fundamental building block of many system designs, consensus
-abstractions such as Paxos are a common technique for maintaining consistent
-data versions, and are a useful system to expose.
-
-The default behavior of the monitor can be seen as a Paxos-based notification
-system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing
-clients to identify when new values (termed epochs in Ceph) are associated to
-given maps.  Since Ceph does not expose this service directly, as part of our Malacology implementation, we expose a key-value
-service designed for managing configuration metadata that is built on top of
-the consensus engine. Since the monitor is intended to be out of the
-high-performance I/O path, a general guideline is to make use of this
-functionality infrequently and to assign small values to maps.\\
-
-\noindent{\it{\textbf{Malacology: }}} Malacology exposes a strongly-consistent view of
-time-varying service metadata as a service rather than a hidden internal
-component. Malacology provides a generic API for adding arbitrary values to
-existing subsystem cluster maps. As a consequence of this, applications can
-define simple but useful service-specific logic to the strongly-consistent
-interface, such as authorization control (just specific clients can write new
-values) or triggering actions based on specific values (e.g. sanitize values).
-The higher level services we implement in \S\ref{sec:services} make use of this
-functionality to register, version and propagate dynamic code (Lua scripts) for
-new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
-policies in the load balancer \S\ref{malacology:mds}.  Using this service
-guarantees that interface definitions are not only made durable, but are
-transparently and consistently propagated throughout the cluster so that
-clients are properly synchronized with the latest interfaces.
-
-\newcomment{\\ \noindent \it{\textbf{Impact: }} provides core functionality
-because it lets daemons come to consensus on system critical state. Bugs in the
-internal subsystems or ommitting this from services that need this type of
-consistency affects correctness.}
-\addressesissue{3, 5}
-
-%And finally, it is important to not underestimate the importance of the
-%preservation of the modifications that enable a new system service. In many
-%(if not all) cases, interfaces defining access to data are just as important
-%as the data itself by virtue of inherently providing structural context. Thus,
-%all components of a system service must be kept to the same standards of
-%protection as data in the system itself.
-
-\oldcomment{\noindent 4.2 Object I/O Intefrace}
-\subsection{Data I/O Interface}
-\label{sec:data-io-interface}
-\label{object-data-interface}
-Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
-supports application-specific object
-interfaces~\cite{weil_rados_2007}. The ability to offload computation can
-reduce data movement, and transactional interfaces can significantly simplify
-construction of complex storage interfaces that require uncoordinated parallel
-access.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-An object interface is a plugin structured similarly to that of an RPC
-in which a developer creates a named function within the cluster that clients
-may invoke. In the case of Ceph each function is implicitly run within the
-context of an object specified when the function is called. Developers of
-object interfaces express behavior by creating a composition of native
-interfaces or other custom object interfaces, and handle serialization of
-function input and output.  A wide range of native interfaces are available to
-developers such as reading and writing to a byte stream, controlling object
-snapshots and clones, and accessing a sorted key-value database. These native
-interfaces may be transactionally composed along with application
-specific logic to create semantically rich interfaces. An example would be
-an interface that atomically updates a matrix stored in the bytestream and an
-index of the matrix stored in the key-value database.\\
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation.png}
-\caption{Malacology allows users to dynamically define object/metadata
-interfaces by extending the object storage daemon (OSD) and metadata server
-(MDS) subsystems with an embedded Lua VM.  It uses the service metadata
-interface (M) to propagate interfaces/versions across the cluster
-\label{fig:implementation}}
-\end{figure}
-
-\noindent {\it{\textbf{Malacology: }}} The implementation of Ceph's object abstraction,
-although powerful, does not readily support programmability. Supporting only
-C/C++ for object interface developers, Ceph requires distribution of compiled
-binaries for the correct architecture, adding a large barrier of entry for
-developers and system administrators. Second, having no way to dynamically
-unload modules, any changes require a full restart of a storage daemon which
-may have serious performance impacts due to loss of cached data. And finally,
-the security limitations of the framework limit the use of object interfaces to
-all but those with administrative level access and deep technical expertise.
-
-To address these concerns, Malacology takes advantage of Lua extensions
-contributed by the Ceph community. This allows new object interfaces to be
-dynamically loaded into the system and modified at runtime, resulting in a
-object storage API with economy of expression, which at the same time provides
-the full set of features of the base object class. New object interfaces that
-are expressed in thousands of lines of code can be implemented in approximately
-an order of magnitude less code~\cite{geambasu_comet_2010}. While the use of
-Lua does not prevent deployment of malicious code, certain types of coding
-mistakes can be handled gracefully, and access policies are used to limit access
-to trusted users~\cite{ierusalimschy1996lua}.
-
-\oldcomment{All of these modifications work in tandem to implement the desired behavior,
-and are typically co-designed together such that each depends on the other to
-behave as expected.}
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by pushing behavior to lower parts of the storage stack, thereby
-minimizing hops and distributing computation.}
-\addressesissue{3, 5}
-
-\subsection{Distributed Metadata Interfaces}
-\begin{notes}
-\textcolor{red}{This section has been re-organized}
-\end{notes}
-\label{sec:distributed-metadata-interfaces}
-\label{malacology:mds}
-
-\oldcomment{The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction~\cite{weil:sc2004-dyn-metadata}.}
-\newcomment{File systems provide clients with the familiar POSIX file abstraction.
-While this guarantees strong consistency it comes at the cost of scalability,
-increased complexity, and lower performance.}
-\addressesissue{1}
-In general, distributed file systems protect resources by providing
-hierarchical indexing and distributed locking services.
-
-%The metadata cluster moves units of the namespace called directory fragments
-%and assigns them to servers using its own metadata load balancer. It then uses
-%a hard-coded policy to balance these fragments using a scalarized load metric
-%based on CPU, workload, and file system operation metrics. Although the CephFS
-%balancer has been shown to be complicated, its complexity is justifiable.
-%Having the flexibility to choose what to move, where to move it, and how much
-%to move is very powerful. The problem is that it is difficult to pick a
-%one-size-fits-all policy, system administrators may want to partition load
-%based on SLAs, or applications are better equipped to make these decisions,
-%etc. -- but we choose instead to expose a balancing API at strategic points in
-%the balancing logic using Lua hooks. ~\cite{sevilla:sc15-mantle}.
-
-\oldcomment{\it{\textbf{Malacology:}} Interfaces are added in strategic spots in the
-metadata service for guarding shared resources, defining new file types,
-and load balancing.}
-
-\oldcomment{\noindent Shared resource interface.}
-\subsubsection{Shared Resource Interface}
-
-In general, file system metadata servers manages client sessions, allowing
-clients to obtain locks (e.g. file byte ranges), and capabilities (e.g. to
-cache file data). Clients and metadata servers use a cooperative protocol in
-which clients voluntarily release resources back to the metadata service in
-order to implement sharing policies.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In Ceph, the locking service implements a capability-based system that
-expresses what data and metadata clients are allowed to access as well as what
-state they may cache and modify locally.  While designed for the file
-abstraction, indexing, locking, and caching are all common services that are
-useful to a broad spectrum of applications.  Distributed applications that
-share centralized resources (e.g. a database or directory) face similar
-challenges which are often solved using application-specific sharding.
-
-\newcomment{\\ \noindent\it{\textbf{Malacology: }}}
-While the current policy for sharing access and voluntarily releasing resources
-is largely best-effort, Malacology supports generalized policies between
-metadata servers and clients that can be used to implement fairness or
-priority.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} provides core functionality
-to protect and provide exclusive access for any shared resource. May hurt
-performance if the resource in question does not require strong consistency.}
-\addressesissue{3}
-
-\oldcomment{\noindent File type interface.}
-\subsubsection{File Type Interface}
-The metadata service manages a POSIX file system hierarchy where files and
-directories are represented as inode data structures and expose a POSIX file
-interface. Applications that manage large amounts of metadata (e.g. users or
-database snapshots) often require a naming service. We allow new inode types to
-be defined such that applications can create domain-specific interfaces to
-inodes that may modify locking and capability policies. We will show how this
-is used in Section~\ref{sec:seq} when we discuss a distributed shared-log built
-on Malacology.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this interface is both a feature
-and a performance optimization. It is a feature because it allows developers to
-add support for different  storage types, such as how to read new file formats
-or what consistency semantics to use for a specific subtree in the hierarchical
-namespace. It is also a performance optimization because future programmers can
-add optimizations for processing specific types of files into the inode itself.}
-\addressesissue{3}
-
-%%% TODO: striping strategy is stored with the inode
-
-\oldcomment{\noindent \bf Load balancing interface.} 
-\subsubsection{Load Balancing Interface}
-
-Ceph also addresses the challenge of balancing metadata load with a separate
-metadata cluster. This cluster uses load balancing policies to migrate
-directory inodes around the cluster to alleviate overload at single
-nodes~\cite{weil:sc2004-dyn-metadata}. The policies use metrics based on system
-state (e.g.  CPU and memory utilization) and statistics collected by the
-cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
-partitioning to move variable sized namespace subtrees. These units can be
-shipped anywhere (i.e., to any metadata server of any capacity) at any time for
-any reason. The original balancer was designed with hard-coded policies and
-tunables.\\
-
-\noindent {\it{\textbf{Malacology: }}}
-Mantle~\cite{sevilla:sc15-mantle}, a programmable metadata load balancer, has
-been implemented on Malacology so that its infrastructure can enjoy all the
-properties of existing internal abstractions.  The load balancing logic is
-intact but the infrastructure has been improved to safely load, manage, and
-persist its load balancing policies.  Composing Mantle with the service
-metadata and durability interfaces, Mantle can safely version balancer
-policies, save balancer policies in the backend object store and centralize
-warnings/errors. When combined with file types, load balancing policies can
-express a variety of rules such as multi-tenant and multi-application
-balancing.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by allowing them to specify how to partition, replicate, and
-distribute metadata in response to overloaded servers.}
-\addressesissue{3}
-
-\subsection{Durability Interface}
-\label{sec:durability}
-
-Ceph provides storage by striping and replicating data across
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-protects data using common techniques such as erasure
-coding, replication, and scrubbing.  Furthermore, many of these features
-are implemented using a peer-to-peer protocol that allows object storage
-devices to operate autonomously without a centralized coordinator.  For
-example, when the number of placement groups change, the OSDs re-balance and
-re-shard data in the background in a process called placement group splitting
-in which OSDs communicate directly with each other to cover a new state.
-In order to reduce load on the monitoring service, Ceph OSDs use a gossip
-protocol to efficiently propagate changes to cluster maps throughout the
-system, and autonomously initiate recovery mechanisms when failures are
-discovered.\\
-
-\noindent {\it{\textbf{Malacology: }}} Metadata service policies and object storage
-interfaces are stored durability within RADOS and managed by storing references
-with the object maps. Since the cluster already propagates a consistent view of
-these data structures, we use this service to automatically install interfaces
-in OSDs, and install policies within the metadata server daemon (MDS) such that clients and daemons are
-synchronized on correct implementations without restarting.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this is a feature because it
-adds data safety and persistence to system metadata; while nice to have it does
-effect correctness.}
-\addressesissue{3}
-
-\section{Services Built on Malacology}
-\label{sec:services}
-\label{services-built-on-malacology}
-
-\label{services}
-%\begin{table}
-%\centering\small
-%\begin{tabular}{  l | l | l }
-%\textbf{Service}              &
-%\textbf{Used to...}           &
-%\textbf{Malacology Interface} \\ \hline
-%Mantle & load balance metadata     & Load Balancing  \\
-%       & agree on current balancer & Service Metadata\\
-%       & control client write lock & Shared Resource \\
-%       & persist policy logic      & Durability      \\ \hline
-%ZLog   & implement CORFU           & Object \& Data  \\ 
-%       & install ZLog interfaces   & Service Metadata\\
-%       & persist log entries       & Durability      \\ \hline
-%ZLog   & load balance sequencers   & Load Balancing  \\
-%Seqr.  & manage access to log tail & Shared Resource \\ 
-%\end{tabular}
-%\caption{Mantle and ZLog are built using Malacology interfaces so they inherit the
-%robustness of the subsystems of the underlying stack.}
-%\label{table:implementation}
-%\end{table}
-
-In this section we describe two services built on top of Malacology. The first
-is Mantle, a framework for dynamically specifying metadata load balancing
-policies. The second system, ZLog, is a high-performance distributed shared-log.
-In addition to these services, we'll demonstrate how we combine ZLog and Mantle
-to implement service-aware metadata load balancing policies.
-
-\subsection{Mantle: Programmable Load Balancer}
-\label{sec:mantle}
-
-Mantle is a programmable load balancer that separates the metadata balancing
-policies from their mechanisms. Administrators inject code to change how the
-metadata cluster distributes metadata. In ~\cite{sevilla:sc15-mantle} the
-authors showed how using Mantle, one can implement a single node metadata
-service, a distributed metadata service with hashing, and a distributed
-metadata service with dynamic subtree partitioning.
-
-The original implementation was ``hard-coded" into Ceph and lacked robustness
-(no versioning, durability, or policy distribution).  Re-implemented using
-Malacology, Mantle now enjoys (1) the versioning provided by the metadata
-service interface provided by the monitor daemons and (2) the durability and
-distribution provided by Ceph's reliable object store.  Re-using the internal
-abstractions with Malacology resulted in a 2\(\times\) reduction in source code
-compared to the original implementation.
-
-\subsubsection{Versioning Balancer Policies}
-
-Ensuring that the version of the current load balancer is consistent across the
-physical servers in the metadata cluster was not addressed in the original
-implementation. The user had to set the version on each individual server and
-it was trivial to make the versions inconsistent. Maintaining consistent
-versions is important for balancing policies that are cooperative in which local
-decisions are made assuming properties about other instances of the same
-policy.
-
-With Malacology, Mantle stores the version of the current load balancer in the
-Ceph service metadata interface. The version of the load balancer corresponds
-to an object name in the balancing policy. Using the service metadata interface
-ensures that all metadata servers use the same version of the load balancer. As
-a result, the policy version inherits the consistency of Ceph's internal
-monitor daemons. The user changes the version of the load balancer using a new
-CLI command.
-
-\subsubsection{Making Balancer Policies Durable}
-
-The load balancer version described above corresponds to the name of an object
-in RADOS that holds the actual Lua balancing code.  When MDS nodes start
-balancing load, they first check the latest version from the MDS map and
-compare it to the balancer they have loaded. If the version has changed, they
-dereference the pointer to the balancer version by reading the corresponding
-object in RADOS. This is in contrast to the original balancer which stored load
-balancer code on the local file system -- a technique which is unreliable and
-may result in silent corruption.
-
-The balancer pulls the Lua code from RADOS synchronously; asynchronous reads
-are not possible because of the architecture of the MDS. The synchronous
-behavior is not the default behavior for RADOS operations, so we achieve this
-with a timeout: if the asynchronous read does not come back within half the
-balancing tick interval the operation is canceled and a Connection Timeout
-error is returned. By default, the balancing tick interval is 10 seconds, so
-Mantle will use a 5 second second timeout.
-
-This design allows Mantle to immediately return an error if anything
-RADOS-related goes wrong.  We use this implementation because we do not want to
-do a blocking OSD read from inside the global MDS lock. Doing so would bring
-down the MDS cluster if any of the OSDs are not responsive.
-
-Storing the balancers in RADOS is simplified by the use of an interpreted
-language for writing balancer code. If we used a language that needs to be
-compiled, like the C++ object classes in the OSD, we would need to ensure
-binary compatibility, which is complicated by different operating systems,
-distributions, and compilers.
-
-\subsubsection{Logging, Debugging, and Warnings}
-
-In the original implementation, Mantle would log all errors, warnings, and
-debug messages to a log stored locally on each MDS server. To get the simplest
-status messages or to debug problems, the user would have to log into each MDS
-individually, look at the logs, and reason about causality and ordering.
-
-With Malacology, Mantle re-uses the centralized logging features of the
-monitoring service. Important errors, warnings, and info messages are collected
-by the monitoring subsystem and appear in the monitor cluster log so instead of
-users going to each node, they can watch messages appear at the monitor.
-Messages are logged sparingly, so as not to spam the monitor with frivolous
-debugging but important events, like balancer version changes or failed
-subsystems show up in the centralized log.
-
-\subsection{ZLog: A Fast Distributed Shared Log}
-\label{sec:zlog}
-
-The second service implemented on Malacology is ZLog, a high-performance
-distributed shared-log that is based on the CORFU
-protocol~\cite{balakrishnan_corfu_2012}. The shared-log is a
-powerful abstraction used to construct distributed systems, such as
-metadata management~\cite{balakrishnan:sosp13} and elastic database
-systems~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
-However, existing implementations that rely on consensus algorithms such as
-Paxos funnel I/O through a single point introducing a bottleneck that restricts
-throughput. In contrast, the CORFU protocol is able to achieve high throughput
-using a network counter called a \emph{sequencer}, that
-decouples log position assignment from log I/O.
-
-While a full description of the CORFU system is beyond the scope of this
-paper, we briefly describe the custom storage device interface, sequencer
-service, and recovery protocol, and how these services are instantiated in the
-Malacology framework.
-
-\subsubsection{Sequencer}
-\label{sec:seq}
-
-High-performance in CORFU is achieved using a sequencer service that assigns
-log positions to clients by reading from a volatile, in-memory counter which
-can run at a very high throughput and at low latency. Since the sequencer is
-centralized, ensuring serializability in the common case is trivial.  The
-primary challenge in CORFU is handling the failure of the sequencer in a way
-that preserves correctness. Failure of the sequencer service in CORFU is
-handled by a recovery algorithm that recomputes the new sequencer state using a
-CORFU-specific custom storage interface to discover the tail of the log, while
-simultaneously invalidating stale client requests using an epoch-based
-protocol.
-
-{\bf Sequencer interface.} The sequencer resource supports the ability to
-\emph{read()} the current tail value and get the \emph{next()} position in
-the log which also atomically increments the tail position.
-We implement the sequencer service in Malacology as an application-specific
-file type that associates a small amount of metadata
-with the inode state managed by the Ceph metadata service.
-This approach has the added
-benefit of allowing the metadata service to handle naming, by
-representing each sequencer instance in the standard POSIX hierarchical
-namespace. The primary challenge in mapping the sequencer resource to the
-metadata service is handling serialization correctly to maintain the global
-ordering provided by the CORFU protocol.
-
-Initially we sought to directly model the sequencer service in Ceph as a
-non-exclusive, non-cacheable resource, forcing clients to perform a round-trip
-access to the resource at the authoritative metadata server for the sequencer
-inode.  Interestingly, we found that the capability system in Ceph uses a
-strategy to reduce metadata service load by allowing clients that open a
-shared file to temporarily obtain an exclusive cached copy of the resource,
-resulting in a round-robin, best-effort batching behavior. When a single
-client is accessing the sequencer resource it is able to increment the
-sequencer locally. Any competing client cannot query the sequencer until the
-metadata service has granted it access.
-
-While unexpected, this discovery allowed us to explore an implementation
-strategy that we had not previously considered. In particular, for bursty
-clients, and clients that can tolerate increased latency, this mode of
-operation may allow a system to achieve much higher throughput than a system
-with a centralized sequencer service.  We utilize the programmability of the
-metadata service to define a new policy for handling capabilities that controls
-the amount of time that clients are able to cache the sequencer resource. This
-allows an administrator or application to control the trade-off between latency
-and throughput beyond the standard best-effort policy that is present in Ceph
-by default.
-
-In Section~\ref{sec:evaluation} we quantify the trade-offs of throughput and
-latency for an approach based on a round-robin batching mode, and compare this
-mode to one in which the metadata server mediates access to the sequencer state
-when it is being shared among multiple clients. Quantifying these trade-offs
-should provide administrators with guidelines for setting the tunables for
-different ``caching" modes of the sequencer.
-
-{\bf Balancing policies.} As opposed to the batching mode for controlling access to the sequencer
-resource, more predictable latency can be achieved by treating the sequencer
-inode as a shared non-cacheable resource, forcing clients to make a round-trip
-to the metadata service. However, the shared nature of the metadata service
-may prevent the sequencer from achieving maximum throughput. To provide a
-solution to this issue we have taken advantage of the programmability of the
-metadata service load balancing to construct a service-specific load balancing
-policy. As opposed to a balancing policy that strives for uniform
-load distribution, a ZLog-specific policy may utilize knowledge of inode types
-to migrate the sequencer service to provisioned hardware during periods of
-contention or high demand.
-
-\subsubsection{Storage Interface}
-
-The storage interface is a critical component in the CORFU protocol. Clients
-independently map log positions that they have obtained from the sequencer
-service (described in detail in the next section) onto storage devices, while
-storage devices provide an intelligent \emph{write}-once, random \emph{read}
-interface for accessing log entries. The key to correctness in CORFU lies with
-the enforcement of up-to-date epoch tags on client requests; requests tagged
-with out-of-date epoch values are rejected, and clients are expected to
-request a new tail from the sequencer after refreshing state from an auxiliary
-service.  This mechanism forms the basis for sequencer recovery.
-
-In order to repopulate the sequencer state (i.e. the cached, current tail of
-the log) during recovery of a sequencer, the maximum position in the log must
-be obtained. To do this, the storage interface exposes an additional
-\emph{seal} method that atomically installs a new epoch value and returns the
-maximum log position that has been written.
-
-Since the sequencer service does not resume until the recovery process has
-completed, there cannot be a race with clients appending to the log, and the
-immutability of the log allow reads to never block during a sequencer failure.
-Recovery of the sequencer process itself may be handled in many ways, such as
-leader election using an auxiliary service such as Paxos. In our
-implementation, the recovery is the same as (and is inherited from) the Ceph
-metadata service. Handling the failure of a client that holds the sequencer
-state is similar, although a timeout is used to determine when a client should
-be considered unavailable.
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-%We wanted to implement CORFU and we had two choices: build it from the ground
-%up or do it on top of Malacology. To show the feasability of building on
-%Malacology our evaluation focuses on the
-
-Our evaluation demonstrates the feasibility of building new service
-abstractions atop programmable storage, focusing on the performance of the
-internal abstractions exposed by Malacology and used to construct the Mantle
-and ZLog services. We also discuss latent capabilities we discovered in this
-process that let us navigate different trade-offs within the services
-themselves. First, we benchmark scenarios with high sequencer contention by
-examining the interfaces used to map ZLog onto Malacology; specifically, we
-describe the sequencer implementation and the propagation of object and data
-interfaces interfaces.  Next, we benchmark scenarios in which the storage
-system manages multiple logs by using Mantle to balance sequencers across a
-cluster.
-
-% should save this for future work, otherwise it undercuts the eval.
-%Detailed system-wide performance measurements and analysis is forthcoming for
-%both Mantle and ZLog, as we try to find the best metadata balancing policies
-%and physical design implementation trade-offs.
-
-Since this work focuses on the programmability of Malacology, the goal of this
-section is to show that the components and subsystems that support the
-Malacology interfaces provide reasonable relative performance, as well as to
-give examples of the flexibility that Malacology provides to programmers.
-
-%As we search for the best metadata balancing policies
-%and physical design implementation trade-offs,
-%we plan to rel
-
-\subsection{Mapping ZLog onto Malacology}
-\label{sec:mapping-zlog-onto-malacology}
-
-We evaluate Malacology by exploring one possible mapping of the ZLog
-implementation of CORFU onto Ceph in which we re-use (1) the metadata service
-to manage naming and synchronization of the sequencer resource by treating the
-resource as an inode, and (2) the monitoring sub-subsystem to distribute and
-install application-specific I/O interface required of the CORFU protocol. In
-Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the inode
-abstraction for implementing the sequencer resource enables load balancing policies
-to migrate the sequencer resource in heavy-load situation.
-
-\subsubsection{Sequencer Implementation}
-\label{sec:sequencer-implementation}
-
-We evaluate the feasibility of using the metadata service to implement a
-sequencer resource that is responsible for maintaining a total ordering of
-the log. Clients contact the sequencer to obtain the tail of the log and then
-independently initiate I/O, thus we measure both the throughput and latency of
-obtaining new tail positions which bounds client append performance.
-
-The sequencer is implemented as a custom file type in which the sequencer
-state (a 64-bit integer) is embedded in the inode of a file representing the
-sequencer. A total ordering of the log is imposed by the re-use of the
-capability service that can be used to grant exclusive access of inode state
-to clients. The metadata service is responsible for maintaining exclusivity
-and granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
-behavior of the system in which a best-effort policy is used.  The two colors
-represent points in time that the clients were able to access the resource.
-The best-effort policy shows a high degree of interleaving
-between clients but the system spends a large portion of time re-distributing
-the capability, reducing overall throughput.
-
-In order to control the performance of the system we implement a
-policy that (1) restricts the length of time that a client may maintain
-exclusive access and (2) limits the number of log positions that a client may
-generate without yielding to other clients waiting for access. The behavior of
-these two modes is illustrated in Figures~\ref{fig:capdelay-quota-behavior}
-(b) and (c), respectively.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/capdelay-quota-behavior-a.png}
-\includegraphics{figures/capdelay-quota-behavior-b.png}
-\includegraphics{figures/capdelay-quota-behavior-c.png}
-\caption{Each dot is an individual request, spread randomly along the \(y\)
-axis. The default behavior is unpredictable, ``delay" lets clients hold the
-lease longer, and ``quota" gives clients the lease for a number of operations.}
-\label{fig:capdelay-quota-behavior}
-\end{figure}
-
-Figure~\ref{fig:captp} demonstrates a configurable trade-off between
-throughput and latency. In the experiment two clients are run each with a
-fixed 0.25 second maximum reservation on the capability, and we vary the size
-of the log position quota running each configuration for two minutes. The
-total operations per second is the combined throughput of the two clients, and
-the average latency is number of microseconds required to obtain a new log
-position. With a small quota more time is spent exchanging exclusive access,
-while a large quota reservation allows clients to experience a much lower
-latency because they experience uncontended access for a longer period of time.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/tradeoff.png}
-\caption{Sequencer throughput by re-using various services.  The highest performance is
-achieved using a single client with exclusive, cacheable privilege. Round-robin
-sharing of the sequencer resource is affected by the amount of time the
-resource is held, with best-effort performing the worst.}
-\label{fig:captp}
-\end{figure}
-
-To get a better picture of latency, Figure~\ref{fig:capcdf} shows the CDF of
-latency for each client in all experiment configurations. At the 99th
-percentile clients accessed the sequencer is less than a millisecond. The CDF
-is cropped at the 99.999th percentile due to large outliers that we believe
-occur in instances in which the MDS is performing I/O while it is in the
-process of re-distributing the capability to another client.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/caps-delay-latency.png}
-\caption{
-The latency distribution shows that if the clients hold their capabilities longer,
-performance decreases. Changing the delay can help the sequencer and
-clients strike a balance between throughput and latency.}
-\label{fig:capcdf}
-\end{figure}
-
-\textbf{Malacology exposes} the internal capability management service and
-allows users to navigate latency and throughput trade-offs.  Other approaches to
-designing the sequencer service also exist, such as using a centralized service
-in which each access is a network round-trip. In contrast to the mechanism we
-explored which is appropriate for clients with bursty workloads, it may be
-easier to provide predictable performance using a centralized service, and we
-will be exploring in the future how this can be achieved using the capability
-system.
-
-\subsubsection{Interface Propagation}
-
-Domain-specific data interfaces (Section~\ref{sec:application-specifc-storage-stacks}) allow
-co-design between applications and the storage system. Malacology supports
-custom object interfaces in RADOS that require interface implementations to
-be installed on the storage devices in the system, supporting the evolution of
-interfaces through automatic system-wide versioning and installation through
-the service metadata interface (Section~\ref{service-metadata}). We evaluate
-the performance of installing a new interface version in the cluster,
-which is an important metric for applications that frequently evolve
-interfaces.
-
-We demonstrate the feasibility of utilizing the Ceph monitoring sub-system by
-evaluating the performance of installing and distributing interface updates.
-Figure~\ref{fig:propdelay} shows the CDF of the latency of interface updates.
-The interfaces are Lua scripts embedded in the cluster map and distributed
-using a peer-to-peer gossip protocol.  The latency is defined as the elapsed
-time following the Paxos proposal for an interface update until each OSD makes
-the update live (the cost of the Paxos proposal is configurable and is
-discussed below). The latency measurements were taken on the nodes running
-Ceph server daemons, and thus exclude the client round-trip cost. In each of
-the experiments 1000 interface updates were observed.
-
-Figure~\ref{fig:propdelay} shows the lower bound cost for updates in a large
-cluster. In the experiment labeled ``120 OSD (RAM)" a cluster of 120 OSDs (10
-OSDs x 12 servers) using an in-memory data store were deployed, showing a
-latency of less than 54 ms with a probability of 90\% and a worst case latency
-of 194 ms. These costs demonstrate the penalty of distributing the interface
-in a large cluster. In practice the costs include, in addition to cluster-wide
-propagation of interface updates, the network round-trip to the
-interface management service, the Paxos commit protocol itself, and other factors
-such as system load. By default Paxos proposals occur periodically with a 1
-second interval in order to accumulate updates. In a minimum, realistic quorum
-of 3 monitors using HDD-based storage, we were able to decrease this interval
-to an average of 222 ms.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=60mm,trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
-\caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
-interface commit.}
-\label{fig:propdelay}
-\end{figure}
-
-\subsection{Load Balancing ZLog Sequencers with Mantle}
-\label{sec:zlog-balancing}
-
-In practice, a storage system implementing CORFU will support a multiplicity of
-independent totally-ordered logs for each application.  For this scenario
-co-locating sequencers on the same physical node is not ideal but building a load
-balancer that can migrate the shared resource ({\it e.g.,}, the resource that
-mediates access to the tail of the log) is a time-consuming, non-trivial task.
-It requires building subsystems for migrating resources, monitoring the
-workloads, collecting metrics that describe the utilization on the physical
-nodes, partitioning resources, maintaining cache coherence, and managing
-multiple sequencers. The following experiments demonstrate the feasibility of
-using the mechanisms of the Malacology load balancing interface to inherit
-these features and to alleviate load from overloaded servers.
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-behaviors.png}
-\caption{CephFS/Mantle load balancing have better throughput than co-locating
-all sequencers on the same server.  Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify this improvement;
-Section~\ref{sec:feature-backoff} examines the migration at 0-60 seconds.
-}\label{fig:mantle-balancer-behaviors}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-performance.png}
-\includegraphics{figures/mantle-mode-performance.png}
-\caption{In (a) all CephFS balancing modes have the same performance; Mantle uses a
-balancer designed for sequencers. In (b) the best combination of mode and
-migration units can have up to a 2\(\times\) improvement.}\label{fig:mantle-balancer-performance}
-\end{figure}
-
-The experiments are run on a cluster with 10 nodes to store objects, one node
-to monitor the cluster, and 3 nodes that can accommodate sequencers. 
-Instead of measuring contention at the clients like
-Section~\ref{sec:sequencer-implementation}, these experiments measure
-contention at the sequencers by forcing clients to make round-trips for every
-request. We implement this using the shared resource interface that forces
-round-trips.  Because the sequencer's only function is to  hand out positions
-for the tail of the log, the workload is read-heavy.
-
-First, we show how the ZLog service can orchestrate multiple sequencers using
-the Malacology load balancing interface.
-Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput over time of
-different load balancers as they migrate 3 sequencers (with 4 clients) around
-the cluster; ``No Balancing" keeps all sequencers on one server, ``CephFS"
-migrates sequencers using the CephFS load balancers, and ``Mantle" uses a
-custom load balancer we wrote specifically for sequencers. The CephFS load
-balancers are hard-coded into Ceph while Mantle is an API for balancing load
-presented in~Mantle~\cite{sevilla:sc15-mantle} and re-implemented on top of
-Malacology in this paper.  The increased throughput for the CephFS and Mantle
-curves between 0 and 60 seconds are a result of migrating the sequencer(s) off
-overloaded servers.
-
-In addition to showing that migrating sequencers improves performance,
-Figure~\ref{fig:mantle-balancer-behaviors} also demonstrates features that we
-will explore in the rest of this section.
-Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify the differences in performance
-when the cluster stabilizes at time 100 seconds and
-Section~\ref{sec:feature-backoff} examines the slope and start time of the
-rebalancing phase between 0 and 60 seconds by comparing the aggressiveness of
-the balancers.
-
-\subsubsection{Feature: Balancing Modes}
-\label{sec:feature-balancing-modes}
-
-Next, we quantify the performance benefits shown in
-Figure~\ref{fig:mantle-balancer-behaviors}.  To understand why load balancers
-perform differently we need to explain the different balancing modes that the
-load balancer service uses and how they stress the subsystems that receive and
-forward client requests in different ways. In
-Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
-performance of the balancing mode that CephFS falls into {\it most of the time}.
-CephFS has 3 modes for balancing load: CPU mode, workload mode, and hybrid
-mode. All three have the same structure for making migration decisions but vary
-based on the metric used to calculate load. For this sequencer workload the 3
-different modes all have the same performance, shown in
-Figure~\ref{fig:mantle-balancer-performance} (a), because the load balancer falls
-into the same mode a majority of the time.  The high variation in performance
-for the CephFS CPU Mode bar reflects the uncertainty of using something as
-dynamic and unpredictable as CPU utilization to make migration decisions. In
-addition to the suboptimal performance and unpredictability, because
-CephFS balancers all fall into the same mode despite using different balancers
-is a problem because it prevents administrators from properly exploring the
-balancing state space.
-
-Mantle gives the administrator more control over balancing policies; for
-the Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the load
-balancing interface to program logic for balancing read-heavy workloads,
-resulting in better throughput and stability.  When we did this we also
-identified two balancing modes relevant for making migration decisions
-for sequencers.
-
-Using Mantle, the administrator can put the load balancing service into ``proxy
-mode" or ``client mode". In proxy mode one server receives all requests and
-farms off the requests to slave servers; the slaves servers do the actual tail
-finding operation. In client mode, clients interact directly with the server
-that has their sequencer.  These modes are illustrated in
-Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
-co-located on one physical server -- performance for that mode is shown by the
-No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
-Mode" in the lower left shows how clients for sequencer 2 are re-directed to
-server B. The two ``Proxy mode" diagrams on the right show how clients continue
-sending requests to server A even though sequencer 2 has migrated to server B
-(the meaning of ``(Half)" is discussed in the next section).
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-modes.png}
-\caption{ In client mode clients sending requests to the server that houses
-their sequencer. In proxy mode clients continue sending their requests to the
-first server.  }\label{fig:mantle-modes}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-mode-behavior.png}
-\caption{The performance of proxy mode achieves the highest throughput but at
-the cost of lower throughput for one of the sequencers. Client mode is more
-fair but results in lower cluster throughput.
-}\label{fig:mantle-mode-behavior}
-\end{figure}
-
-Figure~\ref{fig:mantle-mode-behavior} shows the throughput over time of the two
-different modes for an environment with only 2 sequencers (again 4 clients
-each) and 2 servers. The curves for both sequencers in
-Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second and
-at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
-Performance of Sequencer 2 decreases because it stayed on the proxy which now
-handles all client requests, processes requests for Sequencer 2, and forwards
-requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
-because distributing the sequencers in this way separates (1) the handling of
-the client requests and (2) finding the tail of the log and responding to
-clients.  Doing both steps is too heavy weight for one server and sequencers on
-slave nodes can go faster if work is split up; this phenomenon is not
-uncommon and has been observed in chain replication~\cite{chain_rep}.
-
-Cluster
-throughput improves at the cost of decreased throughput for Sequencer 2.
-Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer mode manually (no
-balancing phase) and shows that the cluster throughput is worse than the
-cluster throughput of proxy mode. That graph also shows that Sequencer 2 has
-less throughput than Sequencer 2. In this case, the scatter-gather process used
-for cache coherence in the metadata protocols causes strain on the server
-housing Sequencer 2 resulting in this uneven performance.
-
-\subsubsection{Feature: Migration Units}
-\label{sec:feature-migration-units}
-
-Another factor that affects performance in this environment is how much load
-is on each server; these experiments quantify that effect by programming the
-load balancing interface to control the amount of load to migrate. We call this
-metric a ``migration unit".  Expressing this heuristic is not easily achievable
-with outward facing tunable parameters (i.e. system knobs) but with Mantle's
-programmable interface, we can force the load balancer to change its migration
-units. To force the balancer into the Proxy Mode (Half) scenario in
-Figure~\ref{fig:mantle-modes}, which uses migration units equal to half the
-load on the current server, we can use: \texttt{ targets[whoami+1] = mds[whoami]["load"]/2
-}.
-
-This code snippet uses globally defined variables and tables from the Mantle
-API to send half of the load on the current server (whoami) to the next ranked
-server (whoami + 1); the \texttt{targets} array is a globally defined table that
-the balancer uses to do the migrations. Alternatively, to migrate all load a
-time step, we can remove the division by 2.
-
-%\begin{figure}[t!]
-%\centering
-%\includegraphics{figures/mantle-mode-performance.png}
-%\caption{The performance of the different modes using various migration units
-%shows almost a 2\(\times\) improvement in the best case.
-%}\label{fig:mantle-mode-performance}
-%\end{figure}
-
-Figure~\ref{fig:mantle-balancer-performance} (b) shows the performance of the modes
-using different migration units. Recall that this setup only has 2 sequencers
-and 2 servers, so performance may be different at scale. Even so, it is clear
-that client mode does not perform as well for read-heavy workloads. We even see
-a throughput improvement when migrating all load off the first server, leaving
-the first server to do administrative tasks (this is common in the metadata
-cluster because the first server does a lot of the cache coherence work) while
-the second server does all the client work. Proxy mode does the best in both
-cases and shows large performance gains when completely decoupling client
-request handling and operation processing in Proxy Mode (Full).  The parameter
-that controls the migration units helps the administrator control the sequencer
-co-location or distribution across the cluster. This trade-off was explored
-extensively in the Mantle paper but the experiments we present here are
-indicative of an even richer set of states to explore.
-
-\subsubsection{Feature: Backoff}
-\label{sec:feature-backoff}
-
-Tuning the aggressiveness of the load balancer decision making is also a
-trade-off that administrators can control and explore. The balancing phase from
-0 to 60 seconds in Figure~\ref{fig:mantle-balancer-behaviors} shows different
-degrees of aggressiveness in making migration decisions; CephFS makes a
-decision quickly and throughput jumps to 2500 ops/second within 10 seconds
-while Mantle takes more time to stabilize. This conservative behavior is
-controlled by programming the balancer to (1) use different conditions for when
-to migrate and (2) using a threshold for sustained overload.
-
-We control the conditions for when to migrate using \texttt{when()}, a callback
-in the Mantle API.  For the Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} we program \texttt{when()} to wait
-for load on the receiving server to fall below a threshold. This makes the
-balancer more conservative because it takes 60 seconds for cache coherence
-messages to settle.  The Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} also takes longer to reach peak
-throughput because we want the policy to wait to see how migrations affect the
-system before proceeding; the balancer does a migration right before 50
-seconds, realizes that there is a third underloaded server, and does another
-migration.
-
-The other way to change aggressiveness of the decision making is to program
-into the balancer a threshold for sustained overload. This forces the balancer
-to wait a certain number of iterations after a migration before proceeding. In
-Mantle, the policy would use the save state function to do a countdown after a
-migration.  Behavior graphs and performance numbers for this backoff feature is
-omitted for space considerations, but our experiments confirm that the more
-conservative the approach the less overall throughput.
-
-\textbf{Malacology pulls} the load balancing service out of the storage system
-to balance sequencers across a cluster. This latent capability also gives
-future programmers the ability to explore the different load balancing
-trade-offs including: load balancing modes to control forwarding vs. client
-redirection, load migration units to control sequencer distribution vs.
-co-location, and backoffs to control conservative vs. aggressive decision
-making.
-
-%\subsubsection{Experiment: Object Class
-%Programmability}\label{experiment-object-class-programmability}
-%
-%Add quote from CORFU paper that talks about what they had to hack
-%
-%\subsubsection{Experiment: Multi-Client
-%Burstiness}\label{experiment-multi-client-burstiness}
-%
-%\begin{figure}[htbp] \centering
-%\includegraphics{figures/caps-delay-thruput.png} \caption{Forcing the client
-%to drop their capabilities later (delay) improves throughput} \end{figure}
-
-\section{Related Work}
-
-% IOStack: Software-Defined Object Storage
-% sRoute: Treating the Storage Stack Like a Network
-   % - lots of good refs to check out
-% Information and control in gray-box systems
-
-Programmability of operating systems and networking resources, including
-distributed storage systems is not new, but we are not aware of work that makes
-generalization of existing services into programmable resources a key principle
-in storage systems design.
-
-Programmable storage systems can be viewed as an infrastructure for creating
-abstractions to better separate policies from mechanisms. This idea is not new.
-Software-defined networks (SDNs) create such an abstraction by separating the
-control plane from the data plane (see for example~\cite{jain:sigcomm13}). This
-notion of control/data separation was also applied in software-defined storage
-(SDS)~\cite{thereska:sosp13,stefanovici:fast16}. Similarly,
-IOStack~\cite{gracia:internet16} is providing policy-based provisioning and
-filtering in OpenStack Swift. According to a SNIA white
-paper~\cite{carlson:snia2014}, the primary goal of SDS is to control and
-facilitate flexible and dynamic provisioning of storage resources of different
-kinds, including flash memory and disk drives, to create a virtualized mapping
-between common storage abstractions (e.g. files, objects, and blocks) and
-storage devices taking data service objectives in terms of protection,
-availability, performance, and security into account. A programmable storage
-system exposes internal abstractions so that end users (not necessarily
-operators) can create new services on top of the storage stack. Thus, our
-notion of programmable storage differs from ``software-defined storage'' (SDS)
-in terms of goals and scope, although definitions of SDS are still in flux.
-
-Another view of programmable storage systems is one of tailoring systems
-resources to applications~\cite{arpaci:sosp01}. Related efforts include the
-Exokernel~\cite{engler:sosp95}, SPIN~\cite{bershad:sosp95} and
-Vino~\cite{seltzer:osdi96} projects; the latter two addressed the ability of injecting
-code into the kernel to specialize resource management. Another approach is to
-pass hints between the different layers of the IO stack to bridge the semantic
-gap between applications and
-storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}).
-
-Malacology uses the same Active and Typed Storage module presented in
-DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File Manifolds
-can be implemented with small changes to the Malacology framework, namely
-asynchronous object calls and Lua stubs in the inode, respectively.
-
-\section{Conclusion and Future Work}
+\section{Conclusion}
 \label{conclusion-and-future-work}
@@
 Programmable storage is a viable method for eliminating duplication of complex
-error-prone software used as workarounds for storage system
-deficiencies. We propose that systems expose their services in a safe way
-allowing application developers to customize system behavior to meet their
-needs while not sacrificing correctness. To illustrate the benefits of this
-approach we presented Malacology, a programmable storage system that
-facilitates the construction of new services by re-purposing existing subsystem
-abstractions of the storage stack.
+error-prone software used as workarounds for storage system deficiencies. We
+propose that systems expose their services in a safe way allowing application
+developers to customize system behavior to meet their needs while not
+sacrificing correctness. To illustrate the benefits of this approach we
+presented Malacology, a programmable storage system that facilitates the
+construction of new services by re-purposing existing subsystem abstractions of
+the storage stack.
 
-Future work will focus on constructing more interfaces to support a wide
-variety of storage system services that can be configured on-the-fly in existing
-systems. This work is one point along that path to producing general-purpose
-storage systems that can
-target  special-purpose applications.  Ultimately we want to utilize
-declarative methods for expressing new services.
-
-\bibliography{paper}
-\bibliographystyle{plain}
-%\printbibliography
-
-\end{document}
+\oldcommentone{Future work will focus on constructing more interfaces to
+support a wide variety of storage system services that can be configured
+on-the-fly in existing systems. This work is one point along that path to
+producing general-purpose storage systems that can target  special-purpose
+applications.  Ultimately we want to utilize declarative methods for expressing
+new services.}

### 2017-02-23 paper/paper.tex
paper: cleanup line breaks and break into separate files
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-02-09 paper/paper.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Uncomment for diff
-\includecomment{notes}
-\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
-\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
-\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
+% Uncomment for diff 0
+%\includecomment{notes}
+%\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
+%\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
+%\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
+
+% Uncomment for diff 1
+%\includecomment{notes}
+%\newcommand{\newcommentone}[1]{{\textcolor{red}{\textbf{#1}}}}
+%\newcommand{\oldcommentone}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
+%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
+%\newcommand{\oldcomment}[1]{}
+%\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}
 
 % Uncomment for camera-ready
-%\excludecomment{notes}
-%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
-%\newcommand{\oldcomment}[1]{}
-%\newcommand{\addressesissue}[1]{}
-
+\excludecomment{notes}
+\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
+\newcommand{\oldcomment}[1]{}
+\newcommand{\newcommentone}[1]{{\textcolor{black}{#1}}}
+\newcommand{\oldcommentone}[1]{}
+\newcommand{\addressesissue}[1]{}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{notes}
-\textcolor{red}{
-\noindent Issues raised by reviewers:
-\begin{enumerate}
-  \item clarify what was implemented: diagram of interfaces and prose
-  specifying the difference between Malacology and prior work (reviewers B, C,
-  D, E)
-  \item describe what makes a programmable storage system (reviewers C, D)
-  \item organize motivation: group interfaces into categories (e.g., core
-  functionality, features, and performance optimizations) and describe other
-  applications we can build with Malacology (reviewers B, C)
-  \item clarify terminology: service vs. file system metadata, storage vs.
-  file system, dirty vs. clean slate approaches, soft-state caching (reviewers
-  B, C)
-  \item address security, access control, and the safety; specify how it
-  affects composability, performance, and multi-tenancy (reviewer D)
-\end{enumerate}
-}
-\end{notes}
-
-\begin{abstract} Storage systems are caught between the need to evolve data
-processing applications efficiently and quickly, and the increasing velocity
-with which storage device technology evolves. This puts tremendous pressure on
-storage systems to support rapid change both in terms of their interfaces and
-their performance. But adapting storage systems can be difficult because
-unprincipled changes might jeopardize years of code-hardening and performance
-optimization efforts that were necessary for users to entrust their data to the
-storage system. We introduce the concept of a programmable storage system, a
-system that exposes internal services and abstractions with the goal of
-re-using and building new services on top of the storage stack. We present
-Malacology, a prototype implementation of a programmable storage system to
-explore how existing abstractions of common storage system services can be
-leveraged to adapt to the needs of new data processing systems and the
-increasing variety of storage devices. We show how this approach allows a large
-degree of flexibility for storage systems to evolve without sacrificing the
-robustness of their code-hardened subsystems. We illustrate the advantages and
-challenges of programmability by composing existing primitives into two new
-higher-level services: a metadata load balancer and a high-performance
-distributed shared-log. The evaluation demonstrates that our services inherit
-desirable qualities of the back-end storage system, including the ability to
-balance load, efficiently propagate cluster metadata, recover from failure, and
-to navigate trade-offs between latency and throughput using leases.
-\end{abstract}
-
-\begin{figure}[tb]
-\centering
-\includegraphics{figures/overview.png}
-\caption{Scalable storage systems have storage daemons which store data,
-monitor daemons (M) that maintain cluster state, and service-specific daemons
-({\it e.g.}, file system metadata servers). Malacology enables the programmability of
-internal abstractions (bold arrows) to re-use and compose
-existing subsystems.  With Malacology, we add new services, ZLog and
-Mantle, that sit alongside traditional user-facing APIs (file, block,
-object).}\label{fig:overview}
-\end{figure}
-
-\section{Introduction}
-\label{introduction}
-\label{sec:intro}
-
-A storage system implements abstractions designed to persistently store data
-and must exhibit a high level of correctness to prevent data loss.  Storage
-systems have evolved around storage devices that often were orders of magnitude
-slower than CPU and memory, and therefore could dominate overall performance if
-not used carefully. Over the last few decades members of the storage systems
-community have developed clever strategies to meet correctness requirements
-while somewhat hiding the latency of traditional storage
-media~\cite{brewer_disks_2016}. To avoid lock-in by a particular vendor, users
-of storage systems have preferred systems with highly standardized APIs and
-lowest common denominator abstract data types such as blocks of bytes and byte
-stream files~\cite{armbrust_view_2010}.
-
-A number of recent developments have disrupted traditional storage systems.
-First, the falling prices of flash storage and the availability of new types of
-non-volatile memory that are orders of magnitude faster than traditional
-spinning media are moving overall performance bottlenecks away from storage
-devices to CPUs and networking, and pressure storage systems to shorten their
-code paths and incorporate new
-optimizations~\cite{gray_tape_2007,gray_flash_2008}.  Second, emerging ``big
-data'' applications demand interface evolution to support flexible consistency
-as well as flexible structured data 
-representations.~\cite{apache_contributors_parquet_2014}.  Finally, production-quality scalable
-storage systems available as open source software have established and are
-continuing to establish new, \emph{de-facto} API standards at a faster pace
-than traditional standards
-bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.
-
-The evolutionary pressure placed on storage systems by these trends raises the
-question of whether there are principles that storage systems designers can
-follow to evolve storage systems efficiently, without jeopardizing years of
-code-hardening and performance optimization efforts.  In this paper we
-investigate an approach that focuses on identifying and exposing existing
-storage system resources, services, and abstractions that in a generalized form
-can be used to \emph{program} new services. By composing higher-level services
-over existing abstractions one can reuse subsystems and leverage their
-optimizations, established correctness, robustness, and efficiency.
-
-{\it \textbf{Contribution 1:}} We define a programmable storage system to be a
-storage system that facilitates the re-use and extension of existing storage
-abstractions provided by the underlying software stack, to enable the creation
-of new services via composition.  A programmable storage system can be realized
-by exposing existing functionality (such as metadata services and
-synchronization and monitoring capabilities) as interfaces that can be ``glued
-together'' in a variety of ways using a high-level language. Programmable
-storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection
-and execution of code within a storage system or storage device---in that the
-former is applicable to any component of the storage system, while the latter
-focuses at the data access level. Given this contrast, we can say that active
-storage is an example of how one internal component (the storage layer) is
-exposed in a programmable storage system.
-
-To illustrate the benefits and challenges of this approach we have designed and
-evaluated Malacology, a programmable storage system that facilitates the
-construction of new services by re-purposing existing subsystem abstractions of
-the storage stack.  We build Malacology in Ceph, a popular open source software
-storage stack.  We choose Ceph to demonstrate the concept of programmable
-storage because it offers a broad spectrum of existing services, including
-distributed locking and caching services provided by metadata servers,
-durability and object interfaces provided by the backend object store, and
-propagation of consistent cluster state provided by the monitoring service (see
-Figure~\ref{fig:overview}).  Malacology is expressive enough to provide the
-functionality necessary for implementing new services.
-
-Malacology includes a non-exhaustive set of interfaces that can be used as
-building blocks for constructing novel storage abstractions, including:
-
-\begin{enumerate}
-
-\item An interface for managing strongly-consistent time-varying \textbf{service
-metadata}.
-
-\item An interface for installing and evolving domain-specific, cluster-wide
-\textbf{data interfaces}.
-
-\item An interface for managing access to \textbf{shared resources} using a
-variety of optimization strategies.
-
-\item An interface for \textbf{load balancing} resources across the cluster.
-
-\item An interface for \textbf{durability} that persists policies using the
-underlying storage stack's object store.
-
-\end{enumerate}
-
-{\it \textbf{Contribution 2:}} We implement two distributed services using Malacology to
-demonstrate the feasibility of the programmable storage approach:
-
-\begin{enumerate}
-
-\item A high-performance distributed shared log service called ZLog, that is an
-implementation of CORFU.~\cite{balakrishnan_corfu_2012}
-
-\item An implementation of Mantle, the programmable load balancing
-service~\cite{sevilla:sc15-mantle}
-
-\end{enumerate}
-
-The remainder of this paper is structured as follows. First, we describe and
-motivate the need for programmable storage by describing current practices in
-the open source software community. Next we describe Malacology by presenting the
-subsystems within the underlying storage system that we re-purpose, and briefly
-describe how those system are used within Malacology (\S\ref{sec:malacology}).
-Then we describe the services that we have constructed within the Malacology
-framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
-implementation (\S\ref{sec:evaluation}).  We conclude by discussing related and future work.
-% and
-%conclude.
-
-\section{Application-Specific Storage Stacks}
-\label{sec:application-specifc-storage-stacks}
-
-Building storage stacks from the ground up for a specific purpose results in
-the best performance. For example, GFS~\cite{ghemawat:sosp03} and
-HDFS~\cite{shvachko:msst10} were designed specifically to serve Map/Reduce and
-Hadoop jobs, and use techniques like exposing data locality and relaxing POSIX
-constraints to achieve application-specific I/O optimizations. Another example
-is Boxwood~\cite{maccormick:osdi04}, which experimented with B-trees
-and chunk stores as storage abstractions to simplify application building.
-Alternatively, general-purpose storage stacks are built with the flexibility to
-serve many applications by providing a variety of interfaces and tunable
-parameters.  Unfortunately, managing competing forces in these systems is
-difficult and users want more control from the general-purpose storage stacks
-without going as far as building their storage system from the ground up.
-
-To demonstrate a recent trend towards more application-specific storage systems
-we examine the state of programmability in Ceph. Something of a storage Swiss
-army knife, Ceph simultaneously supports file, block, and object interfaces on
-a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
-Distributed Object Storage (RADOS) system is a cluster of object storage
-devices (OSDs) that provide Ceph with data durability and integrity using
-replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already
-provides some degree of programmability; the OSDs support domain-specific
-object interfaces that are implemented by composing existing low-level storage
-interfaces that execute atomically. These interfaces are written in C++ and are
-statically loaded into the system.
-
-The Ceph community provides empirical evidence that developers are already
-beginning to embrace programmable storage. Figure~\ref{fig:obj-int-dev-growth}
-shows a dramatic growth in the production use of domain-specific interfaces in
-the Ceph community since 2010. In that figure, classes are functional groupings
-of methods on storage objects (e.g. remotely computing and caching the checksum
-of an object extent).  What is most remarkable is that this trend contradicts
-the notion that API changes are a burden for users.  Rather it appears that
-gaps in existing interfaces are being addressed through ad-hoc approaches to
-programmability. In fact, Table~\ref{table:objclasses} categorizes existing
-interfaces and we clearly see a trend towards reusable services.
-
-\begin{figure}[ht]
-\centering
-\includegraphics{figures/obj-int-dev-growth.png}
-\caption{Since 2010, the growth in the number of co-designed object storage interfaces
-in Ceph has been accelerating. This plot is the number of object classes (a
-group of interfaces), and the total number of methods (the actual API
-end-points).}
-\label{fig:obj-int-dev-growth}
-\end{figure}
-
-\begin{table}[ht]
-\centering
-  \begin{tabular}{l|l|l}
-    Category & Example & \# \\ \hline
-    Logging  & Geographically distribute replicas & 11 \\ \hdashline
-    Metadata & Snapshots in the block device OR  & \multirow{2}{*}{74} \\
-    Management & Scan extents for file system repair & \\ \hdashline
-    Locking  & Grants clients exclusive access & 6 \\ \hdashline
-    Other & Garbage collection, reference counting  & 4\\
-\end{tabular}
-\caption{A variety of object storage classes exist to expose interfaces
-    to applications. \# is the number of methods that implement these categories.
-}
-\label{table:objclasses}
-\end{table}
-
-The takeaway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
-already trying to use programmability because their needs, whether they be
-related to performance, availability, consistency, convenience, etc., are not
-satisfied by the existing default set of interfaces. The popularity of the
-custom object interface facility of Ceph could be due to a number of reasons,
-such as the default algorithms/tunables of the storage system being
-insufficient for the application's performance goals, programmers wanting to
-exploit application-specific semantics, and/or programmers knowing how to
-manage resources to improve performance. A solution based on
-application-specific object interfaces is a way to work around the
-traditionally rigid storage APIs because custom object interfaces give
-programmers the ability to tell the storage system about their application: if
-the application is CPU or IO bound, if it has locality, if its size has the potential to
-overload a single node, etc.  Programmers often know what the problem is
-and how to solve it, but until the ability to modify object interfaces, they
-had no way to express to the storage system how to handle their data.
-
-Our approach is to expose more of the commonly used, code-hardened subsystems
-of the underlying storage system as interfaces. The intent is that these
-interfaces, which can be as simple as a redirection to the persistent data
-store or as complicated as a strongly consistent directory service, should be
-used and re-used in many contexts to implement a wide range of services. By
-making programmability a `feature', rather than a `hack' or `workaround', we
-help standardize a development process that now is largely ad-hoc.
-
-\section{Challenges}
-\label{sec:challenges}
-
-Implementing the infrastructure for programmability into existing services and
-abstractions of distributed storage systems is challenging, even if one assumes
-that the source code of the storage system and the necessary expertise for
-understanding it is available.  Some challenges include:
-
-\begin{itemize}
-
-\item Storage systems are generally required to be highly available so that any
-complete restarts of the storage system to reprogram them is usually
-unacceptable.
-
-\item Policies and optimizations are usually hard-wired into the services and
-one has to be careful when factoring them to avoid introducing additional bugs.
-These policies and optimizations are usually cross-cutting solutions to
-concerns or trade-offs that cannot be fully explored at the time the code is
-written (as they relate to workload or hardware). Given these policies and
-optimizations, decomposition of otherwise orthogonal internal abstractions can
-be difficult or dangerous.
-
-\item Mechanisms that are often only exercised according to hard-wired policies
-and not in their full generality have hidden bugs that are revealed as soon as
-those mechanisms are governed by different policies. In our experience
-introducing programmability into a storage system proved to be a great
-debugging tool.
-
-\item Programmability, especially in live systems, implies changes that need to
-be carefully managed by the system itself, including versioning and propagation
-of those changes without affecting correctness.
-
-\end{itemize}
-
-To address these challenges we present Malacology
-\oldcomment{. Malacology is both a prototype for a programmable storage system
-and a}
-\newcomment{, our prototype programmable storage system. It uses the
-programmable storage}
-\addressesissue{2}
-design approach to evolve storage systems efficiently and
-without jeopardizing years of code-hardening and performance optimization
-efforts.  Although Malacology uses the internal abstractions of the underlying
-storage system, including its subsystems, components, and implementations, we
-emphasize that our system still addresses the general challenges outlined
-above.
-
-The main challenge of designing a programmable storage system is choosing the
-right internal abstractions and picking the correct layers for exposing them.
-\oldcomment{In this paper, we do not present an exhaustive list of possible
-internal abstractions nor do we contend that the abstractions we choose provide
-the best trade-offs for all applications.} 
-\newcomment{A programmable storage is not defined by what abstractions are
-exposed, rather a programmable storage system adheres to the design approach of
-choosing interfaces so administrators can have better control of the storage
-system. The interfaces presented in this paper are abstractions that we found
-useful for building ZLog and Mantle, yet they may not provide the best
-trade-offs for all higher-level services.}
-\addressesissue{2}
-For example, if consensus is correctly exposed one could implement
-high-level features like versioning, serialization, or various flavors of
-strongly consistent data management on top; but perhaps a low-level consensus
-interface is suited well for a particular set of applications.  These questions
-are not answered in this paper and instead we focus on showing the feasibility
-of building such a system, given advances in the quality and robustness of
-today's storage stacks.
-
-The Malacology prototype we present has been implemented on Ceph.
-While there are other systems on top of which Malacology could be implemented (see Table~\ref{table:examples}), we choose Ceph because it is a production quality system and because it
-is open source. The large developer community ensures that code is
-robust and the visibility of the code lets us expose any interface we want. In
-the next section we describe the Ceph components that we expose as Malacology
-interfaces.
-
-\section{Malacology: A Programmable Storage System}
-\label{sec:malacology}
-
-\begin{table*}
-\centering\small
-\begin{tabular}{ l | c | l | l | l }
-%\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
-\multicolumn{4}{c}{} \\
-\textbf{Interface}                      &
-\textbf{Sec}                            &
-\textbf{Example in Production Systems}  &
-\textbf{Example in Ceph}                &
-\textbf{Provided Primitives}            \\ \hline
-Service Metadata
-  & \S\ref{sec:service-metadata-interface}
-  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
-  & cluster state management~\cite{website:ceph-mon}
-  & consensus/consistency
-  \\
-Data I/O
-  & \S\ref{sec:data-io-interface}
-  & Swift in situ storage/compute~\cite{website:zerocloud}
-  & object interface classes~\cite{website:cls-lua}
-  & transaction/atomicity
-  \\
-Shared Resource
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & MPI collective IO, burst
-  & POSIX metadata protocols
-  & serialization/batching
-  \\
-File Type
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & ?????
-  & file striping strategy
-  & data/metadata access
-  \\
-Load Balancing
-  & \S\ref{sec:distributed-metadata-interfaces}
-  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
-  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
-  & migration/sampling
-  \\
-Durability
-  & \S\ref{sec:durability}
-  & S3/Swift interfaces (RESTful API)
-  & object store library~\cite{weil_rados_2007}
-  & persistence/safety
-  \\
-\end{tabular}
-\caption{Common internal abstractions. The same ``internal abstractions" are
-common across large-scale systems because they provide primitives that solve
-general distributed systems problems.  Here we list examples of what these
-internal abstractions are used for in ``production systems" and in Ceph.
-Malacology provides these internal abstractions as interfaces
-\newcomment{(Section~\ref{sec:malacology})}
-that higher level
-\oldcomment{applications}
-\newcomment{services (Section~\ref{sec:services})}
-can use.
-\oldcomment{; {\it italicized abstractions} are interfaces discussed in this
-paper.}}
-\addressesissue{1}
-\label{table:examples}
-\end{table*}
-
-The guiding principle is to re-use existing services and extend them so that
-these services can be \emph{programmed}. We accomplish programmability of a
-service by exporting bindings (or ``hooks'') for an interpreted programming
-language so that programming can occur without having to restart the storage
-system (see also below,~\S\ref{sec:durability}). Table~\ref{table:examples}
-shows the internal services from Ceph that we expose in the Malacology
-prototype via Lua~\cite{ierusalimschy1996lua} bindings
-\oldcomment{.}
-\newcomment{and Figure~\ref{fig:implementation-overview} compares what was
-already present in Ceph (gray boxes) to the Malacology interfaces we added.
-Section~\ref{sec:services} will describe the higher-level services we built
-with these interfaces.}
-\addressesissue{1}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation-overview.png}
-\caption{\newcomment{Malacology is implemented on the daemons and clients that
-run in a Ceph cluster. Interfaces expose internal subsystems and are used as
-building blocks for higher-level services.}\addressesissue{1}
-\label{fig:implementation-overview}}
-\end{figure}
-
-Lua is a portable embedded scripting language and we choose it as the
-interpreted language for Malacology because it offers superior performance and
-productivity trade-offs, including a JIT-based implementation that is well
-known for near native performance. Additionally, Lua has been used extensively
-in game engines, and systems research \cite{neto:dls14-luaos}, including
-storage systems where it has been effectively used both on
-\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010}
-and off \cite{sevilla:sc15-mantle} the performance critical path. Finally, the
-flexibility of the runtime allows execution sandboxing in order to address
-security and performance concerns. We will now discuss the common subsystems
-used to manage storage system and how Malacology makes them programmable.
-
-\subsection{Service Metadata Interface}
-\label{sec:mon}
-\label{sec:service-metadata-interface}
-\label{service-metadata}
-
-% straw man example
-Keeping track of state in a distributed system is an essential part of any
-successful service and a necessary component in order to diagnose and detect
-failures, when they occur. This is further complicated by variable propagation
-delays and heterogeneous hardware in dynamic environments.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In the case of Ceph, a consistent view of cluster state among server daemons
-and clients is critical to provide strong consistency guarantees to clients.
-Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
-metadata server) data structures called ``maps'' that record membership and
-status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
-responsible for integrating state changes into cluster maps, responding to
-requests from out-of-date clients and synchronizing members of the cluster
-whenever there is a change in a map so that they all observe the same system
-state. As a fundamental building block of many system designs, consensus
-abstractions such as Paxos are a common technique for maintaining consistent
-data versions, and are a useful system to expose.
-
-The default behavior of the monitor can be seen as a Paxos-based notification
-system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing
-clients to identify when new values (termed epochs in Ceph) are associated to
-given maps.  Since Ceph does not expose this service directly, as part of our Malacology implementation, we expose a key-value
-service designed for managing configuration metadata that is built on top of
-the consensus engine. Since the monitor is intended to be out of the
-high-performance I/O path, a general guideline is to make use of this
-functionality infrequently and to assign small values to maps.\\
-
-\noindent{\it{\textbf{Malacology: }}} Malacology exposes a strongly-consistent view of
-time-varying service metadata as a service rather than a hidden internal
-component. Malacology provides a generic API for adding arbitrary values to
-existing subsystem cluster maps. As a consequence of this, applications can
-define simple but useful service-specific logic to the strongly-consistent
-interface, such as authorization control (just specific clients can write new
-values) or triggering actions based on specific values (e.g. sanitize values).
-The higher level services we implement in \S\ref{sec:services} make use of this
-functionality to register, version and propagate dynamic code (Lua scripts) for
-new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
-policies in the load balancer \S\ref{malacology:mds}.  Using this service
-guarantees that interface definitions are not only made durable, but are
-transparently and consistently propagated throughout the cluster so that
-clients are properly synchronized with the latest interfaces.
-
-\newcomment{\\ \noindent \it{\textbf{Impact: }} provides core functionality
-because it lets daemons come to consensus on system critical state. Bugs in the
-internal subsystems or ommitting this from services that need this type of
-consistency affects correctness.}
-\addressesissue{3, 5}
-
-%And finally, it is important to not underestimate the importance of the
-%preservation of the modifications that enable a new system service. In many
-%(if not all) cases, interfaces defining access to data are just as important
-%as the data itself by virtue of inherently providing structural context. Thus,
-%all components of a system service must be kept to the same standards of
-%protection as data in the system itself.
-
-\oldcomment{\noindent 4.2 Object I/O Intefrace}
-\subsection{Data I/O Interface}
-\label{sec:data-io-interface}
-\label{object-data-interface}
-Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
-supports application-specific object
-interfaces~\cite{weil_rados_2007}. The ability to offload computation can
-reduce data movement, and transactional interfaces can significantly simplify
-construction of complex storage interfaces that require uncoordinated parallel
-access.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-An object interface is a plugin structured similarly to that of an RPC
-in which a developer creates a named function within the cluster that clients
-may invoke. In the case of Ceph each function is implicitly run within the
-context of an object specified when the function is called. Developers of
-object interfaces express behavior by creating a composition of native
-interfaces or other custom object interfaces, and handle serialization of
-function input and output.  A wide range of native interfaces are available to
-developers such as reading and writing to a byte stream, controlling object
-snapshots and clones, and accessing a sorted key-value database. These native
-interfaces may be transactionally composed along with application
-specific logic to create semantically rich interfaces. An example would be
-an interface that atomically updates a matrix stored in the bytestream and an
-index of the matrix stored in the key-value database.\\
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/implementation.png}
-\caption{Malacology allows users to dynamically define object/metadata
-interfaces by extending the object storage daemon (OSD) and metadata server
-(MDS) subsystems with an embedded Lua VM.  It uses the service metadata
-interface (M) to propagate interfaces/versions across the cluster
-\label{fig:implementation}}
-\end{figure}
-
-\noindent {\it{\textbf{Malacology: }}} The implementation of Ceph's object abstraction,
-although powerful, does not readily support programmability. Supporting only
-C/C++ for object interface developers, Ceph requires distribution of compiled
-binaries for the correct architecture, adding a large barrier of entry for
-developers and system administrators. Second, having no way to dynamically
-unload modules, any changes require a full restart of a storage daemon which
-may have serious performance impacts due to loss of cached data. And finally,
-the security limitations of the framework limit the use of object interfaces to
-all but those with administrative level access and deep technical expertise.
-
-To address these concerns, Malacology takes advantage of Lua extensions
-contributed by the Ceph community. This allows new object interfaces to be
-dynamically loaded into the system and modified at runtime, resulting in a
-object storage API with economy of expression, which at the same time provides
-the full set of features of the base object class. New object interfaces that
-are expressed in thousands of lines of code can be implemented in approximately
-an order of magnitude less code~\cite{geambasu_comet_2010}. While the use of
-Lua does not prevent deployment of malicious code, certain types of coding
-mistakes can be handled gracefully, and access policies are used to limit access
-to trusted users~\cite{ierusalimschy1996lua}.
-
-\oldcomment{All of these modifications work in tandem to implement the desired behavior,
-and are typically co-designed together such that each depends on the other to
-behave as expected.}
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by pushing behavior to lower parts of the storage stack, thereby
-minimizing hops and distributing computation.}
-\addressesissue{3, 5}
-
-\subsection{Distributed Metadata Interfaces}
-\begin{notes}
-\textcolor{red}{This section has been re-organized}
-\end{notes}
-\label{sec:distributed-metadata-interfaces}
-\label{malacology:mds}
-
-\oldcomment{The distributed metadata service in Ceph provides clients with a POSIX file
-system abstraction~\cite{weil:sc2004-dyn-metadata}.}
-\newcomment{File systems provide clients with the familiar POSIX file abstraction.
-While this guarantees strong consistency it comes at the cost of scalability,
-increased complexity, and lower performance.}
-\addressesissue{1}
-In general, distributed file systems protect resources by providing
-hierarchical indexing and distributed locking services.
-
-%The metadata cluster moves units of the namespace called directory fragments
-%and assigns them to servers using its own metadata load balancer. It then uses
-%a hard-coded policy to balance these fragments using a scalarized load metric
-%based on CPU, workload, and file system operation metrics. Although the CephFS
-%balancer has been shown to be complicated, its complexity is justifiable.
-%Having the flexibility to choose what to move, where to move it, and how much
-%to move is very powerful. The problem is that it is difficult to pick a
-%one-size-fits-all policy, system administrators may want to partition load
-%based on SLAs, or applications are better equipped to make these decisions,
-%etc. -- but we choose instead to expose a balancing API at strategic points in
-%the balancing logic using Lua hooks. ~\cite{sevilla:sc15-mantle}.
-
-\oldcomment{\it{\textbf{Malacology:}} Interfaces are added in strategic spots in the
-metadata service for guarding shared resources, defining new file types,
-and load balancing.}
-
-\oldcomment{\noindent Shared resource interface.}
-\subsubsection{Shared Resource Interface}
-
-In general, file system metadata servers manages client sessions, allowing
-clients to obtain locks (e.g. file byte ranges), and capabilities (e.g. to
-cache file data). Clients and metadata servers use a cooperative protocol in
-which clients voluntarily release resources back to the metadata service in
-order to implement sharing policies.
-
-\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
-In Ceph, the locking service implements a capability-based system that
-expresses what data and metadata clients are allowed to access as well as what
-state they may cache and modify locally.  While designed for the file
-abstraction, indexing, locking, and caching are all common services that are
-useful to a broad spectrum of applications.  Distributed applications that
-share centralized resources (e.g. a database or directory) face similar
-challenges which are often solved using application-specific sharding.
-
-\newcomment{\\ \noindent\it{\textbf{Malacology: }}}
-While the current policy for sharing access and voluntarily releasing resources
-is largely best-effort, Malacology supports generalized policies between
-metadata servers and clients that can be used to implement fairness or
-priority.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} provides core functionality
-to protect and provide exclusive access for any shared resource. May hurt
-performance if the resource in question does not require strong consistency.}
-\addressesissue{3}
-
-\oldcomment{\noindent File type interface.}
-\subsubsection{File Type Interface}
-The metadata service manages a POSIX file system hierarchy where files and
-directories are represented as inode data structures and expose a POSIX file
-interface. Applications that manage large amounts of metadata (e.g. users or
-database snapshots) often require a naming service. We allow new inode types to
-be defined such that applications can create domain-specific interfaces to
-inodes that may modify locking and capability policies. We will show how this
-is used in Section~\ref{sec:seq} when we discuss a distributed shared-log built
-on Malacology.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this interface is both a feature
-and a performance optimization. It is a feature because it allows developers to
-add support for different  storage types, such as how to read new file formats
-or what consistency semantics to use for a specific subtree in the hierarchical
-namespace. It is also a performance optimization because future programmers can
-add optimizations for processing specific types of files into the inode itself.}
-\addressesissue{3}
-
-%%% TODO: striping strategy is stored with the inode
-
-\oldcomment{\noindent \bf Load balancing interface.} 
-\subsubsection{Load Balancing Interface}
-
-Ceph also addresses the challenge of balancing metadata load with a separate
-metadata cluster. This cluster uses load balancing policies to migrate
-directory inodes around the cluster to alleviate overload at single
-nodes~\cite{weil:sc2004-dyn-metadata}. The policies use metrics based on system
-state (e.g.  CPU and memory utilization) and statistics collected by the
-cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
-partitioning to move variable sized namespace subtrees. These units can be
-shipped anywhere (i.e., to any metadata server of any capacity) at any time for
-any reason. The original balancer was designed with hard-coded policies and
-tunables.\\
-
-\noindent {\it{\textbf{Malacology: }}}
-Mantle~\cite{sevilla:sc15-mantle}, a programmable metadata load balancer, has
-been implemented on Malacology so that its infrastructure can enjoy all the
-properties of existing internal abstractions.  The load balancing logic is
-intact but the infrastructure has been improved to safely load, manage, and
-persist its load balancing policies.  Composing Mantle with the service
-metadata and durability interfaces, Mantle can safely version balancer
-policies, save balancer policies in the backend object store and centralize
-warnings/errors. When combined with file types, load balancing policies can
-express a variety of rules such as multi-tenant and multi-application
-balancing.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} helps applications optimize
-performance by allowing them to specify how to partition, replicate, and
-distribute metadata in response to overloaded servers.}
-\addressesissue{3}
-
-\subsection{Durability Interface}
-\label{sec:durability}
-
-Ceph provides storage by striping and replicating data across
-RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
-protects data using common techniques such as erasure
-coding, replication, and scrubbing.  Furthermore, many of these features
-are implemented using a peer-to-peer protocol that allows object storage
-devices to operate autonomously without a centralized coordinator.  For
-example, when the number of placement groups change, the OSDs re-balance and
-re-shard data in the background in a process called placement group splitting
-in which OSDs communicate directly with each other to cover a new state.
-In order to reduce load on the monitoring service, Ceph OSDs use a gossip
-protocol to efficiently propagate changes to cluster maps throughout the
-system, and autonomously initiate recovery mechanisms when failures are
-discovered.\\
-
-\noindent {\it{\textbf{Malacology: }}} Metadata service policies and object storage
-interfaces are stored durability within RADOS and managed by storing references
-with the object maps. Since the cluster already propagates a consistent view of
-these data structures, we use this service to automatically install interfaces
-in OSDs, and install policies within the metadata server daemon (MDS) such that clients and daemons are
-synchronized on correct implementations without restarting.
-
-\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this is a feature because it
-adds data safety and persistence to system metadata; while nice to have it does
-effect correctness.}
-\addressesissue{3}
-
-\section{Services Built on Malacology}
-\label{sec:services}
-\label{services-built-on-malacology}
-
-\label{services}
-%\begin{table}
-%\centering\small
-%\begin{tabular}{  l | l | l }
-%\textbf{Service}              &
-%\textbf{Used to...}           &
-%\textbf{Malacology Interface} \\ \hline
-%Mantle & load balance metadata     & Load Balancing  \\
-%       & agree on current balancer & Service Metadata\\
-%       & control client write lock & Shared Resource \\
-%       & persist policy logic      & Durability      \\ \hline
-%ZLog   & implement CORFU           & Object \& Data  \\ 
-%       & install ZLog interfaces   & Service Metadata\\
-%       & persist log entries       & Durability      \\ \hline
-%ZLog   & load balance sequencers   & Load Balancing  \\
-%Seqr.  & manage access to log tail & Shared Resource \\ 
-%\end{tabular}
-%\caption{Mantle and ZLog are built using Malacology interfaces so they inherit the
-%robustness of the subsystems of the underlying stack.}
-%\label{table:implementation}
-%\end{table}
-
-In this section we describe two services built on top of Malacology. The first
-is Mantle, a framework for dynamically specifying metadata load balancing
-policies. The second system, ZLog, is a high-performance distributed shared-log.
-In addition to these services, we'll demonstrate how we combine ZLog and Mantle
-to implement service-aware metadata load balancing policies.
-
-\subsection{Mantle: Programmable Load Balancer}
-\label{sec:mantle}
-
-Mantle is a programmable load balancer that separates the metadata balancing
-policies from their mechanisms. Administrators inject code to change how the
-metadata cluster distributes metadata. In ~\cite{sevilla:sc15-mantle} the
-authors showed how using Mantle, one can implement a single node metadata
-service, a distributed metadata service with hashing, and a distributed
-metadata service with dynamic subtree partitioning.
-
-The original implementation was ``hard-coded" into Ceph and lacked robustness
-(no versioning, durability, or policy distribution).  Re-implemented using
-Malacology, Mantle now enjoys (1) the versioning provided by the metadata
-service interface provided by the monitor daemons and (2) the durability and
-distribution provided by Ceph's reliable object store.  Re-using the internal
-abstractions with Malacology resulted in a 2\(\times\) reduction in source code
-compared to the original implementation.
-
-\subsubsection{Versioning Balancer Policies}
-
-Ensuring that the version of the current load balancer is consistent across the
-physical servers in the metadata cluster was not addressed in the original
-implementation. The user had to set the version on each individual server and
-it was trivial to make the versions inconsistent. Maintaining consistent
-versions is important for balancing policies that are cooperative in which local
-decisions are made assuming properties about other instances of the same
-policy.
-
-With Malacology, Mantle stores the version of the current load balancer in the
-Ceph service metadata interface. The version of the load balancer corresponds
-to an object name in the balancing policy. Using the service metadata interface
-ensures that all metadata servers use the same version of the load balancer. As
-a result, the policy version inherits the consistency of Ceph's internal
-monitor daemons. The user changes the version of the load balancer using a new
-CLI command.
-
-\subsubsection{Making Balancer Policies Durable}
-
-The load balancer version described above corresponds to the name of an object
-in RADOS that holds the actual Lua balancing code.  When MDS nodes start
-balancing load, they first check the latest version from the MDS map and
-compare it to the balancer they have loaded. If the version has changed, they
-dereference the pointer to the balancer version by reading the corresponding
-object in RADOS. This is in contrast to the original balancer which stored load
-balancer code on the local file system -- a technique which is unreliable and
-may result in silent corruption.
-
-The balancer pulls the Lua code from RADOS synchronously; asynchronous reads
-are not possible because of the architecture of the MDS. The synchronous
-behavior is not the default behavior for RADOS operations, so we achieve this
-with a timeout: if the asynchronous read does not come back within half the
-balancing tick interval the operation is canceled and a Connection Timeout
-error is returned. By default, the balancing tick interval is 10 seconds, so
-Mantle will use a 5 second second timeout.
-
-This design allows Mantle to immediately return an error if anything
-RADOS-related goes wrong.  We use this implementation because we do not want to
-do a blocking OSD read from inside the global MDS lock. Doing so would bring
-down the MDS cluster if any of the OSDs are not responsive.
-
-Storing the balancers in RADOS is simplified by the use of an interpreted
-language for writing balancer code. If we used a language that needs to be
-compiled, like the C++ object classes in the OSD, we would need to ensure
-binary compatibility, which is complicated by different operating systems,
-distributions, and compilers.
-
-\subsubsection{Logging, Debugging, and Warnings}
-
-In the original implementation, Mantle would log all errors, warnings, and
-debug messages to a log stored locally on each MDS server. To get the simplest
-status messages or to debug problems, the user would have to log into each MDS
-individually, look at the logs, and reason about causality and ordering.
-
-With Malacology, Mantle re-uses the centralized logging features of the
-monitoring service. Important errors, warnings, and info messages are collected
-by the monitoring subsystem and appear in the monitor cluster log so instead of
-users going to each node, they can watch messages appear at the monitor.
-Messages are logged sparingly, so as not to spam the monitor with frivolous
-debugging but important events, like balancer version changes or failed
-subsystems show up in the centralized log.
-
-\subsection{ZLog: A Fast Distributed Shared Log}
-\label{sec:zlog}
-
-The second service implemented on Malacology is ZLog, a high-performance
-distributed shared-log that is based on the CORFU
-protocol~\cite{balakrishnan_corfu_2012}. The shared-log is a
-powerful abstraction used to construct distributed systems, such as
-metadata management~\cite{balakrishnan:sosp13} and elastic database
-systems~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
-However, existing implementations that rely on consensus algorithms such as
-Paxos funnel I/O through a single point introducing a bottleneck that restricts
-throughput. In contrast, the CORFU protocol is able to achieve high throughput
-using a network counter called a \emph{sequencer}, that
-decouples log position assignment from log I/O.
-
-While a full description of the CORFU system is beyond the scope of this
-paper, we briefly describe the custom storage device interface, sequencer
-service, and recovery protocol, and how these services are instantiated in the
-Malacology framework.
-
-\subsubsection{Sequencer}
-\label{sec:seq}
-
-High-performance in CORFU is achieved using a sequencer service that assigns
-log positions to clients by reading from a volatile, in-memory counter which
-can run at a very high throughput and at low latency. Since the sequencer is
-centralized, ensuring serializability in the common case is trivial.  The
-primary challenge in CORFU is handling the failure of the sequencer in a way
-that preserves correctness. Failure of the sequencer service in CORFU is
-handled by a recovery algorithm that recomputes the new sequencer state using a
-CORFU-specific custom storage interface to discover the tail of the log, while
-simultaneously invalidating stale client requests using an epoch-based
-protocol.
-
-{\bf Sequencer interface.} The sequencer resource supports the ability to
-\emph{read()} the current tail value and get the \emph{next()} position in
-the log which also atomically increments the tail position.
-We implement the sequencer service in Malacology as an application-specific
-file type that associates a small amount of metadata
-with the inode state managed by the Ceph metadata service.
-This approach has the added
-benefit of allowing the metadata service to handle naming, by
-representing each sequencer instance in the standard POSIX hierarchical
-namespace. The primary challenge in mapping the sequencer resource to the
-metadata service is handling serialization correctly to maintain the global
-ordering provided by the CORFU protocol.
-
-Initially we sought to directly model the sequencer service in Ceph as a
-non-exclusive, non-cacheable resource, forcing clients to perform a round-trip
-access to the resource at the authoritative metadata server for the sequencer
-inode.  Interestingly, we found that the capability system in Ceph uses a
-strategy to reduce metadata service load by allowing clients that open a
-shared file to temporarily obtain an exclusive cached copy of the resource,
-resulting in a round-robin, best-effort batching behavior. When a single
-client is accessing the sequencer resource it is able to increment the
-sequencer locally. Any competing client cannot query the sequencer until the
-metadata service has granted it access.
-
-While unexpected, this discovery allowed us to explore an implementation
-strategy that we had not previously considered. In particular, for bursty
-clients, and clients that can tolerate increased latency, this mode of
-operation may allow a system to achieve much higher throughput than a system
-with a centralized sequencer service.  We utilize the programmability of the
-metadata service to define a new policy for handling capabilities that controls
-the amount of time that clients are able to cache the sequencer resource. This
-allows an administrator or application to control the trade-off between latency
-and throughput beyond the standard best-effort policy that is present in Ceph
-by default.
-
-In Section~\ref{sec:evaluation} we quantify the trade-offs of throughput and
-latency for an approach based on a round-robin batching mode, and compare this
-mode to one in which the metadata server mediates access to the sequencer state
-when it is being shared among multiple clients. Quantifying these trade-offs
-should provide administrators with guidelines for setting the tunables for
-different ``caching" modes of the sequencer.
-
-{\bf Balancing policies.} As opposed to the batching mode for controlling access to the sequencer
-resource, more predictable latency can be achieved by treating the sequencer
-inode as a shared non-cacheable resource, forcing clients to make a round-trip
-to the metadata service. However, the shared nature of the metadata service
-may prevent the sequencer from achieving maximum throughput. To provide a
-solution to this issue we have taken advantage of the programmability of the
-metadata service load balancing to construct a service-specific load balancing
-policy. As opposed to a balancing policy that strives for uniform
-load distribution, a ZLog-specific policy may utilize knowledge of inode types
-to migrate the sequencer service to provisioned hardware during periods of
-contention or high demand.
-
-\subsubsection{Storage Interface}
-
-The storage interface is a critical component in the CORFU protocol. Clients
-independently map log positions that they have obtained from the sequencer
-service (described in detail in the next section) onto storage devices, while
-storage devices provide an intelligent \emph{write}-once, random \emph{read}
-interface for accessing log entries. The key to correctness in CORFU lies with
-the enforcement of up-to-date epoch tags on client requests; requests tagged
-with out-of-date epoch values are rejected, and clients are expected to
-request a new tail from the sequencer after refreshing state from an auxiliary
-service.  This mechanism forms the basis for sequencer recovery.
-
-In order to repopulate the sequencer state (i.e. the cached, current tail of
-the log) during recovery of a sequencer, the maximum position in the log must
-be obtained. To do this, the storage interface exposes an additional
-\emph{seal} method that atomically installs a new epoch value and returns the
-maximum log position that has been written.
-
-Since the sequencer service does not resume until the recovery process has
-completed, there cannot be a race with clients appending to the log, and the
-immutability of the log allow reads to never block during a sequencer failure.
-Recovery of the sequencer process itself may be handled in many ways, such as
-leader election using an auxiliary service such as Paxos. In our
-implementation, the recovery is the same as (and is inherited from) the Ceph
-metadata service. Handling the failure of a client that holds the sequencer
-state is similar, although a timeout is used to determine when a client should
-be considered unavailable.
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-%We wanted to implement CORFU and we had two choices: build it from the ground
-%up or do it on top of Malacology. To show the feasability of building on
-%Malacology our evaluation focuses on the
-
-Our evaluation demonstrates the feasibility of building new service
-abstractions atop programmable storage, focusing on the performance of the
-internal abstractions exposed by Malacology and used to construct the Mantle
-and ZLog services. We also discuss latent capabilities we discovered in this
-process that let us navigate different trade-offs within the services
-themselves. First, we benchmark scenarios with high sequencer contention by
-examining the interfaces used to map ZLog onto Malacology; specifically, we
-describe the sequencer implementation and the propagation of object and data
-interfaces interfaces.  Next, we benchmark scenarios in which the storage
-system manages multiple logs by using Mantle to balance sequencers across a
-cluster.
-
-% should save this for future work, otherwise it undercuts the eval.
-%Detailed system-wide performance measurements and analysis is forthcoming for
-%both Mantle and ZLog, as we try to find the best metadata balancing policies
-%and physical design implementation trade-offs.
-
-Since this work focuses on the programmability of Malacology, the goal of this
-section is to show that the components and subsystems that support the
-Malacology interfaces provide reasonable relative performance, as well as to
-give examples of the flexibility that Malacology provides to programmers.
-
-%As we search for the best metadata balancing policies
-%and physical design implementation trade-offs,
-%we plan to rel
-
-\subsection{Mapping ZLog onto Malacology}
-\label{sec:mapping-zlog-onto-malacology}
-
-We evaluate Malacology by exploring one possible mapping of the ZLog
-implementation of CORFU onto Ceph in which we re-use (1) the metadata service
-to manage naming and synchronization of the sequencer resource by treating the
-resource as an inode, and (2) the monitoring sub-subsystem to distribute and
-install application-specific I/O interface required of the CORFU protocol. In
-Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the inode
-abstraction for implementing the sequencer resource enables load balancing policies
-to migrate the sequencer resource in heavy-load situation.
-
-\subsubsection{Sequencer Implementation}
-\label{sec:sequencer-implementation}
-
-We evaluate the feasibility of using the metadata service to implement a
-sequencer resource that is responsible for maintaining a total ordering of
-the log. Clients contact the sequencer to obtain the tail of the log and then
-independently initiate I/O, thus we measure both the throughput and latency of
-obtaining new tail positions which bounds client append performance.
-
-The sequencer is implemented as a custom file type in which the sequencer
-state (a 64-bit integer) is embedded in the inode of a file representing the
-sequencer. A total ordering of the log is imposed by the re-use of the
-capability service that can be used to grant exclusive access of inode state
-to clients. The metadata service is responsible for maintaining exclusivity
-and granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
-behavior of the system in which a best-effort policy is used.  The two colors
-represent points in time that the clients were able to access the resource.
-The best-effort policy shows a high degree of interleaving
-between clients but the system spends a large portion of time re-distributing
-the capability, reducing overall throughput.
-
-In order to control the performance of the system we implement a
-policy that (1) restricts the length of time that a client may maintain
-exclusive access and (2) limits the number of log positions that a client may
-generate without yielding to other clients waiting for access. The behavior of
-these two modes is illustrated in Figures~\ref{fig:capdelay-quota-behavior}
-(b) and (c), respectively.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/capdelay-quota-behavior-a.png}
-\includegraphics{figures/capdelay-quota-behavior-b.png}
-\includegraphics{figures/capdelay-quota-behavior-c.png}
-\caption{Each dot is an individual request, spread randomly along the \(y\)
-axis. The default behavior is unpredictable, ``delay" lets clients hold the
-lease longer, and ``quota" gives clients the lease for a number of operations.}
-\label{fig:capdelay-quota-behavior}
-\end{figure}
-
-Figure~\ref{fig:captp} demonstrates a configurable trade-off between
-throughput and latency. In the experiment two clients are run each with a
-fixed 0.25 second maximum reservation on the capability, and we vary the size
-of the log position quota running each configuration for two minutes. The
-total operations per second is the combined throughput of the two clients, and
-the average latency is number of microseconds required to obtain a new log
-position. With a small quota more time is spent exchanging exclusive access,
-while a large quota reservation allows clients to experience a much lower
-latency because they experience uncontended access for a longer period of time.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/tradeoff.png}
-\caption{Sequencer throughput by re-using various services.  The highest performance is
-achieved using a single client with exclusive, cacheable privilege. Round-robin
-sharing of the sequencer resource is affected by the amount of time the
-resource is held, with best-effort performing the worst.}
-\label{fig:captp}
-\end{figure}
-
-To get a better picture of latency, Figure~\ref{fig:capcdf} shows the CDF of
-latency for each client in all experiment configurations. At the 99th
-percentile clients accessed the sequencer is less than a millisecond. The CDF
-is cropped at the 99.999th percentile due to large outliers that we believe
-occur in instances in which the MDS is performing I/O while it is in the
-process of re-distributing the capability to another client.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics{figures/caps-delay-latency.png}
-\caption{
-The latency distribution shows that if the clients hold their capabilities longer,
-performance decreases. Changing the delay can help the sequencer and
-clients strike a balance between throughput and latency.}
-\label{fig:capcdf}
-\end{figure}
-
-\textbf{Malacology exposes} the internal capability management service and
-allows users to navigate latency and throughput trade-offs.  Other approaches to
-designing the sequencer service also exist, such as using a centralized service
-in which each access is a network round-trip. In contrast to the mechanism we
-explored which is appropriate for clients with bursty workloads, it may be
-easier to provide predictable performance using a centralized service, and we
-will be exploring in the future how this can be achieved using the capability
-system.
-
-\subsubsection{Interface Propagation}
-
-Domain-specific data interfaces (Section~\ref{sec:application-specifc-storage-stacks}) allow
-co-design between applications and the storage system. Malacology supports
-custom object interfaces in RADOS that require interface implementations to
-be installed on the storage devices in the system, supporting the evolution of
-interfaces through automatic system-wide versioning and installation through
-the service metadata interface (Section~\ref{service-metadata}). We evaluate
-the performance of installing a new interface version in the cluster,
-which is an important metric for applications that frequently evolve
-interfaces.
-
-We demonstrate the feasibility of utilizing the Ceph monitoring sub-system by
-evaluating the performance of installing and distributing interface updates.
-Figure~\ref{fig:propdelay} shows the CDF of the latency of interface updates.
-The interfaces are Lua scripts embedded in the cluster map and distributed
-using a peer-to-peer gossip protocol.  The latency is defined as the elapsed
-time following the Paxos proposal for an interface update until each OSD makes
-the update live (the cost of the Paxos proposal is configurable and is
-discussed below). The latency measurements were taken on the nodes running
-Ceph server daemons, and thus exclude the client round-trip cost. In each of
-the experiments 1000 interface updates were observed.
-
-Figure~\ref{fig:propdelay} shows the lower bound cost for updates in a large
-cluster. In the experiment labeled ``120 OSD (RAM)" a cluster of 120 OSDs (10
-OSDs x 12 servers) using an in-memory data store were deployed, showing a
-latency of less than 54 ms with a probability of 90\% and a worst case latency
-of 194 ms. These costs demonstrate the penalty of distributing the interface
-in a large cluster. In practice the costs include, in addition to cluster-wide
-propagation of interface updates, the network round-trip to the
-interface management service, the Paxos commit protocol itself, and other factors
-such as system load. By default Paxos proposals occur periodically with a 1
-second interval in order to accumulate updates. In a minimum, realistic quorum
-of 3 monitors using HDD-based storage, we were able to decrease this interval
-to an average of 222 ms.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=60mm,trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
-\caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
-interface commit.}
-\label{fig:propdelay}
-\end{figure}
-
-\subsection{Load Balancing ZLog Sequencers with Mantle}
-\label{sec:zlog-balancing}
-
-In practice, a storage system implementing CORFU will support a multiplicity of
-independent totally-ordered logs for each application.  For this scenario
-co-locating sequencers on the same physical node is not ideal but building a load
-balancer that can migrate the shared resource ({\it e.g.,}, the resource that
-mediates access to the tail of the log) is a time-consuming, non-trivial task.
-It requires building subsystems for migrating resources, monitoring the
-workloads, collecting metrics that describe the utilization on the physical
-nodes, partitioning resources, maintaining cache coherence, and managing
-multiple sequencers. The following experiments demonstrate the feasibility of
-using the mechanisms of the Malacology load balancing interface to inherit
-these features and to alleviate load from overloaded servers.
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-behaviors.png}
-\caption{CephFS/Mantle load balancing have better throughput than co-locating
-all sequencers on the same server.  Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify this improvement;
-Section~\ref{sec:feature-backoff} examines the migration at 0-60 seconds.
-}\label{fig:mantle-balancer-behaviors}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-balancer-performance.png}
-\includegraphics{figures/mantle-mode-performance.png}
-\caption{In (a) all CephFS balancing modes have the same performance; Mantle uses a
-balancer designed for sequencers. In (b) the best combination of mode and
-migration units can have up to a 2\(\times\) improvement.}\label{fig:mantle-balancer-performance}
-\end{figure}
-
-The experiments are run on a cluster with 10 nodes to store objects, one node
-to monitor the cluster, and 3 nodes that can accommodate sequencers. 
-Instead of measuring contention at the clients like
-Section~\ref{sec:sequencer-implementation}, these experiments measure
-contention at the sequencers by forcing clients to make round-trips for every
-request. We implement this using the shared resource interface that forces
-round-trips.  Because the sequencer's only function is to  hand out positions
-for the tail of the log, the workload is read-heavy.
-
-First, we show how the ZLog service can orchestrate multiple sequencers using
-the Malacology load balancing interface.
-Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput over time of
-different load balancers as they migrate 3 sequencers (with 4 clients) around
-the cluster; ``No Balancing" keeps all sequencers on one server, ``CephFS"
-migrates sequencers using the CephFS load balancers, and ``Mantle" uses a
-custom load balancer we wrote specifically for sequencers. The CephFS load
-balancers are hard-coded into Ceph while Mantle is an API for balancing load
-presented in~Mantle~\cite{sevilla:sc15-mantle} and re-implemented on top of
-Malacology in this paper.  The increased throughput for the CephFS and Mantle
-curves between 0 and 60 seconds are a result of migrating the sequencer(s) off
-overloaded servers.
-
-In addition to showing that migrating sequencers improves performance,
-Figure~\ref{fig:mantle-balancer-behaviors} also demonstrates features that we
-will explore in the rest of this section.
-Sections~\ref{sec:feature-balancing-modes}
-and~\ref{sec:feature-migration-units} quantify the differences in performance
-when the cluster stabilizes at time 100 seconds and
-Section~\ref{sec:feature-backoff} examines the slope and start time of the
-rebalancing phase between 0 and 60 seconds by comparing the aggressiveness of
-the balancers.
-
-\subsubsection{Feature: Balancing Modes}
-\label{sec:feature-balancing-modes}
-
-Next, we quantify the performance benefits shown in
-Figure~\ref{fig:mantle-balancer-behaviors}.  To understand why load balancers
-perform differently we need to explain the different balancing modes that the
-load balancer service uses and how they stress the subsystems that receive and
-forward client requests in different ways. In
-Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
-performance of the balancing mode that CephFS falls into {\it most of the time}.
-CephFS has 3 modes for balancing load: CPU mode, workload mode, and hybrid
-mode. All three have the same structure for making migration decisions but vary
-based on the metric used to calculate load. For this sequencer workload the 3
-different modes all have the same performance, shown in
-Figure~\ref{fig:mantle-balancer-performance} (a), because the load balancer falls
-into the same mode a majority of the time.  The high variation in performance
-for the CephFS CPU Mode bar reflects the uncertainty of using something as
-dynamic and unpredictable as CPU utilization to make migration decisions. In
-addition to the suboptimal performance and unpredictability, because
-CephFS balancers all fall into the same mode despite using different balancers
-is a problem because it prevents administrators from properly exploring the
-balancing state space.
-
-Mantle gives the administrator more control over balancing policies; for
-the Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the load
-balancing interface to program logic for balancing read-heavy workloads,
-resulting in better throughput and stability.  When we did this we also
-identified two balancing modes relevant for making migration decisions
-for sequencers.
-
-Using Mantle, the administrator can put the load balancing service into ``proxy
-mode" or ``client mode". In proxy mode one server receives all requests and
-farms off the requests to slave servers; the slaves servers do the actual tail
-finding operation. In client mode, clients interact directly with the server
-that has their sequencer.  These modes are illustrated in
-Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
-co-located on one physical server -- performance for that mode is shown by the
-No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
-Mode" in the lower left shows how clients for sequencer 2 are re-directed to
-server B. The two ``Proxy mode" diagrams on the right show how clients continue
-sending requests to server A even though sequencer 2 has migrated to server B
-(the meaning of ``(Half)" is discussed in the next section).
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-modes.png}
-\caption{ In client mode clients sending requests to the server that houses
-their sequencer. In proxy mode clients continue sending their requests to the
-first server.  }\label{fig:mantle-modes}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figures/mantle-mode-behavior.png}
-\caption{The performance of proxy mode achieves the highest throughput but at
-the cost of lower throughput for one of the sequencers. Client mode is more
-fair but results in lower cluster throughput.
-}\label{fig:mantle-mode-behavior}
-\end{figure}
-
-Figure~\ref{fig:mantle-mode-behavior} shows the throughput over time of the two
-different modes for an environment with only 2 sequencers (again 4 clients
-each) and 2 servers. The curves for both sequencers in
-Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second and
-at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
-Performance of Sequencer 2 decreases because it stayed on the proxy which now
-handles all client requests, processes requests for Sequencer 2, and forwards
-requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
-because distributing the sequencers in this way separates (1) the handling of
-the client requests and (2) finding the tail of the log and responding to
-clients.  Doing both steps is too heavy weight for one server and sequencers on
-slave nodes can go faster if work is split up; this phenomenon is not
-uncommon and has been observed in chain replication~\cite{chain_rep}.
-
-Cluster
-throughput improves at the cost of decreased throughput for Sequencer 2.
-Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer mode manually (no
-balancing phase) and shows that the cluster throughput is worse than the
-cluster throughput of proxy mode. That graph also shows that Sequencer 2 has
-less throughput than Sequencer 2. In this case, the scatter-gather process used
-for cache coherence in the metadata protocols causes strain on the server
-housing Sequencer 2 resulting in this uneven performance.
-
-\subsubsection{Feature: Migration Units}
-\label{sec:feature-migration-units}
-
-Another factor that affects performance in this environment is how much load
-is on each server; these experiments quantify that effect by programming the
-load balancing interface to control the amount of load to migrate. We call this
-metric a ``migration unit".  Expressing this heuristic is not easily achievable
-with outward facing tunable parameters (i.e. system knobs) but with Mantle's
-programmable interface, we can force the load balancer to change its migration
-units. To force the balancer into the Proxy Mode (Half) scenario in
-Figure~\ref{fig:mantle-modes}, which uses migration units equal to half the
-load on the current server, we can use: \texttt{ targets[whoami+1] = mds[whoami]["load"]/2
-}.
-
-This code snippet uses globally defined variables and tables from the Mantle
-API to send half of the load on the current server (whoami) to the next ranked
-server (whoami + 1); the \texttt{targets} array is a globally defined table that
-the balancer uses to do the migrations. Alternatively, to migrate all load a
-time step, we can remove the division by 2.
-
-%\begin{figure}[t!]
-%\centering
-%\includegraphics{figures/mantle-mode-performance.png}
-%\caption{The performance of the different modes using various migration units
-%shows almost a 2\(\times\) improvement in the best case.
-%}\label{fig:mantle-mode-performance}
-%\end{figure}
-
-Figure~\ref{fig:mantle-balancer-performance} (b) shows the performance of the modes
-using different migration units. Recall that this setup only has 2 sequencers
-and 2 servers, so performance may be different at scale. Even so, it is clear
-that client mode does not perform as well for read-heavy workloads. We even see
-a throughput improvement when migrating all load off the first server, leaving
-the first server to do administrative tasks (this is common in the metadata
-cluster because the first server does a lot of the cache coherence work) while
-the second server does all the client work. Proxy mode does the best in both
-cases and shows large performance gains when completely decoupling client
-request handling and operation processing in Proxy Mode (Full).  The parameter
-that controls the migration units helps the administrator control the sequencer
-co-location or distribution across the cluster. This trade-off was explored
-extensively in the Mantle paper but the experiments we present here are
-indicative of an even richer set of states to explore.
-
-\subsubsection{Feature: Backoff}
-\label{sec:feature-backoff}
-
-Tuning the aggressiveness of the load balancer decision making is also a
-trade-off that administrators can control and explore. The balancing phase from
-0 to 60 seconds in Figure~\ref{fig:mantle-balancer-behaviors} shows different
-degrees of aggressiveness in making migration decisions; CephFS makes a
-decision quickly and throughput jumps to 2500 ops/second within 10 seconds
-while Mantle takes more time to stabilize. This conservative behavior is
-controlled by programming the balancer to (1) use different conditions for when
-to migrate and (2) using a threshold for sustained overload.
-
-We control the conditions for when to migrate using \texttt{when()}, a callback
-in the Mantle API.  For the Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} we program \texttt{when()} to wait
-for load on the receiving server to fall below a threshold. This makes the
-balancer more conservative because it takes 60 seconds for cache coherence
-messages to settle.  The Mantle curve in
-Figure~\ref{fig:mantle-balancer-behaviors} also takes longer to reach peak
-throughput because we want the policy to wait to see how migrations affect the
-system before proceeding; the balancer does a migration right before 50
-seconds, realizes that there is a third underloaded server, and does another
-migration.
-
-The other way to change aggressiveness of the decision making is to program
-into the balancer a threshold for sustained overload. This forces the balancer
-to wait a certain number of iterations after a migration before proceeding. In
-Mantle, the policy would use the save state function to do a countdown after a
-migration.  Behavior graphs and performance numbers for this backoff feature is
-omitted for space considerations, but our experiments confirm that the more
-conservative the approach the less overall throughput.
-
-\textbf{Malacology pulls} the load balancing service out of the storage system
-to balance sequencers across a cluster. This latent capability also gives
-future programmers the ability to explore the different load balancing
-trade-offs including: load balancing modes to control forwarding vs. client
-redirection, load migration units to control sequencer distribution vs.
-co-location, and backoffs to control conservative vs. aggressive decision
-making.
-
-%\subsubsection{Experiment: Object Class
-%Programmability}\label{experiment-object-class-programmability}
-%
-%Add quote from CORFU paper that talks about what they had to hack
-%
-%\subsubsection{Experiment: Multi-Client
-%Burstiness}\label{experiment-multi-client-burstiness}
-%
-%\begin{figure}[htbp] \centering
-%\includegraphics{figures/caps-delay-thruput.png} \caption{Forcing the client
-%to drop their capabilities later (delay) improves throughput} \end{figure}
-
-\section{Related Work}
-
-% IOStack: Software-Defined Object Storage
-% sRoute: Treating the Storage Stack Like a Network
-   % - lots of good refs to check out
-% Information and control in gray-box systems
-
-Programmability of operating systems and networking resources, including
-distributed storage systems is not new, but we are not aware of work that makes
-generalization of existing services into programmable resources a key principle
-in storage systems design.
-
-Programmable storage systems can be viewed as an infrastructure for creating
-abstractions to better separate policies from mechanisms. This idea is not new.
-Software-defined networks (SDNs) create such an abstraction by separating the
-control plane from the data plane (see for example~\cite{jain:sigcomm13}). This
-notion of control/data separation was also applied in software-defined storage
-(SDS)~\cite{thereska:sosp13,stefanovici:fast16}. Similarly,
-IOStack~\cite{gracia:internet16} is providing policy-based provisioning and
-filtering in OpenStack Swift. According to a SNIA white
-paper~\cite{carlson:snia2014}, the primary goal of SDS is to control and
-facilitate flexible and dynamic provisioning of storage resources of different
-kinds, including flash memory and disk drives, to create a virtualized mapping
-between common storage abstractions (e.g. files, objects, and blocks) and
-storage devices taking data service objectives in terms of protection,
-availability, performance, and security into account. A programmable storage
-system exposes internal abstractions so that end users (not necessarily
-operators) can create new services on top of the storage stack. Thus, our
-notion of programmable storage differs from ``software-defined storage'' (SDS)
-in terms of goals and scope, although definitions of SDS are still in flux.
-
-Another view of programmable storage systems is one of tailoring systems
-resources to applications~\cite{arpaci:sosp01}. Related efforts include the
-Exokernel~\cite{engler:sosp95}, SPIN~\cite{bershad:sosp95} and
-Vino~\cite{seltzer:osdi96} projects; the latter two addressed the ability of injecting
-code into the kernel to specialize resource management. Another approach is to
-pass hints between the different layers of the IO stack to bridge the semantic
-gap between applications and
-storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}).
-
-Malacology uses the same Active and Typed Storage module presented in
-DataMods~\cite{watkins_datamods_2012}; Asynchronous Service and File Manifolds
-can be implemented with small changes to the Malacology framework, namely
-asynchronous object calls and Lua stubs in the inode, respectively.
-
-\section{Conclusion and Future Work}
-\label{conclusion-and-future-work}
-
-Programmable storage is a viable method for eliminating duplication of complex
-error-prone software used as workarounds for storage system
-deficiencies. We propose that systems expose their services in a safe way
-allowing application developers to customize system behavior to meet their
-needs while not sacrificing correctness. To illustrate the benefits of this
-approach we presented Malacology, a programmable storage system that
-facilitates the construction of new services by re-purposing existing subsystem
-abstractions of the storage stack.
-
-Future work will focus on constructing more interfaces to support a wide
-variety of storage system services that can be configured on-the-fly in existing
-systems. This work is one point along that path to producing general-purpose
-storage systems that can
-target  special-purpose applications.  Ultimately we want to utilize
-declarative methods for expressing new services.
+  \input{./0_abstract}
+  \input{./1_intro}
+  \input{./2_application_specific_storage_stacks}
+  \input{./3_challenges}
+  \input{./4_malacology_a_programmable_storage_system}
+  \input{./5_services_built_on_malacology}
+  \input{./6_evaluation}
+  \input{./7_future_work}
+  \input{./8_related_work}
+  \input{./9_conclusion}
 
@@
 \bibliographystyle{plain}
-%\printbibliography
 

### 2017-02-23 paper/4_malacology_a_programmable_storage_system.tex
paper: integrate Shel's future work comments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 traditional file system metadata which is information about files. For the rest
-of the paper when we refer to metadata server or service, we mean a daemon that
-manages file system metadata.} \addressesissue{4}
+of the paper when we use the phrase ``metadata server" or ``metadata service",
+we are referring to the daemon(s) that manages file system metadata.}
+\addressesissue{4}
 
@@
 install interfaces in OSDs, and install policies within the metadata server
-daemon \oldcommentone{(MDS)} such that clients and daemons are synchronized on
+daemon (MDS) such that clients and daemons are synchronized on
 correct implementations without restarting.

### 2017-02-23 paper/7_future_work.tex
paper: integrate Shel's future work comments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\newcommentone{This work is one point along the path towards producing
-general-purpose storage systems that can target special-purpose applications.
-Continuing to expore existing storage system functionality as reusable building
-blocks will enable the proliferation of storage-centric APIs, allowing
-developers to custom-fit their applications to take advantage of the
-capabilities of the underlying system. When adding new interfaces and composing
-them together, careful consideration should be made to address security, access
-control, and safety. Correctness and performance will surely be affected if
-interfaces are combined poorly and these drawbacks will be quantified.}
-\addressesissue{5}
+\newcommentone{Malacology is a first step towards showing how general-purpose
+storage systems can be adapted to target special-purpose applications.  By
+encapsulating storage system functionality as reusable building blocks, we
+enable application developers to leverage storage capabilities based on
+interfaces that are proven and understandable.  However, creation and
+composition of interfaces is complex; constructs must be combined safely in
+order to provide correctness, performance and security. We will study
+additional Malacology-based services in order to learn techniques that support
+safe composition.}\addressesissue{5}
 
-\newcommentone{The next higher-level services we plan to build using the
+\newcommentone{Some higher-level services that we plan to build using the
 interfaces in Table~\ref{table:examples} are: an elastic cloud database, a data
-layout manager, and a data processing engine. Some cloud databases aggressively
-cache data to hide the overheads and weaker consistency semantics of the
-distributed file systems they sit on~\cite{pavlo:sigmod16} but using Malacology
-we could build a database that works well because of the storage system not in
-spite of it. We plan to use the data IO interface for push down predicates, the
-file type interface for maintaining indices and stats, and the durability
-interface internals for ingest and data movement. The data layout manager would
-be similar to DRepl~\cite{ionkov:msst2013} and would use the file type,
-durability, and data I/O interface to manage different views of data. Finally,
-the data processing engine would be similar to MapReduce and would use the data
-I/O, shared resource, and load balancing interface to implement the resource
-manager, computation engine, and storage platform in the Hadoop ecosystem.}
+processing engine, and a data layout manager.  The elastic database could use
+the data I/O interface to push down predicates, the file type interface to
+maintain access paths and statistics efficiently, and the durability interface
+to manage ingestion and movement.  The data processing engine would be similar
+to MapReduce, using the data I/O interface to send computation to the data, the
+load balancing interface as a proxy for farming off work, and the durability
+interface for persisting to input or output sinks.  Finally, the data layout
+manager would be similar to DRepl~\cite{ionkov:msst2013}, using the file type,
+durability, and data I/O interfaces to provide multiple views of data.}
 
-\newcommentone{ Our experience with Zlog and Mantle illustrates that the labor
-of wrapping existing services in interfaces is justified by the power and
-flexibility they afford to programmers.  In exchange for this flexibility,
-however, programmable storage systems force programmers to forfeit the
-\emph{protection from change} afforded by narrow storage interfaces like the
-POSIX API.  To implement applications built atop systems such as Malacology,
-programmers must navigate a complex design space, simultaneously addressing
-often orthogonal concerns including functional correctness, performance and
-fault tolerance. Worse still, these compositions can be sensitive to changes in
-the underlying environment (such as hardware upgrades, software version changes
-and evolving workloads).  For example, a major version change in Ceph required
-us to rewrite significant parts of Zlog to maintain acceptable performance --
-each such evolution costs developer time and risks introducing bugs.}
+\newcommentone{ Our experience with ZLog and Mantle demonstrates that the labor
+of wrapping existing services in reusable interfaces is justified by the power
+and flexibility that this encapsulation affords to programmers.  In exchange
+for this flexibility, however, programmers may forfeit the \emph{protection
+from change} afforded by narrow storage interfaces such as the POSIX API.  To
+implement applications on programmable storage systems such as Malacology,
+programmers must find solutions by navigating a complex design space,
+simultaneously addressing functional correctness, performance and fault
+tolerance. Worse still, their solutions may be sensitive to changes in the
+underlying environment, such as hardware upgrades, software version changes and
+evolving workloads.  For example, a major version change in Ceph required us to
+rewrite significant parts of ZLog to maintain acceptable performance. Each such
+evolution costs developer time and risks introducing bugs.}
 
@@
 \emph{declarative} languages based on Datalog~\cite{alvaro:cidr11} to program
-storage APIs.  In this way, a programmer can specify the functional behavior of
-a storage interface once in a relational (or algebraic) language, allowing an
-optimizer to model search the space of functionally equivalent physical
-implementations, and re-optimizing when characteristics change.  Much like
-query planning and optimization in database systems~\cite{hellerstein:cidr15},
-this approach will separate the concerns of correctness and performance, and
-protect applications (which evolve slowly) against changes in highly dynamic
-storage system.} \addressesissue{3}
+data access and storage APIs.  Using this approach, a systems programmer can
+specify the functional behavior in a relational (or algebraic) language,
+allowing an optimizer to search through the space of functionally equivalent
+physical implementations and select a good execution plan, re-optimizing when
+storage characteristics or statistics change.  Much like query planning and
+optimization in database systems~\cite{hellerstein:cidr15}, this approach will
+separate the concerns of correctness and performance, protecting applications
+(which usually evolve slowly) against changes in more dynamic storage system.}
+\addressesissue{3}
+
+%\section{Future Work}
+%
+%\newcommentone{This work is one point along the path towards producing
+%general-purpose storage systems that can target special-purpose applications.
+%Continuing to expore existing storage system functionality as reusable building
+%blocks will enable the proliferation of storage-centric APIs, allowing
+%developers to custom-fit their applications to take advantage of the
+%capabilities of the underlying system. When adding new interfaces and composing
+%them together, careful consideration should be made to address security, access
+%control, and safety. Correctness and performance will surely be affected if
+%interfaces are combined poorly and these drawbacks will be quantified.}
+%\addressesissue{5}
+%
+%\newcommentone{The next higher-level services we plan to build using the
+%interfaces in Table~\ref{table:examples} are: an elastic cloud database, a data
+%layout manager, and a data processing engine. Some cloud databases aggressively
+%cache data to hide the overheads and weaker consistency semantics of the
+%distributed file systems they sit on~\cite{pavlo:sigmod16} but using Malacology
+%we could build a database that works well because of the storage system not in
+%spite of it. We plan to use the data IO interface for push down predicates, the
+%file type interface for maintaining indices and stats, and the durability
+%interface internals for ingest and data movement. The data layout manager would
+%be similar to DRepl~\cite{ionkov:msst2013} and would use the file type,
+%durability, and data I/O interface to manage different views of data. Finally,
+%the data processing engine would be similar to MapReduce and would use the data
+%I/O, shared resource, and load balancing interface to implement the resource
+%manager, computation engine, and storage platform in the Hadoop ecosystem.}
+%
+%\newcommentone{ Our experience with Zlog and Mantle illustrates that the labor
+%of wrapping existing services in interfaces is justified by the power and
+%flexibility they afford to programmers.  In exchange for this flexibility,
+%however, programmable storage systems force programmers to forfeit the
+%\emph{protection from change} afforded by narrow storage interfaces like the
+%POSIX API.  To implement applications built atop systems such as Malacology,
+%programmers must navigate a complex design space, simultaneously addressing
+%often orthogonal concerns including functional correctness, performance and
+%fault tolerance. Worse still, these compositions can be sensitive to changes in
+%the underlying environment (such as hardware upgrades, software version changes
+%and evolving workloads).  For example, a major version change in Ceph required
+%us to rewrite significant parts of Zlog to maintain acceptable performance --
+%each such evolution costs developer time and risks introducing bugs.}
+%
+%\newcommentone{We are actively exploring the use of high-level
+%\emph{declarative} languages based on Datalog~\cite{alvaro:cidr11} to program
+%storage APIs.  In this way, a programmer can specify the functional behavior of
+%a storage interface once in a relational (or algebraic) language, allowing an
+%optimizer to model search the space of functionally equivalent physical
+%implementations, and re-optimizing when characteristics change.  Much like
+%query planning and optimization in database systems~\cite{hellerstein:cidr15},
+%this approach will separate the concerns of correctness and performance, and
+%protect applications (which evolve slowly) against changes in highly dynamic
+%storage system.} \addressesissue{3}

### 2017-02-23 paper/7_future_work.tex
paper: integrate Carlos' future work comments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 interfaces in Table~\ref{table:examples} are: an elastic cloud database, a data
-processing engine, and a data layout manager.  The elastic database could use
-the data I/O interface to push down predicates, the file type interface to
-maintain access paths and statistics efficiently, and the durability interface
-to manage ingestion and movement.  The data processing engine would be similar
-to MapReduce, using the data I/O interface to send computation to the data, the
-load balancing interface as a proxy for farming off work, and the durability
-interface for persisting to input or output sinks.  Finally, the data layout
-manager would be similar to DRepl~\cite{ionkov:msst2013}, using the file type,
-durability, and data I/O interfaces to provide multiple views of data.}
+processing engine, and a data layout manager. Approaches proposed so far use
+the data I/O interface to push down predicates and computation, the file type
+interface to maintain access paths and metadata efficiently, and the durability
+interface to manage ingestion and movement. Using the programmable storage
+approach helps us build higher-level services that work well with the storage
+system not in spite of it.}
 

### 2017-02-23 paper/0_abstract.tex
paper: integrate Jeff's comments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 optimization efforts that were necessary for users to entrust their data to the
-storage system. We introduce the concept of a programmable storage system, a
-system that exposes internal services and abstractions with the goal of
-re-using and building new services on top of the storage stack. We present
-Malacology, a prototype implementation of a programmable storage system to
-explore how existing abstractions of common storage system services can be
-leveraged to adapt to the needs of new data processing systems and the
-increasing variety of storage devices. We show how this approach allows a large
-degree of flexibility for storage systems to evolve without sacrificing the
-robustness of their code-hardened subsystems. We illustrate the advantages and
-challenges of programmability by composing existing primitives into two new
-higher-level services: a \newcommentone{file system} metadata load balancer and
-a high-performance distributed shared-log. The evaluation demonstrates that our
-services inherit desirable qualities of the back-end storage system, including
-the ability to balance load, efficiently propagate cluster metadata, recover
-from failure, and to navigate trade-offs between latency and throughput using
-leases.  \addressesissue{4}
+storage system. We introduce the programmable storage approach, which exposes
+internal services and abstractions of the storage stack as building blocks for
+higher-level services. We also build a prototype to explore how existing
+abstractions of common storage system services can be leveraged to adapt to the
+needs of new data processing systems and the increasing variety of storage
+devices.  We illustrate the advantages and challenges of this approach by
+composing existing internal abstractions into two new higher-level services: a
+\newcommentone{file system} metadata load balancer and a high-performance
+distributed shared-log.  The evaluation demonstrates that our services inherit
+desirable qualities of the back-end storage system, including the ability to
+balance load, efficiently propagate service metadata, recover from failure, and
+to navigate trade-offs between latency and throughput using leases.
 \end{abstract}

### 2017-02-23 paper/2_application_specific_storage_stacks.tex
paper: integrate Jeff's comments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Application-Specific Storage Stacks}
\label{sec:application-specifc-storage-stacks}

Building storage stacks from the ground up for a specific purpose results in
the best performance. For example, GFS~\cite{ghemawat:sosp03} and
HDFS~\cite{shvachko:msst10} were designed specifically to serve Map/Reduce and
Hadoop jobs, and use techniques like exposing data locality and relaxing POSIX
constraints to achieve application-specific I/O optimizations. Another example
is Boxwood~\cite{maccormick:osdi04}, which experimented with B-trees and chunk
stores as storage abstractions to simplify application building.
Alternatively, general-purpose storage stacks are built with the flexibility to
serve many applications by providing a variety of interfaces and tunable
parameters.  Unfortunately, managing competing forces in these systems is
difficult and users want more control from the general-purpose storage stacks
without going as far as building their storage system from the ground up.

To demonstrate a recent trend towards more application-specific storage systems
we examine the state of programmability in Ceph. Something of a storage Swiss
army knife, Ceph simultaneously supports file, block, and object interfaces on
a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
Distributed Object Storage (RADOS) system is a cluster of object storage
devices (OSDs) that provide Ceph with data durability and integrity using
replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already
provides some degree of programmability; the OSDs support domain-specific
object interfaces that are implemented by composing existing low-level storage
interfaces that execute atomically. These interfaces are written in C++ and are
statically loaded into the system.

The Ceph community provides empirical evidence that developers are already
beginning to embrace programmable storage. Figure~\ref{fig:obj-int-dev-growth}
shows a dramatic growth in the production use of domain-specific interfaces in
the Ceph community since 2010. In that figure, classes are functional groupings
of methods on storage objects (e.g. remotely computing and caching the checksum
of an object extent).  What is most remarkable is that this trend contradicts
the notion that API changes are a burden for users.  Rather it appears that
gaps in existing interfaces are being addressed through ad-hoc approaches to
programmability. In fact, Table~\ref{table:objclasses} categorizes existing
interfaces and we clearly see a trend towards reusable services.

\begin{figure}[ht]
\centering
\includegraphics{figures/obj-int-dev-growth.png}
\caption{Since 2010, the growth in the number of co-designed object storage
interfaces in Ceph has been accelerating. This plot is the number of object
classes (a group of interfaces), and the total number of methods (the actual
API end-points).}
\label{fig:obj-int-dev-growth}
\end{figure}

\begin{table}[ht]
\centering
  \begin{tabular}{l|l|l}
    Category & Example & \# \\ \hline
    Logging  & Geographically distribute replicas & 11 \\ \hdashline
    Metadata & Snapshots in the block device OR  & \multirow{2}{*}{74} \\
    Management & Scan extents for file system repair & \\ \hdashline
    Locking  & Grants clients exclusive access & 6 \\ \hdashline
    Other & Garbage collection, reference counting  & 4\\
\end{tabular}
\caption{A variety of object storage classes exist to expose interfaces
    to applications. \# is the number of methods that implement these categories.
}
\label{table:objclasses}
\end{table}

The takeaway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
already trying to use programmability because their needs, whether they be
related to performance, availability, consistency, convenience, etc., are not
satisfied by the existing default set of interfaces. The popularity of the
custom object interface facility of Ceph could be due to a number of reasons,
such as the default algorithms/tunables of the storage system being
insufficient for the application's performance goals, programmers wanting to
exploit application-specific semantics, and/or programmers knowing how to
manage resources to improve performance. A solution based on
application-specific object interfaces is a way to work around the
traditionally rigid storage APIs because custom object interfaces give
programmers the ability to tell the storage system about their application: if
the application is CPU or IO bound, if it has locality, if its size has the
potential to overload a single node, etc.  Programmers often know what the
problem is and how to solve it, but until the ability to modify object
interfaces, they had no way to express to the storage system how to handle
their data.

Our approach is to expose more of the commonly used, code-hardened subsystems
of the underlying storage system as interfaces. The intent is that these
interfaces, which can be as simple as a redirection to the persistent data
store or as complicated as a strongly consistent directory service, should be
used and re-used in many contexts to implement a wide range of services. By
making programmability a `feature', rather than a `hack' or `workaround', we
help standardize a development process that now is largely ad-hoc.

### 2017-02-23 paper/4_malacology_a_programmable_storage_system.tex
paper: integrate Jeff's comments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \textbf{Example in Ceph}                &
-\textbf{Provided Primitives}            \\ \hline
+\textbf{Provided Functionality}            \\ \hline
 Service Metadata
@@
 
-\noindent {\it{\textbf{Malacology: }}} Mantle~\cite{sevilla:sc15-mantle}, a
-programmable \newcomment{file system}metadata load balancer, has been
-implemented on Malacology so that its infrastructure can enjoy all the
-properties of existing internal abstractions.  The load balancing logic is
-intact but the infrastructure has been improved to safely load, manage, and
-persist its load balancing policies.  Composing Mantle with the service
-metadata and durability interfaces, Mantle can safely version balancer
-policies, save balancer policies in the backend object store and centralize
-warnings/errors. When combined with file types, load balancing policies can
-express a variety of rules such as multi-tenant and multi-application
-balancing.
+\noindent {\it{\textbf{Malacology: }}} the existing load balancing mechanisms
+are exposed through an API and programmers can customize the behavior through a
+domain specific language. These mechanisms include the ability to migrate,
+partition, and measure load. Using the Service Metadata and Durability
+interfaces, this Load Balancing interface can safely version balancer policies,
+save balancer policies in the back-end object store and centralize
+warnings/errors. When combined with the File Type interface, load balancing
+policies can express a variety of rules for handling multi-tenant and
+workloads.
 

### 2017-02-23 paper/5_services_built_on_malacology.tex
paper: integrate Jeff's comments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 In this section we describe two services built on top of Malacology. The first
-is Mantle, a framework for dynamically specifying metadata load balancing
+is Mantle~\cite{sevilla:sc15-mantle}, a framework for dynamically specifying metadata load balancing
 policies. The second system, ZLog, is a high-performance distributed
@@
 
-Mantle is a programmable load balancer that separates the metadata balancing
+Mantle~\cite{sevilla:sc15-mantle} is a programmable load balancer that separates the metadata balancing
 policies from their mechanisms. Administrators inject code to change how the
-metadata cluster distributes metadata. In ~\cite{sevilla:sc15-mantle} the
-authors showed how using Mantle, one can implement a single node metadata
+metadata cluster distributes metadata. Our previous work showed how 
+to use Mantle to implement a single node metadata
 service, a distributed metadata service with hashing, and a distributed
 metadata service with dynamic subtree partitioning. 
+We have re-implmented Mantle on Malacology so that Mantle can enjoy the
+properties of existing internal abstractions.
 

### 2017-02-23 paper/6_evaluation.tex
paper: integrate Jeff's comments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 balancers are hard-coded into Ceph while Mantle is an API for balancing load
-presented in~Mantle~\cite{sevilla:sc15-mantle} and re-implemented on top of
+presented in~Mantle and re-implemented on top of
 Malacology in this paper.  The increased throughput for the CephFS and Mantle
@@
 performance of the balancing mode that CephFS falls into {\it most of the
-time}.  CephFS has 3 modes for balancing load: CPU mode, workload mode, and
+time}.  CephFS currently has 3 modes for balancing load: CPU mode, workload mode, and
 hybrid mode. All three have the same structure for making migration decisions
@@
 degrees of aggressiveness in making migration decisions; CephFS makes a
-decision quickly and throughput jumps to 2500 ops/second within 10 seconds
+decision 10 seconds into the run and throughput jumps to 2500 ops/second within 10 seconds
 while Mantle takes more time to stabilize. This conservative behavior is

### 2017-02-23 paper/1_intro.tex
paper: wordsmithing - remove confusing OSD/MDS acronyms - capatilize interface names - spell check and fix colors in figures
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 system} metadata servers, durability and object interfaces provided by the
-backend object store, and propagation of consistent cluster state provided by
+back-end object store, and propagation of consistent cluster state provided by
 the monitoring service (see Figure~\ref{fig:overview}).  Malacology is

### 2017-02-23 paper/2_application_specific_storage_stacks.tex
paper: wordsmithing - remove confusing OSD/MDS acronyms - capatilize interface names - spell check and fix colors in figures
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 Distributed Object Storage (RADOS) system is a cluster of object storage
-devices (OSDs) that provide Ceph with data durability and integrity using
+devices that provide Ceph with data durability and integrity using
 replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already
-provides some degree of programmability; the OSDs support domain-specific
+provides some degree of programmability; the object storage daemons support domain-specific
 object interfaces that are implemented by composing existing low-level storage
@@
 programmers the ability to tell the storage system about their application: if
-the application is CPU or IO bound, if it has locality, if its size has the
+the application is CPU or I/O bound, if it has locality, if its size has the
 potential to overload a single node, etc.  Programmers often know what the

### 2017-02-23 paper/4_malacology_a_programmable_storage_system.tex
paper: wordsmithing - remove confusing OSD/MDS acronyms - capatilize interface names - spell check and fix colors in figures
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
   & \S\ref{sec:shared-resource-interface}
-  & MPI collective IO, burst
+  & MPI collective I/O, burst
   & POSIX metadata protocols
@@
 
-\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}} In the
-case of Ceph, a consistent view of cluster state among server daemons and
+\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}} 
+a consistent view of cluster state among server daemons and
 clients is critical to provide strong consistency guarantees to clients.  Ceph
@@
 
-\noindent{\it{\textbf{Malacology: }}} Malacology exposes a strongly-consistent
-view of time-varying service metadata as a service rather than a hidden
+\noindent{\it{\textbf{Malacology: }}} we expose a strongly-consistent
+view of time-varying service metadata as an interface rather than a hidden
 internal component. Malacology provides a generic API for adding arbitrary
@@
 because it lets daemons come to consensus on system critical state. Bugs in the
-internal subsystems or ommitting this from services that need this type of
+internal subsystems or omitting this from services that need this type of
 consistency affects correctness.} \addressesissue{3, 5}
@@
 
-\oldcomment{\noindent 4.2 Object I/O Intefrace}
+\oldcomment{\noindent 4.2 Object I/O Interfrace}
 \subsection{Data I/O Interface}
@@
 
-\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}} An object
+\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}} an object
 interface is a plugin structured similarly to that of an RPC in which a
@@
 storage daemon (OSD) and metadata server (MDS) subsystems with an embedded Lua
-VM.  It uses the service metadata interface (M) to propagate
+VM.  It uses the Service Metadata interface in the monitor (M) to propagate
 interfaces/versions across the cluster \label{fig:implementation}}
@@
 
-\noindent {\it{\textbf{Malacology: }}} The implementation of Ceph's object
+\noindent {\it{\textbf{Malacology: }}} the implementation of Ceph's object
 abstraction, although powerful, does not readily support programmability.
@@
 
-\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}} In Ceph,
+\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}}
 the locking service implements a capability-based system that expresses what
@@
 
-\newcomment{\\ \noindent\it{\textbf{Malacology: }}} While the current policy
+\newcomment{\\ \noindent\it{\textbf{Malacology: }}} while the current policy
 for sharing access and voluntarily releasing resources is largely best-effort,
@@
 \oldcomment{\noindent File type interface.}
-\subsubsection{Malacology: File Type Interface}
+\subsubsection{File Type Interface}
 \label{sec:file-type-interface}
@@
 
-\newcomment{\\ \noindent {\it{\textbf{Malacology: }}}} We allow new inode types
+\newcomment{\\ \noindent {\it{\textbf{Malacology: }}}} allows new inode types
 to be defined such that applications can create domain-specific interfaces to
@@
 \newcomment{Many large scale storage systems separate file system metadata and
-data IO so that the corresponding services can scale independently. Metadata
+data I/O so that the corresponding services can scale independently. Metadata
 requests transfer small amounts of data and they happen relatively frequently
@@
 
-\newcomment{\\ \noindent {\it{\textbf{Existing Ceph Implementation: }}}} Ceph
-also addresses the challenge of balancing \newcommentone{file system} metadata
+\newcomment{\\ \noindent {\it{\textbf{Existing Ceph Implementation: }}}} 
+addresses the challenge of balancing \newcommentone{file system} metadata
 load with a separate metadata cluster. This cluster uses load balancing
@@
 save balancer policies in the back-end object store and centralize
-warnings/errors. When combined with the File Type interface, load balancing
-policies can express a variety of rules for handling multi-tenant and
+warnings/errors. When combined with the File Type interface, the Load Balancing
+interface can express policies for handling a variety of multi-tenant
 workloads.
@@
 
-\noindent{\it{\textbf{Existing Ceph Implementation: }}} Ceph provides storage
+\noindent{\it{\textbf{Existing Ceph Implementation: }}} provides storage
 by striping and replicating data across RADOS~\cite{weil_rados_2007}, the
 reliable distributed object store. RADOS protects data using common techniques
-such as erasure coding, replication, and scrubbing.  example, when the number
-of placement groups change, the OSDs re-balance and re-shard data in the
-background in a process called placement group splitting in which OSDs
+such as erasure coding, replication, and scrubbing. For example, when the number
+of placement groups change, the object storage daemons re-balance and re-shard data in the
+background in a process called placement group splitting in which object storage daemons
 communicate directly with each other to cover a new state.  In order to reduce
-load on the monitoring service, Ceph OSDs use a gossip protocol to efficiently
+load on the monitoring service, the object storage daemons use a gossip protocol to efficiently
 propagate changes to cluster maps throughout the system, and autonomously
@@
 
-\noindent {\it{\textbf{Malacology: }}} Metadata service policies and object
+\noindent {\it{\textbf{Malacology: }}} metadata service policies and object
 storage interfaces are stored durability within RADOS and managed by storing
@@
 consistent view of these data structures, we use this service to automatically
-install interfaces in OSDs, and install policies within the metadata server
-daemon (MDS) such that clients and daemons are synchronized on
+install interfaces in object storage daemons, and install policies within the metadata server
+daemon such that clients and daemons are synchronized on
 correct implementations without restarting.

### 2017-02-23 paper/5_services_built_on_malacology.tex
paper: wordsmithing - remove confusing OSD/MDS acronyms - capatilize interface names - spell check and fix colors in figures
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 In this section we describe two services built on top of Malacology. The first
-is Mantle~\cite{sevilla:sc15-mantle}, a framework for dynamically specifying metadata load balancing
+is Mantle~\cite{sevilla:sc15-mantle}, a framework for dynamically specifying file system metadata load balancing
 policies. The second system, ZLog, is a high-performance distributed
@@
 metadata service with dynamic subtree partitioning. 
-We have re-implmented Mantle on Malacology so that Mantle can enjoy the
+We have re-implemented Mantle on Malacology so that Mantle can enjoy the
 properties of existing internal abstractions.
@@
 With Malacology, Mantle stores the version of the current load balancer in the
-Ceph service metadata interface. The version of the load balancer corresponds
-to an object name in the balancing policy. Using the service metadata interface
+Service Metadata interface. The version of the load balancer corresponds
+to an object name in the balancing policy. Using the Service Metadata interface
 ensures that all metadata servers use the same version of the load balancer. As
@@
 The load balancer version described above corresponds to the name of an object
-in RADOS that holds the actual Lua balancing code.  When MDS nodes start
-balancing load, they first check the latest version from the MDS map and
+in RADOS that holds the actual Lua balancing code.  When metadata server nodes start
+balancing load, they first check the latest version from the metadata server map and
 compare it to the balancer they have loaded. If the version has changed, they
@@
 The balancer pulls the Lua code from RADOS synchronously; asynchronous reads
-are not possible because of the architecture of the MDS. The synchronous
+are not possible because of the architecture of the metadata server. The synchronous
 behavior is not the default behavior for RADOS operations, so we achieve this
@@
 RADOS-related goes wrong.  We use this implementation because we do not want to
-do a blocking OSD read from inside the global MDS lock. Doing so would bring
-down the MDS cluster if any of the OSDs are not responsive.
+do a blocking object storage daemon read from inside the global metadata server lock. Doing so would bring
+down the metadata server cluster if any of the object storage daemons are not responsive.
 
@@
 language for writing balancer code. If we used a language that needs to be
-compiled, like the C++ object classes in the OSD, we would need to ensure
+compiled, like the C++ object classes in the object storage daemon, we would need to ensure
 binary compatibility, which is complicated by different operating systems,
@@
 In the original implementation, Mantle would log all errors, warnings, and
-debug messages to a log stored locally on each MDS server. To get the simplest
-status messages or to debug problems, the user would have to log into each MDS
+debug messages to a log stored locally on each metadata server. To get the simplest
+status messages or to debug problems, the user would have to log into each metadata server
 individually, look at the logs, and reason about causality and ordering.
@@
 users going to each node, they can watch messages appear at the monitor.
-Messages are logged sparingly, so as not to spam the monitor with frivolous
+Messages are logged sparingly, so as not to overload the monitor with frivolous
 debugging but important events, like balancer version changes or failed
@@
 can run at a very high throughput and at low latency. Since the sequencer is
-centralized, ensuring serializability in the common case is trivial.  The
+centralized, ensuring serialization in the common case is trivial.  The
 primary challenge in CORFU is handling the failure of the sequencer in a way
@@
 log which also atomically increments the tail position.  We implement the
-sequencer service in Malacology as an application-specific file type that
-associates a small amount of metadata with the inode state managed by the Ceph
-metadata service.  This approach has the added benefit of allowing the metadata
+sequencer service in Malacology using the File Type interface.
+This approach has the added benefit of allowing the metadata
 service to handle naming, by representing each sequencer instance in the
@@
 nature of the metadata service may prevent the sequencer from achieving maximum
-throughput. To provide a solution to this issue we have taken advantage of the
-programmability of the metadata service load balancing to construct a
+throughput. To address this issue we use the Load Balancing interface to construct a
 service-specific load balancing policy. As opposed to a balancing policy that
@@
 
-In order to repopulate the sequencer state (i.e. the cached, current tail of
+In order to repopulate the sequencer state ({\it i.e.} the cached, current tail of
 the log) during recovery of a sequencer, the maximum position in the log must

### 2017-02-23 paper/6_evaluation.tex
paper: wordsmithing - remove confusing OSD/MDS acronyms - capatilize interface names - spell check and fix colors in figures
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-The sequencer is implemented as a custom file type in which the sequencer state
-(a 64-bit integer) is embedded in the inode of a file representing the
+The sequencer is implemented using the File Type interface so that the sequencer state
+(a 64-bit integer) is embedded in the inode of a file; this file represents the
 sequencer. A total ordering of the log is imposed by the re-use of the
@@
 reservation allows clients to experience a much lower latency because they
-experience uncontended access for a longer period of time.
+experience isolated access for a longer period of time.
 
@@
 is cropped at the 99.999th percentile due to large outliers that we believe
-occur in instances in which the MDS is performing I/O while it is in the
+occur in instances in which the metadata server is performing I/O while it is in the
 process of re-distributing the capability to another client.
@@
 using a peer-to-peer gossip protocol.  The latency is defined as the elapsed
-time following the Paxos proposal for an interface update until each OSD makes
+time following the Paxos proposal for an interface update until each object storage daemon makes
 the update live (the cost of the Paxos proposal is configurable and is
@@
 Figure~\ref{fig:propdelay} shows the lower bound cost for updates in a large
-cluster. In the experiment labeled ``120 OSD (RAM)" a cluster of 120 OSDs (10
-OSDs x 12 servers) using an in-memory data store were deployed, showing a
+cluster. In the experiment labeled ``120 OSD (RAM)" a cluster of 120 object storage daemons (OSDs)
+using an in-memory data store were deployed, showing a
 latency of less than 54 ms with a probability of 90\% and a worst case latency
@@
 interval in order to accumulate updates. In a minimum, realistic quorum of 3
-monitors using HDD-based storage, we were able to decrease this interval to an
+monitors using hard drive-based storage, we were able to decrease this interval to an
 average of 222 ms.
@@
 multiple sequencers. The following experiments demonstrate the feasibility of
-using the mechanisms of the Malacology load balancing interface to inherit
+using the mechanisms of the Malacology Load Balancing interface to inherit
 these features and to alleviate load from overloaded servers.
@@
 contention at the sequencers by forcing clients to make round-trips for every
-request. We implement this using the shared resource interface that forces
+request. We implement this using the Shared Resource interface that forces
 round-trips.  Because the sequencer's only function is to  hand out positions
@@
 First, we show how the ZLog service can orchestrate multiple sequencers using
-the Malacology load balancing interface.
+the Malacology Load Balancing interface.
 Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput over time of
@@
 Section~\ref{sec:feature-backoff} examines the slope and start time of the
-rebalancing phase between 0 and 60 seconds by comparing the aggressiveness of
+re-balancing phase between 0 and 60 seconds by comparing the aggressiveness of
 the balancers.
@@
 
-Using Mantle, the administrator can put the load balancing service into ``proxy
+Using Mantle, the administrator can put the load balancer into ``proxy
 mode" or ``client mode". In proxy mode one server receives all requests and
@@
 metric a ``migration unit".  Expressing this heuristic is not easily achievable
-with outward facing tunable parameters (i.e. system knobs) but with Mantle's
+with outward facing tunable parameters ({\it i.e.} system knobs) but with Mantle's
 programmable interface, we can force the load balancer to change its migration

### 2017-02-23 paper/7_future_work.tex
paper: wordsmithing - remove confusing OSD/MDS acronyms - capatilize interface names - spell check and fix colors in figures
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 processing engine, and a data layout manager. Approaches proposed so far use
-the data I/O interface to push down predicates and computation, the file type
-interface to maintain access paths and metadata efficiently, and the durability
+the Data I/O interface to push down predicates and computation, the File Type
+interface to maintain access paths and metadata efficiently, and the Durability
 interface to manage ingestion and movement. Using the programmable storage
@@
 \addressesissue{3}
-
-%\section{Future Work}
-%
-%\newcommentone{This work is one point along the path towards producing
-%general-purpose storage systems that can target special-purpose applications.
-%Continuing to expore existing storage system functionality as reusable building
-%blocks will enable the proliferation of storage-centric APIs, allowing
-%developers to custom-fit their applications to take advantage of the
-%capabilities of the underlying system. When adding new interfaces and composing
-%them together, careful consideration should be made to address security, access
-%control, and safety. Correctness and performance will surely be affected if
-%interfaces are combined poorly and these drawbacks will be quantified.}
-%\addressesissue{5}
-%
-%\newcommentone{The next higher-level services we plan to build using the
-%interfaces in Table~\ref{table:examples} are: an elastic cloud database, a data
-%layout manager, and a data processing engine. Some cloud databases aggressively
-%cache data to hide the overheads and weaker consistency semantics of the
-%distributed file systems they sit on~\cite{pavlo:sigmod16} but using Malacology
-%we could build a database that works well because of the storage system not in
-%spite of it. We plan to use the data IO interface for push down predicates, the
-%file type interface for maintaining indices and stats, and the durability
-%interface internals for ingest and data movement. The data layout manager would
-%be similar to DRepl~\cite{ionkov:msst2013} and would use the file type,
-%durability, and data I/O interface to manage different views of data. Finally,
-%the data processing engine would be similar to MapReduce and would use the data
-%I/O, shared resource, and load balancing interface to implement the resource
-%manager, computation engine, and storage platform in the Hadoop ecosystem.}
-%
-%\newcommentone{ Our experience with Zlog and Mantle illustrates that the labor
-%of wrapping existing services in interfaces is justified by the power and
-%flexibility they afford to programmers.  In exchange for this flexibility,
-%however, programmable storage systems force programmers to forfeit the
-%\emph{protection from change} afforded by narrow storage interfaces like the
-%POSIX API.  To implement applications built atop systems such as Malacology,
-%programmers must navigate a complex design space, simultaneously addressing
-%often orthogonal concerns including functional correctness, performance and
-%fault tolerance. Worse still, these compositions can be sensitive to changes in
-%the underlying environment (such as hardware upgrades, software version changes
-%and evolving workloads).  For example, a major version change in Ceph required
-%us to rewrite significant parts of Zlog to maintain acceptable performance --
-%each such evolution costs developer time and risks introducing bugs.}
-%
-%\newcommentone{We are actively exploring the use of high-level
-%\emph{declarative} languages based on Datalog~\cite{alvaro:cidr11} to program
-%storage APIs.  In this way, a programmer can specify the functional behavior of
-%a storage interface once in a relational (or algebraic) language, allowing an
-%optimizer to model search the space of functionally equivalent physical
-%implementations, and re-optimizing when characteristics change.  Much like
-%query planning and optimization in database systems~\cite{hellerstein:cidr15},
-%this approach will separate the concerns of correctness and performance, and
-%protect applications (which evolve slowly) against changes in highly dynamic
-%storage system.} \addressesissue{3}

### 2017-02-23 paper/8_related_work.tex
paper: wordsmithing - remove confusing OSD/MDS acronyms - capatilize interface names - spell check and fix colors in figures
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \section{Related Work}
-
-% IOStack: Software-Defined Object Storage
-% sRoute: Treating the Storage Stack Like a Network
-   % - lots of good refs to check out
-% Information and control in gray-box systems
 
@@
 facilitate flexible and dynamic provisioning of storage resources of different
-kinds, including flash memory and disk drives, to create a virtualized mapping
+kinds, including flash memory and disk drives, to create a virtual mapping
 between common storage abstractions (e.g. files, objects, and blocks) and
@@
 injecting code into the kernel to specialize resource management. Another
-approach is to pass hints between the different layers of the IO stack to
+approach is to pass hints between the different layers of the I/O stack to
 bridge the semantic gap between applications and

### 2017-02-23 paper/1_intro.tex
paper: more clarifications on terminology
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 storage system resources, services, and abstractions that in a generalized form
-can be used to \emph{program} new services. By composing higher-level services
-over existing abstractions one can reuse subsystems and leverage their
-optimizations, established correctness, robustness, and efficiency.
+can be used to \emph{program} new services. While this `dirty-slate' approach
+of factoring out useful code is time consuming, it lets programmers reuse
+subsystems and leverage their optimizations, established correctness,
+robustness, and efficiency. `Clean-slate' approaches could be implemented
+faster but they do so at the expense of ``throwing away" proven code.
 
@@
 
-Malacology includes a non-exhaustive set of interfaces that can be used as
+Malacology is a set of interfaces that can be used as
 building blocks for constructing novel storage abstractions, including:

### 2017-02-23 paper/4_malacology_a_programmable_storage_system.tex
paper: more clarifications on terminology
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \caption{Common internal abstractions. The same ``internal abstractions" common
-across large-scale systems because they provide primitives that solve general
+across large-scale systems because they provide functionality that solves general
 distributed systems problems.  Here we list examples of what these internal
@@
 
-The guiding principle is to re-use existing services and extend them so that
+The guiding principle is to re-use existing services and compose them so that
 these services can be \emph{programmed}. We accomplish programmability of a
-service by exporting bindings (or ``hooks'') for an interpreted programming
+service by exporting bindings for an interpreted programming
 language so that programming can occur without having to restart the storage
@@
 \caption{Malacology allows users to dynamically define
-object/\newcommentone{file system} metadata interfaces by extending the object
+object/\newcommentone{file system} metadata interfaces by composing the object
 storage daemon (OSD) and metadata server (MDS) subsystems with an embedded Lua

### 2017-02-24 paper/0_abstract.tex
paper: wordsmithing and fix colors
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 balance load, efficiently propagate service metadata, recover from failure, and
-to navigate trade-offs between latency and throughput using leases.
+navigate trade-offs between latency and throughput using leases.
 \end{abstract}

### 2017-02-24 paper/1_intro.tex
paper: wordsmithing and fix colors
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 monitor daemons (M) that maintain cluster state, and service-specific daemons
-({\it e.g.}, file system metadata servers). Malacology enables the programmability of
+(e.g., file system metadata servers). Malacology enables the programmability of
 internal abstractions (bold arrows) to re-use and compose
-existing subsystems.  With Malacology, we add new services, ZLog and
+existing subsystems.  With Malacology, we built new higher-level services, ZLog and
 Mantle, that sit alongside traditional user-facing APIs (file, block,
@@
 storage system resources, services, and abstractions that in a generalized form
-can be used to \emph{program} new services. While this `dirty-slate' approach
-of factoring out useful code is time consuming, it lets programmers reuse
-subsystems and leverage their optimizations, established correctness,
+can be used to \emph{program} new services. This `dirty-slate' approach of
+factoring out useful code lets programmers re-use subsystems of the back-end
+storage system, thus inheriting their optimizations, established correctness,
 robustness, and efficiency. `Clean-slate' approaches could be implemented
@@
 
-Malacology is a set of interfaces that can be used as
+Malacology includes a set of interfaces that can be used as
 building blocks for constructing novel storage abstractions, including:
@@
 \item An interface for installing and evolving domain-specific, cluster-wide
-\textbf{data interfaces}.
+\textbf{data I/O} functionality.
 
@@
 briefly describe how those system are used within Malacology
-(\S\ref{sec:malacology}).  Then we describe the services that we have
-constructed within the Malacology framework (\S\ref{sec:services}), and
+(Section~\ref{sec:malacology}).  Then we describe the services that we have
+constructed in the Malacology framework (Section~\ref{sec:services}), and
 evaluate our ideas within our prototype implementation
-(\S\ref{sec:evaluation}).  We conclude by discussing related and future work.
+(Section~\ref{sec:evaluation}).  We conclude by discussing future and related work.

### 2017-02-24 paper/2_application_specific_storage_stacks.tex
paper: wordsmithing and fix colors
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 the best performance. For example, GFS~\cite{ghemawat:sosp03} and
-HDFS~\cite{shvachko:msst10} were designed specifically to serve Map/Reduce and
+HDFS~\cite{shvachko:msst10} were designed specifically to serve MapReduce and
 Hadoop jobs, and use techniques like exposing data locality and relaxing POSIX
@@
 Alternatively, general-purpose storage stacks are built with the flexibility to
-serve many applications by providing a variety of interfaces and tunable
+serve many applications by providing standardized interfaces and tunable
 parameters.  Unfortunately, managing competing forces in these systems is
@@
 Distributed Object Storage (RADOS) system is a cluster of object storage
-devices that provide Ceph with data durability and integrity using
+daemons that provide Ceph with data durability and integrity using
 replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already
 provides some degree of programmability; the object storage daemons support domain-specific
-object interfaces that are implemented by composing existing low-level storage
-interfaces that execute atomically. These interfaces are written in C++ and are
+code that can manipulate objects on the server that has the data local. 
+These ``interfaces" are implemented by composing existing low-level storage
+abstractions that execute atomically. They are written in C++ and are
 statically loaded into the system.

### 2017-02-24 paper/3_challenges.tex
paper: wordsmithing and fix colors
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-To address these challenges we present Malacology \oldcomment{. Malacology is
-both a prototype for a programmable storage system and a} \newcomment{, our
+To address these challenges we present Malacology,\oldcomment{. Malacology is
+both a prototype for a programmable storage system and a} \newcomment{ our
 prototype programmable storage system. It uses the programmable storage}
@@
 is not defined by what abstractions are exposed, rather a programmable storage
-system adheres to the design approach of choosing interfaces so administrators
+system adheres to the design approach of exposing interfaces so administrators
 can have better control of the storage system. The interfaces presented in this

### 2017-02-24 paper/4_malacology_a_programmable_storage_system.tex
paper: wordsmithing and fix colors
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
   & \S\ref{sec:file-type-interface}
-  & ?????
+  & MPI supports architecture-specific code~\cite{thakur:iopads1999}
   & file striping strategy
@@
 \end{tabular}
-\caption{Common internal abstractions. The same ``internal abstractions" common
+\caption{Common internal abstractions. The same ``internal abstractions" are common
 across large-scale systems because they provide functionality that solves general
@@
 language so that programming can occur without having to restart the storage
-system (see also below,~\S\ref{sec:durability}). Table~\ref{table:examples}
+system (see also below, Section~\ref{sec:durability}). Table~\ref{table:examples}
 shows the internal services from Ceph that we expose in the Malacology
@@
 security and performance concerns. We will now discuss the common subsystems
-used to manage storage system and how Malacology makes them programmable.
+used to manage storage systems and how Malacology makes them programmable.
 
@@
 \newcommentone{Service metadata is information about the daemons in the system
-and includes membership details, hardware layout ({\it e.g.}, racks, power
-supplies, etc.), data layout, and daemon state/configuration. It differs from
+and includes membership details, hardware layout (e.g., racks, power
+supplies, etc.), data layout, and daemon state and configuration. It differs from
 traditional file system metadata which is information about files. For the rest
 of the paper when we use the phrase ``metadata server" or ``metadata service",
-we are referring to the daemon(s) that manages file system metadata.}
+we are referring to the daemon(s) that manages file system metadata (not service metadata).}
 \addressesissue{4}
@@
 view of time-varying service metadata as an interface rather than a hidden
-internal component. Malacology provides a generic API for adding arbitrary
+internal component. This is shown in Figure~\ref{fig:implementation}, 
+where object interfaces and load balancer policies use the Service Metadata
+interface.
+Malacology provides a generic API for adding arbitrary
 values to existing subsystem cluster maps. As a consequence of this,
@@
 clients can write new values) or triggering actions based on specific values
-(e.g. sanitize values).  The higher-level services we implement in
-\S\ref{sec:services} make use of this functionality to register, version and
+(e.g. sanitize values).  The higher-level services we implement in Section~\ref{sec:services} make use of this functionality to register, version and
 propagate dynamic code (Lua scripts) for new object interfaces defined in
-storage daemons (\S\ref{object-data-interface}) and policies in the load
-balancer \S\ref{malacology:mds}.  Using this service guarantees that interface
+storage daemons (Section~\ref{object-data-interface}) and policies in the load
+balancer (Section~\ref{malacology:mds}).  Using this service guarantees that interface
 definitions are not only made durable, but are transparently and consistently
@@
 \noindent\newcomment{{\it{\textbf{Impact: }}} provides core functionality
-because it lets daemons come to consensus on system critical state. Bugs in the
+because it lets daemons come to a consensus on system critical state. Bugs in the
 internal subsystems or omitting this from services that need this type of
@@
 object storage API with economy of expression, which at the same time provides
-the full set of features of the base object class. New object interfaces that
+the full set of features of the original object interface implementation. New object interfaces that
 are expressed in thousands of lines of code can be implemented in approximately
@@
 
-In general, file system metadata servers manages client sessions, allowing
+File system metadata servers manage client sessions, allowing
 clients to obtain locks (e.g. file byte ranges), and capabilities (e.g. to
@@
 (e.g. users or database snapshots) often require a naming service.  The
-metadata service manages a POSIX file system hierarchy where files and
-directories are represented as inode data structures and expose a POSIX file
-interface.
+metadata service exposes a POSIX file system hierarchy where files and
+directories are represented as inode data structures.
 
@@
 load with a separate metadata cluster. This cluster uses load balancing
-policies to migrate directory inodes around the cluster to alleviate overload
-at single nodes~\cite{weil:sc2004-dyn-metadata}.  The policies use metrics
+policies to migrate directory inodes around the cluster to alleviate overloaded servers~\cite{weil:sc2004-dyn-metadata}.  The policies use metrics
 based on system state (e.g.  CPU and memory utilization) and statistics
@@
 replication, and data scrubbing. For scalability, many of these features are
-implemented using a peer-to-peer protocol that allows object storage devices to
+implemented using a peer-to-peer protocol that allows object storage daemons to
 operate autonomously without a centralized coordinator.\\}
@@
 of placement groups change, the object storage daemons re-balance and re-shard data in the
-background in a process called placement group splitting in which object storage daemons
-communicate directly with each other to cover a new state.  In order to reduce
+background in a process called placement group splitting. During placement group splitting, object storage daemons
+communicate directly with each other to converge on a new data layout.  In order to reduce
 load on the monitoring service, the object storage daemons use a gossip protocol to efficiently
@@
 \noindent {\it{\textbf{Malacology: }}} metadata service policies and object
-storage interfaces are stored durability within RADOS and managed by storing
-references with the object maps. Since the cluster already propagates a
+storage interfaces are stored durability within RADOS and are managed by storing
+references in the object server daemon maps. Since the cluster already propagates a
 consistent view of these data structures, we use this service to automatically
 install interfaces in object storage daemons, and install policies within the metadata server
-daemon such that clients and daemons are synchronized on
+daemons such that clients and daemons are synchronized on
 correct implementations without restarting.

### 2017-02-24 paper/5_services_built_on_malacology.tex
paper: wordsmithing and fix colors
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 In this section we describe two services built on top of Malacology. The first
-is Mantle~\cite{sevilla:sc15-mantle}, a framework for dynamically specifying file system metadata load balancing
+is Mantle, a framework for dynamically specifying file system metadata load balancing
 policies. The second system, ZLog, is a high-performance distributed
-shared-log.  In addition to these services, we'll demonstrate how we combine
+shared-log.  In addition to these services, we will demonstrate how we combine
 ZLog and Mantle to implement service-aware metadata load balancing policies.
@@
 metadata service with dynamic subtree partitioning.
-We have re-implemented Mantle on Malacology so that Mantle can enjoy the
-properties of existing internal abstractions.
 
@@
 (no versioning, durability, or policy distribution).  Re-implemented using
-Malacology, Mantle now enjoys (1) the versioning provided by the metadata
-service interface provided by the monitor daemons and (2) the durability and
+Malacology, Mantle now enjoys (1) the versioning provided by Ceph's monitor daemons and (2) the durability and
 distribution provided by Ceph's reliable object store.  Re-using the internal
@@
 it was trivial to make the versions inconsistent. Maintaining consistent
-versions is important for balancing policies that are cooperative in which
-local decisions are made assuming properties about other instances of the same
-policy.
+versions is important for cooperative balancing policies, where
+local decisions are made assuming properties about other instances in the cluster.
 
@@
 to an object name in the balancing policy. Using the Service Metadata interface
-ensures that all metadata servers use the same version of the load balancer. As
-a result, the policy version inherits the consistency of Ceph's internal
+means Mantle inherits the consistency of Ceph's internal
 monitor daemons. The user changes the version of the load balancer using a new
@@
 dereference the pointer to the balancer version by reading the corresponding
-object in RADOS. This is in contrast to the original balancer which stored load
+object in RADOS. This is in contrast to the original Mantle implementation which stored load
 balancer code on the local file system -- a technique which is unreliable and
@@
 by the monitoring subsystem and appear in the monitor cluster log so instead of
-users going to each node, they can watch messages appear at the monitor.
+users going to each node, they can watch messages appear at the monitor daemon.
 Messages are logged sparingly, so as not to overload the monitor with frivolous
 debugging but important events, like balancer version changes or failed
-subsystems show up in the centralized log.
+subsystems, show up in the centralized log.
 
@@
 access to the resource at the authoritative metadata server for the sequencer
-inode.  Interestingly, we found that the capability system in Ceph uses a
-strategy to reduce metadata service load by allowing clients that open a shared
+inode.  Interestingly, we found that the capability system in Ceph 
+reduces metadata service load by allowing clients that open a shared
 file to temporarily obtain an exclusive cached copy of the resource, resulting
@@
 Recovery of the sequencer process itself may be handled in many ways, such as
-leader election using an auxiliary service such as Paxos. In our
-implementation, the recovery is the same as (and is inherited from) the Ceph
+leader election using an auxiliary service like Paxos. In our
+implementation, the recovery is the same as (and is inherited from) the CephFS
 metadata service. Handling the failure of a client that holds the sequencer

### 2017-02-24 paper/6_evaluation.tex
paper: wordsmithing and fix colors
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 give examples of the flexibility that Malacology provides to programmers.  
+This section uses a principled approach for evaluating tunables of the
+interfaces and the trade-offs we discuss should be acknowledged when building
+higher-level services.
 
@@
 resource as an inode, and (2) the monitoring sub-subsystem to distribute and
-install application-specific I/O interface required of the CORFU protocol. In
+install application-specific I/O interfaces required of the CORFU protocol. In
 Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the
@@
 The sequencer is implemented using the File Type interface so that the sequencer state
-(a 64-bit integer) is embedded in the inode of a file; this file represents the
-sequencer. A total ordering of the log is imposed by the re-use of the
+(a 64-bit integer) is embedded in the inode of a file.
+A total ordering of the log is imposed by the re-use of the
 capability service that can be used to grant exclusive access of inode state to
@@
 per second is the combined throughput of the two clients, and the average
-latency is number of microseconds required to obtain a new log position. With a
+latency is the number of microseconds required to obtain a new log position. With a
 small quota more time is spent exchanging exclusive access, while a large quota
@@
 latency for each client in all experiment configurations. At the 99th
-percentile clients accessed the sequencer is less than a millisecond. The CDF
+percentile clients accessed the sequencer in less than a millisecond. The CDF
 is cropped at the 99.999th percentile due to large outliers that we believe
@@
 the update live (the cost of the Paxos proposal is configurable and is
-discussed below). The latency measurements were taken on the nodes running Ceph
+discussed below). The latency measurements were taken on the nodes running object
 server daemons, and thus exclude the client round-trip cost. In each of the
@@
 co-locating sequencers on the same physical node is not ideal but building a
-load balancer that can migrate the shared resource ({\it e.g.,}, the resource
+load balancer that can migrate the shared resource (e.g., the resource
 that mediates access to the tail of the log) is a time-consuming, non-trivial
@@
 the cluster; ``No Balancing" keeps all sequencers on one server, ``CephFS"
-migrates sequencers using the CephFS load balancers, and ``Mantle" uses a
-custom load balancer we wrote specifically for sequencers. The CephFS load
-balancers are hard-coded into Ceph while Mantle is an API for balancing load
-presented in~Mantle and re-implemented on top of
-Malacology in this paper.  The increased throughput for the CephFS and Mantle
+migrates sequencers using the hard-coded CephFS load balancers, and ``Mantle" uses a
+custom load balancer we wrote specifically for sequencers. 
+The increased throughput for the CephFS and Mantle
 curves between 0 and 60 seconds are a result of migrating the sequencer(s) off
@@
 something as dynamic and unpredictable as CPU utilization to make migration
-decisions. In addition to the suboptimal performance and unpredictability,
-because CephFS balancers all fall into the same mode despite using different
-balancers is a problem because it prevents administrators from properly
+decisions. 
+In addition to the suboptimal performance and unpredictability,
+another problem is that all the CephFS balancers behave the same. This
+prevents administrators from properly
 exploring the balancing state space.
@@
 Mantle gives the administrator more control over balancing policies; for the
-Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the load
-balancing interface to program logic for balancing read-heavy workloads,
+Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the Load
+Balancing interface to program logic for balancing read-heavy workloads,
 resulting in better throughput and stability.  When we did this we also
@@
 
-Using Mantle, the administrator can put the load balancer into ``proxy
-mode" or ``client mode". In proxy mode one server receives all requests and
-farms off the requests to slave servers; the slaves servers do the actual tail
-finding operation. In client mode, clients interact directly with the server
-that has their sequencer.  These modes are illustrated in
-Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
-co-located on one physical server -- performance for that mode is shown by the
-No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
-Mode" in the lower left shows how clients for sequencer 2 are re-directed to
-server B. The two ``Proxy mode" diagrams on the right show how clients continue
-sending requests to server A even though sequencer 2 has migrated to server B
-(the meaning of ``(Half)" is discussed in the next section).
+Using Mantle, the administrator can put the load balancer into ``proxy mode" or
+``client mode". In proxy mode one server receives all requests and farms off
+the requests to slave servers; the slave servers do the actual tail finding
+operation. In client mode, clients interact directly with the server that has
+their sequencer.  These modes are illustrated in Figure~\ref{fig:mantle-modes}.
+``No Balancing" is when all sequencers are co-located on one physical server --
+performance for that mode is shown by the ``No Balancing" curve in
+Figure~\ref{fig:mantle-balancer-behaviors}. In ``Proxy Mode", clients continue
+sending requests to server A even though some of the sequencers have been
+migrated to another server. Server A redirects client requests for sequencer 2
+to server B.  ``Proxy Mode (Half)" is shown in
+Figure~\ref{fig:mantle-balancer-beahviors}; in this scenario, half of the
+sequencers have migrated off the first server. Alternatively, ``Proxy Mode
+(Full)", which is not pictured, is when all the sequencers migrate off the
+first server.  ``Client Mode", shown on the far right of
+Figure~\ref{fig:mantle-modes}, shows how clients for sequencer 2 contact server
+B without a redirect from server A.
 
@@
 Performance of Sequencer 2 decreases because it stayed on the proxy which now
-handles all client requests, processes requests for Sequencer 2, and forwards
+processes requests for Sequencer 2, and forwards
 requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
@@
 cluster throughput of proxy mode. That graph also shows that Sequencer 2 has
-less throughput than Sequencer 2. In this case, the scatter-gather process used
+less throughput than Sequencer 1. In this case, the scatter-gather process used
 for cache coherence in the metadata protocols causes strain on the server
@@
 Another factor that affects performance in this environment is how much load is
-on each server; these experiments quantify that effect by programming the load
-balancing interface to control the amount of load to migrate. We call this
+on each server; these experiments quantify that effect by programming the Load
+Balancing interface to control the amount of load to migrate. We call this
 metric a ``migration unit".  Expressing this heuristic is not easily achievable
@@
 This code snippet uses globally defined variables and tables from the Mantle
-API to send half of the load on the current server (whoami) to the next ranked
+API~\ref{sevilla:sc15-mantle} to send half of the load on the current server (whoami) to the next ranked
 server (whoami + 1); the \texttt{targets} array is a globally defined table
@@
 metadata cluster because the first server does a lot of the cache coherence
-work) while the second server does all the client work. Proxy mode does the
+work) while the second server does all the processing. Proxy mode does the
 best in both cases and shows large performance gains when completely decoupling
@@
 degrees of aggressiveness in making migration decisions; CephFS makes a
-decision 10 seconds into the run and throughput jumps to 2500 ops/second within 10 seconds
+decision 10 seconds into the run and throughput jumps to 2500 ops/second 
 while Mantle takes more time to stabilize. This conservative behavior is

### 2017-03-07 paper/9_conclusion.tex
paper: add acknowledgements and pointers
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 construction of new services by re-purposing existing subsystem abstractions of
-the storage stack.
+the storage stack. For information on Malacology, Mantle, and ZLog, visit our
+website at \href{programmability.us}{programmability.us}.
 
@@
 new services.}
+
+\section{Acknowledgements}
+
+We thank the EuroSys reviewers for their hard work, attentiveness, and
+genuinely helpful suggestions. We also thank Mahesh Balakrishnan for
+championing and shepherding the paper. This work was partially funded by the
+Center for Research in Open Source Software
+(\href{cross.ucsc.edu}{cross.ucsc.edu}), the DOE Award DE-SC0016074, and the
+NSF Award 1450488.

### 2017-03-07 images/build.sh
image: tweak
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -4,8 +4,7 @@ set -ex
 
 # build builder image
 . docker-cephdev/aliases.sh
-docker build -t cephbuilder/ceph:jewel docker-cephdev/builder-base
-docker build -t cephbuilder/zlog-mantle:jewel docker-zlogmantledev
+docker build -t cephbuilder/ceph:latest docker-cephdev/builder-base
 
 # pull base image from ceph (we will layer on top of this)
 docker pull ceph/daemon:tag-build-master-jewel-ubuntu-14.04
@@ -18,16 +17,16 @@ cd $SRC
 
 # layer custom binaries on top of the base image
 dmake \
-  -e SHA1_OR_REF="remotes/origin/zlog-mantle" \
+  -e SHA1_OR_REF="remotes/origin/malacology/masterish" \
   -e GIT_URL="https://github.com/michaelsevilla/ceph.git" \
-  -e RECONFIGURE="true" \
   -e BUILD_THREADS=`grep processor /proc/cpuinfo | wc -l` \
   -e CONFIGURE_FLAGS="-DWITH_TESTS=OFF" \
-  cephbuilder/zlog-mantle:jewel
+  -e RECONFIGURE="true" \
+  cephbuilder/ceph:latest
 cd -
  
-docker rm -f new || true
-docker run -it -d --name new --entrypoint=/bin/bash ceph-heads/remotes/origin/zlog-mantle-base
+docker run -it -d --name new --entrypoint=/bin/bash ceph-heads/remotes/origin/malacology/masterish-base
 docker exec new sudo apt-get update -y || true
 docker exec new sudo apt-get install gdb -y
 docker commit --change='ENTRYPOINT ["/entrypoint.sh"]' new ceph/zlog-mantle:jewel
+docker rm -f new

### 2017-03-07 images/docker-zlogmantledev/entrypoint.sh

@@ -27,14 +27,7 @@ if [ "$INSTALL_HEADERS" = "true" ] ; then
   cp -L /ceph/src/include/rados/* /ceph/build/install/usr/include/rados/
 fi
 
-cd /ceph/src/zlog
-make
-cp seq-client /ceph/build/install/usr/bin
-
-apt-get update
-apt-get install -y gdb valgrind
-rm -rf /var/lib/apt/lists/* /tmp/* /var/tmp/* debian/
-
+cp /ceph/build/bin/seq-client /ceph/build/install/usr/bin
 
 export CEPH_INSTALL_PATH=/ceph/build/install
 . generate_daemon_image

### 2017-03-07 experiments/mds-zlog-seq-migrate-redux-waves/run.sh
experiments: remove unused experiments and popperize docker build
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -4,6 +4,8 @@ set -ex
 
 ANSIBLE="ansible-playbook --forks 50 --skip-tags with_pkg"
 
+# check if the user grabbed the correct images
+
 cd site
 
 export ANSIBLE_LOG_PATH="../logs/setup.log"

### 2017-03-07 experiments/run-all.sh

@@ -1,5 +1,0 @@
-#!/bin/bash
-set -ex
-cd scale-clients
-./run.sh
-cd -

### 2017-03-07 images/build.sh

@@ -1,6 +1,14 @@
 #!/bin/bash
 
+
+if [ -z $1 ]; then
+  echo "ERROR: please pass an endpoint for storing Docker images, like michaelsevilla/myimage"
+  echo "USAGE: $0 <endpoint>"
+  exit
+fi
+
 set -ex
+docker login
 
 # build builder image
 . docker-cephdev/aliases.sh
@@ -17,16 +25,24 @@ cd $SRC
 
 # layer custom binaries on top of the base image
 dmake \
-  -e SHA1_OR_REF="remotes/origin/malacology/masterish" \
+  -e SHA1_OR_REF="9bfe60e820893f8f2931d81e2fc31002b9a17f32" \
   -e GIT_URL="https://github.com/michaelsevilla/ceph.git" \
   -e BUILD_THREADS=`grep processor /proc/cpuinfo | wc -l` \
   -e CONFIGURE_FLAGS="-DWITH_TESTS=OFF" \
   -e RECONFIGURE="true" \
-  cephbuilder/ceph:latest
+  cephbuilder/ceph:latest build-cmake
 cd -
  
-docker run -it -d --name new --entrypoint=/bin/bash ceph-heads/remotes/origin/malacology/masterish-base
+docker run -it -d --name new --entrypoint=/bin/bash ceph-9bfe60e820893f8f2931d81e2fc31002b9a17f32-base
 docker exec new sudo apt-get update -y || true
 docker exec new sudo apt-get install gdb -y
 docker commit --change='ENTRYPOINT ["/entrypoint.sh"]' new ceph/zlog-mantle:jewel
 docker rm -f new
+
+docker tag ceph/zlog-mantle:jewel $1
+docker push $1
+
+set +x
+
+echo ""
+echo "DONE! Now cd to your experiment and run the run.sh script"

### 2017-03-08 paper/3_challenges.tex
paper: camera-ready template and url fixes
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 internal abstractions nor do we contend that the abstractions we choose provide
-the best trade-offs for all applications.} \newcomment{A programmable storage
+the best trade-offs for all applications.} \newcomment{A programmable storage system
 is not defined by what abstractions are exposed, rather a programmable storage

### 2017-03-08 paper/9_conclusion.tex
paper: camera-ready template and url fixes
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-02-23 paper/9_conclusion.tex
@@
 sacrificing correctness. To illustrate the benefits of this approach we
-presented Malacology, a programmable storage system that facilitates the
-construction of new services by re-purposing existing subsystem abstractions of
-the storage stack.
+presented Malacology\footnote{\url{http://programmability.us}}, a
+programmable storage system that facilitates the construction of new services
+by re-purposing existing subsystem abstractions of the storage stack.
 
@@
 new services.}
+
+\acks
+
+We thank the EuroSys reviewers for their hard work, attentiveness, and
+genuinely helpful suggestions. We also thank Mahesh Balakrishnan for
+championing and shepherding the paper. This work was partially funded by the
+Center for Research in Open Source
+Software\footnote{\url{http://cross.ucsc.edu}}, the DOE Award
+DE-SC0016074, and the NSF Award 1450488.
+
+\textbf{Note}: this paper follows The Popper
+Convention~\cite{jimenez_popper_2016}. The Popper
+Convention\footnote{\url{http://falsifiable.us}} is equivalent to
+the ACM's Result and Artifiact Review and Badging procedures.
+\includegraphics[width=3.5mm]{figures/popper.png}

### 2017-03-08 paper/paper.tex
paper: camera-ready template and url fixes
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-\documentclass[preprint]{sigplanconf-eurosys}
+%-----------------------------------------------------------------------------
+%
+%               Template for sigplanconf LaTeX Class
+%
+% Name:         sigplanconf-template.tex
+%
+% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
+%               file for SIGPLAN conference proceedings.
+%
+% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
+%               sigplanconf-guide.pdf
+%
+% Author:       Paul C. Anagnostopoulos
+%               Windfall Software
+%               978 371-2316
+%               paul@windfall.com
+%
+% Created:      15 February 2005
+% Source:       http://www.sigplan.org/Resources/Author/
+%-----------------------------------------------------------------------------
+
+\documentclass[numbers]{sigplanconf}
+
+% The following \documentclass options may be useful:
+
+% preprint       Remove this option only once the paper is in final form.
+%  9pt           Set paper in  9-point type (instead of default 10-point)
+% 11pt           Set paper in 11-point type (instead of default 10-point).
+% numbers        Produce numeric citations with natbib (instead of default author/year).
+% authorversion  Prepare an author version, with appropriate copyright-space text.
+
+\usepackage{amsmath}
 \usepackage{multirow}
@@
 
+\newcommand{\cL}{{\cal L}}
+
+\begin{document}
+
+\special{papersize=8.5in,11in}
+\setlength{\pdfpageheight}{\paperheight}
+\setlength{\pdfpagewidth}{\paperwidth}
+
+\conferenceinfo{CONF'yy}{Month d--d, 20yy, City, ST, Country}
+\copyrightyear{20yy}
+\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}\reprintprice{\$15.00}
+\copyrightdoi{nnnnnnn.nnnnnnn}
+
+% For compatibility with auto-generated ACM eRights management
+% instructions, the following alternate commands are also supported.
+%\CopyrightYear{2016}
+%\conferenceinfo{CONF'yy,}{Month d--d, 20yy, City, ST, Country}
+%\isbn{978-1-nnnn-nnnn-n/yy/mm}\acmPrice{\$15.00}
+%\doi{http://dx.doi.org/10.1145/nnnnnnn.nnnnnnn}
+
+% Uncomment the publication rights used.
+%\setcopyright{acmcopyright}
+\setcopyright{acmlicensed}  % default
+%\setcopyright{rightsretained}
+
+\titlebanner{This is a draft!}        % These are ignored unless
+\preprintfooter{short description of paper}   % 'preprint' option specified.
+
 \title{Malacology: A Programmable Storage System}
-%\authorinfo{Paper ID: 165}          {Total Pages: 12}
+
 \authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre, Carlos Maltzahn}
-           {\{msevilla, jayhawk, ivo\}@soe.ucsc.edu, \{palvaro, shel, jlefevre, carlosm\}@ucsc.edu}
-\date{}
-\begin{document}
+           {University of California, Santa Cruz}
+           {\{msevilla, jayhawk, ivo\}@soe.ucsc.edu, \{palvaro, shel\}@ucsc.edu, \{jlefevre, carlosm\}@soe.ucsc.edu}
 
@@
 \input{./0_abstract}
+
+% 2012 ACM Computing Classification System (CSS) concepts
+% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10011007.10011006.10011008</concept_id>
+<concept_desc>Software and its engineering~General programming languages</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10003752.10010124.10010138.10010143</concept_id>
+<concept_desc>Theory of computation~Program analysis</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Software and its engineering~General programming languages}
+\ccsdesc[300]{Theory of computation~Program analysis}
+% end generated code
+
+% Legacy 1998 ACM Computing Classification System categories are also
+% supported, but not recommended.
+%\category{CR-number}{subcategory}{third-level}[fourth-level]
+%\category{D.3.0}{Programming Languages}{General}
+%\category{F.3.2}{Logics and Meanings of Programs}{Semantics of Programming Languages}[Program analysis]
+
+\keywords
+keyword1, keyword2
+
 \input{./1_intro}
@@
 
+% The 'abbrvnat' bibliography style is recommended.
+\bibliographystyle{abbrvnat}
+
+% The bibliography should be embedded for final submission.
 \bibliography{paper}
-\bibliographystyle{plain}
 

### 2017-03-08 paper/2_application_specific_storage_stacks.tex
popper: add source links to graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics{figures/obj-int-dev-growth.png}
-\caption{Since 2010, the growth in the number of co-designed object storage
-interfaces in Ceph has been accelerating. This plot is the number of object
-classes (a group of interfaces), and the total number of methods (the actual
-API end-points).}
-\label{fig:obj-int-dev-growth}
+\caption{
+[\href{https://github.com/michaelsevilla/malacology-popper/blob/v2.1/experiments/objclass-dev/visualize.ipynb}{source}]
+Since 2010, the growth in the number of co-designed object storage interfaces
+in Ceph has been accelerating. This plot is the number of object classes (a
+group of interfaces), and the total number of methods (the actual API
+end-points).} \label{fig:obj-int-dev-growth}
 \end{figure}

### 2017-03-08 paper/6_evaluation.tex
popper: add source links to graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics{figures/capdelay-quota-behavior-c.png}
-\caption{Each dot is an individual request, spread randomly along the \(y\)
-axis. The default behavior is unpredictable, ``delay" lets clients hold the
-lease longer, and ``quota" gives clients the lease for a number of operations.}
+\caption{
+[\href{https://github.com/michaelsevilla/malacology-popper/blob/v2.1/experiments/mds-zlog-seq-thruput-v-latency/results-reqdots-kill0-sameclient-capdelay-100-quota-100000/visualize.ipynb}{source}]
+Each dot is an individual request, spread randomly along the \(y\) axis. The
+default behavior is unpredictable, ``delay" lets clients hold the lease longer,
+and ``quota" gives clients the lease for a number of operations.}
 \label{fig:capdelay-quota-behavior}
@@
 \includegraphics{figures/tradeoff.png}
-\caption{Sequencer throughput by re-using various services.  The highest
-performance is achieved using a single client with exclusive, cacheable
-privilege. Round-robin sharing of the sequencer resource is affected by the
-amount of time the resource is held, with best-effort performing the worst.}
+\caption{
+[\href{https://github.com/michaelsevilla/malacology-popper/blob/v2.1/experiments/zlog-seqr-redux/viz.ipynb}{source}]
+Sequencer throughput by re-using various services.  The highest performance is
+achieved using a single client with exclusive, cacheable privilege. Round-robin
+sharing of the sequencer resource is affected by the amount of time the
+resource is held, with best-effort performing the worst.}
 \label{fig:captp}
@@
 \includegraphics{figures/caps-delay-latency.png}
-\caption{ The latency distribution shows that if the clients hold their
-capabilities longer, performance decreases. Changing the delay can help the
-sequencer and clients strike a balance between throughput and latency.}
+\caption{
+[\href{https://github.com/michaelsevilla/malacology-popper/blob/v2.1/experiments/zlog-seqr-redux/viz.ipynb}{source}]
+The latency distribution shows that if the clients hold their capabilities
+longer, performance decreases. Changing the delay can help the sequencer and
+clients strike a balance between throughput and latency.}
 \label{fig:capcdf}
@@
 \includegraphics[width=60mm,trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
-\caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
+\caption{
+[\href{https://github.com/michaelsevilla/malacology-popper/tree/v2.1/experiments/mon-paxos-update/viz.py}{source}]
+Cluster-wide interface update latency, excluding the Paxos proposal cost for
 interface commit.}
@@
 \includegraphics{figures/mantle-balancer-behaviors.png}
-\caption{CephFS/Mantle load balancing have better throughput than co-locating
-all sequencers on the same server.  Sections~\ref{sec:feature-balancing-modes}
+\caption{
+[\href{https://github.com/michaelsevilla/malacology-popper/blob/v2.1/experiments/mds-zlog-seq-migrate-redux-3client/results-mantle-runs/visualize.ipynb}{source}]
+CephFS/Mantle load balancing have better throughput than co-locating all
+sequencers on the same server.  Sections~\ref{sec:feature-balancing-modes}
 and~\ref{sec:feature-migration-units} quantify this improvement;
@@
 \includegraphics{figures/mantle-mode-performance.png}
-\caption{In (a) all CephFS balancing modes have the same performance; Mantle
-uses a balancer designed for sequencers. In (b) the best combination of mode
-and migration units can have up to a 2\(\times\)
+\caption{
+[\href{https://github.com/michaelsevilla/malacology-popper/blob/v2.1/experiments/mds-zlog-seq-migrate-redux-3client/results-mantle-runs/visualize.ipynb}{source},
+\href{https://github.com/michaelsevilla/malacology-popper/blob/v2.1/experiments/mds-zlog-seq-migrate-redux-waves/results-paper/visualize.ipynb}{source}]
+In (a) all CephFS balancing modes have the same performance; Mantle uses a
+balancer designed for sequencers. In (b) the best combination of mode and
+migration units can have up to a 2\(\times\)
 improvement.}\label{fig:mantle-balancer-performance}
@@
 to server B.  ``Proxy Mode (Half)" is shown in
-Figure~\ref{fig:mantle-balancer-beahviors}; in this scenario, half of the
+Figure~\ref{fig:mantle-balancer-behaviors}; in this scenario, half of the
 sequencers have migrated off the first server. Alternatively, ``Proxy Mode
@@
 \includegraphics{figures/mantle-mode-behavior.png}
-\caption{The performance of proxy mode achieves the highest throughput but at
-the cost of lower throughput for one of the sequencers. Client mode is more
-fair but results in lower cluster throughput.
-}\label{fig:mantle-mode-behavior}
+\caption{
+[\href{https://github.com/michaelsevilla/malacology-popper/blob/v2.1/experiments/mds-zlog-seq-migrate-redux-waves/results-paper/visualize.ipynb}{source}]
+The performance of proxy mode achieves the highest throughput but at the cost
+of lower throughput for one of the sequencers. Client mode is more fair but
+results in lower cluster throughput.  }\label{fig:mantle-mode-behavior}
 \end{figure}
@@
 This code snippet uses globally defined variables and tables from the Mantle
-API~\ref{sevilla:sc15-mantle} to send half of the load on the current server (whoami) to the next ranked
+API~\cite{sevilla:sc15-mantle} to send half of the load on the current server (whoami) to the next ranked
 server (whoami + 1); the \texttt{targets} array is a globally defined table

### 2017-03-08 experiments/objclass-dev/cloc.sh

@@ -1,0 +1,33 @@
+#!/bin/bash
+set -e
+set -x
+
+output=`pwd`/$1
+
+sha1s="
+14f2c78614ba2572022d02f91a46ca6fe62c4aee
+789cb9a403839c395db652d226f411aa7fc14a61
+f8929bad58968af04ebf7ce54dd716cc1e98195f
+e10267b531fecb7e75dc99255a98fdbb913e6049
+40018fcc845e494aabfcbacd52ce6ac1d8197b3e
+020983b0a53fd8791637702ef9a82ec7369fd448
+8ef0dfb30e636454891a5a5c8fdd77dde3417953
+"
+
+dir=`mktemp -d`
+pushd $dir
+git clone --recursive git://github.com/ceph/ceph.git
+pushd ceph
+
+for sha1 in $sha1s; do
+  git checkout $sha1
+  git submodule sync || true
+  git submodule update --force --init --recursive || true
+  git clean -dxf  
+  echo "########### $sha1" | tee -a $output
+  cloc --exclude-dir=src/cls/hello src/cls_rbd.cc src/cls_acl.cc src/cls_crypto.cc src/cls/ | tee -a $output
+done
+
+popd
+popd
+rm -rf $dir

### 2017-03-08 paper/6_evaluation.tex
paper: tweaks
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 Cluster-wide interface update latency, excluding the Paxos proposal cost for
-interface commit.}
+committing the Service Metadata interface.}
 \label{fig:propdelay}

### 2017-03-08 paper/9_conclusion.tex
paper: tweaks
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 We thank the EuroSys reviewers for their hard work, attentiveness, and
-genuinely helpful suggestions. We also thank Mahesh Balakrishnan for
+genuinely helpful suggestions. We especially thank Mahesh Balakrishnan for
 championing and shepherding the paper. This work was partially funded by the
@@
 Convention\footnote{\url{http://falsifiable.us}} is equivalent to
-the ACM's Result and Artifiact Review and Badging procedures.
+the ACM's Result and Artifact Review and Badging procedures.
 \includegraphics[width=3.5mm]{figures/popper.png}

### 2017-03-10 paper/9_conclusion.tex
Adds sentence related to Popper
fixes #107

--- 2017-03-08 paper/9_conclusion.tex
@@
 We thank the EuroSys reviewers for their hard work, attentiveness, and
-genuinely helpful suggestions. We also thank Mahesh Balakrishnan for
-championing and shepherding the paper. This work was partially funded by the
+genuinely helpful suggestions. We especially thank Mahesh Balakrishnan for
+shepherding the paper. This work was partially funded by the
 Center for Research in Open Source
@@
 \textbf{Note}: this paper follows The Popper
-Convention~\cite{jimenez_popper_2016}. The Popper
-Convention\footnote{\url{http://falsifiable.us}} is equivalent to
-the ACM's Result and Artifiact Review and Badging procedures.
-\includegraphics[width=3.5mm]{figures/popper.png}
+Convention\footnote{\url{http://falsifiable.us}}~\cite{jimenez_popper_2016}. All 
+the experiments presented here are available on the repository associated to 
+this 
+article\footnote{\url{https://github.com/michaelsevilla/malacology-popper/tree/v2.1}}. 
+For every figure, a \texttt{[source]} link points to a Jupyter notebook that 
+shows the analysis from where the graph was obtained; its parent folder contains 
+all the associated artifacts.

### 2017-03-14 paper/0_abstract.tex
paper: fix abstract sentence, spell check, add CCS/keywords
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\begin{abstract} Storage systems are caught between the need to evolve data
-processing applications efficiently and quickly, and the increasing velocity
-with which storage device technology evolves. This puts tremendous pressure on
-storage systems to support rapid change both in terms of their interfaces and
-their performance. But adapting storage systems can be difficult because
-unprincipled changes might jeopardize years of code-hardening and performance
-optimization efforts that were necessary for users to entrust their data to the
-storage system. We introduce the programmable storage approach, which exposes
-internal services and abstractions of the storage stack as building blocks for
+\begin{abstract} Storage systems need to support high-performance for
+special-purpose data processing applications that run on an evolving storage
+device technology landscape.  This puts tremendous pressure on storage systems
+to support rapid change both in terms of their interfaces and their
+performance. But adapting storage systems can be difficult because unprincipled
+changes might jeopardize years of code-hardening and performance optimization
+efforts that were necessary for users to entrust their data to the storage
+system. We introduce the programmable storage approach, which exposes internal
+services and abstractions of the storage stack as building blocks for
 higher-level services. We also build a prototype to explore how existing

### 2017-03-14 paper/6_evaluation.tex
paper: fix abstract sentence, spell check, add CCS/keywords
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 metric a ``migration unit".  Expressing this heuristic is not easily achievable
-with outward facing tunable parameters ({\it i.e.} system knobs) but with Mantle's
+with outward facing tunable parameters (i.e. system knobs) but with Mantle's
 programmable interface, we can force the load balancer to change its migration

### 2017-03-14 paper/paper.tex
paper: fix abstract sentence, spell check, add CCS/keywords
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\conferenceinfo{CONF'yy}{Month d--d, 20yy, City, ST, Country}
-\copyrightyear{20yy}
-\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}\reprintprice{\$15.00}
-\copyrightdoi{nnnnnnn.nnnnnnn}
+\CopyrightYear{2017} 
+\setcopyright{acmcopyright}
+\conferenceinfo{EuroSys '17,}{April 23-26, 2017, Belgrade, Serbia}
+\isbn{978-1-4503-4938-3/17/04}\acmPrice{\$15.00}
+\doi{http://dx.doi.org/10.1145/3064176.3064208}
+
+%\conferenceinfo{CONF'yy}{Month d--d, 20yy, City, ST, Country}
+%\copyrightyear{20yy}
+%\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}\reprintprice{\$15.00}
+%\copyrightdoi{nnnnnnn.nnnnnnn}
 
@@
 %\setcopyright{acmcopyright}
-\setcopyright{acmlicensed}  % default
+%\setcopyright{acmlicensed}  % default
 %\setcopyright{rightsretained}
@@
 <concept>
-<concept_id>10011007.10011006.10011008</concept_id>
-<concept_desc>Software and its engineering~General programming languages</concept_desc>
+<concept_id>10002951.10003152.10003517.10003519</concept_id>
+<concept_desc>Information systems~Distributed storage</concept_desc>
 <concept_significance>500</concept_significance>
@@
 <concept>
-<concept_id>10003752.10010124.10010138.10010143</concept_id>
-<concept_desc>Theory of computation~Program analysis</concept_desc>
+<concept_id>10011007.10010940.10010941.10010949.10003512</concept_id>
+<concept_desc>Software and its engineering~File systems management</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10011007.10010940.10010992</concept_id>
+<concept_desc>Software and its engineering~Software functional properties</concept_desc>
 <concept_significance>300</concept_significance>
@@
 
-\ccsdesc[500]{Software and its engineering~General programming languages}
-\ccsdesc[300]{Theory of computation~Program analysis}
+\ccsdesc[500]{Information systems~Distributed storage}
+\ccsdesc[500]{Software and its engineering~File systems management}
+\ccsdesc[300]{Software and its engineering~Software functional properties}
 % end generated code
@@
 \keywords
-keyword1, keyword2
+Distributed Storage, Programmability, Ceph
 

### 2017-03-14 paper/4_malacology_a_programmable_storage_system.tex
paper: fix abstract sentence, spell check, add CCS/keywords
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: fix line breaks
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \end{tabular}
-\caption{Common internal abstractions. The same ``internal abstractions" are common
-across large-scale systems because they provide functionality that solves general
-distributed systems problems.  Here we list examples of what these internal
-abstractions are used for in ``production systems" and in Ceph.  Malacology
-provides these internal abstractions as interfaces
+\caption{Common internal abstractions. The same ``internal abstractions" are
+common across large-scale systems because they provide functionality that
+solves general distributed systems problems.  Here we list examples of what
+these internal abstractions are used for in ``production systems" and in Ceph.
+Malacology provides these internal abstractions as interfaces
 \newcomment{(Section~\ref{sec:malacology})} that higher level
@@
 these services can be \emph{programmed}. We accomplish programmability of a
-service by exporting bindings for an interpreted programming
-language so that programming can occur without having to restart the storage
-system (see also below, Section~\ref{sec:durability}). Table~\ref{table:examples}
-shows the internal services from Ceph that we expose in the Malacology
-prototype via Lua~\cite{ierusalimschy1996lua} bindings\oldcomment{.}
-\newcomment{and Figure~\ref{fig:implementation-overview} compares what was
-already present in Ceph (gray boxes) to the Malacology interfaces we added.
+service by exporting bindings for an interpreted programming language so that
+programming can occur without having to restart the storage system (see also
+below, Section~\ref{sec:durability}). Table~\ref{table:examples} shows the
+internal services from Ceph that we expose in the Malacology prototype via
+Lua~\cite{ierusalimschy1996lua} bindings\oldcomment{.} \newcomment{and
+Figure~\ref{fig:implementation-overview} compares what was already present in
+Ceph (gray boxes) to the Malacology interfaces we added.
 Section~\ref{sec:services} will describe the higher-level services we built
@@
 \newcommentone{Service metadata is information about the daemons in the system
-and includes membership details, hardware layout (e.g., racks, power
-supplies, etc.), data layout, and daemon state and configuration. It differs from
+and includes membership details, hardware layout (e.g., racks, power supplies,
+etc.), data layout, and daemon state and configuration. It differs from
 traditional file system metadata which is information about files. For the rest
 of the paper when we use the phrase ``metadata server" or ``metadata service",
-we are referring to the daemon(s) that manages file system metadata (not service metadata).}
-\addressesissue{4}
+we are referring to the daemon(s) that manages file system metadata (not
+service metadata).} \addressesissue{4}
 
@@
 
-\noindent{\it{\textbf{Malacology: }}} we expose a strongly-consistent
-view of time-varying service metadata as an interface rather than a hidden
-internal component. This is shown in Figure~\ref{fig:implementation}, 
-where object interfaces and load balancer policies use the Service Metadata
-interface.
-Malacology provides a generic API for adding arbitrary
-values to existing subsystem cluster maps. As a consequence of this,
-applications can define simple but useful service-specific logic to the
-strongly-consistent interface, such as authorization control (just specific
-clients can write new values) or triggering actions based on specific values
-(e.g. sanitize values).  The higher-level services we implement in Section~\ref{sec:services} make use of this functionality to register, version and
-propagate dynamic code (Lua scripts) for new object interfaces defined in
-storage daemons (Section~\ref{object-data-interface}) and policies in the load
-balancer (Section~\ref{malacology:mds}).  Using this service guarantees that interface
+\noindent{\it{\textbf{Malacology: }}} we expose a strongly-consistent view of
+time-varying service metadata as an interface rather than a hidden internal
+component. This is shown in Figure~\ref{fig:implementation}, where object
+interfaces and load balancer policies use the Service Metadata interface.
+Malacology provides a generic API for adding arbitrary values to existing
+subsystem cluster maps. As a consequence of this, applications can define
+simple but useful service-specific logic to the strongly-consistent interface,
+such as authorization control (just specific clients can write new values) or
+triggering actions based on specific values (e.g. sanitize values).  The
+higher-level services we implement in Section~\ref{sec:services} make use of
+this functionality to register, version and propagate dynamic code (Lua
+scripts) for new object interfaces defined in storage daemons
+(Section~\ref{object-data-interface}) and policies in the load balancer
+(Section~\ref{malacology:mds}).  Using this service guarantees that interface
 definitions are not only made durable, but are transparently and consistently
@@
 \noindent\newcomment{{\it{\textbf{Impact: }}} provides core functionality
-because it lets daemons come to a consensus on system critical state. Bugs in the
-internal subsystems or omitting this from services that need this type of
+because it lets daemons come to a consensus on system critical state. Bugs in
+the internal subsystems or omitting this from services that need this type of
 consistency affects correctness.} \addressesissue{3, 5}
@@
 
-\oldcomment{\noindent 4.2 Object I/O Interfrace}
+\oldcomment{\noindent 4.2 Object I/O Interface}
 \subsection{Data I/O Interface}
@@
 object storage API with economy of expression, which at the same time provides
-the full set of features of the original object interface implementation. New object interfaces that
-are expressed in thousands of lines of code can be implemented in approximately
-an order of magnitude less code~\cite{geambasu_comet_2010}. While the use of
-Lua does not prevent deployment of malicious code, certain types of coding
-mistakes can be handled gracefully, and access policies are used to limit
-access to trusted users~\cite{ierusalimschy1996lua}.
+the full set of features of the original object interface implementation. New
+object interfaces that are expressed in thousands of lines of code can be
+implemented in approximately an order of magnitude less
+code~\cite{geambasu_comet_2010}. While the use of Lua does not prevent
+deployment of malicious code, certain types of coding mistakes can be handled
+gracefully, and access policies are used to limit access to trusted
+users~\cite{ierusalimschy1996lua}.
 
@@
 \newcomment{\\ \noindent {\it{\textbf{Existing Ceph Implementation: }}}}
-addresses the challenge of balancing \newcommentone{file system} metadata
-load with a separate metadata cluster. This cluster uses load balancing
-policies to migrate directory inodes around the cluster to alleviate overloaded servers~\cite{weil:sc2004-dyn-metadata}.  The policies use metrics
-based on system state (e.g.  CPU and memory utilization) and statistics
-collected by the cluster (e.g. the popularity of an inode). Ceph uses dynamic
-subtree partitioning to move variable sized namespace subtrees. These units can
-be shipped anywhere (i.e., to any metadata server of any capacity) at any time
-for any reason. The original balancer was designed with hard-coded policies and
+addresses the challenge of balancing \newcommentone{file system} metadata load
+with a separate metadata cluster. This cluster uses load balancing policies to
+migrate directory inodes around the cluster to alleviate overloaded
+servers~\cite{weil:sc2004-dyn-metadata}.  The policies use metrics based on
+system state (e.g.  CPU and memory utilization) and statistics collected by the
+cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
+partitioning to move variable sized namespace subtrees. These units can be
+shipped anywhere (i.e., to any metadata server of any capacity) at any time for
+any reason. The original balancer was designed with hard-coded policies and
 tunables.\\
@@
 
-\noindent{\it{\textbf{Existing Ceph Implementation: }}} provides storage
-by striping and replicating data across RADOS~\cite{weil_rados_2007}, the
-reliable distributed object store. RADOS protects data using common techniques
-such as erasure coding, replication, and scrubbing. For example, when the number
-of placement groups change, the object storage daemons re-balance and re-shard data in the
-background in a process called placement group splitting. During placement group splitting, object storage daemons
-communicate directly with each other to converge on a new data layout.  In order to reduce
-load on the monitoring service, the object storage daemons use a gossip protocol to efficiently
-propagate changes to cluster maps throughout the system, and autonomously
-initiate recovery mechanisms when failures are discovered.\\
+\noindent{\it{\textbf{Existing Ceph Implementation: }}} provides storage by
+striping and replicating data across RADOS~\cite{weil_rados_2007}, the reliable
+distributed object store. RADOS protects data using common techniques such as
+erasure coding, replication, and scrubbing. For example, when the number of
+placement groups change, the object storage daemons re-balance and re-shard
+data in the background in a process called placement group splitting. During
+placement group splitting, object storage daemons communicate directly with
+each other to converge on a new data layout.  In order to reduce load on the
+monitoring service, the object storage daemons use a gossip protocol to
+efficiently propagate changes to cluster maps throughout the system, and
+autonomously initiate recovery mechanisms when failures are discovered.\\
 

### 2017-03-16 paper/1_intro.tex
paper: fix line breaks
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: camera ready
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 (e.g., file system metadata servers). Malacology enables the programmability of
-internal abstractions (bold arrows) to re-use and compose
-existing subsystems.  With Malacology, we built new higher-level services, ZLog and
-Mantle, that sit alongside traditional user-facing APIs (file, block,
+internal abstractions (bold arrows) to re-use and compose existing subsystems.
+With Malacology, we built new higher-level services, ZLog and Mantle, that sit
+alongside traditional user-facing APIs (file, block,
 object).}\label{fig:overview}
@@
 
-A storage system implements abstractions designed to persistently store data
-and must exhibit a high level of correctness to prevent data loss.  Storage
-systems have evolved around storage devices that often were orders of magnitude
-slower than CPU and memory, and therefore could dominate overall performance if
-not used carefully. Over the last few decades members of the storage systems
-community have developed clever strategies to meet correctness requirements
-while somewhat hiding the latency of traditional storage
+A{\let\thefootnote\relax\footnote{$^{\dag}$ These authors contributed equally to
+this work.}} storage system implements abstractions designed to persistently
+store data and must exhibit a high level of correctness to prevent data loss.
+Storage systems have evolved around storage devices that often were orders of
+magnitude slower than CPU and memory, and therefore could dominate overall
+performance if not used carefully. Over the last few decades members of the
+storage systems community have developed clever strategies to meet correctness
+requirements while somewhat hiding the latency of traditional storage
 media~\cite{brewer_disks_2016}. To avoid lock-in by a particular vendor, users
@@
 
-\item An interface for managing strongly-consistent time-varying \textbf{service
-metadata}.
+\item An interface for managing strongly-consistent time-varying
+\textbf{service metadata}.
 
@@
 \item A high-performance distributed shared log service called ZLog, that is an
-implementation of CORFU.~\cite{balakrishnan_corfu_2012}
+implementation of CORFU~\cite{balakrishnan_corfu_2012}
 
@@
 evaluate our ideas within our prototype implementation
-(Section~\ref{sec:evaluation}).  We conclude by discussing future and related work.
+(Section~\ref{sec:evaluation}).  We conclude by discussing future and related
+work.

### 2017-03-16 paper/paper.tex
paper: camera ready
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \usepackage{comment}
+\usepackage{marvosym}
 
@@
 
-\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre, Carlos Maltzahn}
+\authorinfo{Michael A. Sevilla$^{\dag}$, Noah Watkins$^{\dag}$, Ivo Jimenez, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre, Carlos Maltzahn}
            {University of California, Santa Cruz}
